### 2017-04-21 headers/commands/todo.tex
Template for exercise sheets

\newtoggle{isDraft} % Documents that have this set to false won't contain any todo's.

\toggletrue{isDraft} % Diese Datei wurde aus einem anderen Projekt übernommen.
					 % Solange wir noch nicht genau ausgeknobelt haben, wie das mit dem Drafting läuft, ignorieren wir es erst mal und zeigen TODO's immer an.

% For todo's in text mode:
%\newcommand{\todo}[2]{\iftoggle{isDraft}{\textcolor{red}  {\textbf{TODO\textsubscript{#1}: #2}}}{} \PackageWarning{TODO}{TODO: #2} }

\usepackage{xargs} %use more than 1 argument
%\usepackage[pdftex,dvipsnames]{xcolor} %colors and stuff
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes} %todo structure
\newcommandx{\vtodo}[3][1=]{\todo[author=#2,#1]{#3}} %basic todo-note, only allowed to authors of this section
\newcommandx{\figtodo}[3][1=]{\missingfigure[author=#2,figcolor=white,#1]{#3}}

% For stuff that should only appear in a draft:
\NewEnviron{draft}{\iftoggle{isDraft}{\textcolor{darkgray}\BODY}{}}

% For margin todo's:
\setlength{\marginparsep}{2mm}
\newcommand{\mtodo}[2]{\iftoggle{isDraft}{\marginpar{\tiny\textcolor{red}  {(#1): #2}}}{} \PackageWarning{TODO}{TODO: #2}}
% For indicating that the current version is a draft:
\fancyfoot[L,R]{\iftoggle{isDraft}{\textcolor{red} {\textbf{E N T W U R F}}}{}}

### 2017-04-21 headers/sheetHeader.tex
Template for exercise sheets

\documentclass{article}

\input{\pathToRoot headers/sharedHeader}

\geometry{a4paper,left=2cm,right=2cm,top=2cm,bottom=3cm}

\newcommand{\licenseccjuliachristian}{\def\islicenseccjuliachristian{}}
\newcommand{\suppresslicense}{\def\issuppresslicense{}}

\AtBeginDocument{
    \pagestyle{fancy}
    \renewcommand{\headrulewidth}{0pt}
    \renewcommand{\footrulewidth}{1pt}
    \fancyhead{}
    \fancyfoot[C]{\thepage~/~\pageref{LastPage}}
}

\newcommand{\pgbreakhere}{\Needspace*{4\baselineskip}}
\newcommand{\pgbreakHere}{\Needspace*{10\baselineskip}}
\newcommand{\pgbreakHERE}{\Needspace*{15\baselineskip}}

\newcommand{\raisedrule}[2][0em]{\leavevmode\leaders\hbox{\rule[#1]{1pt}{#2}}\hfill\kern0pt}

% inspired by http://tex.stackexchange.com/questions/242294/suppress-parskip-only-after-a-specific-paragraph
\makeatletter
\newlength\noparskip@parskip % used to store a backup of the parskip value
\newboolean{noparskip@triggered} % flag to indicate that noparskip was run in the current paragraph
\setboolean{noparskip@triggered}{false}
\newboolean{noparskip@active} % flag to indicate that parskip should be restored after this paragraph
\setboolean{noparskip@active}{false}
\let\noparskip@par\par % store a backup of the \par command
\@setpar{% redefine \par with the means of ltpar.dtx to stay compatible to enumerate and itemize
    \ifhmode% since we're counting occurrences of \par, \par\par would be a problem, so check that we are actually ending a paragraph
        \ifthenelse{\boolean{noparskip@active}}{%
            \setlength\parskip\noparskip@parskip% restore parskip
            \setboolean{noparskip@active}{false}% remember not the restore parskip again
        }{}%
        \ifthenelse{\boolean{noparskip@triggered}}{%
            \ifthenelse{\boolean{noparskip@active}}{}{
                % we are triggering noparskip and not currently in a noparskip already
                \setlength\noparskip@parskip\parskip % copy the current parskip into the backup variable
            }%
            \setboolean{noparskip@triggered}{false}% paragraph is ending, so noparskip is no longer triggered
            \setlength\parskip{0pt}% no parskip when the next paragraph begins
            \setboolean{noparskip@active}{true}% parskip must be restored by the next par
        }{}%
    \fi%
    \noparskip@par% run the original par command
}
\def\noparskip@backout{%
    \ifthenelse{\boolean{noparskip@active}}{%
        % a list is beginning and parskip is currently set to zero, wich would mess up the list
        \setlength\parskip{\noparskip@parskip}% restore parskip before the list begins
        \setboolean{noparskip@active}{false}%
    }{}%
    \setboolean{noparskip@triggered}{false}% there's no sense in keeping noparskip triggered throughout a list
}
\xpretocmd\begin{%
    \ifstrequal{#1}{enumerate}{\noparskip@backout}{}%
    \ifstrequal{#1}{itemize}{\noparskip@backout}{}%
    \ifstrequal{#1}{list}{\noparskip@backout}{}%
    \ifstrequal{#1}{proof}{\noparskip@backout}{}%
}{}{}
\def\noparskip{%
    \leavevmode% ensure that we are within a paragraph
    \setboolean{noparskip@triggered}{true}% trigger noparskip
}
\makeatother

\newcommand{\noparskipworkaround}{} % DEPRECATED and no longer needed

\newcommand{\head}[1]{
    {
        \setlength{\parskip}{0pt}
        \hrule height 1pt
        \vspace{.2cm}
        Saarland University \hfill Category Theory Seminar 2017\par
        Programming Systems Lab \hfill \small\url{https://courses.ps.uni-saarland.de/ct_ss17/}\par
        \tiny\raisedrule[0mm]{1pt}
        \vspace{2ex}
        \begin{center}
            \Large
            \textbf{#1}\par
            \raisedrule[2mm]{1pt}
        \end{center}
        \vspace{3ex}
    }
}

\newenvironment{leftframedparagraph}{\begin{mdframed}[hidealllines = true, leftline = true, innerleftmargin = 2ex, innerrightmargin = 0pt,
innertopmargin = 0pt, innerbottommargin = 2pt, skipabove=2ex, skipbelow=1ex, outerlinewidth = 0ex, innerlinewidth = 0.5ex]}{\end{mdframed}}
\newenvironment{leftframed}{\begin{mdframed}[hidealllines = true, leftline = true, innerleftmargin = 2ex, innerrightmargin = 0pt,
innertopmargin = 0pt, innerbottommargin = 0pt, skipabove=2ex, skipbelow=1ex, outerlinewidth = 0ex, innerlinewidth = 0.5ex]}{\end{mdframed}}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-04-21 sheets/basis-0/sheetTemplate.tex
Template for exercise sheets

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 0}{Sets, functions, groups and orders}

\section{Sets and functions}

\begin{exercise}
  The composition of injective functions is injective.
\end{exercise}

\begin{exercise}
  If $f \circ g$ is surjective, what do you know about $g$? Is it injective, surjective, bijective?
\end{exercise}

\section{Semigroups, monoids and groups}

\begin{definition}{Semigroup}
  A semigroup $(S, \cdot)$ consists out of a non-empty type $S$ with an operation $\cdot : S \to S \to S$ which is associative.
  $S$ is called the carrier.
\end{definition}

\begin{definition}{Monoid}
  A monoid is a semi-group with an identity $\alpha$, i.e. $\forall x : S, \alpha \cdot x = x = x \cdot \alpha$.
\end{definition}

\begin{exercise}
  List semi-groups with the following carriers:
  \begin{multicols}{3}
    \begin{itemize}
    \item $\mathbb{N}$
    \item $\mathcal{P}(W)$ for a finite set $W$
    \item $S$ for any non-empty set $S$
    \end{itemize}
  \end{multicols}

  Which of them is a monoid?
\end{exercise}

\begin{definition}{Canonical pre-order}
  A semi-group can be ordered in two canonical ways:
  \begin{itemize}
  \item $x \leq^R_\cdot y := \exists z, y = x \cdot z$
  \item $x \leq^L_\cdot y := \exists z, y = z \cdot x$
  \end{itemize}
\end{definition}

\begin{definition}{Canonically ordered semi-group}
  A semi-group is canonically ordered if $\leq^R_\cdot$ and $\leq^L_\cdot$ are partial orders.
\end{definition}

\begin{definition}{Group}
  A group is a monoid $(S, \cdot, \alpha)$ with inverse elements, i.e. $\forall x, \exists x^{-1}, x \cdot x^{-1} = \alpha$.
  A group is trivial if $S$ is a singleton.
\end{definition}

\begin{exercise}
  Only a trivial group is canonically ordered.
\end{exercise}

\section{Diagrams}

\begin{hint}
  This is how you can draw diagrams:
  \[
    \begin{tikzcd}
                                    & A \arrow{dr}{g}    & \\
      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
    \end{tikzcd}
  \]
\end{hint}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-04-21 sheets/basis-1/sheet-basis-1.tex
Template for exercise sheets

--- 2017-04-21 sheets/basis-0/sheetTemplate.tex
@@
 % Use Basis x or Talk x, where x is the number of the session
-\uebunghead{Basis 0}{Sets, functions, groups and orders}
+\uebunghead{Basis 1}{What is a Category?}
 
-\section{Sets and functions}
-
-\begin{exercise}
-  The composition of injective functions is injective.
-\end{exercise}
-
-\begin{exercise}
-  If $f \circ g$ is surjective, what do you know about $g$? Is it injective, surjective, bijective?
-\end{exercise}
-
-\section{Semigroups, monoids and groups}
-
-\begin{definition}{Semigroup}
-  A semigroup $(S, \cdot)$ consists out of a non-empty type $S$ with an operation $\cdot : S \to S \to S$ which is associative.
-  $S$ is called the carrier.
-\end{definition}
-
-\begin{definition}{Monoid}
-  A monoid is a semi-group with an identity $\alpha$, i.e. $\forall x : S, \alpha \cdot x = x = x \cdot \alpha$.
-\end{definition}
-
-\begin{exercise}
-  List semi-groups with the following carriers:
-  \begin{multicols}{3}
-    \begin{itemize}
-    \item $\mathbb{N}$
-    \item $\mathcal{P}(W)$ for a finite set $W$
-    \item $S$ for any non-empty set $S$
-    \end{itemize}
-  \end{multicols}
-
-  Which of them is a monoid?
-\end{exercise}
-
-\begin{definition}{Canonical pre-order}
-  A semi-group can be ordered in two canonical ways:
-  \begin{itemize}
-  \item $x \leq^R_\cdot y := \exists z, y = x \cdot z$
-  \item $x \leq^L_\cdot y := \exists z, y = z \cdot x$
-  \end{itemize}
-\end{definition}
-
-\begin{definition}{Canonically ordered semi-group}
-  A semi-group is canonically ordered if $\leq^R_\cdot$ and $\leq^L_\cdot$ are partial orders.
-\end{definition}
-
-\begin{definition}{Group}
-  A group is a monoid $(S, \cdot, \alpha)$ with inverse elements, i.e. $\forall x, \exists x^{-1}, x \cdot x^{-1} = \alpha$.
-  A group is trivial if $S$ is a singleton.
-\end{definition}
-
-\begin{exercise}
-  Only a trivial group is canonically ordered.
-\end{exercise}
+\begin{hint}
+  Read Chapter 1.1. For initial and terminal objects refer to Definition 2.17 on pages 48ff. For Monics and Epics, see the respective sections in Chapters 5.1 and 5.2.
+\end{hint}
 

### 2017-05-03 sheets/basis-1/sheet-basis-1.tex
basis1 ex1-3

@@
 
-\section{Diagrams}
+%\section{Diagrams}
+%
+%\begin{hint}
+%  This is how you can draw diagrams:
+%  \[
+%    \begin{tikzcd}
+%                                    & A \arrow{dr}{g}    & \\
+%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+%    \end{tikzcd}
+%  \]
+%\end{hint}
 
-\begin{hint}
-  This is how you can draw diagrams:
-  \[
-    \begin{tikzcd}
-                                    & A \arrow{dr}{g}    & \\
-      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
-    \end{tikzcd}
-  \]
-\end{hint}
+\section{Categories - Definition and Basics}
+
+\begin {exercise}
+Show there can be at most one inverse for a map $f \from A \to B$.
+\end{exercise}
+
+\begin {definition}{Rel}
+The objects of the category \textbf{Rel} are sets, an arrow $f \from A \to B$ is a subset $f \sub A \times B$. 
+The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\}$. 
+Composition of two maps $f \sub A \times B$, $g \sub B \times C$ is defined as 
+\[ g \of f = \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\} \]
+\end{definition}
+
+\begin {exercise}
+Show that \textbf{Rel} is a category.
+\end{exercise}
+
+\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets 
+(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A_\cdot $ ) 
+An arrow $ m \from A \to B $ function that is monotone, i.e. $\forall a, a' \in A$, 
+\[a \leq^A_\cdot a' \implies m(a) \leq^A_\cdot m(a')\]
+\end{definition}
+
+\begin {exercise}
+Show that \textbf{Pos} is a category.
+\end{exercise}
 

### 2017-05-03 sheets/basis-1/sheet-basis-1s.tex
basis1 ex1-3

--- 2017-05-03 sheets/basis-1/sheet-basis-1.tex
@@
 \begin {exercise}
-Show there can be at most one inverse for a map $f \from A \to B$.
+We have to show that for $f \from A \to B$, there is at most one $g \from B \to A$ s.t. $gf = 1_A$ and $fg = 1_B$. 
+\\Assume that $g$ exists and that there is some $h \from B \to A$: $hf = 1_A$, $fh = 1_B$. 
+\\We know from the identity axiom: $1_A \of g = g$ and $h \of 1_B = h$. Using this, our assumtions and the associativity axiom, we get:
+ \[  1_A \of g = g \textiff (h \of f) \of g = g \textiff h \of (f \of g) = g \textiff h \of 1_B = g \textiff h=g. \]
 \end{exercise}
@@
 The objects of the category \textbf{Rel} are sets, an arrow $f \from A \to B$ is a subset $f \sub A \times B$. 
-The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\}$. 
+The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\} := 1_A$. 
 Composition of two maps $f \sub A \times B$, $g \sub B \times C$ is defined as 
@@
 \begin {exercise}
-Show that \textbf{Rel} is a category.
+\textbf{identity laws:} for $f \sub A \times B$, 
+\[f \of 1_A = \{ \angles{a, b} \in A \times B \such \exists a' \emptybk (\angles {a, a'} \in 1_A \emptybk \& \emptybk \angles{a', b} \in f)\} = \{ \angles{a, b} \in A \times B \such \exists a' \in A.
+\emptybk \angles{a', b} \in f)\} = f.\] Analogously, $1_B \of f = f$.
+
+\textbf{associativity:} for $f \sub A \times B, g\sub B \times C, h \sub C \times D$, 
+\[
+h \of (g \of f) = h \of \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\}
+\]
+\[= \{ \angles{a, d} \in A \times D \such \exists b, c \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g \& \angles{c, d} \in h \} = (h \of g) \of f\]
+
 \end{exercise}
@@
 \begin {exercise}
-Show that \textbf{Pos} is a category.
+\textbf{identity laws:} $1_A$ is the identity function, which is monotone since $a \leq^A_\cdot a' \implies a' \leq^A_\cdot a.$ 
+\\The axiom holds because for $f \from A \in B, f \of 1_A = f = 1_B \of f$ (same as in \textbf{Set}).
+\\
+\textbf{associativity:} If $ f \from A \to B, g \from B \to C $ are monotone, so is $g \of f$ because \[a \leq^A_\cdot a' \implies f(a) \leq^A_\cdot f(a') \implies  g(f(a)) \leq^A_\cdot g(f(a'))  .\] Associativity follows from assiociativity of function composition.
 \end{exercise}

### 2017-05-04 sheets/basis-1/sheet1.tex
Add files via upload

--- 2017-05-03 sheets/basis-1/sheet-basis-1.tex
@@
 
+\begin{exercise}
+For a fixed set X with Powerset P(X), does P(X) $\cong$ $P(X)^{op}$ hold?
+\end{exercise}
+
+\begin{exercise}
+  Show that REL is a pointed category.
+   \\ Show that the category GRP of groups has both an initial and
+a final object, and these are the same.
+\\ Show that the category RING of unital rings has both an initial and a
+final object, and these are not the same. 
+\end{exercise}
+
+\begin{exercise}
+Prove that any isomorphism is both monic and epic.
+\end{exercise}
+\begin{exercise}
+Let $A, B, C$ be objects in any category
+  and let $f : A \rightarrow B$ and $g : B \rightarrow C$
+  be maps in that category. Prove the following facts.
+  \[
+    \begin{tikzcd}
+      A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
+                                   & C & 
+    \end{tikzcd}
+  \]
+
+  \begin{itemize}
+  \item If both $f$ and $g$ are isomorphisms,
+    then $h$ is also an isomorphism.
+  \item If $h$ is monic, $f$ is monic.
+  \item If $h$ is epic, $g$ is epic.
+  \end{itemize}
+\end{exercise}
+\begin{exercise}
+  Prove the following for a map $f \from A \to B$
+  in the category \textbf{Set},
+  \begin{itemize}
+  \item If $f$ is monic, it is injective.
+  \item If $f$ is epic, it is surjective. \textit{(Challenge 1)}
+  \end{itemize}
+\end{exercise}
+\begin{exercise}[Challenge 2]
+  Prove that in the category \textbf{Ring}, there are maps
+  which are epic but not surjective.
+  \begin{hint}
+    Consider the inclusion map $\mathbb{Z} \to \mathbb{Q}$, and
+    use the properties of ring homomorphisms.
+  \end{hint}
+\end{exercise}
+
 \end{document}

### 2017-05-04 sheets/basis-1/sheet-basis-1.tex
Merged pull request

@@
 \uebunghead{Basis 1}{What is a Category?}
+
+\author{Sarah Mameche, Andreas Meyer, Leonhard Staut}
 
@@
 The objects of the category \textbf{Rel} are sets, an arrow $f \from A \to B$ is a subset $f \sub A \times B$.
-The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\}$. 
+The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\} := 1_A$.
 Composition of two maps $f \sub A \times B$, $g \sub B \times C$ is defined as
-\[ g \of f = \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\} \]
+\[ g \of f = \{ \angles{a, c} \in A \times C \such \exists b. \angles {a, b} \in f \land \angles{b, c} \in g\} \]
 \end{definition}
@@
 \begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
-(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A_\cdot $ ) 
-An arrow $ m \from A \to B $ function that is monotone, i.e. $\forall a, a' \in A$, 
-\[a \leq^A_\cdot a' \implies m(a) \leq^A_\cdot m(a')\]
+(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A $ )
+An arrow $ m \from A \to B $ is a monotone function, i.e. $\forall a, a' \in A$,
+\[a \leq^A a' \implies m(a) \leq^A m(a')\]
 \end{definition}
@@
 Show that \textbf{Pos} is a category.
+\end{exercise}
+
+\begin{exercise}
+For a fixed set X with Powerset $\mathcal{P}(X)$, does $\mathcal{P}(X) \cong \mathcal{P}(X)^{op}$ hold?
+\end{exercise}
+
+\begin{exercise}
+  \begin{enumerate}
+  \item Show that \textbf{Rel} is a pointed category.
+  \item Show that the category \textbf{Grp} of groups has both an initial and a final object, and these are the same.
+  \item Show that the category \textbf{Ring} of unital rings has both an initial and a final object, and these are not the same.
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+Prove that any isomorphism is both monic and epic.
+\end{exercise}
+
+\newpage
+
+\begin{exercise}
+Let $A, B, C$ be objects in any category
+  and let $f : A \rightarrow B$ and $g : B \rightarrow C$
+  be maps in that category. Prove the following facts.
+  \[
+    \begin{tikzcd}
+      A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
+                                   & C &
+    \end{tikzcd}
+  \]
+
+  \begin{enumerate}
+  \item If both $f$ and $g$ are isomorphisms,
+    then $h$ is also an isomorphism.
+  \item If $h$ is monic, $f$ is monic.
+  \item If $h$ is epic, $g$ is epic.
+  \end{enumerate}
+\end{exercise}
+\begin{exercise}
+  Prove the following for a map $f \from A \to B$
+  in the category \textbf{Set},
+  \begin{enumerate}
+  \item If $f$ is monic, it is injective.
+  \item If $f$ is epic, it is surjective. \textit{(Challenge 1)}
+  \end{enumerate}
+\end{exercise}
+\begin{exercise}[Challenge 2]
+  Prove that in the category \textbf{Ring}, there are maps
+  which are epic but not surjective.
+  \begin{hint}
+    Consider the inclusion map $\mathbb{Z} \to \mathbb{Q}$, and
+    use the properties of ring homomorphisms.
+  \end{hint}
 \end{exercise}

### 2017-05-09 sheets/basis-2/sheet-basis-2.tex
Template for exercise sheets
Added some presheaf/contravariant excercises

--- 2017-04-21 sheets/basis-1/sheet-basis-1.tex
@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}
+
+\DeclareMathOperator{\A}{\mathscr A}
+\DeclareMathOperator{\B}{\mathscr B}
+\DeclareMathOperator{\M}{\mathscr M}
 
@@
 % Use Basis x or Talk x, where x is the number of the session
-\uebunghead{Basis 1}{What is a Category?}
+\uebunghead{Basis 2}{Maps between Categories -- Functors}
+
+\author{Mostafa Ahmed Abdelfattah Abouhamra, Felix Rech, Dominik Wagner}
 
 \begin{hint}
-  Read Chapter 1.1. For initial and terminal objects refer to Definition 2.17 on pages 48ff. For Monics and Epics, see the respective sections in Chapters 5.1 and 5.2.
+  Read Chapter 1.2.
 \end{hint}
 
-\section{Diagrams}
+\section{Contravariant Functors and Presheaves}
 
-\begin{hint}
-  This is how you can draw diagrams:
+\begin{exercise}
+  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories.
+Show that a contravariant functor $F:\A\rightarrow \B$ corresponds (uniquely) to a monotonically increasing function $f:A\rightarrow B$
+\end{exercise}
+
+\begin{exercise}
+  Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:
+
+  \begin{enumerate}
+  \item $\mathbf {Rel}^{\op}\cong \mathbf {Rel}$;
+  \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
+  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Not entirely trivial 1)}
+  \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
+  \end{enumerate}
+
+\end{exercise}
+
+%\section{Diagrams}
+%
+%\begin{hint}
+%  This is how you can draw diagrams:
+%  \[
+%    \begin{tikzcd}
+%                                    & A \arrow{dr}{g}    & \\
+%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+%    \end{tikzcd}
+%  \]
+%\end{hint}
+
+\section{Categories - Definition and Basics}
+
+\begin {exercise}
+Show there can be at most one inverse for a morphism $f \from A \to B$.
+\end{exercise}
+
+\begin {definition}{Rel}
+The objects of the category \textbf{Rel} are sets. The morphisms $f \from A \to B$ are subsets $f \sub A \times B$.
+The identity morphism on set $A$ is the equality relation $\{\round{a,a} \such a \in A\} := 1_A$.
+Composition of two morphisms $f \sub A \times B$, $g \sub B \times C$ is defined as
+\[ g \of f = \{ \round{a, c} \in A \times C \such \exists b. \round{a, b} \in f \land \round{b, c} \in g\} \]
+\end{definition}
+
+\begin {exercise}
+Show that \textbf{Rel} is a category.
+\end{exercise}
+
+\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
+(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A$).
+The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
+\end{definition}
+
+\begin {exercise}
+Show that \textbf{Pos} is a category.
+\end{exercise}
+
+% \begin{exercise}
+% For a fixed set X with Powerset $\mathcal{P}(X)$, does $\mathcal{P}(X) \cong \mathcal{P}(X)^{op}$ hold?
+% \end{exercise}
+
+\begin{definition}{Pointed Category} If an object is both \emph{initial} and \emph{terminal}, it is called a \emph{zero object}. A \emph{pointed category} is one with a zero object.
+\end{definition}
+
+\begin{exercise}
+  \begin{enumerate}
+  \item Show that \textbf{Rel} is a pointed category.
+  \item Show that the category \textbf{Grp} of groups has both an initial and a final object, and that these are the same.
+  \item Show that the category \textbf{Ring} of unital rings has both an initial and a final object, and that these are \emph{not} the same.
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+Show that any isomorphism is both monic and epic.
+\end{exercise}
+
+\newpage
+
+\begin{exercise}
+  Let $A, B, C$ be objects in a category and let $f \from A \to B$ and $g \from B \to C$ be morphisms such that the following diagram commutes:
   \[
     \begin{tikzcd}
-                                    & A \arrow{dr}{g}    & \\
-      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+      A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
+      & C &
     \end{tikzcd}
   \]
+  Show the following facts:
+  \begin{enumerate}
+  \item If both $f$ and $g$ are isomorphisms, then $h$ is also an isomorphism.
+  \item If $h$ is monic, then $f$ is monic.
+  \item If $h$ is epic, then $g$ is epic.
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+  Show the following for a morphism $f \from A \to B$ in the category \textbf{Set}:
+  \begin{enumerate}
+  \item If $f$ is monic, then it is injective.
+  \item If $f$ is epic, then it is surjective. \textit{(Challenge 1)}
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}[Challenge 2]
+  Show that in the category \textbf{Ring}, there are morphisms which are epic but not surjective.
+  \begin{hint}
+    Consider the inclusion morphism $\mathbb{Z} \to \mathbb{Q}$, and use the properties of ring homomorphisms.
   \end{hint}
+\end{exercise}
 

### 2017-05-10 sheets/basis-2/sheet-basis-2.tex
Added extended version of exercises.

--- 2017-04-21 sheets/basis-1/sheet-basis-1.tex
@@
 % Use Basis x or Talk x, where x is the number of the session
-\uebunghead{Basis 1}{What is a Category?}
+\uebunghead{Basis 2}{Maps between Categories -- Functors}
+
+\author{Mostafa Ahmed Abdelfattah Abouhamra, Felix Rech, Dominik Wagner}
 
 \begin{hint}
-  Read Chapter 1.1. For initial and terminal objects refer to Definition 2.17 on pages 48ff. For Monics and Epics, see the respective sections in Chapters 5.1 and 5.2.
+  Read Chapter 1.2.
 \end{hint}
 
-\section{Diagrams}
+\section{Maps between Categories -- Functors}
+
+\begin{exercise}
+  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories.
+Show that a contravariant functor $F:\A\rightarrow \B$ corresponds (uniquely) to a monotonically increasing function $f:A\rightarrow B$
+\end{exercise}
+
+\begin{exercise}
+  Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:
+
+  \begin{enumerate}
+  \item $\mathbf {Rel}^{\op}\cong \mathbf {Rel}$;
+  \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
+  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Not entirely trivial 1)}
+  \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
+  \end{enumerate}
+
+\end{exercise}
+
+\begin{definition}{Coslice Category}
+  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{coslice category} $A/\A$ of $\A$ under $A$ consists of
+  \begin{itemize}
+  \item the arrows to $A$ (i.e. the codomain of which are $A$) as objects
+  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $h\of f=g$.
+  \end{itemize}
+\end{definition}
+
+\begin{definition}{$\mathbf{Sets_*}$}
+  The category $\mathbf{Sets_*}$ of \emph{pointed sets} consists of
+  \begin{itemize}
+  \item pairs $(A,a)$ such that $A$ is a set and $a\in A$ as objects and
+  \item arrows objects between objects $(A,a)$ and $(B,b)$ are functions $f\from A\to B$ such that $f(a)=b$.
+  \end{itemize}
+\end{definition}
+
+\begin{exercise}
+Show that $\mathbf{Sets_*\cong 1/\mathbf{Sets}}$ where $1$ is an arbitrary set with exactly one element.  
+\end{exercise}
+
+\begin{definition}{Slice Category}
+  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{slice category} $\A/A$ of $\A$ over $A$ consists of
+  \begin{itemize}
+  \item the arrows from $A$ (i.e. the domain of which are $A$) as objects
+  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $g\of h=f$.
+  \end{itemize}
+\end{definition}
+
+\begin{exercise}[\textit{Challenge}]
+  Let $\A$ be a category. Construct a functor $F\from \A\to\mathbf{Cat}$ such that $F(A)=\A/A$ and show that this indeed a functor. Is this also an isomorphism?
 
   \begin{hint}
-  This is how you can draw diagrams:
-  \[
-    \begin{tikzcd}
-                                    & A \arrow{dr}{g}    & \\
-      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
-    \end{tikzcd}
-  \]
+    For arrows $f\in\A(A,B)$ consider functors $F_f\from\A/A\to\A/B$ that act on objects $g\in \A(C,A)$ by $F_f(g)=f\of g$. 
   \end{hint}
+\end{exercise}
+
+% \begin{exercise}[\textit{Challenge}]
+% Let $\A$ be a category. Consider again the functor $F_B\from \A\to\mathbf{Hom}(-,B)$ that is defined for a fixed $B\in\ob(A)$ by
+% \begin{align*}
+%   F_B(A)&=\A(A,B)\\
+%   F_B(f)&=F_{B,f}
+% \end{align*}
+% for all $A,A'\in\ob(\A)$, $f\in \A(A,A')$ and
+% \begin{align*}
+%   F_{B,f}(g)=g\of f
+% \end{align*}
+% for all $g\in\A(A,B)$.
+
+% Fill in the details of the construction of the functor $F\from \A^{\op} \to \mathbf{CAT}$ that satisfies $F(B)=\mathbf{Hom}(-,B)$, i.e.
+define its behaviour on arrows and verify that this is indeed a functor.
+
+% \begin{hint}
+%   Consider the functors $F_f\from\mathbf{Hom}(-,B)\to\mathbf{Hom}(-,B')$ for $f\in\A(B',B)$ given by
+%   \begin{align*}
+%     F_f(\A(A,B))&=\{f\of g\mid g\in \A(A,B)\}\\
+%     F_f(g)&=F_{f,g}
+%   \end{align*}
+%   where $g\in\mathbf{Hom}(-,B)(\A(A,B),\A(A',B))$ and the function $F_{f,g}$ is defined by
+%   \begin{align*}
+%     F_{f,g}(h)= g(f\of h).
+%   \end{align*}
+%   for each arrow $h\from\A(A,B')$.
+% \end{hint}
+
+% % Is this functor also an isomorphism
+
+% % given by
+% %   \begin{align*}
+% %     F(B)&=\mathbf{Hom}(-,B)\\
+% %     F(f)&=F_f
+% %   \end{align*}
+% % where $A\in\ob(\A)$ and $f\in\A(B',B)$, each functor $F_f\from \mathbf{Hom}(-,B)\to\mathbf{Hom}(-,B')$ is given by
+% % \begin{align*}
+% %   F_f(\A(A,B))&=\{f\of g\mid g\in \A(A,B)\}\\
+% %   F_f(g)&=F_{f,g}
+% % \end{align*}
+% % where $g\in\mathbf{Hom}(-,B)(\A(A,B),\A(A',B))$ and
+% % \begin{align*}
+% %   F_{f,g}(h)= g(f\of h).
+% % \end{align*}
+% % for each function $h\from\A(A,B)\to\A(A',B)$
+% % Check that this is indeed a functor, i.e. satisfies the axioms, and determine whether it is also an isomorphism.
+% \end{exercise}
+
+% %\section{Diagrams}
+% %
+% %\begin{hint}
+% %  This is how you can draw diagrams:
+% %  \[
+% %    \begin{tikzcd}
+% %                                    & A \arrow{dr}{g}    & \\
+% %      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+% %    \end{tikzcd}
+% %  \]
+% %\end{hint}
+
+% \section{Categories - Definition and Basics}
+
+% \begin {exercise}
+% Show there can be at most one inverse for a morphism $f \from A \to B$.
+% \end{exercise}
+
+% \begin {definition}{Rel}
+% The objects of the category \textbf{Rel} are sets. The morphisms $f \from A \to B$ are subsets $f \sub A \times B$.
+% The identity morphism on set $A$ is the equality relation $\{\round{a,a} \such a \in A\} := 1_A$.
+% Composition of two morphisms $f \sub A \times B$, $g \sub B \times C$ is defined as
+% \[ g \of f = \{ \round{a, c} \in A \times C \such \exists b. \round{a, b} \in f \land \round{b, c} \in g\} \]
+% \end{definition}
+
+% \begin {exercise}
+% Show that \textbf{Rel} is a category.
+% \end{exercise}
+
+% \begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
+% (recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A$).
+The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
+% \end{definition}
+
+% \begin {exercise}
+% Show that \textbf{Pos} is a category.
+% \end{exercise}
+
+% % \begin{exercise}
+% % For a fixed set X with Powerset $\mathcal{P}(X)$, does $\mathcal{P}(X) \cong \mathcal{P}(X)^{op}$ hold?
+% % \end{exercise}
+
+% \begin{definition}{Pointed Category} If an object is both \emph{initial} and \emph{terminal}, it is called a \emph{zero object}. A \emph{pointed category} is one with a zero object.
+% \end{definition}
+
+% \begin{exercise}
+%   \begin{enumerate}
+%   \item Show that \textbf{Rel} is a pointed category.
+%   \item Show that the category \textbf{Grp} of groups has both an initial and a final object, and that these are the same.
+%   \item Show that the category \textbf{Ring} of unital rings has both an initial and a final object, and that these are \emph{not} the same.
+%   \end{enumerate}
+% \end{exercise}
+
+% \begin{exercise}
+% Show that any isomorphism is both monic and epic.
+% \end{exercise}
+
+% \newpage
+
+% \begin{exercise}
+%   Let $A, B, C$ be objects in a category and let $f \from A \to B$ and $g \from B \to C$ be morphisms such that the following diagram commutes:
+%   \[
+%     \begin{tikzcd}
+%       A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
+%       & C &
+%     \end{tikzcd}
+%   \]
+%   Show the following facts:
+%   \begin{enumerate}
+%   \item If both $f$ and $g$ are isomorphisms, then $h$ is also an isomorphism.
+%   \item If $h$ is monic, then $f$ is monic.
+%   \item If $h$ is epic, then $g$ is epic.
+%   \end{enumerate}
+% \end{exercise}
+
+% \begin{exercise}
+%   Show the following for a morphism $f \from A \to B$ in the category \textbf{Set}:
+%   \begin{enumerate}
+%   \item If $f$ is monic, then it is injective.
+%   \item If $f$ is epic, then it is surjective. \textit{(Challenge 1)}
+%   \end{enumerate}
+% \end{exercise}
+
+% \begin{exercise}[Challenge 2]
+%   Show that in the category \textbf{Ring}, there are morphisms which are epic but not surjective.
+%   \begin{hint}
+%     Consider the inclusion morphism $\mathbb{Z} \to \mathbb{Q}$, and use the properties of ring homomorphisms.
+%   \end{hint}
+% \end{exercise}
 
@@
 
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:
+% %%% Local Variables:
+% %%% mode: latex
+% %%% TeX-master: t
+% %%% End:

### 2017-05-10 sheets/basis-2/sheet-basis-2.tex
Add some exercises

@@
 \end{exercise}
+
+\begin{exercise}
+Show that functors preserve isomorphism. That is, prove that if $F : \cat{A} \to \cat{B}$ is a functor and $A, A' \in \cat{A}$ with $A \iso A'$, then $F(A) \iso F(A')$.
+\end{exercise}
+
+\begin{exercise}
+  Find an example of a functor $F : \cat{A} \to \cat{B}$ such that $F$ is faithful but there exist distinct maps $f_1$ and $f_2$ in $\cat{A}$ with $F(f_1) = F(f_2)$.
+\end{exercise}
+
+\begin{exercise}
+  Find interesting examples for a functor that is both full and faithful, one that is full but not faithful, one that is fathful but not full, and one that is neither.
+\end{exercise}
+
+\begin{exercise}
+  \begin{enumerate}
+    \item What are the subcategories of an ordered set? Which are full?
+    \item What are the subcategories of a group? (Careful!)
+      Which are full?
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+  \begin{enumerate}
+    \item Define an endofunctor on $\Set$ that maps every set to its powerset.
+      Check that is satisfies the functor laws.
+      Is it full/faithful?
+    \item Define a contravariant endofunctor on $\Set$, i.e.\ a functor from $\Set^\op$ to $\Set$, that maps every set to its powerset.
+      Check that is satisfies the functor laws.
+      Is it full/faithful?
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+  Check that there is a category $\Cat$ with categories as objects and functors as arrows.
+\end{exercise}
+
+\begin{definition}{Free Monoid}
+  For every set $A$, we define the \emph{free monoid} $A^* \coloneqq \lbrace \text{words over $A$} \rbrace$.
+  We define the binary operation by $w \cdot w' \coloneqq ww'$ as the concatenation of words and use the empty word $\varepsilon$ as neutral element.
+\end{definition}
+
+\begin{exercise}
+  Show that for every set $A$, the free monoid $A^*$ satisfies the following \emph{universal mapping property}:
+  There is a function $i : A \to A^*$ and given any monoid $M$ and any function $j : A \to M$, there is a \emph{unique} monoid homomorphism $f : A^* \to M$ such that $f \circ i = j$.
+\end{exercise}
+
+\begin{exercise}
+  Denote by $\Mon$ the category of monoids with monoid morphisms as arrows.
+  Define a functor $F : \Set \to \Mon$ that maps every set $A$ to the free monoid $A^*$.
+  Is $F$ full/faithful?
+\end{exercise}
+
+\begin{exercise}
+  \begin{enumerate}
+    \item A directed graph consists of a set of vertices $V$, a set of edges $E$ and two functions $s, t : E \to V$ that assign a source and target vertex to each edge.
+      Define a category $\cat{A}$ such that functors from $\cat{A}$ to $\Set$ correspond to directed graphs.
+    \item Define a category $\cat{B}$ such that the functors from $\cat{B}$ to $\Set$ correspond to injections between an arbitrary pair of sets.
+  \end{enumerate}
+\end{exercise}
+
+\end{document}
 

### 2017-05-10 sheets/basis-2/sheet-basis-2.tex
Some changes to order and fixes of typos.

@@
 
-\section{Maps between Categories -- Functors}
-
 \begin{exercise}
-  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories.
-Show that a contravariant functor $F:\A\rightarrow \B$ corresponds (uniquely) to a monotonically increasing function $f:A\rightarrow B$
+Show that functors preserve isomorphism. That is, prove that if $F : \cat{A} \to \cat{B}$ is a functor and $A, A' \in \cat{A}$ with $A \iso A'$, then $F(A) \iso F(A')$.
 \end{exercise}
@@
 \begin{exercise}
-  Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:
-
-  \begin{enumerate}
-  \item $\mathbf {Rel}^{\op}\cong \mathbf {Rel}$;
-  \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
-  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Not entirely trivial 1)}
-  \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
-  \end{enumerate}
-
-\end{exercise}
-
-\begin{definition}{Coslice Category}
-  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{coslice category} $A/\A$ of $\A$ under $A$ consists of
-  \begin{itemize}
-  \item the arrows to $A$ (i.e. the codomain of which are $A$) as objects
-  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $h\of f=g$.
-  \end{itemize}
-\end{definition}
-
-\begin{definition}{$\mathbf{Sets_*}$}
-  The category $\mathbf{Sets_*}$ of \emph{pointed sets} consists of
-  \begin{itemize}
-  \item pairs $(A,a)$ such that $A$ is a set and $a\in A$ as objects and
-  \item arrows objects between objects $(A,a)$ and $(B,b)$ are functions $f\from A\to B$ such that $f(a)=b$.
-  \end{itemize}
-\end{definition}
-
-\begin{exercise}
-Show that $\mathbf{Sets_*\cong 1/\mathbf{Sets}}$ where $1$ is an arbitrary set with exactly one element.  
-\end{exercise}
-
-\begin{definition}{Slice Category}
-  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{slice category} $\A/A$ of $\A$ over $A$ consists of
-  \begin{itemize}
-  \item the arrows from $A$ (i.e. the domain of which are $A$) as objects
-  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $g\of h=f$.
-  \end{itemize}
-\end{definition}
-
-\begin{exercise}[\textit{Challenge}]
-  Let $\A$ be a category. Construct a functor $F\from \A\to\mathbf{Cat}$ such that $F(A)=\A/A$ and show that this indeed a functor. Is this also an isomorphism?
-
-  \begin{hint}
-    For arrows $f\in\A(A,B)$ consider functors $F_f\from\A/A\to\A/B$ that act on objects $g\in \A(C,A)$ by $F_f(g)=f\of g$. 
-  \end{hint}
-\end{exercise}
-
-\begin{exercise}
-Show that functors preserve isomorphism. That is, prove that if $F : \cat{A} \to \cat{B}$ is a functor and $A, A' \in \cat{A}$ with $A \iso A'$, then $F(A) \iso F(A')$.
+  Check that there is a category $\Cat$ with categories as objects and functors as arrows.
 \end{exercise}
@@
 \begin{exercise}
-  Find interesting examples for a functor that is both full and faithful, one that is full but not faithful, one that is fathful but not full, and one that is neither.
+  Find interesting examples for a functor that is both full and faithful, one that is full but not faithful, one that is faithful but not full, and one that is neither.
 \end{exercise}
@@
     \item Define an endofunctor on $\Set$ that maps every set to its powerset.
-      Check that is satisfies the functor laws.
+      Check that it satisfies the functor laws.
       Is it full/faithful?
     \item Define a contravariant endofunctor on $\Set$, i.e.\ a functor from $\Set^\op$ to $\Set$, that maps every set to its powerset.
-      Check that is satisfies the functor laws.
+      Check that it satisfies the functor laws.
       Is it full/faithful?
@@
 \begin{exercise}
-  Check that there is a category $\Cat$ with categories as objects and functors as arrows.
+  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories.
+Consider a contravariant functor $F:\A\rightarrow \B$.
+Describe properties of the corresponding function $f:A\rightarrow B$.
+\end{exercise}
+
+\begin{exercise}
+  Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:
+
+  \begin{enumerate}
+  \item $\mathbf {Rel}^{\op}\cong \mathbf {Rel}$;
+  \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
+  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Not entirely trivial 1)}
+  \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
+  \end{enumerate}
+
 \end{exercise}
@@
   \end{enumerate}
+\end{exercise}
+
+\begin{definition}{Coslice Category}
+  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{coslice category} $A/\A$ of $\A$ under $A$ consists of
+  \begin{itemize}
+  \item the arrows to $A$ (i.e. the domain of which is $A$) as objects
+  \item and for $f\in \A(A,B)$, $g\in\A(A,B')$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $h\of f=g$.
+  \end{itemize}
+\end{definition}
+
+\begin{definition}{$\mathbf{Sets_*}$}
+  The category $\mathbf{Sets_*}$ of \emph{pointed sets} consists of
+  \begin{itemize}
+  \item pairs $(A,a)$ such that $A$ is a set and $a\in A$ as objects and
+  \item arrows between objects $(A,a)$ and $(B,b)$ are functions $f\from A\to B$ such that $f(a)=b$.
+  \end{itemize}
+\end{definition}
+
+\begin{exercise}
+Show that $\mathbf{Sets_*\cong 1/\mathbf{Sets}}$ where $1$ is an arbitrary set with exactly one element.  
+\end{exercise}
+
+\begin{definition}{Slice Category}
+  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{slice category} $\A/A$ of $\A$ over $A$ consists of
+  \begin{itemize}
+  \item the arrows from $A$ (i.e. the codomain of which is $A$) as objects
+  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $g\of h=f$.
+  \end{itemize}
+\end{definition}
+
+\begin{exercise}[\textit{Not entirely trivial 2}]
+  Let $\A$ be a category. Construct a functor $F\from \A\to\mathbf{Cat}$ such that $F(A)=\A/A$ and show that this is indeed a functor. Is this also an isomorphism?
+
+  \begin{hint}
+    For arrows $f\in\A(A,B)$ consider functors $F_f\from\A/A\to\A/B$ that act on objects $g\in \A(C,A)$ by $F_f(g)=f\of g$. 
+  \end{hint}
 \end{exercise}

### 2017-05-11 sheets/basis-1/sheet-basis-1.tex

--- 2017-04-21 sheets/basis-1/sheet-basis-1.tex
@@
 
+\author{Sarah Mameche, Andreas Meyer, Leonhard Staut}
+
 \begin{hint}
-  Read Chapter 1.1. For initial and terminal objects refer to Definition 2.17 on pages 48ff. For Monics and Epics, see the respective sections in Chapters 5.1 and 5.2.
+  Read Chapter 1.1. For initial and terminal objects refer to Definition 2.17 on pages 48ff. For Monics and Epics, refer to the respective sections in Chapters 5.1 and 5.2.
 \end{hint}
 
-\section{Diagrams}
+%\section{Diagrams}
+%
+%\begin{hint}
+%  This is how you can draw diagrams:
+%  \[
+%    \begin{tikzcd}
+%                                    & A \arrow{dr}{g}    & \\
+%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+%    \end{tikzcd}
+%  \]
+%\end{hint}
 
-\begin{hint}
-  This is how you can draw diagrams:
+\section{Categories - Definition and Basics}
+
+\begin {exercise}
+Show there can be at most one inverse for a morphism $f \from A \to B$.
+\end{exercise}
+
+\begin {definition}{Rel}
+The objects of the category \textbf{Rel} are sets. The morphisms $f \from A \to B$ are subsets $f \sub A \times B$.
+The identity morphism on set $A$ is the equality relation $\{\round{a,a} \such a \in A\} := 1_A$.
+Composition of two morphisms $f \sub A \times B$, $g \sub B \times C$ is defined as
+\[ g \of f = \{ \round{a, c} \in A \times C \such \exists b. \round{a, b} \in f \land \round{b, c} \in g\} \]
+\end{definition}
+
+\begin {exercise}
+Show that \textbf{Rel} is a category.
+\end{exercise}
+
+\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
+(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A$).
+The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
+\end{definition}
+
+\begin {exercise}
+Show that \textbf{Pos} is a category.
+\end{exercise}
+
+% \begin{exercise}
+% For a fixed set X with Powerset $\mathcal{P}(X)$, does $\mathcal{P}(X) \cong \mathcal{P}(X)^{op}$ hold?
+% \end{exercise}
+
+\begin{definition}{Pointed Category} If an object is both \emph{initial} and \emph{terminal}, it is called a \emph{zero object}. A \emph{pointed category} is one with a zero object.
+\end{definition}
+
+\begin{exercise}
+  \begin{enumerate}
+  \item Show that \textbf{Rel} is a pointed category.
+  \item Show that the category \textbf{Grp} of groups has both an initial and a final object, and that these are the same.
+  \item Show that the category \textbf{Ring} of unital rings has both an initial and a final object, and that these are \emph{not} the same.
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+Show that any isomorphism is both monic and epic.
+\end{exercise}
+
+\newpage
+
+\begin{exercise}
+  Let $A, B, C$ be objects in a category and let $f \from A \to B$ and $g \from B \to C$ be morphisms such that the following diagram commutes:
   \[
     \begin{tikzcd}
-                                    & A \arrow{dr}{g}    & \\
-      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+      A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
+      & C &
     \end{tikzcd}
   \]
+  Show the following facts:
+  \begin{enumerate}
+  \item If both $f$ and $g$ are isomorphisms, then $h$ is also an isomorphism.
+  \item If $h$ is monic, then $f$ is monic.
+  \item If $h$ is epic, then $g$ is epic.
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+  Show the following for a morphism $f \from A \to B$ in the category \textbf{Set}:
+  \begin{enumerate}
+  \item If $f$ is monic, then it is injective.
+  \item If $f$ is epic, then it is surjective. \textit{(Challenge 1)}
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}[Challenge 2]
+  Show that in the category \textbf{Ring}, there are morphisms which are epic but not surjective.
+  \begin{hint}
+    Consider the inclusion morphism $\mathbb{Z} \to \mathbb{Q}$, and use the properties of ring homomorphisms.
   \end{hint}
+\end{exercise}
 

### 2017-05-11 sheets/basis-2/sheet-basis-2.tex

@@
   \begin{itemize}
-  \item the arrows to $A$ (i.e. the domain of which is $A$) as objects
+  \item the arrows from $A$ (i.e. the domain of which is $A$) as objects
   \item and for $f\in \A(A,B)$, $g\in\A(A,B')$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $h\of f=g$.
@@
   \begin{itemize}
-  \item the arrows from $A$ (i.e. the codomain of which is $A$) as objects
-  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $g\of h=f$.
+  \item the arrows to $A$ (i.e. the codomain of which is $A$) as objects
+  \item and for $f\in \A(B,A)$, $g\in\A(B',A)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $g\of h=f$.
   \end{itemize}

### 2017-05-11 sheets/basis-2/sheet-basis-2.tex
small stilistic change.

@@
   \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
-  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Not entirely trivial 1)}
+  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Challenge 1)}
   \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
@@
 
-\begin{exercise}[\textit{Not entirely trivial 2}]
+\begin{exercise}[\textit{Challenge 2}]
   Let $\A$ be a category. Construct a functor $F\from \A\to\mathbf{Cat}$ such that $F(A)=\A/A$ and show that this is indeed a functor. Is this also an isomorphism?

### 2017-05-15 sheets/basis-2/sheet-basis-2.tex
Add answer environment for testing
Solved odd exercises by Felix.

--- 2017-05-11 sheets/basis-2/sheet-basis-2.tex
@@
 
+\begin{answer}
+  Since $A\cong A'$ there are $f\in\mathscr A(A,A')$ and $g\in\mathscr A(A',A)$ such that $g \of f=1_A$ and $f \of g=1_A'$.
+Hence, $F(g)\of F(f)=F(g\of f)=F(1_A)=1_{F(A)}$ and $F(f)\of F(g)=F(f\of g)=F(1_A')=1_{F(A')}$.
+As a consequence, $F(f)$ is an isomorphism between $F(A)$ and $F(A')$.
+In particular, $F(A)\cong F(A')$.
+\end{answer}
+
 \begin{exercise}
@@
 \end{exercise}
+
+\begin{answer}
+  Suppose the discrete categories $\A$ and $\B$ with two and one objects respectively.
+Then, any functor $F\from A\to B$ is trivially faithful but $F(1_A)=F(1_B)$ (where $A$ and $B$ are the only objects of $\A$).
+\end{answer}
 
@@
 \end{exercise}
+
+\begin{answer}
+  \begin{enumerate}
+  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$.
+  \item The subcategories of a group are either the category with no objects or submonoids of the group (which is of course a monoid).
+The only full subcategories are the group itself and the category with no objects.
+  \end{enumerate}
+\end{answer}
 
@@
 
+\begin{answer}
+  Define
+  \begin{align*}
+    i(a)&=a &(a\in A)
+  \end{align*}
+  and
+  \begin{align*}
+    f(\epsilon)&=0_M&\\
+    f(a_1\cdots a_{n-1}a_n)&=f(a_1\cdots a_{n-1})\of j(a_n) &\text{for }n>0.
+  \end{align*}
+  Clearly, for each $a\in A$, $f(i(a))=j(a)$, thus $f\of i=j$, and $f$ is a group homomorphism.
+
+  Now, suppose $g$ is a group homomorphism satisfying $g\of i=j$. 
+  Let $a_1\cdots a_n\in A^*$. Proof by induction on $n$ that $g(n)=f(n)$.
+
+  Base case: $g(\epsilon)=0_M$ because $\epsilon$ is the neutral element in $A^*$).
+
+  Inductive step: 
+  \begin{align*}
+    g(a_1\cdots a_{n-1}a_n)&=g(a_1\cdots a_{n-1})\of g(a_n)&g\text{ group homomorphism}\\
+                           &=f(a_1\cdots a_{n-1})\of g(a_n)&\text{inductive hypothesis}\\
+                           &=f(a_1\cdots a_{n-1})\of g(i(a_n)) & a_n\in A\\
+                           &=f(a_1\cdots a_{n-1})\of f(i(a_n)) & g\of i=j=f\of i\\
+                           &=f(a_1\cdots a_{n-1})\of f(a_n) & a_n\in A\\
+                           &=f(a_1\cdots a_{n-1}a_n)&f\text{ group homomorphism}
+  \end{align*}
+  Hence, $f=g$, so $f$ is indeed unique.
+
+\end{answer}
+
 \begin{exercise}
@@
 \end{exercise}
+
+\begin{answer}
+  \begin{enumerate}
+  \item Consider $\A$ given by
+    \[
+      \begin{tikzcd}[column sep=huge,row sep=huge]
+        E\arrow[r,yshift=.85ex,"s"] \arrow[r,below,yshift=-.85ex,"e"]&V
+      \end{tikzcd}
+    \]
+    Then a functor $F\from \A\to\Set$ assigns to both $E$ and $V$ sets $E'$ and $V'$ (which correspond to the set of edges and vertices respectively) and to $s,e$ functions $s'$ and $e'$ that take edges $d$ from $E'$ and return elements $s'(d)$ and $e'(d)$ which correspond to the start and end node of the edge $d$ respectively.
+  \item Consider $\B$ given by
+    \[
+      \begin{tikzcd}[column sep=huge,row sep=huge]
+        A\arrow[r,yshift=.85ex,"f"] & B\arrow[l,below,yshift=-.85ex,"g"]
+      \end{tikzcd}
+    \]
+    such that $g\of f=1_A$ and $f\of g=1_B$.
+  \end{enumerate}
+\end{answer}
 

### 2017-05-17 sheets/basis-3/sheet-basis-3.tex
Template for exercise sheets
Add exercises for the third sheet

--- 2017-04-21 sheets/basis-1/sheet-basis-1.tex
@@
 
+\author{Joachim Bard, Maximilian Wuttke, Nikita Ziuzin}
+
 % Use Basis x or Talk x, where x is the number of the session
-\uebunghead{Basis 1}{What is a Category?}
+\uebunghead{Basis 3}{Maps between Functors -- Natural Transformations}
 
 \begin{hint}
-  Read Chapter 1.1. For initial and terminal objects refer to Definition 2.17 on pages 48ff. For Monics and Epics, see the respective sections in Chapters 5.1 and 5.2.
+  Read Chapter 1.3.
 \end{hint}
 
-\section{Diagrams}
+\begin{exercise}
+	Let $\cat{A}$ and $\cat{B}$ be categories.  Prove that
+	$\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$.
+\end{exercise}
 
-\begin{hint}
-  This is how you can draw diagrams:
+\begin{exercise}
+
+  \begin{definition}{Canonical according to Leinster}
+     "Canonical is an informal word, meaning something like 'God-given' or
+    'defined without making arbitrary choices'.  For example, for any two sets
+    $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
+    defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
+    \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
+    defined by 'choose an element $a_0 \in A$ and send everything to $a_0$' is
+    not canonical, because the choice of $a_0$ is arbitrary."
+  \end{definition}
+	Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
+	to $B$.  Write down:
+	\begin{itemize}
+		\item[(a)] a canonical function $A \to B^{(B^A)}$,
+		\item[(b)] a canonical function $A \times B^{A} \to B$.
+	\end{itemize}
+\end{exercise}
+
+\begin{exercise}
+  For a fixed set $A$, the map taking $B$ to $B^A \times A$ can be extended to
+  a functor $F_A:\mathbf{Set} \to \mathbf{Set}$ as follows:
+  \begin{align*}
+    F_A(B) & =  B^A \times A\\
+    F_A(f: B \to C) & =  (F \circ \_) \times id_A
+  \end{align*}
+  The fact that $eval: F_A \to I_{\mathbf{Set}}$ is a natural transformation follows from the commutativity of the diagram
 \[
   \begin{tikzcd}
-                                    & A \arrow{dr}{g}    & \\
-      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+    F_A(C) = C^A \times A \arrow{rrr}{eval_{AC}} \arrow{dd}{F_A(g)=(g\circ \_) \times id_A} & & & C = I_{\mathbf{Set}}(C) \arrow{dd}{g = I_{\mathbf{Set}(g)}} \\
+                                                                                           & & & \\
+    F_A(B) = B^A \times A \arrow{rrr}{eval_{AB}} & & & B = I_{\mathbf{Set}}(B)
   \end{tikzcd}
 \]
-\end{hint}
+for each $g: C \to B$.
+
+Check that this diagram commutes.
+\end{exercise}
+
+\begin{exercise}
+	Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
+	\begin{itemize}
+		\item[(a)] Suppose $\alpha$ is a natural isomorphism.
+		Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
+		\item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
+		$A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
+	\end{itemize}
+\end{exercise}
+
+\begin{exercise}
+	Let the following natural transformations be given:
+	\[
+\xymatrix{
+\cat{A}
+\ruppertwocell<8>^F{\alpha}
+\ar[r]|G
+\rlowertwocell<-8>_H{\beta} &
+\cat{A}'
+\ruppertwocell<8>^{F'}{\hspace{.2em}\alpha'}
+\ar[r]|{G'}
+\rlowertwocell<-8>_{H'}{\hspace{.2em}\beta'} &
+\cat{A}''
+}
+\]
+Now prove that the interchange law holds, that is
+\[
+(\beta' \of \alpha') * (\beta \of \alpha)
+=
+(\beta' * \beta) \of (\alpha' * \alpha)
+\]
+\end{exercise}
+
+\begin{exercise}
+A \emph{permutation} of a set $X$ is a bijection $X \to X$.
+Write $\Sym(X)$ for the set of permutations of $X$.
+A \emph{total order} on a set $X$ is an order $\leq$ such that for all $x, y \in X$,
+either $x \leq y$ or $y \leq x$.
+Write $\Ord(X)$ for the set of total orders on $X$.
+
+Let $\cat{B}$ denote the category of finite sets and bijections.
+
+	\begin{itemize}
+	\item[(a)]
+Give a definition of $\Sym$ on maps in $\cat{B}$ in such a way that $\Sym$
+becomes a functor $\cat{B} \to \Set$.  Do the same for $\Ord$.  Both your
+definitions should be canonical (no arbitrary choices).
+
+	\item[(b)]
+Show that there is no natural transformation $\Sym \to \Ord$.
+(Hint: consider identity permutations.)
+
+	\item[(c)]
+For an $n$-element set $X$, how many elements do the sets $\Sym(X)$
+and $\Ord(X)$ have?
+	\end{itemize}
+
+Conclude that $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$, but not
+\emph{naturally} in $X \in \cat{B}$.  (The moral is that for each finite
+set $X$, there are exactly as many permutations of $X$ as there are total
+orders on $X$, but there is no natural way of matching them up.)
+\end{exercise}
+
+\begin{exercise}
+Let $F\from \cat{A} \to \cat{B}$ be a functor.
+
+	\begin{itemize}
+	\item[(a)]
+Suppose that $F$ is an equivalence.  Prove that $F$ is full, faithful and
+essentially surjective on objects.  (Hint: prove faithfulness before
+fullness.)
+
+	\item[(b)]
+Now suppose instead that $F$ is full, faithful and essentially surjective
+on objects.  For each $B \in \cat{B}$, choose an object $G(B)$ of $\cat{A}$
+and an isomorphism $\varepsilon\from F(G(B)) \to B$.  Prove that $G$ extends
+to a functor in such a way that $(\varepsilon)_{B \in \cat{B}}$ is a natural
+isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
+$1_{\cat{A}} \to G \circ F$, thus proving that $F$ is an equivalence.
+	\end{itemize}
+\end{exercise}
+
+\begin{exercise}
+This exercise makes precise the idea that linear algebra can equivalently
+be done with matrices or with linear maps.
+
+Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
+numbers and with
+\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
+Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
+finite-dimensional vector spaces over $K$.
+Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
+\end{exercise}
 

### 2017-05-18 sheets/basis-2/sheet-basis-2.tex
Fixed part of my solution.

@@
       \begin{tikzcd}[column sep=huge,row sep=huge]
-        A\arrow[r,yshift=.85ex,"f"] & B\arrow[l,below,yshift=-.85ex,"g"]
+        A\arrow[r,yshift=.85ex,"f"] & B\arrow[l,below,yshift=-.85ex,"g"] \arrow[loop right,"h"]
       \end{tikzcd}
     \]
-    such that $g\of f=1_A$ and $f\of g=1_B$.
+    (plus the identities) such that $g\of f=1_A$, $f\of g=h$, $g\of h=g$ and $g\of f = h$.
+Then any functor $F\from\A\to\Set$ maps $A$ and $B$ to sets $A'$ and $B'$ respectively and $f$ is mapped to a function $f\from A'\to B'$ and $g$ is mapped to a function $g\from B'\to A'$ such that for all $a\in A'$, $g(f(a))=a$.
+Thus $f$ is an injection.
   \end{enumerate}

### 2017-05-18 sheets/basis-2/sheet-basis-2.tex
Fix bug in solution.
Fix more bugs in solution.

--- 2017-05-15 sheets/basis-2/sheet-basis-2.tex
@@
       \begin{tikzcd}[column sep=huge,row sep=huge]
-        A\arrow[r,yshift=.85ex,"f"] & B\arrow[l,below,yshift=-.85ex,"g"]
+        A\arrow[r,yshift=.85ex,"f"] & B\arrow[l,below,yshift=-.85ex,"g"] \arrow[loop right,"h"]
       \end{tikzcd}
     \]
-    such that $g\of f=1_A$ and $f\of g=1_B$.
+    (plus the identities) such that $g\of f=1_A$, $f\of g=h$, $g\of h=g$ and $h\of f=f$.
+Then any functor $F\from\A\to\Set$ maps $A$ and $B$ to sets $A'$ and $B'$ respectively and $f$ is mapped to a function $f'\from A'\to B'$ and $g$ is mapped to a function $g'\from B'\to A'$ such that for all $a\in A'$, $g'(f'(a))=a$.
+Thus $f$ is an injection.
   \end{enumerate}

### 2017-05-18 sheets/basis-2/sheet-basis-2.tex
More extensions.
Yet more fixes.

@@
   \begin{enumerate}
-  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$.
+  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$. In the case of full subcategories even $\leq'\;=\;\leq$ holds.
   \item The subcategories of a group are either the category with no objects or submonoids of the group (which is of course a monoid).
@@
   Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories.
-Consider a contravariant functor $F:\A\rightarrow \B$.
+Consider a contravariant functor $F:\A^{\op}\rightarrow \B$.
 Describe properties of the corresponding function $f:A\rightarrow B$.
@@
   Now, suppose $g$ is a group homomorphism satisfying $g\of i=j$. 
-  Let $a_1\cdots a_n\in A^*$. Proof by induction on $n$ that $g(n)=f(n)$.
+  Let $a_1\cdots a_n\in A^*$. Proof by induction on $n$ that $g(a_1\cdots a_n)=f(a_1\cdots a_n)$.
 
-  Base case: $g(\epsilon)=0_M$ because $\epsilon$ is the neutral element in $A^*$).
+  Base case: $g(\epsilon)=0_M=f(\epsilon)$ because $\epsilon$ is the neutral element in $A^*$).
 
@@
 Then any functor $F\from\A\to\Set$ maps $A$ and $B$ to sets $A'$ and $B'$ respectively and $f$ is mapped to a function $f'\from A'\to B'$ and $g$ is mapped to a function $g'\from B'\to A'$ such that for all $a\in A'$, $g'(f'(a))=a$.
-Thus $f$ is an injection.
+Thus $f'$ is an injection.
+
+    Indeed, the reverse is also true: Given any injection $f'\from A'\to B'$ there is functor $F\from \A\to\Set$ such that $f'$ corresponds to $F$ (in the way specified above).
   \end{enumerate}

### 2017-05-18 sheets/basis-2/sheet-basis-2.tex
Fixed further bug.

@@
   \begin{enumerate}
-  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$. In the case of full subcategories even $\leq'\;=\;\leq$ holds.
+  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$.
+In the case of full subcategories even for all $s,t\in S'$, $s\leq t$ if and only if $s\leq' t$ holds.
   \item The subcategories of a group are either the category with no objects or submonoids of the group (which is of course a monoid).

### 2017-05-18 sheets/basis-3/sheet-basis-3.tex
minor stylistic changes to sheet 3.
Remove exercise 5 and add a new exercise about category equivialences to the top

@@
 \begin{exercise}
+  Define categories $\cat{A}$ and $\cat{B}$ such that either $\cat{A} \iso \cat{B}$ in $\CAT$ or $\cat{A} \eqv \cat{B}$, but not both.
+  Do not use the example given during the lecture.
+\end{exercise}
+
+\begin{exercise}
         Let $\cat{A}$ and $\cat{B}$ be categories.  Prove that
@@
 \end{exercise}
-
-\begin{exercise}
 
@@
 \end{definition}
+
+\begin{exercise}
   Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
@@
         \end{itemize}
-\end{exercise}
-
-\begin{exercise}
-	Let the following natural transformations be given:
-	\[
-\xymatrix{
-\cat{A}
-\ruppertwocell<8>^F{\alpha}
-\ar[r]|G
-\rlowertwocell<-8>_H{\beta} &
-\cat{A}'
-\ruppertwocell<8>^{F'}{\hspace{.2em}\alpha'}
-\ar[r]|{G'}
-\rlowertwocell<-8>_{H'}{\hspace{.2em}\beta'} &
-\cat{A}''
-}
-\]
-Now prove that the interchange law holds, that is
-\[
-(\beta' \of \alpha') * (\beta \of \alpha)
-=
-(\beta' * \beta) \of (\alpha' * \alpha)
-\]
 \end{exercise}

### 2017-05-18 sheets/basis-2/sheet-basis-2.tex
Add solutions for exercises by Dominik

--- 2017-05-18 sheets/basis-2/sheet-basis-2.tex
@@
 
-\DeclareMathOperator{\A}{\mathscr A}
-\DeclareMathOperator{\B}{\mathscr B}
-\DeclareMathOperator{\M}{\mathscr M}
+\newcommand{\A}{\cat A}
+\newcommand{\B}{\cat B}
+\newcommand{\M}{\cat M}
+\newcommand{\Rel}{\fcat{Rel}}
 
@@
 
+\begin{answer}
+  For every arrow $x \to y$ in $\A$, there is an arrow $f : y \to x$ in $\A^\op$ and hence an arrow $F(f) : F(y) \to F(x)$ in $\B$.
+  In the language of posets this means that $x \leq y$ implies $f(x) \geq f(y)$, i.e.\ $f$ is monotonically decreasing.
+  The functor laws hold tautologically in posets and don't tell us anything about $f$.
+\end{answer}
+
 \begin{exercise}
@@
 \end{exercise}
+
+\begin{answer}
+  \begin{enumerate}
+    \item The categories are isomorphic.
+      We define functors $F : \Rel^\op \to \Rel$ and $G : \Rel \to \Rel^\op$ in both directions by
+      \begin{align*}
+        F(X) &\coloneqq X  &  G(X) &\coloneqq X \\
+        F(R) &\coloneqq R^{-1}  &  G(R) &\coloneqq R^{-1}
+      \end{align*}
+      for all sets $X$ and $Y$ and relations $R \in X \times Y$.
+      Both functors are mutually inverse bacause $(R^{-1})^{-1} = R$.
+      Thus $\Rel^\op \iso \Rel$.
+
+    \item We refute the statement by contradiction.
+      Assume that we have an isomorphism $F : \Set^\op \to \Set$.
+      Let $1$ be a one-element set.
+      It is final in $\Set$ and hence initial in $\Set^\op$ and will be mapped to the initial object $\emptyset$ in $\Set$, because isomorphisms preserve initial objects.
+      We define $f : 1 \to X$ as a function to some set $X$ that is not isomorphic to $1$.
+      Then $F(X)$ is not isomorphic to $F(1) = \emptyset$, but there is a function $F(f) : F(X) \to \emptyset$, which leads to a contradiction.
+
+    \item Let $X$ be a finite set with at least two elements.
+      We define the monoid $\M \coloneqq \lbrace f : X \to X \rbrace$ with function composition as binary operation and the identity function as neutral element.
+      Let $c \in \M$ be some constant function.
+      Such a function is left-absorbing, which means that for all functions $f \in \M$, we have $c \circ f = c$.
+      Left-absorbing elements in the opposite monoid correspond to right-absorbing functions, i.e., fuctions $d \in \M$ such that for all $f \in M$, we have $f \circ d = d$.
+      Obviously such functions do not exist.
+      Since left-absorbing elements are preserved by isomorphisms, $\M$ cannot be isomorphic to its opposite monoid.
+
+    \item We define mutually inverse functors $F : \CAT_i^\op \to \CAT_i$ and $G : \CAT_i \to \CAT_i^\op$ by
+      \begin{align*}
+        F(\A) &\coloneqq \A  &  G(\A) &\coloneqq \A \\
+        F(H) &\coloneqq H^{-1}  &  G(H) &\coloneqq H^{-1}
+      \end{align*}
+      for all categories $\A$ and $\B$ and functors $H : \A \to B$.
+      The functors $F$ and $G$ are mutually inverse because $(H^{-1})^{-1} = H$.
+      Thus $\CAT_i^\op \iso \CAT_i$.
+  \end{enumerate}
+\end{answer}
 
@@
 
+\begin{answer}
+  We define a functor $F : \Set_* \to 1/\Set$ by
+  \begin{align*}
+    F((A, a)) &\coloneqq (x \mapsto a) \\
+    F(h) &\coloneqq h
+  \end{align*}
+  for all pointed sets $(A, a)$ and $(B, b)$ and functions $h : A \to B$ with $f(a) = b$.
+  The second line is valid because
+  \[ \begin{split}
+    h \circ F((A, a))
+    &= (x \mapsto h(a)) \\
+    &= (x \mapsto b) \\
+    &= F((B, b)).
+  \end{split} \]
+  Now we define the inverse functor $G : 1/\Set \to \Set_*$ by
+  \begin{align*}
+    G(f) &\coloneqq (A, f(\star)) \\
+    G(h) &\coloneqq h
+  \end{align*}
+  for all functions $f : 1 \to A$, $g : 1 \to B$ and $h : A \to B$ with $h \circ f = g$.
+  The second line is valid because
+  \[ \begin{split}
+    h(f(\star))
+    &= (h \circ f)(\star) \\
+    &= g(\star).
+  \end{split} \]
+  Both functors are mutually inverse on objects because
+  \[ \begin{split}
+    F(G(f))
+    &= F((A, f(\star))) \\
+    &= (x \mapsto f(\star)) \\
+    &= f
+  \end{split} \]
+  and
+  \[ \begin{split}
+    G(F((A, a)))
+    &= G(x \mapsto a) \\
+    &= (A, a).
+  \end{split} \]
+  On arrows they are trivially inverse.
+\end{answer}
+
 \begin{definition}{Slice Category}
@@
 \end{exercise}
+
+\begin{answer}
+  As suggested, we define
+  \begin{align*}
+    F(A) &\coloneqq \A / A \\
+    F(f) &\coloneqq F_f,
+  \end{align*}
+  where $F_f : \A / A \to \A / B$ is given by
+  \begin{align*}
+    F_f(g) &\coloneqq f \circ g \\
+    F_f(h) &\coloneqq h
+  \end{align*}
+  for all functions $g$, $g'$ and $h$ such that the following diagram commutes.
+  \[ \begin{tikzcd}[column sep=tiny]
+    C \arrow[dr, "g"'] && C' \arrow[ll, "h"] \arrow[dl, "g'"] \\
+    & A \arrow[d, "f"] \\
+    & B
+  \end{tikzcd} \]
+  The second line is valid because
+  \[ \begin{split}
+    F_f(g) \circ h
+    &= f \circ g \circ h \\
+    &= f \circ g' \\
+    &= F_f(g')
+  \end{split} \]
+  Our definition of $F$ satisfies both functor laws:
+  \begin{enumerate}
+    \item For $f : A_2 \to A_3$ and $f' : A_1 \to A_2$, we have $F(f \circ f') = F(f) \circ F(f')$:
+      \begin{align*}
+        F_{f \circ f'}(g)
+          &= f \circ f' \circ g \\
+          &= F_f(F_f'(g)) \\
+        F_{f \circ f'}(h)
+          &= h \\
+          &= F_f(F_f'(h))
+      \end{align*}
+    \item We have $F(1_A) = 1_{F(A)}$:
+      \begin{align*}
+        F_{1_A}(g) &= g \\
+        F_{1_A}(h) &= h
+      \end{align*}
+  \end{enumerate}
+  In general the functor $F$ is not an isomorphism.
+  For example, if $\A$ is the empty category then it is not isomorphic to $\Cat$.
+\end{answer}
 

### 2017-05-19 sheets/basis-3/sheet-basis-3.tex

@@
 
+\begin{definition}{Product map application}
+  If $(A, C), (B, D)$ are elements of $ob(\cat{A} \times \cat{C})$, then for
+  every pair of arrows $f: A \to B$ and $g: C \to D$, the application of the
+  product map $(f, g): (A, C) \to (B, D)$ on any $(a, c): (A, C)$ is defined
+  as follows:
+  \begin{equation*}
+    (f, g) (a, c) := (f(a), g(c))
+  \end{equation*}
+\end{definition}
+
 \begin{exercise}
@@
     F_A(B) & =  B^A \times A\\
-    F_A(f: B \to C) & =  (F \circ \_) \times id_A
+    F_A(f: B \to C) & =  (f \circ \_, 1_A)
   \end{align*}
-  The fact that $eval: F_A \to I_{\mathbf{Set}}$ is a natural transformation follows from the commutativity of the diagram
+  The fact that $eval: F_A \to 1_{\mathbf{Set}}$ (defined by the rule $eval(f,
+  a) = f(a)$) is a natural transformation follows from the commutativity of the
+  diagram
 \[
   \begin{tikzcd}
-    F_A(C) = C^A \times A \arrow{rrr}{eval_{AC}} \arrow{dd}{F_A(g)=(g\circ \_) \times id_A} & & & C = I_{\mathbf{Set}}(C) \arrow{dd}{g = I_{\mathbf{Set}(g)}} \\
+    F_A(C) = C^A \times A \arrow{rrr}{eval_{AC}} \arrow{dd}{F_A(g)=(g\circ \_, 1_A)} & & & C = 1_{\mathbf{Set}}(C) \arrow{dd}{g = 1_{\mathbf{Set}(g)}} \\
                                                                                            & & & \\
-    F_A(B) = B^A \times A \arrow{rrr}{eval_{AB}} & & & B = I_{\mathbf{Set}}(B)
+    F_A(B) = B^A \times A \arrow{rrr}{eval_{AB}} & & & B = 1_{\mathbf{Set}}(B)
   \end{tikzcd}

### 2017-05-19 sheets/basis-3/sheet-basis-3.tex
Fixed the diagram of 3.4 to match the orientation of n.t. diagrams in Leinster.

--- 2017-05-18 sheets/basis-3/sheet-basis-3.tex
@@
 
+\begin{definition}{Product of Morphisms inside a Category}
+  Let $\cat{A}$ be a category and $A, B, C, D \in \ob(\cat{A})$ with $f \from A \to C$ and $g \from B \to D$.
+  The product morphism of $f$ and $g$, written $f \times g$, is a function
+  \begin{align*}
+    A \times B &\qquad\to\qquad C \times D\\
+    (a, b) &\qquad\mapsto\qquad (f(a), g(b))
+  \end{align*}
+\end{definition}
+
+\newcommand{\eval}[1]{\ensuremath{\mathsf{eval}^{#1}\xspace}}
+
 \begin{exercise}
@@
     F_A(B) & =  B^A \times A\\
-    F_A(f: B \to C) & =  (F \circ \_) \times id_A
+    F_A(f: B \to C) & =  (f \circ \_) \times 1_A
   \end{align*}
-  The fact that $eval: F_A \to I_{\mathbf{Set}}$ is a natural transformation follows from the commutativity of the diagram
+  The function $\eval{A} \from F_A \to 1_{\mathbf{Set}}$, defined as $\eval{A}(f,a) := f(a)$, is a natural transformation if the following diagram commutes for each $g \from B \to C$.
 \[
   \begin{tikzcd}
-    F_A(C) = C^A \times A \arrow{rrr}{eval_{AC}} \arrow{dd}{F_A(g)=(g\circ \_) \times id_A} & & & C = I_{\mathbf{Set}}(C) \arrow{dd}{g = I_{\mathbf{Set}(g)}} \\
+    F_A(B) = B^A \times A \arrow{rrr}{F_A(g)=(g\circ \_) \times 1_A} \arrow{dd}[swap]{\eval{A}_B} & & & 1_{\mathbf{Set}}(B) = B \arrow{dd}{\eval{A}_C} \\
                                                                                            & & & \\
-    F_A(B) = B^A \times A \arrow{rrr}{eval_{AB}} & & & B = I_{\mathbf{Set}}(B)
+    F_A(C) = C^A \times A \arrow{rrr}[swap]{1_{\mathbf{Set}}(g) = g} & & & 1_{\mathbf{Set}}(C) = C
   \end{tikzcd}
 \]
-for each $g: C \to B$.
 

### 2017-05-19 sheets/basis-2/sheet-basis-2.tex
test

@@
 
+\begin{answer}
+In $\mathbf{CAT}$, objects are categories and the self loops are trivial endofunctors and so they trivially follow identity laws.\\
+This counts as a trivial answer to the question. However, we can also include all functors.\\
+For associativity, consider three functors $F:A\rightarrow B$, $G:B\rightarrow C$ and $H:C\rightarrow D$.\\
+\begin{itemize}
+\item $(H \of G) \of F(a)$ = $H \of G(F(a))$ = $H(G(F(a)))$
+\item $H \of (G \of F) (a)$ : $H \of (G \of F(a))$ = $G \of G(F(a))$ = H(G(F(a)))
+\item Therefore, $(H \of G) \of F(a) = H \of (G \of F) (a)$ and we satisfy associativity for objects.
+\item The same works for morphisms because $F(f' \of f)  = F(f') \of F(f)$ by definition. Therefore, we satisfy associativity for objects.
+\item Since identities are mapped to identities, we satisfy identity laws
+\item Therefore, $\mathbf{CAT}$ is a category
+\end{itemize}
+\end{answer}
+
 \begin{exercise}
@@
 \end{exercise}
+
+\begin{answer}
+\begin{itemize}
+\item Any bijective functor from a discrete category to another is a full and faithful functor.
+\item A functor that maps a category with two objects and two morphisms (both have the same domain and codomain) to a category with the same two objects and one morphism is full, but not faithful.
+\item A functor that maps a category with two objects and one morphism to a category with the same two objects and two morphisms is faithful, but not full.
+\item Combining the last two to have a category with four objects which are divided into two pairs as mentioned before which are mapped to another category with four objects that are divided into two pairs with the same mentioned structures.
+\end{itemize}
+\end{answer}
 
@@
 \end{exercise}
+
+\begin{answer}
+\begin{enumerate}
+\item
+\begin{itemize}
+\item Since every set is mapped to another set (its powerset), then all objects are mapped.
+\item A function $f: X \rightarrow Y$ is mapped to $F: P(X) \rightarrow P(Y)$, such that $F(x)  = \{x \in X\}$.
+\item This makes the axioms follow easily from composition in $\mathbf {Set}$. Therefore, it satisfies functor laws.
+\item It is faithful (A set has only one powerset) and not full (consider a function that takes the entire powerset as an element and maps it to the other powerset).
+\end{itemize}
+\item
+The only difference is mapping elements in the powerset to their preimage set. The rest of the argument is the same.
+\end{enumerate}
+\end{answer}
 
@@
 \end{exercise}
+
+\begin{answer}
+\begin{itemize}
+\item We assume mapping a function $f: X \rightarrow Y in$ $\mathbf {Set}$ to a free monoid morphism $F: MonX \rightarrow MonY$ such that $F(Z \in MonX) = (f(z) | z \in Z)$
+\item F is faithful since if two sets have the same free monoids (All the lists that can be generated from the elements of those sets), then the two sets are the same.
+\item F is not full because we can have a morphism $F(Z \in MonX) = (f(z)f(z) | z \in Z)$
+\end{itemize}
+\end{answer}
 

### 2017-05-22 sheets/basis-2/sheet-basis-2.tex
Fix solution for exercise 2.12

--- 2017-05-18 sheets/basis-2/sheet-basis-2.tex
@@
 \newcommand{\Rel}{\fcat{Rel}}
+
+\newcommand{\const}[1]{\mathsf{const}_{#1}}
 
@@
   \begin{align*}
-    F((A, a)) &\coloneqq (x \mapsto a) \\
+    F((A, a)) &\coloneqq \const{a} \\
     F(h) &\coloneqq h
   \end{align*}
-  for all pointed sets $(A, a)$ and $(B, b)$ and functions $h : A \to B$ with $f(a) = b$.
+  for all pointed sets $(A, a)$ and $(B, b)$ and functions $h : A \to B$ with $h(a) = b$.
+  Here $\const{x}$ stands for the constant function with value $x$.
   The second line is valid because
@@
     h \circ F((A, a))
-    &= (x \mapsto h(a)) \\
-    &= (x \mapsto b) \\
-    &= F((B, b)).
+    &= \const{h(a)} \\
+    &= \const{b} \\
+    &= F((B, b))
   \end{split} \]
+  holds for $h$ as above.
   Now we define the inverse functor $G : 1/\Set \to \Set_*$ by
@@
     &= F((A, f(\star))) \\
-    &= (x \mapsto f(\star)) \\
+    &= \const{f(\star)} \\
     &= f
@@
     G(F((A, a)))
-    &= G(x \mapsto a) \\
+    &= G(\const{a}) \\
     &= (A, a).

### 2017-05-22 sheets/basis-2/sheet-basis-2.tex
Fixed my solutions.

@@
   \end{align*}
-  Clearly, for each $a\in A$, $f(i(a))=j(a)$, thus $f\of i=j$, and $f$ is a group homomorphism.
+  Clearly, for each $a\in A$, $f(i(a))=j(a)$, thus $f\of i=j$, and $f$ is a monoid homomorphism.
 
-  Now, suppose $g$ is a group homomorphism satisfying $g\of i=j$.
+  Now, suppose $g$ is a monoid homomorphism satisfying $g\of i=j$.
   Let $a_1\cdots a_n\in A^*$. Proof by induction on $n$ that $g(a_1\cdots a_n)=f(a_1\cdots a_n)$.
@@
   \begin{align*}
-    g(a_1\cdots a_{n-1}a_n)&=g(a_1\cdots a_{n-1})\of g(a_n)&g\text{ group homomorphism}\\
+    g(a_1\cdots a_{n-1}a_n)&=g(a_1\cdots a_{n-1})\of g(a_n)&g\text{ monoid homomorphism}\\
                            &=f(a_1\cdots a_{n-1})\of g(a_n)&\text{inductive hypothesis}\\
@@
                            &=f(a_1\cdots a_{n-1})\of f(a_n) & a_n\in A\\
-                           &=f(a_1\cdots a_{n-1}a_n)&f\text{ group homomorphism}
+                           &=f(a_1\cdots a_{n-1}a_n)&f\text{ monoid homomorphism}
   \end{align*}
@@
 
-    Indeed, the reverse is also true: Given any injection $f'\from A'\to B'$ there is functor $F\from \A\to\Set$ such that $f'$ corresponds to $F$ (in the way specified above).
+    Indeed, assuming the axiom of choice the reverse is also true: Given any injection $f'\from A'\to B'$ there is functor $F\from \A\to\Set$ such that $f'$ corresponds to $F$ (in the way specified above).
+($g$ is mapped to the left inverse of $f$, which can be constructed by picking for any $b'$ in the image of $f'$ an arbitrary preimage of $b'$ under $f'$ (axiom of choice!) and for any $b'$ not in the image of $f'$ an arbitrary element of $A'$ (again axiom of choice!)).
   \end{enumerate}

### 2017-05-22 sheets/basis-3/sheet-basis-3.tex
Some solutions

@@
 \end{exercise}
+\begin{answer}
+    \begin{minipage}{0.5\textwidth}
+        $\cat{A}$:
+        \[
+            \begin{tikzcd}
+                A \arrow{d}{f}\\
+                B
+            \end{tikzcd}
+        \]
+    \end{minipage}
+    \begin{minipage}{0.5\textwidth}
+        $\cat{B}$:
+        \[
+            \begin{tikzcd}
+                A' \arrow{dr}{f'} \arrow[bend left]{rr}{g} &    & A'' \arrow{dl}{f''} \arrow{ll}{g^{-1}} \\
+                                                           & B' &
+            \end{tikzcd}
+        \]
+    \end{minipage}
+    The identity morphisms have been left out for simplicity.
+    First of all note that there is no bijection between the objects of $\cat{A}$ and $\cat{B}$, so $\cat{A} \not\iso \cat{B}$.\\
+    We now show $\cat{A} \eqv \cat{B}$.
+    Therefore we define a functor $F$:
+    \begin{align*}
+        F(A) := A'\\
+        F(B) := B'\\
+        F(f) := f'
+    \end{align*}
+    In addition $F$ maps the identity morhisms to their respective identity morphisms.
+    It is easy to check that $F$ is full and faithful.
+    Furthermore $F$ is surjective on the objects, because $A'$ and $A''$ are isomorph in $\cat{B}$.
+    Thus $\cat{A} \eqv \cat{B}$.
+\end{answer}
 
@@
 \end{exercise}
+\begin{answer}
+    \begin{itemize}
+        \item[(a)] $\lambda\, a\, \lambda\,f.\, f\,a$
+        \item[(b)] $\lambda\, (a, f).\, f\,a$
+    \end{itemize}
+\end{answer}
 
@@
 \end{exercise}
+\begin{answer}
+    \begin{itemize}
+        \item[(a)]
+            By definition of natural isomorphism we have a natural transformation $\beta: G \Rightarrow F$, such that
+            \[\beta \circ \alpha = 1_{F} \text{ and } \alpha \circ \beta = 1_{G}. \]
+            We show that $\forall A \in \cat{A}.\, \beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}$.
+            Suppose $A \in \cat{A}$.
+            We know by definition of the identity natural transformation that $(\beta \circ \alpha)_A = 1_{F(A)}$.
+            The claim $\beta_A \circ \alpha_A = 1_{F(A)}$ follows by definition of composition of natural transformations.
+            $\alpha_A \circ \beta_A = 1_{G(A)}$ is shown analogously.
+        \item[(ab)]
+            We know that for all $A \in \cat{A}$ there is a $\beta_A$ such that
+            \[\beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}.\]
+            First we prove that $\beta$ is a natural transformation.
+            Suppose $f\from A \to A'$.
+            We know that $\alpha$ is a natural transformation:
+            \begin{align*}
+                & G(f) \circ \alpha_A && = \alpha_{A'} \circ F(f)\\
+                \Rightarrow\, & G(f) \circ \alpha_A \circ \beta_A && = \alpha_{A'} \circ F(f) \circ \beta_A\\
+                \Rightarrow\, & \beta_{A'} \circ G(f) && = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
+                \Rightarrow\, & \beta_{A'} \circ G(f) && = F(f) \circ \beta_A
+            \end{align*}
+            By similar reasoning to (a) we get that $\beta$ is the inverse of $\alpha$.
+            Therefore $\alpha$ is a natural isomorpism.
+    \end{itemize}
+\end{answer}
 
@@
 \end{exercise}
+\begin{answer}
+    Define a functor $F: \Mt \to \FDVect_K$:
+    \begin{align*}
+        F(n) & := K^n && \forall n \in \NN\\
+        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{ matrices } A
+    \end{align*}
+    Note that $F$ is indeed a functor, since matrix multiplication is associative.
+    \begin{description}
+        \item[$F$ is faithful:]
+            Let $A, B$ be two $n \times m$ matrices over $K$ such that $F(A) = F(B)$.
+            By using the standard basis in $K^m$ we get, that each column of $A$ and $B$ are the same.
+            Thus $A = B$.
+        \item[$F$ is full:]
+            Let $f: K^m \to K^n$ be a linear function.
+            We define the $n \times m$ matrix $A$ columnwise.
+            The $j$-th column of $A$ is $f(e_j)$, where $e_j$ is the $j$-th standard basis vector.
+            By this definition $F(A)$ and $f$ conincide on the standard basis.
+            Linear algebra tells us that in this case $F(A)$ and $f$ coincide on all vectors.
+            So $A$ is mapped to $f$ under $F$.
+        \item[$F$ is surjective on the objects:]
+            Let $V$ be a finite-dimensional $K$-vectorspace.
+            Then $V$ has some dimension $n \in \NN$.
+            Thus $V \iso K^n = F(n)$.
+    \end{description}
+\end{answer}
 

### 2017-05-23 sheets/basis-1/sheet-basis-1.tex

@@
 
-%\section{Diagrams}
-%
-%\begin{hint}
-%  This is how you can draw diagrams:
-%  \[
-%    \begin{tikzcd}
-%                                    & A \arrow{dr}{g}    & \\
-%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
-%    \end{tikzcd}
-%  \]
-%\end{hint}
-
 \section{Categories - Definition and Basics}
@@
 \end{exercise}
+
+\begin{answer}
+  We have to show that for $f \from A \to B$, there is at most one $g
+  \from B \to A$ s.t. $gf = 1_A$ and $fg = 1_B$.
+
+  Assume that $g$ exists and that there is some $h \from B \to A$: $hf = 1_A$, $fh = 1_B$.
+
+  We know from the identity axiom: $1_A \of g = g$ and $h \of 1_B = h$. Using this, our assumtions and the associativity axiom, we get:
+ \[  1_A \of g = g \textiff (h \of f) \of g = g \textiff h \of (f \of g) = g \textiff h \of 1_B = g \textiff h=g. \]
+\end{answer}
 
@@
 
+\begin{answer}
+  textbf{identity laws:} for $f \sub A \times B$,
+\[f \of 1_A = \{ \angles{a, b} \in A \times B \such \exists a' \emptybk (\angles {a, a'} \in 1_A \emptybk \& \emptybk \angles{a', b} \in f)\} = \{ \angles{a, b} \in A \times B \such \exists a' \in A.
+\emptybk \angles{a', b} \in f)\} = f.\] Analogously, $1_B \of f = f$.
+
+\textbf{associativity:} for $f \sub A \times B, g\sub B \times C, h \sub C \times D$,
+\[
+h \of (g \of f) = h \of \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\}
+\]
+\[= \{ \angles{a, d} \in A \times D \such \exists b, c \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g \& \angles{c, d} \in h \} = (h \of g) \of f\]
+
+\end{answer}
+
 \begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
@@
 \end{exercise}
+
+\begin{answer}
+  \textbf{identity laws:} $1_A$ is the identity function, which is monotone since $a \leq^A_\cdot a' \implies a' \leq^A_\cdot a.$
+\\The axiom holds because for $f \from A \in B, f \of 1_A = f = 1_B \of f$ (same as in \textbf{Set}).
+\\
+\textbf{associativity:} If $ f \from A \to B, g \from B \to C $ are monotone, so is $g \of f$ because \[a \leq^A_\cdot a' \implies f(a) \leq^A_\cdot f(a') \implies  g(f(a)) \leq^A_\cdot g(f(a'))  .\] Associativity follows from assiociativity of function composition.
+\end{answer}
 
@@
 
+\begin{answer}
+  \begin{enumerate}
+  \item $\emptyset$ is the zero object. Since there is nothing, that could be related, the null relation is unique relation from $\emptyset$ to any other set S and vice versa.
+  \item The trivial groups are zero objects.
+
+    Initial oject:
+
+    Let $\left({G, \circ}\right)$ be a group with identity element $e_G$ and $1={e}$ a trivial group, with operation $*: e*e := e$.
+
+    Since group homomorphism preserves identity, any hypothetical group homomorphism $\phi: 1 \to G$ must satisfy:
+
+    $\phi (e) = e_G$
+
+    Show  that $\phi$ is  a group homomorphism:
+
+    The initial object $1$ has only the element $e$,
+
+    $\phi (e) \circ \phi (e) = e_G \circ e_G $ , Definition of $\phi$
+    \\$= e_G$ , Definition of identity
+    \\$= \phi (e)$
+    \\$= \phi (e * e)$, Terminal object:
+
+    Because singleton is terminal object on sets, there is precisely one mapping:
+
+    $!: G \to \left\{{e}\right\}$
+
+    defined by:
+
+    $\forall g \in G: ! (g) = e$
+
+    Show that $!$ is a group homomorphism.
+
+    For any $g, h \in G$, we have:
+    $!(g)*!(h) = e_G \circ e_G $ , Definition of !
+    \\$= e$ , Definition of *
+    \\$= !(g \circ h$ Defintion of !
+
+  \item Initial is the ring of Integers $\mathbb{Z}$, terminal the zero ring consisting of only one element.
+
+    Initial: Assume some $ \phi: \mathbb{Z} \mapsto R $, where R is a Ring. Since $\phi$ is a ring homomorphism, $\phi(n) = \phi(n*1) = n*\phi(1)$ holds and therefore, $\phi$ is unique.
+
+    Terminal: Like in set, there is only the constant mapping between any ring and the one-element ring.
+  \end{enumerate}
+\end{answer}
+
+\section{Monomorphisms and Epimorphisms}
+
 \begin{exercise}
@@
 
-\newpage
+\begin{answer}
+  Consider the diagram:
+  \[
+    \begin{tikzcd}
+      A \arrow[r,shift left, "y"] \arrow[r,shift right,swap,"x"] &
+      B \arrow[dr, shift left, "1"]  \arrow{r}{m}  &
+      C \arrow{d}{e} \arrow[r,shift left, "i"] \arrow[r,shift right,swap,"j"] & D\\
+     &  & B \arrow[ul, shift left, "1"]&
+    \end{tikzcd}
+  \]
+  $m \from B \to C$ is an isomorphism with inverse $e \from C \to B$.
+  Then:\\
+  \begin{minipage}{.5\linewidth}
+    \vspace{4mm}
+    \centering $m$ is monic
+    \[
+      \begin{aligned}
+        m \of x &= m \of y\\
+        e \of (m \of x) &= e \of (m \of y)\\
+        (e \of m) \of x &= (e \of m) \of y\\
+        1_B \of x &= 1_B \of y\\
+        x &= y
+      \end{aligned}
+    \]
+  \end{minipage}%
+  \begin{minipage}{.5\linewidth}
+    \vspace{4mm}
+    \centering $m$ is epic
+    \[
+      \begin{aligned}
+        i \of m &= j \of m \\
+        (i \of m) \of e &= (j \of m) \of e\\
+        i \of (m \of e) &= j \of (m \of e)\\
+        i \of 1_C &= j \of 1_C\\
+        i &= j
+      \end{aligned}
+    \]
+
+  \end{minipage}%
+\end{answer}
+
+\exercisesOnly{\newpage}
 
@@
 
+\begin{answer}
+    \begin{itemize}
+  \item If $f$ and $g$ are isomorphic, their inverses
+    $f^{-1}$ and $g^{-1}$ exist.
+    To show that $h = g \of f$ is also an isomorphism,
+    we need to provide an inverse.\\
+
+    \begin{minipage}{.5\linewidth}
+      \[
+      \begin{aligned}
+        & \ \ \ \ (g \of f) \of (f^{-1} \of g^{-1})\\
+        &= g \of (f \of f^{-1}) \of g^{-1} \\
+        &= (g \of 1_B) \of g^{-1} \\
+        &= g \of g^{-1} \\
+        &= 1_C
+      \end{aligned}
+      \]
+    \end{minipage}%
+    \begin{minipage}{.5\linewidth}
+      \[
+      \begin{aligned}
+        & \ \ \ \ (f^{-1} \of g^{-1}) \of (g \of f)\\
+        &= f^{-1} \of (g^{-1} \of g) \of f\\
+        &= (f^{-1} \of 1_B) \of f\\
+        &= f^{-1} \of f\\
+        &= 1_A
+      \end{aligned}
+      \]
+    \end{minipage}%
+    \vspace{4mm}
+    \\
+    $(f^{-1} \of g^{-1})$ is the inverse of $(g \of f)$.
+    Therefore $h = (g \of f)$ is an isomorphism.
+    \vspace{5mm}
+  \item
+    \raggedright Let $x, y \from D \to A$ be any morphisms
+    from any object $D$ in the category. \\
+    \centering
+      \[
+      \begin{aligned}
+        f \of x &= f \of y \\
+        g \of (f \of x) &= g \of (f \of y) \\
+        (g \of f) \of x &= (g \of f) \of y \\
+        x &= y \\
+      \end{aligned}
+      \]
+  \item
+    \raggedright Let $i, j \from C \to E$ be any morphisms
+    to any object $E$ in the category.\\
+    \centering
+      \[
+      \begin{aligned}
+        i \of g &= j \of g \\
+        (i \of g) \of f &= (j \of g) \of f \\
+        i \of (g \of f) &= j \of (g \of f) \\
+        i &= j \\
+      \end{aligned}
+      \]
+  \end{itemize}
+\end{answer}
+
 \begin{exercise}
@@
 
+\begin{answer}
+  \begin{itemize}
+  \item Let $g,h \from \{x\} \to A$ two different functions
+    from the one element set
+    $\{x\}$ to $A$, with $g(x) = a_1$ and $g(x) = a_2$,
+    for any $a_1,a_2 \in A$, with $a_1 \neq a_2$.\\
+    We have: $g \neq h \xRightarrow{f \ mono} f \of g \neq f \of h$.
+    Therefore: $f(a_1) = f(g(x)) \neq f(h(x)) = f(a_2)$.\\
+    $\Rightarrow f$ is injective.
+  \item Let $B$ be a two element set, e.g.
+    $\{ \texttt{true}, \texttt{false}\}$.
+    Let $g$ be the characteristic function of \texttt{Im}$(f)$,
+    the image of f, which is defined by
+    $\forall y \in Y. \ g(y) = \texttt{true}
+    \Leftrightarrow y \in \texttt{Im}(f)$.
+    And let $h$ be the constant \texttt{true}-function.
+    $\forall y \in Y. \ h(y) = \texttt{true}$.
+    Note that $g = h$, exactly if $f$ is surjective.\\
+    We have: $g \of f = h \of f \xRightarrow{f \ epic} g = h$.\\
+    $\Rightarrow$ f is surjective.
+  \end{itemize}
+\end{answer}
+
 \begin{exercise}[Challenge 2]
@@
 
+\begin{answer}
+  The inclusion map $f \from \mathbb{Z} \to \mathbb{Q}$ is defined by
+  $f(x) = x$ and is clearly not surjective on $\mathbb{Q}$, because
+  the $\texttt{Im}(f)$ doesn't contain any fractions like $\frac{1}{2}$.
+  Let $g,h \from \mathbb{Q} \to \bf{R}$ be two morphisms from
+  $\mathbb{Q}$ to any ring $\bf{R}$.
+  $g,h$ agree on $\mathbb{Z}$, because there is exactly one ring
+  homomorphism $\mathbb{Z} \to \bf{R}$ (In other words $\mathbb{Z}$
+  is an initial object in the category \textbf{Ring}).
+  Then we have $\forall p,q \in \mathbb{Z}, q \neq 0$:
+  \[
+    \begin{aligned}
+      & \ \ \ \ g(\frac{p}{q})\\
+      &= g(p) g(p^{-1})\\
+      &= g(p) g(q)^{-1}\\
+      &= h(p) h(q)^{-1}\\
+      &= h(p) h(p^{-1})\\
+      &= h(\frac{p}{q})
+    \end{aligned}
+  \]
+  The first two steps are justified by the
+  properties of ring homomorphisms to respect the structure.
+  The third step uses the fact that $g$ and $h$ agree on
+  $\mathbb{Z}$.\\
+  Therefore: $g \of f = h \of f \Rightarrow g = h$, and $f$ is epic.
+\end{answer}
+
 \end{document}

### 2017-05-23 sheets/basis-2/sheet-basis-2.tex
Fixed part of Mostafa's solutions.

@@
    \item Since every set is mapped to another set (its powerset), then all objects are mapped.
-\item A function $f: X \rightarrow Y$ is mapped to $F: P(X) \rightarrow P(Y)$, such that $F(x)  = \{x \in X\}$.
-\item This makes the axioms follow easily from composition in $\mathbf {Set}$. Therefore, it satisfies functor laws.
+   \item A function $f: X \rightarrow Y$ is mapped to $F: P(X) \rightarrow P(Y)$, such that $F(x)  = \{f(x) | x \in X\}$.
+   \item This makes the axioms follow easily from composition in $\mathbf {Set}$:
+   \begin{itemize}
+   \item $F(f'\of f) (a) = F(f' \of f(a)) = F(f'(f(a)))$
+   \item $F(f') \of F(f) (a) = F(f') \of F(f(a)) = F(f'(f(a)))$
+   \item $F(f' \of f) = F(f') \of F(f)$
+   \item Identity law can be verified easily
+   \item Therefore, it satisfies functor laws and it is a functor.
+   \end{itemize}
    \item It is faithful (A set has only one powerset) and not full (consider a function that takes the entire powerset as an element and maps it to the other powerset).
@@
 \end{enumerate}
-\end{answer}
-
-\begin{exercise}
-  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories.
-Consider a contravariant functor $F:\A^{\op}\rightarrow \B$.
-Describe properties of the corresponding function $f:A\rightarrow B$.
-\end{exercise}
-
-\begin{answer}
-  For every arrow $x \to y$ in $\A$, there is an arrow $f : y \to x$ in $\A^\op$ and hence an arrow $F(f) : F(y) \to F(x)$ in $\B$.
-  In the language of posets this means that $x \leq y$ implies $f(x) \geq f(y)$, i.e.\ $f$ is monotonically decreasing.
-  The functor laws hold tautologically in posets and don't tell us anything about $f$.
 \end{answer}

### 2017-05-23 sheets/basis-2/sheet-basis-2.tex
Fixed exercise Mostafa (and not me\!) was responsible for.

@@
 \begin{exercise}
+  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories.
+Consider a contravariant functor $F:\A^{\op}\rightarrow \B$.
+Describe properties of the corresponding function $f:A\rightarrow B$.
+\end{exercise}
+
+\begin{answer}
+  For every arrow $x \to y$ in $\A$, there is an arrow $f : y \to x$ in $\A^\op$ and hence an arrow $F(f) : F(y) \to F(x)$ in $\B$.
+  In the language of posets this means that $x \leq y$ implies $f(x) \geq f(y)$, i.e.\ $f$ is monotonically decreasing.
+  The functor laws hold tautologically in posets and don't tell us anything about $f$.
+\end{answer}
+
+\begin{exercise} 
   Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:
@@
 \begin{answer}
-\begin{itemize}
-\item We assume mapping a function $f: X \rightarrow Y in$ $\mathbf {Set}$ to a free monoid morphism $F: MonX \rightarrow MonY$ such that $F(Z \in MonX) = (f(z) | z \in Z)$
-\item F is faithful since if two sets have the same free monoids (All the lists that can be generated from the elements of those sets), then the two sets are the same.
-\item F is not full because we can have a morphism $F(Z \in MonX) = (f(z)f(z) | z \in Z)$
-\end{itemize}
+
+We define $F\from\Set\to\Mon$ in the natural way:
+
+\begin{align*}
+  F(A)&=A^*\\
+  F(f)&=f^*
+\end{align*}
+where
+\begin{align*}
+  f^*(a_1\cdots a_n)=f(a_1)\cdots f(a_n).
+\end{align*}
+Clearly, this is a functor.
+
+Now, suppose $A,A'\in\Set$ and $f,f'\from A\to A'$ such that $F(f)=F(f')$. In particular, for all $a\in A$
+\begin{align*}
+  f(a)=F(f)(a)=F(f')(a)=f'(a)
+\end{align*}
+Hence, $f=f'$ and $F$ is faithful.
+
+Next, suppose $A\in\Set$.
+Define $g\from F(A)\to F(A)$ by $g(a_1\cdots a_n)=a_1a_1\cdots a_na_n$ (i.e.
+each ``character'' is duplicated).
+$g$ is a monoid monomorphism but $F(f)\neq g$ for all $f\from A\to A$ since such $F(f)$ in particular preserve lengths of words.
+
+Ergo, $F$ is not full.
 \end{answer}

### 2017-05-23 sheets/basis-3/sheet-basis-3.tex
Sample solution of Exercise Basis 3.4
Add exercise 6 solution sketch

@@
 \begin{definition}{Canonical according to Leinster}
-  "Canonical is an informal word, meaning something like 'God-given' or
-  'defined without making arbitrary choices'.  For example, for any two sets
+  "Canonical is an informal word, meaning something like `God-given' or
+  `defined without making arbitrary choices'.  For example, for any two sets
   $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
@@
   \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
-  defined by 'choose an element $a_0 \in A$ and send everything to $a_0$' is
+  defined by `choose an element $a_0 \in A$ and send everything to $a_0$' is
   not canonical, because the choice of $a_0$ is arbitrary."
@@
 
+\begin{answer}
+  \begin{itemize}
+    \item[(a)]
+      $\Sym(f) = \lambda g.f \circ g$, where $g$ is a permutation\\
+      $\Ord(f) = \lambda (a, b).(f(a), f(b))$, where $a, b \in X$
+    \item[(b)]
+      Let's consider a non-identity permutation $h$. We can build an ordering
+      from it by taking $(x, h(x))$ as an element of ordering relation for
+      $\forall x \in \text{ some finite set } X$, then filling the gaps by
+      reflexivity and transitivity.
+
+      However, it's not the case for an identity permutation $i$, as for that
+      permutation the $i(x) = x$ which leaves us only with reflexivity pairs if
+      we try to build an ordering from that bijection. Thus, we cannot define
+      formation of an ordering from an identity permutation and therefore
+      there's no natural transformation from $\Sym$ to $\Ord$.
+    \item[(c)]
+      $|\Sym(X)| = n$!\\
+      $|\Ord(X)| = n$!
+  \end{itemize}
+  We have seen that there's no natural transformation $\Sym \to \Ord$.
+  Nonetheless, we have also seen that they have the same number of
+  elements and an isomorphism between them can actually be formed: it's not
+  hard to construct all permutations or orderings of some finite set knowing
+  its elements.
+
+  This leads us to the conclusion that $\Sym(X) \iso \Ord(X)$ for all $X \in
+  \cat{B}$, but not \emph{naturally} in $X \in \cat{B}$.
+\end{answer}
+
 \begin{exercise}
@@
         F(n) & := K^n && \forall n \in \NN\\
-        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{ matrices } A
+        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{matrices} A
     \end{align*}

### 2017-05-24 sheets/basis-1/sheet-basis-1.tex
resolving merge conflicts

@@
 \begin{document}
-
-% Use Basis x or Talk x, where x is the number of the session
+\[% Use Basis x or Talk x, where x is the number of the session
 \uebunghead{Basis 1}{What is a Category?}
@@
 \end{hint}
+
+%\section{Diagrams}
+%
+%\begin{hint}
+%  This is how you can draw diagrams:
+%  \[
+%    \begin{tikzcd}
+%                                    & A \arrow{dr}{g}    & \\
+%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+%    \end{tikzcd}
+%  \]
+%\end{hint}
+
+\section{Categories - Definition and Basics}
+
+\begin {exercise}
+Show there can be at most one inverse for a map $f \from A \to B$.
+\end{exercise}
+
+\begin {definition}{Rel}
+The objects of the category \textbf{Rel} are sets, an arrow $f \from A \to B$ is a subset $f \sub A \times B$. 
+The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\}$. 
+Composition of two maps $f \sub A \times B$, $g \sub B \times C$ is defined as 
+\[ g \of f = \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\} \]
+
+\section{Categories - Definition and Basics}
 
@@
 \[ g \of f = \{ \round{a, c} \in A \times C \such \exists b. \round{a, b} \in f \land \round{b, c} \in g\} \]
+
 \end{definition}
@@
 \end{exercise}
+
+\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets 
+(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A_\cdot $ ) 
+An arrow $ m \from A \to B $ function that is monotone, i.e. $\forall a, a' \in A$, 
+\[a \leq^A_\cdot a' \implies m(a) \leq^A_\cdot m(a')\]
 
@@
 The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
+
 \end{definition}
@@
 
+\]
 \end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2017-05-26 sheets/talk-4/sheet4.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}
\[\uebunghead{Talk 4}{Galois Connections and Adjunctions}
\author {Sarah Mameche}
\section{Galois connections}

%Solutions preliminary

\begin{exercise}
Let $\mathscr{P} = (P, \preceq_P)$ and $\mathscr{Q} = (Q, \preceq_Q)$ be partially ordered sets, and $F \from P \to Q, G \from Q \to P$ monotone functions with $F \dashv G$.
\begin{enumerate}
\item Show that the following characterizations of the Galois connection are equivalent.
\begin{enumerate}
\item $p \preceq_P G (F (p)) \land q \preceq_Q F(G (q))$
\item $p \preceq_P G(q) \Leftrightarrow q \preceq_Q F(p)$
\end{enumerate}
\item Show that the following property holds for $F \dashv G$. 
\[  
\begin{aligned}
 \forall q \in Q.\; G(q) = min\{p \in P \such F(p) \preceq_Q q\}
\end{aligned}
\]
Formulate a similar statement $ \forall p \in P, F(p)$.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item ...
\item %For some $q \in Q$, set $S = \{p \in P \such F(p) \preceq_Q q}.
Dual property: $\forall p \in P.\; F(p) = min\{q \in Q \such p \preceq_P G(q)\} $
\end{enumerate}
\end{answer}

\begin{exercise}
Consider the posets $\mathscr{N} = (\mathbb{N}, \leqslant), \mathscr{Q} = (\mathbb{Q}^+, \leqslant)$ 
and the inclusion function $I : \mathscr{N} \to  \mathscr{Q}$.
\begin{enumerate}
\item Find Galois connections $H \dashv I$ and $I \dashv G$.
\item Show (with the example from this exercise or a new one) that in general, Galois connections are not symmetric.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate} \item right adjunct: floor function  $\mathcal{F} : \mathcal{Q} \to  \mathcal{N}$ \\
left adjunct: ceiling function $\mathcal{C} : \mathcal{Q} \to  \mathcal{N}$
\item There can be no galois connection $F \dashv G$ from $\mathscr{Q}$ to $\mathscr{N} . F(q)=1 \Leftrightarrow 1 q<2,$ so $\{q\in \mathbb{Q}^+ \such F(q) = 1\} $ has no maximum. The statement follows with Ex.1b
\end{enumerate}\end{answer}

\section{Adjunctions}
\begin{exercise}
Recall the naturality axiom that has to hold for the isomorphism $\mathscr{B}(F(A), B) \; \cong \; \mathscr{A}(A, G(B)) $:

\begin{enumerate}[label=(\roman*)]
\item for any $g \from F(A) \to B$ and $q \from B \to B': \overline{q \of g} = G(q) \of \bar{g} $
\item for any $f \from A \to G(B) $ and $ p \from A' \to A: \overline{f \of p} = \bar{f} \of F(p)$
\end{enumerate}
Show that (i) and (ii) can be replaced by (iii) $\forall f, q, p. \;\overline{G(q) \of f \of p} = q \of \bar{f} \of F(p)$
\end{exercise}

\begin{answer}
(i), (ii) $\Rightarrow$ (iii)\\
composing any $q \from B \to B'$ with (ii): $\forall f, p: \forall q: q \of \bar{f} \of F(p) = q \of \overline{f \of p} $ (by ii). \\Since $f \from p \to F(A') \to B$, its transpose $h=:\overline{f \of p} \from A' \to B.$ \\So from (i) and quantif. over A: $q \of \overline{f \of p}  = q \of h = \overline{G(q) \of \bar{h}} = \overline{G(q) \of f \of p}$\\
(iii) $\Rightarrow$ (i)\\
%(iii) quantifies over all $p$, so set $p = id_A$ (exists for each object). Then $f \of p \from A \to G(B)$, and $\bar{f} \of F(p) \from F(A) \to B.$ 
Assume (i) is false. 
Taking a $g \from F(A) \to B$, you get $h = g  \of id_B  \from F(A) \to B$. 
Then the equality (i) does not hold for h, that is \\ 
$\forall q, \overline{q \of h} \not= G(q) \of \bar{h}$ 
\\$ \Leftrightarrow \overline{q \of g  \of id_B} \not= G(q)  \of \overline{h}$ ( Def h)\\
$\Leftrightarrow q \of g  \of id_B \not= \overline{G(q)  \of \overline{h}}$ (Def transposition)

$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{h}} $  (define f as $\bar {g}.$ Also, $  F(id_A) = id_B$ by Def of functors)
\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{g}} \; (h = g \of id_B = g)$
\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of f \of id_A}\;  (f = \bar{g} $ by def. and $f = f \of id_A)$
\\So (iii) is false for $f = \bar {g}$ and $ p= id_A$. \\(iii)  $\Rightarrow$ (i) similarly.
\end{answer}

\begin{exercise}
Find a right adjunct to the forgetful functor $F \from $\textbf{Set }$ \to $\textbf{ Rel}.
\end{exercise}
\begin{answer}
\end{answer}

\begin{exercise}
\begin{enumerate}
\item Find a functor that is both left and right adjoint to another functor.
\item Find a functor that is self-adjoint.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item We have seen that initial (terminal) objects of categories correspond to left (right) adjoints. So for any category that has a zero object, there is a functor that is both left and right adjoint. Also, an equivalence between two categories is both a left and right adjoint.
\item The contravariant powerset functor $\mathcal{P} \from $\textbf {Set }$ \to$\textbf{ Set}$^{op}$ (see Ex. 7) is self-adjoint. Functions $A \to \mathcal{P}(B)$ in \textbf{Set} correspond to subsets of $A \times B$ and thus to functions $B \to \mathcal{P}(A)$.
\end{enumerate}
\end{answer}

\begin{exercise}
Show that adjunctions preserve initial objects. I.e., if  
\begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
\mathscr{A} \arrow[r, shift left,"F"] 
\& \mathscr{B} \arrow[l, shift left, "G"] 
\end{tikzcd}
with $F\dashv G$ and $I$ is initial in $ \mathscr{A}$, then $F(I)$ is initial in $\mathscr{B}$. Dualize the statement.
\end{exercise}

\begin{answer}
Assume $I$ is initial in $ \mathscr{A}$, then maps $I \to A$ are unique $\forall A \in A$. This implies that maps  $I \to G(B)$ are unique $\forall B$ because $G(B) \in ob(\mathscr{A})$. From $F\dashv G$, we get that the corresponding maps $F(I) \to B$ are unique $\forall B$, so $F(I)$ initial in $\mathscr{B}$.\\ It can be shown in the same way that terminal objects are preserved by $G$.\end{answer}

\begin{exercise}
For sets $X, Y$ and a function $f: X \to Y$, show that the following operations are adjunctions:\\
$ \forall A \in \mathcal{P}(X).\; F \from \mathcal {P}(X) \to \mathcal {P}(Y ) $ with $ F(A) := \{f(a) \such a \in A \}$ and\\
$\forall B\in \mathcal{P}(Y). \; F^{-1} \from \mathcal {P}(Y) \to \mathcal {P}(X)$ with $ F^{-1}(B):=\{a\in X \such f(a)\in B\}$
\\Adapt the construction so that it works for groups (alternatively topological spaces) instead of sets. 
\end{exercise}

\begin{answer}
$F \dashv F^{-1}$, meaning $F(f)(X) \subseteq Y \Leftrightarrow X \subseteq F^{-1}$.\\
Assume $F \dashv F^{-1}$. Then $\forall x \in X, f(x) \in Y$ and thus $X \subseteq F^{-1}$\\
Assume $X \subseteq F^{-1}$ and take $b \in F^{-1}$. By Def $f(x) = b$ for some x, and with $X \subseteq F^{-1}, b = f(x) \in Y$\\
Groups: f homomorphism, $\mathcal{P}(G)$ set of subgroups, same proof \\
Top. Sp: continuous maps and subspaces

\end{answer}

\begin{exercise}
Consider a functor $Ob \from $\textbf {CAT }$  \to $\textbf { Set} which maps a category to its set of objects, and a functor to the mapping induced on objects. Are there left and right adjoints?
\end{exercise}

\begin{answer}
There is a left adjoint $D$, where $D(A)$ is a discrete category with objects $A$. There is also a right adjoint $I$, where $I(A)$ is the category with objects $A$ and a unique arrow $a \to b$ for each $a,b \in A$.
\end{answer}

\begin{exercise}
Let $G$ be a group. \begin{enumerate}
\item Find functors between \textbf{Set} and the functor category $[G,$ \textbf{Set}$]$ of left $G$-sets. Which adjunnctions are there between them?
\item Similarly, find functors \textbf{Vect}$_k$ and $[G,$ \textbf{Vect}$_k]$ and look for adjunctions.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item Objects of $[G,$ \textbf{Set}$]$ are functors $F_S$, with one object (set S) and with maps $f_g \from S\to S$ (for each $g \in G$, left group action). Maps in  $[G,$ \textbf{Set}$]$ are natural transformations between the functors. The canonical functor $F \from$ \textbf{Set}$ \to [G,$ \textbf{Set}$]$ maps a set $S$ to the functor with $S$ as the underlying set. Functions $S\to S' $ are mapped to the natural transformation between the resp. functors $F_S, F_S'$. A functor G which goes back to \textbf{Set} is defined similarly. Then $F \dashv G$ and $G \dashv F$, because every function $S \to S'$ corresponds by construction to exactly one natural transformation $\alpha \from F_S \to F_S'$. The correspondence is natural in S, S' because $\alpha$ preserves the group action: $\alpha (g . s) = g . (\alpha . s)\; \forall s \in S$ (because $\alpha$ is a natural transformation).

\item The Functors $F_V$ in  $[G,$ \textbf{Vect}$_k]$ select a vector space $V$, and have linear maps $f_g: V \to W$ for each $g \in G$. Construct $F$ and $G$ as in (a), then  $F \dashv G$ and $G \dashv F$, because the natural transformations preserve the group action.
\end{enumerate}
\end{answer}

\]
\end{document}

### 2017-05-26 sheets/basis-3/sheet-basis-3.tex
resolving merge conflicts
Fix Typos in solution of Basis 3.4
Partialy Solution for Basis 3.7

@@
     It is easy to check that $F$ is full and faithful.
-    Furthermore $F$ is surjective on the objects, because $A'$ and $A''$ are isomorph in $\cat{B}$.
+    Furthermore $F$ is essentially surjective on objects because $A'$ and $A''$ are isomorphic in $\cat{B}$.
     Thus $\cat{A} \eqv \cat{B}$.
@@
     \begin{itemize}
-        \item[(a)] $\lambda\, a\, \lambda\,f.\, f\,a$
+        \item[(a)] $\lambda\, a.\, \lambda\,f.\, f\,a$
         \item[(b)] $\lambda\, (a, f).\, f\,a$
@@
   For a fixed set $A$, the map taking $B$ to $B^A \times A$ can be extended to
-  a functor $F_A:\mathbf{Set} \to \mathbf{Set}$ as follows:
+  a functor $F_A\from \mathbf{Set} \to \mathbf{Set}$ as follows:
   \begin{align*}
@@
 \end{exercise}
+\begin{answer}
+  Let $(x,y) \in B^A \times A$. We have to show that $\eval{A}_C (F_A(g) (x,y)) = 1_{\mathbf {Set}}(g) (\eval{A}_B (x, y))$.
+  By unfolding the definitions of $\times$, $\eval{A}$ and $F_A$ we get:
+  \[
+    \eval{A}_C (F_A(g) (x,y)) =
+    \eval{A}_C ((g \circ \_)\,x, 1_A\,y) =
+    \eval{A}_C ((g \circ x, y)) =
+    g (x (y))
+    \]
+  By unfolding the definitions of $1_{\mathbf {Set}}$ and $\eval{A}$ we get:
+  \[
+    1_{\mathbf {Set}}(g) (\eval{A}_B (x, y)) =
+    g (\eval{A}_B (x, y)) =
+    g (x (y))
+    \]
+\end{answer}
 
@@
         \item[(a)]
-            By definition of natural isomorphism we have a natural transformation $\beta: G \Rightarrow F$, such that
+            By definition of natural isomorphism we have a natural transformation $\beta\from G \to F$, such that
             \[\beta \circ \alpha = 1_{F} \text{ and } \alpha \circ \beta = 1_{G}. \]
@@
             $\alpha_A \circ \beta_A = 1_{G(A)}$ is shown analogously.
-        \item[(ab)]
+        \item[(b)]
             We know that for all $A \in \cat{A}$ there is a $\beta_A$ such that
@@
             We know that $\alpha$ is a natural transformation:
-            \begin{align*}
-                & G(f) \circ \alpha_A && = \alpha_{A'} \circ F(f)\\
-                \Rightarrow\, & G(f) \circ \alpha_A \circ \beta_A && = \alpha_{A'} \circ F(f) \circ \beta_A\\
-                \Rightarrow\, & \beta_{A'} \circ G(f) && = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
-                \Rightarrow\, & \beta_{A'} \circ G(f) && = F(f) \circ \beta_A
-            \end{align*}
+            \begin{alignat*}{2}
+                & & G(f) \circ \alpha_A & = \alpha_{A'} \circ F(f) \\
+                &\Rightarrow &\quad G(f) \circ \alpha_A \circ \beta_A & = \alpha_{A'} \circ F(f) \circ \beta_A\\
+                &\Rightarrow & G(f) & = \alpha_{A'} \circ F(f) \circ \beta_A\\
+                &\Rightarrow & \beta_{A'} \circ G(f) & = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
+                &\Rightarrow & \beta_{A'} \circ G(f) & = F(f) \circ \beta_A
+            \end{alignat*}
             By similar reasoning to (a) we get that $\beta$ is the inverse of $\alpha$.
@@
 on objects.  For each $B \in \cat{B}$, choose an object $G(B)$ of $\cat{A}$
-and an isomorphism $\varepsilon\from F(G(B)) \to B$.  Prove that $G$ extends
-to a functor in such a way that $(\varepsilon)_{B \in \cat{B}}$ is a natural
+and an isomorphism $\epsln_B\from F(G(B)) \to B$.  Prove that $G$ extends
+to a functor in such a way that $(\epsln)_{B \in \cat{B}}$ is a natural
 isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
@@
 \end{exercise}
+\begin{answer}
+    \begin{itemize}
+        \item[(a)]
+            Since $F$ is an equivalance there is a functor $G\from \cat{B} \to \cat{A}$ and a natural isomorphism $\eta\from 1_{\cat{A}} \to G \circ F$.
+            So the following diagram commutes for all $f\from A \to A'$:
+            \[
+                \begin{tikzcd}
+                    A \arrow{r}{f} \arrow{d}{\eta_A} & A' \arrow{d}{\eta_{A'}} \\
+                    G(F(A)) \arrow{r}{G(F(f))}       & G(F(A'))
+                \end{tikzcd}
+            \]
+            Also we have a natural isomorphism $\epsln\from F \circ G \to 1_{\cat{B}}$ such that for all $g\from B \to B'$ the diagram commutes:
+            \[
+                \begin{tikzcd}
+                    F(G(B)) \arrow{d}{\epsln_{B}} \arrow{r}{F(G(g))} & F(G(B')) \arrow{d}{\epsln_{B'}} \\
+                    B \arrow{r}{g}  & B'
+                \end{tikzcd}
+            \]
+            \begin{description}
+                \item[$F$ is faithful:]
+                    Let $A, A' \in \cat{A}$ and $f,f'\from A \to A'$ such that $F(f) = F(f')$.
+                    Using this commutativity for $\eta$ we obtain our goal.
+                    \begin{align*}
+                        f & = \eta_{A'}^{-1} \circ \eta_{A'} \circ f\\
+                          & = \eta_{A'}^{-1} \circ G(F(f)) \circ \eta_A\\
+                          & = \eta_{A'}^{-1} \circ G(F(f')) \circ \eta_A\\
+                          & = \eta_{A'}^{-1} \circ \eta_{A'} \circ f'\\
+                          & = f'
+                    \end{align*}
+                \item[$F$ is full:]
+                    %TODO
+                \item[$F$ is essentially surjective on objects:]
+                    Let $B \in \cat{B}$.
+                    Then $F(G(B)) \iso B$ by $\epsln_B$.
+            \end{description}
+        \item[(b)]
+            $F$ is essentially surj. on obj. so $\forall B \in \cat{B}.\, \exists A \in \cat{A}.\, F(A) \iso B$.
+            We denote by $h(A, B) \from F(A) \to B$ this isomorphism.
+            For each $B \in \cat{B}$ we define $G(B)$ such that $F(G(B)) \iso B$ under $h(G(B), B)$ and $\epsln_B\from F(G(B)) \to B := h(G(B), B)$.
+            Let $g\from B \to B'$ be a morphism in $\cat{B}$.
+            Then $g' := \epsln_{B'}^{-1} \circ g \circ \epsln_{B} \from F(G(B)) \to F(G(B'))$ is a morphism in the image of $F$.
+            We extend $G$ to a functor by mapping each morphism $g\from B \to B'$ to the morphism $f\from G(B) \to G(B')$ such that $g' = F(f)$.
+            This morphism $f$ exists and is unique because $F$ is full and faithful.
+            We have to show that $\epsln$ is a natural transformation.
+            \[
+                \begin{tikzcd}
+                    F(G(B)) \arrow{d}{\epsln_{B}} \arrow{r}{F(G(g))} & F(G(B')) \arrow{d}{\epsln_{B'}} \\
+                    B \arrow{r}{g}                                   & B'
+                \end{tikzcd}
+            \]
+            The diagram commutes since $F(G(g)) = g' = \epsln_{B'}^{-1} \circ g \circ \epsln_{B}$.\\
+            Next we define $\eta_A$ such that $F(\eta_A) = \epsln_{F(A)}^{-1}$.
+            Again this is possible because $\epsln_{F(A)}^{-1}\from F(A) \to F(G(F(A)))$ is in the image of $F$.
+            \[
+                \begin{tikzcd}
+                    A \arrow{d}{\eta_A} \arrow{r}{f} & A' \arrow{d}{\eta_{A'}} \\
+                    G(F(A)) \arrow{r}{G(F(f))}       & G(F(A'))
+                \end{tikzcd}
+            \]
+            \begin{align*}
+                F(G(F(f))) & = \epsln_{F(A')}^{-1} \circ F(f) \circ \epsln_{F(A)}\\
+                           & = F(\eta_{A'}) \circ F(f) \circ F(\eta_A^{-1})\\
+                           & = F(\eta_{A'} \circ f \circ \eta_A^{-1})
+            \end{align*}
+            We obtain $G(F(f)) = \eta_{A'} \circ f \circ \eta_A^{-1}$ because $F$ is faithfull.
+            This shows that $\eta$ is a natural isomorphism.
+            Therefore $F$ is an equivalence.
+    \end{itemize}
+\end{answer}
 
@@
 \begin{answer}
-    Define a functor $F: \Mt \to \FDVect_K$:
+    Define a functor $F\from \Mt \to \FDVect_K$:
     \begin{align*}
-        F(n) & := K^n && \forall n \in \NN\\
-        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{matrices} A
+        F(n) & := K^n && n \in \NN\\
+        F(A\from m \to n) & := \lambda\, x. \, A \cdot x && A \text{ is a } n \times m \text{ matrix}
     \end{align*}
@@
         \item[$F$ is faithful:]
-            Let $A, B$ be two $n \times m$ matrices over $K$ such that $F(A) = F(B)$.
-            By using the standard basis in $K^m$ we get, that each column of $A$ and $B$ are the same.
+            Let $A, B\from m \to n$, that is $n \times m$ matrices over $K$, such that $F(A) = F(B)$.
+            By using the standard basis $(e_j)_{j = 1}^m$ in $K^m$ we get:
+            \[A \cdot e_j = F(A)(e_j) = F(B)(e_j) = B \cdot e_j \quad\forall 1 \le j \le m\]
+            So each column of $A$ and $B$ are the same.
             Thus $A = B$.
         \item[$F$ is full:]
-            Let $f: K^m \to K^n$ be a linear function.
+            Let $f\from K^m \to K^n$ be a linear function.
             We define the $n \times m$ matrix $A$ columnwise.
@@
             So $A$ is mapped to $f$ under $F$.
-        \item[$F$ is surjective on the objects:]
+        \item[$F$ is essentially surjective on objects:]
             Let $V$ be a finite-dimensional $K$-vectorspace.
@@
     \end{description}
+    Thus $F$ is an equivalence and so $\Mt \eqv \FDVect_K$.
+    The functor $F$ is even canonical since we didn't make arbitrary choices.
+    But the functor from $\FDVect_K$ to $\Mt$ involves choosing a basis like in the case for showing that $F$ is full.
 \end{answer}

### 2017-05-29 sheets/basis-3/sheet-basis-3.tex

@@
 and an isomorphism $\epsln_B\from F(G(B)) \to B$.  Prove that $G$ extends
-to a functor in such a way that $(\epsln)_{B \in \cat{B}}$ is a natural
+to a functor in such a way that $(\epsln_B)_{B \in \cat{B}}$ is a natural
 isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
@@
                     Let $A, A' \in \cat{A}$ and $f,f'\from A \to A'$ such that $F(f) = F(f')$.
-                    Using this commutativity for $\eta$ we obtain our goal.
+                    Using the naturality axiom for $\eta$ we obtain our goal.
                     \begin{align*}
@@
                     %TODO
+                    Let $A, A' \in \cat{A}$ and $g \from F(A) \to F(A')$.
                 \item[$F$ is essentially surjective on objects:]
@@
         \item[(b)]
-            $F$ is essentially surj. on obj. so $\forall B \in \cat{B}.\, \exists A \in \cat{A}.\, F(A) \iso B$.
-            We denote by $h(A, B) \from F(A) \to B$ this isomorphism.
-            For each $B \in \cat{B}$ we define $G(B)$ such that $F(G(B)) \iso B$ under $h(G(B), B)$ and $\epsln_B\from F(G(B)) \to B := h(G(B), B)$.
+            For each $B \in \cat{B}$ the class of objects $A \in \cat{A}$ which fulfill $F(A) \iso B$ is non empty, because $F$ is essentially surjective on objects.
+            We define $G(B)$ as such an $A$ using the axiom of choice.
+            In addition we have an isomorphism $\epsln_B\from F(G(B)) \to B$.
             Let $g\from B \to B'$ be a morphism in $\cat{B}$.
@@
             \end{align*}
-            We obtain $G(F(f)) = \eta_{A'} \circ f \circ \eta_A^{-1}$ because $F$ is faithfull.
+            We obtain $G(F(f)) = \eta_{A'} \circ f \circ \eta_A^{-1}$ because $F$ is faithful.
             This shows that $\eta$ is a natural isomorphism.

### 2017-05-29 sheets/basis-3/sheet-basis-3.tex
WIP: Solution for 3.2

@@
 \end{exercise}
+
+\begin{answer}
+  We have to give two functors
+  $H \from \ftrcat{\cat{A}^\op}{\cat{B}^\op} \to \ftrcat{\cat{A}}{\cat{B}}^\op$
+  and
+  $J \from \ftrcat{\cat{A}}{\cat{B}}^\op \to \ftrcat{\cat{A}^\op}{\cat{B}^\op}$
+  such that $H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op}$ and $J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}$.
+
+  First we define $H$. For that we first map objects of
+  $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$ (being itself functors from $\cat{A}^\op$ to $\cat{B}^\op$)
+  to objects of $\ftrcat{\cat{A}}{\cat{B}}^\op$ (being itself functors from $\cat{A}$ to $\cat{B}$).
+
+  Let $F\from \cat{A}^\op \to \cat{B}^\op$ be an object of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
+  Because every object $A : \cat{A}^\op$ is also an object of $\cat{A}$ we can define
+  $H(F)(A) := F(A)$.
+  Now let $f \from A \to B$ be an morphism in $\cat{A}^\op$. We have $F(f^\op)=g^\op$ for some morphism $g$ in $\cat{B}^\op$.
+  We define $H(F)(f) := F(f^\op)^\op = g$.
+
+  Now let $F, G \from \cat{A}^\op \to \cat{B}^\op$ be objects and
+  $\alpha \from F \to G$ be an morphism of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
+  $\alpha = (F(A) \stackrel{\alpha_A}{\to} G(A))_A$ is an natural transformation.
+  We define $H(\alpha) := \beta \from H(G) \to H(F) := (\alpha_A^\op)_A$ and have to show that $\beta$ is indeed such a natural transformation.
+  Let $A, A' : \cat{A}$ and $f \from A \to A'$. We have to show that the following diagram commutes:
+  \[
+    \begin{tikzcd}
+      H(G)(A) \arrow{r}{H(G)(f)} \arrow{d}{H(\alpha)_A} & H(G)(A') \arrow{d}{H(\alpha)_A'} \\
+      H(F)(A) \arrow{r}{H(F)(f)} & H(F)(A')
+    \end{tikzcd}
+  \]
+  By unfolding the definition of $H$ this diagram commutes if and only if the following diagram commutes:
+  \[
+    \begin{tikzcd}
+      G(A) & G(A') \arrow{l}{G(f^\op)} \\
+      F(A) \arrow{u}{\alpha_A} & F(A') \arrow{l}{F(f^\op)} \arrow{u}{\alpha_{A'}}
+    \end{tikzcd}
+  \]
+  This diagram commutes because $\alpha$ is a natural transformation.
+
+  Secondly we define $J$ using a similar construction as above and get
+  $H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op}$ and $J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}$.
+
+  Now by definition of $\iso$,
+  $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$
+  holds.
+\end{answer}
 
@@
 \begin{answer}
+  In the following we will consider a representation of the permutation as a
+  tuple of two elements $(a, b)$, where $a \mapsto b$ is an element of a
+  bijection.
   \begin{itemize}
     \item[(a)]
-      $\Sym(f) = \lambda g.f \circ g$, where $g$ is a permutation\\
-      $\Ord(f) = \lambda (a, b).(f(a), f(b))$, where $a, b \in X$
+      For any $A \in ob(\cat{B}), C \in ob(\Set), f: A \to C$
+      \begin{align*}
+        \Sym(f) &= \lambda (a, b). (f(a), f(b))\\
+        \Ord(f) &= \lambda (a, b).(f(a), f(b))
+      \end{align*}
     \item[(b)]
-      Let's consider a non-identity permutation $h$. We can build an ordering
-      from it by taking $(x, h(x))$ as an element of ordering relation for
-      $\forall x \in \text{ some finite set } X$, then filling the gaps by
-      reflexivity and transitivity.
-
-      However, it's not the case for an identity permutation $i$, as for that
-      permutation the $i(x) = x$ which leaves us only with reflexivity pairs if
-      we try to build an ordering from that bijection. Thus, we cannot define
-      formation of an ordering from an identity permutation and therefore
-      there's no natural transformation from $\Sym$ to $\Ord$.
+      Let's fix a finite set $A$ and define $\Sym \to \Ord$ on it.\\
+      \begin{align*}
+        A &= \{a, b\}\\
+        \Sym(A) &= \{\{(a, a), (b, b)\}, \{(a, b), (b, a)\}\}\\
+        \Ord(A) &= \{\{(a, b), (a, a), (b, b)\}, \{(b, a), (a, a), (b, b)\}\},\\
+      \end{align*}
+      Natural transformation $\alpha: \Sym(A) \to \Ord(A)$:\\
+      \begin{align*}
+        \alpha_A(\{(a, a), (b, b)\}) &= \{(a, b), (a, a), (b, b)\}\\
+        \alpha_A(\{(a, b), (b, a)\}) &= \{(b, a), (a, a), (b, b)\}
+      \end{align*}
+      And a permutation of $A$:
+      \begin{align*}
+        f(a) &= b\\
+        f(b) &= a
+      \end{align*}
+      Now, for $\alpha$ to indeed be a natural transformation, the  naturality
+      axiom should hold for all arrows $A \to A$, including $f$. Let's check if
+      it holds by applying $f$.\\
+      \begin{align*}
+        &\alpha_A(f(\{(a, a), (b, b)\})) &= \alpha_A(\{(b, b), (a, a)\}) =&& \{(b, a), (b, b), (a, a)\}\\
+        &\alpha_A(f(\{(a, b), (b, a)\})) &= \alpha_A(\{(b, a), (a, b)\}) =&& \{(a, b), (b, b), (a, a)\}
+      \end{align*}
+      Here $\alpha_A$ contradicts our definition of it and we can conclude that
+      it is not a natural transformation. The only other possibility of
+      defining $\alpha_A$ would be to switch the images of sets in the
+      definition, but the applying $f$ will yield us a similar contradictory
+      result.  From this we see that no matter how we define $\alpha_A$, the
+      naturality axiom will not hold for our $f$, hence we cannot have a
+      natural transformation for $\Sym(A) \to \Ord(A)$ and, in general, a
+      natural transformation $\Sym \to \Ord$.
     \item[(c)]
-      $|\Sym(X)| = n$!\\
-      $|\Ord(X)| = n$!
+      From combinatorics we know that the number of all permutations of a
+      finite set of size $n$ is exactly $n$!:\\
+      \begin{equation*}
+        |\Sym(X)| = n!
+      \end{equation*}
+      We can also notice that we may consider permutations as an order too,
+      hence the number of all orders is:
+      \begin{equation*}
+        |\Ord(X)| = n!
+      \end{equation*}
   \end{itemize}
-  We have seen that there's no natural transformation $\Sym \to \Ord$.
-  Nonetheless, we have also seen that they have the same number of
-  elements and an isomorphism between them can actually be formed: it's not
-  hard to construct all permutations or orderings of some finite set knowing
-  its elements.
-
-  This leads us to the conclusion that $\Sym(X) \iso \Ord(X)$ for all $X \in
-  \cat{B}$, but not \emph{naturally} in $X \in \cat{B}$.
+  So we know that $\Sym$ and $\Ord$ have exactly the same number of elements
+  and generally we can come up with a mapping between those elements, such that
+  $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$. However, we have also shown
+  that we cannot define a natural transformation $\Sym \to \Ord$ which gives
+  that there's no natural isomorphism between those categories.
 \end{answer}
@@
                     \end{align*}
+                    By analogous reasoning using $\epsln$ we obtain that $G$ is also faithful.
                 \item[$F$ is full:]
-                    %TODO
                     Let $A, A' \in \cat{A}$ and $g \from F(A) \to F(A')$.
+                    We define $f := \eta_{A'}^{-1} \circ G(g) \circ \eta_A$ and use the naturality axiom for $\eta$ in order to get:
+                    \begin{align*}
+                        G(F(f)) & = \eta_A' \circ f \circ \eta_A^{-1}\\
+                                & = \eta_A' \circ \eta_{A'}^{-1} \circ G(g) \circ \eta_A \circ \eta_A^{-1}\\
+                                & = G(g)
+                    \end{align*}
+                    Since $G$ is faithful we obtain $F(f) = g$.
+                    Thus $F$ is full.
                 \item[$F$ is essentially surjective on objects:]

### 2017-05-30 sheets/basis-3/sheet-basis-3.tex
Fix exercise 3.6

@@
 \end{exercise}
-
-\begin{answer}
-  We have to give two functors
-  $H \from \ftrcat{\cat{A}^\op}{\cat{B}^\op} \to \ftrcat{\cat{A}}{\cat{B}}^\op$
-  and
-  $J \from \ftrcat{\cat{A}}{\cat{B}}^\op \to \ftrcat{\cat{A}^\op}{\cat{B}^\op}$
-  such that $H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op}$ and $J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}$.
-
-  First we define $H$. For that we first map objects of
-  $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$ (being itself functors from $\cat{A}^\op$ to $\cat{B}^\op$)
-  to objects of $\ftrcat{\cat{A}}{\cat{B}}^\op$ (being itself functors from $\cat{A}$ to $\cat{B}$).
-
-  Let $F\from \cat{A}^\op \to \cat{B}^\op$ be an object of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
-  Because every object $A : \cat{A}^\op$ is also an object of $\cat{A}$ we can define
-  $H(F)(A) := F(A)$.
-  Now let $f \from A \to B$ be an morphism in $\cat{A}^\op$. We have $F(f^\op)=g^\op$ for some morphism $g$ in $\cat{B}^\op$.
-  We define $H(F)(f) := F(f^\op)^\op = g$.
-
-  Now let $F, G \from \cat{A}^\op \to \cat{B}^\op$ be objects and
-  $\alpha \from F \to G$ be an morphism of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
-  $\alpha = (F(A) \stackrel{\alpha_A}{\to} G(A))_A$ is an natural transformation.
-  We define $H(\alpha) := \beta \from H(G) \to H(F) := (\alpha_A^\op)_A$ and have to show that $\beta$ is indeed such a natural transformation.
-  Let $A, A' : \cat{A}$ and $f \from A \to A'$. We have to show that the following diagram commutes:
-  \[
-    \begin{tikzcd}
-      H(G)(A) \arrow{r}{H(G)(f)} \arrow{d}{H(\alpha)_A} & H(G)(A') \arrow{d}{H(\alpha)_A'} \\
-      H(F)(A) \arrow{r}{H(F)(f)} & H(F)(A')
-    \end{tikzcd}
-  \]
-  By unfolding the definition of $H$ this diagram commutes if and only if the following diagram commutes:
-  \[
-    \begin{tikzcd}
-      G(A) & G(A') \arrow{l}{G(f^\op)} \\
-      F(A) \arrow{u}{\alpha_A} & F(A') \arrow{l}{F(f^\op)} \arrow{u}{\alpha_{A'}}
-    \end{tikzcd}
-  \]
-  This diagram commutes because $\alpha$ is a natural transformation.
-
-  Secondly we define $J$ using a similar construction as above and get
-  $H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op}$ and $J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}$.
-
-  Now by definition of $\iso$,
-  $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$
-  holds.
-\end{answer}
 
@@
 and an isomorphism $\epsln_B\from F(G(B)) \to B$.  Prove that $G$ extends
-to a functor in such a way that $(\epsln_B)_{B \in \cat{B}}$ is a natural
+to a functor in such a way that $(\epsln)_{B \in \cat{B}}$ is a natural
 isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
@@
                     Let $A, A' \in \cat{A}$ and $f,f'\from A \to A'$ such that $F(f) = F(f')$.
-                    Using the naturality axiom for $\eta$ we obtain our goal.
+                    Using this commutativity for $\eta$ we obtain our goal.
                     \begin{align*}
@@
                     \end{align*}
-                    By analogous reasoning using $\epsln$ we obtain that $G$ is also faithful.
                 \item[$F$ is full:]
-                    Let $A, A' \in \cat{A}$ and $g \from F(A) \to F(A')$.
-                    We define $f := \eta_{A'}^{-1} \circ G(g) \circ \eta_A$ and use the naturality axiom for $\eta$ in order to get:
-                    \begin{align*}
-                        G(F(f)) & = \eta_A' \circ f \circ \eta_A^{-1}\\
-                                & = \eta_A' \circ \eta_{A'}^{-1} \circ G(g) \circ \eta_A \circ \eta_A^{-1}\\
-                                & = G(g)
-                    \end{align*}
-                    Since $G$ is faithful we obtain $F(f) = g$.
-                    Thus $F$ is full.
+                    %TODO
                 \item[$F$ is essentially surjective on objects:]
@@
         \item[(b)]
-            For each $B \in \cat{B}$ the class of objects $A \in \cat{A}$ which fulfill $F(A) \iso B$ is non empty, because $F$ is essentially surjective on objects.
-            We define $G(B)$ as such an $A$ using the axiom of choice.
-            In addition we have an isomorphism $\epsln_B\from F(G(B)) \to B$.
+            $F$ is essentially surj. on obj. so $\forall B \in \cat{B}.\, \exists A \in \cat{A}.\, F(A) \iso B$.
+            We denote by $h(A, B) \from F(A) \to B$ this isomorphism.
+            For each $B \in \cat{B}$ we define $G(B)$ such that $F(G(B)) \iso B$ under $h(G(B), B)$ and $\epsln_B\from F(G(B)) \to B := h(G(B), B)$.
             Let $g\from B \to B'$ be a morphism in $\cat{B}$.
@@
             \end{align*}
-            We obtain $G(F(f)) = \eta_{A'} \circ f \circ \eta_A^{-1}$ because $F$ is faithful.
+            We obtain $G(F(f)) = \eta_{A'} \circ f \circ \eta_A^{-1}$ because $F$ is faithfull.
             This shows that $\eta$ is a natural isomorphism.

### 2017-05-30 sheets/basis-3/sheet-basis-3.tex
Update exercise 3.6

@@
 \begin{answer}
-  In the following we will consider a representation of the permutation as a
-  tuple of two elements $(a, b)$, where $a \mapsto b$ is an element of a
-  bijection.
+  In the following we will consider a graph representation of the permutation
+  function which is a tuple of two elements $(a, b)$, where $a \mapsto b$ is an
+  element of a bijection.
   \begin{itemize}
@@
       \begin{align*}
-        \Sym(f) &= \lambda (a, b). (f(a), f(b))\\
-        \Ord(f) &= \lambda (a, b).(f(a), f(b))
+        \Sym(f) &= \lambda X. \{(f(a), f(b)) | (a, b) \in X \}\\
+        \Ord(f) &= \lambda X. \{(f(a), f(b)) | (a, b) \in X \}\\
       \end{align*}
     \item[(b)]
-      Let's fix a finite set $A$ and define $\Sym \to \Ord$ on it.\\
+      Let's fix a finite set $A$ and try to define $\Sym \to \Ord$ on it.\\
       \begin{align*}
@@
       \end{align*}
-      Natural transformation $\alpha: \Sym(A) \to \Ord(A)$:\\
+      Natural transformation $\alpha_A: \Sym(A) \to \Ord(A)$:\\
       \begin{align*}
@@
       \end{align*}
-      Here $\alpha_A$ contradicts our definition of it and we can conclude that
-      it is not a natural transformation. The only other possibility of
-      defining $\alpha_A$ would be to switch the images of sets in the
-      definition, but the applying $f$ will yield us a similar contradictory
-      result.  From this we see that no matter how we define $\alpha_A$, the
-      naturality axiom will not hold for our $f$, hence we cannot have a
-      natural transformation for $\Sym(A) \to \Ord(A)$ and, in general, a
-      natural transformation $\Sym \to \Ord$.
+      Here we get a contradiction with the definition of $\alpha_A$ and we can
+      therefore conclude that it is not a natural transformation (naturality
+      axiom doesn't hold). The only other possibility of defining $\alpha_A$
+      (we can define this natural transformation in only two ways, as there are
+      only two arrows possible between $\Sym(A)$ and $\Ord(A)$ in $\Set$) would
+      be to swap the images in the definition of $\alpha_A$, but the
+      application of $f$ will yield us a similar contradictory result. From
+      this we see that no matter how we define $\alpha_A$, the naturality axiom
+      will not hold for our $f$, hence we cannot have a natural transformation
+      for $\Sym(A) \to \Ord(A)$ and, since in general we need the naturality
+      axiom to hold for all objects and arrows between them, a natural
+      transformation $\Sym \to \Ord$.
     \item[(c)]
@@
   So we know that $\Sym$ and $\Ord$ have exactly the same number of elements
-  and generally we can come up with a mapping between those elements, such that
-  $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$. However, we have also shown
-  that we cannot define a natural transformation $\Sym \to \Ord$ which gives
-  that there's no natural isomorphism between those categories.
+  and we can come up with a mapping between those elements, such that $\Sym(X)
+  \iso \Ord(X)$ for all $X \in \cat{B}$. However, we have also shown that we
+  cannot define a natural transformation $\Sym \to \Ord$ which gives that
+  there's no natural isomorphism between those categories.
 \end{answer}
@@
                     \end{align*}
+                    By analogous reasoning using $\epsln$ we obtain that $G$ is also faithful.
                 \item[$F$ is full:]
-                    %TODO
+                    Let $A, A' \in \cat{A}$ and $g \from F(A) \to F(A')$.
+                    We define $f := \eta_{A'}^{-1} \circ G(g) \circ \eta_A$ and use the naturality axiom for $\eta$ in order to get:
+                    \begin{align*}
+                        G(F(f)) & = \eta_A' \circ f \circ \eta_A^{-1}\\
+                                & = \eta_A' \circ \eta_{A'}^{-1} \circ G(g) \circ \eta_A \circ \eta_A^{-1}\\
+                                & = G(g)
+                    \end{align*}
+                    Since $G$ is faithful we obtain $F(f) = g$.
+                    Thus $F$ is full.
                 \item[$F$ is essentially surjective on objects:]

### 2017-05-30 sheets/basis-3/sheet-basis-3.tex
Revision of solution for 3.2

--- 2017-05-29 sheets/basis-3/sheet-basis-3.tex
@@
   Let $F\from \cat{A}^\op \to \cat{B}^\op$ be an object of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
+  $H(F)$ should be a functor $\cat{A} \to \cat{B}$.
   Because every object $A : \cat{A}^\op$ is also an object of $\cat{A}$ we can define
   $H(F)(A) := F(A)$.
-  Now let $f \from A \to B$ be an morphism in $\cat{A}^\op$. We have $F(f^\op)=g^\op$ for some morphism $g$ in $\cat{B}^\op$.
-  We define $H(F)(f) := F(f^\op)^\op = g$.
+  For every $A, A' : \cat{A}^\op$ and $f \from A \to B$ in $\cat{A}^\op$,
+  $f$ is by definition of $(.)^\op$ also a morphism $A' \to A$ in $\cat{A}$ and
+  $F_{A,A'}(f)$ yields a morphism $F(A) \to F(A')$ in $\cat{B}^\op$, being a morphism $F(A') \to F(A)$ in $\cat{B}$.
+  For given objects $A, A' : \cat{A}$ and a morphism $f \from A \to A'$ in $\cat{A}$, $H(F)(f)_{A,A'}$ is
+  supposed to yield a morphism $H(F)(A) \to H(F)(A') = F(A) \to F(A')$ in $\cat{B}$.
+  We can define $H(F)(f)_{A,A'} := F_{A',A}(f)$.
+  Note that this definition is in particular well typed and that the indices of the functors are normally implicit.
 
-  Now let $F, G \from \cat{A}^\op \to \cat{B}^\op$ be objects and
-  $\alpha \from F \to G$ be an morphism of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
-  $\alpha = (F(A) \stackrel{\alpha_A}{\to} G(A))_A$ is an natural transformation.
-  We define $H(\alpha) := \beta \from H(G) \to H(F) := (\alpha_A^\op)_A$ and have to show that $\beta$ is indeed such a natural transformation.
-  Let $A, A' : \cat{A}$ and $f \from A \to A'$. We have to show that the following diagram commutes:
-  \[
-    \begin{tikzcd}
-      H(G)(A) \arrow{r}{H(G)(f)} \arrow{d}{H(\alpha)_A} & H(G)(A') \arrow{d}{H(\alpha)_A'} \\
-      H(F)(A) \arrow{r}{H(F)(f)} & H(F)(A')
-    \end{tikzcd}
-  \]
-  By unfolding the definition of $H$ this diagram commutes if and only if the following diagram commutes:
-  \[
-    \begin{tikzcd}
-      G(A) & G(A') \arrow{l}{G(f^\op)} \\
-      F(A) \arrow{u}{\alpha_A} & F(A') \arrow{l}{F(f^\op)} \arrow{u}{\alpha_{A'}}
-    \end{tikzcd}
-  \]
-  This diagram commutes because $\alpha$ is a natural transformation.
+  Now let $F, G \from \cat{A}^\op \to \cat{B}^\op$ be concidered as objects of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$ and
+  $\alpha \from F \to G$ be concidered as a morphism of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$ from $F$ to $G$.
+  Now $\alpha = (F(A) \stackrel{\alpha_A}{\to} G(A))_{A : \cat{A}^\op}$ is a natural transformation.
+  Note that for every object $A : \cat{A}^\op$ (and in particular for every object $A : \cat{A}$)
+  the morphism $\alpha_{A} \from F(A) \to G(A)$ in $\cat{B}^\op$ is also a morphism $G(A) \to F(A)$ in $\cat{B}$.
+  $H(\alpha) = \beta$ is supposed to be morphism in ${\ftrcat{\cat{A}}{\cat{B}}}^\op$ from $F$ to $G$ and thus a natural transformation $\beta \from G \to A$.
+  As we have noted we can just define $H(\alpha) := \beta := (G(A) \stackrel{\alpha_A}{\to} F(A))_{A : \cat{A}} = \alpha$.
 
-  Secondly we define $J$ using a similar construction as above and get
-  $H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op}$ and $J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}$.
+  Secondly we define $J$ using the same constructions as above.
+  We observe that the only non-identical part of the definition is the part $H(F)(f)_{A,A'} := F_{A',A}(f)$ were we swap the (implic) indices $A$ and $A'$.
+  Because we do the same swap in the definition of $J(F)(f)$, the identities hold:
+  $$H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op} \quad\land\quad  J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}.$$
 
@@
   $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$
-  holds.
+  holds. Note that the constructions of $H$ and $J$ are canonical in the sense of the following definition.
 \end{answer}

### 2017-05-31 sheets/talk-4/sheet4.tex
fixed ex.1a

@@
 \begin{document}
-\[\uebunghead{Talk 4}{Galois Connections and Adjunctions}
+	\uebunghead{Talk 4}{Galois Connections and Adjunctions}
 	\author {Sarah Mameche}
+
 	\section{Galois connections}
-
-%Solutions preliminary
 
@@
 			\begin{enumerate}
-\item $p \preceq_P G (F (p)) \land q \preceq_Q F(G (q))$
-\item $p \preceq_P G(q) \Leftrightarrow q \preceq_Q F(p)$
+				\item $p \preceq_P G (F (p)) \land F(G (q)) \preceq_Q q$
+				\item $p \preceq_P G(q) \Leftrightarrow F(p) \preceq_Q q$
 			\end{enumerate}
@@
 			\begin{aligned}
- \forall q \in Q.\; G(q) = min\{p \in P \such F(p) \preceq_Q q\}
+			\forall q \in Q.\; G(q) = max\{p \in P \such F(p) \preceq_Q q\}
 			\end{aligned}
@@
 		\begin{enumerate}
-\item ...
-\item %For some $q \in Q$, set $S = \{p \in P \such F(p) \preceq_Q q}.
-Dual property: $\forall p \in P.\; F(p) = min\{q \in Q \such p \preceq_P G(q)\} $
+			\item (i) $\Rightarrow$ (ii):
+			\\Assume (i) and $p \preceq_P G(q)$. Since $F$ is monotone, $F(p) \preceq_Q F(G(q))$, and with (i):  $F(p) \preceq_Q F(G(q)) \preceq_Q q$.
+			\\Assume (i) and $F(p) \preceq_Q q$. Applying monotone $G$ and using (i), $ p \preceq_P G(F(p)) \preceq_P G(q)$.\\ \\
+			(ii) $\Rightarrow$ (i): \\Let $q := F(p)$. Since $F(p) \preceq_Q F(p)$, (ii) implies $p \preceq_P G (F (p))$. 
+			\\Similarly, with $p := G(q)$,  (ii) implies $F(G (q)) \preceq_Q q$.
+
+			\item  For some $q \in Q$, set $D = \{p \in P \such F(p) \preceq_Q q\}$. 
+			\\We know from (a) that $\forall q, F(G(q)) \preceq_Q q$, so $G(q) \in D$.
+			\\To show that $G(q)$ is also the maximum of D, take some $d \in D$, i.e. $F(d) \preceq_Q q$. 
+			\\With monotonicity of G, $G(F(d)) \preceq_Q G(q)$. From (a), $d \preceq_Q G(F(d))$, so $d \preceq_Q G(q)$.
+			\\
+			\\Dual property: $\forall p \in P.\; F(p) = min\{q \in Q \such p \preceq_P G(q)\} $
 		\end{enumerate}
@@
 	\begin{answer}
-\begin{enumerate} \item right adjunct: floor function  $\mathcal{F} : \mathcal{Q} \to  \mathcal{N}$ \\
-left adjunct: ceiling function $\mathcal{C} : \mathcal{Q} \to  \mathcal{N}$
-\item There can be no galois connection $F \dashv G$ from $\mathscr{Q}$ to $\mathscr{N} .
-F(q)=1 \Leftrightarrow 1 q<2,$ so $\{q\in \mathbb{Q}^+ \such F(q) = 1\} $ has no maximum.
-The statement follows with Ex.1b
-\end{enumerate}\end{answer}
+		\begin{enumerate} 
+			\item The right adjunct $G \from \mathbb{Q}^+ \to  \mathbb{N}$ is the floor function, because for $x \in \mathbb{N}, y \in \mathbb{Q}^+:$
+			\\$ x = \lfloor x \rfloor = G(I(x))  \land I(G(y)) =  \lfloor y \rfloor \leqslant y$.
+			\\The left adjunct $H \from \mathbb{Q}^+ \to  \mathbb{N}$ is the ceiling function.\\ For $x \in \mathbb{N}, y \in \mathbb{Q}^+: y \leqslant \lceil y \rceil = I(H(x))  \land H(I(x)) =  \lceil x \rceil = x$.\\
+
+			\item $I \dashv G$ but not $G \dashv I$. G has no right adjoint:
+			\\$\{q\in \mathbb{Q}^+ \such G(q) \leqslant 1\} $ has no maximum, because $G(q)= 1 \; \forall 1 \leqslant q < 2$. 
+			The statement follows with Ex.1 (b).
+		\end{enumerate}
+	\end{answer}
 
 	\section{Adjunctions}
+
 	\begin{exercise}
@@
 	\begin{answer}
-(i), (ii) $\Rightarrow$ (iii)\\
-composing any $q \from B \to B'$ with (ii): $\forall f, p: \forall q: q \of \bar{f} \of F(p) = q \of \overline{f \of p} $ (by ii).
-\\Since $f \from p \to F(A') \to B$, its transpose $h=:\overline{f \of p} \from A' \to B.$ \\So from (i) and quantif.
-over A: $q \of \overline{f \of p}  = q \of h = \overline{G(q) \of \bar{h}} = \overline{G(q) \of f \of p}$\\
+		(i), (ii) $\Rightarrow$ (iii)
+		\\$\forall q \from B \to B', \forall f, p. \;$\\$ q \of \bar{f} \of F(p) = q \of \overline{f \of p}\;\;$ ((ii) and naturality of composition)\\
+		$\Rightarrow q \of \bar{f} \of F(p)  = q \of h\;\;$ (*)\\
+		$\Leftrightarrow  q \of \bar{f} \of F(p) = \overline{G(q) \of \bar{h}}\;\;$ (i)\\
+		$\Leftrightarrow  q \of \bar{f} \of F(p)= \overline{G(q) \of f \of p}\;\;$(*)\\
+		In (*), $F \dashv G$ is used, so the function  $\overline{f \of p} \from A' \to G(B)$ has a transpose $ h =: f \of p \from F(A') \to B$.\\\\
 		(iii) $\Rightarrow$ (i)\\
-%(iii) quantifies over all $p$, so set $p = id_A$ (exists for each object). Then $f \of p \from A \to G(B)$, and $\bar{f} \of F(p) \from F(A) \to B.$ 
-Assume (i) is false. 
-Taking a $g \from F(A) \to B$, you get $h = g  \of id_B  \from F(A) \to B$. 
-Then the equality (i) does not hold for h, that is \\ 
-$\forall q, \overline{q \of h} \not= G(q) \of \bar{h}$ 
-\\$ \Leftrightarrow \overline{q \of g  \of id_B} \not= G(q)  \of \overline{h}$ ( Def h)\\
-$\Leftrightarrow q \of g  \of id_B \not= \overline{G(q)  \of \overline{h}}$ (Def transposition)
+		Assume (iii) and let $p = id_A. \; F(p) = id_B$ by the functor axioms. Due to $F \dashv G$, every $f\from A \to G(B)$ has a transpose $g := \bar{f}$. Then \\\\
+		$\forall q, f. \; \overline{G(q) \of f \of p} = q \of \bar{f} \of F(p)$\\
+		$\Rightarrow \forall q, g.  \;\overline{G(q) \of \bar{g} \of id_A} = q \of g \of id_B$ (instantiations) \\
+		$\Leftrightarrow \forall q, g.  \;\overline{G(q) \of \bar{g}} = q \of g$ (identities)\\
+		$\Leftrightarrow \forall q, g. \; G(q) \of \bar{g} = \overline{q \of g}$ (transposition)\\
 
-$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{h}} $  (define f as $\bar {g}.$ Also, $  F(id_A) = id_B$ by Def of functors)
-\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{g}} \; (h = g \of id_B = g)$
-\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of f \of id_A}\;  (f = \bar{g} $ by def. and $f = f \of id_A)$
-\\So (iii) is false for $f = \bar {g}$ and $ p= id_A$. \\(iii)  $\Rightarrow$ (i) similarly.
+		(iii)  $\Rightarrow$ (ii) analogously.
 	\end{answer}
@@
 	\end{exercise}
+
 	\begin{answer}
+		The right adjunct is the powerset functor $P \from $\textbf{Rel }$ \to $ \textbf{ Set}. 
+		It maps a set $A$ to its powerset $\mathcal{P}(A)$. 
+		A Relation $R \subseteq A \times B$ is mapped to the function $f_R \from \mathcal{P}(A) \to \mathcal{P}(B), f_R(X) = \{b \in B \such \exists a \in X. (a,b) \in R \}$. 
+		\\Every function $f \from A \to \mathcal{P}(A)$ can be seen as a relation $R_f \subseteq F(A) \times B$, with $(x,y) \in R \Leftrightarrow y \in f(x)$, and vice versa. 
+		So \textbf{Rel }$(F(A), B) \cong $\textbf{ Set}$(A, P(B))$.\\
+		Naturality in $A$: for functions $h \from A' \to A$,  the constructed relation for $f \of h$ is $R_{f \of h} = \{(x,y) \such  y \in f(h(x))\}.$ 
+		The relation $R_f \of P(h)$ is the same one:  $R_f \of P(h)= R_f \of R_h = \{(x,z) \such \exists y. \; z \in f(y) \land y \in h(x)\} = R_{f \of h}$. 
+		\\Naturality in $B$: for relations $R \subseteq B \times B', f_{B \times B'}$ is the same as $f_B \of F(B').$
 	\end{answer}
@@
 Also, an equivalence between two categories is both a left and right adjoint.
-\item The contravariant powerset functor $\mathcal{P} \from $\textbf {Set }$ \to$\textbf{ Set}$^{op}$ (see Ex.
-7) is self-adjoint.
-Functions $A \to \mathcal{P}(B)$ in \textbf{Set} correspond to subsets of $A \times B$ and thus to functions $B \to \mathcal{P}(A)$.
+
+			\item The contravariant powerset functor $\mathcal{P} \from $\textbf {Set }$ \to$\textbf{ Set}$^{op}$ (see Ex. 7) is self-adjoint. 
+			As seen in Exercise 4, Functions $A \to \mathcal{P}(B)$ in \textbf{Set} correspond to subsets of $A \times B$ and thus also to functions $B \to \mathcal{P}(A)$.
 		\end{enumerate}
@@
 	\begin{answer}
-Assume $I$ is initial in $ \mathscr{A}$, then maps $I \to A$ are unique $\forall A \in A$.
-This implies that maps  $I \to G(B)$ are unique $\forall B$ because $G(B) \in ob(\mathscr{A})$.
-From $F\dashv G$, we get that the corresponding maps $F(I) \to B$ are unique $\forall B$, so $F(I)$ initial in $\mathscr{B}$.\\ It can be shown in the same way that terminal objects are preserved by $G$.\end{answer}
+		Assume $I$ is initial in $ \mathscr{A}$, then $\forall A \in ob(\mathscr{A})$, the maps $I \to A$ are unique. 
+		This implies that $\forall B \in ob(\mathscr{B})$, the maps  $I \to G(B)$ are unique because $G(B) \in ob(\mathscr{A})$. 
+		From $F\dashv G$, we get that the corresponding maps $F(I) \to B$ are unique, so $F(I)$ initial in $\mathscr{B}$.\\\\
+		Dually, terminal objects are also preserved by $G$.
+	\end{answer}
 
@@
 	\begin{answer}
-$F \dashv F^{-1}$, meaning $F(f)(X) \subseteq Y \Leftrightarrow X \subseteq F^{-1}$.\\
-Assume $F \dashv F^{-1}$. Then $\forall x \in X, f(x) \in Y$ and thus $X \subseteq F^{-1}$\\
-Assume $X \subseteq F^{-1}$ and take $b \in F^{-1}$. By Def $f(x) = b$ for some x, and with $X \subseteq F^{-1}, b = f(x) \in Y$\\
-Groups: f homomorphism, $\mathcal{P}(G)$ set of subgroups, same proof \\
-Top. Sp: continuous maps and subspaces
-
+		$F \dashv F^{-1}$, meaning $F(X) \subseteq Y \Leftrightarrow X \subseteq F^{-1}(Y)$.\\
+		Assume $F(X) \subseteq Y$.  $\forall x \in X, f(x) \in Y$ and thus $X \subseteq F^{-1}(Y)$\\
+		Assume $X \subseteq F^{-1}(Y)$, then $\forall x \; x \in F(Y)$, which means $F(X) \subseteq Y$\\
+		For Groups, f is a homomorphism, and $\mathcal{P}(G)$ the set of subgroups.\\
+		For Topological Spaces, we have continuous maps and subspaces.
 	\end{answer}
@@
 	\begin{answer}
-There is a left adjoint $D$, where $D(A)$ is a discrete category with objects $A$.
-There is also a right adjoint $I$, where $I(A)$ is the category with objects $A$ and a unique arrow $a \to b$ for each $a,b \in A$.
+		There is a left adjoint $D$, where $D(A)$ is a discrete category with objects $A$ (for some set $A$). \\
+		For a category $\mathscr{B}$, the functors in \textbf{CAT} $(D(A), \mathscr{B})$ are fully defined by the object mapping, since $D(A)$ has only identity arrows.
+So the functors can be seen as functions in \textbf{Set} $(A$, ob($\mathscr{B}))$.
+		\\There is also a right adjoint $I$, where $I(A)$ is the category with objects $A$ and a unique arrow $a \to b$ for each $a,b \in A$.
+Here, functors in \textbf{CAT} $(\mathscr{B}, I(A))$ can also be seen as functions in \textbf{Set} (ob($\mathscr{B}), A)$.
+That is because a functor $F$ must map arrows $B \to B'$ to $F(B) \to F(B')$, and if there is a unique arrow $F(B) \to F(B')$, we need only the object mapping to describe the functor.
 	\end{answer}
@@
 
-\]
 \end{document}
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2017-06-01 sheets/basis-5/sheet-basis-5.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\DeclareMathOperator{\A}{\mathscr A}
\DeclareMathOperator{\B}{\mathscr B}
\DeclareMathOperator{\M}{\mathscr M}
\DeclareMathOperator{\Sc}{\mathscr S}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 5}{Adjunctions 2}

\author{Mostafa Abouhamra}

\begin{hint}
Read Chapters 2.2 and 2.3.
\end{hint}

\section{Units and Counits}

\begin{exercise}
\begin{enumerate}
\item Show that for any adjunction, the right adjoint is full and faithful if and only if the counit is an isomorphism.
\item An adjunction satisfying the equivalent conditions of part (a) is called a reflection. Can you give an example of a reflection?
\end{enumerate}
\end{exercise}

\begin{exercise}
Given a functor $F: \A \rightarrow \B$ and a category $\Sc$, there is a functor $F^*: [\B,\Sc] \rightarrow [\A, \Sc]$ defined on objects $Y \in [\B,\Sc]$ by $F^*(Y) = Y \of F$ and on maps $\alpha$ by $F^*(\alpha) = \alpha F$. Show that any adjunction 
\begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\A \arrow[r, shift left,"F"] 
	\& \B \arrow[l, shift left, "G"] 
\end{tikzcd}
with  $F \dashv G$ and category $\Sc$ give rise to an adjunction 
\begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	[\A,\Sc] \arrow[r, shift left,"G^*"] 
	\&  \arrow[l, shift left, "F^*"]  [\B,\Sc]
\end{tikzcd} 
with $G^* \dashv F^*$.
\end{exercise}

\section{Adjunctions by Initial Objects}

\begin{exercise}
Given two functors $F: \A \rightarrow \B$ and $G: \B \rightarrow \A$ and a natural transformation $\eta: 1_{\A} \rightarrow GF$ such that $\eta_A: A \rightarrow GF(A)$ is initial in ($A \Rightarrow G$) for every $A \in \A$, prove there is an adjunction 
\begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\A \arrow[r, shift left,"F"] 
	\& \B \arrow[l, shift left, "G"] 
\end{tikzcd}
with $F \dashv G$.
\end{exercise}

\begin{exercise}
Give an example of an adjunction and formalize it in 3 different ways:
\begin{enumerate}
\item Define the adjunction according to definition 2.2.1.
\item Give the unit and counit. Why do they satisfy the triangle identities?
\item Define the adjunction via the initial object.
\end{enumerate}
\end{exercise}

\begin{exercise}
What can be said about adjunctions between groups (regarded as one-object categories)?
\end{exercise}

\textbf{Corollary}: Let $G: \B \rightarrow \A$ be a functor. Then G has a left adjoint if and only if for each $A \in \A$, the category ($A \Rightarrow G$) has an initial object.\\

\begin{exercise}
State and prove the dual of this corollary.
\end{exercise}

\begin{exercise}
\begin{enumerate}
\item Give an example of an adjunction that is not an equivalence.
\item Let ($F, G, \eta, \epsilon$) be an equivalence of categories. Prove that F is left adjoint to G.
\end{enumerate}
\end{exercise}

\end{document}

### 2017-06-01 sheets/basis-5/sheet-basis-5.tex
fixed one exercise

@@
 
-\textbf{Corollary}: Let $G: \B \rightarrow \A$ be a functor. Then G has a left adjoint if and only if for each $A \in \A$, the category ($A \Rightarrow G$) has an initial object.\\
+\textbf{Fact}: Let $G: \B \rightarrow \A$ be a functor. Then G has a left adjoint if and only if for each $A \in \A$, the category ($A \Rightarrow G$) has an initial object.\\
 
 \begin{exercise}
-State and prove the dual of this corollary.
+State and prove the dual of the above fact.
 \end{exercise}
@@
 \begin{exercise}
+  Recall Definition 1.3.15 of the script, which states that an equivalence of categories $\A$ and $\B$ consists of two functors $F$ and $G$ and two natural isomorphisms $\eta$ and $\epsilon$ which relate their compositions to the respective identity functors.
 \begin{enumerate}

### 2017-06-04 sheets/talk-6/sheet-6.tex
My own eight exercises for talk 6.
More exercises are to come frome the script.

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Talk 6}{Special Constructions in Categories}
\author{Maximilian Wuttke}

\hint{Read the slides for the definitions of the six constructions}

\section{Products and Coproducts}

\begin{exercise}[Product is unique up-to isomorphism]
  Proof that for every category $\cat{A}$ all products are isomorphic. Thus we can speak of \emph{the} product.
\end{exercise}

\begin{definition}{Arbitrary\ product}
  Let $\cat{A}$ be a category, $I$ a set, and $(X_i)_{i \in I}$ a family of objects of $\cat{A}$.
  A \demph{product} of $(X_i)_{i \in I}$ consists of an object $P$ and a family of maps
  $\Bigl(P \toby{p_i} X_i\Bigr)_{i \in I}$
  with the property that for all objects $A$ and families of maps
  $\Bigl(A \toby{f_i} X_i\Bigr)_{i \in I}$
  there exists a unique map $\bar{f}\from A \to P$ such that
  $\forall i \in I: p_i \of \bar{f} = f_i$.
\end{definition}

\begin{exercise}[Greatest lower bound]
  Define the \emph{greatest lower bound} for an family of objects in a poset and proof that it is the arbitrary product of the family when the poset is seen as a category.
\end{exercise}

\begin{definition}{Arbitrary\ coproduct}
  Let $\cat{A}$ be a category, $I$ a set, and $(X_i)_{i \in I}$ a family of objects of $\cat{A}$.
  A \demph{coproduct} of $(X_i)_{i \in I}$ consists of an object $S$ and a family of maps
  $\Bigl(X_i \toby{p_i} S\Bigr)_{i \in I}$
  with the property that for all objects $A$ and families of maps
  $\Bigl(X_i \toby{f_i} A\Bigr)_{i \in I}$
  there exists a unique map $\bar{f}\from S \to A$ such that
  $\forall i \in I: \bar{f} \of p_i = f_i$.
\end{definition}

\begin{exercise}[Least upper bound]
  Define the \emph{least upper bound} for a class of objects in a poset and proof that it is the arbitrary binary coproduct of the poset seen as a category.
\end{exercise}

\begin{exercise}[Disjoint union]
  \begin{itemize}
    \item[(a)]For an index set $I$ and a family of sets $(X_i)_{i \in I}$, refine the arbitrary disjoint union $\coprod_{i \in I}{X_i}$.
    \item[(b)]Show that this arbitrary disjoint union is indeed the arbitrary coproduct in $\Set$.
  \end{itemize}
\end{exercise}

\section{Equalizer and Coequalizer}

\begin{definition}{Minus-Core of Vector Spaces}
  Let $V, W$ be Vector Spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
  We define the linear function $s - t := \lambda v\in V. \; v(s) - t(s)$.
  Remember the definition $\forall g \from V \to W: core(g) := \set{ v \in V | g(v) = 0 }$.
  Now $core (s - t) = \set{v \in V | s(v) = t (v) }$.
\end{definition}

\begin{exercise}[Equalizer of Vector Spaces]
  Let
  $\xymatrix{
      X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  }$
  be a diagramm in $Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagramm.
\end{exercise}

\begin{definition}{Reflexive,\ symmetric\ and transitive\ closure}
  Let $X$ be a arbitrary set and $R \subseteq X \times X$ a binary relation over $X$.
  We define $R_0 := R \cup \set{ (y, x) | (x, y) \in R} \cup \set{ (x, x) | x \in X}$,
  the the \emph{reflexive, symmetric closure} of R.
  Now we define inductively $\forall n \in \nat$:
  \[R_{n+1} := \set{(x_1, x_3) \in R \times R \; | \; \exists x_2 \in X, (x_1, x_2) \in R_n \land (x_2, x_3) \in R_n}.\]
  Now $S := \bigcup_{n \in \nat}{R_n}$ is the reflexive, symmetric and transitive closure of $R$.
\end{definition}

\begin{exercise}[Coequalizer of sets]
  In this exercise we will first define \emph{reflexive symmetric transitive closures} and will define the coequalizer of $\Set$.
  \begin{itemize}
    \item[(a)]First we will show that the reflexive, symmetric and transitive closure of a relation $R$ is the \emph{least} equivalence relation containing $R$.
      Let $R$ be a relation over a set $X$ and $S$ be the reflexive, symmetric and transitive closure of $R$. Show following lemmas:
      \begin{itemize}
        \item[(i)]Show $\forall n\in\nat: R_n$ and $S$ are reflexive and symmetric.
        \item[(ii)]Show $\forall n\in\nat: R \subseteq R_n$ and $R \subseteq S$.
        \item[(iii)]Show that $R_n$ is \emph{monotone}, i.\,e. $\forall n,m\in\nat: n\le m \Rightarrow R_n \subseteq R_m$.
        \item[(iv)]Now show that $S$ is transitive.
          \begin{hint}
            Let $m\in\nat$. Show $\forall x_1, x_2, x_3: (x_1, x_2) \in S \Rightarrow (x_2, x_3) \in R_m \Rightarrow (x_1, x_3) \in S$ via $\nat$-induction.
          \end{hint}
        \item[(v)]Now show that $R$ is the \emph{least} equivalence relation containing $R$, i.\,e.
          $$\forall R': \text{$R'$ is a equivalence relation} \land R \subseteq R' \Rightarrow S \subseteq R'.$$
      \end{itemize}
    \item[(b)]Now we will define the coequalizer of $\Set$.
      Let $X, Y \in \Set, s, t \from X \to Y$.
      We make the following definitions:
      \begin{alignat*}{2}
        R    &:=& \set{(s(a), t(a)) | a\in Y} \\
        \sim &:=& \text{reflexive, symmetric and transitive closure of $R$} \\
        C    &:=& Y / \sim
      \end{alignat*}
      Note that $[y] := \set{ y' \in Y | (y, y') \in \sim}$ is the equivalence class of $y$ and $Y / \sim := \set{ [y] | y \in Y}$ is the set of all equivalence classes of $\sim$.
      Also note that no equivalence class can be empty and for $[y]$ we call $y$ a representant of this class.
      We define $i \from Y \to E$ with $i(y) := [y]$.
      \begin{itemize}
        \item[(i)]Show that the following diagramm commutes:
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$
        \item[(ii)]Let
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^f & A
          }$
          be a cummutative diagramm. Show that $f$ respects equivalence, i.\,e. $\forall (y_1, y_2) \in \sim: f(y_1) = f(y_2)$.
          \begin{hint}
            ``Destruct'' $(y_1, y_2) \in \sim$ and do a $\nat$-induction.
          \end{hint}
        \item[(iii)]Now define $\bar{f} \from C \to A$, such that the following diagramm commutes and show that there can only be one:
          \[ \xymatrix{
            & & A  \\
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[ru]^f \ar[rd]_i \\
            & & C \ar@{.>}[uu]|{\bar{f}}
          } \]
        \item[(iv)]Show that the fork
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
          }$.
          is the coequalizer of
          $\xymatrix{
            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
          }$.
      \end{itemize}
  \end{itemize}
\end{exercise}

\section{Pullbacks and Pushouts}

\begin{exercise}[Pullbacks\,vs.\,producs]
  Show that if $Z$ is terminal and
  \[ \xymatrix{
    P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
    X \ar[r]^s & Z
  } \]
  is the pullback square in $\cat{A}$ of $X, Y, Z, s, t$ then $X, Y$ is the product in $\cat{A}$.
\end{exercise}

\begin{exercise}[Pushouts\,vs.\,coproducs]
  Show that if $Z$ is initial and
  \[ \xymatrix{
    Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \\
    Y \ar[r]^{p_2} & P
  } \]
  is the pullback square in $\cat{A}$ of $X, Y, Z, s, t$, then $X + Y$ is the coproduct in $\cat{A}$.
\end{exercise}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
% vim: ts=2 sts=2 sw=2 expandtab

### 2017-06-04 sheets/talk-6/sheet-6.tex
Solution for 6: Product is unique up-to isomorphism

@@
 \end{exercise}
+
+\begin{answer}
+  \begin{proof}
+    Let $X, Y \in \cat{A}$ and $P$ be a product of $X$ and $Y$ together with $p_1 \from P \to X, p_2 \from P \to Y$
+    and $P'$ be a product of $X$ and $Y$ together with $p'_1 \from P' \to X, p'_2 \from P' \to Y$.
+
+    Lets reconcider the definition of the \emph{product}:
+    For every object $A \in \cat{A}$ and arrows $f_1 \from A \to X, f_2 \from A \to Y$ we get that there exists a unique map
+    $\bar{f} \from A \to P$, such that
+    \[ \xymatrix{
+        &A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}}
+        \ar[rdd]^{f_2}&       \\
+                &P \ar[ld]^{p_1} \ar[rd]_{p_2}                  &       \\
+        X       &                                               &Y
+    } \]
+    commutes.
+    Now we can instanciate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instanciations we get the commuting diagramm
+    \[ \xymatrix{
+      & P \ar@/_/[ldd]_{p_1} \ar@/^/[rdd]^{p_2} & \\
+      & P \ar[ld]^{p_1} \ar@/_/@{.>}[dd]|{\bar{f}} \ar@{.>}[u]|{f} \ar[rd]_{p_2} & \\
+      X & & Y \\
+      & P' \ar[lu]_{p_1'} \ar@/_/@{.>}[uu]|{\tilde{f}} \ar@{.>}[d]|{f'} \ar[ru]^{p'_2} & \\
+      & P' \ar@/^/[luu]^{p'_1} \ar@/_/[ruu]_{p'_2} &
+    } \]
+    for exactly one $f, f', \tilde{f}, \bar{f}$. With the inner triangles we get:
+    \begin{alignat*}{2}
+      p_1  \of \tilde{f} &= p'_1 &\qquad p_2  \of \tilde{f} &= p'_2 \\
+      p'_1 \of \bar{f}   &= p_1  &       p'_2 \of \bar{f}   &= p_2
+    \end{alignat*}
+    By substitution we get
+    \begin{alignat*}{2}
+      p_1   \of (\tilde{f} \of \bar{f}) &= p_1  &\qquad p_2  \of (\tilde{f} \of \bar{f}) &= p_2 \\
+      p'_1  \of (\bar{f} \of \tilde{f}) &= p'_1 &       p'_2 \of (\bar{f} \of \tilde{f}) &= p'_2
+    \end{alignat*}
+    With the outer triangles we get:
+    \begin{alignat*}{5}
+      p_1  &\;\of&&\; f  &= p_1  &\qquad p_2  &\;\of&\; f  &= p_2 \\
+      p'_1 &\;\of&&\; f' &= p'_1 &\qquad p'_2 &\;\of&\; f' &= p'_2
+    \end{alignat*}
+    Note that $f$ and $f'$ are \emph{unique} and the first equations hold for $f = 1_P$ and $f = \tilde{f} \of \bar{f}$, thus $\tilde{f} \of \bar{f} = 1_P$.
+    The second equations holds for $f' = 1_{P'}$ and for $f' = \bar{f} \of \tilde{f}$, thus $\bar{f} \of \tilde{f} = 1_{P'}$.
+    Thus $P$ and $P'$ are isomorphic. \qedhere
+  \end{proof}
+\end{answer}
 

### 2017-06-04 sheets/talk-6/sheet-6.tex
Solutions for 6: Greatest lower bound and least upper bound

@@
 
+\begin{answer}
+  Let $(A, \le)$ be a poset concidered as a category and $I$ a set.
+  A \demph{lower bound} of a family $(x_i)_{i\in I}$ is an element $a \in A$, s.\,t. $\forall i \in I: a \le x_i$.
+  $z \in A$ is a \demph{greatest lower bound} of this family if $\forall a \in A: \text{$a$ is a lower bound} \Rightarrow a \le z$.
+
+  Let $z$ be a greatest lower bound of the family $(x_i)_{i\in I}$. We show that $z$ is a product of $(x_i)_{i\in I}$.
+  \begin{proof}
+    We first have to define a family of maps $\Bigl(z \toby{p_i} x_i\Bigr)_{i \in I}$.
+    This is trivial since we have $\forall i \in I: z \le x_i$.
+    Now let $a \in A$ be and objects and let $\Bigl(a \toby{f_i} x_i\Bigr)_{i \in I}$ be a family of maps from $a$.
+    The existing of this maps means that $a$ is a upper bound of $(x_i)_{i\in I}$.
+    Now by definition of least upper bound, $a \le z$, thus we have a arrow $\bar{f} \from a \to z$.
+    All arrows are unique, thus $\bar{f}$ is unique and the diagram commutes. \qedhere
+  \end{proof}
+\end{answer}
+
 \begin{definition}{Arbitrary\ coproduct}
@@
 \end{exercise}
+
+\begin{answer}
+  Let $(A, \le)$ be a poset concidered as a category and $I$ a set.
+  A \demph{upper bound} of a family $(x_i)_{i\in I}$ is an element $a \in A$, s.\,t. $\forall i \in I: x_i \le z$.
+  $z \in A$ is a \demph{least upper bound} of this family if $\forall a \in A: \text{$a$ is a upper bound} \Rightarrow z \le a$.
+
+  Let $z$ be a least upper bound of the family $(x_i)_{i\in I}$. We show that $z$ is a coproduct of $(x_i)_{i\in I}$.
+  \begin{proof}
+    We first have to define a family of maps $\Bigl(x_i \toby{p_i} z\Bigr)_{i \in I}$.
+    This is trivial since we have $\forall i \in I: x_i \le a$.
+    Now let $a \in A$ be and objects and let $\Bigl(x_i \toby{f_i} a\Bigr)_{i \in I}$ be a family of maps to $a$.
+    The existing of this maps means that $a$ is a lower bound of $(x_i)_{i\in I}$.
+    Now by definition of least lower bound, $z \le a$, thus we have a arrow $\bar{f} \from z \to a$.
+    All arrows are unique, thus $\bar{f}$ is unique and the diagram commutes. \qedhere
+  \end{proof}
+\end{answer}
 

### 2017-06-04 sheets/talk-6/sheet-6.tex
Sheet 6: Fix two typos

@@
   \begin{itemize}
-    \item[(a)]For an index set $I$ and a family of sets $(X_i)_{i \in I}$, refine the arbitrary disjoint union $\coprod_{i \in I}{X_i}$.
+    \item[(a)]For an index set $I$ and a family of sets $(X_i)_{i \in I}$, define the arbitrary disjoint union $\coprod_{i \in I}{X_i}$.
     \item[(b)]Show that this arbitrary disjoint union is indeed the arbitrary coproduct in $\Set$.
@@
             X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
-          }$.
+          }$
           is the coequalizer of

### 2017-06-05 sheets/talk-6/sheet-6.tex
New Exercise + Solutions for talk 6: Direct Sum of Vector Spaces

@@
 
+\begin{exercise}[Direct Sum of Vector Spaces]
+  In the talk we have seen that the direct sum $V \oplus W$ of two vector spaces $V, W$ over a fixed field $K$ is the coproduct of $V$ and $W$.
+  Show that it is also the product.
+\end{exercise}
+
+\begin{answer}
+  \begin{proof}
+    We first define the projection functions $p_1 \from V \oplus W \to V, p_2 \from V \oplus W \to W$:
+    \begin{alignat*}{1}
+      p_1(x,y) &:= x \\
+      p_2(x,y) &:= y
+    \end{alignat*}
+    Now let $A$ be a vector space and $f_1 \from A \to V, f_2 \from A \to W$ be linear functions.
+    We define $\bar{f}(a) := (f_1(a), f_(2))$. We clearly have
+    \begin{alignat*}{2}
+      p_1(\bar{f}(a)) &= p_1(f_1(a), f_2(a)) = f_1(a) \\
+      p_2(\bar{f}(a)) &= p_2(f_1(a), f_2(a)) = f_2(a).
+    \end{alignat*}
+    Thus the diagramm
+    \[ \xymatrix{
+        & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}} \ar[rdd]^{f_2} & \\
+        & {V \oplus W} \ar[ld]^{p_1} \ar[rd]_{p_2} & \\
+        V & & W
+    } \]
+    commutes. Now we have to show that $\bar{f}$ is the only map such that the above diagramm commutes.
+    Let $\tilde{f} \from A \to V \oplus W$ be a map such that $p_1 \of \tilde{f} = f_1 \land p_2 \of \tilde{f} = f_2$.
+    \[
+      \tilde{f}(a) = \bigl(p_1(\tilde{f}(a)), p_2(\tilde{f}(a))\bigr) = \bigl(p_1(\bar{f}(a)), p_2(\bar{f}(a))\bigr) = \bar{f}(a)\qedhere
+    \]
+  \end{proof}
+\end{answer}
+
 \section{Equalizer and Coequalizer}

### 2017-06-05 sheets/talk-6/sheet-6.tex
Solution for 6: Arbitrary Disjoint Union

@@
 \end{exercise}
+
+\begin{answer}
+  First let define the arbitrary disjoint union:
+  \begin{definition}{Arbitrary disjoint union}
+    Let $I$ be an index set and let $(x_i)_{i \in I}$ be a family of sets.
+    We define the arbitrary disjoint union of $(x_i)_{i\in I}$:
+    $$\coprod_{i \in I} {x_i} := \bigcup_{i \in I} \set{ (x, i) \;\big|\; i \in I \land x \in x_i}$$
+  \end{definition}
+  Now let $I$ be an index set and let $(x_i)_{i\in I}$ be a family of sets. We show that $\coprod_{i \in I} {x_i}$ is the coproduct of $(x_i)_{i\in I}$.
+  \begin{proof}
+    We first define the projection maps: $p_i(a) := (a, i)$.
+    Now let $A$ be a set and $\Bigl(x_i \toby{f_i} A\Bigr)_{i \in I}$ be a family of functions to $A$.
+    We define $\bar{f} : \coprod_{i \in I} {x_i} \to A$ with $\bar{f}(a, i) := f_i(a)$.
+    We have to show that $\bar{f}$ is the only function such that
+    $\forall i \in I: \bar{f} \of p_i = f_i$.
+    Let $i \in I$ and we check $a \in x_i$. $\bar{f}(p_i(a)) = \bar{f}(a,i) = f_i(a)$.
+    Now let $\tilde{f} : \coprod_{i \in I} {x_i} \to A$ be another function sucht that
+    $\forall i \in I: \tilde{f} \of p_i = f_i$.
+    Let $(a, i) \in \coprod_{i \in I} {x_i}$. $\bar{f}(a,i)= f_i(a) = \tilde{f}(p_i(a)) = \tilde{f}(a,i)$.\qedhere
+  \end{proof}
+\end{answer}
 

### 2017-06-05 sheets/talk-6/sheet-6.tex
Add hint on sheet 6 that "We will now assume that all the 6 constructions are unique up-to isomorphism."
Solution for 6: Minus-Core

@@
 \end{answer}
+
+\hint{We will now assume that all the 6 constructions are unique up-to isomorphism.}
 
@@
 
-\begin{definition}{Minus-Core of Vector Spaces}
+\begin{definition}{Minus-Core of Linear Maps in Vector Spaces}
   Let $V, W$ be Vector Spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
@@
 \end{exercise}
+
+\begin{answer}
+  \begin{proof}
+    Let $i \from core(s-t) \to X$ be a inclusion function. Let
+    $\xymatrix{
+      A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
+    }$
+    be a commuting diagramm. Note that $f A \subseteq core(s-t)$.
+    Thus we can define $\bar{f} \from A \to core(s-t)$ with $\bar{f}(a) := f(a)$.
+    We see that $\forall a \in A: i(\bar{f}(a))=\bar{f}(a)=f(a)$, thus $i \of \bar{f} = f$.
+    Let $\tilde{f} \from A \to core(s-t)$ be another map such that $i \of \tilde{f} = f$.
+    Now $\tilde{f}(a) = i(\tilde{f}(a)) = i(\bar{f}(a)) = \bar{f}(a)$.
+  \end{proof}
+\end{answer}
 

### 2017-06-05 sheets/talk-6/sheet-6.tex
Partial solutions for 6: Coequalizer of sets

@@
 
+\begin{answer}
+  \begin{itemize}
+    \item[(a)]
+      \begin{itemize}
+        \item[(i)]
+          \begin{proof}
+            To show that $S$ is symmetric and reflexive, we choose $n := 0$. $R_0$ is by construction symmetric and reflexive.
+            We will show symmetrie and reflexivity of $S_n$ seperately.
+            \begin{itemize}
+              \item[\emph{Reflexivity:}]
+                Let $x \in X$. We show $\forall n\in\nat: (x, x) \in R_n$ by induction on $n\in\nat$.
+                \begin{itemize}
+                  \item[$n=0$:]
+                    $R_0$ is by construction reflexive.
+                  \item[$n \rightarrow n+1$:]
+                    We know $(x, x) \in R_n$ by the inductive hypositive and show $(x, x) \in R_{n+1}$. By applying the definition of $R_{n+1}$
+                    it is enough tow show $(x, x) \in R_n \land (x, x) \in R_n$ that is true with the inductive hypositive.
+                \end{itemize}
+                \item[\emph{Symmetrie:}]
+                We show $\forall n\in\nat: \forall x_1,x_2 \in X: (x_1, x_2) \in R \Rightarrow (x_2, x_1) \in R$ by induction on $n\in\nat$.
+                \begin{itemize}
+                  \item[$n=0$:]
+                    $R_0$ is by construction symmetric.
+                  \item[$n \rightarrow n+1$:]
+                    We know $\forall x_1,x_2 \in X: (x_1, x_2) \in R_{n} \Rightarrow (x_2, x_1) \in R$ by the inductive hypositive and have to show
+                    $\forall x_1, x_2 \in X: (x_1, x_2) \in R_{n+1} \Rightarrow (x_2, x_1) \in R_{n+1}$.
+                    Let $x_1, x_2 \in X$ with $(x_1, x_2) \in R_{n+1}$. By unfolding the definition of $R_{n+1}$ we get a $x$ with
+                    $(x_1, x) \in R_n \land (x, x_2) \in R_n$ and have to show $(x_2, x_1) \in R_{n+1}$.
+                    We apply the definition of $R_{n+1}$ and show $(x_2, y) \in R_n \land (x, x_1) \in R_n$.
+                    Applying the inductive hypositive we it is enough to show $(x, x_2) \in R_n \land (x_1, y) \in R_n$ that is true by assumption. \qedhere
+                \end{itemize}
+            \end{itemize}
+          \end{proof} % (a) (i)
+
+        \item[(ii)]
+          \begin{proof}
+            To show that $S$ contains $R$ we choose $n := 0$. $R_0$ contains $R$ by construction. \\
+            We show $\forall n\in\nat: \forall x_1,x_2\in X: (x_1, x_2) \in R \Rightarrow (x_1, x_2) \in R_n$ by induction on $n\in\nat$.
+            \begin{itemize}
+              \item[$n=0$:]
+                $R_0$ contains $R$ by construction.
+              \item[$n \rightarrow n+1$:]
+                We assume $\forall x_1, x_2 : X, (x_1, x_2) \in R \Rightarrow (x_1, x_2) \in R_n$ by the induction hypositive.
+                Let $x_1, x_2 \in X$ and $(x_1, x_2) \in R$. We show $(x_1, x_2) \in R_{n+1}$.
+                We apply the definition of $R_{n+1}$ and show $(x_1, x_1) \in R_n \land (x_1, x_2) \in R_n$.
+                The first we know with (a)(i) and the second by assumption. \qedhere
+            \end{itemize}
+          \end{proof} % (a) (ii)
+
+        \item[(iii)]
+          \begin{proof}
+            Let $m,n\in\nat$ and $x_1, x_2 \in X$. We show $n \le m \Rightarrow (x_1, x_2) \in R_n \Rightarrow (x_1, x_2) \in R_m$ by induction on $n\in\nat$.
+            \begin{itemize}
+              \item[$n=0$:]
+                We have $n \le 0$ and thus $n = 0$ and $(x_1, x_2) \in R_0$. We have to show $(x_1, x_2) \in R_0$ that is true by assumption.
+              \item[$n \rightarrow n+1$:]
+                We know $n \le m+1$ and $(x_1, x_2) \in R_n$. By the induction hypositive we also know
+                $n \le m \Rightarrow (x_1, x_2) \in R_n \Rightarrow (x_1, x_2) \in R_m$.
+                We have to show $(x_1, x_2) \in R_{m+1}$. We have two cases:
+                \begin{itemize}
+                  \item[$n = m+1$:]
+                    The goal follows with the assumption $(x_1, x_2) \in R_{n=m+1}$.
+                  \item[$n \le m$:]
+                    We show $(x_1, x_2) \in R_{m+1}$ by applying the definition of $R_{m+1}$ and have to show $(x_1, x_1) \in R_m \land (x_1, x_2) \in R_m$.
+                    The first we know with (a)(i) and the second by assumption. \qedhere
+                \end{itemize}
+            \end{itemize}
+          \end{proof} % (a) (iii)
+
+        \item[(iv)]
+          \begin{proof}
+          % TODO
+          \end{proof} % (a) (iv)
+
+        \item[(v)]
+          \begin{proof}
+          % TODO
+          \end{proof} % (a) (iv)
+
+      \end{itemize}
+
+      % TODO Do we need part (a) at all?
+    \item[(b)]
+      \begin{itemize}
+        \item[(i)]
+          \begin{proof}
+            Let $x \in X$. We have $i(s(x)=[s(x)]$ and $i(t(x))=[t(x)]$. It remains to show that $[s(x)]=[t(x)]$. Therefore it is enough to show that $s(x) \sim t(x)$.
+            This is true since $\sim$ contains $R$ and $(s(x), t(x)) \in R$ by the construction of $R$.
+          \end{proof} % (b) (i)
+
+        \item[(ii)]
+          \begin{proof}
+            Let $(y_1, y_2) \in \sim$. Then be definition of $\sim$ we have a $n\in\nat$ such that $(y_1, y_2) \in R_n$.
+            We show $\forall n\in\nat: \forall y_1,y_2 \in Y: (y_1, y_2) \in R_n \Rightarrow f (y_1) = f (y_2)$ by induction on $n\in\nat$.
+            \begin{itemize}
+              \item[$n=0$:]
+                We have $(y_1, y_2) \in R_0$. This means $(y_1, y_2) \in R \lor (y_2, y_1) \in R \lor y_1 = y_2)$.
+                We have to show $f(y_1) = f(y_2)$. Case analysis.
+                \begin{itemize}
+                  \item[1.] $(y_1,y_2) \in R$. Then we have $y_1 = s(a) \land y_2 = t(t)$ for some $a \in X$. $f (s (a)) = f (t (a))$ holds since $f$ commutes.
+                  \item[2.] $(y_1,y_2) \in R$. Then we have $y_2 = s(a) \land y_1 = t(t)$ for some $a \in X$. $f (t (a)) = f (s (a))$ holds since $f$ commutes.
+                  \item[3.] $y_1=y_2$. Then we have $f(y_1) = f(y_2)$.
+                \end{itemize}
+              \item[$n \rightarrow n+1$:]
+                We have $(y_1, y_2) \in R_{n+1}$. Therefore we have a $y$ such that $(y_1, y) \in R_n \land (y, y_2) \in R_n$.
+                With the inductive hypositive we know $\forall y_1, y_2 \in Y: (y_1, y_2) \in R_n \Rightarrow f (y_1) = f (y_2)$.
+                By instanciating the inductive hypositive we know $f (y_1) = f (y) \land f (y) = f (y_2)$.
+                Thus we get $f (y_1) = f (y_2)$ \qedhere
+            \end{itemize}
+          \end{proof} % (b) (ii)
+
+        \item[(iii)]
+          \newcommand{\choice}[1]{\set{#1}}
+          We have to assume the axiom of choice. Because every equivalence class is non-empty, we get a function
+          $\choice{ \_ } \from Y/\sim \to Y$ such that $\choice{ e } \sim x$ for every equivalence class $e$ and every object $x \in e$.
+
+          We now define $\bar{f}(e) := f (\choice e)$.
+
+          We have to show that the diagramm commutes. Note that commutativity of this diagramm is equivalent to $\bar{f} \of i =f$.
+          Let $y \in Y$. We have to show $\bar{f}(\choice{i (y)}) = f(y)$. With (ii) it is enough to show $\choice{i (y)} \sim y$.
+          This holds since $\choice{i (y)} = \choice{[y]} \sim y$, since $y \sim y$. \qedhere
+
+          We finally have to show that $\bar{f}$ is the only function with that property.
+          Let $\tilde{f} \from C \to A$ with $\tilde{f} \of i = f$. Let $c \in Y /\sim$.
+          Then $\bar{f}(c) = f(\choice{c}) = \tilde{f}(i(\choice{c}) = \tilde{f}([\choice{c}]) = \tilde{f}(c)$.
+
+        \item[(iv)]We know that there is only \emph{one} coequalizer (up-to isomorphism). By definition we have shown that our construction is \emph{a} coequalizer.
+      \end{itemize}
+
+  \end{itemize}
+\end{answer}
+
 \section{Pullbacks and Pushouts}

### 2017-06-06 sheets/talk-6/sheet-6.tex

--- 2017-06-05 sheets/talk-6/sheet-6.tex
@@
 
+\newcommand{\choice}[1]{\{ #1 \}}
+\newcommand{\class} [1]{[ #1 ]}
+
 \begin{exercise}[Coequalizer of sets]
-  In this exercise we will first define \emph{reflexive symmetric transitive closures} and will define the coequalizer of $\Set$.
-  \begin{itemize}
-    \item[(a)]First we will show that the reflexive, symmetric and transitive closure of a relation $R$ is the \emph{least} equivalence relation containing $R$.
-      Let $R$ be a relation over a set $X$ and $S$ be the reflexive, symmetric and transitive closure of $R$. Show following lemmas:
-      \begin{itemize}
-        \item[(i)]Show $\forall n\in\nat: R_n$ and $S$ are reflexive and symmetric.
-        \item[(ii)]Show $\forall n\in\nat: R \subseteq R_n$ and $R \subseteq S$.
-        \item[(iii)]Show that $R_n$ is \emph{monotone}, i.\,e. $\forall n,m\in\nat: n\le m \Rightarrow R_n \subseteq R_m$.
-        \item[(iv)]Now show that $S$ is transitive.
-          \begin{hint}
-            Let $m\in\nat$. Show $\forall x_1, x_2, x_3: (x_1, x_2) \in S \Rightarrow (x_2, x_3) \in R_m \Rightarrow (x_1, x_3) \in S$ via $\nat$-induction.
-          \end{hint}
-        \item[(v)]Now show that $R$ is the \emph{least} equivalence relation containing $R$, i.\,e.
-          $$\forall R': \text{$R'$ is a equivalence relation} \land R \subseteq R' \Rightarrow S \subseteq R'.$$
-      \end{itemize}
-    \item[(b)]Now we will define the coequalizer of $\Set$.
+  In this exercise we will will define the coequalizer of $\Set$ with the help of reflexive, symmetric and transitive closures.
   Let $X, Y \in \Set, s, t \from X \to Y$.
@@
   \begin{alignat*}{2}
-        R    &:=& \set{(s(a), t(a)) | a\in Y} \\
-        \sim &:=& \text{reflexive, symmetric and transitive closure of $R$} \\
-        C    &:=& Y / \sim
+    R    &:= \set{(s(a), t(a)) | a\in Y} \\
+    \sim &:= \text{reflexive, symmetric and transitive closure of $R$} \\
+    C    &:= Y / \sim
   \end{alignat*}
-      Note that $[y] := \set{ y' \in Y | (y, y') \in \sim}$ is the equivalence class of $y$ and $Y / \sim := \set{ [y] | y \in Y}$ is the set of all equivalence classes of $\sim$.
+  Note that $\class{y} := \set{ y' \in Y | (y, y') \in \sim}$ is the equivalence class of $y$ and $Y / \sim := \set{ \class{y} | y \in Y}$ is the set of all equivalence classes of $\sim$.
   Also note that no equivalence class can be empty and for $[y]$ we call $y$ a representant of this class.
-      We define $i \from Y \to E$ with $i(y) := [y]$.
+  We define $i \from Y \to E$ with $i(y) := \class{y}$.
   \begin{itemize}
-        \item[(i)]Show that the following diagramm commutes:
+    \item[(a)]Show that $\sim$ is reflexive and symmetric and contains $R$.
+    \item[(b)]Show that the following diagramm commutes:
       $\xymatrix{
@@
       }$
-        \item[(ii)]Let
+    \item[(c)]Let
       $\xymatrix{
@@
       \end{hint}
-        \item[(iii)]Now define $\bar{f} \from C \to A$, such that the following diagramm commutes and show that there can only be one:
+    \item[(d)]Now define $\bar{f} \from C \to A$, such that the following diagramm commutes and show that there can only be one:
       \[ \xymatrix{
@@
       } \]
-        \item[(iv)]Show that the fork
+      \hint{Assume the \demph{axiom of choice.}}
+    \item[(e)]Show that the fork
       $\xymatrix{
@@
       }$
-          is the coequalizer of
+      is \emph{the} coequalizer of
       $\xymatrix{
@@
   \end{itemize}
+\end{exercise}
+
+\begin{answer}
+  \begin{itemize}
+    \item[(a)]
+      This follows because for $n:=0$, $R_0$ is by construction reflexive, symmetric and cointains $R$.
+
+    \item[(b)]
+      \begin{proof}
+        Let $x \in X$. We have $i(s(x)=\class{s(x)}$ and $i(t(x))=\class{t(x)}$. It remains to show that $\class{s(x)}=\class{t(x)}$. Therefore it is enough to show that $s(x) \sim t(x)$.
+        This is true since $\sim$ contains $R$ and $(s(x), t(x)) \in R$ by the construction of $R$.
+      \end{proof} % (b) (i)
+
+    \item[(c)]
+      \begin{proof}
+        Let $(y_1, y_2) \in \sim$. Then be definition of $\sim$ we have a $n\in\nat$ such that $(y_1, y_2) \in R_n$.
+        We show $\forall n\in\nat: \forall y_1,y_2 \in Y: (y_1, y_2) \in R_n \Rightarrow f (y_1) = f (y_2)$ by induction on $n\in\nat$.
+        \begin{itemize}
+          \item[$n=0$:]
+            We have $(y_1, y_2) \in R_0$. This means $(y_1, y_2) \in R \lor (y_2, y_1) \in R \lor y_1 = y_2)$.
+            We have to show $f(y_1) = f(y_2)$. Case analysis.
+            \begin{itemize}
+              \item[1.] $(y_1,y_2) \in R$. Then we have $y_1 = s(a) \land y_2 = t(t)$ for some $a \in X$. $f (s (a)) = f (t (a))$ holds since $f$ commutes.
+              \item[2.] $(y_1,y_2) \in R$. Then we have $y_2 = s(a) \land y_1 = t(t)$ for some $a \in X$. $f (t (a)) = f (s (a))$ holds since $f$ commutes.
+              \item[3.] $y_1=y_2$. Then we have $f(y_1) = f(y_2)$.
             \end{itemize}
-\end{exercise}
+          \item[$n \rightarrow n+1$:]
+            We have $(y_1, y_2) \in R_{n+1}$. Therefore we have a $y$ such that $(y_1, y) \in R_n \land (y, y_2) \in R_n$.
+            With the inductive hypositive we know $\forall y_1, y_2 \in Y: (y_1, y_2) \in R_n \Rightarrow f (y_1) = f (y_2)$.
+            By instanciating the inductive hypositive we know $f (y_1) = f (y) \land f (y) = f (y_2)$.
+            Thus we get $f (y_1) = f (y_2)$ \qedhere
+        \end{itemize}
+      \end{proof} % (b) (ii)
+
+    \item[(d)]
+      We have to assume the axiom of choice. Because every equivalence class is non-empty, we get a function
+      $\choice{ \_ } \from Y/\sim \to Y$ such that $\choice{ e } \sim x$ for every equivalence class $e$ and every object $x \in e$.
+
+      We now define $\bar{f}(e) := f (\choice e)$.
+
+      We have to show that the diagramm commutes. Note that commutativity of this diagramm is equivalent to $\bar{f} \of i =f$.
+      Let $y \in Y$. We have to show $\bar{f}(\choice{i (y)}) = f(y)$. With (c) it is enough to show $\choice{i (y)} \sim y$.
+      This holds since $\choice{i (y)} = \choice{\class{y}} \sim y$, since $y \sim y$. \qedhere
+
+      We finally have to show that $\bar{f}$ is the only function with that property.
+      Let $\tilde{f} \from C \to A$ with $\tilde{f} \of i = f$. Let $c \in Y /\sim$.
+      Then $\bar{f}(c) = f(\choice{c}) = \tilde{f}(i(\choice{c}) = \tilde{f}(\class{\choice{c}}) = \tilde{f}(c)$.
+
+    \item[(e)]We know that there is only \emph{one} coequalizer (up-to isomorphism). By definition we have shown that our construction is \emph{a} coequalizer.
+  \end{itemize}
+\end{answer}
 

### 2017-06-06 sheets/talk-6/sheet-6.tex
Sheet 6: Rewording, Removal of Exercise 3, 7 and 8

--- 2017-06-05 sheets/talk-6/sheet-6.tex
@@
 
-\hint{Read the slides for the definitions of the six constructions}
+\hint{Read the slides for the definitions of the six constructions.}
 
@@
 \begin{exercise}[Product is unique up-to isomorphism]
-  Proof that for every category $\cat{A}$ all products are isomorphic. Thus we can speak of \emph{the} product.
+  Prove that for every category $\cat{A}$ and every objects $X, Y \in \cat{A}$
+  all products of $X$ and $Y$ are isomorphic. Thus we can speak of \emph{the} product of $X$ and $Y$.
 \end{exercise}
@@
 \begin{exercise}[Greatest lower bound]
-  Define the \emph{greatest lower bound} for an family of objects in a poset and proof that it is the arbitrary product of the family when the poset is seen as a category.
+  Define the \emph{greatest lower bound} for a family of objects in a poset and prove that it is the arbitrary product of the family when the poset is seen as a category.
 \end{exercise}
@@
 \end{definition}
-
-\begin{exercise}[Least upper bound]
-  Define the \emph{least upper bound} for a class of objects in a poset and proof that it is the arbitrary binary coproduct of the poset seen as a category.
-\end{exercise}
-
-\begin{answer}
-  Let $(A, \le)$ be a poset concidered as a category and $I$ a set.
-  A \demph{upper bound} of a family $(x_i)_{i\in I}$ is an element $a \in A$, s.\,t. $\forall i \in I: x_i \le z$.
-  $z \in A$ is a \demph{least upper bound} of this family if $\forall a \in A: \text{$a$ is a upper bound} \Rightarrow z \le a$.
-
-  Let $z$ be a least upper bound of the family $(x_i)_{i\in I}$. We show that $z$ is a coproduct of $(x_i)_{i\in I}$.
-  \begin{proof}
-    We first have to define a family of maps $\Bigl(x_i \toby{p_i} z\Bigr)_{i \in I}$.
-    This is trivial since we have $\forall i \in I: x_i \le a$.
-    Now let $a \in A$ be and objects and let $\Bigl(x_i \toby{f_i} a\Bigr)_{i \in I}$ be a family of maps to $a$.
-    The existing of this maps means that $a$ is a lower bound of $(x_i)_{i\in I}$.
-    Now by definition of least lower bound, $z \le a$, thus we have a arrow $\bar{f} \from z \to a$.
-    All arrows are unique, thus $\bar{f}$ is unique and the diagram commutes. \qedhere
-  \end{proof}
-\end{answer}
 
@@
   We define the linear function $s - t := \lambda v\in V. \; v(s) - t(s)$.
-  Remember the definition $\forall g \from V \to W: core(g) := \set{ v \in V | g(v) = 0 }$.
+  Remember that for $g \from V \to W$ we have defined $core(g) := \set{ v \in V | g(v) = 0 }$.
   Now $core (s - t) = \set{v \in V | s(v) = t (v) }$.
@@
 
-\begin{exercise}[Coequalizer of sets]
-  In this exercise we will first define \emph{reflexive symmetric transitive closures} and will define the coequalizer of $\Set$.
-  \begin{itemize}
-    \item[(a)]First we will show that the reflexive, symmetric and transitive closure of a relation $R$ is the \emph{least} equivalence relation containing $R$.
-      Let $R$ be a relation over a set $X$ and $S$ be the reflexive, symmetric and transitive closure of $R$. Show following lemmas:
-      \begin{itemize}
-        \item[(i)]Show $\forall n\in\nat: R_n$ and $S$ are reflexive and symmetric.
-        \item[(ii)]Show $\forall n\in\nat: R \subseteq R_n$ and $R \subseteq S$.
-        \item[(iii)]Show that $R_n$ is \emph{monotone}, i.\,e. $\forall n,m\in\nat: n\le m \Rightarrow R_n \subseteq R_m$.
-        \item[(iv)]Now show that $S$ is transitive.
-          \begin{hint}
-            Let $m\in\nat$. Show $\forall x_1, x_2, x_3: (x_1, x_2) \in S \Rightarrow (x_2, x_3) \in R_m \Rightarrow (x_1, x_3) \in S$ via $\nat$-induction.
-          \end{hint}
-        \item[(v)]Now show that $R$ is the \emph{least} equivalence relation containing $R$, i.\,e.
-          $$\forall R': \text{$R'$ is a equivalence relation} \land R \subseteq R' \Rightarrow S \subseteq R'.$$
-      \end{itemize}
-    \item[(b)]Now we will define the coequalizer of $\Set$.
-      Let $X, Y \in \Set, s, t \from X \to Y$.
-      We make the following definitions:
-      \begin{alignat*}{2}
-        R    &:=& \set{(s(a), t(a)) | a\in Y} \\
-        \sim &:=& \text{reflexive, symmetric and transitive closure of $R$} \\
-        C    &:=& Y / \sim
-      \end{alignat*}
-      Note that $[y] := \set{ y' \in Y | (y, y') \in \sim}$ is the equivalence class of $y$ and $Y / \sim := \set{ [y] | y \in Y}$ is the set of all equivalence classes of $\sim$.
-      Also note that no equivalence class can be empty and for $[y]$ we call $y$ a representant of this class.
-      We define $i \from Y \to E$ with $i(y) := [y]$.
-      \begin{itemize}
-        \item[(i)]Show that the following diagramm commutes:
-          $\xymatrix{
-            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
-          }$
-        \item[(ii)]Let
-          $\xymatrix{
-            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^f & A
-          }$
-          be a cummutative diagramm. Show that $f$ respects equivalence, i.\,e. $\forall (y_1, y_2) \in \sim: f(y_1) = f(y_2)$.
-          \begin{hint}
-            ``Destruct'' $(y_1, y_2) \in \sim$ and do a $\nat$-induction.
-          \end{hint}
-        \item[(iii)]Now define $\bar{f} \from C \to A$, such that the following diagramm commutes and show that there can only be one:
-          \[ \xymatrix{
-            & & A  \\
-            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[ru]^f \ar[rd]_i \\
-            & & C \ar@{.>}[uu]|{\bar{f}}
-          } \]
-        \item[(iv)]Show that the fork
-          $\xymatrix{
-            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
-          }$
-          is the coequalizer of
-          $\xymatrix{
-            X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
-          }$.
-      \end{itemize}
-  \end{itemize}
-\end{exercise}
-
 \section{Pullbacks and Pushouts}
 
-\begin{exercise}[Pullbacks\,vs.\,producs]
-  Show that if $Z$ is terminal and
-  \[ \xymatrix{
-    P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
-    X \ar[r]^s & Z
-  } \]
-  is the pullback square in $\cat{A}$ of $X, Y, Z, s, t$ then $X, Y$ is the product in $\cat{A}$.
-\end{exercise}
-
-\begin{exercise}[Pushouts\,vs.\,coproducs]
-  Show that if $Z$ is initial and
+\begin{exercise}[Pushouts\,vs.\,coproducts]
+  Show that if $Z$ is initial and the pullback square
   \[ \xymatrix{
@@
   } \]
-  is the pullback square in $\cat{A}$ of $X, Y, Z, s, t$, then $X + Y$ is the coproduct in $\cat{A}$.
+  is the pullback of
+  $ \xymatrix{
+    X & Z \ar[l]_s \ar[r]^t & Y \\
+  } $
+  then $P$ is the coproduct of $X$ and $Y$.
+
+  Remark: Note that the dual statement holds for pullbacks of terminal objects and products.
 \end{exercise}

### 2017-06-07 sheets/talk-6/sheet-6.tex
Sheet 6: Import exercises from the script

@@
 
+\begin{exercise}
+  Let $\parpair{X}{Y}{s}{t}$ be maps in some category.  Prove that $s = t$ if
+  and only if the equalizer of $s$ and $t$ exists and is an isomorphism, if
+  and only if the coequalizer of $s$ and $t$ exists and is an isomorphism.
+\end{exercise}
+
+\begin{exercise}
+  Let $X$ be a set and $f\from X \to X$ a map.  Describe the coequalizer of
+  $\parpairi{X}{X}{f}{1}$ in $\Set$ as explicitly as possible.
+\end{exercise}
+
 \section{Pullbacks and Pushouts}
@@
 
+\begin{exercise}
+  Take objects and maps $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f}
+  \ar@<-.5ex>[r]_g &Y}$ in some category.  If this is an equalizer, is the
+  square
+  \[ \xymatrix{
+    E \ar[r]^i \ar[d]_i &
+    X \ar[d]^g \\
+    X \ar[r]_f &
+    Y
+  } \]
+  necessarily a pullback?
+  What about the converse?  Give proofs or counterexamples.
+\end{exercise}
+
+\begin{exercise}
+  Take a commutative diagram
+  \[ \xymatrix{
+    \cdot \ar[r] \ar[d] & \cdot \ar[r] \ar[d] & \cdot \ar[d] \\
+    \cdot \ar[r] & \cdot \ar[r] & \cdot
+  } \]
+  in some category.  Suppose that the right-hand square is a pullback.  Show
+  that the left-hand square is a pullback if and only if the outer rectangle is
+  a pullback.
+\end{exercise}
+
 \end{document}

### 2017-06-08 sheets/talk-6/sheet-6.tex

@@
 
-\begin{exercise}[Direct Sum of Vector Spaces]
+\begin{exercise}[Direct sum of vector spaces]
   In the talk we have seen that the direct sum $V \oplus W$ of two vector spaces $V, W$ over a fixed field $K$ is the coproduct of $V$ and $W$.
@@
 
-\begin{definition}{Minus-Core of Linear Maps in Vector Spaces}
-  Let $V, W$ be Vector Spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
+\begin{definition}{Minus-core of linear maps in vector spaces}
+  Let $V, W$ be vector spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
   We define the linear function $s - t := \lambda v\in V. \; v(s) - t(s)$.
@@
 
-\begin{exercise}[Equalizer of Vector Spaces]
+\begin{exercise}[Equalizer of vector spaces]
   Let
@@
 \end{answer}
-
-\begin{definition}{Reflexive,\ symmetric\ and transitive\ closure}
-  Let $X$ be a arbitrary set and $R \subseteq X \times X$ a binary relation over $X$.
-  We define $R_0 := R \cup \set{ (y, x) | (x, y) \in R} \cup \set{ (x, x) | x \in X}$,
-  the the \emph{reflexive, symmetric closure} of R.
-  Now we define inductively $\forall n \in \nat$:
-  \[R_{n+1} := \set{(x_1, x_3) \in R \times R \; | \; \exists x_2 \in X, (x_1, x_2) \in R_n \land (x_2, x_3) \in R_n}.\]
-  Now $S := \bigcup_{n \in \nat}{R_n}$ is the reflexive, symmetric and transitive closure of $R$.
-\end{definition}
 

### 2017-06-08 sheets/talk-6/sheet-6.tex
minor fixes

@@
   Let $V, W$ be vector spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
-  We define the linear function $s - t := \lambda v\in V. \; v(s) - t(s)$.
+  We define the linear function $s - t := \lambda v\in V. \; s(v) - t(v)$.
   Remember that for $g \from V \to W$ we have defined $core(g) := \set{ v \in V | g(v) = 0 }$.
@@
 \begin{exercise}[Pushouts\,vs.\,coproducts]
-  Show that if $Z$ is initial and the pullback square
+  Show that if $Z$ is initial and the pushout square
   \[ \xymatrix{
@@
   } \]
-  is the pullback of
+  is the pushout of
   $ \xymatrix{

### 2017-06-08 sheets/talk-6/sheet-6.tex
Sample solution Talk 6.8

@@
 
+\begin{answer}
+  \begin{proof}
+    We show that $S := P$ is the coproduct of $X$ and $Y$ together with the projection functions $p_1$ and $p_2$.
+    First note that because $Z$ is initial $s$ and $t$ are unique maps.
+    Let
+    $ \xymatrix{
+      X \ar[r]^{f_1} & A & Y \ar[l]_{f_2}
+    } $
+    be a diagramm in $\cat{A}$.
+    We have to show that there is a unique map $\bar{f}' \from P \to A$ shuch that
+    \[ \xymatrix{
+      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
+      & P \ar@{.>}[d]|{\bar{f}'} & \\
+      & A &
+    } \]
+    commutes. We know that the diagramm
+    \[ \xymatrix{
+      Z \ar[r]_s \ar[d]^t & X \ar[d]_{f_1} \\
+      Y \ar[r]^{f_2} & A
+    } \]
+    commutes because $Z$ is initial.
+    Now because this square is then a pushout square there exists a unique map $\bar{f}\from P \to A$ such that
+    \[ \xymatrix{
+      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
+      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\bar{f}} & \\
+      & & A  \\
+    } \]
+    commutes. We see that we can choose $\bar{f}' := \bar{f}$.
+    Suppose we have another $\tilde{f} \from P \to A$ such that
+    \[ \xymatrix{
+      X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
+      & P \ar@{.>}[d]|{\tilde{f}} & \\
+      & A &
+    } \]
+    commutes. Then
+    \[ \xymatrix{
+      Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
+      Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\tilde{f}} & \\
+      & & A  \\
+    } \]
+    would also commute (because $S$ is initial) but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
+  \end{proof}
+
+\end{answer}
+
 \begin{exercise}

### 2017-06-08 sheets/talk-6/sheet-6.tex
Sample solution Talk 6.8

@@
     commutes.
-    Now we can instanciate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instanciations we get the commuting diagramm
+    Now we can instanciate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instanciations we get the commuting diagram
     \[ \xymatrix{
@@
     \end{alignat*}
-    Thus the diagramm
+    Thus the diagram
     \[ \xymatrix{
@@
     } \]
-    commutes. Now we have to show that $\bar{f}$ is the only map such that the above diagramm commutes.
+    commutes. Now we have to show that $\bar{f}$ is the only map such that the above diagram commutes.
     Let $\tilde{f} \from A \to V \oplus W$ be a map such that $p_1 \of \tilde{f} = f_1 \land p_2 \of \tilde{f} = f_2$.
@@
   }$
-  be a diagramm in $Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagramm.
+  be a diagram in $Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagram.
 \end{exercise}
@@
     }$
-    be a commuting diagramm. Note that $f A \subseteq core(s-t)$.
+    be a commuting diagram. Note that $f A \subseteq core(s-t)$.
     Thus we can define $\bar{f} \from A \to core(s-t)$ with $\bar{f}(a) := f(a)$.

### 2017-06-08 sheets/talk-6/sheet-6.tex
Typo hotfix: s/diagram*/diagram/
Sample solution 6.9

--- 2017-06-08 sheets/talk-6/sheet-6.tex
@@
     } $
-    be a diagramm in $\cat{A}$.
+    be a diagram in $\cat{A}$.
     We have to show that there is a unique map $\bar{f}' \from P \to A$ shuch that
@@
     } \]
-    commutes. We know that the diagramm
+    commutes. We know that the diagram
     \[ \xymatrix{
@@
 
+\begin{answer}
+  \begin{itemize}
+    \item[``$\Rightarrow$'']
+      \emph{Counterexample:}
+      Let $\cat{A} := \Set, E := \emptyset, X := \nat, Y := \mathbb{B}$.
+      Define $f(n) := \text{even}(n)$ and $g(n) := \text{odd}(n)$. $i$ is the (empty) inclusion function.
+      Then
+      \[ \xymatrix{
+        E \ar[r]^i \ar[d]_i &
+        X \ar[d]^g \\
+        X \ar[r]_f &
+        Y
+      } \]
+      is clearly a pullback because no number is even and odd at the same time.
+      Now let $A := \nat$, $f_1(n) := n$ and $f_2(n) := n+1$.
+      Then the diagram
+      \[ \xymatrix{
+        A \ar[r]_{f_2} \ar[d]^{f_1} & Y \ar[d]_{g} \\
+        X \ar[r]^f & Z
+      } \]
+      commutes because $\text{even}(n) \Leftrightarrow \text{odd}(n+1)$.
+      But there exists no function $\bar{f} \from \nat \to \emptyset$ at all.
+    \item[``$\Leftarrow$'']
+      \begin{proof}
+        Let
+        \[ \xymatrix{
+          E \ar[r]^i \ar[d]_i &
+          X \ar[d]^g \\
+          X \ar[r]_f &
+          Y
+        } \]
+        be the pullback of the diagram
+        $ \xymatrix{
+          X \ar[r]^f & Y & X\ar[l]_g \\
+        } $.
+        We show that $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is the equalizer of $\xymatrix@1{X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$.
+        First we note that the diagramm $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is indeed a fork as it commutes by assumption.
+        Let
+        $ \xymatrix{
+          A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
+        } $
+        be a fork. Then
+        \[ \xymatrix{
+          A \ar[r]_{f} \ar[d]^{f} & X \ar[d]_{t} \\
+          X \ar[r]^s & Y
+        } \]
+        commutes. Therefore exists a unique map $\bar{f}\from A \to P$ such that
+        \[ \xymatrix{
+          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\bar{f}} & & \\
+          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
+          & X \ar[r]^s & Y
+        } \]
+        commutes. We show that this $\bar{f}$ is the only map such that
+        \[ \xymatrix{
+          A \ar[rd]^f \ar@{.>}[dd]|{\bar{f}} & & \\
+          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
+          P \ar[ru]_i & &
+        } \]
+        commutes. First it is clear that this diagramm commutes for $\bar{f}$. Now let $\tilde{f}$ be another map such that
+        \[ \xymatrix{
+          A \ar[rd]^f \ar@{.>}[dd]|{\tilde{f}} & & \\
+          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
+          P \ar[ru]_i & &
+        } \]
+        commutes. But then
+        \[ \xymatrix{
+          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\tilde{f}} & & \\
+          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
+          & X \ar[r]^s & Y
+        } \]
+        also commutes. By the assumption that $\bar{f}$ is unique we have $\bar{f} = \tilde{f}$. \qedhere
+      \end{proof}
+  \end{itemize}
+\end{answer}
+
 \begin{exercise}

### 2017-06-09 sheets/talk-6/sheet-6.tex
First part of sample solution 6.10

@@
         We show that $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is the equalizer of $\xymatrix@1{X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$.
-        First we note that the diagramm $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is indeed a fork as it commutes by assumption.
+        First we note that the diagram $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is indeed a fork as it commutes by assumption.
         Let
@@
         } \]
-        commutes. First it is clear that this diagramm commutes for $\bar{f}$. Now let $\tilde{f}$ be another map such that
+        commutes. First it is clear that this diagram commutes for $\bar{f}$. Now let $\tilde{f}$ be another map such that
         \[ \xymatrix{
@@
 
+\begin{answer}
+  \begin{proof}
+    We give names to the arrows and objects:
+    \[ \xymatrix{
+      A \ar[r]^h \ar[d]^j & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
+      D \ar[r]^m & E \ar[r]^n & F
+    } \]
+    Now let the pullback square
+    \[ \xymatrix{
+      B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
+      E \ar[r]^n & F
+    } \]
+    be the pullback of $\xymatrix{E \ar[r]^n & F & C \ar[l]_l}$.
+    We show both directions of the equivalence:
+    \begin{itemize}
+      \item[``$\Rightarrow$'']
+        Let the pullback square
+        \[ \xymatrix{
+          A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
+          D \ar[r]^{n \of m} & F
+        } \]
+        be the pullback of $\xymatrix{D \ar[r]^{n \of m} & F & C \ar[l]_l}$.
+        Let
+        \[ \xymatrix{
+          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[d]^k \\
+          D \ar[r]^{m} & E
+        } \]
+        be another pullback square. Now
+        \[ \xymatrix{
+          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[r]^{i} \ar[d]^k & C \ar[d]^l \\
+          D \ar[r]^{m} & E \ar[r]^n & F
+        } \]
+        commutes. Now because the outer square is a pullback there is a unique map $\bar{f} \from Z \to A$, such that
+        \[ \xymatrix{
+          Z \ar@/^/[drrr]^{i \of f_1}
+            \ar@{.>}[dr]|{\bar{f}}
+            \ar@/_/[ddr]_{f_2} & & \\
+            & A \ar[r]_h \ar[d]^j & B \ar[r]_{i} & C \ar[d]_l \\
+            & D \ar[r]^{m} & E \ar[r]^n & F
+        } \]
+        commutes. We show that this $\bar{f}$ is the unique map such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{f_1}
+            \ar@{.>}[dr]|{\bar{f}}
+            \ar@/_/[ddr]_{f_2} \\
+          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
+          & D \ar[r]^m & E
+        } \]
+        commutes. We have to show $j \of \bar{f} = f_2$ and $h \of \bar{f} = f_1$. The first equatation holds by assumption.
+
+        Now we define two maps
+        \begin{alignat*}{4}
+          g_1 &\from Z \to C&  \qquad g_1 &:= i \of f_1 \\
+          g_2 &\from Z \to E&         g_2 &:= m \of j \of \bar{f}
+        \end{alignat*}
+        and show that the diagram
+        \[ \xymatrix{
+          Z \ar[r]^{g_1} \ar[d]^{g_2} & C \ar[d]^l \\
+          E \ar[r]^n & F
+        } \]
+        commutes and is therefore a pullback square of the right-hand square:
+        \begin{alignat*}{2}
+          l \of g_1
+          &= l \of i \of f_1 \\
+          &= l \of i \of h \of \bar{f} \\
+          &= n \of m \of j \of \bar{f} \\
+          &= n \of g_2
+        \end{alignat*}
+        Now because the right-hand square is a pullback there is a unique map $\bar{g} \from Z \to B$, such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{g_1}
+            \ar@{.>}[dr]|{\bar{g}}
+            \ar@/_/[ddr]_{g_2} \\
+          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
+          & E \ar[r]^n & F
+        } \]
+        commutes.
+        By showing that this diagram comutes for $\bar{g} = f_1$ and for $\bar{g} = h \of \bar{f}$ we show $h \of \bar{f} = f_2$.
+        \begin{itemize}
+          \item[1.]We have $i \of f_1 = g_1$ by definition and $k \of f_1 = m \of f_2 = m \of j \of \bar{f} = g_2$.
+          \item[2.]We have $(i \of h) \of \bar{f} = i \of f_2$ and $(k \of h) \of \bar{f} = (m \of j) \of \bar{f}$.
+        \end{itemize}
+
+        We have shown that the diagram
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{f_1}
+            \ar@{.>}[dr]|{\bar{f}}
+            \ar@/_/[ddr]_{f_2} \\
+          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
+          & D \ar[r]^m & E
+        } \]
+        commutes for the given $\bar{f}$. We now have to prove that this $\bar{f}$ is unique.
+        We assume a map $\tilde{f} \from Z \to A$ such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{f_1}
+            \ar@{.>}[dr]|{\tilde{f}}
+            \ar@/_/[ddr]_{f_2} \\
+          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
+          & D \ar[r]^m & E
+        } \]
+        commutes. But than we have another map such that the diagram for the outer square commutes:
+        \[ \xymatrix{
+          Z \ar@/^/[drrr]^{i \of f_1}
+            \ar@{.>}[dr]|{\tilde{f}}
+            \ar@/_/[ddr]_{f_2} & & \\
+            & A \ar[r]_h \ar[d]^j & B \ar[r]_{i} & C \ar[d]_l \\
+            & D \ar[r]^{m} & E \ar[r]^n & F
+        } \]
+        We conclude that $\bar{f} = \tilde{f}$.
+      \item[``$\Leftarrow$'']
+    \end{itemize}
+  \end{proof}
+\end{answer}
+
 \end{document}

### 2017-06-09 sheets/talk-6/sheet-6.tex
Second part of sample solution 6.10

@@
         \begin{itemize}
-          \item[1.]We have $i \of f_1 = g_1$ by definition and $k \of f_1 = m \of f_2 = m \of j \of \bar{f} = g_2$.
-          \item[2.]We have $(i \of h) \of \bar{f} = i \of f_2$ and $(k \of h) \of \bar{f} = (m \of j) \of \bar{f}$.
+          \item[1.]$\bar{g} = f_1$: We have $i \of f_1 = g_1$ by definition and $k \of f_1 = m \of f_2 = m \of j \of \bar{f} = g_2$.
+          \item[2.]$\bar{g} = h \of \bar{f}$: We have $(i \of h) \of \bar{f} = i \of f_2$ and $(k \of h) \of \bar{f} = (m \of j) \of \bar{f}$.
         \end{itemize}
@@
         We conclude that $\bar{f} = \tilde{f}$.
+
       \item[``$\Leftarrow$'']
+        Let the pullback square
+        \[ \xymatrix{
+          A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
+          D \ar[r]^m & E
+        } \]
+        be the pullback of $\xymatrix{D \ar[r]^m & E & B \ar[l]_k}$.
+        Let
+        $ \xymatrix{
+          D & Z \ar[l]_{f_2} \ar[r]^{f_1} & C
+        } $
+        be a diagram. We note that
+        \[ \xymatrix{
+          Z \ar[r]^{f_1} \ar[d]^{f_2} & C \ar[d]^l \\
+          E \ar[r]_l & F
+        } \]
+        is a pullback square of the right-hand square.
+        Thus we have a unique map $\bar{f} \from Z \to B$ such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{f_1}
+            \ar@{.>}[dr]|{\bar{f}}
+            \ar@/_/[ddr]_{f_2} \\
+          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
+          & E \ar[r]^n & F
+        } \]
+        commutes. Then
+        \[ \xymatrix{
+          Z \ar[r]^{\bar{f}} \ar[d]^{f_2} & B \ar[d]^k \\
+          D \ar[r]^m && E
+        } \]
+        is a pullback square of the left-hand square.
+        Thus we have a unique map $\bar{g} \from Z \to A$ such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{\bar{f}}
+            \ar@{.>}[dr]|{\bar{g}}
+            \ar@/_/[ddr]_{f_2} \\
+          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
+          & D \ar[r]^m & E
+        } \]
+        commutes. We show that this $\bar{g}$ is the unique map such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{f_1}
+            \ar@{.>}[dr]|{\bar{g}}
+            \ar@/_/[ddr]_{f_2} \\
+          & A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
+          & D \ar[r]^{n \of m} & F
+        } \]
+        commutes.
+        The diagram comutes:
+        \begin{align*}
+          i \of h \of \bar{g} = i \of \bar{f} = f_1 \\
+          j \of \bar{g} = f_2
+        \end{align*}
+        Let $\tilde{g}$ be another map such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{f_1}
+            \ar@{.>}[dr]|{\tilde{g}}
+            \ar@/_/[ddr]_{f_2} \\
+          & A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
+          & D \ar[r]^{n \of m} & F
+        } \]
+        commutes. To show $\bar{g} = \tilde{g}$ it is because of the uniqueness enough to show that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{\bar{f}}
+            \ar@{.>}[dr]|{\tilde{g}}
+            \ar@/_/[ddr]_{f_2} \\
+          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
+          & D \ar[r]^m & E
+        } \]
+        commutes.
+        We already have $j \of \tilde{g} = f_2$ and only have to show $h \of \tilde{g} = \bar{f}$.
+        Because of $\bar{f} = h \of \bar{g}$ it suffices to show $h \of \tilde{g} = h \of \bar{g}$.
+        The commuting square
+        \[ \xymatrix{
+          Z \ar[r]^{i \of h \of \tilde{g}} \ar[d]_{m \of j \of \tilde{g}} & C \ar[d]^f \\
+          E \ar[r]^n & F
+        } \]
+        is another pullback square for the right-hand square.
+        Thus we have a unique map $\bar{x} \from Z \to B$ such that
+        \[ \xymatrix{
+          Z \ar@/^/[drr]^{i \of h \of \tilde{g}}
+            \ar@{.>}[dr]|{\bar{x}}
+            \ar@/_/[ddr]_{m \of j \of \tilde{g}} \\
+          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
+          & E \ar[r]^n & F
+        } \]
+        commutes.
+        We get $h \of \tilde{g} = h \of \bar{g}$ when we show that this diagram commutes for $\bar{x} = h \of \tilde{g}$ and $\bar{x} = h \of \bar{g}$.
+
+        \begin{itemize}
+          \item[1.]$\bar{x} = h \of \tilde{g}$:
+            We have $i \of h \of \tilde{g} = i \of h \of \tilde{g}$ and $m \of j \of \tilde{g} = k \of h \of \tilde{g}$.
+          \item[2.]$\bar{x} = h \of \bar{g}$:
+            We have $i \of h \of h = i \of f_1 = i \of h \of \tilde{g}$. Furthermore we have
+            \begin{alignat*}{3}
+              & &\quad&  m \of j \of \tilde{g} \\
+              &=&& m \of f_2 \of \tilde{g} \\
+              &=&& m \of j \of \bar{g} \\
+              &=&& k \of h \of \bar{g}.
+            \end{alignat*}
+        \end{itemize}
+        Thus $h \of \tilde{g} = h \of \bar{g}$. Hence the outer square is a pullback.\qedhere
     \end{itemize}

### 2017-06-09 sheets/talk-6/sheet-6.tex
Sample solution 6.6

@@
 \end{exercise}
+
+\begin{answer}
+  \begin{proof}
+    We numerate the statements:
+    \begin{itemize}
+      \item  [I:]$s=t$
+      \item [II:]The qualizer of $s$ and $t$ exists and is an isomorphism.
+      \item[III:]The coequalizer of $s$ and $t$ exists and is an isomorphism.
+    \end{itemize}
+    We show that all statements are equivalent by showing $I \Leftrightarrow II$ and $I \Leftrightarrow III$.
+    Note that $I \Leftrightarrow III$ is the dualized result of $I \Leftrightarrow II$, so we only need to show $I \Leftrightarrow II$.
+
+    \begin{itemize}
+      \item[``$\Rightarrow$'']
+        Let $s=t$. Then the fork
+        $\xymatrix{
+          X \ar[r]^{1_X} & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
+        }$
+        $1_X$ is an isomorphism.
+        Let
+        $\xymatrix{
+          A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
+        }$
+        be a fork.
+        Clearly $\bar{f} := f$ is the only map $\bar{f} \from A \to X$ such that the diagram
+        \[ \xymatrix{
+          A \ar[rd]^f \ar@{.>}[dd]|{\bar{f}} & & \\
+          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
+          X \ar[ru]_{1_X} & &
+        } \]
+        commutes.
+      \item[``$\Leftarrow$'']
+        Let
+        $\xymatrix{
+          E \ar[r]^i & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
+        }$
+        be the equalizer and $i$ a isomorphism. Then by commutativity we get:
+        \begin{alignat*}{4}
+          && s \of i &= t \of i \\
+          \Rightarrow  &\qquad& s \of i \of i^{-1} &= t \of i \of i^{-1} \\
+          \Rightarrow  &      & s &= t
+        \end{alignat*}\qedhere
+    \end{itemize}
+  \end{proof}
+\end{answer}
 

### 2017-06-09 sheets/talk-6/sheet-6.tex
Fix typo in sample solution 6.6

@@
       \item  [I:]$s=t$
-      \item [II:]The qualizer of $s$ and $t$ exists and is an isomorphism.
+      \item [II:]The equalizer of $s$ and $t$ exists and is an isomorphism.
       \item[III:]The coequalizer of $s$ and $t$ exists and is an isomorphism.
@@
 \end{exercise}
+
+\begin{answer}
+  The coequalizer follows the scheme like in the talk.
+  We define the relation $R := \set{ (1_X (x), f(x) \;\big|\; x \in X} = \set{ (x, f(x) \;\big|\; x \in X} = f$ defined as a set.
+  Now $\sim$ is the reflexive symmetric transitive closure of $R$.
+  We define $i(x) := [x] := \set{ x' \in X | x \sim x'}$
+  and the quotient $C := X / \sim \;=\; i(X)$.
+  Now
+  $ \xymatrix{
+    X \ar@<.5ex>[r]^f \ar@<-.5ex>[r]_{1_X} & X \ar[r]^i & C
+  } $
+  is the coequalizer of $\parpairi{X}{X}{f}{1}$.
+  The equivalence $\sim$ has the property:
+  $\forall x_1, x_2 \in X, (x_1, x_2) \in \sim \;\Leftrightarrow\; f(x_1) = f(x_2).$
+  Thus we have $[x] = \set{x' \in X \;\big|\; f(x) = f(x')}$.
+
+\end{answer}
 

### 2017-06-09 sheets/talk-6/sheet-6.tex
Sheet 6: Fix typos

@@
     commutes.
-    Now we can instanciate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instanciations we get the commuting diagram
+    Now we can instantiate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instantiations we get the commuting diagram
     \[ \xymatrix{
@@
     This is trivial since we have $\forall i \in I: z \le x_i$.
-    Now let $a \in A$ be and objects and let $\Bigl(a \toby{f_i} x_i\Bigr)_{i \in I}$ be a family of maps from $a$.
+    Now let $a \in A$ be an object and let $\Bigl(a \toby{f_i} x_i\Bigr)_{i \in I}$ be a family of maps from $a$.
     The existing of this maps means that $a$ is a upper bound of $(x_i)_{i\in I}$.
@@
 \begin{answer}
-  First let define the arbitrary disjoint union:
+  First we define the arbitrary disjoint union:
   \begin{definition}{Arbitrary disjoint union}
@@
     We define the arbitrary disjoint union of $(x_i)_{i\in I}$:
-    $$\coprod_{i \in I} {x_i} := \bigcup_{i \in I} \set{ (x, i) \;\big|\; i \in I \land x \in x_i}$$
+    $$\coprod_{i \in I} {x_i} := \bigcup_{i \in I} \set{ (x, i) \;\big|\; x \in x_i}$$
   \end{definition}
@@
     Let $i \in I$ and we check $a \in x_i$. $\bar{f}(p_i(a)) = \bar{f}(a,i) = f_i(a)$.
-    Now let $\tilde{f} : \coprod_{i \in I} {x_i} \to A$ be another function sucht that
+    Now let $\tilde{f} : \coprod_{i \in I} {x_i} \to A$ be another function such that
     $\forall i \in I: \tilde{f} \of p_i = f_i$.
@@
     \[
-      \tilde{f}(a) = \bigl(p_1(\tilde{f}(a)), p_2(\tilde{f}(a))\bigr) = \bigl(p_1(\bar{f}(a)), p_2(\bar{f}(a))\bigr) = \bar{f}(a)\qedhere
+      \tilde{f}(a) = \bigl(p_1(\tilde{f}(a)), p_2(\tilde{f}(a))\bigr) = \bigl(f_1(a), f_2(A)\bigr) = \bigl(p_1(\bar{f}(a)), p_2(\bar{f}(a))\bigr) = \bar{f}(a)\qedhere
     \]
@@
     be a diagram in $\cat{A}$.
-    We have to show that there is a unique map $\bar{f}' \from P \to A$ shuch that
+    We have to show that there is a unique map $\bar{f}' \from P \to A$ such that
     \[ \xymatrix{
@@
       Then
-      \[ \xymatrix{
-        E \ar[r]^i \ar[d]_i &
-        X \ar[d]^g \\
-        X \ar[r]_f &
-        Y
-      } \]
-      is clearly a pullback because no number is even and odd at the same time.
+      $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f}\ar@<-.5ex>[r]_g &Y}$
+      is clearly a equalizer because no number is even and odd at the same time.
       Now let $A := \nat$, $f_1(n) := n$ and $f_2(n) := n+1$.
@@
         } \]
-        commutes. Now because the outer square is a pullback there is a unique map $\bar{f} \from Z \to A$, such that
+        commutes. Now because the outer rectangle is a pullback there is a unique map $\bar{f} \from Z \to A$, such that
         \[ \xymatrix{
@@
         } \]
-        commutes. But than we have another map such that the diagram for the outer square commutes:
+        commutes. But than we have another map such that the diagram for the outer rectangle commutes:
         \[ \xymatrix{
@@
         } $
-        be a diagram. We note that
+        be a diagram such that
         \[ \xymatrix{
-          Z \ar[r]^{f_1} \ar[d]^{f_2} & C \ar[d]^l \\
-          E \ar[r]_l & F
+          Z \ar[r]^{f_1} \ar[d]_{f_2} & C \ar[d]^l \\
+          D \ar[r]^{n\of m} & F
         } \]
-        is a pullback square of the right-hand square.
+        commutes. We note that
+        \[ \xymatrix{
+          Z \ar[r]^{f_1} \ar[d]_{m \of f_2} & C \ar[d]^l \\
+          E \ar[r]_n & F
+        } \]
+        is thus a pullback square of the right-hand square.
         Thus we have a unique map $\bar{f} \from Z \to B$ such that
@@
             \ar@{.>}[dr]|{\bar{f}}
-            \ar@/_/[ddr]_{f_2} \\
+            \ar@/_/[ddr]_{m \of f_2} \\
           & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
@@
         commutes.
-        The diagram comutes:
+        The diagram comutes for $\bar{g}$:
         \begin{align*}
@@
         \[ \xymatrix{
-          Z \ar[r]^{i \of h \of \tilde{g}} \ar[d]_{m \of j \of \tilde{g}} & C \ar[d]^f \\
+          Z \ar[r]^{i \of h \of \tilde{g}} \ar[d]_{m \of j \of \tilde{g}} & C \ar[d]^l \\
           E \ar[r]^n & F
@@
           \item[2.]$\bar{x} = h \of \bar{g}$:
-            We have $i \of h \of h = i \of f_1 = i \of h \of \tilde{g}$. Furthermore we have
+            We have $i \of h \of \bar{g} = i \of f_1 = i \of h \of \tilde{g}$. Furthermore we have
             \begin{alignat*}{3}
@@
         \end{itemize}
-        Thus $h \of \tilde{g} = h \of \bar{g}$. Hence the outer square is a pullback.\qedhere
+        Thus $h \of \tilde{g} = h \of \bar{g}$. Hence the outer rectangle is a pullback.\qedhere
     \end{itemize}

### 2017-06-09 sheets/talk-6/sheet-6.tex
Sheet 6: More minor edits

@@
     Now let $a \in A$ be an object and let $\Bigl(a \toby{f_i} x_i\Bigr)_{i \in I}$ be a family of maps from $a$.
-    The existing of this maps means that $a$ is a upper bound of $(x_i)_{i\in I}$.
-    Now by definition of least upper bound, $a \le z$, thus we have a arrow $\bar{f} \from a \to z$.
-    All arrows are unique, thus $\bar{f}$ is unique and the diagram commutes. \qedhere
+    The existence of this maps means that $a$ is an upper bound of $(x_i)_{i\in I}$.
+    Now by definition of least upper bound, $a \le z$, thus we have an arrow $\bar{f} \from a \to z$.
+    All arrows are unique, thus $\bar{f}$ is the only map making the diagram commute. \qedhere
   \end{proof}
@@
     \end{alignat*}
+    Clearly $p_1$ and $p_2$ are linear functions.
     Now let $A$ be a vector space and $f_1 \from A \to V, f_2 \from A \to W$ be linear functions.
-    We define $\bar{f}(a) := (f_1(a), f_(2))$. We clearly have
+    We define $\bar{f}(a) := (f_1(a), f_2(a))$. We clearly have
     \begin{alignat*}{2}
@@
     \[
-      \tilde{f}(a) = \bigl(p_1(\tilde{f}(a)), p_2(\tilde{f}(a))\bigr) = \bigl(f_1(a), f_2(A)\bigr) = \bigl(p_1(\bar{f}(a)), p_2(\bar{f}(a))\bigr) = \bar{f}(a)\qedhere
+      \tilde{f}(a) = \bigl(p_1(\tilde{f}(a)), p_2(\tilde{f}(a))\bigr) = \bigl(f_1(a), f_2(A)\bigr) = \bar{f}(a)\qedhere
     \]
@@
   \begin{proof}
-    Let $i \from core(s-t) \to X$ be a inclusion function. Let
+    Let $i \from core(s-t) \to X$ be the inclusion function. Let
     $\xymatrix{
@@
       \item[``$\Rightarrow$'']
-        Let $s=t$. Then the fork
+        Let $s=t$. $1_X$ is an isomorphism and we show that the fork
         $\xymatrix{
@@
         }$
-        $1_X$ is an isomorphism.
-        Let
+        is the equalizer. Let
         $\xymatrix{
@@
   \begin{proof}
-    We show that $S := P$ is the coproduct of $X$ and $Y$ together with the projection functions $p_1$ and $p_2$.
-    First note that because $Z$ is initial $s$ and $t$ are unique maps.
+    We show that $S := P$ is the coproduct of $X$ and $Y$ together with the coprojection functions $p_1$ and $p_2$.
     Let
@@
     } \]
-    would also commute (because $S$ is initial) but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
+    would also commute (because $Z$ is initial) but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
   \end{proof}
@@
       $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f}\ar@<-.5ex>[r]_g &Y}$
-      is clearly a equalizer because no number is even and odd at the same time.
+      is clearly the equalizer because no number is even and odd at the same time.
       Now let $A := \nat$, $f_1(n) := n$ and $f_2(n) := n+1$.
@@
       commutes because $\text{even}(n) \Leftrightarrow \text{odd}(n+1)$.
-      But there exists no function $\bar{f} \from \nat \to \emptyset$ at all.
+      But there exists no function $\bar{f} \from \nat \to \emptyset = E$ at all.
     \item[``$\Leftarrow$'']
@@
         } \]
-        commutes. Therefore exists a unique map $\bar{f}\from A \to P$ such that
+        commutes. Therefore there exists a unique map $\bar{f}\from A \to P$ such that
         \[ \xymatrix{
@@
         } \]
-        be another pullback square. Now
+        be another pullback square for the left-hand square. Now
         \[ \xymatrix{

### 2017-06-14 sheets/talk-7/sheet-7.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\usepackage{upgreek}

\begin{document}

\uebunghead{Talk 7}{General Limits}
\author{Leonhard Staut}

\section{Limits}

\begin{exercise}[Limit definition]
  \begin{enumerate}
  \item Verify that terminal objects are a valid instance of the general limit definition.
  \item Verify that equalizers are a valid instance of the general limit definition.
  \end{enumerate}
\end{exercise}

\begin{exercise}[Pullback monic]
  Let $\mathscr{C}$ be any category and $X, Y$ objects.
  Show that a map $X \overset{f}{\to} Y$ is monic iff the following square is a pullback.
  \[
    \begin{tikzcd}
      X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
      X \arrow{r}{f}                & Y                  
    \end{tikzcd}
  \]
\end{exercise}

\begin{exercise}[Contained limit]
  Let $D$ be a diagram.
  Find an example of a limit $L$ that is completely contained in $D$,
  meaning the vertex of $L$ is in the image of $D$.
\end{exercise}

\begin{exercise}[Cone category]
  Let $\mathscr{C}$ be any category. Let \texttt{Cone} be the category of cones
  in $\mathscr{C}$ over a fixed diagram.
  What are the limits of $\mathscr{C}$ in \texttt{Cone}? What are the colimits?
\end{exercise}

\begin{definition}
  {Constant Functor}
  For any two categories $\mathscr{C}$ and $\mathscr{D}$,
  the constant functor $\Updelta _ V : \mathscr{C} \to \mathscr{D} $ for some object $V \in \mathscr{D}$
  is defined by:
  \begin{itemize}
    \item $\forall A \in \mathscr{C}. \ \Updelta _ V (A) = V$
    \item $\forall A \overset{f}{\to} B \in \mathscr{C}. \ \Updelta _ V (f) = 1_V$
  \end{itemize}

\end{definition}

\begin{exercise}[Cones as natural transformations]
  Let $\mathscr{C}$ and $\bf{I}$ be any categories, $D : \bf{I} \to \mathscr{C}$ a diagram.
  Verify that a natural transformation $\pi : \Updelta_v \to D(\bf{I})$ is a cone over $D$ with vertex $V$.
\end{exercise}

\begin{exercise}[Uniqueness of limits]
  Let $\mathscr{C}$ be any category. Show that limits in $\mathscr{C}$ are unique up to isomorphism.
\end{exercise}

\begin{exercise}[Limit construction]
  Recall the construction:
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K)
    \end{tikzcd}
  \]
  where $E$ is the equalizer of $s$ and $t$.
  We have shown that $E$ is a cone for any diagram $D$. Show that $E$ is also the limit,
  i.e. a universal cone over $D$.
\end{exercise}
\begin{hint}
  \begin{itemize}
  \item You may use the fact that $E$ is a cone iff $E$ is an equalizer of $s$ and $t$.
  \item You do not need the concrete definitions of $s$ and $t$.
  \end{itemize}
\end{hint}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-06-14 sheets/talk-7/sheet-7.tex

@@
 \author{Leonhard Staut}
-
-\section{Limits}
 
@@
   \]
-\end{exercise}
-
-\begin{exercise}[Contained limit]
-  Let $D$ be a diagram.
-  Find an example of a limit $L$ that is completely contained in $D$,
-  meaning the vertex of $L$ is in the image of $D$.
 \end{exercise}
@@
   Let $\mathscr{C}$ and $\bf{I}$ be any categories, $D : \bf{I} \to \mathscr{C}$ a diagram.
-  Verify that a natural transformation $\pi : \Updelta_v \to D(\bf{I})$ is a cone over $D$ with vertex $V$.
+  Verify that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ is a cone over $D$ with vertex $V$.
 \end{exercise}

### 2017-06-14 sheets/talk-7/sheet-7.tex
improve exercise about cone category

@@
 
+\begin{definition}
+  {Cone Category}
+  Let $\mathscr{C}$ be any category, $D : \bf{I} \to \mathscr{C}$ a diagram.
+  \texttt{Cone(D)} is the category of cones in $\mathscr{C}$ over $D$.
+  \begin{itemize}
+  \item The objects of \texttt{Cone(D)} are cones in $\mathscr{C}$ over $D$, that is, if
+    $(U \overset{f_I}{\to} D(I))_{I \in \bf{I}}$ is a cone in $\mathscr{C}$, then there is a corresponding object
+    in \texttt{Cone(D)}.
+  \item The morphisms of \texttt{Cone(D)} are morphisms between two cones in $\mathscr{C}$, that is,
+    if $(U \overset{f_I}{\to} D(I))_{I \in \bf{I}}, (V \overset{g_I}{\to}~ D(I))_{I \in \bf{I}}$
+    are two cones in $\mathscr{C}$ and there is a morphism $V \overset{m}{\to} U$, then there
+    is a corresponding morphism in \texttt{Cone(D)}.
+  \end{itemize}
+\end{definition}
+
 \begin{exercise}[Cone category]
-  Let $\mathscr{C}$ be any category. Let \texttt{Cone} be the category of cones
-  in $\mathscr{C}$ over a fixed diagram.
-  What are the limits of $\mathscr{C}$ in \texttt{Cone}? What are the colimits?
+  Consider an object $L \in$ \texttt{Cone(D)} that corresponds to a limit in $\mathscr{C}$ over $D$.
+  What kind of object is $L$?\\
+  What about an object $L' \in$ \texttt{Cone(D)} that corresponds to a colimit in $\mathscr{C}$ over $D$?
 \end{exercise}
@@
 \end{exercise}
+
+\newpage
 

### 2017-06-21 sheets/talk-7/sheet-7.tex

@@
 \end{exercise}
+\begin{answer}
+  Recall that general limits are universal cones over a diagram.
+  Therefore, instances of general limits are defined by the shape of their diagram.
+  In the following let $\mathscr{C}$ be any category.
+  \begin{enumerate}
+  \item
+    We choose the empty category as the shape $\bf{I}$.
+    Let $D : \bf{I} \to \mathscr{C}$ a diagram. A cone over $D$ is an object $U$ with
+    maps to every object in the diagram. Since the diagram is empty, the cone is just
+    the object $U$.
+    $U$ is universal if for any other cone $V$ in $\mathscr{C}$, there is a unique map
+    $h : V \to U$.
+    Any other cone $V$ is also just an object in $\mathscr{C}$.
+    Therefore $U$ is a terminal object.
+  \item
+    As the shape $\bf{I}$ we choose the following diagram.
+    \[
+      \begin{tikzcd}
+        I \arrow[r,shift left, "m"] \arrow[r,shift right,swap,"n"] & J
+      \end{tikzcd}
+    \]
+    Let $D : \bf{I} \to \mathscr{C}$ a diagram.
+    A cone over $D$ is an object $U$ with maps to every object in the diagram, such that all triangles commute.
+    \[
+      \begin{tikzcd}
+        U \arrow{d}{f_I} \arrow{dr}{f_J} & \\
+        D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
+      \end{tikzcd}
+    \]
+    Specifically, the commuting condition says that $Dm \circ f_I = f_J = Dn \circ f_I$.
+    Since $f_J$ is already completely determined by $f_I$, we can ignore the arrow
+    and state equivalently that the following diagram commutes.
+    \[
+      \begin{tikzcd}
+        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
+      \end{tikzcd}
+    \]
+    The arrow $f_J$ is simply left implicit.\\
+    $U$ is a universal cone, i.e. a limit, if for any other cone $V$ in $\mathscr{C}$, there is a unique map
+    $h : V \to U$.
+    \[
+      \begin{tikzcd}
+        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J) \\
+        V \arrow{ur}{g_I} \arrow[dashed]{u}{h}& &
+      \end{tikzcd}
+    \]
+    This is exactly the definition of an equalizer.
+  \end{enumerate}
+\end{answer}
 
@@
 \end{exercise}
+
+\begin{answer}
+  $"\Leftarrow":$ \\
+  Let $g_1, g_2 : Z \to X$ be two maps. If $f \circ g_1 = f \circ g_2$
+  we have the following commuting diagram.
+  \[
+    \begin{tikzcd}
+      Z \arrow[bend right=30]{ddr}{g_1} \arrow[bend left=30]{drr}{g_2}&   &\\
+             & X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
+             & X \arrow{r}{f}                & Y
+    \end{tikzcd}
+  \]
+  Therefore $Z$ is a cone. Since is $X$ is a pullback we have a unique map $h : Z \to X$,
+  such that the following diagram commutes.
+  \[
+    \begin{tikzcd}
+      Z \arrow[bend right=30]{ddr}{g_1} \arrow[bend left=30]{drr}{g_2} \arrow[dashed]{dr}{h}&   &\\
+             & X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
+             & X \arrow{r}{f}                & Y
+    \end{tikzcd}
+  \]
+  It follows that $1 \circ h = h = g_1 = g_2$.
+  Because there is only one map $Z \to X$, $f$ is trivially monic.\\
+  $"\Rightarrow":$ \\
+  The following square certainly commutes because $f \circ 1 = f \circ 1$.
+  \[
+    \begin{tikzcd}
+      X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
+      X \arrow{r}{f}                & Y
+    \end{tikzcd}
+  \]
+  Therefore it is a cone and we only need to show its universality.
+  We have to show that, assuming f is monic, for any other cone $Z$ there is a unique map
+  $h : Z \to X$ making the following diagram commute.
+  \[
+    \begin{tikzcd}
+      Z \arrow[bend right=30]{ddr}{g_1} \arrow[bend left=30]{drr}{g_2} \arrow[dashed]{dr}{h}&   &\\
+             & X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
+             & X \arrow{r}{f}                & Y
+    \end{tikzcd}
+  \]
+  We know the outer diagram commutes because $Z$ is a cone, so $f \circ g_1 = f \circ g_2$
+  Therefore we have $g_1 = g_2$ because $f$ is monic.
+  Since all the maps $Z \to X$
+  are the same there is exactly one map which is unique.
+\end{answer}
 
@@
 \end{exercise}
+\begin{answer}
+  If $L$ corresponds to a limit, then by definition any other cone with vertex $V$ in $\mathscr{C}$ has a unique map
+  from $V$ to the limit that corresponds to a map in \texttt{Cone(D)}.
+  Because the cones in $\mathscr{C}$ are exactly the objects in \texttt{Cone(D)},
+  $L$ has a unique arrow from every object \texttt{Cone(D)}.
+  Therefore $L$ is a terminal object.\\
+  By duality $L'$ is an initial object.
+\end{answer}
 
@@
 \end{exercise}
+\begin{answer}
+  Recall that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ has components for every object $I \in \bf{I}$
+  and for every morphism $m : I \to J \in \bf{I}$ the following square commutes.
+  \[
+    \begin{tikzcd}
+      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
+      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
+    \end{tikzcd}
+  \]
+  Now we unfold the definition of $\Updelta_V$.
+  \[
+    \begin{tikzcd}
+      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
+      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
+    \end{tikzcd}
+    \ = \
+    \begin{tikzcd}
+      V \arrow{r}{\pi_I}  \arrow{d}{1_V}  & D(I) \arrow{d}{D(m)}    \\
+      V \arrow{r}{\pi_J}                & D(J)
+    \end{tikzcd}
+    \ = \
+    \begin{tikzcd}
+      V \arrow{r}{\pi_I} \arrow{dr}{\pi_J} & D(I) \arrow{d}{D(m)}    \\
+                     & D(J)
+    \end{tikzcd}
+  \]
+  To summarize, the natural transformation $\pi$ has maps $\pi_I$ for all $I \in \bf{I}$,
+  and the above triangle commutes for all maps $m : I \to J$.
+  This is exactly the definition of a cone with vertex $V$.
+  In fact this yields an alternative equivalent definition of cones.
+\end{answer}
 
@@
 \end{exercise}
-
-\newpage
+\begin{answer}
+  Let $U, V$ be two limits. Then there are unique morphisms $f : U \to V$ and $g : V \to U$.
+  Consider the morphism $g \circ f : U \to U$. As there is a unique morphism $U \to U$, it has to be
+  the identity $g \circ f = 1_U$.
+  Similarly $f \circ g = 1_V$, so $U$ and $V$ are isomorphic.
+\end{answer}
 
@@
 \end{hint}
+\begin{answer}
+  To show that $E$ is a universal cone, we need to show that for any other cone $F$, there is a unique 
+  morphism $h : F \to E$.
+  If $F$ is a cone, then it is also an equalizer of $s$ and $t$.
+  \[
+    \begin{tikzcd}
+      E \arrow{r}{e} &
+      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
+      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
+      F \arrow{ur}{f} & &
+    \end{tikzcd}
+  \]
+  Since $E$ is an equalizer, and $F$ is a cone over the same diagram, we have a unique map $h : F \to E$.
+  \[
+    \begin{tikzcd}
+      E \arrow{r}{e} &
+      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
+      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
+      F \arrow{ur}{f} \arrow[dashed]{u}{h} & &
+    \end{tikzcd}
+  \]
+
+\end{answer}
 

### 2017-06-22 sheets/talk8/talk8.tex

\def\pathToRoot{../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Talk 8}{Cartesian Closed Categories}

\begin{hint}
  Read Awodey Chapter 6 and Leinster Chapter 6.3. 
\end{hint}

%\section{Diagrams}
%
%\begin{hint}
%  This is how you can draw diagrams:
%  \[
%    \begin{tikzcd}
%                                    & A \arrow{dr}{g}    & \\
%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
%    \end{tikzcd}
%  \]
%\end{hint}

\begin{exercise}
  Show that the following identites hold for any cartesian closed category $\mathcal{A}$ with X,Y,Z $\in \ ob(\mathcal{A})$ and terminal object 1:
  \textit{(Hint: it suffices to show, that both sides satisfy the same universal property.)}
  \begin{itemize}
  \item $1^X \cong 1$
  \item $X^1 \cong X$
  \item $(X^Y)^Z \cong X^{Y\times Z}$
  \end{itemize}
\end{exercise}

\begin{definition}{Bicartesian Closed Category}
A category $\mathcal{A}$ is called \emph{bicartesian closed} if it is cartesian closed and has in addition an initial object 0 and binary coproducts A + B, $\forall A,B \in \mathcal{A}$.
\end{definition}

\begin{exercise}
  Assume a bicartesian closed category $\mathcal{A}$. Show $\forall X,Y,Z \in ob(\mathcal{A})$:
  \begin{itemize}
  \item $(X+Y) \times Z \cong (X \times Z) + (Y \times Z)$
  \item $X^0 \cong 1$
  \item $X^{Y+Z} \cong X^Y \times X^Z$
  \end{itemize}
\end{exercise}

\begin{definition}{Pointed Category}
  A pointed category $\mathcal{A}$ is one, that has a terminal object 1 and an initial object 0 and 1 = 0.
\end{definition}

\begin{exercise}
  Show that a pointed category $\mathcal{A}$ is cartesian closed if and only it A is trivial, i.e. all its objects are isomorphic.
\end{exercise}

\begin{exercise}
  Is the category of monoids cartesian closed?
\end{exercise}

\begin{exercise}
  Is the category of relations cartesian closed? What are its products?
\end{exercise}

\begin{definition}{Heyting Algebra}
  A Heyting Algebra $\mathcal{H}$ is a partially ordered set with
  \begin{itemize}
  \item Finite meets: 1 and $p\land q$
  \item Finite joins: 0 and $p \lor q$
  \item Exponentials: for each a,b an element $a \Rightarrow b$ such that $a \land b \leq c$ iff $a \leq (b \rightarrow c)$.
  \end{itemize}
\end{definition}

\begin{exercise}
  Show that every Heyting Algebra, seen as a category is cartesian closed.  
\end{exercise}

%\begin{definition}{Downsets}
%  A subset M of A is called a downset of A, if $\forall \ x,y \in A, x \leq y \rightarrow y \in M \rightarrow x \in M $.
%\end{definition}

\begin{exercise}
  Let A be a set with a preoder $\le$. Show that the downsets of A $P^{\downarrow}(A) := \{M \subseteq A \mid \forall \ x,y \in A, x \leq y \rightarrow y \in M \rightarrow x \in M \}$ ordered by inclusion form a Heyting Algebra.
\end{exercise}

\begin{exercise}
  Let $\mathcal{A}$ be a (small) category. Show that the functor category $\textbf{Set}^{A^{op}}$, that is the category of presheaves on A, is cartesian closed. 
\end{exercise}

\begin{definition}{G-Sets}
  A G-set for a group \textbf{G} is a Set \textbf{A} with a G-action g: $A \rightarrow G \rightarrow A \ , (x,a) \mapsto x\cdot a$ such that $\forall x \in A, a,b \in G \ x\cdot 1=x, x\cdot a \cdot b = x\cdot (ab)$.
  The \emph{category} of G-sets for a fixed group \textbf{G} has G-sets as objects and G-equivariant functions as morphisms. $ Hom(A,B) = \{ f:A \rightarrow B \mid \forall (a \in A)(x \in G), f(a\cdot x) = f(a)\cdot x\}. $
\end{definition}

\begin{exercise}
  Show that the category of G-sets is cartesian closed.
\end{exercise}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-07-02 sheets/talk-10/sheet-10.tex
First version sheet and slides.

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Talk 10}{Presheaves, Representables and the Yoneda Lemma}
\author{Dominik Wagner}

\begin{hint}
  Read chapter 4 in Leinster and chapter 8 in Awodey.
\end{hint}

\begin{definition}{Right $M$-Sets}
  Let $M$ be a monoid. 
  \begin{itemize}
  \item A \emph{right $M$-set} is a set $S$ together with a function
    \begin{align*}
      \cdot\from S\times M&\to S\\
      (s,m)&\mapsto s\cdot m
    \end{align*}
    such that $s\cdot (mm')=(s\cdot m)\cdot m'$ and $s\cdot 1_M=s$ for all $m,m'\in M$ and $s\in S$.
    \item A homomorphism between two right $M$-sets $S_1$ and $S_2$ is a mapping $\alpha\from S_1\to S_2$ satisfying $\alpha(s\cdot_{S_1} m)=\alpha(s)\cdot_{S_2}m$.
    \item The \emph{right regular representation} $\underline{M}$ is the $M$-set $M$ together with the right-operation $s\cdot m=sm$, i.e. the underlying monoid-operation.
  \end{itemize}

\end{definition}

\begin{exercise}[Not difficult despite its length]
  In this exercise we prove the Yoneda Lemma for the special case of one-object categories, i.e. monoids regarded as categories.
  We exploit the correspondence between presheaves on one-object categories (monoids $M$) and right $M$-sets.

  \begin{enumerate}
  \item Make yourself familiar again with this correspondence (e.g. by reading the examples 1.2.14 and 1.2.8 in Leinster).
  \item Let $M$ be a monoid. Show that the $M$-set corresponding to the unique(!) representable functor $M^\op\to \Set$ is the right regular representation $\underline M$.
  \item Now let $S$ be any right $M$-set. Show that for each $s\in S$ there is a unique homomorphism $\alpha\from\underline M\to S$ of $M$-sets such that $\alpha(1)=s$. Deduce that 
    \begin{align*}
      \{\text{homomorphisms }\underline M\to S\}\iso S
    \end{align*}
  \item Deduce that
    \begin{align*}
      [M^\op,\Set](H_A, F)\iso F(A)
    \end{align*}
    holds for all presheaves $F\from M^\op\to\Set$.
  \end{enumerate}
\end{exercise}

\begin{answer}
  \begin{enumerate}
  \item Done!
  \item Let $A$ be the unique object corresponding to $M$. In what follows we identify elements $m$ of the monoid $M$ with arrows $A\xrightarrow{m}A$. Then, $H_A(A)=\{\text{maps } A\to A\}= M$. Furthermore, for all $m,m'\in M$, $m\cdot m'=H_A(m')(m)=m\of m'=mm'$. Hence, $H_A$ contains exactly the same information as $\underline M$.
  \item Let $S$ be a right $M$-set. Assume there is a homomorphism $\alpha\from \underline M\to S$. Then for each $s\in S$, 
    \begin{align*}
      \label{eq:defa}
      \alpha(s)=\alpha(1\cdot_M s)=\alpha(1)\cdot_S s.
    \end{align*}
    Thus, $\alpha$ is completely defined by its action on $1$ if it exists. At the same time, this equation completely defines a homomorphism since
    \begin{align*}
      \alpha(s)\cdot m=(\alpha(1)\cdot s)\cdot m=\alpha(1)\cdot(s\cdot m)=\alpha(s\cdot m)
    \end{align*}
    as required.
    Therefore, the function
    \begin{align*}
      \{\text{homomorphisms }\underline M\to S\}&\to S\\
      \alpha&\mapsto \alpha(1)
    \end{align*}
    is an isomorphism.
  \item To conclude the Yoneda lemma we still have to show that natural transformations $F_1\to F_2$ are the same as homomorphisms between $S_1$ and $S_2$ where $S_1,S_2$ are the $M$-sets corresponding to the presheaves $F_1$ and $F_2$.

    To see this let us instantiate the definition of naturality for our special case. It states that for each arrow $A\xrightarrow{m}A$ the following diagram commutes:

    \begin{align*}
      \xymatrix{
      F_1(A)=S_1 \ar[r]^{F_1(m)=-\cdot m} \ar[d]^{\alpha} & F_1(A)=S_1 \ar[d]_{\alpha} \\
      F_2(A)=S_2 \ar[r]^{F_2(m)=-\cdot m} & F_2(A)=S_2
                   }
    \end{align*}
    (since there is only one object) or more concretely for each $s\in S_1$,
    \begin{align*}
      \alpha(s\cdot m)=\alpha(s)\cdot m.
    \end{align*}

    Now the Yoneda lemma (for our special case) is an easy consequence:
    \begin{align*}
      [M^\op,\Set](H_A, F)\iso \{\text{homomorphisms }\underline M\to S\}\iso S= F(A)
    \end{align*}
    where $S$ is the $M$-Set corresponding to $F$.
  \end{enumerate}
\end{answer}

% \begin{exercise}
%   A presheaf X:C op→SetX : C^{op} \to Set is representable precisely if the comma category (Y,const X)(Y,const_X) has a terminal object. If a terminal object is (d,f:Y(d)→X)≃(d,f∈X(d))(d, f : Y(d) \to X) \simeq (d, f \in X(d)) then X≃Y(d)X \simeq Y(d).

% This follows from unwrapping the definition of morphisms in the comma category (Y,const X)(Y,const_X) and applying the Yoneda lemma to find
% (Y,const X)((c,f∈X(c)),(d,g∈X(d)))≃{u∈C(c,d):X(u)(g)=f}. (Y,const_X)((c,f \in X(c)), (d, g \in X(d))) \simeq \{ u \in C(c,d) : X(u)(g) = f \} \,.

% Hence (Y,const X)((c,f∈X(c),(d,g∈X(d)))≃pt(Y,const_X)((c,f \in X(c), (d, g \in X(d))) \simeq pt says precisely that X(−)(f):C(c,d)→X(c)X(-)(f) : C(c,d) \to X(c) is a bijection.
% \end{exercise}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-07-02 slides/slides-10/slides-10.tex
First version sheet and slides.

\documentclass{beamer}

\usepackage[utf8]{inputenc}

\beamertemplatenavigationsymbolsempty
\setbeamertemplate{footline}{
	\raisebox{0.5em}{\makebox[\paperwidth-0.5em][r]{
		\color{gray}\scriptsize\insertframenumber}}}

\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{cd, shapes, tikzmark}
\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}

\title{Presheaves, Representables and\\the Yoneda Lemma}
\author{Dominik Wagner}
%\date{June 7, 2017}

\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Contravariant Representable Functors}

  Let $\cat A$ be a locally small category and $A\in \cat A$. The \emph{contravariant representable functor} of $A$
  \begin{align*}
    H_A=\cat A(-,A)\from \cat A^\op\to \Set
  \end{align*}
  is defined by
  \begin{itemize}
  \item $H_A(B)=\cat A(B,A)$ for objects $B\in\cat A$ and
  \item for maps $B'\xrightarrow{f} B$:
    \begin{align*}
      H_A(f)=-\of f\from \cat A(B,A)&\to \cat A(B',A)\\
                                      g &\mapsto g\of f\\\\
      \xymatrix{B' \ar[r]^{f} & B \ar[r]^{g} & A }
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Yoneda Embedding}
  Let $\cat A$ be locally small. The \emph{Yoneda embedding} of $\cat A$ is the functor
  \begin{align*}
    H_-\from \cat A\to[\cat A^\op,\Set]
  \end{align*}
  defined by
  \begin{itemize}
  \item $H_-(A)=H_A$ for objects $A\in\cat A$ and
  \item for maps $A\xrightarrow{f}A'$, $H_-(f)$ is the natural transformation defined for $B\in \cat A$ by
    \begin{align*}
      H_A(B)=\cat A(B,A)&\to H_{A'}(B)=\cat A(B,A')\\
      g&\mapsto  f\of g\\
    \end{align*}
  \end{itemize}
  \vspace{-0.8cm}
  \includegraphics{nattrans.pdf}
\end{frame}

\begin{frame}
  \frametitle{Yoneda Lemma and Yoneda Principle}
  \begin{lemma}[Yoneda]
    Let $\cat A$ be locally small. Then
    \begin{align*}
      [\cat A^\op,\Set](H_A,F)\iso F(A)
    \end{align*}
    naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$.
  \end{lemma}
  \begin{theorem}
    The Yoneda embedding $H_-\from \cat A\to[\cat A^\op,\Set]$ is full and faithful.
  \end{theorem}
  \begin{corollary}[Yoneda Principle]
    Let $\cat A$ be locally small and $A,A'\in\cat A$. Then
    \begin{align*}
      H_A\iso H_{A'} \iff A\iso A'
    \end{align*}
  \end{corollary}
\end{frame}
\end{document}

% vim mode line:
% vim: ts=2 sts=2 sw=2 expandtab

### 2017-07-03 sheets/talk-10/sheet-10.tex
More extensions.

@@
 
+\begin{exercise}
+  Let $\cat A$ be a locally small CCC. Use the Yoneda embedding to show that for any objects $A,B,C\in\cat A$,
+  \begin{align*}
+    (A\times B)^C\iso A^C\times B^C
+  \end{align*}
+  If $\cat A$ also has binary coproducts, show that also
+  \begin{align*}
+    A^{B+C}&\iso A^B\times A^C\\
+    (A\times B)+(A\times C)&\iso A\times (B+C)
+  \end{align*}
+\end{exercise}
+
+\begin{exercise}
+  Let $\cat B$ be a category and $J\from\cat C\to D$ a functor. There is an induced functor
+  \begin{align*}
+    J\of -\from[\cat B,\cat C]\to [\cat B,\cat D]
+  \end{align*}
+  defined by composition with $J$.
+  \begin{enumerate}
+  \item Show that if $J$ is full and faithful then so is $J\of -$.
+  \item Deduce that if $J$ is full and faithful and $G,G'\from\cat B\to\cat C$ with $J\of G\iso J\of G'$ then $G\iso G'$.
+  \item Now deduce that right adjoints are unique: if $F\from \cat A\to\cat B$ and $G,G'\from\cat B\to \cat A$ with $F\dashv G$ and $F\dashv G'$ then $G\iso G'$. 
+    \hint{The Yoneda embedding is full and faithful.}
+  \end{enumerate}
+\end{exercise}
+
+\begin{definition}{Preservation of Limits}
+  Let $\mathbf I$ be a small category.
+A functor $F\from \cat A\to\cat B$ \emph{preserves colimits of shape $\mathbf I$} if for all diagrams $D\from\mathbf I\to\cat A$ and all cocones $(D(I)\xrightarrow{p_I}A)_{I\in \mathbf I}$ on $D$,
+  \begin{align*}
+    &(D(I)\xrightarrow{p_I}A)_{I\in \mathbf I} \text{ is a limit cone on $D$ in } \cat A\\
+    \Longrightarrow&(F\of D(I)\xrightarrow{F(p_I)}F(A))_{I\in \mathbf I} \text{ is a limit cone on $F\of D$ in } \cat A
+  \end{align*}
+
+  A functor $F\from \cat A\to\cat B$ \emph{preserves limits} if it preserves limits of shape $\mathbf I$ for all small categories $\mathbf I$.
+\end{definition}
+\begin{exercise}
+  Fill in some of the missing details in the proof that for any small category $\cat A$ the presheaf category $[A^\op,\Set]$ is cartesian closed:
+  \begin{enumerate}
+  \item Show that for any $A\in\cat A$, $H_A$ preserves colimits. As a consequence $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\to\mathbf I}(H_A\of D)$$ for all shapes and diagrams.
+    \hint{It suffices to show that it preserves all products and equalizers.}
+  \item Let $F\from \cat A\to\Set$ be a presheaf.
+Show that the functor $-\times F\from [\cat A^\op,\Set]\to[\cat A^\op,\Set]$ preserves colomits.
+That is: For any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$,
+    \begin{align*}
+      \lim_{\to \mathbf I} (D\times F)\iso (\lim_{\to \mathbf I} D)\times F.
+    \end{align*}
+    You may assume that $[\cat A^\op,\Set]$ has all limits.
+(In particular limits, which are functors $\cat A^\op\to\Set$, are taken ``pointwise'': $$\left(\lim_{\to I\in\mathbf I} D(I)\right)(A) \iso \lim_{\to I\in\mathbf I} \left(D(I)(A)\right)$$ for all shapes and diagrams.)
+    \hint{TODO}
+  \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+  Let $\cat A$ be a small category.
+Prove that the representable functors \emph{generate} the presheaf category $[\cat A^\op,\Set]$ in the folowing sense: Given any objects $F,G\in[\cat A^\op,\Set]$ and natural transformations $\alpha,\beta\from F\to G$, if for every representable functor $H_A$ and natural transformation $\gamma\from H_A\to F$, one has $\alpha\of\gamma=\beta\of\gamma$, then $\alpha=\gamma$.
+Thus, the arrows in $[\cat A^\op,\Set]$ are determined by their effect on generalized elements based at representables.
+\end{exercise}
+
+\begin{exercise}
+  special case of yoneda embedding: groups vs. Caley-representation
+\end{exercise}
+
 % \begin{exercise}

### 2017-07-03 slides/slides-10/slides-10.tex
Extended slides.

@@
 \end{frame}
+
+\begin{frame}
+  \frametitle{Cartesian Closedness of Presheaf Category}
+  \begin{lemma}
+    For any small category $\cat A$, any functor $F\from\cat A^\op\to\Set$ is a colimit of representable functors,
+    \begin{align*}
+      \lim_{\to I\in\mathbf I} H_{\pi(I)}\iso F
+    \end{align*}
+    for some small category $\mathbf I$ and functor $\pi\from \mathbf I\to \cat A$.
+  \end{lemma}
+  \begin{lemma}
+    For any $A\in\cat A$, $H_A$ preserves colimits. In particular $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\to\mathbf I}(H_A\of D)$$ 
+    for any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$.
+  \end{lemma}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Cartesian Closedness of Presheaf Category (Cont.)}
+  \begin{lemma}
+    Let $F\from \cat A\to\Set$ be a presheaf. The functor $-\times F\from [\cat A^\op,\Set]\to[\cat A^\op,\Set]$ preserves colomits. That is:
+    \begin{align*}
+      \lim_{\to \mathbf I} (D\times F)\iso (\lim_{\to \mathbf I} D)\times F
+    \end{align*}
+    for any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$.
+  \end{lemma}
+
+  \begin{theorem}
+    For any small category $\cat A$, the category of presheaves $\cat A^\op,\Set$ is cartesian closed.
+  \end{theorem}
+\end{frame}
 \end{document}

### 2017-07-04 slides/slides-10/notes-10.tex
More extensions and started notes.

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

% \beamertemplatenavigationsymbolsempty
% \setbeamertemplate{footline}{
% 	\raisebox{0.5em}{\makebox[\paperwidth-0.5em][r]{
% 		\color{gray}\scriptsize\insertframenumber}}}

\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{cd, shapes, tikzmark}
\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}

\title{Presheaves, Representables and\\the Yoneda Lemma}
\author{Dominik Wagner}
%\date{June 7, 2017}

\begin{document}

\section{Introduction}
In this chapter we prove the following three propositions, which we mentioned earlier but failed to prove directly:
\begin{itemize}
\item Left adjoints are unique up to isomorphisms;
\item If $\cat A$ is cartesian closed and has coproducts, then for all $A,B,C\in\cat A$
  \begin{align*}
    (A\times B)+(A\times C)\iso A\times (B+C);
  \end{align*}
\item For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
\end{itemize}
Along the way, we prove a very important technical lemma called ``Yoneda lemma''. One of its consequences is that two obects are isomorphic if and only if they ``look the same from all other objects''.

In order to get a first understanding what this is supposed to mean we consider the example of a partially order set $(P,\leq)$. Let $p,p'\in P$ be arbitrary. Then we clearly have 
\begin{align*}
  (p\leq p')\iff (\forall q\in P\ldotp q\leq p\rightarrow q\leq p')
\end{align*}
or in other terms
\begin{align*}
  (p\leq p')\iff \downarrow(p)=\{q\in P\mid q\leq p\}\subseteq \{q\in P\mid q\leq p'\}=\downarrow(p'). *
\end{align*}

How can we lift this to arbitrary categories? We can do so by noting that the two sets correspond to the maps into $p$ and $q$ respectively. However, since maps are not unique in general categories we have to do a little bit more work. Recall the following special functor introduced in \ref[?]:

\begin{definition}[Contravariant Representable Functor]
  Let $\cat A$ be a locally small category and $A\in \cat A$. The \emph{contravariant representable functor} of $A$
  \begin{align*}
    H_A=\cat A(-,A)\from \cat A^\op\to \Set
  \end{align*}
  is defined by
  \begin{itemize}
  \item $H_A(B)=\cat A(B,A)$ for objects $B\in\cat A$ and
  \item for maps $B'\xrightarrow{f} B$:
    \begin{align*}
      H_A(f)=-\of f\from \cat A(B,A)&\to \cat A(B',A)\\
      g &\mapsto g\of f\\\\
      \xymatrix{B' \ar[r]^{f} & B \ar[r]^{g} & A }
    \end{align*}
  \end{itemize}
\end{definition}

Clearly, for each $p$ in a preorder regarded as a category $\downarrow(p)$ is essentially the same as $H_p$. Furthermore, inclusion are essentially natural transformations. Then, * corresponds to the fact that the following functor is full and faithful:

\begin{definition}[Yoneda Embedding]
  Let $\cat A$ be locally small. The \emph{Yoneda embedding} of $\cat A$ is the functor
  \begin{align*}
    H_-\from \cat A\to[\cat A^\op,\Set]
  \end{align*}
  defined by
  \begin{itemize}
  \item $H_-(A)=H_A$ for objects $A\in\cat A$ and
  \item for maps $A\xrightarrow{f}A'$, $H_-(f)$ is the natural transformation defined for $B\in \cat A$ by
    \begin{align*}
      H_A(B)=\cat A(B,A)&\to H_{A'}(B)=\cat A(B,A')\\
      g&\mapsto  f\of g **\\
    \end{align*}
  \end{itemize}
  \vspace{-0.8cm}
  \includegraphics{nattrans.pdf}
\end{definition}
Basically, for preorders ** constructs a proof from $p\leq p'$ that for all $q\leq p$ also $q\leq p'$ holds.

In what follows we prove that this is indeed an embedding, i.e. full and faithful.

\section{Yoneda Lemma}
To do so we state the following technical lemma due to Yoneda:
\begin{lemma}[Yoneda]
  Let $\cat A$ be locally small. Then
  \begin{align*}
    [\cat A^\op,\Set](H_A,F)\iso F(A)
  \end{align*}
  naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$. That is for each $B\xrightarrow{f}A$ in $\cat A$ and for each natural transformation $\theta \from F\to F'$ the following commute
  \begin{align*}
    \xymatrix{
    [\cat A^\op,\Set](H_A,F) \ar[r]^{-\of H_f} & [\cat A^\op,\Set](H_B,F) \\
    F(A) \ar[r]^{F(f)} \ar[u]^{\iso}& F(B) \ar[u]_{\iso}
                         } 
  \end{align*}
  \begin{align*}
    \xymatrix{
    [\cat A^\op,\Set](H_A,F) \ar[r]^{\theta\of -} & [\cat A^\op,\Set](H_A,F') \\
    F(A) \ar[r]^{\theta_A} \ar[u]^{\iso} & F'(A) \ar[u]^{\iso}
                             } 
  \end{align*}
\end{lemma}
It is going to help us understand what the arrows $F_A\to F_{A'}$ are (in the presheaf category). More generally, it states that an element of $F(A)$ ``is the same'' as a natural transformation $H_A\to F$.
\begin{proof}
  In order to show the isomorphism in $\Set$, which is supposed to be natural in both $A$ and $F$, we can proceed by:
  \begin{enumerate}
  \item Define a function $(\widehat{-})\from[\cat A^\op,\Set](H_A,F)\to F(A)$.
  \item Define a function $(\widetilde{-})\from F(A)\to[\cat A^\op,\Set](H_A,F)$.
  \item Show that $(\widehat{\widetilde{-}})=1_{F(A)}$.
  \item Show that $(\widetilde{\widehat{-}})=1_{[\cat A^\op,\Set](H_A,F)}$.
  \item Show that $(\widehat{-})$ is natural in $A$.
  \item Show that $(\widehat{-})$ is natural in $F$.
  \end{enumerate}
  It turns out that the definitions of $(\widehat{-})$ and $(\widetilde{-})$ are straight-forward as there is only one canonical choice that can possibly be made:
  \begin{enumerate}
  \item Let $\alpha\from H_A\to F$ be arbitrary. In order to define $(\widehat{-})\from[\cat A^\op,\Set](H_A,F)\to F(A)$, we have to come up with an element $\widehat\alpha\in F(A)$. However, $\alpha$ gives us a function $\alpha_A\from H_A(A)\to F(A)$ in $\Set$. Hence, we only have to come up with an element of $H_A(A)=\cat A(A,A)$ to obtain an element of $F(A)$ and we can simply choose $1_A$. Thus, we define $\widehat\alpha=\alpha_A(1_A)$.
  \item Let $a\in F(A)$. This time, we have to define a natural transformation $\widetilde a\from H_A\to F$, i.e. for each $B\in\cat A$ we have to define a function $\widetilde a_B\from H_A(B)\to F(B)$ in $\Set$ and show that the naturality condition is satisfied.

    In order to define $\tilde a_B$, let $f\in H_A(B)=\cat A(B,A)$ be arbitrary. We have to construct an element of $F(B)$. However, we have a functor $F\from\cat A^\op\to\Set$ and $F(f)\from F(A)\to F(B)$. This is perfect since we already have $a\in F(A)$. Thus, $F(f)(a)\in F(A)$ and we can define 
    \begin{align*}
      \widetilde a_B(f)=F(f)(a).
    \end{align*}

    We still have to check naturality, i.e. that for any $B'\xrightarrow{g}B$ the following commutes:
    \begin{align*}
    \xymatrix{
    F_A(B)=\cat A(B,A) \ar[r]^{H_A(g)=-\of g} \ar[d]^{\widetilde a_B} & F_A(B')=\cat A(B',A) \ar[d]_{\widetilde a_B'} \\
    F(B) \ar[r]^{F(g)} & F(B').
                         } 
  \end{align*}
  Since this square is in $\Set$, it suffices to check $(\widetilde a_B'\of H_A(g))(f)=(F(g)\of\widetilde a_B)(f)$ for every $f\in F_A(B)=\cat A(B,A)$. Indeed, this holds:
  \begin{align*}
    (\widetilde a_B'\of H_A(g))(f)&=(F(f\of g))(a)\\
    &=(F(f)\of F(g))(a)&F \text{ functor}\\
    &=F(f)(F(g)(a))\\
    &=(F(g)\of\widetilde a_B)(f).
  \end{align*}
\item Let $a\in F(A)$. Clearly, we get $\widehat{\widetilde a}=a$:
  \begin{align*}
    \widehat{\widetilde a}=\widetilde a_A(1_A)=F(1_A)(a)=1_{F(A)}(a)=a.
  \end{align*}
\item[(d)-(f)] The remaining proof obligations are quite straight-forward to show and are thus left as an exercise.
  \end{enumerate}
\end{proof}

Now, it is easy to show that $H_-$ is full and faithful by instantiating $F$ with a covariant representable functor.
\begin{theorem}
  The Yoneda embedding $H_-\from \cat A\to[\cat A^\op,\Set]$ is full and faithful.
\end{theorem}
\begin{proof}
  We have 
  \begin{align*}
    \widetilde H_f=(H_f)_A(1_A)=f\of 1_A=f.
  \end{align*}
  Hence,
  \begin{align*}
    \cat A(A,A')&\to[\cat A^\op,\Set](H_A,H_A')\\
    f&\mapsto H_f
  \end{align*}
  and
  \begin{align*}
    (\widetilde{-})\from F(A)\to[\cat A^\op,\Set](H_A,F)
  \end{align*}
  are the same function, which are bijective due to the proof of the Yoneda lemma (instantiating $F$ with $H_{A'}$).
\end{proof}

By a previous exercise we know the following:
\begin{lemma}
  Let $F\from\cat A\to\cat B$ be a functor and $A\xrightarrow{f} A'$ in $\cat A$.

  Then $f$ is an isomorphism in $\cat A$ if and only if $F(f)$ is an isomorphism in $\cat B$.
\end{lemma}
Thus, we immediately get:
\begin{corollary}[Yoneda Principle]
  Let $\cat A$ be locally small and $A,A'\in\cat A$. Then
  \begin{align*}
    H_A\iso H_{A'} \iff A\iso A'
  \end{align*}
\end{corollary}
Note that the direction from right to left is trivial. Thus, the important consequence of this result is that
\begin{align*}
  H_A(B)=\cat A(B,A)\iso \cat  A(B,A')=H_A'(B) \text{ naturally in } B\in\cat A
\end{align*}
implies $A\iso A'$. Consequently, the following slogan is justified:
\begin{quote}
  \textit{``Two objects are the same if and only if they look the same from all viewpoints.''}
\end{quote}
Note that it is really necessary for the isomorphism to be natural in $B\in\cat A$. As the following example illustrates:
TODO
\section{Consequences of the Yoneda Lemma/Principle}
Next, we use the insights just gained and we consider (further) consequences of the Yoneda lemma/principle. In particular, we are going to proof the facts mentioned in the introduction.

We start by proving the following:

\begin{proposition}
  Let $\cat A$ be cartesian closed such that it has coproducts, then for all $A,B,C\in\cat A$
  \begin{align*}
    (A\times B)+(A\times C)\iso A\times (B+C).
  \end{align*}
\end{proposition}
\begin{proof}
  We have 
  \begin{align*}
    \cat A(A\times (B+C),X)&\iso\cat A(B+C,X^A)\\
                           &\iso\cat A(B,X^A)\times \cat A(C,X^A) &\text{needs justification}\\
                           &\iso\cat A(A\times B,X)\times \cat A(A\times C,X) \\
                           &\iso\cat A((A\times B)+(A\times C),X).&\text{needs justification}
  \end{align*}

  All these isomorphisms are natural in $X\in\cat A$. Hence, we have 
  \begin{align*}
    \cat A^\op(X,A\times (B+C))\iso\cat A^\op(X,(A\times B)+(A\times C))
  \end{align*}
 naturally in $X\in\cat A$ and thus we have $(A\times B)+(A\times C)\iso A\times (B+C)$.
\end{proof}

Another rather easy consequence is that left adjoints are unique:

\begin{proposition}
  Let $F,F'\from \cat A\to\cat B$ and $G\from\cat B\to\cat A$ such that $F\dashv G$ and $F'\dashv G$. Then $F\iso F'$
\end{proposition}
\begin{proof}
  By adjointness we have 
  \begin{align*}
    \cat B^\op(B,F(A))\iso\cat B(F(A),B)\iso\cat A(A,G(B))\iso\cat B(F'(A),B)\iso\cat B^\op(B,F'(A))
  \end{align*}
  naturally in $B\in\cat B$. Hence, by what was stated above we have $F(A)\iso F'(A)$ for all $A\in\cat A$. It can be easily shown that this isomorphsim is natural in $A$ (why?), hence $F\iso F'$.
\end{proof}

\section{Presheaf Category is Cartesian Closed}
In this section we show that for any small category $\cat A$ the presheaf category $[\cat A^\op,\Set]$ is cartesian closed. It is easy to see that it has terminal objects and binary products by defining them ``pointwise''. Showing that exponentials exist is still non-trivial. However the Yoneda lemma provides a straight-forward definition of exponentials and in a special case verifying that this is indeed an exponential is simple. In the general case we need a couple of additional lemmas.

Thus, we are left to show that exponentials exists, i.e. that for functors $G,H\in[\cat A^\op,\Set]$, there is a functor $H^G\in[\cat A^\op,\Set]$ such that
\begin{align*}
  [\cat A^\op,\Set](F,H^G)\iso [\cat A^\op,\Set](F\times G,H)***
\end{align*}
naturally in $F\in[\cat A^\op,\Set]$.

To define $H^G$ on objects $A\in\cat A$ note that by the Yoneda lemma we have
\begin{align*}
  H^G(A)\iso [\cat A^\op,\Set](H_A,H^G)
\end{align*}
and if $H^G$ is the exponential of $G$ and $H$ we certainly have
\begin{align*}
  [\cat A^\op,\Set](H_A,H^G)\iso [\cat A^\op,\Set](H_A\times G,H).
\end{align*}
Clearly this set exists and we can thus define
\begin{align*}
  H^G(A)=[\cat A^\op,\Set](H_A\times G,H)
\end{align*}
on objects.
For maps $A\xrightarrow{f}A'$ we define $H^G(f)=H^G_f\from [\cat A^\op,\Set](H_{A'}\times G,H)\to[\cat A^\op,\Set](H_A\times G,H)$ by
$H^G_f(\alpha)\in[\cat A^\op,\Set](H_A\times G,H)$ for $\alpha\in [\cat A^\op,\Set](H_{A'}\times G,H)$ by
\begin{align*}
  H^G_f(\alpha)_B = (H_g\times 1)\of \alpha_B\from (H_{A}\times G)(B)\to H(B)
\end{align*}
for $B\in\cat A$, which is of course natural.

Therefore, we are left to verify that (***) holds and that it is natural in $F\in[\cat A^\op,\Set]$.
In order to understand the proof better let us first consider the special case where $F=H_A$ is a covariant representable functor for some $A\in\cat A$. Then we immediately have
\begin{align*}
  [\cat A^\op,\Set](H_A,H^G)&\iso H^G(A)&\text{Yoneda lemma}\\
  &=[\cat A^\op,\Set](H_A\times G,H)&\text{definition of }H^G
\end{align*}
naturally in $A\in\cat A$.

Unfortunately, we have to show that this natural in \emph{all} $F\in[\cat A^\op,\Set]$, not just covariant representable functors, and not every functor is isomorphic to some $H_A$. However, it turns out that each presheaf is a \emph{colimit} of representable functors:

\begin{lemma}
  For any small category $\cat A$, any functor $F\from\cat A^\op\to\Set$ is a colimit of representable functors,
  \begin{align*}
    \lim_{\to I\in\mathbf I} H_{\pi(I)}\iso F
  \end{align*}
  for some small category $\mathbf I$ and functor $\pi\from \mathbf I\to \cat A$.
\end{lemma}

\end{document}

% vim mode line:
% vim: ts=2 sts=2 sw=2 expandtab

### 2017-07-04 slides/slides-10/notes-10.tex
Extended notes.

@@
 \end{lemma}
+\begin{proof}
+  First of all, we have to define the shape of the limit, i.e. a small category $\mathbf I$. We take the category of elements of $F$, $F(\cat A)$ with
+  \begin{itemize}
+  \item objects being pairs $(a,A)$ such that $A\in\cat A$ and $a\in F(A)$;
+  \item arrows $f\from(a,A)\to(b,B)$ being arrows $f\from A\to B$ in $\cat A$ such that $F(f)(a)=b$.
+  \end{itemize}
+  TODO: figure
+
+  and identities and composites defined in the natural way. Clearly, $F(\cat A)$ is small since $\cat A$ is small.
+
+  Next, we have to define a diagram $F(\cat A)\to[\cat A^\op,\Set]$ such that $F$ is the corresponding limit.
+We simply take $H_-\of\pi$ where $\pi\from F(\cat A)\to\cat A$ is the projection on $\cat A$ defined by
+  \begin{itemize}
+  \item $\pi(a,A)=A$ on objects and
+  \item $\pi(h)=h$ on arrows in $\cat A$ .
+  \end{itemize}
+
+  Now, we have to verify that $F$ is indeed the limit of $H_-\of\pi$ under $F(\cat A)$, i.e. we have to specify a cocone
+  \begin{align*}
+    (H_{\pi(a,A)}\to F)_{(a,A)\in F(\cat A)}
+  \end{align*}
+  and show that it is universal amongst all cocones.
+
+  To define $H_{\pi(a,A)}=H_A\to F$ for any $(a,A)\in F(\cat A)$ we recall that $[\cat A^\op,\Set](H_A,F)\iso F(A)$, where the isomorphism from the right to the left is given by $(\widetilde -)$ (as defined in the proof of the Yoneda lemma).
+Thus, we define the $(a,A)$-component of the cocone to be $\widetilde a$.
+For this to be a cocone we still have to verify that for each $(b,B)\xrightarrow{f}(a,A)$ (i.e.
+$B\xrightarrow{f}A$ such that $F(f)(b)=a$)
+  \begin{align*}
+    \xymatrix{
+      H_{\pi(b,B)}=H_B \ar[r]^{\widetilde b} \ar[d]_{H_{\pi(f)}=H_f} & F \\
+      H_{\pi(a,A)}=H_A \ar[ru]_{\widetilde a} & 
+                         } 
+  \end{align*}
+  commutes. However, by the naturality statement for $A\in\cat A$ in the Yondea lemma we have that 
+  \begin{align*}
+    \xymatrix{
+    [\cat A^\op,\Set](H_A,F) \ar[r]^{-\of H_f} & [\cat A^\op,\Set](H_B,F) \\
+    F(A) \ar[r]^{F(f)} \ar[u]^{(\widetilde-)}& F(B) \ar[u]_{(\widetilde -)}
+                         } 
+  \end{align*}
+  commutes. In particular for any $a\in F(A)$ we have
+  \begin{align*}
+    \widetilde a\of H_f=\widetilde{F(f)(a)}=\widetilde b
+  \end{align*}
+  by definition of arrows in $F(\cat A)$.
+
+  Thus, we have shown that this indeed constitues a cocone. Similarly, it can be shown that this cocone is universal, i.e. that $F$ is the limit of $H_-\of\pi$ under $F(\cat A)$.
+\end{proof}
 

### 2017-07-04 sheets/talk-10/sheet-10.tex
Extended sheet.

@@
 \begin{exercise}
+  Recall the following re-formulation of the Yoneda principle: For any locally small category $\cat A$ and objects $A,A'\in\cat A$:
+  \begin{align*}
+    \cat A(B,A)\iso \cat  A(B,A')\text{ naturally in } B\in\cat A
+  \end{align*}
+implies $A\iso A'$. Show that naturality is really necessary.
+\hint{There is a counterexample with two objects.}
+\end{exercise}
+
+\begin{exercise}
   Let $\cat A$ be a locally small CCC. Use the Yoneda embedding to show that for any objects $A,B,C\in\cat A$,
@@
   \begin{align*}
-    A^{B+C}&\iso A^B\times A^C\\
-    (A\times B)+(A\times C)&\iso A\times (B+C)
+    A^{B+C}&\iso A^B\times A^C
   \end{align*}
@@
   \end{enumerate}
+\end{exercise}
+
+\begin{exercise}
+  Let $\cat A$ be a small category.
+Prove that the representable functors \emph{generate} the presheaf category $[\cat A^\op,\Set]$ in the folowing sense: Given any objects $F,G\in[\cat A^\op,\Set]$ and natural transformations $\alpha,\beta\from F\to G$, if for every representable functor $H_A$ and natural transformation $\gamma\from H_A\to F$, one has $\alpha\of\gamma=\beta\of\gamma$, then $\alpha=\gamma$.
+Thus, the arrows in $[\cat A^\op,\Set]$ are determined by their effect on generalized elements based at representables.
 \end{exercise}
@@
 (In particular limits, which are functors $\cat A^\op\to\Set$, are taken ``pointwise'': $$\left(\lim_{\to I\in\mathbf I} D(I)\right)(A) \iso \lim_{\to I\in\mathbf I} \left(D(I)(A)\right)$$ for all shapes and diagrams.)
-    \hint{TODO}
+    \hint{Use the fact that $[\cat A^\op,\Set]$ has all limits to deduce that there is an arrow $\lim_{\to \mathbf I} (D\times F)\to (\lim_{\to \mathbf I} D)\times F$ and use the Yoneda principle to show that this is an isomorphism.}
   \end{enumerate}
 \end{exercise}
-
-\begin{exercise}
-  Let $\cat A$ be a small category.
-Prove that the representable functors \emph{generate} the presheaf category $[\cat A^\op,\Set]$ in the folowing sense: Given any objects $F,G\in[\cat A^\op,\Set]$ and natural transformations $\alpha,\beta\from F\to G$, if for every representable functor $H_A$ and natural transformation $\gamma\from H_A\to F$, one has $\alpha\of\gamma=\beta\of\gamma$, then $\alpha=\gamma$.
-Thus, the arrows in $[\cat A^\op,\Set]$ are determined by their effect on generalized elements based at representables.
-\end{exercise}
-
-\begin{exercise}
-  special case of yoneda embedding: groups vs. Caley-representation
-\end{exercise}
-
-% \begin{exercise}
-%   A presheaf X:C op→SetX : C^{op} \to Set is representable precisely if the comma category (Y,const X)(Y,const_X) has a terminal object.
-If a terminal object is (d,f:Y(d)→X)≃(d,f∈X(d))(d, f : Y(d) \to X) \simeq (d, f \in X(d)) then X≃Y(d)X \simeq Y(d).
-
-% This follows from unwrapping the definition of morphisms in the comma category (Y,const X)(Y,const_X) and applying the Yoneda lemma to find
-% (Y,const X)((c,f∈X(c)),(d,g∈X(d)))≃{u∈C(c,d):X(u)(g)=f}. (Y,const_X)((c,f \in X(c)), (d, g \in X(d))) \simeq \{ u \in C(c,d) : X(u)(g) = f \} \,.
-
-% Hence (Y,const X)((c,f∈X(c),(d,g∈X(d)))≃pt(Y,const_X)((c,f \in X(c), (d, g \in X(d))) \simeq pt says precisely that X(−)(f):C(c,d)→X(c)X(-)(f) : C(c,d) \to X(c) is a bijection.
-% \end{exercise}
 

### 2017-07-04 sheets/talk-10/sheet-10.tex
Extensions and fixes to notes, slides and sheet.

@@
 
-\begin{definition}{Preservation of Limits}
-  Let $\mathbf I$ be a small category.
+% \begin{definition}{Preservation of Limits}
+%   Let $\mathbf I$ be a small category.
 A functor $F\from \cat A\to\cat B$ \emph{preserves colimits of shape $\mathbf I$} if for all diagrams $D\from\mathbf I\to\cat A$ and all cocones $(D(I)\xrightarrow{p_I}A)_{I\in \mathbf I}$ on $D$,
-  \begin{align*}
-    &(D(I)\xrightarrow{p_I}A)_{I\in \mathbf I} \text{ is a limit cone on $D$ in } \cat A\\
-    \Longrightarrow&(F\of D(I)\xrightarrow{F(p_I)}F(A))_{I\in \mathbf I} \text{ is a limit cone on $F\of D$ in } \cat A
-  \end{align*}
+%   \begin{align*}
+%     &(D(I)\xrightarrow{p_I}A)_{I\in \mathbf I} \text{ is a limit cone on $D$ in } \cat A\\
+%     \Longrightarrow&(F\of D(I)\xrightarrow{F(p_I)}F(A))_{I\in \mathbf I} \text{ is a limit cone on $F\of D$ in } \cat A
+%   \end{align*}
 
-  A functor $F\from \cat A\to\cat B$ \emph{preserves limits} if it preserves limits of shape $\mathbf I$ for all small categories $\mathbf I$.
-\end{definition}
+%   A functor $F\from \cat A\to\cat B$ \emph{preserves limits} if it preserves limits of shape $\mathbf I$ for all small categories $\mathbf I$.
+% \end{definition}
 \begin{exercise}
@@
   \begin{enumerate}
-  \item Show that for any $A\in\cat A$, $H_A$ preserves colimits. As a consequence $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\to\mathbf I}(H_A\of D)$$ for all shapes and diagrams.
-    \hint{It suffices to show that it preserves all products and equalizers.}
+  \item Show that for any $A\in\cat A$, $H_A$ maps all colimits to limits such that $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\leftarrow\mathbf I}(H_A\of D)$$ for all small categories $\mathbf I$ and diagrams $D\from \mathbf I\to\cat A$.
+What is the dual statement?
+    \hint{It suffices to show that this holds for all coproducts and coequalizers.}
   \item Let $F\from \cat A\to\Set$ be a presheaf.

### 2017-07-04 slides/slides-10/notes-10.tex
Extensions and fixes to notes, slides and sheet.

@@
 \begin{itemize}
-\item Left adjoints are unique up to isomorphisms;
 \item If $\cat A$ is cartesian closed and has coproducts, then for all $A,B,C\in\cat A$
@@
   \end{align*}
+\item Left adjoints are unique up to isomorphisms;
 \item For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
@@
 
-We start by proving the following:
+We start by proving the following ``distributivity''-lemma:
 
 \begin{proposition}
+  \label{prop:dis}
   Let $\cat A$ be cartesian closed such that it has coproducts, then for all $A,B,C\in\cat A$
@@
 \end{proposition}
-\begin{proof}
+
+For the proof we need the following lemma, which we are also going to use in the proof that presheaf categories are cartesian closed:
+\begin{lemma}
+  \label{lem:HApres}
+  For any locally small category $\cat A$ and $A\in\cat A$, $H_A$ maps all colimits to limits such that $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\leftarrow\mathbf I}(H_A\of D)$$ for all small categories $\mathbf I$ and diagrams $D\from \mathbf I\to\cat A$.
+\end{lemma}
+Now the proof of proposition \ref{prop:dis} goes through smoothly:
+\begin{proof}[Proof of proposition \ref{prop:dis}]
   We have 
   \begin{align*}
-    \cat A(A\times (B+C),X)&\iso\cat A(B+C,X^A)\\
-                           &\iso\cat A(B,X^A)\times \cat A(C,X^A) &\text{needs justification}\\
-                           &\iso\cat A(A\times B,X)\times \cat A(A\times C,X) \\
-                           &\iso\cat A((A\times B)+(A\times C),X).&\text{needs justification}
+    \cat A(A\times (B+C),X)&\iso\cat A((B+C)\times A,X)\\
+                           &\iso\cat A(B+C,X^A)&\cat A\text{ cartesian closed}\\
+                           &\iso\cat A(B,X^A)\times \cat A(C,X^A) &\text{by lemma \ref{lem:HApres}}\\
+                           &\iso\cat A(A\times B,X)\times \cat A(A\times C,X) &\cat A\text{ cartesian closed}\\
+                           &\iso\cat A((A\times B)+(A\times C),X).&\text{by the dual of lemma \ref{lem:HApres}}
   \end{align*}
 
-  All these isomorphisms are natural in $X\in\cat A$. Hence, we have 
+  All these isomorphisms are natural in $X\in\cat A$. E.g. the second isomorphism is natural in $X\in\cat A$ because it is induced by a transposition. Hence, we have 
   \begin{align*}
@@
   \end{align*}
-  naturally in $B\in\cat B$. Hence, by what was stated above we have $F(A)\iso F'(A)$ for all $A\in\cat A$. It can be easily shown that this isomorphsim is natural in $A$ (why?), hence $F\iso F'$.
+  naturally in $B\in\cat B$. Hence, by what was stated above we have $F(A)\iso F'(A)$ for all $A\in\cat A$. It can be shown that this isomorphsim is natural in $A$ (why?), hence $F\iso F'$.
 \end{proof}
@@
   [\cat A^\op,\Set](H_A,H^G)&\iso H^G(A)&\text{Yoneda lemma}\\
-  &=[\cat A^\op,\Set](H_A\times G,H)&\text{definition of }H^G
+  &=[\cat A^\op,\Set](H_A\times G,H)&\text{definition of }H^G.
 \end{align*}
-naturally in $A\in\cat A$.
 
@@
 \begin{lemma}
+  \label{lem:lim}
   For any small category $\cat A$, any functor $F\from\cat A^\op\to\Set$ is a colimit of representable functors,
@@
 
+Now, are almost in a position where we can prove that presheaf categories are cartesian closed. All we need is the following lemma which states that we can ``pull limits out of products'':
+\begin{lemma}
+  \label{lem:pull}
+  Let $F\from \cat A\to\Set$ be a presheaf. The functor $-\times F\from [\cat A^\op,\Set]\to[\cat A^\op,\Set]$ preserves colomits. That is:
+  \begin{align*}
+    \lim_{\to \mathbf I} (D\times F)\iso (\lim_{\to \mathbf I} D)\times F
+  \end{align*}
+  for any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$.
+\end{lemma}
+Finally, we can prove
+\begin{proposition}
+  For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
+\end{proposition}
+\begin{proof}
+  It easy to prove that $[\cat A^\op,\Set]$ has terminal objects and binary products (by defining them ``pointwise'').
+
+  With the previous definitions we are left to verify that 
+  \begin{align*}
+    [\cat A^\op,\Set](F,H^G)\iso [\cat A^\op,\Set](F\times G,H)
+  \end{align*}
+  naturally in $F\in[\cat A^\op,\Set]$. Using lemma \ref{lem:lim} we have
+  \begin{align*}
+    F\iso\lim_{\to I\in\mathbf I} H_{\pi(I)}.
+  \end{align*}
+Hence, we conclude:
+  \begin{align*}
+    [\cat A^\op,\Set](F,H^G)&\iso [\cat A^\op,\Set](\lim_{\to I\in\mathbf I} H_{\pi(I)},H^G)&\\
+                            &\iso \lim_{\leftarrow I\in\mathbf I} [\cat A^\op,\Set](H_{\pi(I)},H^G)&\text{by lemma \ref{lem:HApres}}\\
+                            &\iso \lim_{\leftarrow I\in\mathbf I} H^G(\pi(I))&\text{by the Yoneda lemma}\\
+                            &\iso \lim_{\leftarrow I\in\mathbf I} [\cat A^\op,\Set](H_{\pi(I)}\times G,H)&\text{by definition of } H^G\\
+                            &\iso [\cat A^\op,\Set](\lim_{\to I\in\mathbf I}(H_{\pi(I)}\times G),H)&\text{by lemma \ref{lem:HApres}}\\
+                            &\iso [\cat A^\op,\Set](\lim_{\to I\in\mathbf I}(H_{\pi(I)})\times G,H)&\text{by lemma \ref{lem:pull}}\\
+                            &\iso[\cat A^\op,\Set](F\times G,H).
+  \end{align*}
+  All the isomorphisms are natural in $F$. Thus $H^G$ is indeed the exponential of $G$ and $H$.
+\end{proof}
+
 \end{document}

### 2017-07-04 slides/slides-10/slides-10.tex
More extensions and started notes.
Extensions and fixes to notes, slides and sheet.

@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
+
+\newtheorem{proposition}{Proposition}
 
@@
 \begin{frame}
-  \frametitle{Yoneda Lemma and Yoneda Principle}
-  \begin{lemma}[Yoneda]
+  \frametitle{Yoneda Lemma}
   Let $\cat A$ be locally small. Then
@@
   \end{align*}
-    naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$.
-  \end{lemma}
+  naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$. That is for each $B\xrightarrow{f}A$ in $\cat A$ and for each natural transformation $\theta \from F\to F'$ the following commute
+  \begin{align*}
+    \xymatrix{
+    [\cat A^\op,\Set](H_A,F) \ar[r]^{-\of H_f} & [\cat A^\op,\Set](H_B,F) \\
+    F(A) \ar[r]^{F(f)} \ar[u]^{\iso}& F(B) \ar[u]_{\iso}
+                         } 
+  \end{align*}
+  \begin{align*}
+    \xymatrix{
+    [\cat A^\op,\Set](H_A,F) \ar[r]^{\theta\of -} & [\cat A^\op,\Set](H_A,F') \\
+    F(A) \ar[r]^{\theta_A} \ar[u]^{\iso} & F'(A) \ar[u]^{\iso}
+                             } 
+  \end{align*}
+\end{frame}
+\begin{frame}
+\frametitle{The Yoneda Principle}
   \begin{theorem}
@@
   \end{theorem}
+  \pause
   \begin{corollary}[Yoneda Principle]
@@
 \begin{frame}
+  \frametitle{Consequences of the Yoneda Lemma/Principle}
+  \begin{proposition}
+    Let $\cat A$ be cartesian closed such that it has coproducts, then for all $A,B,C\in\cat A$
+    \begin{align*}
+      (A\times B)+(A\times C)\iso A\times (B+C).
+    \end{align*}
+  \end{proposition}
+  \pause
+  \begin{lemma}
+    For any locally small category $\cat A$ and $A\in\cat A$, $H_A$ maps
+ all colimits to limits such that $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\leftarrow\mathbf I}(H_A\of D)$$ for all small categories $\mathbf I$ and diagrams $D\from \mathbf I\to\cat A$.
+  \end{lemma}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Consequences of the Yoneda Lemma/Principle (Cont.)}
+  \begin{proposition}
+    Let $F,F'\from \cat A\to\cat B$ and $G\from\cat B\to\cat A$ such that $F\dashv G$ and $F'\dashv G$. Then $F\iso F'$
+  \end{proposition}
+\end{frame}
+
+\begin{frame}
   \frametitle{Cartesian Closedness of Presheaf Category}
+  \begin{proposition}
+    For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
+  \end{proposition}
+  \pause
   \begin{lemma}
@@
   \end{lemma}
-  \begin{lemma}
-    For any $A\in\cat A$, $H_A$ preserves colimits. In particular $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\to\mathbf I}(H_A\of D)$$ 
-    for any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$.
-  \end{lemma}
 \end{frame}
@@
   \begin{lemma}
-    Let $F\from \cat A\to\Set$ be a presheaf. The functor $-\times F\from [\cat A^\op,\Set]\to[\cat A^\op,\Set]$ preserves colomits. That is:
+    Let $F\from \cat A\to\Set$ be a presheaf. Then
     \begin{align*}
@@
   \end{lemma}
-
-  \begin{theorem}
-    For any small category $\cat A$, the category of presheaves $\cat A^\op,\Set$ is cartesian closed.
-  \end{theorem}
+  \pause
+  \begin{proposition}
+    For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
+  \end{proposition}
 \end{frame}

### 2017-07-05 slides/slides-10/notes-10.tex
Yet more changes.

@@
     (\widetilde a_B'\of H_A(g))(f)&=(F(f\of g))(a)\\
-    &=(F(f)\of F(g))(a)&F \text{ functor}\\
-    &=F(f)(F(g)(a))\\
+    &=(F(g)\of F(f))(a)&F \text{ contravariant functor}\\
+    &=F(g)(F(f)(a))\\
     &=(F(g)\of\widetilde a_B)(f).
@@
   \begin{align*}
-    \widetilde H_f=(H_f)_A(1_A)=f\of 1_A=f.
+    \widehat H_f=(H_f)_A(1_A)=f\of 1_A=f.
   \end{align*}
@@
 TODO
-\section{Consequences of the Yoneda Lemma/Principle}
+\section{Consequences of the Yoneda Principle}
 Next, we use the insights just gained and we consider (further) consequences of the Yoneda lemma/principle. In particular, we are going to proof the facts mentioned in the introduction.
@@
   \item objects being pairs $(a,A)$ such that $A\in\cat A$ and $a\in F(A)$;
-  \item arrows $f\from(a,A)\to(b,B)$ being arrows $f\from A\to B$ in $\cat A$ such that $F(f)(a)=b$.
+  \item arrows $f\from(a,A)\to(b,B)$ being arrows $f\from A\to B$ in $\cat A$ such that $F(f)(b)=a$.
   \end{itemize}
@@
 For this to be a cocone we still have to verify that for each $(b,B)\xrightarrow{f}(a,A)$ (i.e.
-$B\xrightarrow{f}A$ such that $F(f)(b)=a$)
+$B\xrightarrow{f}A$ such that $F(f)(a)=b$)
   \begin{align*}

### 2017-07-05 slides/slides-10/notes-10.tex
more changes

--- 2017-07-04 slides/slides-10/notes-10.tex
@@
   \item $H_-(A)=H_A$ for objects $A\in\cat A$ and
-  \item for maps $A\xrightarrow{f}A'$, $H_-(f)$ is the natural transformation defined for $B\in \cat A$ by
+  \item for maps $A\xrightarrow{f}A'$, $H_f$ is the natural transformation defined for $B\in \cat A$ by
     \begin{align*}
@@
 \end{lemma}
-It is going to help us understand what the arrows $F_A\to F_{A'}$ are (in the presheaf category).
+It is going to help us understand what the arrows $H_A\to H_{A'}$ are (in the presheaf category).
 More generally, it states that an element of $F(A)$ ``is the same'' as a natural transformation $H_A\to F$.
@@
     \xymatrix{
-    F_A(B)=\cat A(B,A) \ar[r]^{H_A(g)=-\of g} \ar[d]^{\widetilde a_B} & F_A(B')=\cat A(B',A) \ar[d]_{\widetilde a_B'} \\
+    H_A(B)=\cat A(B,A) \ar[r]^{H_A(g)=-\of g} \ar[d]^{\widetilde a_B} & H_A(B')=\cat A(B',A) \ar[d]_{\widetilde a_B'} \\
     F(B) \ar[r]^{F(g)} & F(B').
@@
   \end{align*}
-  Since this square is in $\Set$, it suffices to check $(\widetilde a_B'\of H_A(g))(f)=(F(g)\of\widetilde a_B)(f)$ for every $f\in F_A(B)=\cat A(B,A)$. Indeed, this holds:
+  Since this square is in $\Set$, it suffices to check $(\widetilde a_B'\of H_A(g))(f)=(F(g)\of\widetilde a_B)(f)$ for every $f\in H_A(B)=\cat A(B,A)$. Indeed, this holds:
   \begin{align*}
@@
 \begin{lemma}
-  Let $F\from\cat A\to\cat B$ be a functor and $A\xrightarrow{f} A'$ in $\cat A$.
+  Let $F\from\cat A\to\cat B$ be a full and faithful functor and $A\xrightarrow{f} A'$ in $\cat A$.
 
@@
 
-We start by proving the following ``distributivity''-lemma:
+We start by proving the following ``distributivity''-property:
 
@@
   \label{prop:dis}
-  Let $\cat A$ be cartesian closed such that it has coproducts, then for all $A,B,C\in\cat A$
+  Let $\cat A$ be cartesian closed such that it has coproducts. Then for all $A,B,C\in\cat A$
   \begin{align*}

### 2017-07-05 slides/slides-10/slides-10.tex
Yet more changes.
more changes

@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
-
-\newtheorem{proposition}{Proposition}
 
@@
   \frametitle{Consequences of the Yoneda Lemma/Principle}
-  \begin{proposition}
+  \begin{theorem}
     Let $\cat A$ be cartesian closed such that it has coproducts, then for all $A,B,C\in\cat A$
@@
     \end{align*}
-  \end{proposition}
+  \end{theorem}
   \pause
@@
   \frametitle{Consequences of the Yoneda Lemma/Principle (Cont.)}
-  \begin{proposition}
-    Let $F,F'\from \cat A\to\cat B$ and $G\from\cat B\to\cat A$ such that $F\dashv G$ and $F'\dashv G$. Then $F\iso F'$
-  \end{proposition}
+  \begin{theorem}
+    Let $F,F'\from \cat A\to\cat B$ and $G\from\cat B\to\cat A$ such that $F\dashv G$ and $F'\dashv G$. Then $F\iso F'$.
+  \end{theorem}
 \end{frame}
@@
   \frametitle{Cartesian Closedness of Presheaf Category}
-  \begin{proposition}
+  \begin{theorem}
     For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
-  \end{proposition}
+  \end{theorem}
   \pause
@@
   \pause
-  \begin{proposition}
+  \begin{theorem}
     For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
-  \end{proposition}
+  \end{theorem}
 \end{frame}

### 2017-07-05 sheets/talk-10/sheet-10.tex

@@
   Let $\cat A$ be a small category.
-Prove that the representable functors \emph{generate} the presheaf category $[\cat A^\op,\Set]$ in the folowing sense: Given any objects $F,G\in[\cat A^\op,\Set]$ and natural transformations $\alpha,\beta\from F\to G$, if for every representable functor $H_A$ and natural transformation $\gamma\from H_A\to F$, one has $\alpha\of\gamma=\beta\of\gamma$, then $\alpha=\gamma$.
+Prove that the representable functors \emph{generate} the presheaf category $[\cat A^\op,\Set]$ in the folowing sense: Given any objects $F,G\in[\cat A^\op,\Set]$ and natural transformations $\alpha,\beta\from F\to G$, if for every representable functor $H_A$ and natural transformation $\gamma\from H_A\to F$, one has $\alpha\of\gamma=\beta\of\gamma$, then $\alpha=\beta$.
 Thus, the arrows in $[\cat A^\op,\Set]$ are determined by their effect on generalized elements based at representables.
@@
   \item Let $F\from \cat A\to\Set$ be a presheaf.
-Show that the functor $-\times F\from [\cat A^\op,\Set]\to[\cat A^\op,\Set]$ preserves colomits.
+Show that the functor $-\times F\from [\cat A^\op,\Set]\to[\cat A^\op,\Set]$ preserves colimits.
 That is: For any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$,

### 2017-07-05 slides/slides-10/notes-10.tex

@@
                            &\iso\cat A(A\times B,X)\times \cat A(A\times C,X) &\cat A\text{ cartesian closed}\\
-                           &\iso\cat A((A\times B)+(A\times C),X).&\text{by the dual of lemma \ref{lem:HApres}}
+                           &\iso\cat A((A\times B)+(A\times C),X).&\text{by lemma \ref{lem:HApres}}
   \end{align*}
@@
   \item $\pi(a,A)=A$ on objects and
-  \item $\pi(h)=h$ on arrows in $\cat A$ .
+  \item $\pi(f)=f$ on arrows in $\cat A$ .
   \end{itemize}

### 2017-07-08 sheets/talk-10/sheet-10.tex
Extended solutions.

@@
 \end{exercise}
+\begin{answer}
+consider the following category $\cat A$: 
+\begin{align*}
+  \xymatrix{
+  A \ar@(ul,ur)^{1_A,f} \ar@/^0.3pc/@[][r]^{f_1,f_2} & B \ar@/^0.3pc/@[][l]^{g_1,g_2} \ar@(ul,ur)^{1_B,g}
+                                                       } 
+  \end{align*}
+  i.e. between any pair of object there are two arrows, where the composition of $f_i$ and $g_j$ ($i,j\in\{1,2\}$) is given by
+  \begin{align*}
+    f_i\of g_j&=g
+  \end{align*}
+  and the remaining compositions can be chosen arbitrarily.
+
+  Obviously,
+  \begin{align*}
+    \cat A(C,A)\iso\cat A(C,B)
+  \end{align*}
+  for all $C\in\cat A$ because all homsets contain exactly two elements.
+However $A\not\iso B$ because by construction neither $g_1$ nor $g_2$ (which are the only maps between $B$ and $A$) have a left inverse.
+\end{answer}
 
@@
 \begin{exercise}
-  Let $\cat B$ be a category and $J\from\cat C\to D$ a functor. There is an induced functor
+  Let $\cat B$ be a category and $J\from\cat C\to\cat D$ a functor. There is an induced functor
   \begin{align*}
@@
 \end{exercise}
+\begin{answer}
+  \begin{enumerate}
+  \item Clearly, $J\of-$ is defined on arrows $\alpha\from F\to G$ for $F,G\from\cat B\to\cat C$ by
+    \begin{align*}
+      J\of -(\alpha)&=\alpha'
+    \end{align*} 
+    where $\alpha$ is a natural transformation $J\of F\to J\of G$ defined by
+    \begin{align*}
+      \alpha'_B=J(\alpha_B)\from J\of F(B)\to J\of G(B)
+    \end{align*}
+    for $B\in\cat B$.
+
+    Full and faithfulness of $J$ means that for each $C,C'\in\cat C$ the function
+    \begin{align*}
+      J_{C,C'}\from\cat C(C,C)\to \cat D(J(C),J(C'))\\
+      f&\mapsto J(f)
+    \end{align*}
+    is bijective.
+
+    To show that $J\of-$ is full and faithful, let $F,G\from\cat B\to\cat C$ be natural transformations and $\beta\from J\of F\to J\of G$ be a natural transformation.
+    Define, a natural transformation $\alpha\from F\to G$ by 
+    \begin{align*}
+      \alpha_B=J_{F(A),G(A)}^{-1}(\beta_B)
+    \end{align*}
+    for $B\in\cat B$. This is well-defined by full- and faithfulness and defines indeed a natural transformation.
+
+    Clearly, $J\of -(\alpha)=\beta$. This shows fullness of $J\of -$. For faithfulness assume a natural transformation $\alpha'\from F\to G$ and suppose 
+    \begin{align*}
+      J\of-(\alpha)=J\of-(\alpha')
+    \end{align*}
+    By definition this means
+    \begin{align*}
+      J(\alpha_B)=J(\alpha'_B)
+    \end{align*}
+    for any $B\in\cat B$. Faithfulness of $J$ implies $\alpha_B=\alpha'_B$ for any $B\in\cat B$, i.e. $\alpha=\alpha'$.
+  \item \label{ex:lem} $J\of -\from[\cat B,\cat C]\to [\cat B,\cat D]$ is full and faithful by the previous part.
+Hence for $G,G'\in [\cat B,\cat C]$, $G\iso G'$ if and only if $J\of G=J\of-(G)\iso J\of-(G')=J\of G'$.
+  \item By adjointness we get
+    \begin{align*}
+      H_{G(B)}(A)=\cat A(A,G(B))\iso\cat B(F(A),B)\iso\cat A(A,G'(B))=H_{G'(B)}(A)
+    \end{align*}
+    naturally in $A\in\cat A$ and $B\in\cat B$. In particular,
+    \begin{align*}
+      (H_-\of G)(B)=H_{G(B)}\iso H_{G'(B)}=(H_-\of G')(B)
+    \end{align*}
+    naturally in $B$, i.e. $H_-\of G\iso H_-\of G'$. Due to the fact that the Yoneda embedding $H_-$ is full and faithful and part \ref{ex:lem}, $G\iso G'$.
+
+    % By adjointness we get
+%     \begin{align*}
+%       \cat A(A,G(B))\iso\cat B(F(A),B)\iso\cat A(A,G'(B))
+%     \end{align*}
+%     naturally in $A\in\cat A$ and $B\in\cat B$. % This means 
+%     % \begin{align*}
+%     %   H_{G(B)}(A)\iso H_{G'(B)}(A)
+%     % \end{align*}
+%     % naturally in $A$, i.e. $(H_-\of G)(B)=H_{G(B)}\iso H_{G'(B)}=(H_-\of G')(B)$.
+% This means there is a natural transformation $\eta_B$ that is an isomorphism and for which
+%     \begin{align*}
+%                 H_{G(B)}(A)\xrightarrow{\eta_{B,A}}H_{G'(B)}(A)
+%     \end{align*}
+%     naturally in $A$ and $B$.
+
+%     As a consequence $H_-(G(B))=H_{G(B)}\iso H_{G'(B)}=H_-(G'(B))$ and the isomorphism is given by $\eta_B$. By the Yoneda principle, $G(B)\iso G'(B)$.
+
+%     This isomorphism is even natural in $B$: Since $\cat A(A,G(B))\iso\cat A(A,G'(B))$ naturally in $B$ the following commutes for any $f\from B\to B'$:
+%     \begin{align*}
+%       \xymatrix{
+%       H_{G(B)} \ar[r]^{H_{G(f)}}\ar[d]^{\eta_B} & H_{G(B')} \ar[d]_{\eta_{B'}}\\
+%       H_{G(B')} \ar[r]^{H_{G'(f)}} & H_{G'(B')} 
+%                                         } 
+%   \end{align*}
+%   Because the Yoneda embedding $H_-$ is full and faithful this means
+%   \begin{align*}
+%     \xymatrix{
+%     G(B) \ar[r]^{G(f)}\ar[d]^{\zeta_B} & G(B') \ar[d]_{\zeta_{B'}}\\
+%     G(B') \ar[r]^{G'(f)} & G'(B') 
+%                                    } 
+%   \end{align*}
+%   commutes where $\zeta_B$ and $\zeta_{B'}$ are the isomorphism corresponding to $\eta_B$ and $\eta_{B'}$ respectively under $H_-$.
+Another consequence is that $G(B)\iso G'(B)$ naturally in $B$, i.e.
+$$.
+  \end{enumerate}
+\end{answer}
 
@@
 \end{exercise}
+
+\begin{answer}
+  Let $F,G,\alpha,\beta$ be given as stated above. By the naturality statement in the Yoneda lemma we have that
+  \begin{align*}
+    \xymatrix{
+    [\cat A^\op,\Set](H_A,F) \ar[r]^{\alpha\of -} & [\cat A^\op,\Set](H_A,G) \\
+    F(A) \ar[r]^{\alpha_A} \ar[u]^{\iso} & G(A) \ar[u]^{\iso}
+                             } 
+  \end{align*}
+commutes for any $A\in\cat A$. In particular, with the definitions from the proof of the Yoneda lemma we have that 
+\begin{align*}
+    \xymatrix{
+    [\cat A^\op,\Set](H_A,F) \ar[r]^{\alpha\of -} & [\cat A^\op,\Set](H_A,G) \ar[d]^{\widehat{-}} \\
+    F(A) \ar[r]^{\alpha_A} \ar[u]^{\widetilde{-}} & G(A) 
+                             } 
+\end{align*}
+commutes for any $A\in\cat A$. Consequently, for any $A\in\cat A$ and $a\in F(A)$,
+\begin{align*}
+  \alpha_A(a)=\widehat{\alpha\of\widetilde a}.
+\end{align*}
+Analogously,
+\begin{align*}
+  \beta_A(a)=\widehat{\beta\of\widetilde a}.
+\end{align*}
+By assumption ($\widetilde a\from H_A\to F$)
+\begin{align*}
+  \alpha\of\widetilde a=\beta\of\widetilde a
+\end{align*}
+holds and therefore
+\begin{align*}
+  \alpha_A(a)=\beta_A(a)
+\end{align*}
+for any $A\in\cat A$ and $a\in F(A)$, which means $\alpha=\beta$.
+\end{answer}
 

### 2017-07-12 slides/talk-12/header.tex
Add slides for talk 12

--- 2017-07-03 slides/slides-10/slides-10.tex
@@
-\documentclass{beamer}
-
-\usepackage[utf8]{inputenc}
+\usepackage{mathtools}
+\usepackage{mathpartir}
+\usepackage{semantic}
+\usepackage{tikzsymbols}  % This defines \Cat, that we want to redefine
+\usepackage{stmaryrd}  % This defines \mapsfrom, that we want to redefine
+\usepackage{tikz}
+\usetikzlibrary{cd}
+\usepackage{adjustbox}
+\usepackage{transparent}
 
@@
 
-\usepackage{mathtools}
-\usepackage{tikz}
-\usetikzlibrary{cd, shapes, tikzmark}
-\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
-
-\title{Presheaves, Representables and\\the Yoneda Lemma}
-\author{Dominik Wagner}
-%\date{June 7, 2017}
-
-\begin{document}
-
-\frame{\titlepage}
-
-\begin{frame}
-  \frametitle{Contravariant Representable Functors}
-
-  Let $\cat A$ be a locally small category and $A\in \cat A$. The \emph{contravariant representable functor} of $A$
-  \begin{align*}
-    H_A=\cat A(-,A)\from \cat A^\op\to \Set
-  \end{align*}
-  is defined by
-  \begin{itemize}
-  \item $H_A(B)=\cat A(B,A)$ for objects $B\in\cat A$ and
-  \item for maps $B'\xrightarrow{f} B$:
-    \begin{align*}
-      H_A(f)=-\of f\from \cat A(B,A)&\to \cat A(B',A)\\
-                                      g &\mapsto g\of f\\\\
-      \xymatrix{B' \ar[r]^{f} & B \ar[r]^{g} & A }
-    \end{align*}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Yoneda Embedding}
-  Let $\cat A$ be locally small. The \emph{Yoneda embedding} of $\cat A$ is the functor
-  \begin{align*}
-    H_-\from \cat A\to[\cat A^\op,\Set]
-  \end{align*}
-  defined by
-  \begin{itemize}
-  \item $H_-(A)=H_A$ for objects $A\in\cat A$ and
-  \item for maps $A\xrightarrow{f}A'$, $H_-(f)$ is the natural transformation defined for $B\in \cat A$ by
-    \begin{align*}
-      H_A(B)=\cat A(B,A)&\to H_{A'}(B)=\cat A(B,A')\\
-      g&\mapsto  f\of g\\
-    \end{align*}
-  \end{itemize}
-  \vspace{-0.8cm}
-  \includegraphics{nattrans.pdf}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Yoneda Lemma and Yoneda Principle}
-  \begin{lemma}[Yoneda]
-    Let $\cat A$ be locally small. Then
-    \begin{align*}
-      [\cat A^\op,\Set](H_A,F)\iso F(A)
-    \end{align*}
-    naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$.
-  \end{lemma}
-  \begin{theorem}
-    The Yoneda embedding $H_-\from \cat A\to[\cat A^\op,\Set]$ is full and faithful.
-  \end{theorem}
-  \begin{corollary}[Yoneda Principle]
-    Let $\cat A$ be locally small and $A,A'\in\cat A$. Then
-    \begin{align*}
-      H_A\iso H_{A'} \iff A\iso A'
-    \end{align*}
-  \end{corollary}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Cartesian Closedness of Presheaf Category}
-  \begin{lemma}
-    For any small category $\cat A$, any functor $F\from\cat A^\op\to\Set$ is a colimit of representable functors,
-    \begin{align*}
-      \lim_{\to I\in\mathbf I} H_{\pi(I)}\iso F
-    \end{align*}
-    for some small category $\mathbf I$ and functor $\pi\from \mathbf I\to \cat A$.
-  \end{lemma}
-  \begin{lemma}
-    For any $A\in\cat A$, $H_A$ preserves colimits. In particular $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\to\mathbf I}(H_A\of D)$$ 
-    for any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$.
-  \end{lemma}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Cartesian Closedness of Presheaf Category (Cont.)}
-  \begin{lemma}
-    Let $F\from \cat A\to\Set$ be a presheaf. The functor $-\times F\from [\cat A^\op,\Set]\to[\cat A^\op,\Set]$ preserves colomits. That is:
-    \begin{align*}
-      \lim_{\to \mathbf I} (D\times F)\iso (\lim_{\to \mathbf I} D)\times F
-    \end{align*}
-    for any small category $\mathbf I$ and functor $D\from\mathbf I\to [\cat A^\op,\Set]$.
-  \end{lemma}
-
-  \begin{theorem}
-    For any small category $\cat A$, the category of presheaves $\cat A^\op,\Set$ is cartesian closed.
-  \end{theorem}
-\end{frame}
-\end{document}
-
-% vim mode line:
-% vim: ts=2 sts=2 sw=2 expandtab
+\input{../../headers/definitions}

### 2017-07-12 sheets/talk-10/sheet-10.tex
Completed solutions.

@@
 
+\begin{answer}
+  \begin{itemize}
+  \item It holds
+    \begin{align*}
+      \cat A(X,(A\times B)^C&\iso\cat A(X\times C,A\times B)&\cat A\text{ cartesian closed}\\
+                            &\iso\cat A(X\times C,A)\times \cat A(X\times C,B)&\text{dual of \ref{ex:det}-\ref{ex:col}}\\
+                            &\iso\cat A(X,A^C)\times \cat A(X,B^C)&\cat A\text{ cartesian closed}\\
+                            &\iso\cat A(X,A^C\times B^C)&\text{dual of \ref{ex:det}-\ref{ex:col}}
+    \end{align*}
+    naturally in $X\in\cat A$. Hence, $(A\times B)^C\iso A^C\times B^C$.
+
+    To see that the first isomorphsim is natural in $X$, suppose an arrow $f\from Y\to X$. Naturality means that for any $g\from X\to (A\times B)^C$
+    \begin{align*}
+      \overline{g\of f}=\epsilon\of ((g\of f)\times 1_{A\times B})=\epsilon\of ((g\times 1_{A\times B})\of(f\times 1_{A\times B})=\overline g\of (f\times 1).
+    \end{align*}
+    The other isomorphisms can be equally simply shown to be natural in $X$.
+  \item It holds
+    \begin{align*}
+      \cat A(X,A^{B+C}&\iso\cat A(X\times (B+C),A)&\cat A\text{ cartesian closed}\\
+                      &\iso\cat A((X\times B)+(X\times C),A)&\text{``distributivity''-lemma}\\
+                      &\iso\cat A(X\times B,A)\times\cat A(X\times C, A)&\text{\ref{ex:det}-\ref{ex:col}}\\
+                      &\iso\cat A(X,A^B)\times\cat A(X, A^C)&\cat A\text{ cartesian closed}\\
+                      &\iso\cat A(X,A^B\times A^C)&\text{dual of \ref{ex:det}-\ref{ex:col}}
+    \end{align*}
+    naturally in $X\in\cat A$. Hence, $A^{B+C}\iso A^B\times A^C$.
+  \end{itemize}
+\end{answer}
+
 \begin{exercise}
@@
     \begin{align*}
-      J_{C,C'}\from\cat C(C,C)\to \cat D(J(C),J(C'))\\
+      J_{C,C'}&\from\cat C(C,C)\to \cat D(J(C),J(C'))\\
       f&\mapsto J(f)
@@
     To show that $J\of-$ is full and faithful, let $F,G\from\cat B\to\cat C$ be natural transformations and $\beta\from J\of F\to J\of G$ be a natural transformation.
-    Define, a natural transformation $\alpha\from F\to G$ by 
+    Define a natural transformation $\alpha\from F\to G$ by 
     \begin{align*}
-      \alpha_B=J_{F(A),G(A)}^{-1}(\beta_B)
+      \alpha_B=J_{F(B),G(B)}^{-1}(\beta_B)
     \end{align*}
@@
 \begin{exercise}
+  \label{ex:det}
   Fill in some of the missing details in the proof that for any small category $\cat A$ the presheaf category $[A^\op,\Set]$ is cartesian closed:
   \begin{enumerate}
-  \item Show that for any $A\in\cat A$, $H_A$ maps all colimits to limits such that $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\leftarrow\mathbf I}(H_A\of D)$$ for all small categories $\mathbf I$ and diagrams $D\from \mathbf I\to\cat A$.
+  \item \label{ex:col}Show that for any $A\in\cat A$, $H_A$ maps all colimits to limits such that $$H_A(\lim_{\to\mathbf I} D)\iso \lim_{\leftarrow\mathbf I}(H_A\of D)$$ for all small categories $\mathbf I$ and diagrams $D\from \mathbf I\to\cat A$.
 What is the dual statement?
@@
 
+\begin{answer}
+  \begin{enumerate}
+  \item
+    By the hint it suffices that coproducts and coequalizers are mapped to the corresponding products and equalizers, respectively (because each colimit can be obtained from coproducts and coequalizers).
+    \begin{itemize}
+    \item Coproducts are mapped to products: Let $(B_i)_{i\in I}$ be an arbitrary family of objects and suppose $B=\sum\limits_{i\in I}B_i$ exists.
+Then for any object $A$ and family of maps $(f_i\from B_i\to A)_{i\in I}$ there is a unique $\bar f\from B\to A$ s.t.
+$f_i=\bar f\of p_i$ for all $i\in I$ because of the universality property of sums.
+Hence there is a bijection $\prod\limits_{i\in I} \cat A(B_i,A)\iso \cat A(B,A)$, i.e.
+$\prod\limits_{i\in I}H_A(B_i)\iso H_A(\sum\limits_{i\in I}B_i)$.
+In particular $H_A(0)=1$ if $\cat A$ has an initial object $0$.
+    \item Coequalizers are mapped to equalizers: Let $e$ be the equalizer of $f$ and $g$:
+      \begin{align*}
+        \xymatrix{
+  A \ar@/^0.3pc/@[][r]^{f} \ar@/_0.3pc/@[][r]_{g} & B  \ar[r]^{e} & E
+                                                       } 
+      \end{align*}
+      i.e. $e\of f=e\of g$. We are going to show that $H_C(e)$ is the equalizer of $H_C(f)$ and $H_C(g)$ for any $C\in\cat A$.
+
+      \begin{align*}
+        \xymatrix{
+  H_C(A) & H_C(B) \ar@/^0.3pc/@[][l]^{H_C(f)} \ar@/_0.3pc/@[][l]_{H_C(g)}  & H_C(E) \ar[l]^{H_C(e)}
+                                                       } 
+      \end{align*}
+
+      Let $C$ be arbitrary. It holds $H_C(f)\of H_C(e)=H_C(g)\of H_C(e)$ since for any $h\from E\to C$, 
+      \begin{align*}
+        H_C(f)\of H_C(e)(h)=h\of e\of f=h\of e\of g=H_C(g)\of H_C(e)(h).
+      \end{align*}
+      Next, let $u\from S\to H_C(B)$ be another fork, i.e.
+for any $s\in S$, $H_C(f)\of u (s)=u(s)\of f=u(s)\of g=H_C(g)\of u(s)$.
+Hence, by the universality property, for any $s\in S$ there is a unique $\overline{u(s)}$ s.t.
+$u(s)=\overline{u(s)}\of e$.
+This immediately implies that there is a unique $\overline u\from S\to H_C(E)$ such that $u=H_C(e)\of \overline u$ (namely $\overline u(s)=\overline{u(s)}$).
+    \end{itemize}
+    The dual statement is that for any $A\in\cat A$, small category $\mathbf I$ and diagram $D\from \mathbf I\to\cat A$,
+    $$\cat A(A,\lim_{\leftarrow I\in\mathbf I} D(I))\iso \lim_{\leftarrow I\in\mathbf I}(\cat A(A, D(I)))$$
+    i.e. that the ``covariant representable functor preserves limits''.
+  \item Let
+    \begin{align*}
+      (D(I)\xrightarrow{\alpha_I} \lim_{\to I\in\mathbf I} D(I))_{I\in\mathbf I}
+    \end{align*}
+    be the cocone constituting the colimit. By applying the functor $-\times F$ we get a cocone 
+    \begin{align*}
+      (D(I)\times F\xrightarrow{\alpha_I\times F} (\lim_{\to I\in\mathbf I} D(I))\times F)_{I\in\mathbf I}
+    \end{align*} 
+    since functors obviously preserve cocones. Now, due to the hint $[\cat A^\op,\Set]$ has all limits. Therefore, there is a unique natural transformation 
+    \begin{align*}
+      (\lim_{\to I\in\mathbf I} D(I)\times F)\xrightarrow{\overline\alpha} (\lim_{\to I\in\mathbf I} D(I))\times F
+    \end{align*}
+    such that the corresponding arrows commute. We are going to show that $\alpha$ is an isomorphism. By exercise Basis 3.5 it suffices to show that each component
+    \begin{align*}
+      (\lim_{\to I\in\mathbf I} D(I)\times F)(A)\xrightarrow{\overline\alpha_A} ((\lim_{\to I\in\mathbf I} D(I))\times F)(A)
+    \end{align*}
+ ($A\in\cat A$) is an isomorphism in $\Set$. 
+    This can be shown with the Yoneda principle.
+
+    Let $S\in\Set$. Then we have
+    \begin{align*}
+      \Set((\lim_{\to I\in\mathbf I} D(I)\times F)(A), S)&\iso \Set((\lim_{\to I\in\mathbf I} D(I)(A)\times F(A)), S)&\text{limits are computed pointwise}\\
+      &\iso \lim_{\leftarrow I\in\mathbf I} \Set((D(I)(A)\times F(A)), S)&\text{\ref{ex:det}-\ref{ex:col}}\\
+      &\iso \lim_{\leftarrow I\in\mathbf I} \Set(D(I)(A), S^{F(A)})&\Set \text{ is cartesian closed}\\
+      &\iso \Set(\lim_{\to I\in\mathbf I} D(I)(A), S^{F(A)})&\text{\ref{ex:det}-\ref{ex:col}}\\
+      &\iso \Set((\lim_{\to I\in\mathbf I} D(I)(A))\times F(A), S)&\Set \text{ is cartesian closed}\\
+      &\iso \Set(((\lim_{\to I\in\mathbf I} D(I))\times F)(A), S)&\text{limits are computed pointwise}
+    \end{align*}
+    natural in $S$.
+  \end{enumerate}
+\end{answer}
 \end{document}

### 2017-07-13 sheets/talk-11/sheet-11.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Talk 11}{Dependent Types}
\author{Nikita Ziuzin}

\begin{hint}
  Read Martin Hofmann's Syntax and Semantics of Dependent Types and section $1$
  in Simon Huber's A Model of Type Theory in Cubical Sets.
\end{hint}

\begin{exercise}
  Read Hofmann's section on type for natural numbers (section 2.1.3).

  Familiarize yourself with identity types (ibid., section 2.1.5)

  By analogy to the type of natural numbers define the rules for a list type
  former which to any type $A$ associates a type $List(A)$ consisting of finite
  sequences of elements of $A$.

  \emph{Hint}: think of lists as inductively
  generated from the empty list by successive additions of elements of $A$
  ("cons").

  Define a length function of type $List(A) \to \mathbb{N}$ and
  define a type $Vec_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
  using lists, the identity type and the $\Sigma$-type.
\end{exercise}

\begin{exercise}
  After reading the material for the previous exercise, read a section on
  universes in Hofmann's notes (2.1.6).

  Give the rules for universe $U$ containing a code for $\hat{0}$ for the empty
  type and a code $\hat{1}$ for the unit type $1$. Show that in a type theory
  which supports natural numbers, this universe, and the empty type itself the
  following type in the empty context is inhabited
  \[
    \diamond \vdash Id_\mathbb{N}(0, Suc(0)) \to 0 \text{ type}
  \]
  corresponding to Peano's fourth axiom $1 \neq 0$.

  \emph{Hint}: define using $R^\mathbb{N}$ a function $f \from \mathbb{N} \to
  U$ such that $\diamond \vdash f 0 = \hat{1}: U$ and $\diamond \vdash
  f(Suc(0)) = \hat{0} : U$
\end{exercise}

\begin{exercise}
  Recall the notation for the category of elements of a presheaf $\Gamma$:
  $\int_\cat{C}\Gamma$ Check that substitution of $A\from
  (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from \Delta \to \Gamma$
  corresponds to precomposing with $\int_\cat{C} \sigma : \int_\cat{C} \Delta
  \to \int_\cat{C} \Gamma$ induced by $\sigma$.

  Then verify that this construction imposes a functor $\int_\cat{C} \from
  Psh(\cat{C}) \to \CAT$.
\end{exercise}
\begin{answer}
\end{answer}

\begin{exercise}
  Define dependent sum using presheaf model notation

  \emph{Hint}: Read section $1.2$ in Huber's thesis to recall presheaf model
  construction. Try to come up with your definition before checking the answer
  given in the section.
\end{exercise}
\begin{answer}
  Section $1.2.2$.
\end{answer}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-07-14 sheets/talk-11/sheet-11.tex

@@
 \begin{hint}
-  Read Martin Hofmann's Syntax and Semantics of Dependent Types and section $1$
-  in Simon Huber's A Model of Type Theory in Cubical Sets.
+  Read sections $2.1.3, 2.1.5, 2.1.6$ in Martin Hofmann's Syntax and Semantics
+  of Dependent Types and section $1$ in Simon Huber's A Model of Type Theory in
+  Cubical Sets.
 \end{hint}
@@
 \begin{exercise}
-  Read Hofmann's section on type for natural numbers (section 2.1.3).
+  Describe the elements of $Ty_F(\Gamma)$ and $Ter_F(\Gamma, A)$.
+\end{exercise}
+\begin{answer}
+  $Ty_F(\Gamma)$ is the first projection in $F(\Gamma)$, that is a set with some
+  objects (possibly empty).
+  \[
+    Ty_F(\Gamma) := \{A_\Gamma, B_\Gamma, \dots\}
+  \]
+  $Ter_F(\Gamma, A)$ is then an element of the second projection of $F(\Gamma)$,
+  that is a member of the family of sets indexed over $Ty_F(\Gamma)$, which has
+  A as its index and also may be empty.
+  \[
+    Ter_F(\Gamma, A) := \{a_A, b_A, \dots\}_A
+  \]
+\end{answer}
+
+\begin{exercise}
+  Check that the following square is a pullback in Categories with Families.
+  \[
+    \begin{tikzcd}
+      \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
+      \Delta \arrow{r}{\sigma} & \Gamma
+    \end{tikzcd}
+  \]
+  where $\Delta.A \vdash q : A p_\Delta$
+\end{exercise}
+
+\begin{exercise}
+  Recall the notation for the category of elements of a presheaf $\Gamma$
+  (section $1.2$ p. $16$ in Huber's work): $\int_\cat{C}\Gamma$ Check that
+  substitution of $A\from (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from
+  \Delta \to \Gamma$ corresponds to precomposing with $\int_\cat{C} \sigma :
+  \int_\cat{C} \Delta \to \int_\cat{C} \Gamma$ induced by $\sigma$.
+
+  Then verify that this construction imposes a functor $\int_\cat{C} \from
+  Psh(\cat{C}) \to \CAT$.
+\end{exercise}
+\begin{answer}
+\end{answer}
+
+\begin{exercise}
+  From the slides we know that If $B$ in $\Pi x: A. B$ does not depend on $A$,
+  then $\Pi x: A. B$ corresponds to the ordinary non-dependent function type:
+  \begin{equation*}
+    A \to B \vcentcolon= \Pi x: A. B
+  \end{equation*}
+
+  Can you give a term of the type $A \to A$ using the presheaf model?
+
+  \emph{Hint}: Read section $1.2.1$ in Huber's work for the general case.
+
+\end{exercise}
+
+\begin{exercise}
+  Suppose we have two independent of each other types $\Gamma \vdash A \text{
+  type}$ and $\Gamma \vdash B \text{ type}$.
+  Then we can define a product type in our presheaf model in the following way:
+  \[
+    (A \times B)\rho := \{(a,b) | a \in A\rho \land b \in B\rho\}
+  \]
+
+  Now suppose that we have dependent types $\Gamma \vdash C \text{ type}$ and
+  $\Gamma, C \vdash D \text{ type}$.
+
+  How would you define in the presheaf model a product type that involves a
+  dependent type (i.e. the dependent sum type)?
+
+  \[
+    (\Sigma AB)\rho := ???
+  \]
+
+  \emph{Hint}: The definition is \emph{almost} identical to the one given
+  above. But what do you need to change? To recall the presheaf model, read
+  section $1.2$ in Huber's work.
+\end{exercise}
+\begin{answer}
+  Section $1.2.2$.
+\end{answer}
+
+\begin{exercise}
+  Read Hofmann's section on type for the natural numbers (section 2.1.3).
 
@@
   Define a length function of type $List(A) \to \mathbb{N}$ and
-  define a type $Vec_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
+  define a type $\mathit{Vec_A}(n)$ of lists of length $n$ for each $n : \mathbb{N}$
   using lists, the identity type and the $\Sigma$-type.
@@
 
-\begin{exercise}
-  Recall the notation for the category of elements of a presheaf $\Gamma$:
-  $\int_\cat{C}\Gamma$ Check that substitution of $A\from
-  (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from \Delta \to \Gamma$
-  corresponds to precomposing with $\int_\cat{C} \sigma : \int_\cat{C} \Delta
-  \to \int_\cat{C} \Gamma$ induced by $\sigma$.
-
-  Then verify that this construction imposes a functor $\int_\cat{C} \from
-  Psh(\cat{C}) \to \CAT$.
-\end{exercise}
-\begin{answer}
-\end{answer}
-
-\begin{exercise}
-  Define dependent sum using presheaf model notation
-
-  \emph{Hint}: Read section $1.2$ in Huber's thesis to recall presheaf model
-  construction. Try to come up with your definition before checking the answer
-  given in the section.
-\end{exercise}
-\begin{answer}
-  Section $1.2.2$.
-\end{answer}
-
 \end{document}

### 2017-07-14 sheets/talk-11/sheet-11.tex
Made formulations precise and fixed typographic issues

--- 2017-07-13 sheets/talk-11/sheet-11.tex
@@
 \begin{hint}
-  Read Martin Hofmann's Syntax and Semantics of Dependent Types and section $1$
-  in Simon Huber's A Model of Type Theory in Cubical Sets.
+  Read sections $2.1.3, 2.1.5, 2.1.6$ in Martin Hofmann's \href{http://www.irif.fr/~mellies/mpri/mpri-ens/articles/hofmann-syntax-and-semantics-of-dependent-types.pdf}{Syntax and Semantics of Dependent Types} and section $1$ in Simon Huber's \href{http://www.cse.chalmers.se/~simonhu/misc/lic.pdf}{A Model of Type Theory in Cubical Sets}.
 \end{hint}
@@
 \begin{exercise}
-  Read Hofmann's section on type for natural numbers (section 2.1.3).
+  Define the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for the presheaf model.
+\end{exercise}
+\begin{answer}
+  $\mathit{Ty}_F(\Gamma)$ is the first projection in $F(\Gamma)$, that is a set with some
+  objects (possibly empty).
+  \[
+    \mathit{Ty}_F(\Gamma) := \{A_\Gamma, B_\Gamma, \dots\}
+  \]
+  $\mathit{Ter}_F(\Gamma, A)$ is then an element of the second projection of $F(\Gamma)$,
+  that is a member of the family of sets indexed over $\mathit{Ty}_F(\Gamma)$, which has
+  A as its index and also may be empty.
+  \[
+    \mathit{Ter}_F(\Gamma, A) := \{a_A, b_A, \dots\}_A
+  \]
+\end{answer}
 
-  Familiarize yourself with identity types (ibid., section 2.1.5)
+\begin{exercise}
+  Let $(\cat C, \mathcal F)$ be a category with families.
+  Check that the following square is a pullback in $\cat{C}$:
+  \[
+    \begin{tikzcd}
+      \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
+      \Delta \arrow{r}{\sigma} & \Gamma
+    \end{tikzcd}
+  \]
+  where $\Delta.A \vdash q : A p_\Delta$
+\end{exercise}
+
+\begin{exercise}
+  Recall the notation $\int_\cat{C}\Gamma$ for the category of elements of a presheaf $\Gamma$ (section $1.2$ p.$16$ in Huber's thesis).
+  Check that substitution in $A\from (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from \Delta \to \Gamma$ corresponds to precomposing with $\int_\cat{C} \sigma : \int_\cat{C} \Delta \to \int_\cat{C} \Gamma$ induced by $\sigma$.
+
+  Then verify that this construction imposes a functor $\int_\cat{C} \from \mathit{Psh}(\cat{C}) \to \CAT$.
+\end{exercise}
+\begin{answer}
+\end{answer}
+
+\begin{exercise}
+  From the slides we know that if $\Gamma \vdash A, B ~\text{type}$, i.e. $B$ does not depend on $A$,
+  then $\Pi x: A. B$ corresponds to the ordinary non-dependent function type:
+  \begin{equation*}
+    A \to B \vcentcolon= \Pi x: A. B
+  \end{equation*}
+
+  Give a term of type $A \to A$, and the elements in $\mathit{Ty}(\Gamma)$ and $\mathit{Ter}(\Gamma, A \to A)$ corresponding to the type and the term.
+
+  \emph{Hint}: Read section $1.2.1$ in Huber's thesis for the general case.
+
+\end{exercise}
+
+\begin{exercise}
+  Suppose we have two types $\Gamma \vdash A,B \text{ type}$ each not depending on the other.
+  Then we can define a product type in the presheaf model in the following way:
+  \[
+    (A \times B)\rho := \{(a,b) | a \in A\rho \land b \in B\rho\}
+  \]
+
+  Now suppose that we have dependent types $\Gamma \vdash C \text{ type}$ and
+  $\Gamma, C \vdash D \text{ type}$.
+
+  Define the product type that involves a
+  dependent type (i.e. the dependent sum type) in the presheaf model.
+
+  \[
+    (\Sigma AB)\rho := ???
+  \]
+
+  \emph{Hint}: The definition is \emph{almost} identical to the one given
+  above. But what do you need to change? To recall the presheaf model, read
+  section $1.2$ in Huber's thesis.
+\end{exercise}
+\begin{answer}
+  Section $1.2.2$.
+\end{answer}
+
+\begin{exercise}
+  Read Hofmann's section on the type of natural numbers (section 2.1.3) and familiarize yourself with identity types (ibid., section 2.1.5)
 
   By analogy to the type of natural numbers define the rules for a list type
-  former which to any type $A$ associates a type $List(A)$ consisting of finite
+  former which to any type $A$ associates a type $\mathit{List}(A)$ consisting of finite
   sequences of elements of $A$.
@@
 
-  Define a length function of type $List(A) \to \mathbb{N}$ and
-  define a type $Vec_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
+  Define a length function of type $\mathit{List}(A) \to \mathbb{N}$ and
+  define a type $\mathit{Vec}_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
   using lists, the identity type and the $\Sigma$-type.
@@
 \begin{exercise}
-  After reading the material for the previous exercise, read a section on
-  universes in Hofmann's notes (2.1.6).
+  Read the section on universes in Hofmann's notes (2.1.6).
 

### 2017-07-14 sheets/talk-11/sheet-11.tex
Update last exercise
Adjust line lengths

@@
 \begin{exercise}
-  Define the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for the presheaf model.
+  Define the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
+  the presheaf model.
 \end{exercise}
 \begin{answer}
-  $\mathit{Ty}_F(\Gamma)$ is the first projection in $F(\Gamma)$, that is a set with some
-  objects (possibly empty).
+  $\mathit{Ty}_F(\Gamma)$ is the first projection in $F(\Gamma)$, that is a set
+  with some objects (possibly empty).
   \[
@@
   \]
-  $\mathit{Ter}_F(\Gamma, A)$ is then an element of the second projection of $F(\Gamma)$,
-  that is a member of the family of sets indexed over $\mathit{Ty}_F(\Gamma)$, which has
-  A as its index and also may be empty.
+  $\mathit{Ter}_F(\Gamma, A)$ is then an element of the second projection of
+  $F(\Gamma)$, that is a member of the family of sets indexed over
+  $\mathit{Ty}_F(\Gamma)$, which has A as its index and also may be empty.
   \[
@@
 \begin{exercise}
-  Let $(\cat C, \mathcal F)$ be a category with families.
-  Check that the following square is a pullback in $\cat{C}$:
+  Let $(\cat C, \mathcal F)$ be a category with families.  Check that the
+  following square is a pullback in $\cat{C}$:
   \[
@@
 \begin{exercise}
-  Recall the notation $\int_\cat{C}\Gamma$ for the category of elements of a presheaf $\Gamma$ (section $1.2$ p.$16$ in Huber's thesis).
-  Check that substitution in $A\from (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from \Delta \to \Gamma$ corresponds to precomposing with $\int_\cat{C} \sigma : \int_\cat{C} \Delta \to \int_\cat{C} \Gamma$ induced by $\sigma$.
+  Recall the notation $\int_\cat{C}\Gamma$ for the category of elements of a
+  presheaf $\Gamma$ (section $1.2$ p.$16$ in Huber's thesis).  Check that
+  substitution in $A\from (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from
+  \Delta \to \Gamma$ corresponds to precomposing with $\int_\cat{C} \sigma :
+  \int_\cat{C} \Delta \to \int_\cat{C} \Gamma$ induced by $\sigma$.
 
-  Then verify that this construction imposes a functor $\int_\cat{C} \from \mathit{Psh}(\cat{C}) \to \CAT$.
+  Then verify that this construction imposes a functor $\int_\cat{C} \from
+  \mathit{Psh}(\cat{C}) \to \CAT$.
 \end{exercise}
@@
 \begin{exercise}
-  From the slides we know that if $\Gamma \vdash A, B ~\text{type}$, i.e. $B$ does not depend on $A$,
-  then $\Pi x: A. B$ corresponds to the ordinary non-dependent function type:
+  From the slides we know that if $\Gamma \vdash A, B ~\text{type}$, i.e. $B$
+  does not depend on $A$, then $\Pi x: A. B$ corresponds to the ordinary
+  non-dependent function type:
   \begin{equation*}
@@
 
-  Give a term of type $A \to A$, and the elements in $\mathit{Ty}(\Gamma)$ and $\mathit{Ter}(\Gamma, A \to A)$ corresponding to the type and the term.
+  Give a term of type $A \to A$, and the elements in $\mathit{Ty}(\Gamma)$ and
+  $\mathit{Ter}(\Gamma, A \to A)$ corresponding to the type and the term.
 
@@
 \begin{exercise}
-  Suppose we have two types $\Gamma \vdash A,B \text{ type}$ each not depending on the other.
-  Then we can define a product type in the presheaf model in the following way:
+  Suppose we have two types $\Gamma \vdash A,B \text{ type}$ each not depending
+  on the other.  Then we can define a product type in the presheaf model in the
+  following way:
   \[
@@
 
-  Define the product type that involves a
-  dependent type (i.e. the dependent sum type) in the presheaf model.
+  Define the product type that involves a dependent type (i.e. the dependent
+  sum type) in the presheaf model.
 
@@
 \begin{exercise}
-  Read Hofmann's section on the type of natural numbers (section 2.1.3) and familiarize yourself with identity types (ibid., section 2.1.5)
+  Read Hofmann's section on the type of natural numbers (section 2.1.3) and
+  familiarize yourself with identity types (ibid., section 2.1.5)
 
   By analogy to the type of natural numbers define the rules for a list type
-  former which to any type $A$ associates a type $\mathit{List}(A)$ consisting of finite
-  sequences of elements of $A$.
+  former which to any type $A$ associates a type $\mathit{List}(A)$ consisting
+  of finite sequences of elements of $A$.
 
-  \emph{Hint}: think of lists as inductively
-  generated from the empty list by successive additions of elements of $A$
-  ("cons").
+  \emph{Hint}: think of lists as inductively generated from the empty list by
+  successive additions of elements of $A$ ("cons").
 
-  Define a length function of type $\mathit{List}(A) \to \mathbb{N}$ and
-  define a type $\mathit{Vec}_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
+  Define a length function of type $\mathit{List}(A) \to \mathbb{N}$ and define
+  a type $\mathit{Vec}_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
   using lists, the identity type and the $\Sigma$-type.
@@
 \begin{exercise}
-  Read the section on universes in Hofmann's notes (2.1.6).
+  Read the section on universes in Hofmann's notes (2.1.6) as well as the
+  material for the previous exercise.
 

### 2017-07-15 notes/notes-2/notes-2.tex
Started notes for talk 2.

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}

% \beamertemplatenavigationsymbolsempty
% \setbeamertemplate{footline}{
% 	\raisebox{0.5em}{\makebox[\paperwidth-0.5em][r]{
% 		\color{gray}\scriptsize\insertframenumber}}}

\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{cd, shapes, tikzmark}
\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}

\title{Presheaves, Representables and\\the Yoneda Lemma}
\author{Dominik Wagner}
%\date{June 7, 2017}

\begin{document}

\begin{definition}
  Let $\cat A$ and $\cat B$ be categories. A \emph{functor} $\from \cat A\to\cat B$ consists of
    \begin{itemize}
    \item a function
      \begin{align*}
        \ob(\cat A)\to\ob(\cat B),
      \end{align*}
      written as $A\mapsto F(A)$ and
    \item for each $A,A'\in\ob(\cat A)$, a function
      \begin{align*}
        \cat A(A,A')\to\cat B(F(A),F(A')),
      \end{align*}
      written as $f\mapsto F(f)$
    \end{itemize}
    such that the following axioms are satisfies:
    \begin{itemize}
    \item $F(1_A)=1_{F(A)}$ whenever $A\in\cat A$ and
    \item $F(f'\of f)=F(f')\of F(f)$ whenever $A\xrightarrow{f}A'\xrightarrow{f'}A''$ in $\cat A$.
    \end{itemize}
\end{definition}

Next, we consider a further example of a $\Set$-valued functor, which unlike the previous ones is not forgetful.
\begin{example}
  \label{ex:Mset}
  Let $\cat M$ be a category with a single object $M$, i.e. a monoid. Then, a functor $F\from\cat M\to\Set$ is a set $S=F(M)$ together with functions $F(m)\from S\to S$ for $m\from M\to M$ in $\cat M$. Hence, we can introduce the short-hand notation
  \begin{align*}
    m\cdot s=F(m)(s).
  \end{align*}
  The functorality axioms for $F$ state that
  \begin{align*}
    1_M\cdot s&=F(1_M)(s)=1_S(s)=s\\
    (m\of m')\cdot s&=F(m\of m')(s)=(F(m)\of F(m'))(s)=F(m)(F(m')(s))=m\cdot (m'\cdot s)
  \end{align*}
  for all $m,m'\from M\to M$ and $s\in S$.

  As a consequence, a functor from a one-object category, which correspond to a monoid $M$, to $\Set$ is essentially a \emph{left $M$-set}. (Recall that a left $M$-set for a monoid $M$ is a set $S$ together with a function
  \begin{align*}
    \cdot\from M\times S&\to S\\
    (m,s)&\mapsto m\cdot s
  \end{align*}
  such that $(m\of m')\cdot s=m\cdot (m'\cdot s)$ and $1_M\cdot s=s$ for all $m,m'\in M$ and $s\in S$.)  
\end{example}

\section{Functors between Preorders}
Next, suppose that $\cat P$ and $\cat Q$ are preorders regarded as categories. A functor $F\from\cat P\to\cat Q$ is a mapping $\ob(\cat P)\to\ob(\cat Q)$ of the objects together with a function
\begin{align*}
  \cat P(P,P')\to\cat Q(F(P),F(P'))
\end{align*}
for each $P,P'\in\ob(\cat P)$. In the special case of preorders the latter function states that
\begin{align*}
  P\leq_\cat P P'\rightarrow F(P)\leq_\cat Q F(P')
\end{align*}
due to the fact that for each pair of objects in a preorder there is either exactly one arrow between them (denoting that the first is smaller than the second) or none at all (denoting incomperatibility). Hence, functors between correspond to monotonically (increasing) functions.

A very natural question to ask is whether there is also a general category theoretic analogue to monotonically decreasing functions, i.e. functions $F\from\ob(\cat P)\to\ob(\cat Q)$ between preorders $\cat P$ and $\cat Q$ such that
\begin{align*}
  P\leq_\cat P P'\rightarrow F(P)\geq_\cat Q F(P')
\end{align*}
for each $P,P'\in\ob(\cat P)$. Again, this exactly the same as to demand that for each $P,P'\in\ob(\cat P)$ there is a function
\begin{align*}
  \cat P(P,P')\to\cat Q(F(P'),F(P)).
\end{align*}
Since there is the unique arrow in $\cat (P,P')$ if and only if $\cat P^\op(P',P)$ contains the unique arrow we see that this amounts to the same as having a function
\begin{align*}
  \cat P^\op(P',P)\to\cat Q(F(P'),F(P)),
\end{align*}
that is we need to reverse the direction of the arrows w.r.t. $\cat P$. Due to the fact that the additional functorality axioms are trivially satisfied for preorders, we see that monotone decreasing functors are exactly the same as functors
\begin{align*}
  F\from\cat P^\op\to\cat Q
\end{align*}
for the corresponding categories $\cat P$ and $\cat Q$.

Such ``operations'' that are essentially functors but reverse the direction of the arrows are going to play a prominent role in later chapters. Therefore, they have a special name:
\begin{definition}
  Let $\cat A$ and $\cat B$ be categories. A \emph{contravariant functor} from $\cat A$ to $\cat B$ is a functor $\cat A^\op\to\cat B$.
\end{definition}
\section{Dual Vector Spaces and Contravariant Representable Functors}
Next, we are going to look at the generalization of a construction from linear algebra that is used in the study of bilinear mappings amongst others. The general idea is that we analyze objects by analyzing the mappings into them.

Let $\mathbb K$ be a field. Given any two $\mathbb K$-vector spaces $V$ and $W$ there is the vector space 
\begin{align*}
  \Hom(V,W)=\{f\from V\to W\mid f \text{ linear}\}
\end{align*}
of linear mappings between $V$ and $W$ where the vector space operations addition and scalar multiplication are defined pointwise. Therefore, for a fixed $\mathbb K$-vector space $W$, we can define a mapping
\begin{align*}
  V\mapsto H_W(V)=\Hom(V,W).
\end{align*}
It seems reasonable that we can extend this to a functor $H_W\from\Vect_{\mathbb K}\to\Vect_{\mathbb K}$. So let us check the details: We have to define its action on linear maps $f\from V\to V'$, that is we have to define a linear mapping 
\begin{align*}
  H_W(f)\from H_W(V)=\Hom(V,W)\to H_W(V')=\Hom(V',W)
\end{align*}
 that satisfies the axioms. To do so, let $g\in H_W(V)=\Hom(V,W)$ be arbitrary and we have to come up with a linear mapping $V'\to W$. This seems to be impossible: There is no canonical way to construct a linear mapping $V'\to W$ from linear mappings $f\from V\to V'$ and $g\from V\to W$. Hence, there is no reasonable way to define a functor $H_W\from\Vect_{\mathbb K}\to\Vect_{\mathbb K}$.

However, it is perfectly possible to canonically construct a linear mapping $V\to W$ from linear mappings $f\from V\to V'$ and $g\from V'\to W$ by just taking its composition $g\of f$ (since composition preserves linearity):
\begin{align*}
  \xymatrix{V' \ar[r]^{f} & V \ar[r]^{g} & W }.
\end{align*}
That is, given arbitrary $f\from V\to V'$ we can define a mapping
\begin{align*}
  H_W(f)\from H_W(V')&\to H_W(V)\\
  g&\mapsto g\of f
\end{align*}
that is moreover trivially linear. This is indeed a contravariant functor since
\begin{align*}
  H_W(1_V)=\lambda g\ldotp g\of 1_V=\lambda g\ldotp g=1_{H_W(V)}\\
  H_W(f'\of f)=\lambda g\ldotp g\of (f'\of f)=\lambda g\ldotp (g\of f')\of f=(\lambda g\ldotp g\of f)\of (\lambda g\ldotp g\of f')=H_W(f)\of H_W(f')
\end{align*}
Thus, we have defined a functor $H_W\from\Vect_{\mathbb K}^\op\to\Vect_{\mathbb K}$.

In linear algebra one frequently takes $\mathbb K$ itself for $W$ (regarded as a one-dimensional vector space over itself) and $H_{\mathbb K}(V)$ is called the dual vector space $V^*$ of $V$. It is a well known fact that the dual vector space $V^*$ reveals a lot about $V$ itself and its relation to other vector spaces. In particular, $V$ is a subspace of the dual of $V^*$.

Since this construction seems very useful indeed for vector spaces it is a natural question whether we can do something similar for arbitrary categories.
The obvious analogue to 
\begin{align*}
  H_W(V)=\Hom(V,W)=\{f\from V\to W\mid f \text{ linear}\}
\end{align*}
 is
\begin{align*}
  H_B(A)=\cat A(A,B)
\end{align*}
for $A$ and $B$ being objects in a category $\cat A$. This suggests a functor $F_A\from\cat A^\op\to\Set$. However, this is only well-defined if all collections $\cat A(A,B)$ are really sets, which is not necessarily the case for arbitrary categories. Hence, in all what follows we assume a \emph{locally small} category $\cat A$ in which all collections $\cat A(A,B)$ are really sets.

Bearing this in mind, the defintion of the functor $H_B$ for $B\in\cat A$ is a straight-forward generalization:
\begin{align*}
  H_B(A)=\cat A(A,B)
\end{align*}
for objects $A\in\cat A$ and for $f\from A'\to A$ the function $H_B(f)\from H_B(A)\to H_B(A')$ is defined by
\begin{align*}
  H_B(f)\from H_B(A)=\cat A(A,B)&\to H_B(A')=\cat A(A',B)\\
  g&\mapsto g\of f.
\end{align*}
\vspace{-0.7cm}
\begin{align*}
  \xymatrix{A' \ar[r]^{f} & A \ar[r]^{g} & B }
\end{align*}
These functors $H_A$ are called the \emph{contravariant representable functor} of $A$. Furthermore, contravariant functors to $\Set$ (like $H_A$) have a special name:
\begin{definition}
  Let $\cat A$ be a category. A \emph{presheaf} on $\cat A$ is a functor $\cat A^\op\to\Set$.
\end{definition}
In a later chapter we are going to discover that they really helps us understand the structure of locally small categories. In particular, we are going to prove that two objects $A$ and $B$ are isomorphic if and only if $H_A$ and $H_B$ are isomorphic\footnote{The notion of mappings between functors is defined in the next chapter.}, i.e. $A$ and $B$ ``look the same from all other objects''.

\begin{example}
  Similarly, as in example \ref{ex:Mset} presheaves on one-object categories (i.e. monoids $M$) correspond to \emph{right} $M$-sets, where a right $M$-set is a set $S$ together with a function
  \begin{align*}
    \cdot\from S\times M&\to S\\
    (s,m)&\mapsto s\cdot m
  \end{align*}
  such that $s\cdot (m\of m')=(s\cdot m)\cdot m'$ and $s\cdot 1_M=s$ for all $m,m'\in M$ and $s\in S$.
\end{example}

\end{document}

% vim mode line:
% vim: ts=2 sts=2 sw=2 expandtab

### 2017-07-17 sheets/talk-11/sheet-11.tex
Minor reformulations
Add first solutions to sheet 11

@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}
+\usepackage{semantic}
 
@@
 \begin{exercise}
-  Define the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
+  Give the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
   the presheaf model.
@@
 \begin{answer}
-  $\mathit{Ty}_F(\Gamma)$ is the first projection in $F(\Gamma)$, that is a set
-  with some objects (possibly empty).
   \[
-    \mathit{Ty}_F(\Gamma) := \{A_\Gamma, B_\Gamma, \dots\}
+    \mathit{Ty}_F(\Gamma) := \mathit{Psh(\Gamma)}
   \]
-  $\mathit{Ter}_F(\Gamma, A)$ is then an element of the second projection of
-  $F(\Gamma)$, that is a member of the family of sets indexed over
-  $\mathit{Ty}_F(\Gamma)$, which has A as its index and also may be empty.
   \[
-    \mathit{Ter}_F(\Gamma, A) := \{a_A, b_A, \dots\}_A
+    \mathit{Ter}_F(\Gamma, A) := \{(a\rho \in A\rho)_{I \in \cat{C}, \rho \in \Gamma(I)} | \forall f \from J \to I, a\rho f = a(\rho f) \}
   \]
@@
 \end{exercise}
+\begin{answer}
+  Let $\Delta'$ be a context and $p' \from \Delta' \to \Delta$ and $\sigma'
+  \from \Delta' \to \Gamma.A$ be two morphisms such that $\sigma \circ p' =
+  p_\Gamma \circ \sigma'$. Therefore the diagram
+  \[
+    \begin{tikzcd}
+      \Delta' \arrow[bend right=30]{ddr}{p'} \arrow[dashed]{dr}{\delta} \arrow[bend left=30]{drr}{\sigma'}&   &\\
+          & \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
+          & \Delta \arrow{r}{\sigma} & \Gamma
+    \end{tikzcd}
+  \]
+  (without $\delta$) is commutative. We must show that there exists a unique
+  $\delta \from \Delta' \to \Delta.A\sigma$ such that $(\sigma \circ p_\Delta,
+  q) \circ \delta = \sigma'$ and $p_\Delta \circ \delta = p'$.
+
+  We can decompose $\sigma'$ as $\sigma' = (p_\Gamma, q_\Gamma) \circ \sigma' =
+  (p_\Gamma \circ \sigma', q_\Gamma\sigma') = (\sigma \circ p', M)$, where $M
+  \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash A(\sigma \circ p')
+  ~\text{type}$). Let $\delta$ be morphism
+  \[
+    \delta \equiv (p', M) \from \Delta' \to \Delta.A\sigma
+  \]
+  which is suitable since $p_\Delta \circ \delta = p'$ and
+  \[
+    (\sigma \circ p_\Delta, q) \circ \delta =
+    (\sigma \circ p_\Delta \circ \delta, q\delta) =
+    (\sigma \circ p', q_\Gamma\sigma') =
+    (p_\Gamma \circ \sigma', q_\Gamma\sigma') =
+    (p_\Gamma, q_\Gamma) \circ \sigma' =
+    \sigma'
+  \]
+
+  For uniqueness assume morphism $\delta' \from \Delta' \to \Delta.A\sigma$
+  such that $(\sigma \circ p_\Delta, q) \circ \delta' = \sigma'$ and $p_\Delta
+  \circ \sigma' = p'$. As before we have $\delta' = (p_\Delta \circ \delta',
+  q\delta') = (p', N)$ where $N \equiv q\delta'$. The following equalities
+  hold:
+  \[
+    N = q\delta' = q_\Gamma(\sigma \circ p_\Delta \circ \delta', q\delta') = q_\Gamma((\sigma \circ p_\Delta, q) \circ \delta') = q_\Gamma\delta' = M
+  \]
+  and therefore $\delta' = \delta$.
+\end{answer}
 
@@
 \begin{answer}
+  Indeed, the operation of $\sigma$ on such objects is a precomposition
+  satisfying the naturality axiom:
+  \[
+    \int_\cat{C}\sigma(I, \rho) := (I, \sigma_I \rho)
+  \]
+
+  Let's check that $\int_\cat{C}$ indeed takes us into a category.
+  \begin{itemize}
+    \item objects are pairs $(I, \rho)$ where $I \in \cat{C}$ and $\rho \in \Gamma(I)$
+    \item a morphism $(J, \rho') \to (I, \rho)$ is a morphism $f \from J \to I$
+      in $\cat{C}$ such that $\rho' = \rho f$.
+  \end{itemize}
+  We need to check that we have composition and identities.
+
+  With our definition of objects, we can consider first and second projections
+  of the pairs separately. We can observe, that elements of first projection
+  are elements in $\cat{C}$ which have required identities and composition
+  under morphisms in $\cat{C}$ (which are also our morphisms). Then what
+  remains is the second projection. Identities and composition are exactly our
+  conditions on presheaves which also hold.
+
+  Now we know the operation of $\int_\cat{C}$ on objects and arrows in
+  $Psh(\cat{C})$.
 \end{answer}
@@
   \emph{Hint}: Read section $1.2.1$ in Huber's thesis for the general case.
+\end{exercise}
+\begin{answer}
+  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we the elements of $\mathit{Ty}(\Gamma)$:
+  \[
+    (A \to A)\rho = \{w | w_f \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
+  \]
+  And required equalities hold trivially for any $g \from K \to J$: $((\lambda
+  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.
 
-\end{exercise}
+  Then we give a term by the following equation:
+  \[
+    ((\lambda q) \rho)_f u = q(\rho f, u) = u
+  \]
+\end{answer}
 
@@
 \begin{answer}
-  Section $1.2.2$.
+  \[
+    (\Sigma AB)\rho := \{(a,b) | a \in A\rho \land b \in B(\rho, a)\}
+  \]
 \end{answer}
@@
 \end{exercise}
+\begin{answer}
+  First let's define the type formers:
+  \[
+    \inference[List-F]{\Gamma \vdash A ~\text{type}}{\Gamma \vdash \mathbf{List_A} ~\text{type}}\quad
+    \inference[List-I-nil]{\Gamma \vdash A ~\text{type}}{\Gamma \vdash \mathit{nil_A} : \mathbf{List_A}}\quad
+    \inference[List-I-cons]{\Gamma \vdash a : A \quad \Gamma \vdash x : \mathbf{List_A}}{\Gamma \vdash \mathit{Cons(a, x)} : \mathbf{List_A}}
+  \]
+
+  \[
+    \inference[List-E]{
+      \Gamma, l: \mathbf{List_A} \vdash B ~\text{type}\\
+      \Gamma \vdash H_{nil} : B[nil_A/l]\\
+      \Gamma, a : A, l : \mathbf{List_A}, x: B \vdash H_{cons} : B[Cons(a, l)/l]\\
+      \Gamma \vdash l : \mathbf{List_A}
+    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, l) : B[l/l]}
+  \]
+
+  \[
+    \inference[List-C-nil]{
+      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, nil) : B[nil/l]
+    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B]H_{cons}, nil) = H_{nil} : B[nil/l]}
+  \]
+
+  \[
+    \inference[List-C-cons]{
+      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(a, l)) : B[Cons(a, l)/l]
+    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(a, l)) = D}
+  \]
+  where $D = H_{cons}[a/a, l/l, R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, l)/x] : B[Cons(a, l)/l]$.
+
+  Now we can define the length function:
+  \[
+    \mathit{Length(l) \vcentcolon= R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]\mathbb{N}}(0, [a : A, l: \mathbf{List_A}, x : \mathbb{N}] Suc(x), l) : \mathbb{N}}
+  \]
+
+  And finally the $\mathit{Vec_A(n)}$ type:
+  \[
+    \mathit{Vec_A(n)} \vcentcolon= \Sigma l: \mathbf{List_A}. Id_\mathbb{N}(n, \mathit{Length}(l))
+  \]
+\end{answer}
 
@@
 \end{exercise}
+\begin{answer}
+  \[
+    \inference[U-0]{\vdash \Gamma~\text{ctxt}}{\Gamma \vdash \hat{0} : U}\quad
+    \inference[U-0-Ty]{\Gamma \vdash \hat{0}: U}{\Gamma \vdash El(\hat{0}) = 0~\text{type}}
+  \]
+
+  \[
+    \inference[U-1]{\vdash \Gamma~\text{ctxt}}{\Gamma \vdash \hat{1} : U}\quad
+    \inference[U-1-T]{\Gamma \vdash \hat{1} : U}{\Gamma \vdash El(\hat{1}) = 1~\text{type}}
+  \]
+
+  \[
+    f(n) = R^\mathbb{N}_{[n:\mathbb{N}]U}(\hat{1}, [n: \mathbb{N}, x: U]\hat{0}, n)
+  \]
+
+  Now to get an inhabitant of that type let's use $\mathit{Subst}$:
+  \[
+    P : Id(0, Suc(0)) \vdash \mathit{Subst}(P, El(f(0))) : El(f(Suc(0)))
+  \]
+  Since the inhabitant to $El(f(0)) = El(\hat{1}) = 1~\text{type}$, namely
+  $\star$, is always present by the formation rules, and $El(f(Suc(0))) =
+  0~\text{type}$ we obtain an inhabitant to the required type.
+\end{answer}
 

### 2017-07-18 papers/base-3/paper-3.tex
started with equivalences

%/papers/base-3/paper-3.tex
\def\pathToRoot{../../}\input{\pathToRoot headers/header_lecturenotes}

\newcounter{counter}[section]
\newtheorem{defn}[counter]{Definition}
\newtheorem{exmp}[counter]{Example}
\newtheorem{lemma}[counter]{Lemma}

\newcommand{\catA}[0]{\cat{A}}
\newcommand{\catB}[0]{\cat{B}}

\title{Natural transformations}
\author{Joachim Bard, Maximilian Wuttke, Nikita Ziuzin}

\begin{document}

\maketitle

\section{Equivalences of Categories}
\begin{exmp}
  \[
    \begin{tikzcd}
      \catA: & & \catB:\\
      A \arrow[loop left] & & B \arrow[bend left]{d} \arrow[loop left] \\
                          & & B' \arrow[bend left]{u} \arrow[loop left]
    \end{tikzcd}
  \]
\end{exmp}

\begin{defn}
  \label{def:eqv}
  Tow categories $\catA$ and $\catB$ are \emph{equivalent} if:
  \begin{itemize}
    \item There are functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ and
    \item there are natural transformations $\eta \from 1_{\catA} \to G \circ F$ and $\varepsilon \from F \circ G \to 1_\catB$.
  \end{itemize}
  The functors $F$ and $G$ are called \emph{equivalences}.
  We write $\catA \eqv \catB$ if $\catA$ and $\catB$ are equivalent.
\end{defn}

\begin{defn}
  \label{def:surj-on-obj}
  A functor $F \from \catA \to \catB$ is \emph{essentially surjective on objects}
  if for every object $B \in \catB$ there is an object $A \in \catA$ such that $F(A) \iso B$.
\end{defn}

\begin{lemma}
  \label{lem:equiv}
  A functor $F \from \catA \to \catB$ is an equivalence if and only if $F$ is full, faithful and surjective on objects.
\end{lemma}

\begin{exmp}
  Let $K$ be a field.
  We define the category $\catA$ such that objects are of the form $K^n$ for $n \in \NN$ and morphisms are linear functions.
  In order to show that $\catA \eqv \FDVect_K$ we have to define a functor $F \from \catA \to \FDVect_K$.
\end{exmp}

\section{Horizontal Composition}

\end{document}

% vim: ts=2 sts=2 sw=2 expandtab

### 2017-07-18 papers/base-3/paper-3.tex
First draft for writeup 3

@@
-%/papers/base-3/paper-3.tex
 \def\pathToRoot{../../}\input{\pathToRoot headers/header_lecturenotes}
+% \input{\pathToRoot headers/theoremenvts.tex}
+
+\newcommand{\catA}[0]{\cat{A}}
+\newcommand{\catB}[0]{\cat{B}}
 
@@
 \newtheorem{exmp}[counter]{Example}
-\newtheorem{lemma}[counter]{Lemma}
-
-\newcommand{\catA}[0]{\cat{A}}
-\newcommand{\catB}[0]{\cat{B}}
 
@@
 
-\section{Equivalences of Categories}
-\begin{exmp}
+\section*{Motivation, definitions and examples}
+Morphisms are maps between objects of a category, functors are maps between categories, and finally, natural transformations are maps between functors of the same domain and codomain.
+
+For example, concider a discrete category $\catA$ and an arbitrary category $\catB$.
+Now let $\parpair{\catA}{\catB}{F}{G}$ be functors from $\catA$ to $\catB$.
+Now because $\catA$ is discrete, we can ignore the morphism-mappings of $F$ and $G$ (since there are only identity morphisms in $\catA$ that are mapped to identity morphisms in $\catB$) and write $F$ and $G$ as families
+of objects in $\catB$
+$\bigl(F(A)\bigr)_{A\in\catA}$ and $\bigl(G(A)\bigr)_{A\in\catA}$.
+When we have a mapping $\alpha$ of $\catB$-morphisms
+$\bigl(F(A)\toby{\alpha_A}G(A)\bigr)_{A\in\catA}$
+this $\alpha$ could be regarded as a mapping between the functors $F$ and $G$.
+Indeed this construction is a special case of the more general definition of natural transformations.
+
+We first define what a natural transformation in general is precisely:
+\begin{defn}[Natural transformation]
+  \label{def:nat}
+  Let $\catA$ and $\catB$ be categories and $\parpair{\catA}{\catB}{F}{G}$ be two functors from $\catA$ to $\catB$.
+  Let $\alpha$ be a family of maps
+  $\bigl(F(a)\toby{\alpha_A}G(A)\bigr)_{A\in\catA}$.
+  Then we define that $\alpha$ is a \demph{natural transformation} from $F$ to $G$ iff
+  for every objects and maps $A, A' \in \catA, f \from A \to A'$, the sqare
+  \begin{equation}
+    \label{eq:nat}
+    \begin{array}{c}
+      \xymatrix{
+        F(A) \ar[r]^{F(f)} \ar[d]_{\alpha_A} & F(A') \ar[d]^{\alpha_{A'}} \\
+        G(A) \ar[r]_{G(f)} & G(A')
+      }
+    \end{array}
+  \end{equation}
+  commutes.
+  The maps $\alpha_A$ are then called the \demph{components} of $\alpha$.
+  The commuting square (\ref{eq:nat}) is also called the \demph{naturality axiom} (throught this is not a axiom in logical sense but a defined property for some $\alpha$).
+\end{defn}
+
+If $\alpha$ is a natural transformation from $F$ to $G$ we can also write this as a diagram:
+\[\xymatrix{
+  \catA \rtwocell^F_G{\alpha} & \catB
+}\]
+
+\begin{exmp}[Discreate Categories]
+  For the first example, the mapping
+  $\bigl(F(A)\toby{\alpha_A}G(A)\bigr)_{A\in\catA}$
+  the naturality axiom (\ref{eq:nat}) holds, because $\catA$ is discrete, so every morphism $f \from A \to A'$ in $\catA$ is an identy morphism (hence $A=A'$). The diagramm
   \[
-    \begin{tikzcd}
-      \catA: & & \catB:\\
-      A \arrow[loop left] & & B \arrow[bend left]{d} \arrow[loop left] \\
-                          & & B' \arrow[bend left]{u} \arrow[loop left]
-    \end{tikzcd}
-  \]
+    \xymatrix{
+      F(A) \ar[r]^{F(I)} \ar[d]_{\alpha_A} & F(A) \ar[d]^{\alpha_{A}} \\
+      G(A) \ar[r]_{G(I)} & G(A)
+  }\]
+  commutes trivially for every identity morphism $I \from A \to A$. Hence $\alpha$ is a natural transformation according to definition~\ref{def:nat}.
 \end{exmp}
 
-\begin{defn}
-  \label{def:eqv}
-  Tow categories $\catA$ and $\catB$ are \emph{equivalent} if:
-  \begin{itemize}
-    \item There are functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ and
-    \item there are natural transformations $\eta \from 1_{\catA} \to G \circ F$ and $\varepsilon \from F \circ G \to 1_\catB$.
-  \end{itemize}
-  The functors $F$ and $G$ are called \emph{equivalences}.
-  We write $\catA \eqv \catB$ if $\catA$ and $\catB$ are equivalent.
-\end{defn}
+\begin{exmp}[Determinants and commutative rings]
+  Let $n\in\nat$ and $R$ be a commutative ring.
+  We have a forgetful functor $U \from \CRing \to \Mon$ by forgetting the addition of the commutative ring.
+  Furthermore the set of all $n \times n$ matrices forms a monoid under matrix multiplication.
+  If we map $R \in \CRing$ to $R^{n \times n}$ and ring homomorphisms $f \from R \to S$ to
+  $\lambda A \in R^{n \times n}. \bigl( f(A_{i,j}) \bigr)_{1 \le i, j \le n}$,
+  i.~e. applying $f$ element-wise to the matrix $A$, we get another functor $M_n \from \CRing \to \Mon$.
+  We will now see that the \emph{determinant} is a natural transformation between this two functors $M_n$ and $U$.
 
-\begin{defn}
-  \label{def:surj-on-obj}
-  A functor $F \from \catA \to \catB$ is \emph{essentially surjective on objects}
-  if for every object $B \in \catB$ there is an object $A \in \catA$ such that $F(A) \iso B$.
-\end{defn}
-
-\begin{lemma}
-  \label{lem:equiv}
-  A functor $F \from \catA \to \catB$ is an equivalence if and only if $F$ is full, faithful and surjective on objects.
-\end{lemma}
-
-\begin{exmp}
-  Let $K$ be a field.
-  We define the category $\catA$ such that objects are of the form $K^n$ for $n \in \NN$ and morphisms are linear functions.
-  In order to show that $\catA \eqv \FDVect_K$ we have to define a functor $F \from \catA \to \FDVect_K$.
+  First we note that for every ring the determinant of a matrix over this ring is defined canonically.
+  Furthermore by definition if $A \in R^{n \times n}$ then $\det A \in R$ for every ring $R$.
+  We therefore can establish a family of mappings $\det$:
+  $\bigl(M_n(R)\toby{\det_R}U(R)\bigr)_{R\in\CRing}$, because $U(R) = R$.
+  Now, to show that $\det$ is indeed a natural transformation, we have to show that the following diagramm commutes for every ring homomorphism $f \from R \to S$:
+  \[\xymatrix{
+    M_n(R) \ar[r]^{M_n(f)} \ar[d]_{\det_R} & M_n(S) \ar[d]^{\det_S} \\
+    U(R) \ar[r]_{U(f)} & U(s)
+  }\]
+  By unfolding the definitions of $U$ and $M_n$ and applying functional extentionality, this equivalent to
+  $f(\det_R(f(A))) = \det_S (\bigl( f(A_{i,j}) \bigr)_{1 \le i, j \le n})$
+  for every $A \in R^{n \times n}$.
+  This already is a standard result of linear algebra.
+  Hence $\det$ is really a natural transformation between the functors $M_n$ and $U$. Or, written as a diagram:
+  \[\xymatrix{
+    \CRing \rtwocell<5>^{M_n}_{U}{\alpha} & \Mon
+  }\]
 \end{exmp}
-
-\section{Horizontal Composition}
 

### 2017-07-18 papers/base-3/paper-3.tex
typo

@@
 
-\begin{exmp}[Discreate Categories]
+\begin{exmp}[Discrete Categories]
   For the first example, the mapping

### 2017-07-18 sheets/talk-12/sheet-12.tex
Add exercises for talk 12

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\graphicspath{{graphics/}}

\begin{document}
\uebunghead{Talk 12}{Monoidal categories and string diagrams}
\author{Felix Rech}

\begin{exercise}
  Let $A$, $B$, $C$, $D$, be objects in a monoidal category.
  Construct a morphism
  \[
    \roundddd{\rounddd{(A \otimes I) \otimes B} \otimes C} \otimes D
    \to A \otimes \roundddd{B \otimes \rounddd{C \otimes (I \otimes D)}}.
  \]
  Can you find another?
\end{exercise}

\begin{exercise}
  Convert the following algebraic equations into graphical language.
  Which would you expect to be true in any symmetric monoidal category?
  \begin{enumerate}
    \item
      $(g \otimes 1) \of \sigma \of (f \otimes 1)
        = (f \otimes 1) \of \sigma \of (g \otimes 1)$ \\
      for $A \xrightarrow{f, g} A$.

    \item
      $h \of (1 \otimes \lambda) \of (1 \otimes (f \otimes 1)) \of (1 \otimes \lambda^{-1}) \of g
        = h \of g \of \lambda \of (f \otimes 1) \of \lambda^{-1}$ \\
      for $A \xrightarrow{g} B \otimes C$, $I \xrightarrow{f} I$ and $B \otimes C \xrightarrow{h} D$.

    \item
      $\rho \of (1 \otimes f) \of \alpha \of (\sigma \otimes 1)
        = \lambda \of (f \otimes 1) \of \alpha^{-1} \of (1 \otimes \sigma) \of \alpha$ \\
      for $A \otimes B \xrightarrow{f} f$.
  \end{enumerate}
\end{exercise}

\begin{exercise}
    Consider the following diagrams in the graphical calculus:

    \centering \includegraphics{monoidal-braided-symmetric-comparison}

    \begin{enumerate}
      \item Which diagrams are equal as morphisms in a monoidal category?
      \item Which diagrams are equal as morphisms in a braided monoidal category?
      \item Which diagrams are equal as morphisms in a symmetric monoidal category?
    \end{enumerate}
\end{exercise}

\pagebreak
\begin{exercise}
  Verify the correctness of the following graphical equations by
  \begin{enumerate}
    \item the correctness theorem of the graphical calculus,
    \item substitution in string diagrams using the axioms of braided monoidal categories,
    \item construction of a commuting diagram from the axioms.
  \end{enumerate}
  \begin{enumerate}[label=(\roman*)]
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_1}}}$
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_2}}}$
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_3}}}$
      ($\sigma_{A, I} = \lambda_{A}^{-1} \of \rho_A$)
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Show that $\Set$ is a symmetric monoidal category under $I \coloneqq \emptyset$ and $A \otimes B \coloneqq A + B + (A \times B)$, where we write $\times$ for Cartesian product of sets and $+$ for disjoint union of sets.
\end{exercise}

\begin{exercise}
  Assume that we have morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon, \varepsilon'} I$ in a monoidal category such that $\eta$ and $\varepsilon$ satisfy one snake equation and $\eta$ and $\varepsilon'$ satisfy the other snake equation.
  Show that this already implies $\varepsilon = \varepsilon'$.
\end{exercise}

\begin{exercise}
  In a monoidal category, show that
  \begin{enumerate}
    \item if an initial object $0$ exists and $L \dashv R$, then $L \otimes 0 \iso 0 \iso 0 \otimes R$;
    \item if a terminal object $1$ exists and $L \dashv R$, then $R \otimes 1 \iso 1 \iso 1 \otimes L$.
  \end{enumerate}
\end{exercise}

\begin{definition}{Idempotent, splitting}
  An endomorphism $A \xrightarrow{f} A$ is called \demph{idempotent} when $f \of f = f$. An idempotent $A \xrightarrow{f} A$ \demph{splits} when there exists an object $B$ and morphisms $A \xrightarrow{g} B$ and $B \xrightarrow{h} A$ such that $f = h \of g$ and $g \of h = 1$.
\end{definition}

\begin{exercise}
  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying one of the snake equations, then $L$ has a right dual (not necessarily $R$).
\end{exercise}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-07-18 sheets/talk-11/sheet-11.tex
Fix typos, notation and extend some explantations

@@
   \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash A(\sigma \circ p')
-  ~\text{type}$). Let $\delta$ be morphism
+  ~\text{type}$). Let $\delta$ be the morphism
   \[
@@
   \[
-    N = q\delta' = q_\Gamma(\sigma \circ p_\Delta \circ \delta', q\delta') = q_\Gamma((\sigma \circ p_\Delta, q) \circ \delta') = q_\Gamma\delta' = M
+    N = q\delta' = q_\Gamma(\sigma \circ p_\Delta \circ \delta', q\delta') = q_\Gamma((\sigma \circ p_\Delta, q) \circ \delta') = q_\Gamma\sigma' = M
   \]
@@
 \begin{answer}
-  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we the elements of $\mathit{Ty}(\Gamma)$:
+  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we give the
+  element of $\mathit{Ty}(\Gamma)$ corresponding to the type $A \to A$:
   \[
@@
   \]
-  And required equalities hold trivially for any $g \from K \to J$: $((\lambda
+  The required equalities hold trivially for any $g \from K \to J$, $((\lambda
   I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.
 
-  Then we give a term by the following equation:
+  $\Gamma, x : A \vdash x : A$ corresponds to $q$ in the model. Thus $\Gamma
+  \vdash \lambda x.x : A -> A$ corresponds to $\lambda q$ being defined as
+  follows:
   \[
@@
       \Gamma, a : A, l : \mathbf{List_A}, x: B \vdash H_{cons} : B[Cons(a, l)/l]\\
-      \Gamma \vdash l : \mathbf{List_A}
-    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, l) : B[l/l]}
+      \Gamma \vdash M : \mathbf{List_A}
+    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, M) : B[M/l]}
   \]
@@
     \inference[List-C-cons]{
-      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(a, l)) : B[Cons(a, l)/l]
-    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(a, l)) = D}
+      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(N, M)) : B[Cons(N, M)/l]
+    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(N, M)) = D}
   \]
-  where $D = H_{cons}[a/a, l/l, R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, l)/x] : B[Cons(a, l)/l]$.
+  where $D = H_{cons}[N/a, M/l, R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, M)/x] : B[Cons(N, M)/l]$.
 
@@
   \]
+
+  Note that it is not the only way to define $\mathit{Vec_A(n)}$, as one could
+  define it with the type formers analogously to our type $\mathit{List_A}$.
 \end{answer}

### 2017-07-19 sheets/talk-11/sheet-11.tex

@@
   \]
+  where $\rho \in \Gamma(I), I \in \cat{C}$
 
-  Let's check that $\int_\cat{C}$ indeed takes us into a category.
+  Let's check that $\int_\cat{C}$ takes us into a category. Recall the
+  definition of $\int_\cat{C}\Gamma$:
   \begin{itemize}
@@
   \end{itemize}
-  We need to check that we have composition and identities.
 
-  With our definition of objects, we can consider first and second projections
-  of the pairs separately. We can observe, that elements of first projection
-  are elements in $\cat{C}$ which have required identities and composition
-  under morphisms in $\cat{C}$ (which are also our morphisms). Then what
-  remains is the second projection. Identities and composition are exactly our
-  conditions on presheaves which also hold.
+  We know our objects and morphisms, we get identities and composition by
+  the equations $\rho 1 = \rho$, $(\rho f) g = \rho (f g)$ and the fact that
+  morphisms $f$ are valid morphisms in the category $\cat{C}$.
 
-  Now we know the operation of $\int_\cat{C}$ on objects and arrows in
-  $Psh(\cat{C})$.
+  Let's now see that $\int_\cat{C}\sigma$ is a functor in $\CAT$. We have a
+  mapping for objects $\int_\cat{C}\sigma(I, \rho)$ given above. The operation
+  on arrows is as follows, where $f \from J \to I$, $g \from K \to J$:
+  \[
+    \int_\cat{C}\sigma(f) := \sigma_{I \to J} \Gamma_f
+  \]
+  Then, keeping in mind that composition $fg$ is applied on the right we get
+  $\int_\cat{C}\sigma(fg) = \sigma_{I \to K} \Gamma_{fg} = \sigma_{I \to K}
+  \lambda \rho.\rho (fg) = \sigma_{I \to K} \lambda \rho. (\rho f)g = \sigma_{J
+  \to K} (\sigma_{I \to J} \lambda \rho.\rho f) g = (\sigma_{J \to K} \lambda
+  \rho.\rho g) (\sigma_{I \to J} \lambda.\rho f) = (\sigma_{J \to K} \Gamma_g)
+  (\sigma_{I \to J} \Gamma_f)$. Which, together with $\rho 1 = \rho$ confirms
+  that $\int_\cat{C}\sigma$ is indeed a functor.
+
+  Thus, we conclude that $\int_\cat{C} \from \mathit{Psh}(\cat{C}) \to \CAT$ is
+  a functor.
 \end{answer}

### 2017-07-20 papers/base-3/paper-3.tex
more stuff for the paper

--- 2017-07-18 papers/base-3/paper-3.tex
@@
 
-\section{Equivalences of Categories}
+\section{Equivalence of Categories}
 \begin{exmp}
@@
 \end{exmp}
+%todo explain example
 
@@
   We define the category $\catA$ such that objects are of the form $K^n$ for $n \in \NN$ and morphisms are linear functions.
-  In order to show that $\catA \eqv \FDVect_K$ we have to define a functor $F \from \catA \to \FDVect_K$.
+  Moreover we denote the category of finite dimensional vector spaces by $\FDVect_K$.
+  Objects in this category are finite dimensional vector spaces and morphisms are linear functions.
+  Let $F \from \catA \to \FDVect_K$ be the embedding functor from $\catA$ into $\FDVect_K$, i.e. $F$ maps objects and morphisms to themselves.
+  Clearly $F$ is full and faithful.
+  For surjectivity on objects we fix an arbitrary object $V \in \FDVect_K$.
+  Let $n \in \NN$ be the dimension of $V$.
+  From linear algebra we know that $V \iso K^n = F(K^n)$.
+  Thus $F$ is essentially surjectivity on objects.
+  By lemma~\ref{lem:equiv} we obtain that $\catA \eqv \FDVect_K$.
+  %In order to show that $\catA \eqv \FDVect_K$ we have to define a functor $F \from \catA \to \FDVect_K$ which is full, faithful and surjective on objects.
 \end{exmp}
@@
 
+\[
+\xymatrix@C+.5em{
+\cat{A} \rtwocell<4>^F_G{\alpha}   &
+\cat{A}' \rtwocell<4>^{F'}_{G'}{\alpha'}   &
+\cat{A}''
+}
+\]
+\[
+\xymatrix@C+.5em{
+  \cat{A} \rtwocell<4>^{F' \of F}_{G' \of G} &\cat{A}'',
+}
+\]
+
+This horizontal composition is written $\alpha' * \alpha$.
+The following diagram commutes because $\alpha'$ is a natural transformation.
+\[
+  \begin{tikzcd}
+    F'(F(A)) \arrow{r}{F'(\alpha_A)} \arrow{d}{\alpha'_{F(A)}} \arrow{dr}{(\alpha' * \alpha)_A} & F'(G(A)) \arrow{d}{\alpha'_{G(A)}} \\
+    G'(F(A)) \arrow{r}{G'(\alpha_A)}                                                            & G'(G(A))
+  \end{tikzcd}
+\]
+So we can define $(\alpha' * \alpha)_A$ either as $G'(\alpha_A) \circ \alpha'_{F(A)}$ or as $\alpha'_{G(A)} \circ F'(\alpha_A)$.
+Now we want to show that $\alpha' * \alpha$ is again a natural transformation.
+Therefore let $f\from A \to A'$ be a morphism in $\catA$.
+\[
+  \begin{tikzcd}
+    F'(F(A)) \arrow{r}{F'(F(f))} \arrow{d}{\alpha'_{F(A)}} & F'(F(A')) \arrow{d}{\alpha'_{F(A')}} \\
+    G'(F(A)) \arrow{r}{G'(F(f))} \arrow{d}{G'(\alpha_A)}   & G'(F(A')) \arrow{d}{G'(\alpha_{A'}}  \\
+    G'(G(A)) \arrow{r}{G'(G(f))}                           & G'(G(A'))
+  \end{tikzcd}
+\]
+
+%todo example
+
+We can also combine horizontal and vertical composition.
+\[
+\xymatrix{
+\cat{A}
+\ruppertwocell<8>^F{\alpha}
+\ar[r]|G
+\rlowertwocell<-8>_H{\beta} &
+\cat{A}'
+\ruppertwocell<8>^{F'}{\hspace{.2em}\alpha'}
+\ar[r]|{G'}
+\rlowertwocell<-8>_{H'}{\hspace{.2em}\beta'} &
+\cat{A}''
+}
+\]
+
+%todo interchange law
+
 \end{document}

### 2017-07-21 sheets/talk-12/sheet-12.tex
Fix some exercises for talk 12

@@
         = \lambda \of (f \otimes 1) \of \alpha^{-1} \of (1 \otimes \sigma) \of \alpha$ \\
-      for $A \otimes B \xrightarrow{f} f$.
+      for $A \otimes B \xrightarrow{f} I$.
   \end{enumerate}
@@
 \begin{exercise}
-  Verify the correctness of the following graphical equations by
+  Verify the correctness of the following graphical equations in braided monoidal categories by
   \begin{enumerate}
     \item the correctness theorem of the graphical calculus,
-    \item substitution in string diagrams using the axioms of braided monoidal categories,
-    \item construction of a commuting diagram from the axioms.
+    \item substitution in string diagrams using the axioms of braided monoidal categories and the correctness theorem for the graphical calculus on general monoidal categories,
+    \item construction of a commuting diagram from the axioms and the coherence theorem for general monoidal categories.
   \end{enumerate}
@@
 \begin{exercise}
-  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying one of the snake equations, then $L$ has a right dual (not necessarily $R$).
+  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying one of the snake equations, then $L$ has a right dual (not necessarily given by $R$ with $\eta$ and $\varepsilon$).
 \end{exercise}

### 2017-07-21 sheets/talk-11/sheet-11.tex

@@
   \[
-    \mathit{Ty}_F(\Gamma) := \mathit{Psh(\Gamma)}
+    \mathit{Ty}_F(\Gamma) := \mathit{Psh(\int_\cat{C} \Gamma)}
   \]
@@
 
-  We can decompose $\sigma'$ as $\sigma' = (p_\Gamma, q_\Gamma) \circ \sigma' =
-  (p_\Gamma \circ \sigma', q_\Gamma\sigma') = (\sigma \circ p', M)$, where $M
-  \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash A(\sigma \circ p')
-  ~\text{type}$). Let $\delta$ be the morphism
+  We can decompose $\sigma'$ as $\sigma' = 1 \circ \sigma' = (p_\Gamma,
+  q_\Gamma) \circ \sigma' = (p_\Gamma \circ \sigma', q_\Gamma\sigma') = (\sigma
+  \circ p', M)$, where $M \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash
+  A(\sigma \circ p') ~\text{type}$). Let $\delta$ be the morphism
   \[
@@
     (p_\Gamma, q_\Gamma) \circ \sigma' =
+    1 \circ \sigma' =
     \sigma'
@@
   \]
-  Then, keeping in mind that composition $fg$ is applied on the right we get
-  $\int_\cat{C}\sigma(fg) = \sigma_{I \to K} \Gamma_{fg} = \sigma_{I \to K}
-  \lambda \rho.\rho (fg) = \sigma_{I \to K} \lambda \rho. (\rho f)g = \sigma_{J
-  \to K} (\sigma_{I \to J} \lambda \rho.\rho f) g = (\sigma_{J \to K} \lambda
-  \rho.\rho g) (\sigma_{I \to J} \lambda.\rho f) = (\sigma_{J \to K} \Gamma_g)
-  (\sigma_{I \to J} \Gamma_f)$. Which, together with $\rho 1 = \rho$ confirms
-  that $\int_\cat{C}\sigma$ is indeed a functor.
+  Recall the naturality axiom for the natural transformation $\sigma$. It
+  requires for each $I \in \cat{C}$ a corresponding $\sigma_I$. Hence the
+  operation of $\sigma$ on the composition of morphisms in $\cat{C}$ can be
+  considered as application of two indexed $\sigma_i$ of the corresponding
+  objects between the morphisms (refer to the naturality square to see this).
+  Below we give the indices on $\sigma$ that correspond to the place in the
+  naturality square for clarity.  Then with associativity and keeping in mind
+  that the composition $fg$ and the individual $f$ and $g$ are applied on the
+  right we get $\int_\cat{C}\sigma(fg) = \sigma_{I \to K} \Gamma_{fg} =
+  \sigma_{I \to K} \lambda \rho.\rho (fg) = \sigma_{I \to K} \lambda \rho.
+  (\rho f)g = \sigma_{J \to K} (\sigma_{I \to J} \lambda \rho.\rho f) g =
+  (\sigma_{J \to K} \lambda \rho.\rho g) (\sigma_{I \to J} \lambda \rho.\rho f)
+  = (\sigma_{J \to K} \Gamma_g) (\sigma_{I \to J} \Gamma_f)$.  Which, together
+  with $\rho 1 = \rho$ confirms that $\int_\cat{C}\sigma$ is indeed a functor.
 
@@
   \[
-    (A \to A)\rho = \{w | w_f \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
+    (A \to A)\rho = \{w | w_f u \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
   \]
@@
 
-  Now suppose that we have dependent types $\Gamma \vdash C \text{ type}$ and
-  $\Gamma, C \vdash D \text{ type}$.
+  Now suppose that we have dependent types $\Gamma \vdash A \text{ type}$ and
+  $\Gamma, A \vdash B \text{ type}$.
 
@@
   \]
+  where $(a, b)f = (af, bf)$ for $f \from J \to I$. The pairing operation
+  $\Gamma \vdash (u, v) : \Sigma AB$ of $\Gamma \vdash u : A$ and $\Gamma.A
+  \vdash v : B(1, u)$ is defined by the componentwise pairing: $(u, v)\rho =
+  (u\rho, v\rho)$. Likewise for the projections $p$ and $q$, $(pw)\rho = a$ and
+  $(qw)\rho = b$. This validates the necessary equations.
 \end{answer}
@@
   \[
-    P : Id(0, Suc(0)) \vdash \mathit{Subst}(P, El(f(0))) : El(f(Suc(0)))
+    P : Id(0, Suc(0)) \vdash \mathit{Subst}(P, (\star : El(f(0)))) : El(f(Suc(0)))
   \]

### 2017-07-21 slides/talk-12/main.tex

\documentclass{beamer}
\input{header}

\begin{document}
\title {Monoidal Categories and Graphical Calculus}
\author {Felix Rech}
\date {July 19, 2017}

\frame[plain, noframenumbering]{\titlepage}

\begin{frame}[fragile]{Strict Monoidal Category}

  A \demph{strict monoidal category} is a category $C$ with the following:
  \begin{description}[labelindent=foo]
    \item[Tensor product:] A functor $\otimes : C \times C -> C$
    \item[Unit object:] An object $I \in \ob(C)$
    \item[Associativity:] An equality \\
      $(A \otimes B) \otimes C = A \otimes (B \otimes C)$
    \item[Left unit law:] An equality $I \otimes A = A$
    \item[Right unit law:] An equality $A \otimes I = A$
  \end{description}

  \transparent{0}
  Those have to satisfy the \demph{triangle} and \demph{pentagon} equalities:

  \vspace{2ex}

  \begin{adjustbox}{max totalsize={\textwidth}{\textheight}, center}
    \transparent{0}
    \begin{tikzcd}[column sep=0]
      (A \otimes I) \otimes B
        \arrow[dr, "\rho_A \otimes 1_B"']
        \arrow[rr, "\alpha_{A, I, B}"] &&
      A \otimes (I \otimes B)
        \arrow[dl, "1_A \otimes \lambda_B"] \\
      %
      & A \otimes B
    \end{tikzcd}
    \begin{tikzpicture}[commutative diagrams/every diagram]
      \node (S) at (0, -2)
          {$(A \otimes B) \otimes (C \otimes D)$};
      \node (E) at (3, 0)
          {$A \otimes (B \otimes (C \otimes D))$};
      \node (NE) at (2.5, 2)
          {$A \otimes ((B \otimes C) \otimes D)$};
      \node (NW) at (-2.5, 2)
          {$(A \otimes (B \otimes C)) \otimes D$};
      \node (W) at (-3, 0)
          {$((A \otimes B) \otimes C) \otimes D$};

      \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C} \otimes 1_D$} (W)
        (W) edge node[anchor=north east] {$\alpha_{A \otimes B, C, D}$} (S)
        (S) edge node[anchor=north west] {$\alpha_{A, B, C \otimes D}$} (E)
        (E) edge node[anchor=west] {$1_A \otimes \alpha_{B, C, D}$} (NE)
        (NW) edge node {$\alpha_{A, B \otimes C, D}$} (NE);
    \end{tikzpicture}
  \end{adjustbox}

\end{frame}

\begin{frame}[fragile]{Monoidal Category}

  A \demph{monoidal category} is a category $C$ with the following:
  \begin{description}[labelindent=foo]
    \item[Tensor product:] A functor $\otimes : C \times C -> C$
    \item[Unit object:] An object $I \in \ob(C)$
    \item[Associator:] A natural isomorphism \\
      $(A \otimes B) \otimes C \xrightarrow{\smash{\alpha_{A, B, C}}} A \otimes (B \otimes C)$
    \item[Left unitor:] A natural isomorphism $I \otimes A \xrightarrow{\smash{\lambda_A}} A$
    \item[Right unitor:] A natural isomorphism $A \otimes I \xrightarrow{\rho_A} A$
  \end{description}

  \pause
  Those have to satisfy the \demph{triangle} and \demph{pentagon} equalities:

  \vspace{2ex}

  \begin{adjustbox}{max totalsize={\textwidth}{\textheight}, center}
    \transparent{0.6}
    \begin{tikzcd}[column sep=0]
      (A \otimes I) \otimes B
        \arrow[dr, "\rho_A \otimes 1_B"']
        \arrow[rr, "\alpha_{A, I, B}"] &&
      A \otimes (I \otimes B)
        \arrow[dl, "1_A \otimes \lambda_B"] \\
      %
      & A \otimes B
    \end{tikzcd}
    \begin{tikzpicture}[commutative diagrams/every diagram]
      \node (S) at (0, -2)
          {$(A \otimes B) \otimes (C \otimes D)$};
      \node (E) at (3, 0)
          {$A \otimes (B \otimes (C \otimes D))$};
      \node (NE) at (2.5, 2)
          {$A \otimes ((B \otimes C) \otimes D)$};
      \node (NW) at (-2.5, 2)
          {$(A \otimes (B \otimes C)) \otimes D$};
      \node (W) at (-3, 0)
          {$((A \otimes B) \otimes C) \otimes D$};

      \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C} \otimes 1_D$} (W)
        (W) edge node[anchor=north east] {$\alpha_{A \otimes B, C, D}$} (S)
        (S) edge node[anchor=north west] {$\alpha_{A, B, C \otimes D}$} (E)
        (E) edge node[anchor=west] {$1_A \otimes \alpha_{B, C, D}$} (NE)
        (NW) edge node {$\alpha_{A, B \otimes C, D}$} (NE);
    \end{tikzpicture}
  \end{adjustbox}

\end{frame}

\begin{frame}

  \begin{theorem}[Coherence for monoidal categories]
    Given the data of a monoidal category, if the pentagon and triangle equations hold, then any well-typed equation built only from $\lambda$, $\rho$ and $\alpha$ and their inverses holds.
  \end{theorem}

\end{frame}

\begin{frame}{Graphical Calculus (1)}

  \begin{definition}
    Two diagrams are \demph{isotopic} when one can be deformed continuously into the other, keeping the boundaries fixed.
  \end{definition}

  \begin{theorem}[Correctness for monoidal categories]
    A well-typed equation between morphisms in a monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{planar} isotopy.
  \end{theorem}

  \invisible{
    \begin{theorem}[Correctness for braided monoidal categories]
      A well-typed equation between morphisms in a braided monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{spatial} isotopy.
    \end{theorem}

    \begin{theorem}[Correctness for symmetric monoidal categories]
      A well-typed equation between morphisms in a symmetric monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{four-dimensional} isotopy.
    \end{theorem}
  }

\end{frame}

\begin{frame}{Braiding}

  A \demph{braided monoidal category} is a monoidal category equipped with a natural isomorphism
  \[ A \otimes B \xrightarrow{\sigma_{A, B}} B \otimes A \]
  satisfying the \demph{hexagon equations}:

  \vspace{2ex}

  \begin{adjustbox}{max totalsize={\textwidth}{\textheight}, center}
    \transparent{0.6}
    \begin{tikzpicture}[commutative diagrams/every diagram]
      \node (SW) at (-1.75, -1.5)
          {$(B \otimes A) \otimes C$};
      \node (SE) at (1.75, -1.5)
          {$B \otimes (A \otimes C)$};
      \node (E) at (2.5, 0)
          {$B \otimes (C \otimes A)$};
      \node (NE) at (1.75, 1.5)
          {$(B \otimes C) \otimes A$};
      \node (NW) at (-1.75, 1.5)
          {$A \otimes (B \otimes C)$};
      \node (W) at (-2.5, 0)
          {$(A \otimes B) \otimes C$};

      \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C}^{-1}$} (W)
        (W) edge node[anchor=east] {$\sigma{A, B} \otimes 1_C$} (SW)
        (SW) edge node[anchor=north] {$\alpha_{B, A, C}$} (SE)
        (SE) edge node[anchor=west] {$1_B \otimes \sigma_{A, C}$} (E)
        (E) edge node[anchor=west] {$\alpha_{B, C, A}^{-1}$} (NE)
        (NW) edge node {$\sigma_{A, B \otimes C}$} (NE);
    \end{tikzpicture}
    \begin{tikzpicture}[commutative diagrams/every diagram]
      \node (SW) at (-1.75, -1.5)
          {$A \otimes (C \otimes B)$};
      \node (SE) at (1.75, -1.5)
          {$(A \otimes C) \otimes B$};
      \node (E) at (2.5, 0)
          {$(C \otimes A) \otimes B$};
      \node (NE) at (1.75, 1.5)
          {$C \otimes (A \otimes B)$};
      \node (NW) at (-1.75, 1.5)
          {$(A \otimes B) \otimes C$};
      \node (W) at (-2.5, 0)
          {$A \otimes (B \otimes C)$};

      \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C}$} (W)
        (W) edge node[anchor=east] {$1_A \otimes \sigma_{B, C}$} (SW)
        (SW) edge node[anchor=north] {$\alpha_{A, C, B}^{-1}$} (SE)
        (SE) edge node[anchor=west] {$\sigma_{A, C} \otimes 1_B$} (E)
        (E) edge node[anchor=west] {$\alpha_{C, A, B}$} (NE)
        (NW) edge node {$\sigma_{A \otimes B, C}$} (NE);
    \end{tikzpicture}
  \end{adjustbox}

  \vspace{2ex}

  \invisible{
    A braided monoidal category is \demph{symmetric} when
    \[ \sigma_{B, A}^{-1} = \sigma_{A, B}. \]
  }

\end{frame}

\begin{frame}{Graphical Calculus (2)}

  \begin{definition}
    Two diagrams are \demph{isotopic} when one can be deformed continuously into the other, keeping the boundaries fixed.
  \end{definition}

  {\transparent{0.2}
    \begin{theorem}[Correctness for monoidal categories]
      A well-typed equation between morphisms in a monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{planar} isotopy.
    \end{theorem}
  }

  \begin{theorem}[Correctness for braided monoidal categories]
    A well-typed equation between morphisms in a braided monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{spatial} isotopy.
  \end{theorem}

  \invisible{
    \begin{theorem}[Correctness for symmetric monoidal categories]
      A well-typed equation between morphisms in a symmetric monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{four-dimensional} isotopy.
    \end{theorem}
  }

\end{frame}

\begin{frame}{Symmetric Braiding}

  A \demph{braided monoidal category} is a monoidal category equipped with a natural isomorphism
  \[ A \otimes B \xrightarrow{\sigma_{A, B}} B \otimes A \]
  satisfying the \demph{hexagon equations}:

  \vspace{2ex}

  \begin{adjustbox}{max totalsize={\textwidth}{\textheight}, center}
    \transparent{0.6}
    \begin{tikzpicture}[commutative diagrams/every diagram]
      \node (SW) at (-1.75, -1.5)
          {$(B \otimes A) \otimes C$};
      \node (SE) at (1.75, -1.5)
          {$B \otimes (A \otimes C)$};
      \node (E) at (2.5, 0)
          {$B \otimes (C \otimes A)$};
      \node (NE) at (1.75, 1.5)
          {$(B \otimes C) \otimes A$};
      \node (NW) at (-1.75, 1.5)
          {$A \otimes (B \otimes C)$};
      \node (W) at (-2.5, 0)
          {$(A \otimes B) \otimes C$};

      \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C}^{-1}$} (W)
        (W) edge node[anchor=east] {$\sigma{A, B} \otimes 1_C$} (SW)
        (SW) edge node[anchor=north] {$\alpha_{B, A, C}$} (SE)
        (SE) edge node[anchor=west] {$1_B \otimes \sigma_{A, C}$} (E)
        (E) edge node[anchor=west] {$\alpha_{B, C, A}^{-1}$} (NE)
        (NW) edge node {$\sigma_{A, B \otimes C}$} (NE);
    \end{tikzpicture}
    \begin{tikzpicture}[commutative diagrams/every diagram]
      \node (SW) at (-1.75, -1.5)
          {$A \otimes (C \otimes B)$};
      \node (SE) at (1.75, -1.5)
          {$(A \otimes C) \otimes B$};
      \node (E) at (2.5, 0)
          {$(C \otimes A) \otimes B$};
      \node (NE) at (1.75, 1.5)
          {$C \otimes (A \otimes B)$};
      \node (NW) at (-1.75, 1.5)
          {$(A \otimes B) \otimes C$};
      \node (W) at (-2.5, 0)
          {$A \otimes (B \otimes C)$};

      \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C}$} (W)
        (W) edge node[anchor=east] {$1_A \otimes \sigma_{B, C}$} (SW)
        (SW) edge node[anchor=north] {$\alpha_{A, C, B}^{-1}$} (SE)
        (SE) edge node[anchor=west] {$\sigma_{A, C} \otimes 1_B$} (E)
        (E) edge node[anchor=west] {$\alpha_{C, A, B}$} (NE)
        (NW) edge node {$\sigma_{A \otimes B, C}$} (NE);
    \end{tikzpicture}
  \end{adjustbox}

  \vspace{2ex}

  A braided monoidal category is \demph{symmetric} when
  \[ \sigma_{B, A}^{-1} = \sigma_{A, B}. \]

\end{frame}

\begin{frame}{Graphical Calculus (3)}

  \begin{definition}
    Two diagrams are \demph{isotopic} when one can be deformed continuously into the other, keeping the boundaries fixed.
  \end{definition}

  {\transparent{0.2}
    \begin{theorem}[Correctness for monoidal categories]
      A well-typed equation between morphisms in a monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{planar} isotopy.
    \end{theorem}

    \begin{theorem}[Correctness for braided monoidal categories]
      A well-typed equation between morphisms in a braided monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{spatial} isotopy.
    \end{theorem}
  }

  \begin{theorem}[Correctness for symmetric monoidal categories]
    A well-typed equation between morphisms in a symmetric monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{four-dimensional} isotopy.
  \end{theorem}

\end{frame}

\begin{frame}[t, fragile]{Dual Objects}

  In a monoidal category, an object $L$ is \demph{left-dual} to an object $R$, and $R$ is \demph{right-dual} to $L$, written $L \dashv R$, when there exist
  \begin{itemize}
    \item a \demph{unit} morphism $I \xrightarrow{\eta} R \otimes L$ and
    \item a \demph{counit} morphism $L \otimes R \xrightarrow{\varepsilon} I$
  \end{itemize}
  satisfying the \demph{snake equations}:

  \vspace{2ex}

  \begin{adjustbox}{max width=\linewidth, center}
  \transparent{0.6}
  \begin{tikzcd}
    L \arrow[r, "\rho_L^{-1}"] \arrow[d, "1_L"] &
    L \otimes I \arrow[r, "1_L \otimes \eta"] &
    L \otimes (R \otimes L) \arrow[d, "\alpha_{L, R, L}^{-1}"] \\
    %
    L &
    I \otimes L \arrow[l, "\lambda_L"] &
    (L \otimes R) \otimes L \arrow[l, "\varepsilon \otimes 1_L"]
  \end{tikzcd}
  \begin{tikzcd}
    R \arrow[r, "\lambda_R^{-1}"] \arrow[d, "1_R"] &
    I \otimes R \arrow[r, "\eta \otimes 1_R"] &
    (R \otimes L) \otimes R \arrow[d, "\alpha_{R, L, R}"] \\
    %
    R &
    R \otimes I \arrow[l, "\rho_R"] &
    R \otimes (L \otimes R) \arrow[l, "1_R \otimes \varepsilon"]
  \end{tikzcd}
  \end{adjustbox}

\end{frame}

\begin{frame}[t]{Properties of Duals}

  \begin{lemma}
    The left- and right-dual of an object are unique.
  \end{lemma}

  \pause

  \begin{lemma}
    The unit of a duality determines the counit and vice-versa.
  \end{lemma}

  \pause

  \begin{lemma}
    In a monoidal category $I \dashv I$.
  \end{lemma}

  \pause

  \begin{lemma}
    In a monoidal category $L \dashv R$ and $L' \dashv R'$ implies $L \otimes L' \dashv R' \otimes R$.
  \end{lemma}

  \pause

  \begin{lemma}
    In a braided monoidal category $L \dashv R$ implies $R \dashv L$.
  \end{lemma}

\end{frame}

\end{document}

### 2017-07-21 papers/base-3/paper-3.tex
merged with Maxi

--- 2017-07-18 papers/base-3/paper-3.tex
@@
 \newtheorem{exmp}[counter]{Example}
+\newtheorem{lemma}[counter]{Lemma}
 
@@
 
+\section{Equivalence of Categories}
+\begin{exmp}
+  \[
+    \begin{tikzcd}
+      \catA: & & \catB:\\
+      A \arrow[loop left] & & B \arrow[bend left]{d} \arrow[loop left] \\
+                          & & B' \arrow[bend left]{u} \arrow[loop left]
+    \end{tikzcd}
+  \]
+\end{exmp}
+%todo explain example
+
+\begin{defn}
+  \label{def:eqv}
+  Tow categories $\catA$ and $\catB$ are \emph{equivalent} if:
+  \begin{itemize}
+    \item There are functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ and
+    \item there are natural transformations $\eta \from 1_{\catA} \to G \circ F$ and $\varepsilon \from F \circ G \to 1_\catB$.
+  \end{itemize}
+  The functors $F$ and $G$ are called \emph{equivalences}.
+  We write $\catA \eqv \catB$ if $\catA$ and $\catB$ are equivalent.
+\end{defn}
+
+\begin{defn}
+  \label{def:surj-on-obj}
+  A functor $F \from \catA \to \catB$ is \emph{essentially surjective on objects}
+  if for every object $B \in \catB$ there is an object $A \in \catA$ such that $F(A) \iso B$.
+\end{defn}
+
+\begin{lemma}
+  \label{lem:equiv}
+  A functor $F \from \catA \to \catB$ is an equivalence if and only if $F$ is full, faithful and surjective on objects.
+\end{lemma}
+
+\begin{exmp}
+  Let $K$ be a field.
+  We define the category $\catA$ such that objects are of the form $K^n$ for $n \in \NN$ and morphisms are linear functions.
+  Moreover we denote the category of finite dimensional vector spaces by $\FDVect_K$.
+  Objects in this category are finite dimensional vector spaces and morphisms are linear functions.
+  Let $F \from \catA \to \FDVect_K$ be the embedding functor from $\catA$ into $\FDVect_K$, i.e. $F$ maps objects and morphisms to themselves.
+  Clearly $F$ is full and faithful.
+  For surjectivity on objects we fix an arbitrary object $V \in \FDVect_K$.
+  Let $n \in \NN$ be the dimension of $V$.
+  From linear algebra we know that $V \iso K^n = F(K^n)$.
+  Thus $F$ is essentially surjectivity on objects.
+  By lemma~\ref{lem:equiv} we obtain that $\catA \eqv \FDVect_K$.
+  %In order to show that $\catA \eqv \FDVect_K$ we have to define a functor $F \from \catA \to \FDVect_K$ which is full, faithful and surjective on objects.
+\end{exmp}
+
+\section{Horizontal Composition}
+
+\[
+\xymatrix@C+.5em{
+\cat{A} \rtwocell<4>^F_G{\alpha}   &
+\cat{A}' \rtwocell<4>^{F'}_{G'}{\alpha'}   &
+\cat{A}''
+}
+\]
+\[
+\xymatrix@C+.5em{
+  \cat{A} \rtwocell<4>^{F' \of F}_{G' \of G} &\cat{A}'',
+}
+\]
+
+This horizontal composition is written $\alpha' * \alpha$.
+The following diagram commutes because $\alpha'$ is a natural transformation.
+\[
+  \begin{tikzcd}
+    F'(F(A)) \arrow{r}{F'(\alpha_A)} \arrow{d}{\alpha'_{F(A)}} \arrow{dr}{(\alpha' * \alpha)_A} & F'(G(A)) \arrow{d}{\alpha'_{G(A)}} \\
+    G'(F(A)) \arrow{r}{G'(\alpha_A)}                                                            & G'(G(A))
+  \end{tikzcd}
+\]
+So we can define $(\alpha' * \alpha)_A$ either as $G'(\alpha_A) \circ \alpha'_{F(A)}$ or as $\alpha'_{G(A)} \circ F'(\alpha_A)$.
+Now we want to show that $\alpha' * \alpha$ is again a natural transformation.
+Therefore let $f\from A \to A'$ be a morphism in $\catA$.
+\[
+  \begin{tikzcd}
+    F'(F(A)) \arrow{r}{F'(F(f))} \arrow{d}{\alpha'_{F(A)}} & F'(F(A')) \arrow{d}{\alpha'_{F(A')}} \\
+    G'(F(A)) \arrow{r}{G'(F(f))} \arrow{d}{G'(\alpha_A)}   & G'(F(A')) \arrow{d}{G'(\alpha_{A'}}  \\
+    G'(G(A)) \arrow{r}{G'(G(f))}                           & G'(G(A'))
+  \end{tikzcd}
+\]
+
+%todo example
+
+We can also combine horizontal and vertical composition.
+\[
+\xymatrix{
+\cat{A}
+\ruppertwocell<8>^F{\alpha}
+\ar[r]|G
+\rlowertwocell<-8>_H{\beta} &
+\cat{A}'
+\ruppertwocell<8>^{F'}{\hspace{.2em}\alpha'}
+\ar[r]|{G'}
+\rlowertwocell<-8>_{H'}{\hspace{.2em}\beta'} &
+\cat{A}''
+}
+\]
+
+%todo interchange law
+
 \end{document}

### 2017-07-23 papers/base-3/paper-3.tex
added interchange law and some explanation

@@
 \section{Equivalence of Categories}
+We now want to examine when two categories are the same.
 \begin{exmp}
@@
 \end{exmp}
-%todo explain example
+todo: explain example
 
@@
 
+Given two natural transformations $\alpha$ and $\alpha'$:
 \[
@@
 \cat{A}' \rtwocell<4>^{F'}_{G'}{\alpha'}   &
-\cat{A}''
+\cat{A}'',
 }
 \]
+we can obtain a natural transformation:
 \[
 \xymatrix@C+.5em{
-  \cat{A} \rtwocell<4>^{F' \of F}_{G' \of G} &\cat{A}'',
+  \cat{A} \rtwocell<4>^{F' \circ F}_{G' \circ G} &\cat{A}'',
 }
@@
     F'(F(A)) \arrow{r}{F'(F(f))} \arrow{d}{\alpha'_{F(A)}} & F'(F(A')) \arrow{d}{\alpha'_{F(A')}} \\
-    G'(F(A)) \arrow{r}{G'(F(f))} \arrow{d}{G'(\alpha_A)}   & G'(F(A')) \arrow{d}{G'(\alpha_{A'}}  \\
+    G'(F(A)) \arrow{r}{G'(F(f))} \arrow{d}{G'(\alpha_A)}   & G'(F(A')) \arrow{d}{G'(\alpha_{A'})}  \\
     G'(G(A)) \arrow{r}{G'(G(f))}                           & G'(G(A'))
@@
 
-%todo example
+The upper diagram commutes because $\alpha'$ is a natural transformation.
+Since $\alpha$ is a natural transformation we have $G(f) \circ \alpha_A = \alpha_{A'} \circ F(f)$.
+By applying $G'$ on both sides we obtain the commutativity of the lower diagram.
+The commutativity of both inner diagrams ensures the commutativity of the outer diagram.
+This shows that $\alpha' * \alpha$ is indeed a natural transformation.
 
@@
 
-%todo interchange law
+This natural transformations follow the interchange law:
+\[
+(\beta' \circ \alpha') * (\beta \circ \alpha)
+=
+(\beta' * \beta) \circ (\alpha' * \alpha)
+\from
+F' \circ F \to H' \circ H.
+\]
+Additionally we have $1_{F'} * 1_{F} = 1_{F' \circ F}$, where $1_{F}$ is the natural transformation which components are the identities of $\cat{A'}$.
 

### 2017-07-25 papers/base-3/paper-3.tex
Add the material on functor categories and natural isomorphisms (#1)

@@
   }\]
+\end{exmp}
+
+\begin{defn}[Functor Category]
+  Natural transformations, as other types of maps that we've encountered earlier
+  can be composed. That is, for any two categories $\cat{A}$ and $\cat{B}$,
+  functors $F, G, H$ between them, natural transformations $\alpha \from F \to
+  G$ and $\beta \from G \to H$:
+  \vspace{-2ex}
+  \[
+  \xymatrix@C+1.5em{
+  \cat{A}
+  \ruppertwocell<8>^F{\alpha}
+  \ar[r]|G
+  \rlowertwocell<-8>_H{\beta}
+  &
+  \cat{B}
+  },
+  \]
+  \vspace{-3ex}\\
+  there is a composite natural transformation $\beta \of \alpha$
+  \[
+  \xymatrix@C+1.5em{
+  \cat{A} \rtwocell^F_H{\hspace{.8em}\beta\of\alpha} &\cat{B}
+  }
+  \]
+  defined by $(\beta\of\alpha)_A = \beta_A \of \alpha_A$ for all $A \in \cat{A}$.
+  There is also an identity natural transformation
+  \[
+  \xymatrix@C+1em{
+  \cat{A} \rtwocell^F_F{\hspace{.5em}1_F} &\cat{B}
+  }
+  \]
+  on any functor $F$, defined by $(1_F)_A = 1_{F(A)}$.
+
+  This gives rives to a category whose objects are the functors from $\cat{A}$ to
+  $\cat{B}$ and whose maps are the natural transformations between them. This is
+  called the \demph{functor category} from $\cat{A}$ to $\cat{B}$, and written as
+  $\ftrcat{\cat{A}}{\cat{B}}$ or $\cat{B}^\cat{A}$.
+\end{defn}
+
+\begin{exmp}
+  Let $2$ be the discrete category with two objects. A functor from $2$ to a
+  category $\cat{B}$ is a pair of objects of $\cat{B}$, and a natural
+  transformation is a pair of maps.  The functor category $\ftrcat{2}{\cat{B}}$
+  is therefore isomorphic to the product category $\cat{B} \times \cat{B}$.
+  This fits well with the alternative notation $\cat{B}^2$ for the functor
+  category.
+\end{exmp}
+
+\begin{exmp}
+  Take ordered sets $A$ and $B$, viewed as categories. Given order-preserving
+  maps $\parpairi{A}{B}{f}{g}$, viewed as functors, there is at most one
+  natural transformation
+  \[
+  \xymatrix{
+  A \rtwocell^f_g &B,
+  }
+  \]
+  and there is one if and only if $f(a) \leq g(a)$ for all $a \in A$.  (The
+  naturality axiom~\eqref{eq:nat} holds automatically, because in an ordered
+  set, all diagrams commute.)  So $\ftrcat{A}{B}$ is an ordered set too; its
+  elements are the order-preserving maps from $A$ to $B$, and $f \leq g$ if and
+  only if $f(a) \leq g(a)$ for all $a \in A$.
+\end{exmp}
+
+\begin{defn}
+  Let $\cat{A}$ and $\cat{B}$ be categories. A \demph{natural isomorphism}
+  between functors from $\cat{A}$ to $\cat{B}$ is an isomorphism in
+  $\ftrcat{\cat{A}}{\cat{B}}$.
+\end{defn}
+
+This definition can also be given as a lemma:
+
+\begin{lemma}
+Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural
+transformation.  Then $\alpha$ is a natural isomorphism if and only if
+$\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
+\end{lemma}
+
+\begin{proof}
+  \begin{itemize}
+    \item[($\rightarrow$)]
+      By definition of natural isomorphism we have a natural transformation
+      $\beta\from G \to F$, such that \[\beta \circ \alpha = 1_{F} \text{ and }
+      \alpha \circ \beta = 1_{G}. \] We show that $\forall A \in \cat{A}.\,
+      \beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A =
+      1_{G(A)}$.  Suppose $A \in \cat{A}$.  We know by definition of the
+      identity natural transformation that $(\beta \circ \alpha)_A = 1_{F(A)}$.
+      The claim $\beta_A \circ \alpha_A = 1_{F(A)}$ follows by definition of
+      composition of natural transformations.  $\alpha_A \circ \beta_A =
+      1_{G(A)}$ is shown analogously.
+    \item[($\leftarrow$)]
+      We know that for all $A \in \cat{A}$ there is a $\beta_A$ such that
+      \[\beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A =
+      1_{G(A)}.\] First we prove that $\beta$ is a natural transformation.
+      Suppose $f\from A \to A'$.  We know that $\alpha$ is a natural
+      transformation:
+      \begin{alignat*}{2}
+          & & G(f) \circ \alpha_A & = \alpha_{A'} \circ F(f) \\
+          &\Rightarrow &\quad G(f) \circ \alpha_A \circ \beta_A & = \alpha_{A'} \circ F(f) \circ \beta_A\\
+          &\Rightarrow & G(f) & = \alpha_{A'} \circ F(f) \circ \beta_A\\
+          &\Rightarrow & \beta_{A'} \circ G(f) & = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
+          &\Rightarrow & \beta_{A'} \circ G(f) & = F(f) \circ \beta_A
+      \end{alignat*}
+      By similar reasoning to (a) we get that $\beta$ is the inverse of
+      $\alpha$.  Therefore $\alpha$ is a natural isomorpism.
+  \end{itemize}
+\end{proof}
+
+Then, we say that functors $F$ and $G$ are \demph{naturally isomorphic} if
+there is a natural isomorphism from $F$ to $G$. Note that this natural
+isomorphism is just isomorphism in $\ftrcat{\cat{A}}{\cat{B}}$, and we already
+have notation for this: $F \iso G$.
+
+\begin{defn}
+  Given functors $\parpair{\cat{A}}{\cat{B}}{F}{G}$, we say that
+  \[
+  F(A) \iso G(A) \text{ \demph{naturally in} } A
+  \]
+  if $F$ and $G$ are naturally isomorphic.
+\end{defn}
+
+This alternative terminology can be understood as follows.  If $F(A) \iso
+G(A)$ naturally in $A$ then certainly $F(A) \iso G(A)$ for each individual
+$A$, but more is true: we can choose isomorphisms $\alpha_A\from F(A) \to
+G(A)$ in such a way that the naturality axiom~\eqref{eq:nat} is satisfied.
+
+\begin{exmp}
+  Let $F, G\from \cat{A} \to \cat{B}$ be functors from a discrete category
+  $\cat{A}$ to a category $\cat{B}$.  Then $F \iso G$ if and only if $F(A) \iso
+  G(A)$ for all $A \in \cat{A}$.
+
+  So in \emph{this} case, $F(A) \iso G(A)$ naturally in $A$ if and only if
+  $F(A) \iso G(A)$ for all $A$.  But this is only true because $\cat{A}$ is
+  discrete.  In general, it is emphatically false.  There are many examples of
+  categories and functors $\parpairi{\cat{A}}{\cat{B}}{F}{G}$ such that $F(A)
+  \iso G(A)$ for all $A \in \cat{A}$, but not \emph{naturally} in $A$.
 \end{exmp}

### 2017-07-26 sheets/talk-12/sheet-12.tex
Add example solutions for talk 12

@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}
+
+\usetikzlibrary{fit}
+\usetikzlibrary{knots}
+\usetikzlibrary{hobby}
+\tikzstyle{morphism}=[draw, minimum size=0.5cm, semithick]
+\tikzstyle{phantom morphism}=[minimum size=0.5cm]
+\tikzstyle{object}=[semithick]
+\tikzstyle{cap object}=[object, out=90, in=90, looseness=2]
+\tikzstyle{cup object}=[object, out=270, in=270, looseness=2]
+\tikzstyle{highlight box}=[draw, dotted]
 
@@
 \end{exercise}
+
+\begin{answer}
+  We construct the morphism as
+  \[ \begin{tikzcd}
+    (((A \otimes I) \otimes B) \otimes C) \otimes D
+      \arrow[d, "((\rho_A \otimes 1_B) \otimes 1_C) \otimes 1_D"] \\
+    ((A \otimes B) \otimes C) \otimes D
+      \arrow[d, "\alpha_{A, B, C} \otimes 1_D"] \\
+    (A \otimes (B \otimes C)) \otimes D
+      \arrow[d, "\alpha_{A, B \otimes C, D}"] \\
+    A \otimes ((B \otimes C) \otimes D)
+      \arrow[d, "1_A \otimes \alpha_{B, C, D}"] \\
+    A \otimes (B \otimes (C \otimes D))
+      \arrow[d, "1_A \otimes (1_B \otimes (1_C \otimes \lambda_D^{-1}))"] \\
+    A \otimes (B \otimes (C \otimes (I \otimes D))).
+  \end{tikzcd} \]
+  We have no more information about the category except that it is monoidal.
+  In this setting we cannot construct two different morphisms of the same type, which follows from the coherence theorem.
+\end{answer}
 
@@
 
+\begin{answer}
+  \begin{enumerate}
+  \item
+    \[ \vcenter{ \hbox{ \begin{tikzpicture}
+      \coordinate (in_f) at (0, 0);
+      \node[morphism] (f) at (0, 1) {$f$};
+      \node[phantom morphism] (f') at (1, 2.5) {};
+      \coordinate (out_f) at (1, 3.5);
+
+      \coordinate (in_g) at (1, 0);
+      \node[phantom morphism] (g') at (1, 1) {};
+      \node[morphism] (g) at (0, 2.5) {$g$};
+      \coordinate (out_g) at (0, 3.5);
+
+      \begin{knot}[clip width=8]
+        \strand[object] (in_f)
+          to node[auto] {$A$} (f.south)
+          (f.north) to[out=90, in=270] (f'.south)
+          to (f'.north)
+          to node[auto] {$A$} (out_f);
+        \strand[object] (in_g)
+          to node[auto] {$A$} (g'.south)
+          to (g'.north)
+          to[out=90, in=270] (g.south)
+          (g.north) to node[auto] {$A$} (out_g);
+      \end{knot}
+    \end{tikzpicture}}}
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \coordinate (in_f) at (0, 0);
+      \node[morphism] (f) at (0, 1) {$g$};
+      \node[phantom morphism] (f') at (1, 2.5) {};
+      \coordinate (out_f) at (1, 3.5);
+
+      \coordinate (in_g) at (1, 0);
+      \node[phantom morphism] (g') at (1, 1) {};
+      \node[morphism] (g) at (0, 2.5) {$f$};
+      \coordinate (out_g) at (0, 3.5);
+
+      \begin{knot}[clip width=8]
+        \strand[object] (in_f)
+          to node[auto] {$A$} (f.south)
+          (f.north) to[out=90, in=270] (f'.south)
+          to (f'.north)
+          to node[auto] {$A$} (out_f);
+        \strand[object] (in_g)
+          to node[auto] {$A$} (g'.south)
+          to (g'.north)
+          to[out=90, in=270] (g.south)
+          (g.north) to node[auto] {$A$} (out_g);
+      \end{knot}
+    \end{tikzpicture}}} \]
+
+  \item
+    \[ \vcenter{ \hbox{ \begin{tikzpicture}
+      \coordinate (in) at (0, 0);
+
+      \node[phantom morphism] (g_left) at (-1, 1) {};
+      \node[phantom morphism] (g_right) at (1, 1) {};
+      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
+      \node at (g) {$g$};
+
+      \node[morphism] (f) at (0, 2) {$f$};
+
+      \node[phantom morphism] (h_left) at (-1, 3) {};
+      \node[phantom morphism] (h_right) at (1, 3) {};
+      \node[morphism, fit=(h_left) (h_right), inner sep=0] (h) {};
+      \node at (h) {$h$};
+
+      \coordinate (out) at (0, 4);
+
+      \draw[object] (in) to node[auto] {$A$} (g);
+      \draw[object] (-1, 0 |- g.north) to node[auto] {$B$} (-1, 0 |- h.south);
+      \draw[object] (1, 0 |- g.north) to node[auto] {$C$} (1, 0 |- h.south);
+      \draw[object] (h) to node[auto] {$D$} (out);
+    \end{tikzpicture}}}
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \coordinate (in) at (0, 0);
+
+      \node[morphism] (f) at (-1, 1) {$f$};
+
+      \node[phantom morphism] (g_left) at (-1, 2) {};
+      \node[phantom morphism] (g_right) at (1, 2) {};
+      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
+      \node at (g) {$g$};
+
+      \node[phantom morphism] (h_left) at (-1, 3) {};
+      \node[phantom morphism] (h_right) at (1, 3) {};
+      \node[morphism, fit=(h_left) (h_right), inner sep=0] (h) {};
+      \node at (h) {$h$};
+
+      \coordinate (out) at (0, 4);
+
+      \draw[object] (in) to node[auto] {$A$} (g);
+      \draw[object] (-1, 0 |- g.north) to node[auto] {$B$} (-1, 0 |- h.south);
+      \draw[object] (1, 0 |- g.north) to node[auto] {$C$} (1, 0 |- h.south);
+      \draw[object] (h) to node[auto] {$D$} (out);
+    \end{tikzpicture}}} \]
+
+  \item
+    \[ \vcenter{ \hbox{ \begin{tikzpicture}
+      \coordinate (A_in) at (0, 0);
+      \coordinate (C_in) at (1, 0);
+      \coordinate (B_in) at (2, 0);
+      \node[phantom morphism] (C_1) at (0, 1.5) {};
+
+      \node[phantom morphism] (f_left) at (1, 1.5) {};
+      \node[phantom morphism] (f_right) at (2, 1.5) {};
+      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+      \node at (f) {$f$};
+
+      \coordinate (out) at (0, 2);
+
+      \draw[object] (B_in) to node[auto, at start] {$B$} (B_in |- f.south);
+
+      \begin{knot}[clip width=8]
+        \strand[object, out=90, in=270] (A_in)
+          to node[auto, at start] {$A$} (C_in |- f.south);
+        \strand[object, out=90, in=270] (C_in)
+          to node[auto, at start] {$C$} (C_1.south)
+          to (out);
+      \end{knot}
+    \end{tikzpicture}}}
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \coordinate (A_in) at (0, 0);
+      \coordinate (C_in) at (1, 0);
+      \coordinate (B_in) at (2, 0);
+      \node[phantom morphism] (C_1) at (2, 1.5) {};
+
+      \node[phantom morphism] (f_left) at (0, 1.5) {};
+      \node[phantom morphism] (f_right) at (1, 1.5) {};
+      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+      \node at (f) {$f$};
+
+      \coordinate (out) at (2, 2);
+
+      \draw[object] (A_in) to node[auto, at start] {$A$} (A_in |- f.south);
+
+      \begin{knot}[clip width=8]
+        \strand[object, out=90, in=270] (C_in)
+          to node[auto, at start] {$C$} (C_1.south)
+          to (out);
+        \strand[object, out=90, in=270] (B_in)
+          to node[auto, at start] {$B$} (C_in |- f.south);
+      \end{knot}
+    \end{tikzpicture}}} \]
+  \end{enumerate}
+
+  The correctness theorem for the graphical calculus on symmetric monoidal categories tells us that two diagrams describe the same morphism if and only if they are isotopic in four-dimensional space.
+  This means essentially that nodes can move accross edges and that edges can move \textit{through} each other.
+  With this we can see that equations (b) and (c) hold, but not equation (a).
+\end{answer}
+
 \begin{exercise}
@@
 
-\pagebreak
+\begin{answer}
+  \begin{enumerate}
+  \item
+    We regard the diagrams up to planar isotopy.
+    Thus diagrams (1) and (2) are equal but (3) and (4) are not equal to any other diagram.
+  \item
+    We regard the diagrams up to spacial isotopy.
+    Thus diagrams (1), (2) and (3) are equal but (4) is not equal to any other diagram.
+  \item
+    We regard the diagrams up to four-dimensional isotopy.
+    Thus all diagrams are equal.
+  \end{enumerate}
+\end{answer}
+
+\ifdefined\issolution \else \pagebreak \fi
 \begin{exercise}
@@
 
+\begin{answer}
+  \begin{enumerate}
+  \item
+    In three-dimensional space, the diagrams are obviously isotopic.
+    The equations follow directly from the correctness theorem.
+  \item
+    \begin{enumerate}
+    \item
+      \[
+        \vcenter{ \hbox{ \begin{tikzpicture}
+          \node[phantom morphism] (in) at (0, -0.5) {};
+          \node[phantom morphism] (f_left) at (1, 2) {};
+          \node[phantom morphism] (f_right) at (2, 2) {};
+          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+          \node[phantom morphism] (out_1) at (0, 4) {};
+          \node[phantom morphism] (out_2) at (1, 4) {};
+          \node[phantom morphism] (out_3) at (2, 4) {};
+
+          \begin{knot}[clip width=8]
+            \strand[object, out=90, in=270] (in)
+              to (in |- f.north)
+              to (out_2);
+            \strand[object, out=90, in=270] (f_left) to (out_1);
+            \strand[object, out=90, in=270] (f_right) to (out_3);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[use Hobby shortcut]
+          \node[phantom morphism] (in) at (2.5, -0.5) {};
+          \node[phantom morphism] (f_left) at (0.5, 1) {};
+          \node[phantom morphism] (f_right) at (1.5, 1) {};
+          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+          \node[phantom morphism] (out_1) at (1, 4) {};
+          \node[phantom morphism] (out_2) at (2, 4) {};
+          \node[phantom morphism] (out_3) at (3, 4) {};
+
+          \begin{knot}[clip width=8]
+            \strand[object, out=90, in=270] (in)
+              to (in |- f.north)
+              to (1, 2.5)
+              to (out_2);
+            \strand[object, in angle=-90, out angle=90, in=-90, out=90] (f_left.north)
+              .. (1.6, 1.92)
+              .. (out_2 |- 0, 2.5)
+              to (out_1);
+            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
+              (f_right.north)
+              .. (1.9, 1.83)
+              .. (out_3 |- 0, 2.5)
+              to (out_3);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}
+          \node[phantom morphism] (in) at (2, 0) {};
+          \node[phantom morphism] (f_left) at (0, 1) {};
+          \node[phantom morphism] (f_right) at (1, 1) {};
+          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+          \node[phantom morphism] (out_1) at (0, 4.5) {};
+          \node[phantom morphism] (out_2) at (1, 4.5) {};
+          \node[phantom morphism] (out_3) at (2, 4.5) {};
+
+          \begin{knot}[clip width=8]
+            \strand[object, out=90, in=270] (in)
+              to (in |- f.north)
+              to (0, 3)
+              to (out_2);
+            \strand[object, in=-90, out=90] (f_left.north)
+              to (0, 2)
+              to (1, 3)
+              to (out_1);
+            \strand[object, in=-90, out=90]
+              (f_right.north)
+              to (2, 2.5)
+              to (out_3);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}
+          \node[phantom morphism] (in) at (2, 0) {};
+          \node[phantom morphism] (f_left) at (0, 1) {};
+          \node[phantom morphism] (f_right) at (1, 1) {};
+          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+          \node[phantom morphism] (out_1) at (0, 4.5) {};
+          \node[phantom morphism] (out_2) at (1, 4.5) {};
+          \node[phantom morphism] (out_3) at (2, 4.5) {};
+
+          \begin{knot}[clip width=8]
+            \strand[object, out=90, in=270] (in)
+              to (in |- f.north)
+              to (1, 2.5)
+              to (out_2);
+            \strand[object, out=90, in=270] (f_left) to (out_1);
+            \strand[object, out=90, in=270] (f_right) to (2, 2.5) to (out_3);
+          \end{knot}
+        \end{tikzpicture} } }
+      \]
+      In step one we apply naturality of $\sigma^{-1}$, in step two we apply one hexagon equality and in step three we cancel inverses.
+      Note that we implicitly use the equality from (iii) to justify that we cross the invisible unit object below the box with another object in step one.
+
+    \item
+      \[
+        \vcenter{ \hbox{ \begin{tikzpicture}
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90]
+              (0, 0)
+              to (1, 1)
+              to (2, 2)
+              to (2, 3);
+            \strand[object, in=-90, out=90]
+              (1, 0)
+              to (0, 1)
+              to (0, 2)
+              to (1, 3);
+            \strand[object, in=-90, out=90]
+              (2, 0)
+              to (2, 1)
+              to (1, 2)
+              to (0, 3);
+          \end{knot}
+
+          % \draw[highlight box] (-0.2, 0) rectangle (2.2, 2);
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[use Hobby shortcut]
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90]
+              (0, 0)
+              to (2, 2)
+              to (2, 3);
+            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
+              (1, 0)
+              .. (0.9, 1)
+              .. (0, 2)
+              to (1, 3);
+            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
+              (2, 0)
+              .. (1.1, 1)
+              .. (1, 2)
+              to (0, 3);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[use Hobby shortcut]
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90]
+              (0, 0)
+              to (0, 1)
+              to (2, 3);
+            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
+              (1, 0)
+              to (2, 1)
+              .. (1.1, 2)
+              .. (1, 3);
+            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
+              (2, 0)
+              to (1, 1)
+              .. (0.9, 2)
+              .. (0, 3);
+          \end{knot}
+
+          % \draw[highlight box] (-0.2, 1) rectangle (2.2, 3);
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90]
+              (0, 0)
+              to (0, 1)
+              to (1, 2)
+              to (2, 3);
+            \strand[object, in=-90, out=90]
+              (1, 0)
+              to (2, 1)
+              to (2, 2)
+              to (1, 3);
+            \strand[object, in=-90, out=90]
+              (2, 0)
+              to (1, 1)
+              to (0, 2)
+              to (0, 3);
+          \end{knot}
+        \end{tikzpicture} } }
+      \]
+      We apply one hexagon equality in step one, naturality of $\sigma$ in step two and the other hexagon equality in step three.
+
+    \item
+      \[
+        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90] (-1, 0) to (1, 4.5);
+            \strand[object, dashed, in=-90, out=90] (1, 0) to (-1, 4.5);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90]
+              (0, 0)
+              to (0, 4.5);
+            \strand[object, dashed, in=-90, out=90]
+              (1, 0)
+              to (-1, 1.5)
+              to (1, 3)
+              to (-1, 4.5);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90]
+              (0, 0)
+              to (0, 4.5);
+            \strand[object, dashed, in=-90, out=90]
+              (1, 0)
+              to (-1, 1.4)
+              to (-1, 1.6)
+              to (1, 3.1)
+              (-1, 1.4)
+              to (1, 2.9)
+              to (1, 3.1)
+              to (-1, 4.5);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
+          \begin{knot}[clip width=8]
+            \strand[object, in=-90, out=90]
+              (0, 0)
+              to (0, 4.5);
+            \strand[object, dashed, in=-90, out=90]
+              (1, 0)
+              to (-1, 1)
+              to (-1, 2)
+              to (1, 3.5)
+              (-1, 1)
+              to (1, 2.5)
+              to (1, 3.5)
+              to (-1, 4.5);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
+          \begin{knot}[clip width=8]
+            \strand[object]
+              (0, 0)
+              to (0, 4.5);
+            \strand[object, dashed, dashed, in=-90, out=90]
+              (1, 0)
+              to (-1, 1)
+              to (1, 2)
+              to[in=0] (0, 2.5)
+              (0, 2)
+              to[out=180] (-1, 2.5)
+              to (1, 3.5)
+              to (-1, 4.5);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
+          \begin{knot}[clip width=8]
+            \strand[object]
+              (0, 0)
+              to (0, 4.5);
+            \strand[object, dashed, in=-90, out=90]
+              (1, 0)
+              to[in=-45] (0, 3)
+              (0, 1.5)
+              to[out=135] (-1, 4.5);
+          \end{knot}
+        \end{tikzpicture} } }
+        =
+        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
+          \begin{knot}[clip width=8]
+            \strand[object]
+              (0, 0)
+              to (0, 4.5);
+            \strand[object, dashed, in=-90, out=90]
+              (1, 0)
+              to[in=-45] (0, 1.5)
+              (0, 3)
+              to[out=135] (-1, 4.5);
+          \end{knot}
+        \end{tikzpicture} } }
+      \]
+      We introduce $\sigma$ with it's inverse in step one.
+      We apply naturality of $\sigma^{-1}$ in step two and one hexagon equation in step three.
+      The rest follows from the correctness of the graphical calculus for general monoidal categories and cancellation of inverses.
+    \end{enumerate}
+  \item
+    \begin{enumerate}
+    \item
+      \[ \begin{tikzcd}[column sep=0.5cm]
+        & B
+          \arrow[ld, "\rho_B^{-1}"']
+          \arrow[rd, "\lambda_B^{-1}"] \\
+        %
+        B \otimes I
+          \arrow[rr, "\sigma_{B, I}"]
+          \arrow[d, "1_B \otimes f"] &&
+        I \otimes B
+          \arrow[d, "f \otimes 1_B"] \\
+        %
+        B \otimes (A \otimes C)
+          \arrow[rr, "\sigma_{B, A \otimes C}"]
+          \arrow[d, "\alpha_{B, A, C}^{-1}"] &&
+        (A \otimes C) \otimes B
+          \arrow[d, "\alpha_{A, C, B}"] \\
+        %
+        (B \otimes A) \otimes C
+          \arrow[d, "\sigma_{B, A} \otimes 1_C"] &&
+        A \otimes (C \otimes B)
+          \arrow[d, "1_A \otimes \sigma_{C, B}"] \\
+        %
+        (A \otimes B) \otimes C
+          \arrow[rr, "\alpha_{A, B, C}"] &&
+        A \otimes (B \otimes C)
+      \end{tikzcd} \]
+      The left outer path describes the first diagram and the right outer path describes the second diagram.
+      From top to bottom, the parts commute by the equality from (iii), naturality of $\sigma$ and one hexagon equality.
+
+    \item
+      \[ \begin{tikzcd}[column sep=1.2cm]
+        (A \otimes B) \otimes C
+          \arrow[r, "\alpha_{A, B, C}"]
+          \arrow[d, "\sigma_{A, B} \otimes 1_C"] &
+        A \otimes (B \otimes C)
+          \arrow[d, "1_A \otimes \sigma_{B, C}"] \\
+        %
+        (B \otimes A) \otimes C
+          \arrow[d, "\alpha_{B, A, C}"] &
+        A \otimes (C \otimes B)
+          \arrow[d, "\alpha_{A, C, B}^{-1}"] \\
+        %
+        B \otimes (A \otimes C)
+          \arrow[r, "\sigma_{B, A \otimes C}"]
+          \arrow[d, "1_B \otimes \sigma_{A, C}"] &
+        (A \otimes C) \otimes B
+          \arrow[d, "\sigma_{A, C} \otimes 1_B"] \\
+        %
+        B \otimes (C \otimes A)
+          \arrow[r, "\sigma_{B, C \otimes A}"]
+          \arrow[d, "\alpha_{B, C, A}^{-1}"] &
+        (C \otimes A) \otimes B
+          \arrow[d, "\alpha_{C, A, B}"] \\
+        %
+        (B \otimes C) \otimes A
+          \arrow[d, "\sigma_{B, C} \otimes 1_A"] &
+        C \otimes (A \otimes B)
+          \arrow[d, "1_C \otimes \sigma_{A, B}"] \\
+        %
+        (C \otimes B) \otimes A
+          \arrow[r, "\alpha_{C, B, A}"] &
+        C \otimes (B \otimes A)
+      \end{tikzcd} \]
+      The left outer path describes the first diagram and the right outer path describes the second diagram.
+      The parts of the diagram commute by the hexagon equalities and naturality of $\sigma$.
+
+    \item
+      \[ \begin{tikzcd}
+        I \otimes A
+          \arrow[rrrrrr, bend left, "\sigma"]
+          \arrow[r, "\sigma"]
+          \arrow[rrdd, "\rho^{-1}"']
+          \arrow[rrrddd, bend right, "\lambda"'] &
+        A \otimes I
+          \arrow[rrrr, bend left, "\sigma"]
+          \arrow[r, "1 \otimes \lambda^{-1}"]
+          \arrow[rd, "\rho^{-1}"'] &
+        A \otimes (I \otimes I)
+          \arrow[rr, "\sigma"]
+          \arrow[d, "\alpha^{-1}"] &&
+        (I \otimes I) \otimes A
+          \arrow[d, "\alpha"'] &
+        I \otimes A
+          \arrow[l, "\lambda^{-1} \otimes 1"']
+          \arrow[ld, "\lambda^{-1}"] &
+        A \otimes I
+          \arrow[l, "\sigma"']
+          \arrow[lldd, "\lambda^{-1}"]
+          \arrow[lllddd, bend left, "\rho"] \\
+        %
+        && (A \otimes I) \otimes I
+          \arrow[d, "\sigma \otimes 1"] &&
+        I \otimes (I \otimes A)
+          \arrow[d, "1 \otimes \sigma"'] \\
+        %
+        && (I \otimes A) \otimes I
+          \arrow[rr, "\alpha"] &&
+        I \otimes (A \otimes I) \\
+        %
+        &&& A
+      \end{tikzcd} \]
+      The rectangle in the middle represents one hexagon equality.
+      The section above it commutes by naturality of $\sigma$.
+      The squares to it's right and left commute by naturality of $\lambda$ and $\rho$.
+      Everything else commutes trivially or by the coherence theorem for general monoidal categories.
+    \end{enumerate}
+  \end{enumerate}
+\end{answer}
+
 \begin{exercise}
@@
 \end{exercise}
+
+\begin{answer}
+  You can define the associator, unitors and symmetry by composition of the corresponding morphisms for products and coproducts as well as distributivity.
+  As compositions of natural isomorphisms they will also be natural isomorphisms.
+  You can then check the triangle, pentagon and hexagon equalities and show that the symmetry is self-inverse (Application of a proof assistant adviced).
+
+  Another approach is to look at a category that includes $\Set$ as subcategory but makes the monoid structure more obvious.
+  We observe that $I$ and $\otimes$ coincide with the terminal object and the product functor from the category of sets with partial functions.
+  Liker every terminal object and product, they give rise to a monoidal structure.
+  Since the associator, unitors and symmetry are isomorphisms, they are total and contained in $\Set$.
+  Thus the monoidal structure exists also $\Set$.
+\end{answer}
 
@@
 \end{exercise}
+
+\begin{answer}
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (epsilon_left) at (0, 1) {};
+      \node[phantom morphism] (epsilon_right) at (1, 1) {};
+      \node[morphism, fit=(epsilon_left) (epsilon_right), inner sep=0] (epsilon) {};
+      \node at (epsilon) {$\varepsilon$};
+
+      \coordinate (in1) at (0, -0.5);
+      \coordinate (in2) at (1, -0.5);
+      \draw[object] (in1) to (epsilon_left);
+      \draw[object] (in2) to (epsilon_right);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (epsilon_left) at (0, 1) {};
+      \node[phantom morphism] (epsilon_right) at (1, 1) {};
+      \node[morphism, fit=(epsilon_left) (epsilon_right), inner sep=0] (epsilon) {};
+      \node at (epsilon) {$\varepsilon$};
+
+      \node[phantom morphism] (eta_left) at (1, 0) {};
+      \node[phantom morphism] (eta_right) at (2, 0) {};
+      \node[morphism, fit=(eta_left) (eta_right), inner sep=0] (eta) {};
+      \node at (eta) {$\eta$};
+
+      \node[phantom morphism] (epsilon'_left) at (2, 1) {};
+      \node[phantom morphism] (epsilon'_right) at (3, 1) {};
+      \node[morphism, fit=(epsilon'_left) (epsilon'_right), inner sep=0] (epsilon') {};
+      \node at (epsilon') {$\varepsilon'$};
+
+      \coordinate (in1) at (0, -0.5);
+      \coordinate (in2) at (3, -0.5);
+      \draw[object] (in1) to (epsilon_left);
+      \draw[object] (eta_left) to (epsilon_right);
+      \draw[object] (eta_right) to (epsilon'_left);
+      \draw[object] (in2) to (epsilon'_right);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (epsilon'_left) at (2, 1) {};
+      \node[phantom morphism] (epsilon'_right) at (3, 1) {};
+      \node[morphism, fit=(epsilon'_left) (epsilon'_right), inner sep=0] (epsilon') {};
+      \node at (epsilon') {$\varepsilon'$};
+
+      \coordinate (in1) at (2, -0.5);
+      \coordinate (in2) at (3, -0.5);
+      \draw[object] (in1) to (epsilon'_left);
+      \draw[object] (in2) to (epsilon'_right);
+    \end{tikzpicture} } }
+  \]
+  In the first step we apply the snake equation for $\varepsilon$ and in the second step the snake equation for $\varepsilon'$.
+\end{answer}
 
@@
 \end{exercise}
+
+\begin{answer}
+  We only show the equivalence $L \otimes 0 \iso 0$.
+  The other equivalences are analogous.
+
+  We define functions in both directions by
+  \[ \begin{tikzpicture}
+    \node[phantom morphism] (f_left) at (1, 1) {};
+    \node[phantom morphism] (f_right) at (2, 1) {};
+    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+    \node at (f) {$!$};
+
+    \node[phantom morphism] (L_mid) at (0, 1) {};
+
+    \coordinate (in_L) at (0, 0);
+    \coordinate (in_0) at (1.5, 0);
+    \coordinate (out) at (2, 2);
+
+    \draw[object] (in_L)
+      to node[auto] {$L$} (L_mid.south)
+      to (L_mid.north)
+      to[cap object] (f_left);
+    \path (f_left) to node[auto, swap] {$R$} (f_left |- out);
+    \draw[object] (in_0) to node[auto] {$0$} (f);
+    \draw[object] (f_right) to node[auto, swap] {$0$} (out);
+  \end{tikzpicture} \]
+  and
+  \[ \begin{tikzpicture}
+    \node[phantom morphism] (f_left) at (0, 0) {};
+    \node[phantom morphism] (f_right) at (1, 0) {};
+    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+    \node at (f) {$!$};
+
+    \draw[object] (0.5, -1) to node[auto] {$0$} (f);
+    \draw[object] (f_left) to node[auto] {$L$}  (0, 1);
+    \draw[object] (f_right) to node[auto] {$0$}  (1, 1);
+  \end{tikzpicture}. \]
+  In one direction they are trivially inverse because morphisms to $0$ are unique.
+  In the other direction it follows by
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (f_left) at (1, 1) {};
+      \node[phantom morphism] (f_right) at (2, 1) {};
+      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+      \node at (f) {$!$};
+
+      \node[phantom morphism] (g_left) at (1.5, 2) {};
+      \node[phantom morphism] (g_right) at (2.5, 2) {};
+      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
+      \node at (g) {$!$};
+
+      \node[phantom morphism] (L_mid) at (0, 1) {};
+
+      \coordinate (in_L) at (0, 0);
+      \coordinate (in_0) at (1.5, 0);
+      \coordinate (out_L) at (1.5, 3);
+      \coordinate (out_0) at (2.5, 3);
+
+      \draw[object] (in_L)
+        to node[auto] {$L$} (L_mid.south)
+        to (L_mid.north)
+        to[cap object] (f_left);
+      \path (f_left) to node[auto, swap] {$R$} (f_left |- g.south);
+      \draw[object] (in_0) to node[auto] {$0$} (f);
+      \draw[object] (f_right) to node[auto, swap] {$0$} (g);
+      \draw[object] (g_left) to node[auto] {$L$} (out_L);
+      \draw[object] (g_right) to node[auto] {$0$} (out_0);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (f_left) at (1, 1) {};
+      \node[phantom morphism] (f_right) at (2.5, 1) {};
+      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+      \node at (f) {$!$};
+
+      \node[phantom morphism] (g_left) at (2, 2) {};
+      \node[phantom morphism] (g_right) at (3, 2) {};
+      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
+      \node at (g) {$!$};
+
+      \node[phantom morphism] (L_mid_1) at (0, 1) {};
+      \node[phantom morphism] (L_mid_2) at (0, 2) {};
+      \node[phantom morphism] (R_mid) at (1, 2) {};
+
+      \coordinate (in_L) at (0, 0);
+      \coordinate (in_0) at (1.75, 0);
+      \coordinate (out_L) at (2, 3);
+      \coordinate (out_0) at (3, 3);
+
+      \draw[object] (in_L)
+        to node[auto] {$L$} (L_mid_1.south)
+        to (L_mid_2.north)
+        to[cap object] (R_mid)
+        to (f_left);
+      \path (R_mid) to node[auto, swap] {$R$} (R_mid |- out_L);
+      \draw[object] (in_0) to node[auto] {$0$} (f);
+      \draw[object] (f_right) to node[auto] {$0$} (g);
+      \draw[object] (g_left) to node[auto, swap] {$L$} (out_L);
+      \draw[object] (g_right) to node[auto, swap] {$0$} (out_0);
+
+      \node[highlight box, fit=(f) (g)] {};
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (f_left) at (1, 1) {};
+      \node[phantom morphism] (f_right) at (2.5, 1) {};
+      \node[phantom morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
+
+      \node[phantom morphism] (g_left) at (2, 2) {};
+      \node[phantom morphism] (g_right) at (3, 2) {};
+      \node[phantom morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
+
+      \node[phantom morphism] (L_mid_1) at (0, 1) {};
+      \node[phantom morphism] (L_mid_2) at (0, 2) {};
+      \node[phantom morphism] (R_mid) at (1, 2) {};
+      \node[phantom morphism] (L_mid_3) at (2, 2) {};
+
+      \coordinate (in_L) at (0, 0);
+      \coordinate (in_0) at (3, 0);
+      \coordinate (out_L) at (2, 3);
+      \coordinate (out_0) at (3, 3);
+
+      \draw[object] (in_L)
+        to node[auto] {$L$} (L_mid_1.south)
+        to (L_mid_2.south)
+        to (L_mid_2.north)
+        to[cap object] (R_mid)
+        to (R_mid.south)
+        to[cup object] (L_mid_3)
+        to (L_mid_3.north)
+        to node[auto, swap] {$L$} (out_L);
+      \path (R_mid) to node[auto, swap] {$R$} (R_mid |- out_L);
+      \draw[object] (in_0)
+        to node[auto] {$0$} (in_0 |- f.south)
+        to (out_0 |- g.south)
+        to (out_0 |- g.north)
+        to node[auto, swap] {$0$} (out_0);
+
+      \node[highlight box, fit=(f_left) (g_right)] {};
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (L_mid) at (0, 1) {};
+      \node[phantom morphism] (0_mid) at (1, 1) {};
+
+      \draw[object] (0, 0)
+        to node[auto] {$L$} (L_mid.south)
+        to (0, 3);
+      \draw[object] (1, 0)
+        to node[auto] {$0$} (0_mid.south)
+        to (1, 3);
+    \end{tikzpicture} } }.
+  \]
+  In the second step we use again that morphisms to $0$ are unique.
+\end{answer}
 
@@
 
+\begin{answer}
+  For clarity, we draw $\eta$ as
+  \[ \begin{tikzpicture}
+    \draw[object] (0, 0.5)
+      to node[auto, swap] {$R$} (0, 0)
+      to[cup object] (1, 0)
+      to node[auto] {$L$} (1, 0.5);
+    \end{tikzpicture} \]
+    and $\varepsilon$ as
+    \[ \begin{tikzpicture}
+      \draw[object] (0, -0.5)
+        to node[auto] {$L$} (0, 0)
+        to[cap object] (1, 0)
+        to node[auto, swap] {$R$} (1, -0.5);
+    \end{tikzpicture}. \]
+
+    Assume that we have the snake equality
+    \[ \vcenter{ \hbox{ \begin{tikzpicture}
+      \draw[object] (0, 2)
+        to (0, 1)
+        to[cup object] (1, 1)
+        to[cap object] (2, 1)
+        to (2, 0);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \draw (0, 2) to (0, 0);
+    \end{tikzpicture} } }. \]
+
+    Then the other snake is idempotent by
+    \[ \vcenter{ \hbox{ \begin{tikzpicture}
+      \draw[object] (0, -1)
+        to (0, 0)
+        to[cap object] (1, 0)
+        to[cup object] (2, 0)
+        to (2, 2)
+        to[cap object] (3, 2)
+        to[cup object] (4, 2)
+        to (4, 3);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \draw[object] (0, -1)
+        to (0, 2)
+        to[cap object] (1, 2)
+        to (1, 1)
+        { [local bounding box=inner snake]
+          node[phantom morphism] {}
+          to[cup object] (2, 1)
+          to[cap object] (3, 1)
+          node[phantom morphism] {}
+        }
+        to (3, 0)
+        to[cup object] (4, 0)
+        to (4, 3);
+      \node[highlight box, fit=(inner snake)] {};
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \draw[object] (0, -1)
+        to (0, 2)
+        to[cap object] (1, 2)
+        to (1, 1)
+        { [local bounding box=inner snake 2]
+          node[phantom morphism] {}
+        }
+        to (1, 0)
+        to[cup object] (2, 0)
+        to (2, 3);
+      % \node[highlight box, fit=(inner snake 2) (inner snake 2 |- inner snake.north) (inner snake 2 |- inner snake.south)] {};
+    \end{tikzpicture} } }.
+  \]
+  Hence there are morphisms $g$ and $h$ such that
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (out) at (0, 3) {};
+      \node[morphism] (g) at (0, 2) {$g$};
+      \node[morphism] (h) at (0, 1) {$h$};
+      \node[phantom morphism, overlay] (in) at (0, 0) {};
+      \draw[object] (in) to (h) to (g) to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (out) at (0, 3) {};
+      \node[phantom morphism, overlay] (in) at (0, 0) {};
+      \draw[object] (in) to (out);
+    \end{tikzpicture} } }
+  \]
+  and
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (out) at (0, 3) {};
+      \node[morphism] (h) at (0, 2) {$h$};
+      \node[morphism] (g) at (0, 1) {$g$};
+      \node[phantom morphism, overlay] (in) at (0, 0) {};
+      \draw[object] (in) to (g) to (h) to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (out) at (2, 3) {};
+      \node[phantom morphism, overlay] (in) at (0, 0) {};
+      \draw[object] (in)
+        to (0, 1.5)
+        to[cap object] (1, 1.5)
+        to[cup object] (2, 1.5)
+        to (out);
+    \end{tikzpicture} } }.
+  \]
+  We use those morphisms to define a unit
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (out_left) at (0, 1) {};
+      \node[phantom morphism, overlay] (out_right) at (1, 1) {};
+      \node[morphism] (g) at (1, 0) {$g$};
+      \draw[object] (out_left)
+        to node[auto, swap] {$R$} (0, 1 |- g.north)
+        to (0, 1 |- g.south)
+        to[cup object] (g.south)
+        (g.north) to node[auto] {$L$} (out_right);
+    \end{tikzpicture} } }
+  \]
+  and a counit
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[morphism] (h) at (0, 1) {$h$};
+      \node[phantom morphism, overlay] (in_left) at (0, 0) {};
+      \node[phantom morphism, overlay] (in_right) at (1, 0) {};
+      \draw[object] (in_left)
+        to node[auto] {$L$} (h)
+        to[cap object] (1, 0 |- h.north)
+        to (1, 0 |- h.south)
+        to node[auto, swap] {$R$} (in_right);
+    \end{tikzpicture} } },
+  \]
+  that exhibit a duality $L \dashv R$.
+  The snake equations hold by
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (out) at (-1, 2.5) {};
+      \node[morphism] (h) at (0, 1) {$h$};
+      \node[morphism] (g) at (0, 0) {$g$};
+      \node[phantom morphism] (in) at (1, -1.5) {};
+      \draw[object] (in)
+        to (in |- h.north)
+        to[cap object] (h.north)
+        (h.south) to (g)
+        to[cup object] (out |- g.south)
+        to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (out) at (-1, 2.5) {};
+      \node[phantom morphism] (h) at (2, 1) {};
+      \node[phantom morphism] (g) at (0, 0) {};
+      \node[phantom morphism] (in) at (3, -1.5) {};
+      \draw[object] (in)
+        to (in |- h.north)
+        to[cap object] (h.north)
+        to (2, 0.5)
+        to[cup object] (1, 0.5)
+        to[cap object] (0, 0.5)
+        to (g.south)
+        to[cup object] (out |- g.south)
+        to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (out) at (0, 4) {};
+      \node[phantom morphism] (in) at (2, 0) {};
+      \draw[object] (in)
+        to (2, 2)
+        to[cap object] (1, 2)
+        to[cup object] (0, 2)
+        to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (out) at (0, 4) {};
+      \node[phantom morphism] (in) at (0, 0) {};
+      \draw[object] (in) to (out);
+    \end{tikzpicture} } }
+  \]
+  and
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (in) at (0, -0.5) {};
+      \node[morphism] (h) at (0, 1) {$h$};
+      \node[morphism] (g) at (2, 3) {$g$};
+      \node[phantom morphism, overlay] (out) at (2, 4.5) {};
+      \draw[object] (in)
+        to (h)
+        to (0, 2)
+        to[cap object] (1, 2)
+        to[cup object] (2, 2)
+        to (g)
+        to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (in) at (0, 0) {};
+      \node[morphism] (h1) at (0, 1) {$h$};
+      \node[morphism] (g1) at (0, 2) {$g$};
+      \node[morphism] (h2) at (0, 3) {$h$};
+      \node[morphism] (g2) at (0, 4) {$g$};
+      \node[phantom morphism, overlay] (out) at (0, 5) {};
+      \draw[object] (in)
+        to (h1)
+        to (g1)
+        to (h2)
+        to (g2)
+        to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (in) at (0, 0) {};
+      \node[phantom morphism, overlay] (out) at (0, 5) {};
+      \draw[object] (in) to (out);
+    \end{tikzpicture} } }.
+  \]
+\end{answer}
+
 \end{document}

### 2017-07-26 papers/base-3/paper-3.tex
added some explanation to the motivating example

@@
 % \input{\pathToRoot headers/theoremenvts.tex}
+
+\usepackage{todonotes}
 
@@
 We now want to examine when two categories are the same.
+To that end let us consider the following example.
 \begin{exmp}
+  \label{exmp:eqv}
+  We take two discrete categories $\catA$ and $\catB$ and define two functors $F \from \catA \to \catB$ and $G \from \catB \to \catA$ as depicted below.
+  They operate on morphisms in the only possible way.
   \[
@@
       \catA:                                               & & \catB:\\
-      A \arrow[loop left] & & B \arrow[bend left]{d} \arrow[loop left] \\
-                          & & B' \arrow[bend left]{u} \arrow[loop left]
+      A \arrow[loop left] \arrow[dashed, bend left]{rr}{F} & & B \arrow[bend left]{d} \arrow[loop above] \arrow[dashed]{ll}{G}\\
+                                                           & & B' \arrow[bend left]{u} \arrow[loop below] \arrow[dashed]{llu}{G}
     \end{tikzcd}
   \]
+  The two objects $B$ and $B'$ of $\catB$ are isomorphic.
+  By "merging" them we obtain the category $\catA$.
+  We now examine how this merging can be described by $F$ and $G$.
+  We observe that $G \circ F = 1_{\catA}$.
+  But $F \circ G$ is only isomorphic to $1_{\catB}$.
+  This isomorphism between the composition of $F$ and $G$ leads to the following definition of equivalence.
 \end{exmp}
-todo: explain example
 
@@
   \label{def:eqv}
-  Tow categories $\catA$ and $\catB$ are \emph{equivalent} if:
+  Two categories $\catA$ and $\catB$ are \demph{equivalent} if:
   \begin{itemize}
@@
   \end{itemize}
-  The functors $F$ and $G$ are called \emph{equivalences}.
+  The functors $F$ and $G$ are called \demph{equivalences}.
   We write $\catA \eqv \catB$ if $\catA$ and $\catB$ are equivalent.
@@
   \label{def:surj-on-obj}
-  A functor $F \from \catA \to \catB$ is \emph{essentially surjective on objects}
+  A functor $F \from \catA \to \catB$ is \demph{essentially surjective on objects}
   if for every object $B \in \catB$ there is an object $A \in \catA$ such that $F(A) \iso B$.
@@
 \]
-Additionally we have $1_{F'} * 1_{F} = 1_{F' \circ F}$, where $1_{F}$ is the natural transformation which components are the identities of $\cat{A'}$.
+Additionally we have $1_{F'} * 1_{F} = 1_{F' \circ F}$, where $1_{F}, 1_{F'}$ and $1_{F' \circ F}$ are the natural transformations which components are the identities on $\cat{A'}, \cat{A'}$ and $\cat{A''}$.
 

### 2017-07-26 papers/base-3/paper-3.tex
Fix typo
Add missing $$
Fix some typos

--- 2017-07-25 papers/base-3/paper-3.tex
@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/header_lecturenotes}
-% \input{\pathToRoot headers/theoremenvts.tex}
 
@@
 
-For example, concider a discrete category $\catA$ and an arbitrary category $\catB$.
+For example, consider a discrete category $\catA$ and an arbitrary category $\catB$.
 Now let $\parpair{\catA}{\catB}{F}{G}$ be functors from $\catA$ to $\catB$.
@@
 
-We first define what a natural transformation in general is precisely:
+First we define precisely what a natural transformation is in general.
 \begin{defn}[Natural transformation]
@@
   $\bigl(F(a)\toby{\alpha_A}G(A)\bigr)_{A\in\catA}$.
-  Then we define that $\alpha$ is a \demph{natural transformation} from $F$ to $G$ iff
-  for every objects and maps $A, A' \in \catA, f \from A \to A'$, the sqare
+  Then we define that $\alpha$ is a \demph{natural transformation} from $F$ to $G$ if
+  for every objects and maps $A, A' \in \catA, f \from A \to A'$, the square
   \begin{equation}
@@
   The maps $\alpha_A$ are then called the \demph{components} of $\alpha$.
-  The commuting square (\ref{eq:nat}) is also called the \demph{naturality axiom} (throught this is not a axiom in logical sense but a defined property for some $\alpha$).
+  The commuting square (\ref{eq:nat}) is also called the \demph{naturality axiom} (through this is not a axiom in logical sense but a defined property for some $\alpha$).
 \end{defn}
@@
 
-\begin{exmp}[Discrete Categories]
-  For the first example, the mapping
-  $\bigl(F(A)\toby{\alpha_A}G(A)\bigr)_{A\in\catA}$
-  the naturality axiom (\ref{eq:nat}) holds, because $\catA$ is discrete, so every morphism $f \from A \to A'$ in $\catA$ is an identy morphism (hence $A=A'$). The diagramm
+\begin{exmp}[Discrete categories]
+  In the first example, the naturality axiom (\ref{eq:nat}) holds for the mapping
+  $\bigl(F(A)\toby{\alpha_A}G(A)\bigr)_{A\in\catA}$,
+  because $\catA$ is discrete, so every morphism $f \from A \to A'$ in $\catA$ is an identity morphism (hence $A=A'$). The diagram
   \[
@@
   i.~e. applying $f$ element-wise to the matrix $A$, we get another functor $M_n \from \CRing \to \Mon$.
-  We will now see that the \emph{determinant} is a natural transformation between this two functors $M_n$ and $U$.
+  We will now see that the \emph{determinant} is a natural transformation between $M_n$ and $U$.
 
   First we note that for every ring the determinant of a matrix over this ring is defined canonically.
-  Furthermore by definition if $A \in R^{n \times n}$ then $\det A \in R$ for every ring $R$.
+  Furthermore by definition $\det A \in R$ holds for $A \in R^{n \times n}$.
   We therefore can establish a family of mappings $\det$:
   $\bigl(M_n(R)\toby{\det_R}U(R)\bigr)_{R\in\CRing}$, because $U(R) = R$.
-  Now, to show that $\det$ is indeed a natural transformation, we have to show that the following diagramm commutes for every ring homomorphism $f \from R \to S$:
+  Now, to show that $\det$ is indeed a natural transformation, we have to show that the following diagram commutes for every ring homomorphism $f \from R \to S$:
   \[\xymatrix{
@@
   }\]
-  By unfolding the definitions of $U$ and $M_n$ and applying functional extentionality, this equivalent to
+  By unfolding the definitions of $U$ and $M_n$ and applying functional extensionality, this is equivalent to
   $f(\det_R(f(A))) = \det_S (\bigl( f(A_{i,j}) \bigr)_{1 \le i, j \le n})$
@@
   \[\xymatrix{
-    \CRing \rtwocell<5>^{M_n}_{U}{\alpha} & \Mon
+    \CRing \rtwocell<5>^{M_n}_{U}{\hspace{.7em}\det} & \Mon
   }\]
@@
 
+\section*{Functor Category}
 \begin{defn}[Functor Category]
-  Natural transformations, as other types of maps that we've encountered earlier
+  Natural transformations, as other types of maps that we have encountered earlier
   can be composed. That is, for any two categories $\cat{A}$ and $\cat{B}$,
@@
       \end{alignat*}
-      By similar reasoning to (a) we get that $\beta$ is the inverse of
-      $\alpha$.  Therefore $\alpha$ is a natural isomorpism.
+      By similar reasoning to ($\leftarrow$) we get that $\beta$ is the inverse
+      of $\alpha$.  Therefore $\alpha$ is a natural isomorphism.
   \end{itemize}
@@
 
-\section{Equivalence of Categories}
+\section*{Equivalence of Categories}
 We now want to examine when two categories are the same.
@@
 
-\section{Horizontal Composition}
+\section*{Horizontal Composition}
 

### 2017-07-26 papers/base-3/paper-3.tex
though

@@
   The maps $\alpha_A$ are then called the \demph{components} of $\alpha$.
-  The commuting square (\ref{eq:nat}) is also called the \demph{naturality axiom} (through this is not a axiom in logical sense but a defined property for some $\alpha$).
+  The commuting square (\ref{eq:nat}) is also called the \demph{naturality axiom} (though this is not a axiom in logical sense but a defined property for some $\alpha$).
 \end{defn}

### 2017-07-26 papers/base-3/paper-3.tex
an axiom

@@
   The maps $\alpha_A$ are then called the \demph{components} of $\alpha$.
-  The commuting square (\ref{eq:nat}) is also called the \demph{naturality axiom} (though this is not a axiom in logical sense but a defined property for some $\alpha$).
+  The commuting square (\ref{eq:nat}) is also called the \demph{naturality axiom} (though this is not an axiom in logical sense but a defined property for some $\alpha$).
 \end{defn}

### 2017-07-27 papers/base-3/paper-3.tex

@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/header_lecturenotes}
+
+\usepackage{todonotes}
 

### 2017-07-28 notes/notes-2/notes-2.tex
Small corrections.

@@
 \end{align*}
-due to the fact that for each pair of objects in a preorder there is either exactly one arrow between them (denoting that the first is smaller than the second) or none at all (denoting incomperatibility).
-Hence, functors between correspond to monotonically (increasing) functions.
+due to the fact that for each pair of objects in a preorder there is either exactly one arrow between them (denoting that the first is smaller than the second) or none at all (denoting incomparability).
+Hence, functors between preorders correspond to monotonically (increasing) functions.
 
@@
 \end{align*}
-for each $P,P'\in\ob(\cat P)$. Again, this exactly the same as to demand that for each $P,P'\in\ob(\cat P)$ there is a function
+for each $P,P'\in\ob(\cat P)$. Again, this is exactly the same as to demand that for each $P,P'\in\ob(\cat P)$ there is a function
 \begin{align*}
@@
 \end{align*}
-Since there is the unique arrow in $\cat (P,P')$ if and only if $\cat P^\op(P',P)$ contains the unique arrow we see that this amounts to the same as having a function
+Since there is the unique arrow in $\cat P(P,P')$ if and only if $\cat P^\op(P',P)$ contains the unique arrow we see that this amounts to the same as having a function
 \begin{align*}
@@
 \section{Dual Vector Spaces and Contravariant Representable Functors}
-Next, we are going to look at the generalization of a construction from linear algebra that is used in the study of bilinear mappings amongst others.
+Next, we are going to look at the generalization of a construction from linear algebra, which is used in the study of bilinear mappings amongst others and which can be regarded as a special contravariant functor.
 The general idea is that we analyze objects by analyzing the mappings into them.
@@
 \end{align*}
-of linear mappings between $V$ and $W$ where the vector space operations addition and scalar multiplication are defined pointwise.
+of linear mappings between $V$ and $W$ where the vector space operations (addition and scalar multiplication) are defined pointwise.
 Therefore, for a fixed $\mathbb K$-vector space $W$, we can define a mapping
@@
 It seems reasonable that we can extend this to a functor $H_W\from\Vect_{\mathbb K}\to\Vect_{\mathbb K}$.
-So let us check the details: We have to define its action on linear maps $f\from V\to V'$, that is we have to define a linear mapping
+So let us check the details: We still have to define its action on linear maps $f\from V\to V'$, that is we have to define a linear mapping
 \begin{align*}
@@
 \end{align*}
-that is moreover trivially linear. This is indeed a contravariant functor since
+that is moreover trivially linear. This is looks like a contravariant functor. To verify that it is indeed a functor we check the axioms:
 \begin{align*}
-  H_W(1_V)=\lambda g\ldotp g\of 1_V=\lambda g\ldotp g=1_{H_W(V)}\\
-  H_W(f'\of f)=\lambda g\ldotp g\of (f'\of f)=\lambda g\ldotp (g\of f')\of f=(\lambda g\ldotp g\of f)\of (\lambda g\ldotp g\of f')=H_W(f)\of H_W(f')
+  H_W(1_V)&=\lambda g\ldotp g\of 1_V=\lambda g\ldotp g=1_{H_W(V)}\\
+  H_W(f'\of f)=\lambda g\ldotp g\of (f'\of f)&=\lambda g\ldotp (g\of f')\of f=(\lambda g\ldotp g\of f)\of (\lambda g\ldotp g\of f')=H_W(f)\of H_W(f').
 \end{align*}
@@
 \begin{align*}
-  H_W(V)=\Hom(V,W)=\{f\from V\to W\mid f \text{ linear}\}
+  H_W(V)=\Hom(V,W)=\{f\from V\to W\mid f \text{ linear}\},
 \end{align*}
- is
+ that is its effect on objects, is
 \begin{align*}
@@
 \end{align*}
-for $A$ and $B$ being objects in a category $\cat A$.
-This suggests a functor $F_A\from\cat A^\op\to\Set$.
-However, this is only well-defined if all collections $\cat A(A,B)$ are really sets, which is not necessarily the case for arbitrary categories.
-Hence, in all what follows we assume a \emph{locally small} category $\cat A$ in which all collections $\cat A(A,B)$ are really sets.
+for $A,B\in\ob(\cat A)$.
+In contrast to the special case of vector spaces, $\cat A(A,B)$ is not a category again but just a collection.
+If all collections $\cat A(A,B)$ are sets this suggests a functor $F_A\from\cat A^\op\to\Set$.
+However, obviously this is only well-defined if all collections $\cat A(A,B)$ are really sets, which is not necessarily the case for arbitrary categories.
+Hence, in all what follows we assume a \emph{locally small} category $\cat A$ in which all collections $\cat A(A,B)$ are indeed sets.
 
@@
 \end{align*}
-These functors $H_A$ are called the \emph{contravariant representable functor} of $A$. Furthermore, contravariant functors to $\Set$ (like $H_A$) have a special name:
+The functor $H_A$ is called the \emph{contravariant representable functor} of $A$. Furthermore, contravariant functors to $\Set$ (like $H_A$) have a special name:
 \begin{definition}
@@
 \end{definition}
-In a later chapter we are going to discover that they really helps us understand the structure of locally small categories.
+In a later chapter we are going to discover that contravariant representable functors really help us understand the structure of locally small categories.
 In particular, we are going to prove that two objects $A$ and $B$ are isomorphic if and only if $H_A$ and $H_B$ are isomorphic\footnote{The notion of mappings between functors is defined in the next chapter.}, i.e.

### 2017-07-28 notes/notes-2/notes-2.tex
Do some minor clean-ups to the notes for talk 2

@@
 \documentclass{article}
+\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
 
-\usepackage[utf8]{inputenc}
-\usepackage{amsthm}
 \newtheorem{lemma}{Lemma}
@@
 
-% \beamertemplatenavigationsymbolsempty
-% \setbeamertemplate{footline}{
-% 	\raisebox{0.5em}{\makebox[\paperwidth-0.5em][r]{
-% 		\color{gray}\scriptsize\insertframenumber}}}
-
-\usepackage{mathtools}
-\usepackage{tikz}
-\usetikzlibrary{cd, shapes, tikzmark}
-\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
-
-\title{Presheaves, Representables and\\the Yoneda Lemma}
-\author{Dominik Wagner}
-%\date{June 7, 2017}
+\usetikzlibrary{cd}
 
 \begin{document}
+\title{Presheaves, Representables and\\the Yoneda Lemma}
+\author{Dominik Wagner, Felix Rech}
+\maketitle
 
 \begin{definition}
-  Let $\cat A$ and $\cat B$ be categories. A \emph{functor} $\from \cat A\to\cat B$ consists of
+  Let $\cat A$ and $\cat B$ be categories. A \demph{functor} $F \from \cat A\to\cat B$ consists of
     \begin{itemize}
@@
     \begin{itemize}
-    \item $F(1_A)=1_{F(A)}$ whenever $A\in\cat A$ and
-    \item $F(f'\of f)=F(f')\of F(f)$ whenever $A\xrightarrow{f}A'\xrightarrow{f'}A''$ in $\cat A$.
+    \item $F(1_A)=1_{F(A)}$ for all objects $A$ and
+    \item $F(f'\of f)=F(f')\of F(f)$ for all morphisms $A\xrightarrow{f}A'\xrightarrow{f'}A''$.
     \end{itemize}
@@
   \label{ex:Mset}
-  Let $\cat M$ be a category with a single object $M$, i.e.
-a monoid.
+  Let $\cat M$ be a category with a single object $M$, i.e.\ a monoid.
 Then, a functor $F\from\cat M\to\Set$ is a set $S=F(M)$ together with functions $F(m)\from S\to S$ for $m\from M\to M$ in $\cat M$.
@@
   \begin{align*}
-    \cdot\from M\times S&\to S\\
+    {} \cdot {} \from M\times S&\to S\\
     (m,s)&\mapsto m\cdot s
@@
 
-A very natural question to ask is whether there is also a general category theoretic analogue to monotonically decreasing functions, i.e.
-functions $F\from\ob(\cat P)\to\ob(\cat Q)$ between preorders $\cat P$ and $\cat Q$ such that
+A very natural question to ask is whether there is also a general category theoretic analogue to monotonically decreasing functions, i.e.\ functions $F\from\ob(\cat P)\to\ob(\cat Q)$ between preorders $\cat P$ and $\cat Q$ such that
 \begin{align*}
@@
 \begin{definition}
-  Let $\cat A$ and $\cat B$ be categories. A \emph{contravariant functor} from $\cat A$ to $\cat B$ is a functor $\cat A^\op\to\cat B$.
+  Let $\cat A$ and $\cat B$ be categories. A \demph{contravariant functor} from $\cat A$ to $\cat B$ is a functor $\cat A^\op\to\cat B$.
 \end{definition}
 \section{Dual Vector Spaces and Contravariant Representable Functors}
-Next, we are going to look at the generalization of a construction from linear algebra, which is used in the study of bilinear mappings amongst others and which can be regarded as a special contravariant functor.
+Next, we are going to look at the generalization of a construction from linear algebra which is used in the study of bilinear mappings amongst others and which can be regarded as a special contravariant functor.
 The general idea is that we analyze objects by analyzing the mappings into them.
@@
 \begin{definition}
-  Let $\cat A$ be a category. A \emph{presheaf} on $\cat A$ is a functor $\cat A^\op\to\Set$.
+  Let $\cat A$ be a category. A \demph{presheaf} on $\cat A$ is a functor $\cat A^\op\to\Set$.
 \end{definition}
@@
 \begin{example}
-  Similarly, as in example \ref{ex:Mset} presheaves on one-object categories (i.e. monoids $M$) correspond to \emph{right} $M$-sets, where a right $M$-set is a set $S$ together with a function
+  Similarly, as in example \ref{ex:Mset} presheaves on one-object categories (i.e.\ monoids $M$) correspond to \emph{right} $M$-sets, where a right $M$-set is a set $S$ together with a function
   \begin{align*}
-    \cdot\from S\times M&\to S\\
+    {} \cdot {} \from S\times M&\to S\\
     (s,m)&\mapsto s\cdot m

### 2017-07-28 notes/notes-2/notes-2.tex
Extend notes for talk 2
Add sections about full/faithful functors and subcategories and many
basic examples, mostly about monoids.

@@
 \maketitle
+
+We want to define structure-preserving maps between categories.
+Since the morphisms are are most interesting part of a category, such a map should not just operate on objects, but on morphisms to.
+In order to preserve the structure, it has to map identity morphisms to identity morphisms and compositions of morphisms to compositions.
+This is made precise in the following definition.
 
@@
 
-Next, we consider a further example of a $\Set$-valued functor, which unlike the previous ones is not forgetful.
+The image of such a functor can be thought of as a \textit{shadow} of category $\cat{A}$ in $\cat{B}$.
+Sometimes we call the first function \demph{action on paths} and the second function \demph{action on morphisms}.
+
+We can easily define composition of functors and an identity functor on every category.
+With those we define a category $\Cat$ that contains categories as objects and functors as morphisms.
+
+Let us look at some examples of functors.
+We recommend to check the axioms yourself in order to verify that they indeed satisfy our definition.
+
+\begin{example}
+  We define a \demph{powerset functor} $\Set \xrightarrow{\mathcal{P}} \Set$ by
+  \begin{align*}
+    \mathcal{P}(A) &\coloneqq \{ X \mid X \subseteq A \} \\
+    \mathcal{P}(f) &\coloneqq (\{x, y, \ldots\} \mapsto \{f(x), f(y), \ldots\})
+  \end{align*}
+  for all objects $A$ and $B$ and morphisms $A \xrightarrow{f} B$.
+  In a similar way we can define a functor that maps a set $A$ to the set of lists or to the set of multisets with elements in $A$ or to many other types of collections.
+  In all those cases the action on morphisms allows us to apply a function element-wise on a collection.
+  If you have experience in functional programing, this may seem familiar to you.
+\end{example}
+
+In the next few examples we will talk about a vary simple type of algebraic structures with a binary operation that we now define.
+
+\begin{definition}
+  A set $M$ is a \demph{monoid} if it has a binary operation ${}*{} \from M \times M \to M$ that satisfies the following axioms.
+  \begin{description}
+    \item [Associativity] $(x * y) * z = x * (y * z)$ for all $x, y, z \in M$.
+    \item [Neutral element] There is an element $e \in M$ such that $e * x = x = x * e$ for all $x \in M$.
+  \end{description}
+\end{definition}
+
+This is the same as the definition of \emph{groups} except that we don't require inverses.
+Similar to homomorphisms on groups, we define homomorphisms on monoids.
+
+\begin{definition}
+  Given two monoids $M$ and $N$, a function $M \xrightarrow{f} N$ is a \demph{monoid homomorphism} if it satisfies the axioms
+  \begin{itemize}
+    \item $f(e) = e$
+    \item $f(x * y) = f(x) * f(y)$
+  \end{itemize}
+  where we write $e$ for the neutral element in both monoids and $*$ for the binary operation in both monoids.
+\end{definition}
+
+You might already spot some similarity to the functoriality axioms.
+We use this in our next example.
+
+\begin{example}
+  There is a one-to-one correspondence between monoids and categories with one object.
+  Elements of the monoid correspond to morphisms in the category and the binary operation is interpreted as composition of morphisms.
+  With this interpretation, the functoriality axioms are the same as the axioms for monoid homomorphisms, which means that the homomorphisms between two groups are exactly the functors between the representing categories.
+\end{example}
+
+\begin{example}
+  Monoids together with monoid homomorphisms form a category.
+  We can define a functor from the category of monoids to $\Set$, that maps every monoid to its element set and every homomorphism to the undelying function.
+  This functor maps into a category that is similar to its domain but with less structure on the objects.
+  Hence one can say that the functor \emph{forgets} the monoid structure.
+  This can be observed on many functors.
+  We call such functors \demph{forgetful}.
+  Forgetfulness is an informal concept and depends on our interpretation of the category, but usually it is obvious if you should call a functor forgetful.
+\end{example}
+
+\begin{example}
+  There is a functor from the category of commutative monoids to the category of general monoids which maps every object and every morphism to itself.
+  It is injective on objects and on arrows, which means that we can always say where a resulting object \textit{came from}.
+  But we still call it \emph{forgetful} because it maps into a category where in general the objects have less structure.
+\end{example}
+
+We saw two examples of functors that forget structure, but there are also functors that do the opposite and create structure.
+We will take a look at one of them.
+
+\begin{example}
+  Given a set $A$, we define the \demph{free monoid} $A^*$ over $A$ as the set of lists with elements in $A$.
+  As binary operation we use the concatenation of lists and as neutral element the empty list.
+  If we have a function $A \xrightarrow{f} B$ then we can lift it to a monoid homomorphism $A^* \xrightarrow{f^*} B^*$ that applies $f$ pointwise on a list, i.e.\ $f^*([x, y, \ldots]) = [f(x), f(y), \ldots]$.
+  This gives us a functor that maps a set to its free monoid and hence \emph{adds structure}.
+  In general we call such a functor a \demph{free functor}.
+\end{example}
+
+Next, we consider a further example of a $\Set$-valued functor which, unlike the previous one, is not forgetful.
 \begin{example}
@@
 
+\section{Full and Faithful functors}
+
+The following definition introduces a concept that is similar to injectivity of functors but restricts only the action on morphisms, not on objects.
+
+\begin{definition}
+  A functor $\cat{A} \xrightarrow{F} \cat{B}$ is \demph{faithful} if for all objects $A, B \in \cat{A}$ the function
+  \begin{align*}
+    \cat{A}(A, B) &\to \cat{B}(F(A), F(B)) \\
+    f &\mapsto F(f)
+  \end{align*}
+  is \emph{injective}.
+\end{definition}
+
+Note that this definition only speaks about what heapens to morphisms between the \emph{same pair} of objects.
+For morphisms $A \xrightarrow{f, g} B$, we can conclude that $F(A) = F(B)$ implies $A = B$ but for a morphism $A \xrightarrow{f} B$ and a morphism $C \xrightarrow{g} D$ between another pair of objects, the faithfulness of $F$ doesn't tell us anything.
+Let us consider a functor like in the following graphic as an example.
+\[ \begin{tikzpicture}
+    \draw (0, 0) rectangle (2, 3);
+    \node (A) at (0.5, 2.5) {$A$};
+    \node (C) at (1, 0.5) {$C$};
+    \node (B) at (1.5, 2.5) {$B$};
+    \draw[->] (A) to node[auto, swap] {$f$} (C);
+    \draw[->] (B) to node[auto] {$g$} (C);
+
+    \draw (3, 0) rectangle (5, 3);
+    \node (X) at (4, 2.5) {$X$};
+    \node (Y) at (4, 0.5) {$Y$};
+    \draw[->] (X) to node[auto] {$u$} (Y);
+
+    \draw[->, dotted, out=45, in=120] (A) to (X);
+    \draw[->, dotted, out=45, in=120] (B) to (X);
+    \draw[->, dotted, out=-45, in=-120] (C) to (Y);
+  \end{tikzpicture} \]
+Assume that a functor $F$ is given by the dotted arrows and maps the morphism $f$ and $g$ both to $u$.
+The functor is faithful nevertheless because the morphisms $f$ and $g$ start from different objects.
+
+Analogously to the definition of faithfulness, we define fullness of a functor, which is similar to the concept of surjectivity.
+
+\begin{definition}
+  A functor $\cat{A} \xrightarrow{F} \cat{B}$ is \demph{full} if for all objects $A, B \in \cat{A}$ the function
+  \begin{align*}
+    \cat{A}(A, B) &\to \cat{B}(F(A), F(B)) \\
+    f &\mapsto F(f)
+  \end{align*}
+  is \emph{surjective}.
+\end{definition}
+
+\section{Subcategories}
+
+With the definition of full and faithful functors we adapted two set theoretic concepts for our work with categories.
+The same can be done with the subset relation.
+This time there is one obvious way to do it which is the following definition.
+
+\begin{definition}
+  Given two categories $\cat{A}$ and $\cat{B}$, we say that $\cat{A}$ is a \demph{subcategory} of $\cat{B}$ if $\ob(\cat{A}) \subseteq \ob(\cat{B})$ and for all objects $A, B \in \cat{A}$, we have the inclusion $\cat{A}(A, B) \subseteq \cat{B}(A, B)$.
+\end{definition}
+
+If we consider the inclusion functor $\cat{A} \to \cat{B}$ that maps every object and every morphism to itself, we can ask the question, whether it is full or faithful.
+The faithfulness is obvious, but only certain subcategories have full inclusion functors.
+This motivates the definition of full subcategories.
+
+\begin{definition}
+  A subcategory $\cat{A}$ of a category $\cat{B}$ is \demph{full} if for all objects $A, B \in \cat{A}$, there is an equality $\cat{A}(A, B) = \cat{B}(A, B)$, not just an inclusion.
+\end{definition}
+
+With our definition, a subcategory is full if and only if the inclusion functor is full.
+An important property is that a full subcategory is completely determined by the selection of objects that it contains.
+
+As the image of a monoid homomorphism is a monoid and the image of a group homomorphism is a group, one could easily assume that the image of a functor is a category, but in general this is wrong as we can see in this example:
+\[ \begin{tikzpicture}
+  \draw (0, 0) rectangle (3, 5);
+  \node (A) at (1, 4.5) {$A$};
+  \node (B) at (1, 2.5) {$B$};
+  \node (C) at (2, 2.5) {$C$};
+  \node (D) at (2, 0.5) {$D$};
+  \draw[->] (A) to node[auto] {$f$} (B);
+  \draw[->] (C) to node[auto] {$g$} (D);
+
+  \draw (4, 0) rectangle (7, 5);
+  \node (X) at (5, 4.5) {$X$};
+  \node (Y) at (5, 2.5) {$Y$};
+  \node (Z) at (5, 0.5) {$Z$};
+  \draw[->] (X) to node[auto] {$u$} (Y);
+  \draw[->] (Y) to node[auto] {$v$} (Z);
+  \draw[->, bend left=45] (X) to node[auto] {$v \of u$} (Z);
+
+  \draw[->, dotted, in=120, out=45] (A) to (X);
+  \draw[->, dotted, in=120, out=45] (B) to (Y);
+  \draw[->, dotted, in=-120, out=-45] (C) to (Y);
+  \draw[->, dotted, in=-120, out=-45] (D) to (Z);
+\end{tikzpicture} \]
+Assume that a functor $F$ is given by the dotted arrows and maps the morphism $f$ to $u$ and $g$ to $v$.
+Since the image of $F$ contains $X \xrightarrow{u} Y$ and $Y \xrightarrow{i} Z$, it should also contain $v \of u$ to be a category.
+But $F$ does not map any morphism to $v \of u$, thus the image is not a category.
+
 \end{document}

### 2017-07-28 notes/notes-2/notes-2.tex
Fixed some obvious bugs.

@@
 \begin{document}
-\title{Presheaves, Representables and\\the Yoneda Lemma}
-\author{Dominik Wagner, Felix Rech}
+\title{Maps between Categories -- Functors}
+\author{Felix Rech \and Dominik Wagner}
 \maketitle
@@
 We want to define structure-preserving maps between categories.
-Since the morphisms are are most interesting part of a category, such a map should not just operate on objects, but on morphisms to.
+Since the morphisms are most interesting part of a category, such a map should not just operate on objects, but on morphisms, too.
 In order to preserve the structure, it has to map identity morphisms to identity morphisms and compositions of morphisms to compositions.
@@
 
-In the next few examples we will talk about a vary simple type of algebraic structures with a binary operation that we now define.
+In the next few examples we will talk about a very simple type of algebraic structure with a binary operation that we now define.
 
@@
 
-This is the same as the definition of \emph{groups} except that we don't require inverses.
+This is the same as the definition of \emph{groups} except that we do not require inverses.
 Similar to homomorphisms on groups, we define homomorphisms on monoids.
@@
 
-Note that this definition only speaks about what heapens to morphisms between the \emph{same pair} of objects.
-For morphisms $A \xrightarrow{f, g} B$, we can conclude that $F(A) = F(B)$ implies $A = B$ but for a morphism $A \xrightarrow{f} B$ and a morphism $C \xrightarrow{g} D$ between another pair of objects, the faithfulness of $F$ doesn't tell us anything.
+Note that this definition only speaks about what happens to morphisms between the \emph{same pair} of objects.
+For morphisms $A \xrightarrow{f, g} B$, we can conclude that $F(A) = F(B)$ implies $A = B$ but for a morphism $A \xrightarrow{f} B$ and a morphism $C \xrightarrow{g} D$ between another pair of objects, the faithfulness of $F$ does not tell us anything.
 Let us consider a functor like in the following graphic as an example.

### 2017-07-28 notes/notes-2/notes-2.tex
Make some stylistic changes to notes for talk 2

@@
 \begin{document}
-\title{Maps between Categories -- Functors}
+\title{Functors \\ Maps between Categories}
 \author{Felix Rech \and Dominik Wagner}
@@
       written as $A\mapsto F(A)$ and
-    \item for each $A,A'\in\ob(\cat A)$, a function
+    \item for all objects $A$ and $A'$, a function
       \begin{align*}
@@
 Sometimes we call the first function \demph{action on paths} and the second function \demph{action on morphisms}.
+The definition of functors turns the class of categories into a category itself.
 
-We can easily define composition of functors and an identity functor on every category.
-With those we define a category $\Cat$ that contains categories as objects and functors as morphisms.
+\begin{definition}
+  We define a category $\Cat$ with categories as elements and functors as morphisms.
+  The identity morphism on an object is the functor that maps every object and every morphism to itself.
+  The composition is defined by composition of both actions on objects and composition of both actions on morphisms.
+\end{definition}
 
@@
 You might already spot some similarity to the functoriality axioms.
-We use this in our next example.
+We capture this in our next example.
 
@@
   Monoids together with monoid homomorphisms form a category.
-  We can define a functor from the category of monoids to $\Set$, that maps every monoid to its element set and every homomorphism to the undelying function.
+  We can define a functor from the category of monoids to $\Set$ that maps every monoid to its element set and every homomorphism to the undelying function.
   This functor maps into a category that is similar to its domain but with less structure on the objects.
   Hence one can say that the functor \emph{forgets} the monoid structure.
-  This can be observed on many functors.
+  This can be observed on many other functors as well.
   We call such functors \demph{forgetful}.
@@
   There is a functor from the category of commutative monoids to the category of general monoids which maps every object and every morphism to itself.
-  It is injective on objects and on arrows, which means that we can always say where a resulting object \textit{came from}.
+  It is injective on objects and on arrows, which means that we can always say, where a resulting object \textit{came from}.
   But we still call it \emph{forgetful} because it maps into a category where in general the objects have less structure.
@@
   As binary operation we use the concatenation of lists and as neutral element the empty list.
-  If we have a function $A \xrightarrow{f} B$ then we can lift it to a monoid homomorphism $A^* \xrightarrow{f^*} B^*$ that applies $f$ pointwise on a list, i.e.\ $f^*([x, y, \ldots]) = [f(x), f(y), \ldots]$.
+  If we have a function $A \xrightarrow{f} B$ between sets then we can lift it to a monoid homomorphism $A^* \xrightarrow{f^*} B^*$ that applies $f$ pointwise on a list, i.e.\ $f^*([x, y, \ldots]) = [f(x), f(y), \ldots]$.
   This gives us a functor that maps a set to its free monoid and hence \emph{adds structure}.
@@
 Note that this definition only speaks about what happens to morphisms between the \emph{same pair} of objects.
-For morphisms $A \xrightarrow{f, g} B$, we can conclude that $F(A) = F(B)$ implies $A = B$ but for a morphism $A \xrightarrow{f} B$ and a morphism $C \xrightarrow{g} D$ between another pair of objects, the faithfulness of $F$ does not tell us anything.
-Let us consider a functor like in the following graphic as an example.
+We can still map two different morphisms onto the same target, as long as they are defined between a different pair of objects.
+Let us illustrate this with an example.
 \[ \begin{tikzpicture}
@@
 With our definition, a subcategory is full if and only if the inclusion functor is full.
-An important property is that a full subcategory is completely determined by the selection of objects that it contains.
+A convenient property is that a full subcategory is completely determined by the selection of objects that it contains.
 
-As the image of a monoid homomorphism is a monoid and the image of a group homomorphism is a group, one could easily assume that the image of a functor is a category, but in general this is wrong as we can see in this example:
+As the image of a monoid homomorphism is a monoid and the image of a group homomorphism is a group, one could easily assume that the image of a functor is a category, but in general this is wrong as we can see in the next example:
 \[ \begin{tikzpicture}
@@
 Assume that a functor $F$ is given by the dotted arrows and maps the morphism $f$ to $u$ and $g$ to $v$.
-Since the image of $F$ contains $X \xrightarrow{u} Y$ and $Y \xrightarrow{i} Z$, it should also contain $v \of u$ to be a category.
+Since the image of $F$ contains $u$ and $v$, it should also contain $v \of u$ to be a category.
 But $F$ does not map any morphism to $v \of u$, thus the image is not a category.

### 2017-08-02 sheets/talk-12/sheet-12.tex

@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}
+
+\DeclareMathOperator{\myIn}{in}
 
@@
       We apply naturality of $\sigma^{-1}$ in step two and one hexagon equation in step three.
-      The rest follows from the correctness of the graphical calculus for general monoidal categories and cancellation of inverses.
+      The rest follows from the naturality of $\lambda$ and $\rho$ with the correctness of the graphical calculus for general monoidal categories and cancellation of inverses.
     \end{enumerate}
@@
       The section above it commutes by naturality of $\sigma$.
-      The squares to it's right and left commute by naturality of $\lambda$ and $\rho$.
+      The squares to its right and left commute by naturality of $\lambda$ and $\rho$.
       Everything else commutes trivially or by the coherence theorem for general monoidal categories.
@@
 \begin{answer}
-  You can define the associator, unitors and symmetry by composition of the corresponding morphisms for products and coproducts as well as distributivity.
-  As compositions of natural isomorphisms they will also be natural isomorphisms.
-  You can then check the triangle, pentagon and hexagon equalities and show that the symmetry is self-inverse (Application of a proof assistant adviced).
-
-  Another approach is to look at a category that includes $\Set$ as subcategory but makes the monoid structure more obvious.
+  First we show that this describes a monoidal category.
+  The tensor product an unit are already given.
+  The next step is to define the associator and unitors.
+  We ignore some parentheses in our notation and denote by $\myIn_i : A_i \to A_1 + \cdots + A_i + \cdots + A_n$ the injection into the $n$-ary disjoint union for all families of sets $A_1, \ldots, A_n$ and indices $1 \leq i \leq n$.
+  We define the associator and unitors by pattern matching:
+  \begin{align*}
+    \alpha_{A, B, C} &: A + B + (A \times B) + C + \roundddd{\rounddd{A + B + (A \times B)} \times C} \\
+      &\qquad \to A + B + C + (B \times C) + \roundddd{A \times \rounddd{B + C + (B \times C)}} \\
+    \alpha(\myIn_1(a)) &= \myIn_1(a) \\
+    \alpha(\myIn_2(b)) &= \myIn_2(b) \\
+    \alpha(\myIn_3(a, b)) &= \myIn_5(a, \myIn_1(b)) \\
+    \alpha(\myIn_4(c)) &= \myIn_3(c) \\
+    \alpha(\myIn_5(\myIn_1(a), c)) &= \myIn_5(a, \myIn_2(c)) \\
+    \alpha(\myIn_5(\myIn_2(b), c)) &= \myIn_4(b, c) \\
+    \alpha(\myIn_5(\myIn_3(a, b), c)) &= \myIn_5(a, \myIn_3(b, c)) \\
+    \\
+    \lambda_{A} &: \emptyset + A + (\emptyset \times A) \to A \\
+    \lambda(\myIn_2(a)) &= a \\
+    \\
+    \rho_{A} &: A + \emptyset + (A \times \emptyset) \to A \\
+    \rho(\myIn_1(a)) &= a.
+  \end{align*}
+  Those are obviously bijections since you can basically swap the terms on both sides of the defining equations and get the definition of the inverse.
+  They describe natural transformations because they do not change the elements of $A$, $B$ and $C$ but only repackage them.
+  To check the triangle and pentagon equalities, is trivial but tedious because you have to make many case distinctions.
+  It is best done with a proof assistant.
+  We omit it here.
+
+  We showed that $\emptyset$ and $\otimes$ define a monoidal structure.
+  Now we will show that it is even symmetric.
+  We define the symmetry as
+  \begin{align*}
+    \sigma_{A, B} &: A + B + (A \times B) \\
+      &\qquad\to B + A + (B \times A) \\
+    \sigma(\myIn_1(a)) &= \myIn_2(a) \\
+    \sigma(\myIn_2(b)) &= \myIn_1(b) \\
+    \sigma(\myIn_3(a, b)) &= \myIn3(b, a).
+  \end{align*}
+  The symmetry is self-inverse because as above, we can swap the argument pattern with the right side of each defining equation and get the same set of equations back.
+  Also as above, the symmetry is natural because it does not modify elements of $A$ or $B$.
+  The hexagon equalities are again trivial to show, but the proof needs many case distinctions and is thus omitted here.
+
+  Another approach is to look at a category that includes $\Set$ as subcategory but makes the symmetric monoidal structure more obvious.
   We observe that $I$ and $\otimes$ coincide with the terminal object and the product functor from the category of sets with partial functions.
-  Liker every terminal object and product, they give rise to a monoidal structure.
+  Liker every terminal object and product, they give rise to a symmetric monoidal structure.
   Since the associator, unitors and symmetry are isomorphisms, they are total and contained in $\Set$.
-  Thus the monoidal structure exists also $\Set$.
+  Thus the symmetric monoidal structure exists also in $\Set$.
 \end{answer}
@@
     \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
-    \node at (f) {$!$};
+    \node at (f) {$0_{R \otimes 0}$};
 
@@
     \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
-    \node at (f) {$!$};
+    \node at (f) {$0_{L \otimes R}$};
 
@@
   \end{tikzpicture}. \]
-  In one direction they are trivially inverse because morphisms to $0$ are unique.
+  Where $0_{R \otimes 0}$ and $0_{L \otimes 0}$ stand for the unique morphisms from the initial object.
+  In one direction they are trivially inverse because morphisms from $0$ are unique.
   In the other direction it follows by
@@
       \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
-      \node at (f) {$!$};
+      \node at (f) {$0_{R \otimes 0}$};
 
@@
       \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
-      \node at (g) {$!$};
+      \node at (g) {$0_{L \otimes 0}$};
 
@@
   \]
-  In the second step we use again that morphisms to $0$ are unique.
+  In the second step we use again that morphisms from $0$ are unique.
 \end{answer}
@@
 \begin{exercise}
-  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying one of the snake equations, then $L$ has a right dual (not necessarily given by $R$ with $\eta$ and $\varepsilon$).
+  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying the first snake equation, then $L$ has a right dual (not necessarily given by $R$ with $\eta$ and $\varepsilon$).
 \end{exercise}
@@
     \[ \vcenter{ \hbox{ \begin{tikzpicture}
-      \draw[object] (0, 2)
-        to (0, 1)
+      \draw[object] (2, 2)
+        to (2, 1)
         to[cup object] (1, 1)
-        to[cap object] (2, 1)
-        to (2, 0);
-    \end{tikzpicture} } }
-    =
-    \vcenter{ \hbox{ \begin{tikzpicture}
-      \draw (0, 2) to (0, 0);
+        to[cap object] (0, 1)
+        to (0, 0);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \draw (0, 2) (0, 0);
     \end{tikzpicture} } }. \]
@@
     \[ \vcenter{ \hbox{ \begin{tikzpicture}
-      \draw[object] (0, -1)
-        to (0, 0)
-        to[cap object] (1, 0)
+      \draw[object] (4, -1)
+        to node[auto, near start] {$R$} (4, 0)
+        to[cap object] (3, 0)
         to[cup object] (2, 0)
         to (2, 2)
+        to[cap object] (1, 2)
+        to[cup object] (0, 2)
+        to node[auto, near end] {$R$} (0, 3);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \draw[object] (4, -1)
+        to node[auto, very near start] {$R$} (4, 2)
         to[cap object] (3, 2)
-        to[cup object] (4, 2)
-        to (4, 3);
-    \end{tikzpicture} } }
-    =
-    \vcenter{ \hbox{ \begin{tikzpicture}
-      \draw[object] (0, -1)
-        to (0, 2)
-        to[cap object] (1, 2)
-        to (1, 1)
+        to (3, 1)
         { [local bounding box=inner snake]
@@
           to[cup object] (2, 1)
-          to[cap object] (3, 1)
+          to[cap object] (1, 1)
           node[phantom morphism] {}
         }
-        to (3, 0)
-        to[cup object] (4, 0)
-        to (4, 3);
+        to (1, 0)
+        to[cup object] (0, 0)
+        to node[auto, very near end] {$R$} (0, 3);
       \node[highlight box, fit=(inner snake)] {};
@@
     \vcenter{ \hbox{ \begin{tikzpicture}
-      \draw[object] (0, -1)
-        to (0, 2)
+      \draw[object] (2, -1)
+        to node[auto, very near start] {$R$} (2, 2)
         to[cap object] (1, 2)
@@
         to (1, 0)
-        to[cup object] (2, 0)
-        to (2, 3);
+        to[cup object] (0, 0)
+        to node[auto, very near end] {$R$} (0, 3);
       % \node[highlight box, fit=(inner snake 2) (inner snake 2 |- inner snake.north) (inner snake 2 |- inner snake.south)] {};
@@
   \]
-  Hence there are morphisms $g$ and $h$ such that
+  Since all idempotents split, there are morphisms $R \xrightarrow{g} X$ and $ X\xrightarrow{h} R$ such that
   \[
@@
       \node[phantom morphism, overlay] (in) at (0, 0) {};
-      \draw[object] (in) to (h) to (g) to (out);
+      \draw[object] (in)
+        to node[auto] {$X$} (h)
+        to (g)
+        to node[auto] {$X$} (out);
     \end{tikzpicture} } }
@@
       \node[phantom morphism, overlay] (in) at (0, 0) {};
-      \draw[object] (in) to (out);
+      \draw[object] (in) to node[auto, very near start] {$X$} (out);
     \end{tikzpicture} } }
@@
       \node[phantom morphism, overlay] (in) at (0, 0) {};
-      \draw[object] (in) to (g) to (h) to (out);
-    \end{tikzpicture} } }
-    =
-    \vcenter{ \hbox{ \begin{tikzpicture}
-      \node[phantom morphism, overlay] (out) at (2, 3) {};
-      \node[phantom morphism, overlay] (in) at (0, 0) {};
       \draw[object] (in)
-        to (0, 1.5)
+        to node[auto] {$R$} (g)
+        to (h)
+        to node[auto] {R$$} (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (out) at (0, 3) {};
+      \node[phantom morphism, overlay] (in) at (2, 0) {};
+      \draw[object] (in)
+        to node[auto, near start] {$R$} (2, 1.5)
         to[cap object] (1, 1.5)
-        to[cup object] (2, 1.5)
-        to (out);
+        to[cup object] (0, 1.5)
+        to node[auto, near end] {$R$} (out);
     \end{tikzpicture} } }.
   \]
-  We use those morphisms to define a unit
+  The object $X$ is right dual to $L$.
+  We use the morphisms $g$ and $h$ to define a unit
   \[
@@
       \node[phantom morphism, overlay] (out_right) at (1, 1) {};
-      \node[morphism] (g) at (1, 0) {$g$};
+      \node[morphism] (g) at (0, 0) {$g$};
       \draw[object] (out_left)
-        to node[auto, swap] {$R$} (0, 1 |- g.north)
-        to (0, 1 |- g.south)
-        to[cup object] (g.south)
-        (g.north) to node[auto] {$L$} (out_right);
+        to node[auto, swap] {$X$} (g)
+        to[cup object] (1, 0 |- g.south)
+        to (1, 0 |- g.north)
+        to node[auto] {$L$} (out_right);
     \end{tikzpicture} } }
@@
     \vcenter{ \hbox{ \begin{tikzpicture}
-      \node[morphism] (h) at (0, 1) {$h$};
+      \node[morphism] (h) at (1, 1) {$h$};
       \node[phantom morphism, overlay] (in_left) at (0, 0) {};
@@
       \draw[object] (in_left)
-        to node[auto] {$L$} (h)
-        to[cap object] (1, 0 |- h.north)
-        to (1, 0 |- h.south)
-        to node[auto, swap] {$R$} (in_right);
+        to node[auto] {$L$} (0, 0 |- h.south)
+        to (0, 0 |- h.north)
+        to[cap object] (h.north)
+        (h.south) to node[auto, swap] {$X$} (in_right);
+      % \draw[object] (in_left)
+      %   to node[auto] {$L$} (h)
+      %   to[cap object] (1, 0 |- h.north)
+      %   to (1, 0 |- h.south)
+      %   to node[auto, swap] {$R$} (in_right);
     \end{tikzpicture} } },
@@
     \vcenter{ \hbox{ \begin{tikzpicture}
-      \node[phantom morphism] (out) at (-1, 2.5) {};
+      \node[phantom morphism] (out) at (1, 2.5) {};
       \node[morphism] (h) at (0, 1) {$h$};
       \node[morphism] (g) at (0, 0) {$g$};
-      \node[phantom morphism] (in) at (1, -1.5) {};
+      \node[phantom morphism] (in) at (-1, -1.5) {};
       \draw[object] (in)
@@
     \vcenter{ \hbox{ \begin{tikzpicture}
-      \node[phantom morphism] (out) at (-1, 2.5) {};
-      \node[phantom morphism] (h) at (2, 1) {};
-      \node[phantom morphism] (g) at (0, 0) {};
-      \node[phantom morphism] (in) at (3, -1.5) {};
+      \node[phantom morphism] (out) at (3, 2.5) {};
+      \node[phantom morphism] (h) at (0, 1) {};
+      \node[phantom morphism] (g) at (2, 0) {};
+      \node[phantom morphism] (in) at (-1, -1.5) {};
       \draw[object] (in)
@@
         to[cap object] (h.north)
-        to (2, 0.5)
+        to (0, 0.5)
         to[cup object] (1, 0.5)
-        to[cap object] (0, 0.5)
+        to[cap object] (2, 0.5)
         to (g.south)
@@
     \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism] (out) at (2, 4) {};
+      \node[phantom morphism] (in) at (0, 0) {};
+      \draw[object] (in)
+        to (0, 2)
+        to[cap object] (1, 2)
+        to[cup object] (2, 2)
+        to (out);
+    \end{tikzpicture} } }
+    =
+    \vcenter{ \hbox{ \begin{tikzpicture}
       \node[phantom morphism] (out) at (0, 4) {};
-      \node[phantom morphism] (in) at (2, 0) {};
+      \node[phantom morphism] (in) at (0, 0) {};
+      \draw[object] (in) to (out);
+    \end{tikzpicture} } }
+  \]
+  and
+  \[
+    \vcenter{ \hbox{ \begin{tikzpicture}
+      \node[phantom morphism, overlay] (in) at (2, -0.5) {};
+      \node[morphism] (h) at (2, 1) {$h$};
+      \node[morphism] (g) at (0, 3) {$g$};
+      \node[phantom morphism, overlay] (out) at (0, 4.5) {};
       \draw[object] (in)
+        to (h)
         to (2, 2)
@@
         to[cup object] (0, 2)
-        to (out);
-    \end{tikzpicture} } }
-    =
-    \vcenter{ \hbox{ \begin{tikzpicture}
-      \node[phantom morphism] (out) at (0, 4) {};
-      \node[phantom morphism] (in) at (0, 0) {};
-      \draw[object] (in) to (out);
-    \end{tikzpicture} } }
-  \]
-  and
-  \[
-    \vcenter{ \hbox{ \begin{tikzpicture}
-      \node[phantom morphism, overlay] (in) at (0, -0.5) {};
-      \node[morphism] (h) at (0, 1) {$h$};
-      \node[morphism] (g) at (2, 3) {$g$};
-      \node[phantom morphism, overlay] (out) at (2, 4.5) {};
-      \draw[object] (in)
-        to (h)
-        to (0, 2)
-        to[cap object] (1, 2)
-        to[cup object] (2, 2)
         to (g)

### 2017-08-03 sheets/talk-6/sheet-6.tex
Merge branch 'uds-psl/master'

@@
 \begin{answer}
-  The coequalizer follows the scheme like in the talk.
-  We define the relation $R := \set{ (1_X (x), f(x) \;\big|\; x \in X} = \set{ (x, f(x) \;\big|\; x \in X} = f$ defined as a set.
+  The definition of the coequalizer follows the scheme like in the talk.
+  We define the relation $R := \set{ (1_X (x), f(x) \;\big|\; x \in X} = \set{ (x, f(x) \;\big|\; x \in X} = f$ seen as a functional relation.
   Now $\sim$ is the reflexive symmetric transitive closure of $R$.
@@
   is the coequalizer of $\parpairi{X}{X}{f}{1}$.
-  The equivalence $\sim$ has the property:
-  $\forall x_1, x_2 \in X, (x_1, x_2) \in \sim \;\Leftrightarrow\; f(x_1) = f(x_2).$
-  Thus we have $[x] = \set{x' \in X \;\big|\; f(x) = f(x')}$.
 
+  We discover that $(x, y) \in \sim$ means that we can ``draw a line of forward and backward steps'' between $x$ and $y$.
+  A ``forward step'' means that if we have $(x, y) \in \sim$ and $f(y)=z$ then $(x, z) \in \sim$.
+  A ``backward step'' means that if we have $(x, y) \in \sim$ and $f(z)=y$ then $(x, z) \in \sim$.
+  We also have $(x, x) \in \sim$ for every $x \in X$.
+  More formally:
+  With $\hat{R} := R \cup R^{-1}$ we have
+  $\sim = \bigcup_{n\in\nat} {(\hat{R})^n}$
+  where $(\hat{R})^n = \underbrace{\hat{R} \of \cdots \of \hat{R}}_{\text{$n$ times}}$ is the $n$th power relation over $\hat{R}$.
 \end{answer}

### 2017-08-03 sheets/talk-7/sheet-7.tex
Merge branch 'uds-psl/master'

@@
   \item
-    As the shape $\bf{I}$ we choose the following diagram.
+    As the shape $\bf{I}$ we choose the following category.
     \[
@@
   What kind of object is $L$?\\
-  What about an object $L' \in$ \texttt{Cone(D)} that corresponds to a colimit in $\mathscr{C}$ over $D$?
 \end{exercise}
@@
   Therefore $L$ is a terminal object.\\
-  By duality $L'$ is an initial object.
+  There is also a dual category $\texttt{Cone(D)}^{Op}$, the category of cocones. By duality,
+  an object $L'$ corresponding to a colimit in $\mathscr{C}$ is an initial object in $\texttt{Cone(D)}^{Op}$.
 \end{answer}
@@
   Let $\mathscr{C}$ and $\bf{I}$ be any categories, $D : \bf{I} \to \mathscr{C}$ a diagram.
-  Verify that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ is a cone over $D$ with vertex $V$.
+  Verify that a natural transformation $\pi : \Updelta_V \to D$ is a cone over $D$ with vertex $V$.
 \end{exercise}
 \begin{answer}
-  Recall that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ has components for every object $I \in \bf{I}$
+  Recall that a natural transformation $\pi : \Updelta_V \to D$ has components for every object $I \in \bf{I}$
   and for every morphism $m : I \to J \in \bf{I}$ the following square commutes.
@@
   This is exactly the definition of a cone with vertex $V$.
-  In fact this yields an alternative equivalent definition of cones.
+  In fact this yields an alternative equivalent characterization of cones.
 \end{answer}
@@
   \begin{itemize}
-  \item You may use the fact that $E$ is a cone iff $E$ is an equalizer of $s$ and $t$.
+  \item You may use the fact that $E$ is a cone iff $E$ is a fork of $s$ and $t$.
   \item You do not need the concrete definitions of $s$ and $t$.
@@
   morphism $h : F \to E$.
-  If $F$ is a cone, then it is also an equalizer of $s$ and $t$.
+  If $F$ is a cone, then it is also a fork of $s$ and $t$.
   \[

### 2017-08-03 papers/base-1/paper-1.tex
Writeup for talk 1

\def\pathToRoot{../}\input{\pathToRoot headers/uebungHeader}

\usepackage{charter}
\fontfamily{bch}\selectfont

\begin{document}
\title{What is a Category?}
\author{Sarah Mameche, Leonhard Staut, Andreas Meyer}
\maketitle
\section {Introduction}
A category consists of mathematical objects which are related to one another via mappings. The objects in a category can be arbitrary ones, but in most cases, they share some structure which is preserved by the morphisms between pairs of objects.\\ Various mathematical objects and mappings can be formalized as a category. The most intuitive example is the category of sets where mappings are simply functions. It is often helpful to think of morphisms between objects in a category as an abstraction of a function. Category theory makes these abstractions to find general statements and proofs that hold for all mathematical objects that can be seen as a category.
\\ In the following, we introduce the basic definition of a category, along with some important properties of its objects and mappings, and give common examples of categories.
\section {Definition of Categories}
\begin{definition}{Category}
	A category $\mathscr{C}$ has the following components:
	\begin{itemize}
		\item a collection of objects $ob(\mathscr{C})$
		\item for every pair of objects $A, B \in ob(\mathscr{C})$ a collection of maps (arrows, morphisms), denoted by $\mathscr{C}(A, B)$
		\item for objects $A, B, C \in ob(\mathscr{C})$ a composition function
		\[\mathscr{C}(B, C) \times \mathscr{C}(A, B) \rightarrow  \mathscr{C}(A, C)
		\]		\[ (g,f) \mapsto (g \of f)
		\]
		\item for every $A \in ob(\mathscr{C})$, an element $1_A \in \mathscr{C}(A, A)$, called the identity on $A$ 
 	\end{itemize}
 The following axioms have to be satisfied:
 \begin{itemize}
 	\item \textbf{identity laws:} $\forall f \in \mathscr{C}(A, B). \; f \of 1_A = f = 1_B \of f$ 
 	\item \textbf{associativity:} $\forall f \in \mathscr{C}(A, B), \; g \in \mathscr{C}(B, C), \; h \in \mathscr{C}(C, D). \; (h \of g) \of f = h \of (g \of f) $.
 \end{itemize}
\end{definition}
    \textit{Remark 2.1.} A category in which all objects are isolated and the identity maps are the only morphisms is called a \textbf{discrete} category.

%\section {Diagrams}
\textit{Remark 2.2.}
It is possible to draw diagrams for a category or parts thereof. This is particularly useful to make statements about how arrow composition should behave. For example, consider the following diagram:
\[
  \begin{tikzcd}
    A \arrow{r}{f} \arrow{d}{g} & B \arrow{d}{h} \\
    C \arrow{r}{i}              & D 
  \end{tikzcd}
\]
This represents a category with objects $\{A, B, C, D\}$ and
morphisms $\{f, g, h, i, 1_A, 1_B, 1_C, 1_D \}$.
Since every object requires to have an identity arrow by definition,
we usually do not include them in diagrams. The morphisms obtained by arrow composition are also left implicit; in the example, you could add diagonal arrows from $A$ to $D$ for the morphisms $ h \circ f $ and $ i \circ g $. \\
We can require these arrows to be the same,  $ h \circ f = i \circ g $, by saying that the diagram \emph{commutes}. In general, a diagram commutes if any two paths between two objects
obtained by composing arrows are the same.
%The example diagram above commutes if $ h \circ f = i \circ g $, meaning there is
%exactly one way to reach $D$ from $A$.

\section {Isomorphisms}
The notion that a function or mapping is invertible can also be defined for maps in categories.
\begin{definition}{Isomorphism}
  Let $\mathscr{C}$ be a category and $A, B \in ob(\mathscr{C})$. \\
  A map $f \from A \to B$ in $\mathscr{C}$ is an isomorphism if there is a $g \from B \to A$ s.t. $g \circ f = 1_A$ and $f \circ g = 1_B$. \\
  $A$ and $B$ are isomorphic in $\mathscr{C}$, written $A \cong B$, if there is an isomorphism between them.
\end{definition}

In short, a morphism is an isomorphism if it has an inverse.
For example, in the category \textbf{Set},
isomorphisms are exactly the invertible functions.\\\\
It can be shown that if inverses exist, they are unique, and thus we speak of
\emph{the} inverse of $f$.\\\\
\textit{Proposition 4.1.} If a map $f$ is an isomorphism, then its inverse is unique.

\begin{proof}
  Let $f \from A \to B$ be a map.
  Assume there exist two inverses $g \from B \to A$, such that $gf = 1_A$, $fg = 1_B$ and \mbox{$h \from B \to A$}, such that $hf = 1_A$, $fh = 1_B$.
  By the identity axiom, $1_A \of g = g$ and $h \of 1_B = h$. Therefore we get:
 \[  1_A \of g = g \textiff (h \of f) \of g = g \textiff h \of (f \of g) = g \textiff h \of 1_B = g \textiff h=g. \]
\end{proof}

\section {Examples of Categories}
\begin {enumerate}

 \item The category of sets, denoted as \textbf{Set}, is the category whose objects $ob$(\textbf{Set}) are sets.
   \\ The arrows in \textbf{Set} are the functions between two sets $A,B\ \in ob(\textbf{Set}).$
   \\ The identity- function is defined as $\forall A: set, \ id_A:\ A \to A,\ \forall x \in A. \ f (x) = x$.
   \\ The composition $\circ$ is the composition of functions and this is associative. \\\textit{Proof}:
   \\ Assume sets $A, B, C,$ and $D$ and functions $f \from A \to B$, $g \from B \to C$, $h \from C \to D$ and $x \in A$.
   \\ $((h\ \circ \ g)\ \circ\ f)\ x = ((h \ \circ g \ )(f \ x) =h(g(f\  (x))) $
   \\ $(h\ \circ \ (g \ \circ f))\ x = h ((g\ \circ \ f) \ x) = h(g(f \ (x)))$
   \\ Thus $(h\ \circ \ g)\ \circ \ f = \ h \ \circ \ (g \ \circ \ f)$ holds. $\Box$
 \item The category of relations, denoted as \textbf{Rel}, is the category whose objects are sets.
   \\The arrows are all binary relations between two $A,B \ \in ob(\textbf{Rel}).$
   \\The identity arrow  is the identity function $\forall A: set, \ id_A:\ A \to A,\ \forall x \in A. \ f (x) = x$
   \\ The composition $R \circ S$, $R  \in \textbf{Rel}(A,B), S \in \textbf{Rel}(B,C), A,B,C \in ob(\textbf{Rel})$ is defined as $ (x,y) \in  R \circ S \leftrightarrow \exists  z.(x,z) \in S \land (z,y) \in R$.

  \item Categories similar to \textbf{Set} can be constructed for sets which have some additional structure and structure-preserving mappings between them. For example, there is a category \textbf{Poset} with partially ordered sets as objects. It is defined similarly to \textbf{Set} but with monotone functions.

  \item A single poset $(P, \leq)$ gives rise to a category $\mathscr{P}$ if we take elements of $P$ to be the objects. There is a unique arrow in $\mathscr{P}(A, B)$ iff $A \leq B$. The reflexivity requirement of $\leq$ ensures that identity arrows exist for all objects. Also, the category has composition because the order is transitive.

  \item \textbf{Vect$_k$} is the category of vector spaces over a field $k$, with linear transformations as arrows.

  \item There is a category \textbf{Grp} whose objects are groups; maps are group homomorphisms between two groups $A,B \in ob(\textbf{Grp}).$
    Similarly there are categories \textbf{Mon} with monoids as objects and \textbf{Ring} with rings as objects.
    The maps are given by homomorphisms between monoids and rings respectively.

  \item
    The objects in a category do not have to be like sets, and the maps do not have to be like functions.
    A single group $(G, \cdot)$ can also be seen as a category $\mathcal{G}$. The category has only one object $G$ which represents the group itself,
    and arrows in $\mathcal{G}(G, G)$  correspond to group elements.
    The identity arrow is the unit element $1$ of the group.\\
    Composition of maps in the category $\circ$ corresponds to applying the group action $\cdot$.\\
    A group requires all elements to have an inverse, thus we need each arrow $\mathcal{G}(G, G)$ to be an isomorphism.\\
    Dropping the last requirement, the category would represent a single monoid.

  \item Categories can be arbitrarily constructed by giving objects, arrows and arrow composition in a way that satisfies the axioms in the definition. For instance, the category \textbf{1} includes a single object and its identity map, without further specification what the object is. \\

 \end {enumerate}

\section {Epics and Monics}
We have seen that sets and fuctions form a category, and that invertibility of a function can be generalized to isomorphism in arbitrary categories. In this section, we generalize injectivity and surjectivity of functions by defining two special kinds of morphisms: monomorphisms and epimorphisms.

\begin{definition}{Monomorphism}
  Let $\mathscr{C}$ be a category. A map $X \overset{f}{\to} Y$ is called a monomorphism if for all objects $Z$ and maps
  $
  \begin{tikzcd}
    Z \arrow[r,shift left, "g"] \arrow[r,shift right,swap,"g'"] & X
  \end{tikzcd}
  $
  \[
    f \circ g = f \circ g' \Rightarrow g = g'
  \]
  If $f$ is a monomorphism, we say that $f$ is monic.
\end{definition}

\begin{definition}{Epimorphism}
  Let $\mathscr{C}$ be a category. A map $X \overset{f}{\to} Y$ is called an epimorphism if for all objects $Z$ and maps
  $
  \begin{tikzcd}
    Y \arrow[r,shift left, "g"] \arrow[r,shift right,swap,"g'"] & Z
  \end{tikzcd}
  $
  \[
    g' \circ f = g' \circ f \Rightarrow g = g'
  \]
  If $f$ is an epimorphism, we say that $f$ is epic.
\end{definition}

%% TODO: We could add a reference to the definition of categories, so that we can link it in the text below

In other words, a monomorphism is cancellable on the left, and an epimorphism is cancellable on the right of compositions.
By the identity law, the identity map of an arbitrary object is always both monic and epic.\\
\\ \textit{Proposition 6.1.} In the category \textbf{Set}, monic maps are exactly the injective functions, and epic maps are exactly the surjective functions.
\begin{proof}
  First we show $f$ monic $\Leftrightarrow$ $f$ injective.
  \begin{itemize}
  \item $"\Rightarrow"$: Let $f \from A \to B$ be monic, $a, a' \in A$ with $a \neq a'$, $g, g' \from {x} \to A$ with $g(x) = a, g'(x) = a'$.
    We show $a \neq a' \Rightarrow f(a) \neq f(a')$.
    We have $g \neq g'$.
    Since $f$ is monic we get $f \circ g \neq f \circ g'$.\\
    We have: $f(a) = f(g(x)) \neq f(g'(x)) = f(a')$. Therefore $f$ is injective.
  \item $"\Leftarrow"$: Let $f \from A \to B$ be injective, $g, g' \from X \to A$ with $g \neq g'$.
    We show $f \circ g \neq f \circ g'$.
    There exists an $x \in X$ with $g(x) \neq g'(x)$.\\
    We have: $f(g(x)) \neq f(g'(x))$ since $f$ is injective.
    Therefore $f$ is monic.
  \end{itemize}
  Secondly we show $f$ epic $\Leftrightarrow$ $f$ surjective.
  \begin{itemize}
  \item $"\Rightarrow"$: Let $f \from A \to B$ be epic, and $X$ a two element set, e.g. $\{ \texttt{true}, \texttt{false}\}$.
    Let $g \from B \to X$ be the characteristic function of \texttt{Im}$(f)$ which is defined by
    $\forall b \in B. \ g(y) = \texttt{true}
    \Leftrightarrow y \in \texttt{Im}(f)$.
    And let $g' \from B \to X$ be the constant \texttt{true}-function.
    $\forall b \in B. \ g'(y) = \texttt{true}$.
    Note that $g = g'$ exactly if $f$ is surjective.\\
    We have: $g \of f = g' \of f \xRightarrow{f \ epic} g = g'$.
    Therefore $f$ is surjective.
  \item $"\Leftarrow"$: Let $f \from A \to B$ be surjective, and $g, g' \from B \to X$ with $g \neq g'$.
    There exists a $b \in B$ with $g(x) \neq g'(x)$.
    Since $f$ is surjective, there is an $a \in A$ with $f(a) = x$.\\
    We have: $g(f(a) \neq g'(f(a)) \Rightarrow g \circ f \neq g' \circ f$.
    Therefore $f$ is epic.
  \end{itemize}
\end{proof}
In ordinary set theory, injectivity and surjectivity are formulated in terms of elements of a set.
Epic and monic maps enable us to abstract from that and
talk about these properties without referring to sets or their elements at all.
We can then use this abstract concept of monic and epic arrows to find these structures
in different categories.
In the categories \textbf{Mon}, \textbf{Grp}, \textbf{Ring}, and \textbf{Vect}
monic maps are exactly the injective homomorphisms,
similar to the category \textbf{Set}.
However, we show the following: \\
\\\textit{Proposition 6.2.} There are epic maps in \textbf{Mon} which are not surjective.
\begin{proof}
  Consider the two monoids $(\mathbb{N}, +, 0)$ and $(\mathbb{Z}, +, 0)$, i.e. the
  additive monoids of the natural numbers and the integers respectively.
  The map $i \from \mathbb{N} \to \mathbb{Z}; \ i(n) = n$ is called inclusion map.
  Obviously this map is a monoid homomorphisms that is not surjective,
  because there are no negative integers in the image of $i$.
  However, $i$ is epic.\\
  Let $(\mathcal{M}, *, e)$ be a monoid and $f, g \from \mathbb{Z} \to \mathcal{M}$ be
  monoid homomorphisms.\\
  We need to show the implication $f \circ i = g \circ i \Rightarrow f = g$, that is,
  if $f$ and $g$ agree on $\mathbb{N}$, then they agree on the entire domain $\mathbb{Z}$.\\
  Note that:
  \begin{align*}
    f(-n) &= f ((-1)_1 + (-1)_2 + \dots + (-1)_n)\\
          &= f(-1)_1 * f(-1)_2 * \dots * f(-1)_n
  \end{align*}
  Therefore it suffices to show that $f(-1) = g(-1)$.
  \begin{align*}
    f(-1) &= f(-1) * e \\
          &= f(-1) * g(0) \\
          &= f(-1) * g(1 - 1) \\
          &= f(-1) * g(1) * g(-1) \\
          &\overset{(*)}{=} f(-1) * f(1) * g(-1) \\
          &= f(-1 + 1) * g(-1) \\
          &= f(0) * g(-1) \\
          &= e * g(-1) \\
          &= g(-1) \\ 
  \end{align*}
  In $(*)$ we use the assumption that $f=g$ if restricted to $\mathbb{N}$.\\
  Therefore $i$ is an epic map, even though it is not surjective.
\end{proof}

Are monomorphisms and epimorphisms are related to isomorphisms?
Recall that in the category \textbf{Set} isomorphisms correspond to the invertible functions.
It can be shown that invertible functions between two sets are
exactly the bijective functions, i.e. functions that are both injective and surjective.
Therefore, isomorphisms in the category \textbf{Set} are exactly those maps that are both monic and epic.\\
In the category \textbf{Grp} isomorphisms are exactly the bijective group homomorphisms
since those are the group homomorphisms that have an inverse homomorphism.
Similarly, in the category \textbf{Ring} isomorphisms are exactly the bijective ring homomorphisms.\\
We now show for a general category $\mathscr{C}$ that every isomorphism is both monic and epic.
\textit{Proposition 6.3.}Every isomorphism in a category is both monic and epic.
\begin{proof}
  Consider the commuting diagram:
  \[
    \begin{tikzcd}
      A \arrow[r,shift left, "y"] \arrow[r,shift right,swap,"x"] &
      B \arrow[dr, shift left, "1"]  \arrow{r}{m}  &
      C \arrow{d}{e} \arrow[r,shift left, "i"] \arrow[r,shift right,swap,"j"] & D\\
      &  & B \arrow[ul, shift left, "1"]&
    \end{tikzcd}
  \]
  Assume $m \from B \to C$ is an isomorphism with inverse $e \from C \to B$.
  Then:\\
  \begin{minipage}{.5\linewidth}
    \vspace{4mm}
    \centering $m$ is monic:
    \[
      \begin{aligned}
        m \of x &= m \of y\\
        \Rightarrow e \of (m \of x) &= e \of (m \of y)\\
        \Leftrightarrow (e \of m) \of x &= (e \of m) \of y\\
        \Leftrightarrow 1_B \of x &= 1_B \of y\\
        \Leftrightarrow x &= y
      \end{aligned}
    \]
  \end{minipage}%
  \begin{minipage}{.5\linewidth}
    \vspace{4mm}
    \centering $m$ is epic:
    \[
      \begin{aligned}
        i \of m &= j \of m \\
        \Rightarrow (i \of m) \of e &= (j \of m) \of e\\
        \Leftrightarrow i \of (m \of e) &= j \of (m \of e)\\
        \Leftrightarrow i \of 1_C &= j \of 1_C\\
        \Leftrightarrow i &= j
      \end{aligned}
    \]
  \end{minipage}%
\end{proof}
Note that the converse of the statement holds for the category \textbf{Set}, but not in general.
If a map is both monic and epic, it is not necessarily an isomorphism.
For a map $f$ to be an isomorphism there needs to exist an inverse map $f^{-1}$.
However monic and epic maps do not necessarily have an inverse.\\

\section {Dual categories}

For any category $\mathscr{C}$, $\mathscr{C}^{op}$ denotes the \textbf{dual category}
which is obtained by reversing the arrows in $\mathscr{C}$.
For every theorem $\Sigma$ in the language of category theory,
the dualized theorem $\Sigma^*$ exists,
and a proof for any theorem trivially yields the proof for the dual theorem by the \textbf{duality principle}.\\
Dual categories are especially useful for structures that are inherently dual.
As an example, epic and monic maps are dual to each other, since any monic map in $\mathscr{C}$
is an epic map in $\mathscr{C}^{op}$ and vice versa.
The following theorem is an example where duality can be used.\\\\
\textit{Proposition 6.1.} The composition of two monic maps is monic.
\begin{proof}
  Consider the commuting diagram:
  \[
    \begin{tikzcd}
      A \arrow[dr, swap, "h"] \arrow{r}{g} & B \arrow{d}{f} \\
      & C
    \end{tikzcd}
  \]
  We assume $f$ and $g$ to be monic and show that $h = f \circ g $ is monic.
  Let $X$ be any object and $x, x' \from X \to A$ two maps in the category.
  \begin{align*}
    h \circ x &= h \circ x' \\
   \Rightarrow (f \circ g) \circ x &= (f \circ g) \circ x' \\
    \Rightarrow f \circ (g \circ x) &= f \circ (g \circ x') \\
    \Rightarrow g \circ x &= g \circ x' \\
    \Rightarrow x &= x' \\
  \end{align*}
  Therefore, $h$ is monic.
\end{proof}
The dual of the statement above states that the composition of epic maps is also epic.
We get this fact immediately using the duality principle, since epic maps in $\mathscr{C}$
are monic maps in $\mathscr{C}^{op}$ whose composition is monic in $\mathscr{C}^{op}$, as we
have just shown, and therefore epic in $\mathscr{C}$.

\section {Terminal and Initial Objects}
In the previous sections, we introduced some properties of the morphisms in a category, namely that a map can be monic or epic, or an isomorphism. Now we will see properties of the objects in a category .
\begin {definition}{Initial and terminal Objects}
  Let $\mathscr{C}$ be a category.\\
  An object $0$ is called \textbf{initial}  iff for any object $A \in \mathscr{C}$,
  there is a unique morphism $0 \to A$.\\
  An object $1$ is called \textbf{terminal} iff for any object $A \in \mathscr{C}$,
  there is a unique morphism $A \to 1$.
\end {definition}

Note that inital and terminal objects are dual to each other.

An initial object in $\mathscr{C}$ is terminal in $\mathscr{C}^{op}$.\\
\\\textit{Proposition 8.1.}
Initial and terminal objects are unique up to isomorphism.

\begin{proof}
Assume $0$ and $0'$ to be inital objects in some category $\mathscr{C}$ and show that
$f \from 0 \to 0', g \from 0' \to 0$ form an isomorphism
$g \circ f$ between $0$ and $0'$. Consider the following diagram:

\[
\begin{tikzcd}
  & 0' \arrow{dr}{g} \arrow{rr}{1_{0'}}    &&  0'\\
  0 \arrow{ur}{f} \arrow{rr}{1_0} && 0 \arrow{ur}{f}
\end{tikzcd}
\]

Since $0$ is initial, we know that $f$ is unique, from the same argument follows the uniqueness of $g = f^{-1}$. \\$f \circ g$ and $1_{0'}$ are both maps in $\mathscr{C}(0', 0')$ and $0'$ is initial, so we have $f \circ g = 1_0'$. Analogously, $g \circ f = 1_0$ because $0$ is initial. \\
By duality, uniqueness also holds for terminal objects. 
\end{proof}
In the special case where the initial and terminal object of a category are identical, we speak of a \textbf{zero object} in a \textbf{pointed category}.\\
The 
\textit{Proposition 8.2.} In \textbf{Set}, $\emptyset$ is initial and the singleton set $\{x\}$ for an arbitrary set element $x$ is terminal.
\begin{proof}
There is only the empty function from $\emptyset$ to any other set, since there are no arguments in the domain.
\\For all sets $A$ there is a unique function $f \from A \to \{x\}$, since all elements of $A$ can only be mapped
  to $x$, that is $\forall y \in A.\ f(y) = x$. Note that it does not make a difference which element $x$ is, because singleton objects are isomorphic to one another; thus, we get a unique terminal object.
\end{proof}
\textit{Proposition 8.3.} In \textbf{Grp}, the trivial group is both initial and terminal.\begin{proof}
Let $1=\{e\}$ the trivial group with operation $*: e*e := e$.
\begin {itemize}
\item Let $\left({G, \circ}\right)$ be any group with identity element $e_G$,
  and let $\phi: 1 \to G$ be a function.
  Since group homomorphisms to preserve the identity,
  there is only one possible mapping for $\phi$ to be a group homomorphism:
  $\phi (e) = e_G$.\\
  Now we only need show that $\phi$ is a group homomorphism.
  \begin{align*}
    \phi (e) \circ \phi (e) &= e_G \circ e_G \\
                            &= e_G \\
                            &= \phi (e) \\
                            &= \phi (e * e)
  \end{align*}
  Therefore $\phi$ is the unique group homomorphism from the trivial group.
\item Recall that the singleton set is terminal in \textbf{Set}.
  Therefore there is exactly one mapping
  $\phi: G \to \{e\}$ defined by:
  $\forall g \in G: \phi(g) = e$.
  Again we need show that $\phi$ is a group homomorphism.
  For any $g, g' \in G$, we have:
  \begin{align*}
    \phi(g)*\phi(g') &= e * e \\
    &= e \\
    &= \phi (g \circ g')
  \end{align*}
  Therefore $\phi$ is the unique group homomorphism into the trivial group.
\end{itemize}
The trivial group is both initial and terminal. It is therefore an example for a
zero object, and \textbf{Grp} is an example for a pointed category.
The idea behind the name is that the shape of \textbf{Grp}, if you were to draw it,
looks like all other objects are pointed at the trivial group.
\[
  \begin{tikzcd}[column sep=large]
    \dots
    & \bigcirc \arrow [drrr, ""]
    & \bigcirc \arrow[drr, ""]
    & \bigcirc \arrow[dr, ""]
    & \dots
    & \bigcirc \arrow[dl, ""]
    & \bigcirc \arrow[dll,""]
    & \bigcirc \arrow[dlll,""]
    & \dots \\[5mm]
    \dots
    & \bigcirc \arrow[rrr, ""]
    & & &
    Triv
    \arrow[dlll, ""]
    \arrow[dll, ""]
    \arrow[dl, ""]
    \arrow[dr, ""]
    \arrow[drr, ""]
    \arrow[drrr, ""]
    \arrow[ulll, ""]
    \arrow[ull, ""]
    \arrow[ul, ""]
    \arrow[ur, ""]
    \arrow[urr, ""]
    \arrow[urrr, ""]
    \arrow[rrr, ""]
    \arrow[lll, ""]
    & &
    &  \bigcirc \arrow[lll, ""]
    & \dots   \\[5mm]
    \dots
    & \bigcirc \arrow [urrr, ""]
    & \bigcirc \arrow[urr, ""]
    & \bigcirc \arrow[ur, ""]
    & \dots
    & \bigcirc \arrow[ul, ""]
    & \bigcirc \arrow[ull,""]
    & \bigcirc \arrow[ulll,""]
    & \dots
  \end{tikzcd}
\]\end{proof}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-08-22 papers/talk-7/paper-7.tex
started on final paper for talk

\def\pathToRoot{../../}\input{\pathToRoot headers/header-7}

\title{General limits}
\author{Leonhard Staut}

\begin{document}

\maketitle
\chapter{Introduction}
We have seen examples of limits.\\
Insert examples here.\\
Now we generalize them...
\chapter{Definition}

\begin{definition}
  Let $\bm{I}, \mathscr{C}$ be categories.
  A \textbf{diagram} is a functor $D \from \bm{I} \to \mathscr{C}$.
  The category $\bm{I}$ is called the \textbf{shape} of the diagram.
\end{definition}

%% TODO: explain reasoning behind this definition and give example

\begin{definition}
  Let $D \from \bm{I} \to \mathscr{C}$ be a diagram.
  A \textbf{cone} $U$ over $D$ is an object $U \in \mathscr{C}$
  with maps $f_I, \forall\ I \in \bm{I}$, such that
  $\forall\ I \overset{m}{\to} J \in \bm{I}$, the following triangle commutes:
  \[
    \begin{tikzcd}
      & U \arrow[dl, swap, "f_I"] \arrow{dr}{f_J}&\\
      D_I \arrow[rr, swap, "D(m)"] & & D_J\\
    \end{tikzcd}
  \]
  That is $f_J = D(m) \circ f_I$.\\
  For a cone we write $(U \overset{f_I}{\to} D(I))_{I\in \bm{I}}$.
  We call $U$ the \textbf{vertex}, and $f_I$ the \textbf{projection maps} of the cone.
\end{definition}
\begin{remark}
  When $U$ is the vertex of a cone, then we sometimes use just $U$ to refer to the whole cone
  for better readability.
\end{remark}

%% TODO: write a sentence like "In other words, ..." and give example. example might be tricky to draw

\begin{definition}
  A cone $(U \overset{f_I}{\to} D(I))_{I\in \bm{I}}$ is universal, if
  for all other cones $(V \overset{g_I}{\to} D(I))_{I\in \bm{I}}$ there is a unique
  map $h \from V \to U$, such that for all $I \in \bm{I}$ the following triangles commute.
  \[
    \begin{tikzcd}
      V \arrow[dr, swap, "g_I"] \arrow{r}{h} & U \arrow{d}{f_I} \\
      & D_I
    \end{tikzcd}
  \]
  That is $g_i = f_i \circ h$.
\end{definition}

In other words, any cone over $D$ has maps into $D$, and a cone $U$ is universal if
for any other cone $V$, you can first go to $U$, that is the unique morphism.
From $U$ you can go to $D_I$, and the commuting condition says that this is equal
to the morphism that immediately takes you from $V$ to $D_I$.

%TODO: give example. might be tricky to draw.

\chapter{Building new limits}
It is possible to build new limits...

\chapter{Colimits}
Just dualize it...

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-08-23 sheets/talk-6/sheet-6.tex
Minor typos in sheet-6

@@
 
-\begin{exercise}[Product is unique up-to isomorphism]
+\begin{exercise}[Products are unique up-to isomorphism]
   Prove that for every category $\cat{A}$ and every objects $X, Y \in \cat{A}$
@@
 
-    Lets reconcider the definition of the \emph{product}:
+    Lets reconsider the definition of the \emph{product}:
     For every object $A \in \cat{A}$ and arrows $f_1 \from A \to X, f_2 \from A \to Y$ we get that there exists a unique map

### 2017-08-24 papers/talk-6/paper-6.tex
First stuff for writeup-6

\def\pathToRoot{../../}\input{\pathToRoot headers/header_lecturenotes}

\usepackage{todonotes}

% Shortcuts
\newcommand{\catA}[0]{\cat{A}}
\newcommand{\catB}[0]{\cat{B}}
\newcommand{\fbar}[0]{\bar{f}}
\newcommand{\ftilde}[0]{\tilde{f}}
\newcommand{\gbar}[0]{\bar{g}}
\newcommand{\gtilde}[0]{\tilde{g}}
\newcommand{\dunion}[0]{\sqcup}
\newcommand{\bigdunion}[0]{\bigsqcup}
\renewcommand{\demph}[1]{\emph{#1}} % Definition

% { _ }
\newcommand{\setCon}[1]{\bigl \{ #1 \bigr \}}
% { _ | _ }
\newcommand{\setMap}[2]{\setCon{#1 \,\big|\, #2}}
% ( _ , _ )
\newcommand{\pair}[2]{\left( #1 , #2 \right)}
% [ _ ]
\newcommand{\class}[1]{\left[ #1 \right]}
% < _ >
\newcommand{\choice}[1]{\left< #1 \right>}

% Theorem Environments
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}

% Title page
\title{Special Constructions in Categories}
\author{Maximilian Wuttke}

\begin{document}

\maketitle

\section*{Motivation}
We want to generalize some concepts of sets like the Cartesian product, disjoint union, solution sets and quotient sets.
In this chapter we define six constructions in categories.
We will introduce the constructions \emph{products}, \emph{equalizer} and \emph{pullbacks} and their duals
\emph{coproducts}, \emph{coequalizer} and \emph{pushouts}, respectively.

\section*{Products and Coproducts}

First we want to generalize the Cartesian product to general categories.

\begin{example}[Cartesian Product]
  \label{ex:prod:cart}
  Let $X$ and $Y$ be sets.
  Then we define the \demph{Cartesian Product}
  $$X \times Y := \setMap{\pair{x}{y}}{x \in X, \; y \in Y}. $$
  Also we define the \demph{projection functions} $p_1 \from X \times Y \to X$, $p_2 \from X \times Y \to Y$ in a canonical way:
  \begin{alignat*}{2}
    p_1(x,y) &:= x \\
    p_2(x,y) &:= y
  \end{alignat*}

  Let $A$ be another set and $f_1 \from A \to X, f_2 \from A \to Y$ functions.
  If we want to define a function $\fbar \from A \to X \times Y$, such that
  $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ for every $a \in A$,
  there is only one possible way to define $\fbar \from A \to X \times Y$:
  $$\fbar(a) := \pair{f_1(a)}{f_2(a)}.$$
  We have $p_1(\fbar(a)) = p_1(f_1(a), f_2(b)) = f_1(a)$ and $p_2(\fbar(a)) = p_2(f_1(a), f_2(a)) = f_2(a)$.
  If we have another $\ftilde \from A \to X \times Y$ with $p_1(\ftilde(a)) = f_1(a)$ and $p_2(\ftilde(a)) = f_2(a)$ we show $\fbar(a) = \ftilde(a)$ for every $a \in A$:
  $$\ftilde(a) = \pair{f_1(a)}{f_2(a)} = \pair{p_1(\ftilde(a)}{p_2(\ftilde(a))} = \ftilde(a).$$
  The last step is because we have $\pair{p_1(z)}{p_2(z)} = x$ for every $z \in X \times Y$.
\end{example}

We can use this knowledge to define products in arbitrary categories.

\begin{definition}[Product]
  \label{def:prod}
  Let $\cat{A}$ be a category and $X, Y \in \cat{A}$ objects.
  A \demph{product} of $X$ and $Y$ consists of an object $P$ and maps
  $ \xymatrix{
    X &P \ar[l]_{p_1} \ar[r]^{p_2} & Y
  } $
  with the property that for all objects and maps
  $ \xymatrix{
    X &A \ar[l]_{f_1} \ar[r]^{f_2} & Y
  } $
  in $\cat{A}$, there exists a unique map $\bar{f}\from A \to P$ such that
  \[ \xymatrix{
    & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}} \ar[rdd]^{f_2} & \\
    & P \ar[ld]^{p_1} \ar[rd]_{p_2} & \\
    X & & Y
  } \]
  commutes. The maps $p_1$ and $p_2$ are called the \demph{projections}.
\end{definition}

\begin{lemma}[Products are unique up-to isomorphism]
  \label{lem:prod:uniq}
  For every category $\cat{A}$ and every objects $X, Y \in \cat{A}$
  all products of $X$ and $Y$ are isomorphic. Thus we can speak of \emph{the} product of $X$ and $Y$.

  This fact is also true for all the following constructions but we omit the proofs and will do them in the next chapter.
\end{lemma}

\begin{proof}
  Let $X, Y \in \cat{A}$ and $P$ be a product of $X$ and $Y$ together with $p_1 \from P \to X, p_2 \from P \to Y$
  and $P'$ be a product of $X$ and $Y$ together with $p'_1 \from P' \to X, p'_2 \from P' \to Y$.

  Lets reconsider the definition of the \emph{product}:
  For every object $A \in \cat{A}$ and arrows $f_1 \from A \to X, f_2 \from A \to Y$ we get that there exists a unique map
  $\bar{f} \from A \to P$, such that
  \[ \xymatrix{
      &A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}}
      \ar[rdd]^{f_2}&       \\
              &P \ar[ld]^{p_1} \ar[rd]_{p_2}                  &       \\
      X       &                                               &Y
  } \]
  commutes.
  Now we can instantiate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instantiations we get the commuting diagram
  \[ \xymatrix{
    & P \ar@/_/[ldd]_{p_1} \ar@/^/[rdd]^{p_2} & \\
    & P \ar[ld]^{p_1} \ar@/_/@{.>}[dd]|{\bar{f}} \ar@{.>}[u]|{f} \ar[rd]_{p_2} & \\
    X & & Y \\
    & P' \ar[lu]_{p_1'} \ar@/_/@{.>}[uu]|{\tilde{f}} \ar@{.>}[d]|{f'} \ar[ru]^{p'_2} & \\
    & P' \ar@/^/[luu]^{p'_1} \ar@/_/[ruu]_{p'_2} &
  } \]
  for exactly one $f, f', \tilde{f}, \bar{f}$. With the inner triangles we get:
  \begin{alignat*}{2}
    p_1  \of \tilde{f} &= p'_1 &\qquad p_2  \of \tilde{f} &= p'_2 \\
    p'_1 \of \bar{f}   &= p_1  &       p'_2 \of \bar{f}   &= p_2
  \end{alignat*}
  By substitution we get
  \begin{alignat*}{2}
    p_1   \of (\tilde{f} \of \bar{f}) &= p_1  &\qquad p_2  \of (\tilde{f} \of \bar{f}) &= p_2 \\
    p'_1  \of (\bar{f} \of \tilde{f}) &= p'_1 &       p'_2 \of (\bar{f} \of \tilde{f}) &= p'_2
  \end{alignat*}
  With the outer triangles we get:
  \begin{alignat*}{5}
    p_1  &\;\of&&\; f  &= p_1  &\qquad p_2  &\;\of&\; f  &= p_2 \\
    p'_1 &\;\of&&\; f' &= p'_1 &\qquad p'_2 &\;\of&\; f' &= p'_2
  \end{alignat*}
  Note that $f$ and $f'$ are \emph{unique} and the first equations hold for $f = 1_P$ and $f = \tilde{f} \of \bar{f}$, thus $\tilde{f} \of \bar{f} = 1_P$.
  The second equations holds for $f' = 1_{P'}$ and for $f' = \bar{f} \of \tilde{f}$, thus $\bar{f} \of \tilde{f} = 1_{P'}$.
  Thus $P$ and $P'$ are isomorphic. \qedhere
\end{proof}

\begin{example}[Reversed Cartesian product]
  \label{ex:prod:cart2}
  For sets we can define another product, where we reverse the order of the elements, i.~e.:
  \begin{alignat*}{2}
    X \times' Y &:= \setMap{\pair{y}{x}}{x \in X, \; y \in Y} \\
    p'_1(x,y)   &:= y \\
    p'_2(x,y)   &:= x
  \end{alignat*}
  Now $X \times Y$ and $X \times' Y$ are isomorphic with the involution $i(a, b) := \pair{b}{a}$.
\end{example}

\begin{example}[Posets]
  \label{ex:prod:poset}
  Let $\pair{A}{\le}$ be a poset.
  A \demph{lower bound} for two objects $x, y \in A$ is a element $a \in A$ such that $a \le x$ and $a \le y$.
  A \demph{greatest lover bound} of $x, y \in A$ is a lower bound $z \in A$ of $x$ and $y$ such that for all lower bounds $a \in A$ of $x$ and $y$, $a \le z$.
  With other words: $z$ is a greatest lower bound of $x$ and $y$ if and only if
  $$z \le x \land z \le y \land \forall a \in A, a \le x \land a \le y \Rightarrow a \le z.$$

  When we see the poset $\pair{A}{\le}$ as a category $\catA$ and have a greatest lower bound $z$ of $x, y \in A$, then $z$ is the product of $x$ and $y$.
  To see this, we apply the definition \ref{def:prod}.
  Let $x, y \in \catA$ and $z \in \catA$ be a greatest lower bound of $x$ and $y$.
  Let $a \in \catA$ be another object.
  If we have $f_1 \from a \to x$ and $f_2 \from a \to y$ then by definition of $\catA$, $a$ is a lower bound of $x$ and $y$.
  Now because $z$ is a greatest lower bound, $a \le z$ holds and therefore is a unique map $\fbar \from a \to z$.
  The diagram commutes trivially.
\end{example}

\begin{example}[Direct sum of vector spaces]
  \label{ex:prod:vspace}
  Let $K$ be a field with the operations $+_K$ and $\cdot_K$ and $V, W$ vector spaces over this field with the operations $+_V, \cdot_V, +_W, \cdot_W$.
  Then we define\footnote{One should verify that the defined structure is indeed a vector field.}
  the \demph{direct sum} $V \oplus W$ of $V$ with $V \oplus W := \bigl(V \times W, +, \cdot \bigr)$ and with the following operations:
  \begin{alignat*}{4}
    \pair{v_1}{w_1} + \pair{v_2}{w_2} &:= \pair{v_1 +_V v_2}{w_1 +_W w_2}           &\quad& \forall v_1, v_2 \in V, \forall w_1, w_2 \in W \\
    \alpha \cdot \pair{v}{w}          &:= \pair{\alpha \cdot_V v}{\alpha \cdot_W w} &     & \forall v \in V, \forall w \in W
  \end{alignat*}
  We can now show that $U \oplus V$ is the product of $U$ and $V$ in $\Vect$.
  Therefore we can forget that $U$ and $V$ are vector spaces and follow the proof in example \ref{ex:prod:cart}.
\end{example}

Now we want to abstract the concept of the disjoint union of sets.

\begin{example}[Disjoint union]
  \label{ex:coprod:dunion}
  Let $X$ and $Y$ be sets.
  Let furthermore $0$ and $1$ be some sets with $0 \ne 1$.
  Then we define the \demph{disjoint union} of the sets $X$ and $Y$:
  $$X \dunion Y := (X \times \setCon 0) \cup (Y \times \setCon 1) = \setMap{\pair{x}{0}}{x \in X} \cup \setMap{\pair{y}{1}}{y \in Y}.$$

  Also we define the \demph{coprojection functions} $p_1 \from A \to X \dunion Y$, $p_2 \from Y \to X \dunion Y$:
  \begin{alignat*}{2}
    p_1(x) &:= \pair{x}{0} \\
    p_2(y) &:= \pair{y}{1}
  \end{alignat*}

  Let $A$ be another set and $f_1 \from X \to A, f_2 \from Y \to A$ functions.
  If we want to define a function $\fbar \from X \dunion Y \to A$, such that
  $\fbar(p_1(x)) = f_1(x)$ and $\fbar(p_2(y)) = f_2(y)$ for every $x \in X, y \in Y$,
  there is only one possible way to define $\fbar$:
  \begin{alignat*}{2}
    \fbar(x, 0) &:= f_1(x) \\
    \fbar(y, 1) &:= f_2(y).
  \end{alignat*}
\end{example}

We can again use this knowledge to define coproducts in arbitrary categories.

\begin{definition}[Coproduct]
  \label{def:coproduct}
  Let $\cat{A}$ be a category and $X, Y \in \cat{A}$.
  A \demph{coproduct} of $X$ and $Y$ consists of an object $S$ and maps
  $ \xymatrix{
    X \ar[r]^{p_1} & S & Y \ar[l]_{p_2}
  } $
  with the property that for all objects and maps
  $ \xymatrix{
    X \ar[r]^{f_1} & A & Y \ar[l]_{f_2}
  } $
  in $\cat{A}$, there exists a unique map $\bar{f}\from S \to A$ such that
  \[ \xymatrix{
    X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
    & S \ar@{.>}[d]|{\bar{f}} & \\
    & A &
  } \]
  commutes. The maps $p_1$ and $p_2$ are called the \demph{coprojections}.
\end{definition}

\begin{example}[Direct sum of vector spaces]
  \label{ex:coprod:vspace}
  In the example \ref{ex:prod:vspace} we have seen that the direct sum of two vectorspaces $U$ and $V$ is the product of $V$ and $W$ in $\Vect$.
  We will now see that it is \emph{also} the coproduct.
  In general products and coproducts are not the same as we have seen in the examples \ref{ex:prod:cart} and \ref{ex:coprod:dunion}.
  First we have to define the two coprojection functions $p_1 \from V \to V \oplus W, p_2 \from W \to V \oplus W$:
  \begin{alignat*}{2}
    p_1(x) &:= \pair{x}{0} \\
    p_2(y) &:= \pair{0}{y}
  \end{alignat*}
  Then let $A \in \Vect$ and $f_1 \from V \to A, f_2 \from W \to A$ be vector morphisms.
  We define the vector morphism $\fbar \from V \oplus W \to A$ with $\fbar(v, w) := f_1(v) + f_2(w)$.
  By construction and because $f_1$ and $f_2$ are vector morphims we have:
  \begin{alignat*}{4}
    \fbar(p_1(u)) &= \fbar(u, 0) &= f_1(u) + f_2(0) &= f_1(u) \\
    \fbar(p_2(v)) &= \fbar(0, v) &= f_1(0) + f_2(v) &= f_2(v).
  \end{alignat*}
  Now we have to show that $\fbar$ is the only vector morphims with this property.
  Let $\ftilde \from V \oplus W \to A$ be another vector morphims with $\ftilde(p_1(u)) = f_1(u)$ and $\ftilde(p_2(v)) = f_2(v)$.
  Then we have for every $v \in V$ and $w \in W$:
  $$ \fbar(v,w) = f_1(v) + f_2(w)\ = \ftilde(p_1(v)) + \ftilde(p_2(w)) = \ftilde(p_1(v) + p_2(w)) = \ftilde(v, w). $$
  Thus $\fbar = \ftilde$ and $V \oplus W$ is indeed the product of $V$ and $W$ in the the category $\Vect$.
\end{example}

\section*{Equalizers and Coequalizer}

%% XXX Weiterhin erst Beispiel und dann Definition oder erst Definition und dann Beispiel?
Now We present abstract concepts and give examples.

\begin{definition}[Equalizer]
  \label{def:equa}
  Let $\cat{A}$ be a category and
  $ \xymatrix{
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  } $
  a diagram in $\cat{A}$.
  An \demph{equalizer} of this diagram is a commuting diagram called \demph{fork} \\
  $ \xymatrix{
    E \ar[r]^i & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  } $
  with the property that for all other forks
  $ \xymatrix{
    A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  } $
  in $\cat{A}$, there exists a unique map $\bar{f}\from A \to E$ such that
  \[ \xymatrix{
    A \ar[rd]^f \ar@{.>}[dd]|{\bar{f}} & & \\
    & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
    E \ar[ru]_i & &
  } \]
  commutes.
\end{definition}

\begin{example}[Solution sets]
  Let $X, Y$ be sets and $s, t \from X \to Y$ functions.
  We define $E := \setMap{x \in X}{s(x)=t(x)}$.
  Because $E \subseteq X$ we can define the \emph{inclusion function} $i \from E \to X$ with $i(e) := e$.
  Now we show that the fork
  $ \xymatrix{
    E \ar[r]^i & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  } $
  is the equalizer of $s$ and $t$. This diagram commutes by construction.
  Let $A$ be another set and
  $ \xymatrix{
    A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  } $
  be another fork.
  We then define $\fbar \from A \to E$ with $\fbar(a) = f(a)$.
  Then $i \of \fbar = f$ holds by definition.
  Let $\ftilde \from A \to E$ be another function such that $i \of \ftilde = f$.
  Then $\ftilde(a) = i (\ftilde(a)) = f(a) = i (\fbar(a)) = \fbar(a)$.
\end{example}

\todo{Mehr Beispiele dazu?}

\begin{definition}[Coequalizer]
  \label{def:coequa}
  Let $\cat{A}$ be a category and
  $ \xymatrix{
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
  } $
  a diagram in $\cat{A}$.
  A \demph{coequalizer} of this diagram is a commuting diagram called \demph{cofork} \\
  $ \xymatrix{
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
  } $
  with the property that for all other coforks
  $ \xymatrix{
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^f & A
  } $
  in $\cat{A}$, there exists a unique map $\bar{f}\from C \to A$ such that
  \[ \xymatrix{
    & & A  \\
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[ru]^f \ar[rd]_i \\
    & & C \ar@{.>}[uu]|{\bar{f}}
  } \]
  commutes.
\end{definition}

For our next example we need the definition of the reflexive transitive closure.

\begin{definition}[Symmetric transitive closure]
  Let $R$ be a binary relation over a set $X$.
  $\hat{R} := R \cup R^{-1}$ is the symmetric closure of $R$ where $R^{-1}$ is defined by
  $R^{-1} := \setMap{\pair{b}{a}}{\pair{a}{b} \in R}$.
  The \demph{symmetric transitive closure} of $R$ is $\sim := \bigcup_{n\in\nat} {(\hat{R})^n}$
  where ${\hat R}^n$ is defined recursively with
  ${\hat R}^0 := \setMap{\pair{a}{a}}{a \in X}$ and
  ${\hat R}^{n+1} := R_n \of \hat R = \setMap{\pair{a}{c}}{\pair{a}{b} \in {\hat R}^n \land \pair{b}{c} \in \hat R}$.
  It can be shown that $\sim$ is the \emph{least equivalence relation} containing $R$.
\end{definition}

\begin{example}[Quotient sets]
  Let $X, Y$ be sets and $s, t \from X \to Y$ be functions.
  We have to give a set $C$ and a function $i$ such that the diagram
  $ \xymatrix{
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
  } $
  commutes.
  Therefore we define the relation
  $R := \setMap{\pair{s(a)}{t(a)}}{a \in Y} \subseteq Y^2$ the relation of values in $Y$ that correspond from the same $x \in X$.
  Let $\sim \subseteq Y^2$ be the symmetric transitive closure of $R$.
  Now we define $C := Y / \sim = \setMap{\class y}{y \in Y}$ where $\class y := \setMap{y'}{y \sim y'}$ are the equivalence classes.
  For the function $i \from Y \to E$ we choose $i(y) := \class y$.
  Then the diagram
  $ \xymatrix{
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^i & C
  } $
  commutes because for $x \in X$, $s(x)$ and $t(x)$ are in $R$ and therefore also in $\sim$; because $s(x) \sim t(x)$ we have $i(s(x)) = \class{s(x)} = \class{t(x)} = i(t(x))$.
  Now let
  $ \xymatrix{
    X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[r]^f & A
  } $
  be a cofork.
  By induction we can show that $f$ ``does not care'' about equal values, i.~e. $f(y_1) = f(y_2)$ whenever $y_1 \sim y_2$.

  To define the function $\fbar \from C \to A$ we need a choice function $\choice \cdot$ that returns for every equivalence class in $\sim$ a member of it, i.~e.
  $\forall c \in C, \forall y \in e, \choice{e} \sim y$.
  Therefore we also know $\forall c \in C, \class {\choice c} = c$.

  Now we can define $\fbar \from C \to A$ by $\fbar(e) := f(\choice e)$.
  We now have  $\fbar(i(y)) = f(\choice{\class y}) = f(y)$ because $\choice{\class y} \sim y$ (by reflexivity of $\sim$).
  Now we only have to show that this $\fbar$ is unique. So we assume another function $\ftilde \from C \to A$ with $\ftilde \of i = f$. But then we have
  $$ \fbar(c) = f(\choice c) = \ftilde(i(\choice c)) = \ftilde(\class{\choice c}) = \ftilde(c).$$
\end{example}

\section*{Pullbacks and Pushouts}

\begin{definition}[Pullback]
  \label{def:pullback}
  Let $\cat{A}$ be a category and
  $ \xymatrix{
    X \ar[r]^s & Z & Y\ar[l]_t \\
  } $
  a diagram in $\cat{A}$.
  A \demph{pullback} of this diagram is a \demph{pullback square}
  \[ \xymatrix{
    P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
    X \ar[r]^s & Z
  } \]
  with the property that for all other pull back squares
  \[ \xymatrix{
    A \ar[r]_{f_2} \ar[d]^{f_1} & Y \ar[d]_{t} \\
    X \ar[r]^s & Z
  } \]
  in $\cat{A}$, there exists a unique map $\bar{f}\from A \to P$ such that
  \[ \xymatrix{
    A \ar[drr]^{f_2} \ar[ddr]_{f_1} \ar@{.>}[dr]|{\bar{f}} & & \\
    & P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
    & X \ar[r]^s & Z
  } \]
  commutes.
\end{definition}

The pullback is also called ``fibred product''. This will be clear in the following example:
\todo{Besser erklären, was das Beispiel damit zu tun hat}

\begin{example}[Fibred product]
  When the pullback square
  \[ \xymatrix{
    P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
    X \ar[r]^s & Z
  } \]
  is the pullback of
  $ \xymatrix{
    X \ar[r]^s & Z & Y\ar[l]_t \\
  } $
  and $Z$ is a terminal object then $P$ is the product of $X$ and $Y$.
\end{example}

\begin{example}[Pullbacks in sets]
  Let $X, Y$ be sets and $s \from X \to Z, t \from Y \to Z$ be functions.
  Then $P := \setMap{\pair{x}{y}}{s(x) = t(y)}$ with the projection functions
  $p_1(x,y) := x$ and $p_2(x, y) = y$ is the pullback of
  $ \xymatrix{
    X \ar[r]^s & Z & Y\ar[l]_t \\
  } $.
  The pullback square commutes by construction.
  Now let $A$ be a set and $f_1 \from A \to X, f_2 \from A \to Y$ be functions with $s \of f_1 = t \of f_2$.
  Then we define $\fbar \from A \to P$ by
  $\fbar(a) := \pair{f_1(a)}{f_2(a)}$.
  This is well defined because for $a \in A$ we have $\fbar(a) = \pair{f_1(a)}{f_2(a)} \in P \Leftrightarrow s(f_1(a) = t(f_2(a))$ wich is true by assumption.
  We also have $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ by constructions.
  We also see that $\fbar$ is unique with this property.
\end{example}

\begin{definition}[Pushout]
  \label{def:pushout}
  Let $\cat{A}$ be a category and
  $ \xymatrix{
    X & Z \ar[l]_s \ar[r]^t & Y \\
  } $
  a diagram in $\cat{A}$.
  A \demph{pushout} of this diagram is a \demph{pushout square} \\
  \[ \xymatrix{
    Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \\
    Y \ar[r]^{p_2} & P
  } \]
  with the property that for all other pushout squares
  \[ \xymatrix{
    Z \ar[r]_s \ar[d]^t & X \ar[d]_{f_1} \\
    Y \ar[r]^{f_2} & A
  } \]
  in $\cat{A}$, there exists a unique map $\bar{f}\from P \to A$ such that
  \[ \xymatrix{
    Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
    Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\bar{f}} & \\
    & & A  \\
  } \]
  commutes.
\end{definition}

\todo{Beispiele}

\section*{Conclusion}
We have seen much analogous definitions and proofs.
Our definitions follow a simple ``pattern'':
\begin{itemize}
  \item First we have a some objects or a commuting diagram ($D_0$).
  \item Then we ``insert'' a object $P$ and get a new commuting diagram ($D_1$) where we ``connect'' $P$ with the objects of $D_0$ with ``projection functions''.
  \item Then we ``replace'' this object with a new object, called $A$, and replace the ``projection functions'' with new functions. We get a new diagram ($D_2$).
  \item For each $A$ there exists a unique map $\fbar$ such if we ``merge'' $D_1$ and $D_2$ we get a new commuting diagram $D_3$, where $A$ and $P$ are connected with $\fbar$.
\end{itemize}

In the next chapter we will see that all this six constructions are instances of a more general concept: limits and colimits.
We will formalize the ``pattern'' above.

We have assumed that all instances of the six special constructions above are unique up-to isomorphism
and have shown this fact only for products. We will see a proof for all limits and colimits.

\end{document}

% vim: ts=2 ss=2 sw=2 expandtab

### 2017-08-24 papers/talk-7/paper-7.tex
first examples and explanations

@@
 
+In category theory we take limits over diagrams.
+We begin by formally defining what a diagram is.
+
 \begin{definition}
@@
 \end{definition}
+\begin{example}
+  \label{diagramexample}
+  This is an example of a diagram of shape $\bm I$. \\[1cm]
+  \begin{minipage}{0.3\linewidth}
+    \hspace{2cm}
+    \begin{tikzcd}
+      & \bullet \arrow{d} \\
+      \bullet\arrow{r} & \bullet
+    \end{tikzcd}\\
+  \end{minipage}%
+  \begin{minipage}{0.2\linewidth}
+    $\qquad\quad\overset{D}{\Longrightarrow}$
+  \end{minipage}%
+  \begin{minipage}{0.1\linewidth}
+    \begin{tikzpicture}
+      \draw(1,0) circle (1.5cm);
+      \begin{tikzcd}
+        & \bullet \arrow{d} \\
+        \bullet\arrow{r} & \bullet
+      \end{tikzcd}
+    \end{tikzpicture}
+  \end{minipage}\\[2em]
+  \begin{minipage}{0.3\linewidth}
+    \hspace{2cm}
+    $\qquad \bm I$
+  \end{minipage}%
+  \begin{minipage}{0.2\linewidth}
+    \
+  \end{minipage}%
+  \begin{minipage}{0.3\linewidth}
+    \hspace{2cm}
+    $\qquad \mathscr C$
+  \end{minipage}\\[1em]
+\end{example}
 
-%% TODO: explain reasoning behind this definition and give example
+The functor $D$ yields
+a part of the category $\mathscr C$ that essentially looks like $\bm I$.
+We also say $D$ \textbf{identifies} diagrams of shape $\bm I$,
+which is why we use $\bm I$ to denote the shape.
+Especially with diagrams of such a small shape
+there might be many parts in $\mathscr C$ that have the shape of $\bm I$.
+We therefore get all different diagrams of that shape by considering
+all different functors $D$.
+Diagrams do not need to be small though. Their shape can also be infinite.\\
+We are typically interested in diagrams of a specific category,
+$\mathscr C$ in the example. Therefore we usually leave out $\bm I$ and $D$,
+and we just draw the identified part of $\mathscr C$ instead.
 
@@
 \end{remark}
+%% TODO: Notation DI, D(I), D_I ??
+\begin{example}
+  A cone over the diagram of
+  \hyperref[diagramexample]{Example \ref*{diagramexample}} looks like this.
+  \[
+    \begin{tikzcd}
+      U \arrow[bend right=30]{ddr}{f_I} \arrow[bend left=30]{drr}{f_k} \arrow[bend left=20, swap]{ddrr}{f_J}&   &\\
+      & & D_K \arrow{d}{D_n} \\
+      & D_I \arrow{r}{D_m} & D_J
+    \end{tikzcd}
+  \]
+  The commuting condition says that $D_m \circ f_I = f_J$ and $D_n \circ f_K = f_J$.
+\end{example}
 
 %% TODO: write a sentence like "In other words, ..." and give example. example might be tricky to draw
+
+Now we define what it means for a cone to be universal. Intuitively,
+a cone is universal if it satisfies a property that makes it universal among cones,
+specifically, that all other cones factor through it uniquely.
 
@@
   for all other cones $(V \overset{g_I}{\to} D(I))_{I\in \bm{I}}$ there is a unique
-  map $h \from V \to U$, such that for all $I \in \bm{I}$ the following triangles commute.
+  map $h \from V \to U$, such that for all $I \in \bm{I}$
+  the following triangles commute.
   \[
@@
 
-%TODO: give example. might be tricky to draw.
+\begin{example}
+    \[
+    \begin{tikzcd}
+      {\color{red}V}
+      \arrow[bend right=40,color=red]{dddrr}{g_I}
+      \arrow[bend left=40, color=red]{ddrrr}{g_k}
+      \arrow[bend left=30, color=red, swap]{dddrrr}{g_J}
+      \arrow[dashed,swap]{dr}{h}
+      & & & \\
+      & {\color{blue}U}
+      \arrow[bend right=30,color=blue]{ddr}{f_I}
+      \arrow[bend left=30, color=blue]{drr}{f_k}
+      \arrow[bend left=20, color=blue, swap]{ddrr}{f_J}&   &\\
+      & & & D_K \arrow{d}{D_n} \\
+      & & D_I \arrow{r}{D_m} & D_J
+    \end{tikzcd}
+  \]
+  The universal cone {\color{blue}U}
+  is drawn in blue and the other cone {\color{red}V} is drawn in red.
+  Both cones obviously have to satisfy the commuting condition for cones.
+  Additionally the commuting condition for universal cones states that
+  $g_I = f_I \circ h$, $g_J = f_J \circ h$, and $g_K = f_k \circ h$.
+\end{example}
+
+\begin{definition}
+  A limit is a universal cone over a diagram.
+\end{definition}
+
+A limit is therefore characterized by the kind of diagram
+over which it is a universal cone, meaning we get different instances of limits
+by choosing diagrams of different shapes.
+
+%% TODO: terminal object, pullback, equalizer, product.
 
@@
 It is possible to build new limits...
+%% TODO: example
+%% TODO: general limits
+%% TODO: possibly other construction mentioned in awodey.
 
 \chapter{Colimits}
-Just dualize it...
+Colimits are the dual of limits.
+\begin{definition}
+  A \textbf{colimit} is a universal co-cone over a diagram.
+\end{definition}
+%% TODO: complete definition
+
+\[
+    \begin{tikzcd}
+      & D_K \arrow{d}{D_n}
+      \arrow[bend left=30,color=blue]{ddr}{f_K}
+      \arrow[bend left=50,color=red]{dddrr}{g_K} & & \\
+      D_I \arrow{r}{D_m}
+      \arrow[bend right=30,color=blue]{drr}{f_I}
+      \arrow[bend right=40,color=red]{ddrrr}{g_I} &
+      D_J \arrow[bend left=20,swap, color=blue]{dr}{f_J}
+      \arrow[bend left=40,color=red]{ddrr}{g_J} & & \\
+      & & {\color{blue}U} \arrow[dashed]{dr}{h}& \\
+      & & & {\color{red}V}
+    \end{tikzcd}
+\]
+%% TODO: example drawing
 

### 2017-08-24 papers/talk-7/paper-7.tex
introduction examples

@@
 \chapter{Introduction}
-We have seen examples of limits.\\
-Insert examples here.\\
+We have seen three special structures in categories.\\[1em]
+\begin{minipage}{.35\linewidth}
+  \begin{tikzcd}
+    & V \arrow[bend right=40]{ddl} \arrow[bend left=40]{ddr} \arrow[dashed]{d} &\\
+    & U \arrow{dl} \arrow{dr} &\\
+    \bullet &&\bullet
+  \end{tikzcd}
+\end{minipage}%
+\begin{minipage}{.35\linewidth}
+  \begin{tikzcd}
+    V
+    \arrow[bend left=40]{drr}
+    \arrow[bend right=40]{ddr}
+    \arrow[dashed,swap]{dr}
+    && \\
+    &U
+    \arrow{r}
+    \arrow{d}
+    &\bullet \arrow{d} \\
+    & \bullet \arrow{r} & \bullet
+  \end{tikzcd}
+\end{minipage}%
+\begin{minipage}{.3\linewidth}
+  \begin{tikzcd}
+    V \arrow[dashed]{dd} \arrow{dr}&& \\
+    & \bullet \arrow[shift left]{r} \arrow[shift right]{r} & \bullet \\\
+    U \arrow{ur} &&
+  \end{tikzcd}
+\end{minipage}\\[2em]
 Now we generalize them...

### 2017-08-24 papers/talk-6/paper-6.tex
writeup6: Erste Korrektur-Iteration

@@
 \newcommand{\choice}[1]{\left< #1 \right>}
+% explanation over relation sign
+\newcommand{\explainRel}[2]{\stackrel{\text{#1}}{#2}}
 
@@
 \title{Special Constructions in Categories}
+\subtitle{Examples for limits and colimits}
+
 \author{Maximilian Wuttke}
@@
   Then we define the \demph{Cartesian Product}
-  $$X \times Y := \setMap{\pair{x}{y}}{x \in X, \; y \in Y}. $$
+  $$X \times Y := \setMap{\pair{x}{y}}{x \in X, \; y \in Y}.$$
   Also we define the \demph{projection functions} $p_1 \from X \times Y \to X$, $p_2 \from X \times Y \to Y$ in a canonical way:
@@
   If we want to define a function $\fbar \from A \to X \times Y$, such that
-  $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ for every $a \in A$,
+  $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ for every $a \in A$ holds,
   there is only one possible way to define $\fbar \from A \to X \times Y$:
   $$\fbar(a) := \pair{f_1(a)}{f_2(a)}.$$
-  We have $p_1(\fbar(a)) = p_1(f_1(a), f_2(b)) = f_1(a)$ and $p_2(\fbar(a)) = p_2(f_1(a), f_2(a)) = f_2(a)$.
+  By construction we have $p_1(\fbar(a)) = p_1(f_1(a), f_2(b)) = f_1(a)$ and $p_2(\fbar(a)) = p_2(f_1(a), f_2(a)) = f_2(a)$.
   If we have another $\ftilde \from A \to X \times Y$ with $p_1(\ftilde(a)) = f_1(a)$ and $p_2(\ftilde(a)) = f_2(a)$ we show $\fbar(a) = \ftilde(a)$ for every $a \in A$:
-  $$\ftilde(a) = \pair{f_1(a)}{f_2(a)} = \pair{p_1(\ftilde(a)}{p_2(\ftilde(a))} = \ftilde(a).$$
-  The last step is because we have $\pair{p_1(z)}{p_2(z)} = x$ for every $z \in X \times Y$.
+  $$\fbar(a) = \pair{f_1(a)}{f_2(a)} = \pair{p_1(\ftilde(a)}{p_2(\ftilde(a))} = \ftilde(a).$$
+  The last step is because we have $\pair{p_1(z)}{p_2(z)} = z$ for every $z \in X \times Y$.
 \end{example}
@@
   } $
-  in $\cat{A}$, there exists a unique map $\bar{f}\from A \to P$ such that
+  in $\cat{A}$, there exists a unique map $\fbar\from A \to P$ such that
   \[ \xymatrix{
-    & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}} \ar[rdd]^{f_2} & \\
+    & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\fbar} \ar[rdd]^{f_2} & \\
     & P \ar[ld]^{p_1} \ar[rd]_{p_2} & \\
@@
   and $P'$ be a product of $X$ and $Y$ together with $p'_1 \from P' \to X, p'_2 \from P' \to Y$.
-
-  Lets reconsider the definition of the \emph{product}:
-  For every object $A \in \cat{A}$ and arrows $f_1 \from A \to X, f_2 \from A \to Y$ we get that there exists a unique map
-  $\bar{f} \from A \to P$, such that
-  \[ \xymatrix{
-      &A \ar[ldd]_{f_1} \ar@{.>}[d]|{\bar{f}\vphantom{\bar{\bar{f}}}}
-      \ar[rdd]^{f_2}&       \\
-              &P \ar[ld]^{p_1} \ar[rd]_{p_2}                  &       \\
-      X       &                                               &Y
-  } \]
-  commutes.
-  Now we can instantiate $A$ with $P'$, $f_1$ with $p_1$, $f_2$ with $p_2$. By analogous instantiations we get the commuting diagram
+  We apply definition \ref{def:prod} multiple times and get the following commuting diagram:
   \[ \xymatrix{
     & P \ar@/_/[ldd]_{p_1} \ar@/^/[rdd]^{p_2} & \\
-    & P \ar[ld]^{p_1} \ar@/_/@{.>}[dd]|{\bar{f}} \ar@{.>}[u]|{f} \ar[rd]_{p_2} & \\
+    & P \ar[ld]^{p_1} \ar@/_/@{.>}[dd]|{\fbar} \ar@{.>}[u]|{f} \ar[rd]_{p_2} & \\
     X & & Y \\
-    & P' \ar[lu]_{p_1'} \ar@/_/@{.>}[uu]|{\tilde{f}} \ar@{.>}[d]|{f'} \ar[ru]^{p'_2} & \\
+    & P' \ar[lu]_{p_1'} \ar@/_/@{.>}[uu]|{\ftilde} \ar@{.>}[d]|{f'} \ar[ru]^{p'_2} & \\
     & P' \ar@/^/[luu]^{p'_1} \ar@/_/[ruu]_{p'_2} &
   } \]
-  for exactly one $f, f', \tilde{f}, \bar{f}$. With the inner triangles we get:
+  for exactly one $f, f', \fbar, \ftilde$. With the inner triangles we get:
   \begin{alignat*}{2}
-    p_1  \of \tilde{f} &= p'_1 &\qquad p_2  \of \tilde{f} &= p'_2 \\
-    p'_1 \of \bar{f}   &= p_1  &       p'_2 \of \bar{f}   &= p_2
+    p_1  \of \ftilde &= p'_1 &\qquad p_2  \of \ftilde &= p'_2 \\
+    p'_1 \of \fbar   &= p_1  &       p'_2 \of \fbar   &= p_2
   \end{alignat*}
-  By substitution we get
+  By substitution we get:
   \begin{alignat*}{2}
-    p_1   \of (\tilde{f} \of \bar{f}) &= p_1  &\qquad p_2  \of (\tilde{f} \of \bar{f}) &= p_2 \\
-    p'_1  \of (\bar{f} \of \tilde{f}) &= p'_1 &       p'_2 \of (\bar{f} \of \tilde{f}) &= p'_2
+    p_1   \of (\ftilde \of \fbar) &= p_1  &\qquad p_2  \of (\ftilde \of \fbar) &= p_2 \\
+    p'_1  \of (\fbar \of \ftilde) &= p'_1 &       p'_2 \of (\fbar \of \ftilde) &= p'_2
   \end{alignat*}
@@
   \end{alignat*}
-  Note that $f$ and $f'$ are \emph{unique} and the first equations hold for $f = 1_P$ and $f = \tilde{f} \of \bar{f}$, thus $\tilde{f} \of \bar{f} = 1_P$.
-  The second equations holds for $f' = 1_{P'}$ and for $f' = \bar{f} \of \tilde{f}$, thus $\bar{f} \of \tilde{f} = 1_{P'}$.
+  Note that $f$ and $f'$ are \emph{unique} and the first equations hold for $f = 1_P$ and $f = \ftilde \of \fbar$, thus $\ftilde \of \fbar = 1_P$.
+  The second equations holds for $f' = 1_{P'}$ and for $f' = \fbar \of \ftilde$, thus $\fbar \of \ftilde = 1_{P'}$.
   Thus $P$ and $P'$ are isomorphic. \qedhere
@@
     X \times' Y &:= \setMap{\pair{y}{x}}{x \in X, \; y \in Y} \\
-    p'_1(x,y)   &:= y \\
-    p'_2(x,y)   &:= x
+    p'_1(y,x)   &:= x \\
+    p'_2(y,x)   &:= y
   \end{alignat*}
@@
   Then we define\footnote{One should verify that the defined structure is indeed a vector field.}
-  the \demph{direct sum} $V \oplus W$ of $V$ with $V \oplus W := \bigl(V \times W, +, \cdot \bigr)$ and with the following operations:
+  the \demph{direct sum} $V \oplus W := \bigl(V \times W, +, \cdot \bigr)$ as a new vector field with the following operations:
   \begin{alignat*}{4}
@@
   \end{alignat*}
-  We can now show that $U \oplus V$ is the product of $U$ and $V$ in $\Vect$.
-  Therefore we can forget that $U$ and $V$ are vector spaces and follow the proof in example \ref{ex:prod:cart}.
+  We can show that $U \oplus V$ is the product of $U$ and $V$ in $\Vect$.
+  The proof follows the same structure as in example \ref{ex:prod:cart}.
 \end{example}
@@
 
-We can again use this knowledge to define coproducts in arbitrary categories.
+Again we can use this knowledge to define coproducts in arbitrary categories.
 
@@
   } $
-  in $\cat{A}$, there exists a unique map $\bar{f}\from S \to A$ such that
+  in $\cat{A}$, there exists a unique map $\fbar\from S \to A$ such that
   \[ \xymatrix{
     X \ar[rdd]_{f_1}\ar[rd]^{p_1} & & Y \ar[ldd]^{f_2} \ar[ld]_{p_2} \\
-    & S \ar@{.>}[d]|{\bar{f}} & \\
+    & S \ar@{.>}[d]|{\fbar} & \\
     & A &
@@
   \label{ex:coprod:vspace}
-  In the example \ref{ex:prod:vspace} we have seen that the direct sum of two vectorspaces $U$ and $V$ is the product of $V$ and $W$ in $\Vect$.
+  In the example \ref{ex:prod:vspace} we have seen that the direct sum of two vector spaces $U$ and $V$ is the product of $V$ and $W$ in $\Vect$.
   We will now see that it is \emph{also} the coproduct.
@@
   We define the vector morphism $\fbar \from V \oplus W \to A$ with $\fbar(v, w) := f_1(v) + f_2(w)$.
-  By construction and because $f_1$ and $f_2$ are vector morphims we have:
+  By construction and because $f_1$ and $f_2$ are vector morphism we have:
   \begin{alignat*}{4}
@@
   \end{alignat*}
-  Now we have to show that $\fbar$ is the only vector morphims with this property.
-  Let $\ftilde \from V \oplus W \to A$ be another vector morphims with $\ftilde(p_1(u)) = f_1(u)$ and $\ftilde(p_2(v)) = f_2(v)$.
+  Now we have to show that $\fbar$ is the only vector morphism with this property.
+  Let $\ftilde \from V \oplus W \to A$ be another vector morphism with $\ftilde(p_1(u)) = f_1(u)$ and $\ftilde(p_2(v)) = f_2(v)$.
   Then we have for every $v \in V$ and $w \in W$:
   $$ \fbar(v,w) = f_1(v) + f_2(w)\ = \ftilde(p_1(v)) + \ftilde(p_2(w)) = \ftilde(p_1(v) + p_2(w)) = \ftilde(v, w). $$
-  Thus $\fbar = \ftilde$ and $V \oplus W$ is indeed the product of $V$ and $W$ in the the category $\Vect$.
+  Thus $\fbar = \ftilde$ and $V \oplus W$ is indeed the product of $V$ and $W$ in the category $\Vect$.
 \end{example}
@@
 %% XXX Weiterhin erst Beispiel und dann Definition oder erst Definition und dann Beispiel?
-Now We present abstract concepts and give examples.
+Now we present more abstract concepts and give examples.
 
@@
   a diagram in $\cat{A}$.
-  An \demph{equalizer} of this diagram is a commuting diagram called \demph{fork} \\
+  An \demph{equalizer} of this diagram is a commuting diagram called \demph{fork}
   $ \xymatrix{
@@
   } $
-  in $\cat{A}$, there exists a unique map $\bar{f}\from A \to E$ such that
+  in $\cat{A}$, there exists a unique map $\fbar\from A \to E$ such that
   \[ \xymatrix{
-    A \ar[rd]^f \ar@{.>}[dd]|{\bar{f}} & & \\
+    A \ar[rd]^f \ar@{.>}[dd]|{\fbar} & & \\
     & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
@@
   } $
-  in $\cat{A}$, there exists a unique map $\bar{f}\from C \to A$ such that
+  in $\cat{A}$, there exists a unique map $\fbar\from C \to A$ such that
   \[ \xymatrix{
@@
     X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \ar[ru]^f \ar[rd]_i \\
-    & & C \ar@{.>}[uu]|{\bar{f}}
+    & & C \ar@{.>}[uu]|{\fbar}
   } \]
@@
 
-For our next example we need the definition of the reflexive transitive closure.
+For our next example we need the definition of the symmetric reflexive transitive closure.
 
-\begin{definition}[Symmetric transitive closure]
+\begin{definition}[Symmetric reflexive transitive closure]
   Let $R$ be a binary relation over a set $X$.
@@
   $R^{-1} := \setMap{\pair{b}{a}}{\pair{a}{b} \in R}$.
-  The \demph{symmetric transitive closure} of $R$ is $\sim := \bigcup_{n\in\nat} {(\hat{R})^n}$
+  The \demph{symmetric transitive closure} of $R$ is $\sim := \bigcup_{n\in\nat} {\hat R}^n$
   where ${\hat R}^n$ is defined recursively with
@@
   Therefore we define the relation
-  $R := \setMap{\pair{s(a)}{t(a)}}{a \in Y} \subseteq Y^2$ the relation of values in $Y$ that correspond from the same $x \in X$.
-  Let $\sim \subseteq Y^2$ be the symmetric transitive closure of $R$.
+  $R := \setMap{\pair{s(a)}{t(a)}}{a \in Y} \subseteq Y^2$.
+  Let $\sim \; \subseteq Y^2$ be the symmetric reflexive transitive closure of $R$.
   Now we define $C := Y / \sim = \setMap{\class y}{y \in Y}$ where $\class y := \setMap{y'}{y \sim y'}$ are the equivalence classes.
@@
 
-  To define the function $\fbar \from C \to A$ we need a choice function $\choice \cdot$ that returns for every equivalence class in $\sim$ a member of it, i.~e.
-  $\forall c \in C, \forall y \in e, \choice{e} \sim y$.
-  Therefore we also know $\forall c \in C, \class {\choice c} = c$.
+  To define the function $\fbar \from C \to A$ we need a choice function $\choice \cdot \from C \to Y$ that returns for every equivalence class in $\sim$ a member of it, i.~e.
+  $\forall c \in C, \forall y \in Y, y \in c \Leftrightarrow y \sim \choice{c}$
+  which is equivalent to $\forall c \in C, \class {\choice c} = c$.
 
   Now we can define $\fbar \from C \to A$ by $\fbar(e) := f(\choice e)$.
-  We now have  $\fbar(i(y)) = f(\choice{\class y}) = f(y)$ because $\choice{\class y} \sim y$ (by reflexivity of $\sim$).
-  Now we only have to show that this $\fbar$ is unique. So we assume another function $\ftilde \from C \to A$ with $\ftilde \of i = f$. But then we have
-  $$ \fbar(c) = f(\choice c) = \ftilde(i(\choice c)) = \ftilde(\class{\choice c}) = \ftilde(c).$$
+  We now have  $\fbar(i(y)) = f(\choice{\class y}) = f(y)$ because $y \sim \choice{\class y}$ (by the choice axiom and reflexivity of $\sim$).
+  Now we only have to show that this $\fbar$ is unique in respect to that property.
+  So we assume another function $\ftilde \from C \to A$ with $\ftilde \of i = f$. But then we have
+  $$ \fbar(c) \explainRel{Def}= f(\choice c) \explainRel{Ass.}= \ftilde(i(\choice c)) \explainRel{Def}= \ftilde(\class{\choice c}) \explainRel{Choice}= \ftilde(c).$$
 \end{example}
@@
   } \]
-  in $\cat{A}$, there exists a unique map $\bar{f}\from A \to P$ such that
+  in $\cat{A}$, there exists a unique map $\fbar\from A \to P$ such that
   \[ \xymatrix{
-    A \ar[drr]^{f_2} \ar[ddr]_{f_1} \ar@{.>}[dr]|{\bar{f}} & & \\
+    A \ar[drr]^{f_2} \ar[ddr]_{f_1} \ar@{.>}[dr]|{\fbar} & & \\
     & P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
@@
   $\fbar(a) := \pair{f_1(a)}{f_2(a)}$.
-  This is well defined because for $a \in A$ we have $\fbar(a) = \pair{f_1(a)}{f_2(a)} \in P \Leftrightarrow s(f_1(a) = t(f_2(a))$ wich is true by assumption.
+  This is well defined because for $a \in A$ we have $\fbar(a) = \pair{f_1(a)}{f_2(a)} \in P \Leftrightarrow s(f_1(a) = t(f_2(a))$ which is true by assumption.
   We also have $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ by constructions.
@@
   } \]
-  in $\cat{A}$, there exists a unique map $\bar{f}\from P \to A$ such that
+  in $\cat{A}$, there exists a unique map $\fbar\from P \to A$ such that
   \[ \xymatrix{
     Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \ar[ddr]^{f_1} & \\
-    Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\bar{f}} & \\
+    Y \ar[r]^{p_2} \ar[drr]_{f_2}& P \ar@{.>}[dr]|{\fbar} & \\
     & & A  \\
@@
 Our definitions follow a simple ``pattern'':
+\todo{Pattern könnte besser erklärt werden aber nicht zu formal!}
 \begin{itemize}

### 2017-08-26 papers/talk-6/paper-6.tex
writeup6: Zweite Korrektur-Iteration

@@
   $$\fbar(a) := \pair{f_1(a)}{f_2(a)}.$$
-  By construction we have $p_1(\fbar(a)) = p_1(f_1(a), f_2(b)) = f_1(a)$ and $p_2(\fbar(a)) = p_2(f_1(a), f_2(a)) = f_2(a)$.
+  By construction we have $p_1(\fbar(a)) = p_1(f_1(a), f_2(a)) = f_1(a)$ and $p_2(\fbar(a)) = p_2(f_1(a), f_2(a)) = f_2(a)$.
   If we have another $\ftilde \from A \to X \times Y$ with $p_1(\ftilde(a)) = f_1(a)$ and $p_2(\ftilde(a)) = f_2(a)$ we show $\fbar(a) = \ftilde(a)$ for every $a \in A$:
@@
   Let $X, Y \in \cat{A}$ and $P$ be a product of $X$ and $Y$ together with $p_1 \from P \to X, p_2 \from P \to Y$
-  and $P'$ be a product of $X$ and $Y$ together with $p'_1 \from P' \to X, p'_2 \from P' \to Y$.
+  and $P'$ be another product of $X$ and $Y$ together with $p'_1 \from P' \to X, p'_2 \from P' \to Y$.
   We apply definition \ref{def:prod} multiple times and get the following commuting diagram:
@@
   A \demph{lower bound} for two objects $x, y \in A$ is a element $a \in A$ such that $a \le x$ and $a \le y$.
-  A \demph{greatest lover bound} of $x, y \in A$ is a lower bound $z \in A$ of $x$ and $y$ such that for all lower bounds $a \in A$ of $x$ and $y$, $a \le z$.
+  A \demph{greatest lower bound} of $x, y \in A$ is a lower bound $z \in A$ of $x$ and $y$ such that for all lower bounds $a \in A$ of $x$ and $y$, $a \le z$.
   With other words: $z$ is a greatest lower bound of $x$ and $y$ if and only if
@@
   Let $a \in \catA$ be another object.
-  If we have $f_1 \from a \to x$ and $f_2 \from a \to y$ then by definition of $\catA$, $a$ is a lower bound of $x$ and $y$.
-  Now because $z$ is a greatest lower bound, $a \le z$ holds and therefore is a unique map $\fbar \from a \to z$.
+  If we have $f_1 \from a \to x$ and $f_2 \from a \to y$ then by definition of $\catA$, $a \le x$ and $a \le y$ and thus $a$ is a lower bound of $x$ and $y$.
+  Now because $z$ is a greatest lower bound, $a \le z$ holds and therefore there is a unique map $\fbar \from a \to z$.
   The diagram commutes trivially.
@@
   \end{alignat*}
-  We can show that $U \oplus V$ is the product of $U$ and $V$ in $\Vect$.
+  We can show that $V \oplus W$ is the product of $V$ and $W$ in $\Vect$.
   The proof follows the same structure as in example \ref{ex:prod:cart}.
 \end{example}
+
+\todo{Soll ich die Verallgemeinerten Produkte über Klassen von Objekten auch einführen?}
 
@@
   \label{ex:coprod:vspace}
-  In the example \ref{ex:prod:vspace} we have seen that the direct sum of two vector spaces $U$ and $V$ is the product of $V$ and $W$ in $\Vect$.
+  In the example \ref{ex:prod:vspace} we have seen that the direct sum of two vector spaces $V$ and $W$ is the product of $V$ and $W$ in the category $\Vect$.
   We will now see that it is \emph{also} the coproduct.
@@
   We define the vector morphism $\fbar \from V \oplus W \to A$ with $\fbar(v, w) := f_1(v) + f_2(w)$.
-  By construction and because $f_1$ and $f_2$ are vector morphism we have:
+  By construction and because $f_1$ and $f_2$ are linear maps we have:
   \begin{alignat*}{4}
-    \fbar(p_1(u)) &= \fbar(u, 0) &= f_1(u) + f_2(0) &= f_1(u) \\
-    \fbar(p_2(v)) &= \fbar(0, v) &= f_1(0) + f_2(v) &= f_2(v).
+    \fbar(p_1(v)) &= \fbar(v, 0) &= f_1(v) + f_2(0) &= f_1(v) \\
+    \fbar(p_2(w)) &= \fbar(0, w) &= f_1(0) + f_2(w) &= f_2(w).
   \end{alignat*}
   Now we have to show that $\fbar$ is the only vector morphism with this property.
-  Let $\ftilde \from V \oplus W \to A$ be another vector morphism with $\ftilde(p_1(u)) = f_1(u)$ and $\ftilde(p_2(v)) = f_2(v)$.
+  Let $\ftilde \from V \oplus W \to A$ be another vector morphism with $\ftilde(p_1(v)) = f_1(v)$ and $\ftilde(p_2(w)) = f_2(w)$.
   Then we have for every $v \in V$ and $w \in W$:
@@
 
-\section*{Equalizers and Coequalizer}
+\section*{Equalizers and Coequalizers}
 
@@
   a diagram in $\cat{A}$.
-  A \demph{coequalizer} of this diagram is a commuting diagram called \demph{cofork} \\
+  A \demph{coequalizer} of this diagram is a commuting diagram called \demph{cofork}
   $ \xymatrix{
@@
   The \demph{symmetric transitive closure} of $R$ is $\sim := \bigcup_{n\in\nat} {\hat R}^n$
-  where ${\hat R}^n$ is defined recursively with
+  where ${\hat R}^n$ is recursively defined with
   ${\hat R}^0 := \setMap{\pair{a}{a}}{a \in X}$ and
-  ${\hat R}^{n+1} := R_n \of \hat R = \setMap{\pair{a}{c}}{\pair{a}{b} \in {\hat R}^n \land \pair{b}{c} \in \hat R}$.
+  ${\hat R}^{n+1} := {\hat R}^n \of \hat R = \setMap{\pair{a}{c}}{\pair{a}{b} \in {\hat R}^n \land \pair{b}{c} \in \hat R}$.
   It can be shown that $\sim$ is the \emph{least equivalence relation} containing $R$.
@@
 
-  To define the function $\fbar \from C \to A$ we need a choice function $\choice \cdot \from C \to Y$ that returns for every equivalence class in $\sim$ a member of it, i.~e.
+  To define the function $\fbar \from C \to A$ we need a choice function $\choice \cdot \from C \to Y$ that returns for every equivalence class of $\sim$ a member of it, i.~e.
   $\forall c \in C, \forall y \in Y, y \in c \Leftrightarrow y \sim \choice{c}$
@@
   a diagram in $\cat{A}$.
-  A \demph{pullback} of this diagram is a \demph{pullback square}
+  A \demph{pullback} of this diagram is a diagram called \demph{pullback square}
   \[ \xymatrix{
@@
 
-The pullback is also called ``fibred product''. This will be clear in the following example:
-\todo{Besser erklären, was das Beispiel damit zu tun hat}
-
-\begin{example}[Fibred product]
+\begin{example}[Pullbacks and products]
   When the pullback square
@@
 \begin{example}[Pullbacks in sets]
-  Let $X, Y$ be sets and $s \from X \to Z, t \from Y \to Z$ be functions.
+  Let $X, Y, Z$ be sets and $s \from X \to Z, t \from Y \to Z$ be functions.
   Then $P := \setMap{\pair{x}{y}}{s(x) = t(y)}$ with the projection functions
-  $p_1(x,y) := x$ and $p_2(x, y) = y$ is the pullback of
+  $p_1(x,y) := x$ and $p_2(x, y) := y$ is the pullback of
   $ \xymatrix{
@@
   $\fbar(a) := \pair{f_1(a)}{f_2(a)}$.
-  This is well defined because for $a \in A$ we have $\fbar(a) = \pair{f_1(a)}{f_2(a)} \in P \Leftrightarrow s(f_1(a) = t(f_2(a))$ which is true by assumption.
+  This is well defined because for $a \in A$ we have $\fbar(a) = \pair{f_1(a)}{f_2(a)} \in P \Leftrightarrow s(f_1(a)) = t(f_2(a))$ which is true by assumption.
   We also have $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ by constructions.
-  We also see that $\fbar$ is unique with this property.
+  We also see that $\fbar$ is unique in respect to that property.
 \end{example}
@@
   a diagram in $\cat{A}$.
-  A \demph{pushout} of this diagram is a \demph{pushout square} \\
+  A \demph{pushout} of this diagram is a commuting diagram called \demph{pushout square} \\
   \[ \xymatrix{
@@
 \section*{Conclusion}
-We have seen much analogous definitions and proofs.
+We have seen much analogous definitions.
 Our definitions follow a simple ``pattern'':
@@
 \begin{itemize}
-  \item First we have a some objects or a commuting diagram ($D_0$).
-  \item Then we ``insert'' a object $P$ and get a new commuting diagram ($D_1$) where we ``connect'' $P$ with the objects of $D_0$ with ``projection functions''.
+  \item First we have some objects or a commuting diagram ($D_0$).
+  \item Then we ``insert'' an object $P$ and get a new commuting diagram ($D_1$) where we ``connect'' $P$ with the objects of $D_0$ with ``projection functions''.
   \item Then we ``replace'' this object with a new object, called $A$, and replace the ``projection functions'' with new functions. We get a new diagram ($D_2$).

### 2017-08-27 papers/talk-6/paper-6.tex
writeup6: example pushouts in sets (and add missing \labels)

@@
 \begin{example}[Solution sets]
+  \label{ex:equalizer:solset}
   Let $X, Y$ be sets and $s, t \from X \to Y$ functions.
@@
 \begin{definition}[Symmetric reflexive transitive closure]
+  \label{def:srtc}
   Let $R$ be a binary relation over a set $X$.
@@
 \begin{example}[Quotient sets]
+  \label{ex:coequa:quotient}
   Let $X, Y$ be sets and $s, t \from X \to Y$ be functions.
@@
   Now we can define $\fbar \from C \to A$ by $\fbar(e) := f(\choice e)$.
-  We now have  $\fbar(i(y)) = f(\choice{\class y}) = f(y)$ because $y \sim \choice{\class y}$ (by the choice axiom and reflexivity of $\sim$).
+  We now have  $\fbar(i(y)) \explainRel{def}= f(\choice{\class y}) = f(y)$ because $y \sim \choice{\class y}$ (by the choice axiom and reflexivity of $\sim$).
   Now we only have to show that this $\fbar$ is unique in respect to that property.
   So we assume another function $\ftilde \from C \to A$ with $\ftilde \of i = f$. But then we have
-  $$ \fbar(c) \explainRel{Def}= f(\choice c) \explainRel{Ass.}= \ftilde(i(\choice c)) \explainRel{Def}= \ftilde(\class{\choice c}) \explainRel{Choice}= \ftilde(c).$$
+  $$ \fbar(c) \explainRel{def}= f(\choice c) \explainRel{ass}= \ftilde(i(\choice c)) \explainRel{def}= \ftilde(\class{\choice c}) \explainRel{choice}= \ftilde(c).$$
 \end{example}
@@
 \begin{example}[Pullbacks and products]
+  \label{ex:pullback:product}
   When the pullback square
@@
 \begin{example}[Pullbacks in sets]
+  \label{ex:pullback:set}
   Let $X, Y, Z$ be sets and $s \from X \to Z, t \from Y \to Z$ be functions.
@@
 
-\todo{Beispiele}
+\begin{example}[Pushouts in sets]
+  \label{ex:pushout:set}
+  To give an example for pushouts in sets we somehow have to combine the coproduct (disjoint union; see example \ref{ex:coprod:dunion}) and the coequalizer (quotient set; see example \ref{ex:coequa:quotient}).
+  Let
+  $ \xymatrix{
+    X & Z \ar[l]_s \ar[r]^t & Y
+  } $
+  be a functions and sets.
+  We want to define a set $P$ and two functions $f_1 \from X \to A, f_2 \from Y \to A$, such that the diagram
+  \[ \xymatrix{
+    Z \ar[r]_s \ar[d]^t & X \ar[d]_{p_1} \\
+    Y \ar[r]^{p_2} & P
+  } \]
+  commutes.
+  We define $P := (X \dunion Y) / \sim$ where $\sim \; \subseteq (X \dunion Y)^2$ is the symmetric reflexive transitive closure (see definition \ref{def:srtc}) over the relation
+  $R := \setMap{\pair{\pair{s(z)}{0}}{\pair{t(z)}{1}}}{z \in Z} \subseteq (X \dunion Y)^2$.
+  When we choose the projection functions
+  \begin{alignat*}{2}
+    p_1(x) &:= \class{\pair{x}{0}} \\
+    p_2(y) &:= \class{\pair{y}{1}}
+  \end{alignat*}
+  the diagram commutes by construction. Now let
+  \[ \xymatrix{
+    Z \ar[r]_s \ar[d]^t & X \ar[d]_{f_1} \\
+    Y \ar[r]^{f_2} & A
+  } \]
+  be another pushout square. To define the function $\fbar \from P \to A$ we first define a helper function $\fbar' \from X \dunion Y \to A$.
+  Therefore we use the choice function
+  $\choice \cdot \from P \to X \dunion Y$:
+  \begin{alignat*}{2}
+    \fbar'(x, 0) &:= f_1(x) \\
+    \fbar'(y, 1) &:= f_2(y) \\
+    \fbar(p)     &:= \fbar'(\choice p)
+  \end{alignat*}
+  By induction we can show that $\fbar'$ ``does not care'' about equal values, i.~e. $\forall a,b \in X \dunion Y, a \sim b \Rightarrow \fbar'(a) = \fbar'(b)$.
+
+  We show $f_1 = \fbar \of p_1$; $f_2 = \fbar \of p_2$ follows analogously.
+  $$\fbar(p_1(x)) \explainRel{def}= \fbar'(\choice {\class {\pair x 0}}) = \fbar'(x, 0) \explainRel{def}= f_1(x).$$
+  The second step follows by $\pair x 0 \sim \choice{\class {\pair x 0}}$ which holds by the choice axiom and reflexivity of $\sim$.
+
+  The last step is again to show that $\fbar$ is unique in respect to that property.
+So we assume another function $\ftilde \from X \dunion Y \to A$ with $f_1 = \ftilde \of p_1$ and $f_2 = \ftilde \of p_2$.
+  Let $p \in P$. We assume without loss of generality $\choice p = \pair x 0$ for some $x \in X$. Then
+  $$\fbar(p)
+    \explainRel{def}    = \fbar'(\choice p)
+    \explainRel{ass}    = \fbar'(x, 0)
+    \explainRel{def}    = f_1(x)
+    \explainRel{ass}    = \ftilde(p_1(x)) \explainRel{def}= \ftilde(\class{\pair x 0})
+    \explainRel{ass}    = \ftilde(\class{\choice p})
+    \explainRel{choice} = \ftilde(p).$$
+\end{example}
 
 \section*{Conclusion}
-We have seen much analogous definitions.
+We have seen much analogous definitions and proofs.
 Our definitions follow a simple ``pattern'':

### 2017-08-27 papers/talk-7/paper-7.tex
limit construction, colimit definition

@@
 \begin{document}
+
+%% TODO: consistent notation, DI D(I) D_I, also for maps D_m Dm
 
@@
 \end{minipage}\\[2em]
-Now we generalize them...
+These three examples follow a pattern.
+In each example there is an object $U$ with some arrows. This object $U$
+has a universal property which states that any other object $V$ which has arrows
+similar to $U$
+factors uniquely through $U$, that is, there is a unique morphism $V \to U$,
+such that the resulting triangles commute.\\
+Products, pullbacks, and equalizers are instances of a more general concept called
+limit which expresses the common pattern between these examples.
+%% TODO: possibly give further applications of limits too
+%% TODO: maybe some more intuition what limits are about
+
 \chapter{Definition}
-
 In category theory we take limits over diagrams.
@@
 The functor $D$ yields
-a part of the category $\mathscr C$ that essentially looks like $\bm I$.
-We also say $D$ \textbf{identifies} diagrams of shape $\bm I$,
-which is why we use $\bm I$ to denote the shape.
+a collection of objects and morphisms in the category $\mathscr C$
+that essentially looks like $\bm I$.
+We also say $D$ \textbf{identifies} diagrams of shape $\bm I$.
 Especially with diagrams of such a small shape
@@
   A \textbf{cone} $U$ over $D$ is an object $U \in \mathscr{C}$
-  with maps $f_I, \forall\ I \in \bm{I}$, such that
+  with maps $U \overset{f_I}{rightarrow} D_I, \forall\ I \in \bm{I}$, such that
   $\forall\ I \overset{m}{\to} J \in \bm{I}$, the following triangle commutes:
@@
 \begin{definition}
-  A cone $(U \overset{f_I}{\to} D(I))_{I\in \bm{I}}$ is universal, if
+  A cone $(U \overset{f_I}{\to} D(I))_{I\in \bm{I}}$ is \textbf{universal}, if
   for all other cones $(V \overset{g_I}{\to} D(I))_{I\in \bm{I}}$ there is a unique
@@
 In other words, any cone over $D$ has maps into $D$, and a cone $U$ is universal if
-for any other cone $V$, you can first go to $U$, that is the unique morphism.
-From $U$ you can go to $D_I$, and the commuting condition says that this is equal
-to the morphism that immediately takes you from $V$ to $D_I$.
+any other cone $V$ has a unique morphism to $U$.
+The commuting condition states that for any cone the projection maps $g_I$ are equivalent
+to the map that is obtained by composing the unique limit map $h$ with the
+projection maps of the limit $f_I$.
 
@@
   $g_I = f_I \circ h$, $g_J = f_J \circ h$, and $g_K = f_k \circ h$.
+  Therefore this diagram commutes in its entirety.
 \end{example}
@@
 \begin{definition}
-  A limit is a universal cone over a diagram.
+  A \textbf{limit} is a universal cone over a diagram.
 \end{definition}
@@
 
-%% TODO: terminal object, pullback, equalizer, product.
+\begin{example}
+  Consider a diagram of the following shape
+  \begin{tikzcd}
+    \bullet &\bullet
+  \end{tikzcd},
+  which is just two objects without non-trivial arrows.
+  A limit over this diagram looks like this.
+  \[
+    \begin{tikzcd}
+      & {\color{red} V}
+      \arrow[bend right=40,color=red, swap]{ddl}{g_I}
+      \arrow[bend left=40,color=red]{ddr}{g_J}
+      \arrow[dashed]{d}{h} &\\
+      & {\color{blue} U}
+      \arrow[color=blue, swap]{dl}{f_I}
+      \arrow[color=blue]{dr}{f_J} &\\
+      I && J
+    \end{tikzcd}
+  \]
+  This is the ordinary binary product.
+  The commuting condition for limits states that
+  $f_I \circ h = g_I$, and $f_J \circ h = g_J$, which
+  is exactly the commuting condition for products.
+  We have therefore verified that the product is indeed an instance of the general limit definition.
+  Using discrete categories with more than two objects we get products of arbitrary size.
+  \[
+    \begin{tikzcd}
+      && {\color{red} V}
+      \arrow[bend right=40,color=red]{ddll}
+      \arrow[bend right=40,color=red]{ddl}
+      \arrow[bend left=40,color=red]{ddrr}
+      \arrow[bend left=40,color=red]{ddr}
+      \arrow[dashed]{d} &&\\
+      && {\color{blue} U}
+      \arrow[color=blue]{dl}
+      \arrow[color=blue]{dll}
+      \arrow[color=blue]{dr}
+      \arrow[color=blue]{drr} &&\\
+      \bullet & \bullet & \ldots & \bullet &\bullet
+    \end{tikzcd}
+  \]  
+\end{example}
+
+\begin{example}
+  If we take the empty category as the shape for a diagram, then a cone
+  over that diagram is just a single object $U$ that has no other required arrows.
+  Then the universal property for limits states that every other cone, i.e. an object $V$,
+  has a unique map $V \to U$. A limit over the empty diagram is therefore just a terminal object,
+  and terminal objects are therefore another instance of the general limit definition.
+\end{example}
+
+\begin{example}
+  Consider the following diagram.
+  \begin{tikzcd}
+    I \arrow{r}{m} & J & K \arrow[swap]{l}{n}
+  \end{tikzcd}
+  A cone over this diagram looks like this:
+  \[
+    \begin{tikzcd}
+      U
+      \arrow[]{r}{\ \ \ f_K}
+      \arrow[swap]{rd}{\ f_J}
+      \arrow[swap]{d}{f_I}
+      &K \arrow{d}{n} \\
+      I \arrow[swap]{r}{m} & J
+    \end{tikzcd}
+  \]
+  The commuting condition for cones states however that
+  $f_J = m\circ f_I$, and $f_J = n\circ f_K$. Since $f_J$ is uniquely determined
+  by composition of other arrows, we do not need to draw it and leave it implicit.
+  With this in mind a limit over this diagram looks like this:
+  \[
+    \begin{tikzcd}
+      {\color{red}V}
+      \arrow[bend left=50,color=red]{drr}{g_K}
+      \arrow[bend right=50,color=red,swap]{ddr}{g_I}
+      \arrow[dashed,swap]{dr}{h}
+      && \\
+      &{\color{blue}U}
+      \arrow[color=blue]{r}{f_K}
+      \arrow[color=blue, swap]{d}{f_I}
+      &K \arrow{d}{n} \\
+      & I \arrow[swap]{r}{m} & J
+    \end{tikzcd}
+  \]
+  The commuting condition for limits results in exactly the known commuting condition for
+  pullbacks. We have therefore verified that pullbacks are a valid instance of the general limit definition.
+  It follows similarly that equalizers are limit instances over diagrams of the shape
+  \begin{tikzcd}
+    I \arrow[r,shift left, "m"] \arrow[r,shift right,swap,"n"] & J
+  \end{tikzcd}.
+\end{example}
 
 \chapter{Building new limits}
-It is possible to build new limits...
-%% TODO: example
-%% TODO: general limits
+It is easy to see that limits do not necessarily exists in all categories.
+Sometimes we want to show that a category has certain limits,
+but it might be difficult to show for this category that the limits
+of interest exist.
+We now study how some limits can be used to construct other limits.
+We start with an example.
+
+\begin{example}
+  Let $\mathscr C$ be a category with all products and all equalizers,
+  that is, for all diagrams of the respective shape, there exists the product
+  and the equalizer.
+  We can use these limits to construct arbitrary pullbacks.
+  We start with the diagram for pullbacks.
+  \[
+    \begin{tikzcd}
+      & K \arrow{d}{n} \\
+      I \arrow[swap]{r}{m} & J
+    \end{tikzcd}
+  \]
+  Now we take the product of $I$ and $K$. This product exists by assumption.
+  \[
+    \begin{tikzcd}
+      I \times K \arrow{r}{p_K} \arrow{d}{p_I} & K \arrow{d}{n} \\
+      I \arrow[swap]{r}{m} & J
+    \end{tikzcd}
+  \]
+  This looks already similar to a pullback.
+  However, we need to ensure the commuting condition.
+  For this purpose we use an equalizer which also exists by assumption.
+  \[
+    \begin{tikzcd}
+      E \arrow{r}{e} &
+      I \times K
+      \arrow[r,shift right,swap, "m \circ p_I"] \arrow[r,shift left, "n \circ p_K"]
+      & J \\
+    \end{tikzcd}
+  \]
+  We now unfold this drawing.
+  \[
+    \begin{tikzcd}
+      E \arrow[swap]{d}{p_I \circ e} \arrow{r}{p_K \circ e} & K \arrow{d}{n} \\
+      I \arrow[swap]{r}{m} & J
+    \end{tikzcd}
+  \]
+  $E$ satisfies the commuting and universality conditions of limits
+  since it is already an equalizer.
+  Therefore it is a pullback.
+\end{example}
+
+We now generalize
+the approach used in the example. It turns out that we can construct all
+limits with it.
+Let $D : \bm I \to \mathscr C$ be any diagram.
+We use the following construction:
+\[
+  \begin{tikzcd}
+    E \arrow{r}{e} &
+    \displaystyle \prod_{I \in \bm{I}} D(I)
+    \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
+    \displaystyle \prod_{K \overset{n}{\to} J \in \bm{I}} D(J)
+  \end{tikzcd}
+\]
+with
+\begin{align*}
+  s_n &= Dn \circ p_K\ \hat{=} \prod_{I \in \bm{I}} D(I)
+        \overset{p_K}{\rightarrow} D_K \overset{Dn}{\rightarrow} D_J \\
+  t_n &= p_J\ \hat{=} \prod_{I \in \bm{I}} D(I)
+        \overset{p_J}{\rightarrow} D_J
+\end{align*}
+The second product has components $p_n$ for each arrow
+$K \overset{n}{\to} J \in \bm{I}$ with a projection on
+the destination of the arrow.
+For example for the diagram of a pullback
+\[
+  \begin{tikzcd}
+    & K \arrow{d}{n} \\
+    I \arrow[swap]{r}{m} & J
+  \end{tikzcd}
+\]
+the mentioned product has projections $p_{1_I} = I$, $p_{1_J} = J$, and $p_{1_K} = K$
+for each identity arrow.
+Additionally there are projections for the remaining arrows $p_n = J$, and $p_m = J$.
+Since $E$ is an equalizer we get
+\begin{align*}
+  s \circ e &= t \circ e\\
+  s_n \circ e &= t_n \circ e, \quad \forall\ K \overset{n}{\rightarrow} J \in \bm I \\
+  Dn \circ p_K \circ e &= p_J \circ e \\
+  Dn \circ \uprho_K &= \uprho_J
+\end{align*}
+For all maps $K \overset{n}{\rightarrow} J$ we get the commuting diagram
+\[
+  \begin{tikzcd}
+    E \arrow{r}{\uprho_K} \arrow{dr}[swap]{\uprho_J} & D_K \arrow{d}{Dn}\\
+    & D_J
+  \end{tikzcd}
+\]
+This is the commuting condition for a cone.
+Since any other cone over $D$
+is also a fork of $s$ and $t$
+it follows that the equalizer $E$ is also a universal cone
+and therefore a limit over $D$.
+
+\begin{remark}
+  We can even construct arbitrary limits over infinite diagrams
+  because we assumed that $\mathscr C$ has all products.
+  If $\mathscr C$ only has finite products, then we can
+  use this construction for
+  limits over finite diagrams only which is still useful since
+  most practical limits are constructed using very small shapes for the diagram.\\
+  Finite products themselves can in turn be constructed if $\mathscr C$
+  has a terminal object and binary products. The terminal object is
+  only needed for unary products. All other finite products
+  can be constructed by cascading binary products.
+\end{remark}
+
 %% TODO: possibly other construction mentioned in awodey.
+%% pullbacks + term constructs fin products + eq, which is the other direction
+%% than the example.
 
 \chapter{Colimits}
-Colimits are the dual of limits.
+To define colimits we dualize the concept of limits.
+\begin{definition}
+  Let $D : \bm I \to \mathscr C$ be a diagram.
+  A \textbf{cocone} $U$ over $D$ is an object $U \in \mathscr{C}$
+  with maps $D_I \overset{ f_I}{\rightarrow} U,\ \forall\ I \in \bm{I}$, such that
+  $\forall\ I \overset{m}{\to} J \in \bm{I}$, the following triangle commutes:
+  \[
+    \begin{tikzcd}
+      D_I \arrow[rr, "D(m)"]\arrow[dr, swap, "f_I"] & & D_J \arrow{dl}{f_J}\\
+      & U  &
+    \end{tikzcd}
+  \]
+  That is $f_I = f_J \circ D(m)$.\\
+\end{definition}
+A cone has maps from the vertex into the diagram,
+while a cocone has maps from the diagram into the vertex.
+\begin{definition}
+  A cocone $(D(I)\overset{f_I}{\to} U)_{I\in \bm{I}}$ is \textbf{universal}, if
+  for all other cones $(\D(I) \overset{g_I}{\to} V)_{I\in \bm{I}}$ there is a unique
+  map $h \from U \to V$, such that for all $I \in \bm{I}$
+  the following triangles commute.
+  \[
+    \begin{tikzcd}
+      D_I \arrow[dr, swap, "g_I"] \arrow{r}{f_I} & U \arrow{d}{h} \\
+      & V
+    \end{tikzcd}
+  \]
+  That is $g_i = f_i \circ h$.
+\end{definition}
+
 \begin{definition}

### 2017-08-28 papers/talk-7/paper-7.tex
small proof clarifications

@@
 
-%% TODO: write a sentence like "In other words, ..." and give example. example might be tricky to draw
-
+In other words a cone is an object that covers the diagram with projection maps
+which commute with the whole diagram.\\
 Now we define what it means for a cone to be universal. Intuitively,
@@
 any other cone $V$ has a unique morphism to $U$.
-The commuting condition states that for any cone the projection maps $g_I$ are equivalent
+The commuting condition states that for any cone
+the projection maps $g_I$ are equivalent
 to the map that is obtained by composing the unique limit map $h$ with the
@@
   is exactly the commuting condition for products.
-  We have therefore verified that the product is indeed an instance of the general limit definition.
-  Using discrete categories with more than two objects we get products of arbitrary size.
+  We have therefore verified that the product is
+  indeed an instance of the general limit definition.
+  Using discrete categories with more than two objects
+  we get products of arbitrary size.
   \[
@@
 \end{align*}
-The second product has components $p_n$ for each arrow
+To clarify, the second product has components $p_n$ for each arrow
 $K \overset{n}{\to} J \in \bm{I}$ with a projection on
@@
 for each identity arrow.
-Additionally there are projections for the remaining arrows $p_n = J$, and $p_m = J$.
+Additionally there are projections for the remaining arrows $p_n = J$, and $p_m = J$.\\
 Since $E$ is an equalizer we get
@@
   s_n \circ e &= t_n \circ e, \quad \forall\ K \overset{n}{\rightarrow} J \in \bm I \\
+  Dn \circ p_K \circ e &= p_J \circ e
+\end{align*}
+The goal is to construct a limit, i.e. a universal cone.
+We take $E$ as the vertex and
+$\uprho_I \coloneqq p_I \circ e, \forall\ I \in \bm I$
+as the projection maps of the cone and we get
+for all maps $K \overset{n}{\rightarrow} J \in \bm I$
+\begin{align*}
   Dn \circ p_K \circ e &= p_J \circ e \\
@@
 \end{align*}
-For all maps $K \overset{n}{\rightarrow} J$ we get the commuting diagram
+which corresponds to the commuting diagram
 \[
@@
 \]
-This is the commuting condition for a cone.
+This is exactly the commuting condition we need for a cone.\\
 Since any other cone over $D$
@@
   A cocone $(D(I)\overset{f_I}{\to} U)_{I\in \bm{I}}$ is \textbf{universal}, if
-  for all other cones $(\D(I) \overset{g_I}{\to} V)_{I\in \bm{I}}$ there is a unique
+  for all other cones $(D(I) \overset{g_I}{\to} V)_{I\in \bm{I}}$ there is a unique
   map $h \from U \to V$, such that for all $I \in \bm{I}$

### 2017-08-28 papers/talk-7/paper-7.tex
added more limit constructions and restructured that section a bit. still some todo's left though

@@
 of interest exist.
-We now study how some limits can be used to construct other limits.
-We start with an example.
 
-\begin{example}
+\section{Construction of pullbacks using products and equalizers}
 Let $\mathscr C$ be a category with all products and all equalizers,
@@
 Therefore it is a pullback.
-\end{example}
 
-We now generalize
-the approach used in the example. It turns out that we can construct all
-limits with it.
+\section{Construction of finite products using pullbacks and terminal objects}
+
+Finite products themselves can in turn be constructed if $\mathscr C$
+has a terminal object and binary products. The terminal object is
+only needed for unary products. All other finite products
+can be constructed by cascading binary products.
+Binary products again can be constructed using pullbacks and a terminal object
+in the following way.
+Suppose $\mathscr C$ has pullbacks and a terminal object $1$. Consider the diagram
+\[
+  \begin{tikzcd}
+    & B \arrow{d} \\
+    A \arrow{r} & 1
+  \end{tikzcd}
+\]
+We have the pullback $U$, such that the following square commutes.
+\[
+  \begin{tikzcd}
+    U \arrow{r}{p_B} \arrow{d}{p_A} & B \arrow{d}{f} \\
+    A \arrow{r}{g} & 1
+  \end{tikzcd}
+\]
+It is easy to see that $U$ is the product of $A$ and $B$ by
+simply leaving out $f$ and $g$. In general, a pullback
+is like a binary product with some additional structure.
+
+\section{Construction of equalizers using pullbacks and finite products}
+
+Suppose we are again in a category $\mathscr C$ with pullbacks and a terminal objects.
+We can therefore construct products, as shown in. %%TODO: add reference
+We now show that we can also construct equalizers for any arrows $f, g \from A \to B$
+in the following way. Consider the diagram.
+\[
+  \begin{tikzcd}
+    & B \arrow{d}{\langle 1_B, 1_B \rangle} \\
+    A \arrow{r}{\langle f,g \rangle} & B \times B
+  \end{tikzcd}
+\]
+The product $B \times B$ exists by construction.
+We then have the pullback $E$, such that the following diagram commutes.
+\[
+  \begin{tikzcd}
+    E \arrow{r}{h} \arrow{d}{e} & B \arrow{d}{\langle 1_B, 1_B \rangle} \\
+    A \arrow{r}{\langle f,g \rangle} & B \times B
+  \end{tikzcd}
+\]
+By the commuting condition we get $f \circ e = h = g \circ e$, and $E$ is therefore
+indeed an equalizer.
+
+%% TODO: possibly construct terminal object from products and equalizers to get the nice equivalence
+
+\section{Construction of arbitrary limits}
+We now study how arbitrary limits can be constructed using only a
+few specific instances.
+We assume a category $\mathscr C$ that has all products and equalizers.
+The approach is similar to the construction of pullbacks in section.
+%% TODO: add reference
 Let $D : \bm I \to \mathscr C$ be any diagram.
@@
 and therefore a limit over $D$.
-
 \begin{remark}
-  We can even construct arbitrary limits over infinite diagrams
+  We can even construct limits over infinite diagrams
   because we assumed that $\mathscr C$ has all products.
@@
   use this construction for
-  limits over finite diagrams only which is still useful since
-  most practical limits are constructed using very small shapes for the diagram.\\
-  Finite products themselves can in turn be constructed if $\mathscr C$
-  has a terminal object and binary products. The terminal object is
-  only needed for unary products. All other finite products
-  can be constructed by cascading binary products.
+  limits over finite diagrams.
 \end{remark}
-
-%% TODO: possibly other construction mentioned in awodey.
-%% pullbacks + term constructs fin products + eq, which is the other direction
-%% than the example.
 
@@
   \]
-  That is $f_I = f_J \circ D(m)$.\\
+  That is $f_I = f_J \circ D(m)$.
 \end{definition}

### 2017-08-29 papers/talk-7/paper-7.tex
finished first draft

--- 2017-08-28 papers/talk-7/paper-7.tex
@@
 \maketitle
-\chapter{Introduction}
+\chapter*{Introduction}
 We have seen three special structures in categories.\\[1em]
@@
 Products, pullbacks, and equalizers are instances of a more general concept called
-limit which expresses the common pattern between these examples.
+limit which expresses the common pattern between these examples.\\
 %% TODO: possibly give further applications of limits too
@@
 
+{
+  %% Nice hack to get rid of page break just for the introduction chapter
+  \renewcommand{\clearpage}{}
   \chapter{Definition}
-In category theory we take limits over diagrams.
-We begin by formally defining what a diagram is.
+}
+In category theory we take limits of diagrams.
+Therefore we first make the notion of a diagram more precise.
 
@@
 We also say $D$ \textbf{identifies} diagrams of shape $\bm I$.
-Especially with diagrams of such a small shape
-there might be many parts in $\mathscr C$ that have the shape of $\bm I$.
+The shape category has no further meaning. Its only purpose
+is to define how the diagram is supposed to look like.
+The reason we use an intermediate
+category $\bm I$ to define the shape is that
+there might be many parts in $\mathscr C$ which have the shape $\bm I$,
+especially when $\bm I$ is so small.
 We therefore get all different diagrams of that shape by considering
 all different functors $D$.
-Diagrams do not need to be small though. Their shape can also be infinite.\\
+Diagrams do not necessarily have to be small. Their shape can even be infinite.
+However most interesting diagrams tend to be very small.\\
 We are typically interested in diagrams of a specific category,
@@
   A \textbf{cone} $U$ over $D$ is an object $U \in \mathscr{C}$
-  with maps $U \overset{f_I}{rightarrow} D_I, \forall\ I \in \bm{I}$, such that
+  with maps $U \overset{f_I}{\rightarrow} D_I, \forall\ I \in \bm{I}$, such that
   $\forall\ I \overset{m}{\to} J \in \bm{I}$, the following triangle commutes:
@@
 \begin{remark}
-  When $U$ is the vertex of a cone, then we sometimes use just $U$ to refer to the whole cone
-  for better readability.
+  When it is clear from the context that $U$ is the vertex of a cone, then we use just
+  the vertex $U$
+  to refer to the whole cone.
 \end{remark}
@@
 
-In other words a cone is an object that covers the diagram with projection maps
-which commute with the whole diagram.\\
+In other words a cone is an object that covers the diagram with projection maps,
+such that all triangles commute.\\
 Now we define what it means for a cone to be universal. Intuitively,
-a cone is universal if it satisfies a property that makes it universal among cones,
-specifically, that all other cones factor through it uniquely.
+a cone is universal if it
+is universally the closest to the diagram,
+and that all other cones factor through it uniquely.
 
@@
   $g_I = f_I \circ h$, $g_J = f_J \circ h$, and $g_K = f_k \circ h$.
-  Therefore this diagram commutes in its entirety.
+  Therefore this example commutes in its entirety.
 \end{example}
@@
 
-A limit is therefore characterized by the kind of diagram
+A limit of a diagram is therefore characterized by the kind of diagram
 over which it is a universal cone, meaning we get different instances of limits
@@
   which is just two objects without non-trivial arrows.
-  A limit over this diagram looks like this.
+  A limit of this diagram looks like this.
   \[
@@
 \begin{example}
+  \label{terminallimit}
   If we take the empty category as the shape for a diagram, then a cone
@@
   Then the universal property for limits states that every other cone, i.e. an object $V$,
-  has a unique map $V \to U$. A limit over the empty diagram is therefore just a terminal object,
+  has a unique map $V \to U$. A limit of the empty diagram is therefore just a terminal object,
   and terminal objects are therefore another instance of the general limit definition.
@@
   by composition of other arrows, we do not need to draw it and leave it implicit.
-  With this in mind a limit over this diagram looks like this:
+  With this in mind a limit of this diagram looks like this:
   \[
@@
 but it might be difficult to show for this category that the limits
-of interest exist.
-We now study how some limits can be used to construct other limits.
-We start with an example.
+of interest exist.\\
+In this chapter we show the following:\\
+\hyperref[equiinstance]{(\ref*{equiinstance})} A category $\mathscr C$ has finite products and equalizers iff
+it has pullbacks and a terminal
+object, that is,
+the limits on the right hand side can be constructed from the left hand side
+and vice versa.\\
+\hyperref[arbitrarylimits]{(\ref*{arbitrarylimits})}
+If a category $\mathscr C$ has products and equalizers, then it has all limits, that is,
+limits of an arbitrary diagram can be constructed using these two instances.
 
-\begin{example}
-  Let $\mathscr C$ be a category with all products and all equalizers,
-  that is, for all diagrams of the respective shape, there exists the product
-  and the equalizer.
-  We can use these limits to construct arbitrary pullbacks.
+\section{Equivalence of instances}
+\label{equiinstance}
+
+In this section we establish that the instances of limits we have seen are related
+in a way that allows us to show how to construct these instances using others.
+This implies that if we know that a category has some of the right instances
+we can construct other limits for free using only category theoretical arguments.
+
+\begin{fact}
+  \label{pullbackconstruction}
+  Let $\mathscr C$ be a category with finite products and equalizers.
+  Then we can construct pullbacks.
+\end{fact}
+\begin{proof}
 We start with the diagram for pullbacks.
@@
 \]
-  Now we take the product of $I$ and $K$. This product exists by assumption.
+We need to construct a limit of this diagram.
+First we the product of $I$ and $K$. This product exists by assumption.
 \[
@@
 Therefore it is a pullback.
-\end{example}
+\end{proof}
+\begin{fact}
+  \label{terminalproduct}
+  Let $\mathscr C$ be a category with finite products and equalizers.
+  Then we have a terminal object.
+\end{fact}
+\begin{proof}
+  Recall a terminal object is a limit of the empty diagram as mentioned in
+  \hyperref[terminallimit]{Example \ref*{terminallimit}}.
+  Since we have finite products we have in particular the empty product
+  which is nothing else than a limit of the empty diagram.
+\end{proof}
 
-We now generalize
-the approach used in the example. It turns out that we can construct all
-limits with it.
+\begin{fact}
+  \label{pullbackproduct}
+  Let $\mathscr C$ be a category with pullbacks and a terminal object $1$.
+  Then we can construct finite products.
+\end{fact}
+\begin{proof}
+Finite products can be constructed
+with the terminal object and binary products. For any object
+$A$, unary products are
+constructed using the binary product $A \times 1$.
+The empty product is just the terminal object $1$ as discussed in
+\hyperref[terminalproduct]{Fact \ref*{terminalproduct}}.
+All other finite products can be constructed by cascading binary products.
+Therefore we only show how to construct binary products.
+Consider following the diagram:
+\[
+  \begin{tikzcd}
+    & B \arrow{d} \\
+    A \arrow{r} & 1
+  \end{tikzcd}
+\]
+The pullback $U$ exists by assumption.
+\[
+  \begin{tikzcd}
+    U \arrow{r}{p_B} \arrow{d}{p_A} & B \arrow{d}{f} \\
+    A \arrow{r}{g} & 1
+  \end{tikzcd}
+\]
+We used the suggestive names $p_A$ and $p_B$ since it
+is easy to see that $U$ is already the product of $A$ and $B$ by
+simply leaving out $f$ and $g$. Their existence does not matter for the product.
+In general, a pullback
+is like a binary product with some additional structure.
+\end{proof}
+
+\begin{fact}
+  Let $\mathscr C$ be a category with pullbacks and a terminal object $1$.
+  Then we can construct equalizers.
+\end{fact}
+\begin{proof}
+This construction uses the fact that $\mathscr C$ has products
+as shown in \hyperref[pullbackproduct]{Fact \ref*{pullbackproduct}}.
+Consider the following diagram.
+\[
+  \begin{tikzcd}
+    & B \arrow{d}{\langle 1_B, 1_B \rangle} \\
+    A \arrow{r}{\langle f,g \rangle} & B \times B
+  \end{tikzcd}
+\]
+The product $B \times B$ exists by construction.
+We then have the pullback $E$, such that the following diagram commutes.
+\[
+  \begin{tikzcd}
+    E \arrow{r}{h} \arrow{d}{e} & B \arrow{d}{\langle 1_B, 1_B \rangle} \\
+    A \arrow{r}{\langle f,g \rangle} & B \times B
+  \end{tikzcd}
+\]
+Since $E$ is a pullback the square above commutes and
+we get $f \circ e = h = g \circ e$, and $E$ is therefore
+indeed an equalizer of $f$ and $g$.
+\end{proof}
+
+\section{Construction of arbitrary limits}
+\label{arbitrarylimits}
+We now study how arbitrary limits can be constructed using only a
+few specific instances.
+The approach is similar to the construction of pullbacks in
+\hyperref[pullbackconstruction]{Fact \ref*{pullbackconstruction}}.
+
+\begin{fact}
+  Let $\mathscr C$ be a category that has all products and equalizers.
   Let $D : \bm I \to \mathscr C$ be any diagram.
+  Then we can construct a limit of $D$.
+\end{fact}
+\begin{proof}
 We use the following construction:
@@
 \]
-with
+with the maps $s, t$ that have components $s_n, t_n$ for all $K \overset{n}{\to} J \in \bm{I}$:
 \begin{align*}
@@
 \end{align*}
-To clarify, the second product has components $p_n$ for each arrow
-$K \overset{n}{\to} J \in \bm{I}$ with a projection on
-the destination of the arrow.
-For example for the diagram of a pullback
-\[
-  \begin{tikzcd}
-    & K \arrow{d}{n} \\
-    I \arrow[swap]{r}{m} & J
-  \end{tikzcd}
-\]
-the mentioned product has projections $p_{1_I} = I$, $p_{1_J} = J$, and $p_{1_K} = K$
-for each identity arrow.
-Additionally there are projections for the remaining arrows $p_n = J$, and $p_m = J$.\\
-Since $E$ is an equalizer we get
+All products exist by assumption.
+To clarify, the product over all maps $K \overset{n}{\to} J \in \bm{I}$
+has components $p_n = J$ for each map $n$.\\
+We take $E$ as the equalizer of $s$ and $t$ which also exists by assumption.
+We therefore have $s \circ e = t \circ e$.
+The maps $s$ and $t$ have components. Therefore the equality has to hold for all
+components and we get
 \begin{align*}
@@
 \end{align*}
-The goal is to construct a limit, i.e. a universal cone.
+The goal is to construct a limit.
+We do this by constructing a cone and showing that it is universal.
+For the construction of a cone
+we need a vertex and projection maps into every object of the diagram,
+such that the commuting condition of cones is satisfied.
 We take $E$ as the vertex and
 $\uprho_I \coloneqq p_I \circ e, \forall\ I \in \bm I$
-as the projection maps of the cone and we get
-for all maps $K \overset{n}{\rightarrow} J \in \bm I$
+as the projection maps of the cone.
+Clearly, we have all necessary projection maps and we only need to verify
+the commuting condition of cones.
+We have all maps $K \overset{n}{\rightarrow} J \in \bm I$
 \begin{align*}
@@
 it follows that the equalizer $E$ is also a universal cone
-and therefore a limit over $D$.
-
+and therefore a limit of $D$.
+\end{proof}
 \begin{remark}
-  We can even construct arbitrary limits over infinite diagrams
+  We can even construct limits of infinite diagrams
   because we assumed that $\mathscr C$ has all products.
@@
   use this construction for
-  limits over finite diagrams only which is still useful since
-  most practical limits are constructed using very small shapes for the diagram.\\
-  Finite products themselves can in turn be constructed if $\mathscr C$
-  has a terminal object and binary products. The terminal object is
-  only needed for unary products. All other finite products
-  can be constructed by cascading binary products.
+  limits of finite diagrams.\\
+  Using \hyperref[equiinstance]{Section \ref*{equiinstance}}
+  we can alternatively require pullbacks and a terminal object instead of finite products.
 \end{remark}
-
-%% TODO: possibly other construction mentioned in awodey.
-%% pullbacks + term constructs fin products + eq, which is the other direction
-%% than the example.
 

### 2017-09-01 papers/talk-6/paper-6.tex
writeup6: Fix spacing in example ex:coprod:vspace
writeup6: third review; add somme prose and generalized (co)products

@@
 
-% { _ }
-\newcommand{\setCon}[1]{\bigl \{ #1 \bigr \}}
-% { _ | _ }
-\newcommand{\setMap}[2]{\setCon{#1 \,\big|\, #2}}
-% ( _ , _ )
-\newcommand{\pair}[2]{\left( #1 , #2 \right)}
-% [ _ ]
-\newcommand{\class}[1]{\left[ #1 \right]}
-% < _ >
-\newcommand{\choice}[1]{\left< #1 \right>}
-% explanation over relation sign
+\newcommand{\setOf}[1]{\bigl \{ #1 \bigr \}}
+\newcommand{\setMap}[2]{\setOf{#1 \,\big|\, #2}}
+\newcommand{\pair}[2]{\bigl( #1 , #2 \bigr)}
+\newcommand{\class}[1]{\bigl[ #1 \bigr]}
+\newcommand{\choice}[1]{\bigl< #1 \bigr>}
 \newcommand{\explainRel}[2]{\stackrel{\text{#1}}{#2}}
+\newcommand{\family}[2]{\bigl( #1 \bigr)_{#2}}
+
+\DeclareMathOperator{\core}{core}
 
@@
 \section*{Motivation}
-We want to generalize some concepts of sets like the Cartesian product, disjoint union, solution sets and quotient sets.
-In this chapter we define six constructions in categories.
-We will introduce the constructions \emph{products}, \emph{equalizer} and \emph{pullbacks} and their duals
-\emph{coproducts}, \emph{coequalizer} and \emph{pushouts}, respectively.
+
+In category theory, we want to unite similar structures from different areas of mathematics.
+For example we can unite the Cartesian product, the direct sum of vector spaces and greatest lower bounds in posets.
+We can also unite solution sets and the core in vector spaces.
+We will see how to generalize structures of sets like the Cartesian product, disjoint union, solution sets and quotient sets and how some of these structures correspond to structures in vector spaces.
+In this chapter we define six constructions in general categories.
+We will introduce the structures \emph{products}, \emph{equalizer} and \emph{pullbacks} and their duals \emph{coproducts}, \emph{coequalizer} and \emph{pushouts}.
 
@@
 
-\begin{example}[Cartesian Product]
+\begin{example}[Cartesian product]
   \label{ex:prod:cart}
   Let $X$ and $Y$ be sets.
-  Then we define the \demph{Cartesian Product}
+  Then we define the \demph{Cartesian product}
   $$X \times Y := \setMap{\pair{x}{y}}{x \in X, \; y \in Y}.$$
@@
 
+We see a simple instance of Lemma \ref{lem:prod:uniq} in the following example for the Cartesian product.
+
 \begin{example}[Reversed Cartesian product]
@@
   $$z \le x \land z \le y \land \forall a \in A, a \le x \land a \le y \Rightarrow a \le z.$$
-
   When we see the poset $\pair{A}{\le}$ as a category $\catA$ and have a greatest lower bound $z$ of $x, y \in A$, then $z$ is the product of $x$ and $y$.
@@
   \label{ex:prod:vspace}
-  Let $K$ be a field with the operations $+_K$ and $\cdot_K$ and $V, W$ vector spaces over this field with the operations $+_V, \cdot_V, +_W, \cdot_W$.
+  Let $K$ be a field and $V, W$ vector spaces over this field.
   Then we define\footnote{One should verify that the defined structure is indeed a vector field.}
@@
   \begin{alignat*}{4}
-    \pair{v_1}{w_1} + \pair{v_2}{w_2} &:= \pair{v_1 +_V v_2}{w_1 +_W w_2}           &\quad& \forall v_1, v_2 \in V, \forall w_1, w_2 \in W \\
-    \alpha \cdot \pair{v}{w}          &:= \pair{\alpha \cdot_V v}{\alpha \cdot_W w} &     & \forall v \in V, \forall w \in W
+    \pair{v_1}{w_1} + \pair{v_2}{w_2} &:= \pair{v_1 + v_2}{w_1 + w_2}           &\quad& \forall v_1, v_2 \in V, \forall w_1, w_2 \in W \\
+    \alpha \cdot \pair{v}{w}          &:= \pair{\alpha \cdot v}{\alpha \cdot w} &     & \forall \alpha \in K, \forall v \in V, \forall w \in W
   \end{alignat*}
-  We can show that $V \oplus W$ is the product of $V$ and $W$ in $\Vect$.
-  The proof follows the same structure as in example \ref{ex:prod:cart}.
+  We can show that $V \oplus W$ is the product of $V$ and $W$ in $\Vect_K$.
+  The proof follows the same structure as in Example \ref{ex:prod:cart}.
 \end{example}
 
-\todo{Soll ich die Verallgemeinerten Produkte über Klassen von Objekten auch einführen?}
+We have seen how for an arbitrary category the product of \emph{two} objects is defined.
+We can generalize this to products of families of objects.
+
+\begin{definition}[Arbitrary product]
+  \label{def:arb:prod}
+  Let $\catA$ be a category, $I$ a set (i.~e. of indices) and $\family{X_i}{i \in I}$ be a family of objects of $\catA$.
+  A product of $\family{X_i}{i \in I}$ consists of an object $P \in \catA$ and the projection maps $\family{P \toby{p_i} X_i}{i \in I}$, such that:
+  For all objects $A \in \catA$ and for all families of maps $\family{A \toby{f_i} X_i}{i \in I}$ there exists a unique map $\fbar \from A \to P$, such that $p_i \of \fbar = f_i$ holds for every index $i \in I$.
+\end{definition}
+
+\begin{example}[Arbitrary Cartesian product]
+  \label{ex:arb:prod:cart}
+  Let $I$ be a set of indices and $\family{X_i}{i \in I}$ be a family of objects.
+  An \demph{arbitrary Cartesian tuple} of $\family{X_i}{i \in I}$ consists of a family of elements $\pi = \family{\pi_i \in X_i}{i \in I}$.
+  The \demph{arbitrary Cartesian product} $\prod_{i \in I}{X_i}$ is then set set of all such tuples.
+  We define the projection functions $p_i(\pi) := \pi_i$ for $i \in I$.
+  Let $A$ be a set and $\family{A \toby{f_i} X_i}{i \in I}$ be a family of maps.
+  Then we define $\fbar(a) := \family{f_i(a)}{i \in I}$. $p_i \of \fbar = f_i$ holds by definition for $i \in I$.
+  When we have another function $\ftilde \from A \to \bigl(\prod_{i \in I}{X_i}\bigr)$ with $p_i \of \ftilde = f_i$ for $i \in I$, then $\fbar = \ftilde$, since for every $i \in I$:
+  $\fbar(a) \explainRel{def}= \family{f_i(a)}{i \in I} \explainRel{ass}= \family{p_i(\ftilde(a))}{i \in I} \explainRel{ext}= \ftilde(a)$.
+\end{example}
 
 Now we want to abstract the concept of the disjoint union of sets.
+We will see that the generalized structure is the dual of the product.
 
@@
   Then we define the \demph{disjoint union} of the sets $X$ and $Y$:
-  $$X \dunion Y := (X \times \setCon 0) \cup (Y \times \setCon 1) = \setMap{\pair{x}{0}}{x \in X} \cup \setMap{\pair{y}{1}}{y \in Y}.$$
-
+  $$X \dunion Y := (X \times \setOf 0) \cup (Y \times \setOf 1) = \setMap{\pair{x}{0}}{x \in X} \cup \setMap{\pair{y}{1}}{y \in Y}.$$
   Also we define the \demph{coprojection functions} $p_1 \from A \to X \dunion Y$, $p_2 \from Y \to X \dunion Y$:
@@
   \end{alignat*}
-
   Let $A$ be another set and $f_1 \from X \to A, f_2 \from Y \to A$ functions.
@@
   \label{ex:coprod:vspace}
-  In the example \ref{ex:prod:vspace} we have seen that the direct sum of two vector spaces $V$ and $W$ is the product of $V$ and $W$ in the category $\Vect$.
+  In Example \ref{ex:prod:vspace} we have seen that the direct sum of two vector spaces $V$ and $W$ is the product of $V$ and $W$ over a field $K$ in the category $\Vect_K$.
   We will now see that it is \emph{also} the coproduct.
-  In general products and coproducts are not the same as we have seen in the examples \ref{ex:prod:cart} and \ref{ex:coprod:dunion}.
+  In general products and coproducts are not the same as we have seen in Examples \ref{ex:prod:cart} and \ref{ex:coprod:dunion}.
   First we have to define the two coprojection functions $p_1 \from V \to V \oplus W, p_2 \from W \to V \oplus W$:
@@
   \end{alignat*}
-  Then let $A \in \Vect$ and $f_1 \from V \to A, f_2 \from W \to A$ be vector morphisms.
+  Then let $A \in \Vect_K$ and $f_1 \from V \to A, f_2 \from W \to A$ be vector morphisms.
   We define the vector morphism $\fbar \from V \oplus W \to A$ with $\fbar(v, w) := f_1(v) + f_2(w)$.
   By construction and because $f_1$ and $f_2$ are linear maps we have:
-  \begin{alignat*}{4}
-    \fbar(p_1(v)) &= \fbar(v, 0) &= f_1(v) + f_2(0) &= f_1(v) \\
-    \fbar(p_2(w)) &= \fbar(0, w) &= f_1(0) + f_2(w) &= f_2(w).
+  \begin{alignat*}{6}
+    & \fbar(p_1(v)) ~&\explainRel{def}=&~ \fbar(v, 0) ~&\explainRel{def}=&~ f_1(v) + f_2(0) ~&\explainRel{lin}=&~ f_1(v) \\
+    & \fbar(p_2(w)) ~&\explainRel{def}=&~ \fbar(0, w) ~&\explainRel{def}=&~ f_1(0) + f_2(w) ~&\explainRel{lin}=&~ f_2(w) .
   \end{alignat*}
@@
   Then we have for every $v \in V$ and $w \in W$:
-  $$ \fbar(v,w) = f_1(v) + f_2(w)\ = \ftilde(p_1(v)) + \ftilde(p_2(w)) = \ftilde(p_1(v) + p_2(w)) = \ftilde(v, w). $$
-  Thus $\fbar = \ftilde$ and $V \oplus W$ is indeed the product of $V$ and $W$ in the category $\Vect$.
+  $$\fbar(v,w) \explainRel{def}= f_1(v) + f_2(w)\ \explainRel{ass}= \ftilde(p_1(v)) + \ftilde(p_2(w)) \explainRel{lin}= \ftilde(p_1(v) + p_2(w)) \explainRel{def}= \ftilde(v, w).$$
+  Thus $\fbar = \ftilde$ and $V \oplus W$ is indeed the product of $V$ and $W$ in the category $\Vect_K$.
+\end{example}
+
+We can also generalize the coproduct as we have generalized the product in Definition \ref{def:arb:prod}.
+
+\begin{definition}[Arbitrary coproduct]
+  \label{def:arb:coprod}
+  Let $\catA$ be a category, $I$ a set (i.~e. of indices) and $\family{X_i}{i \in I}$ be a family of objects of $\catA$.
+  A product of $\family{X_i}{i \in I}$ consists of an object $C \in \catA$ and the coprojection maps $\family{X_i \toby{p_i} C}{i \in I}$, such that:
+  For all objects $A \in \catA$ and for all families of maps $\family{X_i \toby{f_i} A}{i \in I}$ there exists a unique map $\fbar \from P \to A$, such that $\fbar \of p_i = f_i$ holds for every index $i \in I$.
+\end{definition}
+
+In Example \ref{ex:arb:prod:cart} we have generalized the Cartesian product.
+No we will generalize Example \ref{ex:coprod:dunion} and define the disjoint union of a a family of sets.
+
+\begin{example}[Arbitrary disjoint union]
+  \label{ex:arb:coprod:dunion}
+  Let $\family{X_i}{i \in I}$ be a familiy of sets. We define the \demph{arbitrary disjoint union}
+  $$\bigdunion_{i \in I}{X_i} := \bigcup_{i \in I} \setMap{\pair{x}{i}}{x \in X_i}.$$
+  We define the coprojection functions $p_i \from X_i \to \bigl(\bigdunion_{i' \in I}{X_{i'}}\bigr)$ simply by $p_i(x) := \pair{x}{i}$ for every index $i \in I$.
+  Let $A$ be a set and $\family{X_i \toby{f_i} A}{i \in I}$ be a family of maps. We define $\fbar \from \bigl(\bigdunion_{i \in I}{X_i}\bigr) \to A$ by $\fbar(x, i) := f_i(x)$.
+  Now $\fbar \of p_i = f_i$ holds per definition for $i \in I$.
+  Let $\ftilde \from \bigl(\bigdunion_{i \in I}{X_i}\bigr) \to A$ be another function such that $\ftilde \of p_i = f_i$ holds for $i \in I$. Then
+  $\fbar(x, i) \explainRel{def}= f_i(x) \explainRel{ass}= \ftilde(p_i(a)) \explainRel{def}= \ftilde(x, i)$ for $\pair{x}{i} \in \bigl(\bigdunion_{i \in I}{X_i}\bigr)$ and thus $\fbar = \ftilde$.
+  Therefore $\bigdunion_{i \in I}{X_i}$ is the arbitrary disjoint union of $\family{X_i}{i \in I}$.
 \end{example}
@@
 
-%% XXX Weiterhin erst Beispiel und dann Definition oder erst Definition und dann Beispiel?
-Now we present more abstract concepts and give examples.
+Solutions sets and minus cores (as defined bellow) share the same structure.
+We will generalize that structure and it equalizer.
 
@@
 \begin{example}[Solution sets]
-  \label{ex:equalizer:solset}
+  \label{ex:equa:solset}
   Let $X, Y$ be sets and $s, t \from X \to Y$ functions.
@@
   Let $\ftilde \from A \to E$ be another function such that $i \of \ftilde = f$.
-  Then $\ftilde(a) = i (\ftilde(a)) = f(a) = i (\fbar(a)) = \fbar(a)$.
+  Then $\fbar(a) \explainRel{def}= f(a) \explainRel{ass}= i(\ftilde(a)) \explainRel{def}= \ftilde(a)$ holds for every $a \in A$, thus $\fbar = \ftilde$.
 \end{example}
 
-\todo{Mehr Beispiele dazu?}
+\begin{example}[Minus-core of linear maps between vector spaces]
+  \label{ex:equa:core}
+  Let $V$ and $W$ be vector spaces over a field $K$.
+  The \demph{core} of a linear map $h \from V \to W$ is defined by $$\core(h) := \setMap{v \in V}{h(v) = 0}.$$
+
+  Let $s, t \from V \to W$ be linear maps between $V$ and $W$.
+  The \demph{minus core} of $s$ and $t$ $$\core(s - t) \explainRel{def}= \setMap{v \in V}{s(v) = t(v)}$$ is the coequalizer of $s$ and $t$ in $\Vect_K$.
+  The proof follows the structure of Example \ref{ex:equa:solset}.
+\end{example}
+
+We define coequalizer as the dual of the equalizer. We will see that an example for this structure are quotient sets that are generated by a equivalence relation.
 
@@
   Let $R$ be a binary relation over a set $X$.
-  $\hat{R} := R \cup R^{-1}$ is the symmetric closure of $R$ where $R^{-1}$ is defined by
+  $\hat{R} := R \cup R^{-1}$ is the \demph{symmetric closure} of $R$ where $R^{-1}$ is defined by
   $R^{-1} := \setMap{\pair{b}{a}}{\pair{a}{b} \in R}$.
@@
   ${\hat R}^{n+1} := {\hat R}^n \of \hat R = \setMap{\pair{a}{c}}{\pair{a}{b} \in {\hat R}^n \land \pair{b}{c} \in \hat R}$.
-  It can be shown that $\sim$ is the \emph{least equivalence relation} containing $R$.
+  It can be shown that $\sim$ is the least equivalence relation containing $R$.
 \end{definition}
@@
   Therefore we define the relation
-  $R := \setMap{\pair{s(a)}{t(a)}}{a \in Y} \subseteq Y^2$.
+  $R := \setMap{\pair{s(a)}{t(a)}}{a \in X} \subseteq Y^2$.
   Let $\sim \; \subseteq Y^2$ be the symmetric reflexive transitive closure of $R$.
-  Now we define $C := Y / \sim = \setMap{\class y}{y \in Y}$ where $\class y := \setMap{y'}{y \sim y'}$ are the equivalence classes.
+  Now we define $C := Y / \sim ~:= \setMap{\class y}{y \in Y}$ where $\class y := \setMap{y'}{y \sim y'}$ are the equivalence classes.
   For the function $i \from Y \to E$ we choose $i(y) := \class y$.
@@
   be a cofork.
-  By induction we can show that $f$ ``does not care'' about equal values, i.~e. $f(y_1) = f(y_2)$ whenever $y_1 \sim y_2$.
+  By induction we can show that $f$ respects the relation $\sim$, i.~e. $f(y_1) = f(y_2)$ whenever $y_1 \sim y_2$.
 
@@
   We now have  $\fbar(i(y)) \explainRel{def}= f(\choice{\class y}) = f(y)$ because $y \sim \choice{\class y}$ (by the choice axiom and reflexivity of $\sim$).
-  Now we only have to show that this $\fbar$ is unique in respect to that property.
+  Now we only have to show that this $\fbar$ is unique with respect to that property.
   So we assume another function $\ftilde \from C \to A$ with $\ftilde \of i = f$. But then we have
@@
 
+% XXX Insert exercise 6.7 as further example?
+
 \section*{Pullbacks and Pushouts}
+
+In this section we will define our last two structures: pullbacks and their duals, pushouts.
 
@@
 
-\begin{example}[Pullbacks and products]
+\begin{lemma}[Pullbacks and products]
   \label{ex:pullback:product}
@@
   and $Z$ is a terminal object then $P$ is the product of $X$ and $Y$.
-\end{example}
+\end{lemma}
+\begin{proof}
+  We show that $P$ is the product of $X$ and $Y$ together with the projection functions $p_1$ and $p_2$.
+  Let
+  $ \xymatrix{
+    X &A \ar[l]_{f_1} \ar[r]^{f_2} & Y
+  } $
+  be a diagram in $\cat{A}$.
+  We have to show that there is a unique map $\fbar' \from P \to A$ such that
+  \[ \xymatrix{
+    & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\fbar} \ar[rdd]^{f_2} & \\
+    & P \ar[ld]^{p_1} \ar[rd]_{p_2} & \\
+    X & & Y
+  } \]
+  commutes. We know that the diagram
+  \[ \xymatrix{
+    A \ar[r]_{f_2} \ar[d]^{f_1} & Y \ar[d]_{t} \\
+    X \ar[r]^s & Z
+  } \]
+  commutes, since $s \of f_1$ and $t \of f_2$ are both arrows from $A$ to $Z$ and $Z$ is a terminal object.
+  Now because this square is then a pullback square there exists a unique map $\fbar \from A \to P$ such that
+  \[ \xymatrix{
+    A \ar[drr]^{f_2} \ar[ddr]_{f_1} \ar@{.>}[dr]|{\fbar} & & \\
+    & P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
+    & X \ar[r]^s & Z
+  } \]
+  commutes. We see that we can choose $\fbar' := \fbar$.
+  Suppose we have another $\tilde{f} \from P \to A$ such that
+  \[ \xymatrix{
+    & A \ar[ldd]_{f_1} \ar@{.>}[d]|{\ftilde} \ar[rdd]^{f_2} & \\
+    & P \ar[ld]^{p_1} \ar[rd]_{p_2} & \\
+    X & & Y
+  } \]
+  commutes. Then
+  \[ \xymatrix{
+    A \ar[drr]^{f_2} \ar[ddr]_{f_1} \ar@{.>}[dr]|{\ftilde} & & \\
+    & P \ar[r]_{p_2} \ar[d]^{p_1} & Y \ar[d]_{t} \\
+    & X \ar[r]^s & Z
+  } \]
+  would also commute (because $Z$ is terminal), but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
+\end{proof}
 
@@
   This is well defined because for $a \in A$ we have $\fbar(a) = \pair{f_1(a)}{f_2(a)} \in P \Leftrightarrow s(f_1(a)) = t(f_2(a))$ which is true by assumption.
-  We also have $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ by constructions.
-  We also see that $\fbar$ is unique in respect to that property.
+  We also have $p_1(\fbar(a)) = f_1(a)$ and $p_2(\fbar(a)) = f_2(a)$ by construction.
+  We also see that $\fbar$ is unique with respect to that property.
 \end{example}
+
+We define the dual of the pullbacks: pushouts.
 
@@
   \label{ex:pushout:set}
-  To give an example for pushouts in sets we somehow have to combine the coproduct (disjoint union; see example \ref{ex:coprod:dunion}) and the coequalizer (quotient set; see example \ref{ex:coequa:quotient}).
+  To give an example for pushouts in sets we somehow have to combine the coproduct (disjoint union; see Example \ref{ex:coprod:dunion}) and the coequalizer (quotient set; see Example \ref{ex:coequa:quotient}).
   Let
@@
   \end{alignat*}
-  By induction we can show that $\fbar'$ ``does not care'' about equal values, i.~e. $\forall a,b \in X \dunion Y, a \sim b \Rightarrow \fbar'(a) = \fbar'(b)$.
+  By induction we can show that $\fbar'$ respects the relatation $\sim$, i.~e. $\forall a,b \in X \dunion Y, a \sim b \Rightarrow \fbar'(a) = \fbar'(b)$.
 
@@
 
-  The last step is again to show that $\fbar$ is unique in respect to that property.
+  The last step is again to show that $\fbar$ is unique with respect to that property.
 So we assume another function $\ftilde \from X \dunion Y \to A$ with $f_1 = \ftilde \of p_1$ and $f_2 = \ftilde \of p_2$.
@@
 \section*{Conclusion}
-We have seen much analogous definitions and proofs.
+We have seen many analogous definitions and proofs.
 Our definitions follow a simple ``pattern'':
-\todo{Pattern könnte besser erklärt werden aber nicht zu formal!}
 \begin{itemize}

### 2017-09-07 papers/talk-7/paper-7.tex
part 1 of feedback iteration. switched to article class. added separate section on products. added fixes to problematic parts. still need to fix arbitrary construction and colimits

@@
 \maketitle
-\chapter*{Introduction}
-We have seen three special structures in categories.\\[1em]
+\section*{Introduction}
+We have seen three special structures in categories,
+the product, the pullback, and the equalizer.
+As an example, the binary product is constructed in the following way.
+We start with two objects. Then an object $U$ with maps to these objects is the product,
+if it exists an satisfies the universal property for products, that is, all other such objects $V$
+factor through $U$ uniquely.\\[1em]
+\begin{minipage}{.35\linewidth}
+  \begin{tikzcd}
+    \bullet &&\bullet
+  \end{tikzcd}
+\end{minipage}%
+\begin{minipage}{.35\linewidth}
+  \begin{tikzcd}
+    & U \arrow{dl} \arrow{dr} &\\
+    \bullet &&\bullet
+  \end{tikzcd}
+\end{minipage}%
+\begin{minipage}{.35\linewidth}
+  \begin{tikzcd}
+    & V \arrow[bend right=40]{ddl} \arrow[bend left=40]{ddr} \arrow[dashed]{d} &\\
+    & U \arrow{dl} \arrow{dr} &\\
+    \bullet &&\bullet
+  \end{tikzcd}
+\end{minipage}
+\\[1em]
+The pullback and the equalizer are constructed similarly.
+They all have an input diagram, and we call an object $U$ with maps into this diagram
+product/pullback/equalizer if it satisfies the universal property.
+\\[1em]
 \begin{minipage}{.35\linewidth}
@@
 \end{minipage}\\[2em]
-These three examples follow a pattern.
-In each example there is an object $U$ with some arrows. This object $U$
-has a universal property which states that any other object $V$ which has arrows
-similar to $U$
-factors uniquely through $U$, that is, there is a unique morphism $V \to U$,
-such that the resulting triangles commute.\\
-Products, pullbacks, and equalizers are instances of a more general concept called
-limit which expresses the common pattern between these examples.\\
 %% TODO: possibly give further applications of limits too
@@
 
-{
-  %% Nice hack to get rid of page break just for the introduction chapter
-  \renewcommand{\clearpage}{}
-  \chapter{Definition}
-}
-In category theory we take limits of diagrams.
-Therefore we first make the notion of a diagram more precise.
+\section{Definition}
+Products, pullbacks, and equalizers are instances of a more general concept called
+limit which expresses the common pattern between the examples above.
+The goal is to make this abstract common pattern more formal
+and establish relations between concrete instances of it.
+We start with the definition of a limit.\\
+
+\begin{definition}
+  A \textbf{limit} is a universal cone over a diagram.
+\end{definition}
+
+In the following section we explain this definition. We need to clarify three things
+in particular.
+First we need to define the concept of a diagram more formally.
+Then we define what a cone over a diagram is. Lastly we state what it means for a cone
+to be universal.
+We start with diagrams. They are the initial entities of which we take the limit.
+In the examples above the diagrams
+are the various collections of objects (and arrows) that we start with
+and over which the object $U$ has maps.
 
@@
   \label{diagramexample}
-  This is an example of a diagram of shape $\bm I$. \\[1cm]
+  This is an example of a diagram we use to build a pullback. \\[1cm]
   \begin{minipage}{0.3\linewidth}
@@
   \end{minipage}\\[1em]
+  We did not give names to the objects as we are not concerned with them.
 \end{example}
@@
 a collection of objects and morphisms in the category $\mathscr C$
-that essentially looks like $\bm I$.
+that looks just like $\bm I$.
 We also say $D$ \textbf{identifies} diagrams of shape $\bm I$.
@@
 $\mathscr C$ in the example. Therefore we usually leave out $\bm I$ and $D$,
-and we just draw the identified part of $\mathscr C$ instead.
-
+and we just draw the identified part of $\mathscr C$ instead.\\
+The next part on the way to the limit definition is the concept of a cone which
+is what we address now.
 \begin{definition}
@@
       & U \arrow[dl, swap, "f_I"] \arrow{dr}{f_J}&\\
-      D_I \arrow[rr, swap, "D(m)"] & & D_J\\
+      D_I \arrow[rr, swap, "D_m"] & & D_J\\
     \end{tikzcd}
@@
   the vertex $U$
-  to refer to the whole cone.
+  to refer to the whole cone.\\
+  Note that the underlying diagram determines how the cone looks like.
 \end{remark}
@@
 \begin{example}
-  A cone over the diagram of
-  \hyperref[diagramexample]{Example \ref*{diagramexample}} looks like this.
+  A cone over the diagram of pullbacks
+  (\hyperref[diagramexample]{Example \ref*{diagramexample}}) looks like this.
   \[
@@
 such that all triangles commute.\\
-Now we define what it means for a cone to be universal. Intuitively,
-a cone is universal if it
-is universally the closest to the diagram,
-and that all other cones factor through it uniquely.
+The last step we need for the
+general limit definition is
+to define what a universal cone is.
 
@@
 
-In other words, any cone over $D$ has maps into $D$, and a cone $U$ is universal if
-any other cone $V$ has a unique morphism to $U$.
-The commuting condition states that for any cone
-the projection maps $g_I$ are equivalent
-to the map that is obtained by composing the unique limit map $h$ with the
-projection maps of the limit $f_I$.
+Intuitively, a cone is universal if it
+is universally the ``closest'' to the diagram.
 
 \begin{example}
+  This is an example of a universal cone over the
+  diagram of pullbacks
+  (\hyperref[diagramexample]{Example \ref*{diagramexample}}).
     \[
@@
       \arrow[bend right=40,color=red]{dddrr}{g_I}
-      \arrow[bend left=40, color=red]{ddrrr}{g_k}
+      \arrow[bend left=40, color=red]{ddrrr}{g_K}
       \arrow[bend left=30, color=red, swap]{dddrrr}{g_J}
@@
       \arrow[bend right=30,color=blue]{ddr}{f_I}
-      \arrow[bend left=30, color=blue]{drr}{f_k}
+      \arrow[bend left=30, color=blue]{drr}{f_K}
       \arrow[bend left=20, color=blue, swap]{ddrr}{f_J}&   &\\
@@
   Additionally the commuting condition for universal cones states that
-  $g_I = f_I \circ h$, $g_J = f_J \circ h$, and $g_K = f_k \circ h$.
-  Therefore this example commutes in its entirety.
+  $g_I = f_I \circ h$, $g_J = f_J \circ h$, and $g_K = f_K \circ h$.
+  Therefore this example commutes in its entirety.\\
+  Note that there could be several other cones like $V$.
 \end{example}
 
-\begin{definition}
-  A \textbf{limit} is a universal cone over a diagram.
-\end{definition}
-
-A limit of a diagram is therefore characterized by the kind of diagram
+A limit is therefore characterized by the kind of diagram
 over which it is a universal cone, meaning we get different instances of limits
@@
   indeed an instance of the general limit definition.
-  Using discrete categories with more than two objects
-  we get products of arbitrary size.
-  \[
-    \begin{tikzcd}
-      && {\color{red} V}
-      \arrow[bend right=40,color=red]{ddll}
-      \arrow[bend right=40,color=red]{ddl}
-      \arrow[bend left=40,color=red]{ddrr}
-      \arrow[bend left=40,color=red]{ddr}
-      \arrow[dashed]{d} &&\\
-      && {\color{blue} U}
-      \arrow[color=blue]{dl}
-      \arrow[color=blue]{dll}
-      \arrow[color=blue]{dr}
-      \arrow[color=blue]{drr} &&\\
-      \bullet & \bullet & \ldots & \bullet &\bullet
-    \end{tikzcd}
-  \]  
 \end{example}
@@
   over that diagram is just a single object $U$ that has no other required arrows.
-  Then the universal property for limits states that every other cone, i.e. an object $V$,
-  has a unique map $V \to U$. A limit of the empty diagram is therefore just a terminal object,
-  and terminal objects are therefore another instance of the general limit definition.
+  The universal property for limits then states that every other cone, i.e. any object $V$,
+  has a unique map $V \to U$. A limit of the empty diagram is therefore just a terminal object.
 \end{example}
@@
   $f_J = m\circ f_I$, and $f_J = n\circ f_K$. Since $f_J$ is uniquely determined
-  by composition of other arrows, we do not need to draw it and leave it implicit.
+  by composition of other arrows, we do not need to draw it and leave it implicit here
+  and for all other cones over this diagram.
   With this in mind a limit of this diagram looks like this:
@@
   \]
-  The commuting condition for limits results in exactly the known commuting condition for
-  pullbacks. We have therefore verified that pullbacks are a valid instance of the general limit definition.
+  The commuting condition for $U$ being a limit is exactly the commuting condition for
+  $U$ being a pullback.
+  Thus pullbacks are a valid instance of the general limit definition.
   It follows similarly that equalizers are limit instances over diagrams of the shape
@@
     I \arrow[r,shift left, "m"] \arrow[r,shift right,swap,"n"] & J
-  \end{tikzcd}.
+  \end{tikzcd}. For equalizers we use the same approach we used above to ignore the morphism
+  $f_J$ to make the limit look like the conventional equalizer.
 \end{example}
 
-\chapter{Building new limits}
-It is easy to see that limits do not necessarily exists in all categories.
-Sometimes we want to show that a category has certain limits,
-but it might be difficult to show for this category that the limits
-of interest exist.\\
-In this chapter we show the following:\\
+\section{Building new limits}
+
+\begin{definition}
+We say a category \textbf{has limits} over diagrams of shape $\bm I$, if for all
+diagrams of shape $\bm I$, there exists a universal cone over it.
+\end{definition}
+A question that often arises is if a given category $\mathscr C$ has certain limits,
+e.g. products or a terminal object. Now that we have a abstract definition of limits
+we can approach this problem in a more systematic way.
+This leads to two observations:\\
 \hyperref[equiinstance]{(\ref*{equiinstance})} A category $\mathscr C$ has finite products and equalizers iff
 it has pullbacks and a terminal
-object, that is,
-the limits on the right hand side can be constructed from the left hand side
-and vice versa.\\
+object.\\
 \hyperref[arbitrarylimits]{(\ref*{arbitrarylimits})}
 If a category $\mathscr C$ has products and equalizers, then it has all limits, that is,
-limits of an arbitrary diagram can be constructed using these two instances.
+limits of diagrams of arbitrary shape can be constructed using these two instances.
 
-\section{Equivalence of instances}
+\subsection{Products}
+So far we have only been concerned with the product of two objects, i.e. the binary product.
+We now study the concept of products more generally.
+\begin{lemma}
+  \label{emptyproduct}
+  The empty product, i.e. the product of zero objects, is a terminal objects
+\end{lemma}
+\begin{proof}
+  The empty product is a limit over the empty diagram just as the terminal object
+  as discussed in
+  \hyperref[terminallimit]{Example \ref*{terminallimit}}
+\end{proof}
+
+Using discrete categories with more than two objects
+we get products of arbitrary size.
+\begin{example}
+  This is an example of a product of $k$ objects. We use the symbol $\prod$ for
+  arbitrary products.
+  \[
+    \begin{tikzcd}
+      && {\color{red} V}
+      \arrow[bend right=40,color=red]{ddll}
+      \arrow[bend right=40,color=red]{ddl}
+      \arrow[bend left=40,color=red]{ddrr}
+      \arrow[bend left=40,color=red]{ddr}
+      \arrow[dashed]{d} &&\\
+      && {\color{blue} {\displaystyle \prod_{i \in \{1,\ldots,k\} } D_i}}
+      \arrow[color=blue]{dl}
+      \arrow[color=blue]{dll}
+      \arrow[color=blue]{dr}
+      \arrow[color=blue]{drr} &&\\
+      D_1 & D_2 & \ldots & D_{k-1} & D_k
+    \end{tikzcd}
+  \]
+\end{example}
+
+\begin{lemma}
+  \label{finiteproducts}
+  Let $\mathscr C$ be a category with a terminal object $1$ and binary products.
+  Then we can construct all finite products.
+\end{lemma}
+\begin{proof}
+  \begin{itemize}
+  \item The empty product is just the terminal object by
+    \hyperref[emptyproduct]{Example \ref*{emptyproduct}}
+  \item The unary product of any object $A$ is constructed by the binary product
+    $A\times 1$ %% TODO: explain why this works...
+  \item The binary product exists by assumption
+  \item We want to build some finite product over objects $D_1, \dots D_k$.
+    Assume by induction that the following two products exist.
+    \[ P_1 \coloneqq \displaystyle \prod_{i \in \{1, \dots, m \} } D_i \text{ , and } P_2 \coloneqq \displaystyle \prod_{i \in \{m+1, \dots, k\} } D_i \]
+    with projection maps $(\pi_i)_{i \in \{1, \dots, m \} }$ for $P_1$, and $(\tau_i)_{i \in \{m+1, \dots, k\} }$ for $P_2$.
+    Then there exists a binary product of $P_1 \times P_2$ with projection maps $f_{P_1}, f_{P_2}$.
+    $P_1 \times P_2$ has maps $\pi_i \circ f_{P_1}$ and $\tau_i \circ f_{P_2}$.
+    $P_1 \times P_2$ is therefore a cone.
+    Since $P_1$ and $P_2$ are products, any other cone $V$ over them has unique maps $h_{P_1}$ and $h_{P_2}$.
+    The unique morphism $V \rightarrow P_1 \times P_2$ is given by $\langle h_{P_1}, h_{P_2} \rangle$.
+    $P_1 \times P_2$ is therefore universal.
+  \end{itemize}
+  Intuitively the construction works by stacking binary products on top of each other.
+  Since we assume $\mathscr C$ to have all products, the product of two products exists. This way we can create
+  arbitrarily large finite products.
+\end{proof}
+
+\subsection{Equivalence of instances}
 \label{equiinstance}
@@
 
-\begin{fact}
+\begin{theorem}
   \label{pullbackconstruction}
@@
   Then we can construct pullbacks.
-\end{fact}
+\end{theorem}
 \begin{proof}
@@
 We need to construct a limit of this diagram.
-First we the product of $I$ and $K$. This product exists by assumption.
+First we take the binary product of $I$ and $K$ which exists by assumption.
 \[
@@
   \begin{tikzcd}
+    E \arrow{dr}{e}& &\\
+    & I \times K \arrow[swap]{d}{p_I} \arrow{r}{p_K}& K \arrow{d}{n}\\
+    & I \arrow[swap]{r}{m} & J
+  \end{tikzcd}
+  \qquad \bm{\rightsquigarrow} \qquad 
+  \begin{tikzcd}
     E \arrow[swap]{d}{p_I \circ e} \arrow{r}{p_K \circ e} & K \arrow{d}{n} \\
@@
 since it is already an equalizer.
+%% TODO: more detail...
 Therefore it is a pullback.
-\end{proof}
-\begin{fact}
-  \label{terminalproduct}
-  Let $\mathscr C$ be a category with finite products and equalizers.
-  Then we have a terminal object.
-\end{fact}
-\begin{proof}
-  Recall a terminal object is a limit of the empty diagram as mentioned in
-  \hyperref[terminallimit]{Example \ref*{terminallimit}}.
-  Since we have finite products we have in particular the empty product
-  which is nothing else than a limit of the empty diagram.
 \end{proof}
@@
 Finite products can be constructed
-with the terminal object and binary products. For any object
-$A$, unary products are
-constructed using the binary product $A \times 1$.
-The empty product is just the terminal object $1$ as discussed in
-\hyperref[terminalproduct]{Fact \ref*{terminalproduct}}.
-All other finite products can be constructed by cascading binary products.
+with the terminal object and binary products by
+\hyperref[finiteproducts]{Lemma \ref*{finiteproducts}}.
 Therefore we only show how to construct binary products.
@@
 \]
-We used the suggestive names $p_A$ and $p_B$ since it
-is easy to see that $U$ is already the product of $A$ and $B$ by
-simply leaving out $f$ and $g$. Their existence does not matter for the product.
-In general, a pullback
-is like a binary product with some additional structure.
+The arrows $f$ and $g$ are unique and therefore monic.
+Hence $U$ satisfies the universality condition and is indeed the product.
+%% TODO: possibly show this for one calculation
 \end{proof}
 
-\begin{fact}
+Now that we know that a category with pullbacks and a terminal object
+also has finite products, we can proceed to construct equalizers as well
+as their construction build on the existence of products in this category.
+
+\begin{theorem}
   Let $\mathscr C$ be a category with pullbacks and a terminal object $1$.
   Then we can construct equalizers.
-\end{fact}
+\end{theorem}
 \begin{proof}
-This construction uses the fact that $\mathscr C$ has products
-as shown in \hyperref[pullbackproduct]{Fact \ref*{pullbackproduct}}.
-Consider the following diagram.
+  We start with the initial diagram for equalizers.
+  \[
+    \begin{tikzcd}
+      A \arrow[shift left]{r}{f} \arrow[shift right, swap]{r}{g} & B
+    \end{tikzcd}
+  \]
+  We transform this into the following diagram.
+  %%TODO: explain why they're equivalent
   \[
@@
   \]
-The product $B \times B$ exists by construction.
+  The product $B \times B$ exists by \hyperref[pullbackproduct]{Theorem \ref*{pullbackproduct}}.
   We then have the pullback $E$, such that the following diagram commutes.
@@
   Since $E$ is a pullback the square above commutes and
-we get $f \circ e = h = g \circ e$, and $E$ is therefore
+  we get $f \circ e = 1_B \circ h = h = 1_B \circ h = g \circ e$, and $E$ is therefore
   indeed an equalizer of $f$ and $g$.
@@
 
-\section{Construction of arbitrary limits}
+%%TODO: fix this section
+\subsection{Construction of arbitrary limits}
 \label{arbitrarylimits}
@@
 
-\chapter{Colimits}
+%%TODO: fix this section
+\section{Colimits}
 To define colimits we dualize the concept of limits.

### 2017-09-08 papers/talk-7/paper-7.tex
reworked construction of arbitrary limits

@@
   \item The unary product of any object $A$ is constructed by the binary product
-    $A\times 1$ %% TODO: explain why this works...
-  \item The binary product exists by assumption
+    $A \times A$.
+    We have $1_A \from A \to A$, and $\pi_1, \pi_2 \from A \times A \to A$.
+    By the universal property of products there exists a unique map $h \from A \to A \times A$.
+    We have $1_A = \pi_1 \circ h$.\\
+    On the other hand we have $(\pi_1 \circ h) \circ \pi_1 = 1_a \circ \pi_1 = \pi_1 \circ ( h \circ \pi_1)$,
+    and therefore $1_A = h \circ \pi_1$.
+    We have constructed an isomorphism and therefore $A \times A \iso A$.
+  \item The binary product exists by assumption for all objects $A, B$.
   \item We want to build some finite product over objects $D_1, \dots D_k$.
@@
 \]
-The arrows $f$ and $g$ are unique and therefore monic.
-Hence $U$ satisfies the universality condition and is indeed the product.
-%% TODO: possibly show this for one calculation
+Since $U$ is a pullback we have for any other cone $V$ with projections $q_A, q_B$
+a unique map $h \from V \to U$, such that
+$q_A = p_A \circ h$, and $q_B = p_B \circ h$, which is exactly the commuting condition
+for products.
 \end{proof}
@@
 \begin{theorem}
+  \label{pullbackequalizer}
   Let $\mathscr C$ be a category with pullbacks and a terminal object $1$.
@@
   We transform this into the following diagram.
-  %%TODO: explain why they're equivalent
   \[
@@
   The product $B \times B$ exists by \hyperref[pullbackproduct]{Theorem \ref*{pullbackproduct}}.
+  Furthermore we have $B \times B \iso B$ by \hyperref[finiteproducts]{Lemma \ref*{finiteproducts}}.
+  The above diagram is therefore equivalent.\\
   We then have the pullback $E$, such that the following diagram commutes.
@@
 
-\begin{fact}
-  Let $\mathscr C$ be a category that has all products and equalizers.
-  Let $D : \bm I \to \mathscr C$ be any diagram.
+\begin{theorem}
+  \label{arbitrarylimits}
+  Let $\mathscr C$ be a category that has all (finite) products and equalizers.
+  Let $D : \bm I \to \mathscr C$ be a (finite) diagram.
   Then we can construct a limit of $D$.
-\end{fact}
+\end{theorem}
 \begin{proof}
@@
 \end{align*}
-All products exist by assumption.
-To clarify, the product over all maps $K \overset{n}{\to} J \in \bm{I}$
-has components $p_n = J$ for each map $n$.\\
+
+The product $\displaystyle \prod_{K \overset{n}{\to} J \in \bm{I}} D(J)$ has components for each map in the diagram.
+We take the diagram for pullbacks as an example for clarification:
+\begin{tikzcd}
+ I \arrow{r}{m} & J & \arrow[swap]{l}{n} K
+\end{tikzcd}.\\
+For this example diagram the product has the following projections. One for each identity:
+$p_{1_I} \from \prod \to I,\ p_{1_J} \from \prod \to J,\ p_{1_K} \from \prod \to K$,
+and one for each remaining map in the diagram: $p_m \from \prod \to J,\ p_n \from \prod \to J$.\\
+All products in the above construction exist by assumption.
 We take $E$ as the equalizer of $s$ and $t$ which also exists by assumption.
 We therefore have $s \circ e = t \circ e$.
-The maps $s$ and $t$ have components. Therefore the equality has to hold for all
+The maps $s$ and $t$ have components. The equality has to hold for all
 components and we get
@@
 \begin{remark}
-  We can even construct limits of infinite diagrams
-  because we assumed that $\mathscr C$ has all products.
+  If $\mathscr C$ has all products including infinite products then
+  this construction works even for infinite diagrams.
   If $\mathscr C$ only has finite products, then we can
   use this construction for
-  limits of finite diagrams.\\
-  Using \hyperref[equiinstance]{Section \ref*{equiinstance}}
-  we can alternatively require pullbacks and a terminal object instead of finite products.
+  limits of finite diagrams.
 \end{remark}
+
+To summarize, we can construct arbitrary limits in the following way.
+If $\mathscr C$ has all products and equalizers then we can
+construct all limits by
+\hyperref[arbitrarylimits]{Theorem \ref*{arbitrarylimits}}.
+Equalizer can be constructed using finite products and pullbacks
+by \hyperref[arbitrarylimits]{Theorem \ref*{pullbackequalizer}}.
+Finite products can in turn be constructed from binary products and a terminal object
+by \hyperref[finiteproducts]{Lemma \ref*{finiteproducts}}.
+Binary products can be constructed using a pullback and a terminal object
+by \hyperref[arbitrarylimits]{Theorem \ref*{pullbackproduct}}.
+Having pullbacks and a terminal object is therefore the lowest requirement from
+which it is possible to bootstrap the constructions of other limit instances
+and even of arbitrary limits.
 

### 2017-09-08 papers/talk-7/paper-7.tex
reworked colimit section

@@
 factor through $U$ uniquely.\\[1em]
-\begin{minipage}{.35\linewidth}
+\begin{minipage}{.26\linewidth}
   \begin{tikzcd}
@@
 \end{minipage}%
-\begin{minipage}{.35\linewidth}
+\begin{minipage}{.1\linewidth}
+  $\rightsquigarrow \quad$
+\end{minipage}%
+\begin{minipage}{.3\linewidth}
   \begin{tikzcd}
@@
   \end{tikzcd}
+\end{minipage}%
+\begin{minipage}{.1\linewidth}
+  $\rightsquigarrow \quad$
 \end{minipage}%
@@
 and establish relations between concrete instances of it.
-We start with the definition of a limit.\\
-
+We start with the definition of a limit.
 \begin{definition}
@@
 First we need to define the concept of a diagram more formally.
-Then we define what a cone over a diagram is. Lastly we state what it means for a cone
+Then we define cones over a diagram. Lastly we state what it means for a cone
 to be universal.
@@
   \label{diagramexample}
-  This is an example of a diagram we use to build a pullback. \\[1cm]
+  This is an example of the diagram used for the pullback. \\[1cm]
   \begin{minipage}{0.3\linewidth}
@@
 \section{Colimits}
-To define colimits we dualize the concept of limits.
+ Colimits are the dual of limits.
+The motivation to dualize limits is to obtain dualized versions of theorems for limits.
+Many concepts in mathematics turn out to be colimits and it is
+therefore useful to have a general definition for them as well.
+\begin{definition}
+  A \textbf{colimit} is a universal cocone over a diagram.
+\end{definition}
+A cocone is the dualized version of a cone.
+Recall that a cone has maps from the vertex into the diagram. Dually,
+a cocone has maps from the diagram into the vertex.
 \begin{definition}
@@
 \end{definition}
-A cone has maps from the vertex into the diagram,
-while a cocone has maps from the diagram into the vertex.
+The universality condition is also dualized as expected.
+For limits the condition is the existence of a unique map from
+any other cone to the limit cone.
+For colimits the condition is therefore dually the existence of a unique map
+from the colimit to any other cocone.
 \begin{definition}
   A cocone $(D(I)\overset{f_I}{\to} U)_{I\in \bm{I}}$ is \textbf{universal}, if
-  for all other cones $(D(I) \overset{g_I}{\to} V)_{I\in \bm{I}}$ there is a unique
+  for all other cocones $(D(I) \overset{g_I}{\to} V)_{I\in \bm{I}}$ there is a unique
   map $h \from U \to V$, such that for all $I \in \bm{I}$
@@
 \end{definition}
-
-\begin{definition}
-  A \textbf{colimit} is a universal co-cone over a diagram.
-\end{definition}
-%% TODO: complete definition
-
+\begin{example}
 \[
@@
   \]
-%% TODO: example drawing
+\end{example}
+
+\begin{example}
+  We now verify that our three examples are indeed instances of the general colimit definition.
+  The diagram for the colimit is the dualized diagram of the limit.
+  For coproducts the diagram is therefore the same as for products.
+  We just construct a cocone instead of a cone.
+  \[
+    \begin{tikzcd}
+      \bullet \arrow{dr} \arrow[bend right=30]{ddr}&& \bullet \arrow{dl} \arrow[bend left= 30]{ddl}\\
+      & U \arrow[dashed]{d} & \\
+      & V &
+    \end{tikzcd}
+  \]
+  It is easy to see that the commuting condition for universal cocones is exactly the commuting condition
+  for coproducts.\\
+  Next we study the coequalizer.
+  We start with the diagram
+  \[
+    \begin{tikzcd}
+      X \arrow[shift left]{r}{f} \arrow[shift right, swap]{r}{g} & Y
+    \end{tikzcd}
+  \]
+  We construct a cocone over it
+  \[
+    \begin{tikzcd}
+      X \arrow[shift left]{r}{f} \arrow[shift right, swap]{r}{g} \arrow[swap]{d}{p_X} & Y \arrow[bend left=30]{dl}{p_Y}\\
+      U
+    \end{tikzcd}
+  \]
+  Because of the commuting condition for cocones we have
+  $p_1 = p_2 \circ f = p_2 \circ g$.
+  Since $p_1$ is uniquely determined by composition we leave the arrow implicit and do not draw it for all further cones.
+  If $U$ is a colimit, then we have
+  \[
+    \begin{tikzcd}
+      X \arrow[shift left]{r}{f} \arrow[shift right, swap]{r}{g} & Y \arrow{r}{p_Y} \arrow{dr}{q_Y}& U \arrow[dashed]{d}{h} \\
+      && V
+    \end{tikzcd}
+  \]
+  The commuting condition says that the diagram commutes.
+  Specifically we have $h \circ p \circ f= q \circ f$ and $h \circ p \circ g = q \circ g$ which is exactly
+  the commuting condition for coequalizers.\\
+  We have similarly that a pushout is a colimit of the diagram \begin{tikzcd} I & J \arrow{l}{f} \arrow{r}{g} & K \end{tikzcd}
+  by leaving out the projection map $p_J$.
+\end{example}
 

### 2017-09-18 papers/talk-11/paper-11.tex
Add first draft of writeup 11

\def\pathToRoot{../../}\input{\pathToRoot headers/header_lecturenotes}

\usepackage{todonotes}
\usepackage{semantic}

\DeclareMathOperator{\core}{core}

% Theorem Environments
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}

% Title page
\title{Presheaf Model for Dependent Types}

\author{Nikita Ziuzin}

\begin{document}

\maketitle

\section*{Dependent types}

Dependent types are types that unlike simple types like functions and products
may depend on or vary with values.  A classical example of a dependent type is
a type of lists $List_{A}(N)$ of a given length $N : \NN$ over some
type $A$. The elements of such type are $Nil_{A} : List(0)$ and
$Cons_{A}(x, l) : List(N+1)$, where $x : {A}$ and $l : List_{A}(N)$.
Now let us consider a function that takes $n : \NN$ and returns a list of
length $n$ containing $0$ in all entries.  This function would have the type
$\Pi n : \NN.List_\NN(n)$, a type of functions for which the type
of the return value depends on the argument.  The important point of the
dependent typing is that it reveals more information about the function.
Another example is an exception-free head function that returns the first
element of the list:
\[
  hd: \Pi x : \mathbb{N}.List_{A}(Suc(x)) \to A
\]
The type of this function prevents it from being applied to an empty list,
showing clearly that the function is defined only for non empty lists.

\section*{Formalization}

To formalize dependent types we build a system that allows us to make
judgements of the form $a : A$ (term $a$ has type $A$) and $A~type$ ($A$ is a
type). We will also want a notion of definitional equality to be built into the
system, for example we want to consider $0 : \NN$ and $0 + 0 : \NN$ to be
(definitionally) equal terms and $List_A(0)$, $List_A(0 + 0)$ as
(definitionally) equal types: if $l : List_A(0)$ then also $l : List_A(0 + 0)$.
This leads to two more forms of the judgements: $a = b : A$ ($a$ and $b$ are
definitionally equal terms of type $A$) and $A = B~type$ ($A$ and $B$ are
definitionally equal types). Finally, we must know types of the free variables
occurring in a judgement: we cannot assert $x + y : \NN$ unless we know $x :
\NN$ and $y : \NN$. Since such declarations may depend on each other like in $x
: \NN$, $y : List_\NN(x)$ it is convenient to make all judgements relative to a
list of variable declarations including at least the free variables occurring
inside the judgement. Such lists are called contexts and sometimes also type
assignments.

Intuitively, a context $x_1 : A_1, \dots, x_n : A_n$ is well-formed if each
$A_i$ is a type in the context $x_1 : A_1, \dots, x_{n-1} : A_{n-1}$ and all
the variables $x_i$ are pairwise distinct. Thus we arrive to six kinds of
judgements:

\begin{align*}
  &\vdash \Gamma~ctxt && \text{$\Gamma$ is a valid context} \\
  &\Gamma \vdash A~type && \text{$A$ is a type in $\Gamma$} \\
  &\Gamma \vdash a : A && \text{$a$ is a term of type $A$ in context $\Gamma$} \\
  &\vdash \Gamma = \Delta && \text{$\Gamma$ and $\Delta$ are definitionally equal contexts} \\
  &\Gamma \vdash A = B && \text{$A$ and $B$ are definitionally equal types in context $\Gamma$} \\
  &\Gamma \vdash a = b : A && \text{$a$ and $b$ are definitionally equal terms of type $A$ in context $\Gamma$}
\end{align*}

We also have the following structural rules:
\begin{itemize}
  \item The empty context construction, context extension and context equality
    rules
    \[
      \inference[C-Emp]{}{\vdash \diamond~ctxt}
      \qquad
      \inference[C-Ext]{\Gamma \vdash A~type}{\vdash \Gamma, x : A~ctxt}
    \]
    \[
      \inference[C-Ext-Eq]{\vdash \Gamma = \Delta~ctxt \qquad \Gamma \vdash A = B~type}{\vdash \Gamma, x:A = \Delta, y:B~ctxt}
    \]
    where $x$ and $y$ are assumed to be fresh variables.
  \item The variable rule
    \[
      \inference[Var]{\vdash \Gamma, x:A,\Delta~ctxt}{\Gamma,x:A,\Delta \vdash x:A}
    \]
  \item Rules expressing that definitional equality is an equivalence relation:
    \[
      \inference[C-Eq-R]{\vdash \Gamma~ctxt}{\vdash \Gamma = \Gamma~ctxt}
    \]
    \[
      \inference[C-Eq-S]{\vdash \Gamma = \Delta~ctxt}{\vdash \Delta = \Gamma~ctxt}
    \]
    \[
      \inference[C-Eq-T]{\vdash \Gamma = \Delta~ctxt \qquad \vdash \Delta = \Theta~ctxt}{\vdash \Gamma = \Theta~ctxt}
    \]
    with analogous rules for terms and types (omitted).
  \item Rules relating typing and definitional equality:
    \[
      \inference[Tm-Conv]{\Gamma \vdash a : A \qquad \vdash \Gamma = \Delta~ctxt \qquad \Gamma \vdash A = B~type}{\Delta \vdash a : B}
    \]
    \[
      \inference[Ty-Conv]{\vdash \Gamma = \Delta~ctxt \qquad \Gamma \vdash A~type}{\Delta \vdash A~type}
    \]
  \item And rules for weakening and substitution, where $\mathcal{J}$ ranges over judgements of the form $a : A, A~type, a = b: A, A = B~type$:
    \[
      \inference[Weak]{\Gamma, \Delta \vdash \mathcal{J} \qquad \Gamma \vdash C~type}{\Gamma, x : C, \Delta \vdash \mathcal{J}}
    \]
    \[
      \inference[Subst]{\Gamma, x: C, \Delta \vdash \mathcal{J} \qquad \Gamma \vdash c: C}{\Gamma, \Delta[c/x] \vdash \mathcal{J}[c/x]}
    \]
    Here $\mathcal{J}[c/x]$ (and $\Delta[c/x]$) denotes the capture-free
    substitution of $x$ for $c$ in $\mathcal{J}$. This means that bound
    variables in $\mathcal{J}$ are systematically renamed to prevent any free
    variables in $c$ from becoming bound in $\mathcal{J}[c/x]$. The contexts,
    types, and terms will be considered equal if they agree up to names of
    bound variables and we will assume existence of capture-free substitution
    function on these equivalence classes.
\end{itemize}

\section*{Type formers}

Type and term formers are introduced by formation, introduction, elimination,
and equality rules. We present several type formers to give an example of how
they may be constructed.

\subsection*{Dependent product}

Dependent product (also known as dependent function space or $\Pi$-type)
corresponds to set-theoretic notion of cartesian product over a family of sets
$\Pi_{i \in I}B_i$ which has as elements functions mapping an index $i$ to an
element of a corresponding set $B_i$. In type theory this is expressed as
follows:

\[
  \inference[$\Pi$-F]{\Gamma \vdash A~type \qquad \Gamma, x:A \vdash B~type}{\Gamma \vdash \Pi x:A. B~type}
\]

\[
  \inference[$\Pi$-Eq]{\Gamma \vdash A_1 = A_2 ~type \qquad \Gamma, x:A \vdash B_1 = B_2 ~type}{\Gamma \vdash \Pi x:A_1. B_1 = \Pi x:A_2. B_2~type}
\]

The first rule expresses that a dependent product consists of a type $A$ (that
possibly depends on other types in $\Gamma$) and a type depending on $A$ (and
$\Gamma$), namely $B$. The rule $\Pi$-Eq expresses that the definitional
equality is respected by the $\Pi$-former. Note that the variable $x$ becomes
bound in $\Pi x:A. B$.

To form the elements of the $\Pi$-type we have the introduction rules with
associated congruence rule:

\[
  \inference[$\Pi$-I]{\Gamma, x:A \vdash b : B}{\Gamma \vdash \lambda x : A. b : \Pi x:A. B}
\]

\[
  \inference[$\Pi$-I-Eq]{\Gamma, x:A_1 \vdash b_1 = b_2 : B_1 \qquad \Gamma A_1 = A_2~type \qquad \Gamma x: A_1 \vdash B_1 = B_2~type}{\Gamma \vdash \lambda x : A_1. b^{B_1} = \lambda x : A_2. b^{B_2} : \Pi x:A_1. B_1}
\]

So to give an element of $\Pi x:A. B$ one must give an element of $B[x]$ in the
presence of a variable $x$ of type $A$. The congruence rule $\Pi$-I-Eq
expresses that definitional equality preserves $\Pi$-introduction. In the
following we will refrain from writing down the congruence rules explicitly,
having them present implicitly for every type and term formers we introduce.

Elements of $\Pi$-type are consumed using application:
\[
  \inference[$\Pi$-E]{\Gamma \vdash u : \Pi x:A. B \qquad \Gamma \vdash a : A}
  {\Gamma \vdash App_{[x:A]B}(u, a): B[a / x]}
\]

Square brackets in the typing annotation $[x : A] B$ here mean that the
variable $x$ is bound in $B$.

The following rule gives definitional equality to applications --- applying a
function $\lambda x : A. b^B$ to a term $a : A$ results in $b$ with $x$
replaced by $a$:

\[
  \inference[$\Pi$-C]{\Gamma \vdash \lambda x : A. b : \Pi x:A. B \qquad \Gamma \vdash a : A}
  {\Gamma \vdash App_{[x:A]B}(\lambda x : A. b, a) = b[a / x]: B [a / x]}
\]

Note that the substitution plays more prominent role here than in simply typed
lambda calculus: substitution is needed to formulate not only the term
conversion rules, but also the typing rules.

\subsection*{Dependent sum}

The next type former we consider is the $\Sigma$-type, or dependent sum,
corresponding to disjoint union in set theory. If we are given a family of sets
$(B_i)_{i \in I}$ we can form the set $\Sigma_{i \in I} B_i := \{ (i, b)~|~i
\in I \land b \in B_i\}$ whose elements consist of an index $i$ and an element
of the corresponding set $B_i$. In type theory the corresponding formation and
introduction rules look as follows.
\[
  \inference[$\Sigma$-F]{\Gamma \vdash A \text{ type}\\ \Gamma, x:A \vdash B \text{ type}}{\Gamma \vdash \Sigma x : A. B \text{ type}}
\]
\[
  \inference[$\Sigma$-I]{\Gamma \vdash a : A\\ \Gamma\vdash b: B[a / x]}{\Gamma Pair_{[x:A]B} (a,b) : \Sigma x : A. B}
\]

Elimination rule:
\[
  \inference[$\Sigma$-E]{\Gamma, z : \Sigma x: A. B \vdash C \text{ type}\\ \Gamma, x: A, y: B \vdash c : C[Pair_{[x:A]B}(x, y)/z]\\ \Gamma \vdash u : \Sigma x: A. B}
  {R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, u): C[u/z]}
\]

And definitional equality:
\[
  \inference{\Gamma\vdash R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, Pair_{[x:A]B}(a,b)): C[Pair_{[x:A]B}/z]}
  {\Gamma\vdash R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, Pair_{[x:A]B}(a,b)) = c[a/x,b/y] : C [Pair_{[x:A]B}/z}
\]

The projection functions for the dependent sum can be defined using these
rules:
\[
  \mathbf{p}(a) := R^\Sigma_{[z : \Sigma x: A. B]A}([x : A, y: B]x, a) : A
\]
for the first projection and
\[
  \mathbf{q}(a) := R^\Sigma_{[z : \Sigma x: A. B]B[z.1/x]}([x : A, y: B]y, a) : B[\mathbf{p}(a)/x]
\]
for the second projection, assuming $\Gamma \vdash A~type$, $\Gamma, x:A \vdash
B~type$, and $\Gamma \vdash a : \Sigma x: A. B$.

The important special cases of $\Pi x: A. B$ and $\Sigma x : A. B$ arise when
$B$ does not actually depend on $A$. The we write
\[
  A \to B := \Pi x: A. B
\]
and
\[
  A \times B := \Sigma x: A. B
\]
indicating that in this cases $\Pi$- and $\Sigma$-types correspond to ordinary
non-dependent function and cartesian product types respectively.

\section*{Categories with Families}

We will use categories with families (CwF) to model dependent types. First
consider what is a category of families. The category of families of sets,
$\fcat{Fam}$, has as objects $(A, B)$ where $A$ is a set and $B = (B_a~|~a \in
A)$ is an $A$-indexed family of sets $B_a$, $a \in A$. A morphism $(A, B) \to
(A', B')$ is given by a pair $(f, g)$ where $f: A \to A'$ and $g$ is an
$A$-indexed famiy of functions such that $g_a: B_a \to B_{fa}'$.

A \emph{category with families} is given by ($\cat{C}$,$F$) where:

\begin{itemize}
  \item[$\bullet$] $\cat{C}$ is a category, objects: $\Gamma, \Delta, \dots$ are
    \emph{contexts}, morphisms:  $\sigma, \tau, \dots$ are
    \emph{substitutions} or context morphisms

  \item[$\bullet$] A terminal object $1$ in $\cat{C}$ is the empty context ($\diamond$).

  \item[$\bullet$] $F$ is a functor $F\from \cat{C}^\op \to \fcat{Fam}$; For a
    context $\Gamma$ we write $Ty_F(\Gamma)$ for the indexing set of the family
    $F(\Gamma)$ and its family as $(Ter_F(\Gamma;A) | A \in Ty_F(\Gamma))$;
    $\Gamma \vdash A~type$ means that $A \in Ty_F(\Gamma)$ and $\Gamma \vdash a
    : A$ means $a \in Ter_F(\Gamma;A)$. For a context morphism $\sigma \from
    \Delta \to \Gamma$ the morphism $F(\sigma)$ acts on types $\Gamma \vdash
    A~type$ as $A\sigma$ and on terms $\Gamma \vdash a : A$ as $a\sigma$. The
    following equations also hold:

    \[
      A1=A \quad (A\sigma)\tau = A(\sigma\tau) \quad a1=a \quad (a\sigma)\tau=a(\sigma\tau)
    \]

  \item[$\bullet$] There is also the operation of \emph{context extension}: if
    $\vdash \Gamma~ctxt$ and $\Gamma \vdash A~type$, there is a context
    $\Gamma, x:A$ (for brevity also written as $\Gamma.A$ from now on), a
    context morphism $\mathbf{p}: \Gamma.A \to \Gamma$, and a term $\Gamma.A
    \vdash \mathbf{q} : A \mathbf{p}$. This should satisfy the following
    property: for each $\vdash \Delta~ctxt$ and a context morphism
    $\sigma \from \Delta \to \Gamma$ and $\Delta \vdash a : A \sigma$, there is
    a context morphism $(\sigma, a): \Delta \to \Gamma.A$ such that:
    \[
      \mathbf{p}(\sigma, a) = \sigma \quad \mathbf{q}(\sigma, a) = a \quad (\sigma, a) \tau =
      (\sigma \tau, a \tau) \quad (\mathbf{p}, \mathbf{q}) = 1
    \]
    Where in the last equation $1: \Gamma.A \to \Gamma.A$
\end{itemize}

\section*{Presheaf model}

Let $\cat{C}$ be a small category. In the following objects of $\cat{C}$ will
be denoted as $I,J,K$ and morphisms as $f,g,h$.

The context $\Gamma$ is then a presheaf on $\cat{C}$, $\Gamma \from
\cat{C}^\op \to \Set$, that is, we are given a set $\Gamma(I)$ for each
$I$ in $\ob(\cat{C})$, and functions $\Gamma(I) \to \Gamma(J)$, $\rho \mapsto \rho
f$ (called \emph{restriction} and written as $f$ acting on the right) for each
$f: J \to I$ such that
\[
  \rho 1 = \rho \quad \text{and} \quad (\rho f) g = \rho (f g)
\]
for $g: K \to J$ and $f: J \to I$. We write $\Gamma_f$ for $\rho \mapsto \rho
f$.

The empty context then is a terminal presheaf and a context morphism $\sigma
\from \Delta \to \Gamma$ is a natural transformation making the diagram
commute.

\[
  \xymatrix{
    \Delta(I) \ar[r]^{\sigma_I} \ar[d]^{\Delta_f} & \Gamma(I) \ar[d]^{\Gamma_f} \\
    \Delta(J) \ar[r]^{\sigma_J} & \Gamma(J)
  }
\]

If we write $\Gamma_f$ and $\Delta_f$ as $f$ acting on the right, the
equation becomes:
\[
  (\sigma \rho) f = \sigma (\rho f)
\]
for $\rho$ in $\Delta(I)$.

Now we consider how to give a dependent type $\Gamma \vdash A~type$ in context
$\Gamma$. For each object $I \in \ob(\cat{C})$ and $\rho \in \Gamma(I)$ we require a
set $A \rho$, and for each $f \from J \to I$, we require a function $A \rho \to
A (\rho f)$, written as $a \mapsto af$ satisfying $a1 = a$ and $afg = a (f g)$
if $g \from K \to J$. The dependence on $I$ in $A\rho$ and on $I$ and $\rho$ in
$af$ is omitted for brevity. Substitution $\Delta \vdash A \sigma$ with $\sigma
\from \Delta \to \Gamma$ is simply given $(A \sigma) \rho = A (\sigma \rho)$
for $\rho \in \Gamma(I)$, together with the induced map
\[
  (A\sigma)\rho = A (\sigma \rho) \to A ((\sigma \rho)f) = A (\sigma (\rho f)) = (A \sigma) (\rho f)
\]
for $f \from J \to I$.

The definition of a dependent type can be rephrased more categorically: $A$ is
a presheaf on $\int_C \Gamma$, where $\int_C \Gamma$ is a category of elements
of the presheaf $\Gamma$, defined as follows:
\begin{itemize}
  \item objects are pairs $(I, \rho)$, where $I \in \ob(\cat{C})$ and $\rho \in
    \Gamma(I)$
  \item a morphism $(J, \rho') \to (I, \rho)$ is a morphism $f \from J \to I$
    in $\cat{C}$ such that $\rho' = \Gamma_f \rho$
\end{itemize}

A term $\Gamma \vdash a : A$ of a dependent type $\Gamma \vdash A~type$ is
given by a family of elements $a \rho \in A \rho$ for each $I \in \ob(\cat{C})$
and $\rho \in \Gamma(I)$, such that $a \rho f = a (\rho f)$ for each $f \from J
\to I$. The substitution $\Delta \vdash a \sigma : A \sigma$ with $\sigma \from
\Delta \to \Gamma$ is given by the family $(a \sigma) \rho = a (\sigma \rho)$.

For $\Gamma \vdash A~type$ the context extension $\vdash \Gamma.A~ctxt$ is
defined by
\begin{align*}
  &(\Gamma.A)(I) = \{(\rho, u)~|~\rho \in \Gamma(I) \land u \in A \rho\} &&
  \text{for $I \in \cat{C}$}\\
  &(\rho, u)f = (\rho f, u f) && \text{for $f \from J \to I$}
\end{align*}
and the projections are defined by $\mathbf{p} \from \Gamma.A \to \Gamma$,
$\mathbf{p}(\rho, u) = \rho$, and $\Gamma.A \vdash \mathbf{q} : Ap$ by
$\mathbf{q}(\rho, u) = u$.  Now assume $\sigma \from \Delta \to \Gamma$ and
$\Delta \vdash a : A\sigma$; we define $(\sigma, a) \from \Delta \to \Gamma.A$
by $(\sigma, a)\rho = (\sigma \rho, a \rho)$.

\subsection*{Dependent product}

As a motivation for dependent products recall how to define exponents in
presheaf categories. Let $\Gamma$ and $\Delta$ be presheaves and suppose we
already know how the exponent $\Delta^\Gamma$ is constructed. Then we get by
the Yoneda Lemma and the fact that $-^\Gamma$ should be right adjoint to $-
\times \Gamma$, that
\[
  (\Delta^\Gamma)(I) \cong Hom(\mathbf{y}I, \Delta^\Gamma) \cong Hom(\mathbf{y}I \times \Gamma, \Delta)
\]
where $\mathbf{y}$ is Yoneda embedding.

The latter can be taken up as a definition. So an element $w$ of
$(\Delta^\Gamma)(I)$ is a natural transformation $w \from \mathbf{y}I\times
\Gamma \to \Delta$, so it is given by functions
\[
  w_J \from (\mathbf{y}I)(J) \times \Gamma(J) \to \Delta(J),
\]
and the naturality condition becomes $(w_J(f, \rho))g = w_K(fg, \rho g)$ for $f
\from J \to I$, $\rho \in \Gamma(J)$, and $g \from K \to J$.

We will now show that CwF associated to a presheaf category supports
$\Pi$-types. Given $\Gamma \vdash A$ and $\Gamma.A \vdash B$ we have to define
$\Gamma \vdash \Pi x: A. B~type$, that is, we have to define the set $(\Pi x:
A.B) \rho$ for $I \in \cat{C}$ and $\rho \in \Gamma(I)$. The elements $w$ of
$(\Pi x:A.B)\rho$ are families $w = (w_f~|~J\in \cat{C}, f \from J \to I)$ of
(dependent) functions such that
\[
  w_f u \in B(\rho f, u) \qquad \text{for $J \in \cat{C}, f \from J \to I, and u \in A(\rho f)$},
\]
with the requirement that for $g \from K \to J$
\[
  (w_f u)g = w_{fg}(ug).
\]

For such a family $w \in (\Pi x: A. B)\rho$, the restriction $wf \in (\Pi
x:A.B)(\rho f)$ for $f: \from J \to I$ is defined by taking
\[
  (w f)_g u = w_{fg}u \in B(\rho f g, u)
\]
where $g \from K \to J$ and $u \in A(\rho f g)$. This definition satisfies $w 1
= w$ and $w f g = w(f g)$.

Now let $\Gamma.A \vdash b : B$; we have to define $\Gamma \vdash \lambda b :
\Pi x:A.B$, that is , give $(\lambda b)\rho \in (\Pi x:A.B)\rho$. For $f \from
J \to I$ and $u \in A(\rho f)$ we set
\[
  ((\lambda b)\rho)_f u = b (\rho f, u) \in B(\rho f, u)
\]
This satisfies for $g \from K \to J$
\[
  (((\lambda b)\rho)_f u)g = (b(\rho f, u))g = b(\rho(fg), ug) = ((\lambda b)\rho)_{fg}(ug)
\]
and defines a term since
\[
  (((\lambda b)\rho)f)_g u = ((\lambda b) \rho)_{fg} u = b(\rho fg, u) = ((\lambda b)(\rho f))_g u.
\]

To define the application let $\Gamma \vdash u : \Pi x:A.B$ and $\Gamma \vdash
v : A$, and we set $App(u, v)\rho = (u\rho)_1 (v\rho) \in B(\rho, v\rho)$, thus
$App(u, v)\rho \in B(1, v)\rho$ as $(\rho, v \rho) = (1, v)\rho$.
$\beta$-equality is readily checked
\[
  App(\lambda b, v)\rho = ((\lambda b)\rho)_1(v\rho)=b(\rho, v\rho)
\]
and similarly for $\eta$-equality and other equations.

\begin{example}[Ordinary function type]
  Let us consider a term of type $A \to A$, and the elements in $Ty(\Gamma)$
  and $Ter(\Gamma, A \to A)$ for some $\vdash \Gamma~ctxt$, $\Gamma \vdash
  A~type$.

  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we give the
  element of $\mathit{Ty}(\Gamma)$ corresponding to the type $A \to A$:
  \[
    (A \to A)\rho = \{w | w_f u \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
  \]
  The required equalities hold trivially for any $g \from K \to J$, $((\lambda
  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.

  $\Gamma, x : A \vdash x : A$ corresponds to $q$ in the model. Thus $\Gamma
  \vdash \lambda x.x : A -> A$ corresponds to $\lambda q$ being defined as
  follows:
  \[
    ((\lambda q) \rho)_f u = q(\rho f, u) = u
  \]
\end{example}

\subsection*{Dependent sum}
The interpretation of $\Gamma \vdash \Sigma x:A.B~type$ for $\Gamma \vdash
A~type$ and $\Gamma.A \vdash B~type$ is defined by
\[
  (\Sigma x :A. B)\rho = \{(a, b)~|~a \in A\rho \land b \in B(\rho, a)\}
\]
for $\rho \in \Gamma(I), I \in \ob(\cat{C})$. The restrictions are defined
componentwise $(a, b)f = (af, bf)$ for $f \from J \to I$. The pairing operation
$\Gamma \vdash (u, v) : \Sigma x:A.B$ of $\Gamma \vdash u : A$ and $\Gamma.A
\vdash v : B(1, u)$ is defined by the componentwise pairing: $(u, v)\rho =
(u\rho, v\rho)$. Likewise for the projections $\mathbf{p}$ and $\mathbf{q}$,
$(\mathbf{p}w)\rho = a$ and $(\mathbf{q}w)\rho = b$ for $w\rho = (a, b)$. This
validates the necessary equations.

\section*{Conclusion}

\end{document}

% vim: ts=2 ss=2 sw=2 expandtab

### 2017-09-18 papers/talk-7/paper-7.tex
Fixed all explanations and notations. Should be good to submit now

@@
 \begin{document}
-
-%% TODO: consistent notation, DI D(I) D_I, also for maps D_m Dm
 
@@
       & U \arrow[dl, swap, "f_I"] \arrow{dr}{f_J}&\\
-      D_I \arrow[rr, swap, "D_m"] & & D_J\\
+      D_I \arrow[rr, swap, "D_m"] & & D_J
     \end{tikzcd}
   \]
-  That is $f_J = D(m) \circ f_I$.\\
-  For a cone we write $(U \overset{f_I}{\to} D(I))_{I\in \bm{I}}$.
+  That is, $f_J = D_m \circ f_I$.\\
+  For a cone we write $(U \overset{f_I}{\to} D_I)_{I\in \bm{I}}$.
   We call $U$ the \textbf{vertex}, and $f_I$ the \textbf{projection maps} of the cone.
@@
   When it is clear from the context that $U$ is the vertex of a cone, then we use just
-  the vertex $U$
-  to refer to the whole cone.\\
-  Note that the underlying diagram determines how the cone looks like.
+  the vertex $U$ to refer to the whole cone.
 \end{remark}
-%% TODO: Notation DI, D(I), D_I ??
 \begin{example}
@@
     \begin{tikzcd}
-      U \arrow[bend right=30]{ddr}{f_I} \arrow[bend left=30]{drr}{f_k} \arrow[bend left=20, swap]{ddrr}{f_J}&   &\\
+      U \arrow[bend right=30]{ddr}{f_I} \arrow[bend left=30]{drr}{f_K} \arrow[bend left=20, swap]{ddrr}{f_J}&   &\\
       & & D_K \arrow{d}{D_n} \\
@@
 \begin{definition}
-  A cone $(U \overset{f_I}{\to} D(I))_{I\in \bm{I}}$ is \textbf{universal}, if
-  for all other cones $(V \overset{g_I}{\to} D(I))_{I\in \bm{I}}$ there is a unique
+  A cone $(U \overset{f_I}{\to} D_I)_{I\in \bm{I}}$ is \textbf{universal}, if
+  for all other cones $(V \overset{g_I}{\to} D_I)_{I\in \bm{I}}$ there is a unique
   map $h \from V \to U$, such that for all $I \in \bm{I}$
@@
   \]
-  That is $g_i = f_i \circ h$.
+  That is, $g_I = f_I \circ h$.
 \end{definition}
-
 Intuitively, a cone is universal if it
@@
   This is an example of a universal cone over the
-  diagram of pullbacks
-  (\hyperref[diagramexample]{Example \ref*{diagramexample}}).
+  diagram of pullbacks.
   \[
@@
 \begin{proof}
+  The idea is to first show the existence of the empty and the unary product.
+  For larger products we stack binary products on top of each other.
   \begin{itemize}
@@
     \hyperref[emptyproduct]{Example \ref*{emptyproduct}}
-  \item The unary product of any object $A$ is constructed by the binary product
-    $A \times A$.
-    We have $1_A \from A \to A$, and $\pi_1, \pi_2 \from A \times A \to A$.
-    By the universal property of products there exists a unique map $h \from A \to A \times A$.
-    We have $1_A = \pi_1 \circ h$.\\
-    On the other hand we have $(\pi_1 \circ h) \circ \pi_1 = 1_a \circ \pi_1 = \pi_1 \circ ( h \circ \pi_1)$,
-    and therefore $1_A = h \circ \pi_1$.
-    We have constructed an isomorphism and therefore $A \times A \iso A$.
+  \item The unary product of any object $A$ is constructed from the binary product
+    $A \times 1$.
+    We have maps $1_A \from A \to A$, $\pi_A \from A \times 1 \to A$, and $\pi_1 \from A \times 1 \to 1$.
+    Every other object $B$ has a unique map $t_B \from B \to 1$. If $B$ has a map $f \from B \to A$, then
+    it factors uniquely through $A \times 1$ by the universal property of products.
   \item The binary product exists by assumption for all objects $A, B$.
-  \item We want to build some finite product over objects $D_1, \dots D_k$.
-    Assume by induction that the following two products exist.
-    \[ P_1 \coloneqq \displaystyle \prod_{i \in \{1, \dots, m \} } D_i \text{ , and } P_2 \coloneqq \displaystyle \prod_{i \in \{m+1, \dots, k\} } D_i \]
-    with projection maps $(\pi_i)_{i \in \{1, \dots, m \} }$ for $P_1$, and $(\tau_i)_{i \in \{m+1, \dots, k\} }$ for $P_2$.
+  \item We build some finite product over objects $D_1, \dots D_k$.
+    Assume by induction that the following two smaller products exist.
+    \[ P_1 \coloneqq \displaystyle \prod_{i \in \{1, \dots, \frac{k}{2} \} } D_i \text{ , and } P_2 \coloneqq \displaystyle \prod_{i \in \{\frac{k}{2} + 1, \dots, k\} } D_i \]
+    with projection maps $(\pi_i)_{i \in \{1, \dots, \frac{k}{2} \} }$ for $P_1$, and $(\tau_i)_{i \in \{\frac{k}{2}+1, \dots, k\} }$ for $P_2$.
     Then there exists a binary product of $P_1 \times P_2$ with projection maps $f_{P_1}, f_{P_2}$.
@@
   \end{itemize}
-  Intuitively the construction works by stacking binary products on top of each other.
-  Since we assume $\mathscr C$ to have all products, the product of two products exists. This way we can create
-  arbitrarily large finite products.
 \end{proof}
@@
 \]
-$E$ satisfies the commuting and universality conditions of limits
-since it is already an equalizer.
-%% TODO: more detail...
-Therefore it is a pullback.
+Since $E$ is an equalizer we have $e \circ p_K \circ n = e \circ p_I \circ m$.
+Therefore, $E$ satisfies the commuting condition for cones over a pullback square.
+Furthermore note that any other cone $V$ over this pullback square has maps $f \from V \to I$,
+$g \from V \to K$.
+By the universal property of products there is a unique map $m \from V \to I \times K$.
+Since $E$ is an equalizer, there is a unique map $h \from V \to E$,
+and $E$ is indeed a universal cone, that is, a pullback.
 \end{proof}
@@
 \]
-Since $U$ is a pullback we have for any other cone $V$ with projections $q_A, q_B$
-a unique map $h \from V \to U$, such that
-$q_A = p_A \circ h$, and $q_B = p_B \circ h$, which is exactly the commuting condition
-for products.
+Every object has a unique map to $1$ because it is terminal.
+The cones over $A$ and $B$ are therefore exactly the cones
+over $A$, $B$, and $1$.
+Since $U$ is a pullback, it is therefore also the product.
 \end{proof}
@@
 Now that we know that a category with pullbacks and a terminal object
-also has finite products, we can proceed to construct equalizers as well
-as their construction build on the existence of products in this category.
+also has finite products, we can proceed to construct equalizers as well.
+Their construction builds on the existence of products in this category.
 
@@
   \]
-  We transform this into the following diagram.
+  To construct the equalizer of $f$ and $g$ we take the pullback of the following diagram
   \[
@@
   The product $B \times B$ exists by \hyperref[pullbackproduct]{Theorem \ref*{pullbackproduct}}.
-  Furthermore we have $B \times B \iso B$ by \hyperref[finiteproducts]{Lemma \ref*{finiteproducts}}.
-  The above diagram is therefore equivalent.\\
-  We then have the pullback $E$, such that the following diagram commutes.
+  We then have the pullback $E$, such that the following square commutes.
   \[
@@
   \]
-  Since $E$ is a pullback the square above commutes and
-  we get $f \circ e = 1_B \circ h = h = 1_B \circ h = g \circ e$, and $E$ is therefore
-  indeed an equalizer of $f$ and $g$.
+  The product $B \times B$ has projection maps $\pi_1, \pi_2$.
+  By the commuting condition we get
+  \begin{align*}
+    & f \circ e \\
+    =\ & \pi_1 \circ \langle f, g \rangle \circ e\\
+    =\ & \pi_1 \circ \langle 1_B, 1_B \rangle \circ h\\
+    =\ & \pi_2 \circ \langle 1_B, 1_B \rangle \circ h\\
+    =\ & \pi_2 \circ \langle f,g \rangle \circ e\\
+    =\ & g \circ e
+  \end{align*}
+  Therefore, $E$ satisfies the commuting condition for equalizers.
+  Universality follows, since any other fork of $f$ and $g$ is also a pullback
+  cone over the above pullback square.
 \end{proof}
@@
     E \arrow{r}{e} &
-    \displaystyle \prod_{I \in \bm{I}} D(I)
+    \displaystyle \prod_{I \in \bm{I}} D_I
     \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
-    \displaystyle \prod_{K \overset{n}{\to} J \in \bm{I}} D(J)
+    \displaystyle \prod_{K \overset{n}{\to} J \in \bm{I}} D_J
   \end{tikzcd}
@@
 \begin{align*}
-  s_n &= Dn \circ p_K\ \hat{=} \prod_{I \in \bm{I}} D(I)
+  s_n &= Dn \circ p_K\ \hat{=} \prod_{I \in \bm{I}} D_I
         \overset{p_K}{\rightarrow} D_K \overset{Dn}{\rightarrow} D_J \\
-  t_n &= p_J\ \hat{=} \prod_{I \in \bm{I}} D(I)
+  t_n &= p_J\ \hat{=} \prod_{I \in \bm{I}} D_I
         \overset{p_J}{\rightarrow} D_J
@@
 
-The product $\displaystyle \prod_{K \overset{n}{\to} J \in \bm{I}} D(J)$ has components for each map in the diagram.
-We take the diagram for pullbacks as an example for clarification:
+The product $\displaystyle \prod_{K \overset{n}{\to} J \in \bm{I}} D_J$ has components for each map in the diagram.
+We take the diagram for pullbacks as an example:
 \begin{tikzcd}
  I \arrow{r}{m} & J & \arrow[swap]{l}{n} K
-\end{tikzcd}.\\
+\end{tikzcd}.
 For this example diagram the product has the following projections. One for each identity:
@@
 \end{align*}
-The goal is to construct a limit.
-We do this by constructing a cone and showing that it is universal.
-For the construction of a cone
-we need a vertex and projection maps into every object of the diagram,
-such that the commuting condition of cones is satisfied.
+With this setup we show that we have a universal cone over $D$.
 We take $E$ as the vertex and
@@
 is also a fork of $s$ and $t$
-it follows that the equalizer $E$ is also a universal cone
-and therefore a limit of $D$.
+it follows that the equalizer $E$ is also a universal cone,
+that is, a limit of $D$.
 \end{proof}
@@
 \hyperref[arbitrarylimits]{Theorem \ref*{arbitrarylimits}}.
-Equalizer can be constructed using finite products and pullbacks
+Equalizers can be constructed using finite products and pullbacks
 by \hyperref[arbitrarylimits]{Theorem \ref*{pullbackequalizer}}.
-Finite products can in turn be constructed from binary products and a terminal object
+In the case that we want to construct infinite limits
+the existence of infinite products is necessary
+for which we have no construction.
+If we want to construct finite limits, then finite products are sufficient.
+Finite products can be constructed from binary products and a terminal object
 by \hyperref[finiteproducts]{Lemma \ref*{finiteproducts}}.
-Binary products can be constructed using a pullback and a terminal object
+Binary products can in turn be constructed from a pullback and a terminal object
 by \hyperref[arbitrarylimits]{Theorem \ref*{pullbackproduct}}.
-Having pullbacks and a terminal object is therefore the lowest requirement from
-which it is possible to bootstrap the constructions of other limit instances
-and even of arbitrary limits.
+Having pullbacks and a terminal object is therefore the most basic requirement from
+which it is possible to bootstrap the construction of arbitrary finite limits.
 
-%%TODO: fix this section
 \section{Colimits}
@@
     \begin{tikzcd}
-      D_I \arrow[rr, "D(m)"]\arrow[dr, swap, "f_I"] & & D_J \arrow{dl}{f_J}\\
+      D_I \arrow[rr, "D_m"]\arrow[dr, swap, "f_I"] & & D_J \arrow{dl}{f_J}\\
       & U  &
@@
   \]
-  That is $f_I = f_J \circ D(m)$.
+  That is $f_I = f_J \circ D_m$.
 \end{definition}
@@
 \begin{definition}
-  A cocone $(D(I)\overset{f_I}{\to} U)_{I\in \bm{I}}$ is \textbf{universal}, if
-  for all other cocones $(D(I) \overset{g_I}{\to} V)_{I\in \bm{I}}$ there is a unique
+  A cocone $(D_I\overset{f_I}{\to} U)_{I\in \bm{I}}$ is \textbf{universal}, if
+  for all other cocones $(D_I \overset{g_I}{\to} V)_{I\in \bm{I}}$ there is a unique
   map $h \from U \to V$, such that for all $I \in \bm{I}$
@@
 \begin{example}
-  We now verify that our three examples are indeed instances of the general colimit definition.
+  We now verify for coproducts and coequalizers that they are indeed
+  instances of the general colimit definition.\\
   The diagram for the colimit is the dualized diagram of the limit.

### 2017-09-18 papers/talk-7/paper-7.tex
fixed some typos

@@
 We start with two objects. Then an object $U$ with maps to these objects is the product,
-if it exists an satisfies the universal property for products, that is, all other such objects $V$
+if it exists and satisfies the universal property for products, that is, all other such objects $V$
 factor through $U$ uniquely.\\[1em]
@@
 They all have an input diagram, and we call an object $U$ with maps into this diagram
-product/pullback/equalizer if it satisfies the universal property.
+product/pullback/equalizer if makes the diagram commute and satisfies the universal property.
 \\[1em]
@@
 
-In the following section we explain this definition. We need to clarify three things
-in particular.
+In the following section we explain this definition. We need to clarify three things for this.
 First we need to define the concept of a diagram more formally.
@@
 In the examples above the diagrams
-are the various collections of objects (and arrows) that we start with
-and over which the object $U$ has maps.
+are the various collections of objects (and arrows) that we start with.
 
@@
 there might be many parts in $\mathscr C$ which have the shape $\bm I$,
-especially when $\bm I$ is so small.
+especially when $\bm I$ is small.
 We therefore get all different diagrams of that shape by considering
@@
 This leads to two observations:\\
-\hyperref[equiinstance]{(\ref*{equiinstance})} A category $\mathscr C$ has finite products and equalizers iff
+\hyperref[equiinstance]{Section \ref*{equiinstance}:} A category $\mathscr C$ has finite products and equalizers iff
 it has pullbacks and a terminal
 object.\\
-\hyperref[arbitrarylimits]{(\ref*{arbitrarylimits})}
+\hyperref[arbitrarylimits]{Section \ref*{arbitrarylimits}:}
 If a category $\mathscr C$ has products and equalizers, then it has all limits, that is,
@@
   \label{emptyproduct}
-  The empty product, i.e. the product of zero objects, is a terminal objects
+  The empty product, i.e. the product of zero objects, is a terminal object.
 \end{lemma}
@@
   \item The empty product is just the terminal object by
-    \hyperref[emptyproduct]{Example \ref*{emptyproduct}}
+    \hyperref[emptyproduct]{Lemma \ref*{emptyproduct}}
   \item The unary product of any object $A$ is constructed from the binary product
@@
 In this section we establish that the instances of limits we have seen are related
-in a way that allows us to show how to construct these instances using others.
-This implies that if we know that a category has some of the right instances
+in a way that allows us to construct some instances using others.
+This implies that in any category which has the right instances
 we can construct other limits for free using only category theoretical arguments.
@@
   \label{arbitrarylimits}
-  Let $\mathscr C$ be a category that has all (finite) products and equalizers.
-  Let $D : \bm I \to \mathscr C$ be a (finite) diagram.
+  Let $\mathscr C$ be a category that has all products and equalizers.
+  Let $D : \bm I \to \mathscr C$ be a diagram.
   Then we can construct a limit of $D$.
@@
 The product $\displaystyle \prod_{K \overset{n}{\to} J \in \bm{I}} D_J$ has components for each map in the diagram.
-We take the diagram for pullbacks as an example:
+To clarify this, we take the diagram for pullbacks as an example:
 \begin{tikzcd}
@@
 \end{tikzcd}.
-For this example diagram the product has the following projections. One for each identity:
+For this diagram the product has the following projections. One for each identity:
 $p_{1_I} \from \prod \to I,\ p_{1_J} \from \prod \to J,\ p_{1_K} \from \prod \to K$,
@@
 as the projection maps of the cone.
-Clearly, we have all necessary projection maps and we only need to verify
-the commuting condition of cones.
-We have all maps $K \overset{n}{\rightarrow} J \in \bm I$
+Clearly, we have all necessary projection maps. Now we need to verify
+the commuting condition.
+We have for all maps $K \overset{n}{\rightarrow} J \in \bm I$
 \begin{align*}

### 2017-09-18 papers/talk-11/paper-11.tex
Connect CwF with the model
Mention support of pi and sigma types
Correct typo
Add rules for context morphisms

@@
 \[
-  \inference[$\Sigma$-I]{\Gamma \vdash a : A\\ \Gamma\vdash b: B[a / x]}{\Gamma Pair_{[x:A]B} (a,b) : \Sigma x : A. B}
+  \inference[$\Sigma$-I]{\Gamma \vdash a : A\\ \Gamma\vdash b: B[a / x]}{\Gamma \vdash Pair_{[x:A]B} (a,b) : \Sigma x : A. B}
 \]
@@
 
+We say that a category with families supports $\Pi$- and $\Sigma$-types if it's
+closed under the corresponding rules shown in previous sections.
+
+We also add the rules for context morphisms:
+
+\[
+  \inference{\Gamma \vdash A~type \qquad \sigma \from \Delta \to \Gamma}{\Delta \vdash A\sigma} \qquad
+  \inference{\Gamma \vdash a : A \qquad \sigma \from \Delta \to \Gamma}{\Delta \vdash a\sigma : A\sigma}
+\]
+\[
+  \inference{\sigma \from \Delta \to \Gamma \qquad \Gamma \vdash A \qquad \Delta \vdash a : A\sigma}{(\sigma, a) \from \Delta \to \Gamma.A}
+\]
+
 \section*{Presheaf model}
 
-Let $\cat{C}$ be a small category. In the following objects of $\cat{C}$ will
-be denoted as $I,J,K$ and morphisms as $f,g,h$.
+We will now show how any presheaf category gives rise to a category with
+families where the contexts are presheaves.  Let $\cat{C}$ be a small category.
+In the following objects of $\cat{C}$ will be denoted as $I,J,K$ and morphisms
+as $f,g,h$.
 

### 2017-09-20 papers/paper-4/AdjunctionsWriteup.tex

\def\pathToRoot{../}\input{\pathToRoot headers/uebungHeader}
\usepackage{charter}
\fontfamily{bch}\selectfont

\begin{document}
\title{Galois Connections and Adjunctions}
\author{Sarah Mameche}
\maketitle

\section*{Introduction}

	Apart from isomorphism and equivalence, there is another very common relation between categories, called adjunction or adjointness. Though it is a weaker notion than equivalence, it appears very commonly and characterizes important properties of categories.\\\\We will first only consider adjunctions between partially ordered sets, which are known as Galois connections, to get some intuitions. Afterwards, we will see how Galois connections fit into category theory and how they can be generalized to adjunctions between arbitrary categories.\\
\section*{Galois Connections}

\subsection*{Motivation}
	In this section, we concentrate on partially ordered sets, in short posets, i.e. sets together with a binary relation that is reflexive, transitive and antisymmetric.\\ 
\paragraph{Example 1}
	Consider the powerset over the integers, $\pset({\mathbb{Z}})$,  and the set of closed intervals, $\mathit{intervals}({\mathbb{Z}})$. Both are partially ordered by inclusion, and the following monotone functions map between them in a natural way:
	\[\mathit{int} \from \pset({\mathbb{Z}}) \to \mathit{intervals}({\mathbb{Z}}), \,\, x \mapsto [\mathit{inf(x)}, \mathit{sup(x)}]\]
	\[\;\;\;\;\;\;\;\;\;\;\;\mathit{set} \from  \mathit{intervals}({\mathbb{Z}}) \to\pset({\mathbb{Z}}), \,\, [a, b] \mapsto \{x \in \mathbb{Z} \mid a \leq x \leq b \}\]

	Here, $int$ maps from a concrete set of numbers to a less informative interval which only specifies the range of numbers in the set, and $set$ creates a set with all integers contained in some interval's range. \\\\
	While the original posets are not isomorphic, we have $Im(int) = \mathit{intervals}(\mathbb Z) \cong Im(set)$ because $Im(set)$ only contains sets with all integers in a certain range. The following properties show in more detail how the posets are related.\\\\
	Note first that composing $set$ after $int$ on some set $x$ results in a possibly bigger set which contains $x$. Composing the functions the other way on an interval $y$ yields an interval that is contained by (in fact, equal to) $y$.
	\[\text{(i)}\;x \subseteq set(int(x)) \land int(set(y)) \subseteq y\]

	Given an arbitrary set $x$ and interval $y$, we can make a second observation. The interval $int(x)$ is contained in $y$ if no numbers from the set $x$ are beyond the interval borders of y. By construction of $set$, it is equivalent to state that $x$ is contained in $set(y)$.	
	\[\text{(ii)}\;int(x) \subseteq y \Leftrightarrow x \subseteq set(y)\;\;\;\;\;\;\;\;\;\;\;\,\]

	We say that the functions form a Galois connection, denoted by  $int \dashv set$.
	  In the context of this example, you can  think of this as a connection between abstraction and concretization, where sets are the concrete objects and intervals are abstract objects. In general, a Galois connection between a pair of posets always consists of two order-preserving functions which map in both directions between the posets. {\begin{flushright}$\square$\end{flushright}}
	\begin{definition}{Galois Connection} 
	Let $\mathcal{P} = (P, \preceq_P)$ and 	$\mathcal{Q} = (Q, \preceq_Q)$ be two partially ordered sets. \\A Galois connection $F \dashv G$ between $\mathcal{P}$ and $\mathcal{Q}$  consists of two monotone functions $F : P \rightarrow Q$ and $G : Q \rightarrow P$, such that 
	\[F(p) \preceq_Q q \Leftrightarrow p \preceq_P G(q)\] $\;\text{for all } p \in P, \; q \in Q$.
	\end{definition} The following notation can be used:
	\[\begin{tikzcd}	
	[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\mathcal{P} \arrow[r, shift left=2,"F","\bot"']
	\& \mathcal{Q} \arrow[l, shift left=2, "G"]
	\end{tikzcd}\]

	Recall from example 1 that we found two properties of the connection $int \dashv set$. The first shows how composition of the functions behaves with respect to the partial orders: elements of a poset can only get bigger when composing one way, and elements of the other poset only get smaller when composing the other way. The second one specifies how a member of one poset relates to one from the other poset which we applied one of the functions to. We used (ii) in the definition because it will be used for generalizations later, but both (i) and (ii) are common characterizations for a Galois connection, and they can be shown to be equivalent.
	\\

\paragraph{Theorem 2}  
	Let $\mathcal{P} = (P, \preceq_P)$ and $\mathcal{Q} = (Q, \preceq_Q)$ be posets and $F: P \rightarrow Q$ and $G: Q \rightarrow P$ monotone functions that form a Galois connection $F \dashv G$. The following characterizations for $F \dashv G$ are equivalent:
	\[\text{(i)}\;p \preceq_P G (F (p)) \land F(G (q)) \preceq_Q q\]
	\[\text{(ii)}\;  F(p) \preceq_Q q \Leftrightarrow p \preceq_P G(q)\;\;\;\;\;\;\;\;\,\]
%$\text{for all } p \in P, \; q \in Q$.

\begin{proof}  
	(i) $\Rightarrow$ (ii):\\
	Assume (i) and $p \preceq_P G(q)$. Since $F$ is monotone, $F(p) \preceq_Q F(G(q))$, and with (i):  $F(p) \preceq_Q F(G(q)) \preceq_Q q$.
	Assume (i) and $F(p) \preceq_Q q$. Applying monotone $G$ and using (i), $ p \preceq_P G(F(p)) \preceq_P G(q)$.\\ \\
	(ii) $\Rightarrow$ (i): \\Let $q := F(p)$. Since $F(p) \preceq_Q F(p)$, (ii) implies $p \preceq_P G (F (p))$. \\
	Similarly, with $p := G(q)$,  (ii) implies $F(G (q)) \preceq_Q q$.\\
\end{proof}

\vspace{\baselineskip}
\paragraph{Example 3}
	 A logic $\mathcal{L}$ can be seen as a poset, consisting of a set of formulas or propositions and the consequence relation $\rightarrow$. $\alpha \rightarrow \beta$ means that $\beta$ follows from $\alpha$ according to $\mathcal{L}$  for propositions $\alpha, \beta \in \mathcal{L}$. Depending on the definition of $\mathcal L$ and $\rightarrow$, $\alpha \rightarrow \beta$ can for example mean that everything made true by $\alpha$ is made true by $\beta$, or that there is a proof for $\beta$ if there is one for $\alpha$. It is obvious that $\to$ is reflexive and transitive, and if we see equivalent propositions as the same, it is antisymmetric.
	\\\\Fixing some formula $\alpha$, we consider the following functions:
	\[f = (- \land \alpha) \from \mathcal{L} \to \mathcal{L}, \,\, \beta \mapsto \beta \land \alpha\]
	\[\;\;\;g = (\alpha \rightarrow - ) \from \mathcal{L} \to \mathcal{L}, \,\, \gamma \mapsto   \alpha \rightarrow \gamma\\\]

	To see that there is a Galois connection $f \dashv g$, we spell out the properties (i) and (ii) from the previous section.\\
	\[\text{(i)}\;\beta \rightarrow (\alpha \rightarrow (\beta \land \alpha)) \land ((\alpha \rightarrow \gamma) \land \alpha) \rightarrow \gamma\]
	\[\text{(ii)}\;((\beta \land \alpha) \rightarrow \gamma) \Leftrightarrow (\beta \rightarrow (\alpha \rightarrow \gamma))\;\;\;\;\;\;\;\;\;\;\;\;\;\\\]

	It can easily be checked that the statements are logically valid. (i) formulates modus ponens and the introduction rule for $\land$. (ii) it is expresses that assuming two statements subsequently is equivalent to assuming their conjunction. So there is a Galois connection between conjunction and implication.  \\
		\[\begin{tikzcd}	
	[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\mathcal{L} \arrow[r, shift left=2,"\land","\bot"']
	\& \mathcal{L} \arrow[l, shift left=2, "\rightarrow"]
	\end{tikzcd}\]{\begin{flushright}$\square$\end{flushright}}
\paragraph{Example 4}
	Take a non-empty poset $\mathcal{P} = (P, \preceq)$ and the singleton poset $\mathcal{O} = (\{0\}, =)$, with an arbitrary object $0$ and the only possible order relation  $ =$. 
	For monotone functions $F: \mathcal{P} \to \mathcal{O} \text{ and } G, H: \mathcal{O} \to \mathcal{P}$, we look for possible Galois connections $F \dashv G$ and $H \dashv F.$
	Note that $G$ and $H$ map $0$ to some object in $P$, and $F$ is the constant $0$-function.
	If the connection $F \dashv G$ exists, then \[F(p)  = 0   \Leftrightarrow   p  \preceq  G(0) \,\,\,(\forall p \in P)\]
	\[\;\,\Leftrightarrow\;\;\;\;\;\;\top \Leftrightarrow p   \preceq  G(0) \;\;(\forall p \in P),\]
	which means $G(0)$ is the maximum of $P$.\\\\
	Analogously, if $H \dashv F$ exists, $H(0)$ is the minimum of $P$.  {\begin{flushright}$\square$\end{flushright}}\vspace{\baselineskip}

\subsection*{Properties of Galois Connections}

\paragraph{Theorem 5} 
	Given posets $\mathcal P, \mathcal Q, \mathcal R$. If $F \dashv G$ is a Galois connection between $\mathcal P$ and $\mathcal Q$ and $F' \dashv G'$ is one between  $\mathcal Q$ and $\mathcal R$, then also $F'\circ F \dashv G' \circ G$. 
\begin{proof}
		Given any $p \in \mathcal P$, $r \in R$, we get $F(p) \preceq G'(r) \Leftrightarrow p \preceq G(G'(r))$ and $F'(F(p)) \preceq r \Leftrightarrow F(p) \preceq G'(r)$  from the connections. Therefore, $F'(F(p)) \preceq r \Leftrightarrow p \preceq G(G'(r))$ which means $F'\circ F \dashv G' \circ G$.\\
	\end{proof}
\vspace{\baselineskip}
\paragraph{Theorem 6}  
	Let $\mathcal P$ and $\mathcal Q$ be posets and  $F,F' \from P \to Q$ and $G,G' \from Q \to P$ be monotone functions.
\[\text{(i) If } F \dashv G \text{ and }F \dashv G',\text{ then } G = G'. \]
\[\text{(ii) If } F \dashv G \text{ and }F' \dashv G,\text{ then } F = F'. \]
\begin{proof}
	We show (i) by proving $G' \preceq G$ and $G \preceq G'$.
	For a $q \in Q$, it follows from $F \dashv G$ that $F(G'(q)) \preceq q \Leftrightarrow G'(q) \preceq G(q)$. With 1.1.2 applied to  $F \dashv G'$, we get $F(G'(q)) \preceq q$, thus $G'(q) \preceq G(q)$.\\\\ Similarily, $F \dashv G'$ implies $F(G(q)) \preceq q \Leftrightarrow G(q) \preceq G'(q)$ and with 1.1.2, $F(G(q)) \preceq q$. Thus also $G(q) \preceq G'(q)$. (ii) can be shown the same way.\\
\end{proof}

\vspace{\baselineskip}
	According to Theorem 6, if there exists a Galois connection between two functions, one function uniquely determines the other. However, mind that there may be different Galois connections between the same pair of posets, or none altogether.\\\\
	The following shows how exactly one function determines the other: \\
\paragraph{Theorem 7} 
	 If $F \dashv G$ between  posets $\mathcal{P} = (P, \preceq_P)$ and  $\mathcal{Q} = (Q, \preceq_Q)$ , then
\[\text{(i)}\; \forall q \in Q.\; G(q) = max\{p \in P \such F(p) \preceq_Q q\} \]
\[\text{(ii)}\;  \forall p \in P.\; F(p) = min\{q \in Q \such p \preceq_P G(q)\} \]
\begin{proof}
 	(i) For a $q \in Q$, set $D = \{p \in P \such F(p) \preceq_Q q\}$. 
	\\We know from the definition that $\forall q, F(G(q)) \preceq_Q q$, so $G(q) \in D$.
	\\\\To show that $G(q)$ is also the maximum of D, take some $d \in D$, i.e. $F(d) \preceq_Q q$. 
	\\With monotonicity of $G$, $G(F(d)) \preceq_Q G(q)$. From the property 1.1.2 (i), $d \preceq_Q G(F(d))$, so $d \preceq_Q G(q)$.
	\\\\ (ii) can be shown analogously.\\
\end{proof}
\vspace{\baselineskip}\vspace{\baselineskip}\newpage
\subsection*{Galois Connecions in Category Theory}
	A partially ordered set $\mathcal{P} = (P, \preceq_P)$ can be seen as a category $\mathscr P$ where objects are the elements of the set, and there is a unique arrow $f \from p \to p'$  between two objects iff $p \preceq_P p'.$\\\\
	%aturality condition of functors oder so -> monoton? oder nat cond of Def? 
	When looking for Galois connections between two poset categories $\mathscr P = (P, \preceq_P)$ and $\mathscr Q = (Q, \preceq_Q)$, it makes sense to consider functors  $F : \mathscr P \rightarrow \mathscr Q$ and $G : \mathscr Q \rightarrow \mathscr P$ instead of monotone functions. To see that the functors are order-preserving, recall from the definition of functors that there is a function\[\mathscr P(p, p') \to\mathscr Q(F(p), F(p'))\] for all $p, p' \in \mathscr P$. Given that each set of morphisms in a poset category is either empty or contains exactly one map, the above means that $p \preceq_P p' \Rightarrow F(p) \preceq_Q F(p')$. We can argue analogously for $G$. Therefore, there is no need to require monotonicity here. \\\\
	The property that every set of morphisms contains at most one element can also be used to formulate the condition for a Galois connection differently.\\

\paragraph{Lemma 8} 
	A Galois connection $F \dashv G$ between poset categories $\mathcal{P}$ and $\mathcal{Q}$ can be described by a pair of functors $F : \mathscr P \rightarrow \mathscr Q$ and $G : \mathscr Q \rightarrow \mathscr P$ where
	\[\mathscr Q(F(p), q) \cong \mathscr P(p, G(q))\]
	naturally in $p \in \mathscr P, \; q \in \mathscr Q$. 
	\\
\begin{proof}
	The condition that $ \mathscr Q(F(p), q) \cong \mathscr P(p, G(q))$ 
	 is naturally isomorphic in $p \in \mathscr P, \; q \in \mathscr Q$ means that there is an isomorphism $\alpha _{p,q} \from \mathscr Q(F(p), q) \to \mathscr P(p, G(q))$ for arbitrary $p, q$ (Leinster, 1.3.11). We prove that this is equivalent to $F(p) \preceq_Q  q \Leftrightarrow p \preceq_P G(q)$. 
	 \\\\
	$(\Rightarrow)$ Assume $\alpha _{p, q}$ is an isomorphism.
	\\
	If $F(p) \preceq_Q  q$, there is a morphism $f \in  \mathscr Q(F(p), q)$. Then $\alpha _{p, q}(f)$ is a morphism in $\mathscr P(p, G(q))$, i.e. $p \preceq_P G(q)$. 
	\\
	If $p \preceq_P G(q)$, there exists a morphism $g \in \mathscr P(p, G(q))$. Because $\alpha _{p, q}$ is an isomorphism,  $\alpha^{-1} _{p,q} (g) \in  \mathscr Q(F(p), q)$, therefore $F(p) \preceq_Q  q$.
	\\\\
	$(\Leftarrow)$
	Assume $F(p) \preceq_Q q \Leftrightarrow p \preceq_P G(q)$. Since $\mathscr{P}, \mathscr{Q}$ are posets, we can distinguish between two cases: 
	 \[\text{(i)} \;\;|\mathscr Q(F(p), q)| =0 \Leftrightarrow (F(p), q) \notin \; \preceq_Q \; \Leftrightarrow (p, G(q)) \notin \; \preceq_P \; \Leftrightarrow |\mathscr P(p, Q(q))| = 0\]
	 \[\text{(ii)} \;\,|\mathscr Q(F(p), q)| = 1 \Leftrightarrow F(p) \preceq_Q q \Leftrightarrow p \preceq_P G(q) \Leftrightarrow |\mathscr P(p, Q(q))| = 1\;\;\;\;\;\;\;\;\;\;\;\;\;\;\]
	 In both cases, the second equivalence follows from the assumption, the other ones follow from the definition of poset categories. Because the sets are isomorphic, there is a bijection $\alpha_{p,q} \from \mathscr Q(F(p), q) \to \mathscr P(p, G(q))$.
	 \\
\end{proof}

\section*{Adjunctions}
	In the previous section, we characterized Galois connections only with the use of functors and a natural isomorphism between certain sets of morphisms. In particular, we are not tied to partially ordered sets anymore and can generalize the result of Lemma 8 for arbitrary categories. In the general case, we no longer speak of Galois connections but of adjunctions and adjoint functors.\\
\subsection*{Definition}
\begin{definition}{Adjunction}Let $\mathscr A$, $\mathscr B$ be categories. \\
	A pair of functors $F: \mathscr A \rightarrow \mathscr B$ and $G: \mathscr B \rightarrow \mathscr A$ is adjoint, written $F \dashv G$, iff
	\[\mathscr B(F(A), B) \cong \mathscr{A}(A, G(B))\]
 naturally in $A \in \mathscr A, \; B \in \mathscr B$.
\end{definition}
\vspace{\baselineskip}
 	If $F \dashv G$,  $F$ is called left adjoint to $G$, and $G$ right adjoint to $F$. We illustrate adjointness with 
 	\[
 	\begin{tikzcd}	
 		[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
 		\mathscr{A} \arrow[r, shift left=2,"F","\bot"']
 		\& \mathscr{B.} \arrow[l, shift left=2, "G"] 
 	\end{tikzcd}\]\\
 	An adjunction is the natural isomorphism between the sets of morphisms given in the defintition. For every possible choice of objects $A \in \mathscr A$ and $B \in \mathscr B$, the isomorphism fixes a bijecion between $\mathscr B(F(A), B)$ and $\mathscr A(A, G(B))$. Thus, each morphism $g \in \mathscr B(F(A), B)$ has a 
 	corresponding arrow $f \in \mathscr A(A, G(B))$, and vice versa. \\\\Corresponding morphisms $f$ and $g$ are called transposes of each other, which can be indicated by $f = \bar{g}$ and $g = \bar{f}$, with $\bar {\bar{f}} = f$ and $\bar {\bar{g}} = g.$
 	\\\\
\subsection*{Naturality}
	The definition in 2.1 has the requirement that the bijections are natural in $A \in \mathscr A$ and $B \in \mathscr B$. This ensures that the adjuncion agrees with arrow composition in the categories $\mathscr A$ and $\mathscr B$. We will look at the naturality requirement in more detail now.
	 \\\\
	First, naturality only in $B \in \mathscr B$ means that maps in $\mathscr B(F(A), B)$ and their transposes compose with outgoing maps $q : B \to B'$ and their transposes in a natural way, as illustrated in the following diagram.

	\[
	\begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}, row sep = 7ex, column sep =10ex]
	\mathscr{B}(F(A), B) \arrow[r, "q \circ -"]  \arrow[d, "\alpha_{A, B}"]
	\&  \mathscr{B}(F(A), B')  \arrow[d, "\alpha_{A, B'}" ] \\
	\mathscr{A}(A, G(B)) \arrow[r,  "G(q) \circ -"]
	\& \mathscr{A}(A, G(B'))
	\end{tikzcd}
	\]\\\\
	Here, $\alpha_{A, B}$ is the natural transformation which sends a morphism to its transpose.\\\\
	In order to have naturality in $B$, the diagram needs to commute. That is to say, 
	$\alpha_{A, B'}(q \circ g) = G(q) \circ \alpha_{A, B}(g)$ has to hold for any $g \in \mathscr B(F(A), B)$. Using the notation 
	 with transposes, 
	\[\text{(i)}\;\;\overline{q \circ g} = G(q) \circ \bar{g} \;\;\;\;\]\\
	Naturality in $A \in \mathscr A$ works similarly, but we look at ingoing maps $h: A' \to A$:

	\[
	\begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize},, row sep = 7ex,  column sep =10ex]
	\mathscr A(A', G(B)) \arrow[r, "- \circ h"]  \arrow[d, "\alpha_{A', B}"]
	\& \mathscr{A}(A, G(B))  \arrow[d, "\alpha_{A, B}" ] \\
	\mathscr B(F(A'), B) \arrow[r,  "- \circ F(h)"]
	\& \mathscr B(F(A), B)
	\end{tikzcd}
	\]\\\\
	The diagram commutes if $\alpha_{A, B}(f \circ h) = \alpha_{A', B}(f) \circ F(h)$
	for any $ f \in \mathscr A(A, G(B))$. In short,
	\[\text{(ii)}\;\;\overline{f \of h} = \bar{f} \of F(h) \;\;\;\;\]\\
	The above conditions (i) and (ii) can be used equivalently to the condition that the isomorphism is natural, and they are known as the naturality axiom.
	\\\\

	\subsection*{Properties of Adjuncions}
	The definition of adjointness we have seen was based on one of the characterization of Galois connections in Theorem 2. The other one can also be generalized, which leads to an alternative  definition of adjoints.
	 \paragraph{Remark 9} For categories and functors \begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
		\mathscr{A} \arrow[r, shift left=1,"F"]
		\& \mathscr{B} \arrow[l, shift left=1, "G"] 
	\end{tikzcd}, the following statements are equivalent.
	\[\text{(i) There is an adjunction } F \dashv G \text{ with } \mathscr B(F(A), B) \cong \mathscr{A}(A, G(B)) \text{ naturally in } A \in \mathscr A, \; B \in \mathscr B.\]
	\[\text{(ii) There is a pair of natural transformations }
	\eta \from 1_\mathscr{A} \to G \of F, \epsilon \from F \of G \to 1_\mathscr{B}.\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\\\]\\
	The natural transformations $(\eta, \epsilon)$	have to satisfy some additional equations to ensure naturality (Leinster, 2.2.2) but we won't go into detail on this here because we focus on the first definition.\\\\
	We can also genearalize the results of Theorems 5 and 6.\\
\paragraph{Theorem 10} 
	Adjunctions compose. \\Given
\begin{tikzcd}
	[ampersand replacement=\&, every label/.append style = {font = \scriptsize}]
	\vspace{\baselineskip}\mathscr{A} \arrow[r, shift left=2,"F","\bot"'] 
	\& \mathscr{B} \arrow[l, shift left=2, "G"] \arrow[r, shift left=2,"F'","\bot"'] 
	\& \mathscr{C} \arrow[l, shift left=2,"G'"] 
\end{tikzcd}
	it also holds that 
	\begin{tikzcd}
		[ampersand replacement=\&, every label/.append style = {font = \scriptsize}]
		\vspace{\baselineskip}\mathscr{A} \arrow[r, shift left=2,"F' \circ F","\bot"'] 	
		\& \mathscr{C} \arrow[l, shift left=2,"G' \circ G"] 
	\end{tikzcd} with
	\[
	\mathscr C(F'(F(A)), C) \; \cong \; \mathscr B(F(A), G'(C))\; \cong \; \mathscr A(A, G(G'(C)))
	\]
	naturally in $A \in \mathscr A, C \in \mathscr C$.
	\begin{proof}
		If there are adjunctions $F \dashv G$ and $F' \dashv G'$, there are natural transformations $\alpha_{A,B}$ and $\beta_{B',C}$ such that $\mathscr B(F(A), B) \cong \mathscr{A}(A, G(B))$ and $\mathscr C(F'(B'), C) \cong \mathscr{B}(B', G(C))$ for any $A \in \mathscr A, B,B' \in \mathscr B, C \in \mathscr C$. \\\\
		Horizontal composition of natural transformation produces a natural transformation again (Leinster, 1.3.24). With $B = G(C)$ and $B' = F(A)$, composition of $\beta_{F(A), C} $ and $ \alpha_{A, G'(C)}$ results in a natural transformation  $\gamma_{A,C}$ such that $\mathscr C(F'(F(A)), C) \; \cong  \; \mathscr A(A, G(G'(C)))$ naturally in $A$ and $C$.\\
			\end{proof}\vspace{\baselineskip}
\paragraph{Remark 11} 
	Adjoints are unique up to isomorphism. Thus, we speak of "the" left or right adjoint of a functor if the adjoint exists.\\\\
	The proof of remark 11 requires some more results (see Leinster, 4.3.13 and 4.3.18, which use the Yoneda Lemma). \\\\
\paragraph{Theorem 12} 
Adjunctions preserve initial and terminal objects.
\begin{proof}
	We want to show:  If \begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
		\mathscr{A} \arrow[r, shift left=2,"F","\bot"']
		\& \mathscr{B} \arrow[l, shift left=2, "G"] 
	\end{tikzcd}
	and $I$ is the initial object of $ \mathscr A$, then $F(I)$ is initial in $\mathscr B$.\\
	Because $I$ is initial, maps $I \to A$ are unique $\forall A \in \mathscr A$. 
	This implies that maps  $I \to G(B)$ are unique $\forall B \in \mathscr B$ since $G(B) \in \mathscr A$. 
	From $F\dashv G$, we get that the corresponding maps $F(I) \to B$ are also unique, in other words, $F(I)$ is the initial object of $\mathscr B$.\\\\Dually, terminal objects are preserved by $G$.\\
\end{proof}
\subsection*{Forgetful functors}
	Recall the pair of functions in the introductory example, where one constructs an interval from a set whereas the other "forgets" that something is an interval and maps it to the underlying set. The upcoming examples follow a similar intuition. They involve free and forgetful functors, which relate a category of algebraic objects with another category where the objects have some additional structure. In the next section, the free functors are named $F$ and the forgetful functors $U$.\\
\paragraph{Example 13}  For a field $K$, we have
	  \[ \begin{tikzcd}	
	[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\textbf{Set} \arrow[r, shift left=2,"F","\bot"']
	\& \textbf{Vect$_K$} \arrow[l, shift left=2, "U"]
	\end{tikzcd}\]
	The forgetful functor $U$ maps a vector space to its underlying set and a linear map to a function. Its counterpart $F$ maps a set $S$ to the vector space with basis $S$. %F(S)
	In more detail, $F(S)$ is the set of functions $\lambda \from S \to K$ such that $\lambda (s) \neq 0$ only for finitely many $s \in S$. This means $F(S)$ contains all $K$-linear combinations of $S$ of the form $\sum\limits_{s \in S} \lambda_s s$. $F$ maps a function between two sets to a linear map between the vector spaces that result from the sets.  \\\\\\
	To see that $F$ and $U$ are adjoint, we find a bijection
	\[\textbf{Vect}_K (F(S), V) \cong \textbf{Set}(S, U(V)) \] for a given set $S$ and vector space $V$.\\\\
	A linear map $g \from F(S) \to V$ is also a function between the underlying sets of the vector spaces, therefore $\bar g \from S \to U(V)$ is given by \[\bar g (s) = g (s) \; \forall s \in S.\]
	Going the other direction, we want to define a linear map bewteen the vector spaces $F(S)$ and $V$ given a function $f \from S \to U(V)$. Since the vector spaces have basis $S$ and $U(V)$ respectively, $f$ specifies what the basis elements of $F(S)$ are mapped to. This is sufficient to define a linear map between the spaces. We get $\bar f \from F(S) \to V$ with \[\bar f (\sum\limits_{s \in S} \lambda_s s) = \sum\limits_{s \in S} \lambda_s f(s).\]\\\\
	We verify that the mappings $g \mapsto \bar g, f \mapsto \bar f$ defined above are bijective. For $g \from F(S) \to V,\; \bar {\bar g} = g$ because
	\[\bar{\bar g} (\sum\limits_{s \in S} \lambda_s s) = \sum\limits_{s \in S} \lambda_s \bar g(s)
= \sum\limits_{s \in S} \lambda_s  g(s) = g(\sum\limits_{s \in S} \lambda_s  s).\]\\
	And for $f \from S \to U(V),\; \bar {\bar f} = f$ because $\bar {\bar f} (s) = \bar f (s) = f(s) \; \forall s \in S.$  {\begin{flushright}$\square$\end{flushright}}
\paragraph{Example 14} 
	A similar example is the adjunction
	 \[ \begin{tikzcd}	
	[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\textbf{Set} \arrow[r, shift left=2,"F","\bot"']
	\& \textbf{Mon} \arrow[l, shift left=2, "U"]
	\end{tikzcd} \]
	 where the forgetful functor $U$ maps a monoid to the set of its elements, and homomorphisms to functions. The free monoid of a set $S$ is the set of lists over $S$. So an element of $F(S)$ is a number $n$ and a function from the set $\{i \in \mathbb N \such i < n\}$ into $S$. Monoid multiplication is concatenation of lists, and the identity element is the empty list.\\\\
	 There is a correspondence between functions $f \from S \to U(M)$ and homomorphisms $h \from F(S) \to M$. As in the previous example, this is explained by the fact that homomorphisms can be seen as functions between the underlying sets, and that functions between the sets $S$ and $U(M)$ uniquely determine a homomorphism between the respective free monoids $F(S)$ and $M$.{\begin{flushright}$\square$\end{flushright}}\vspace{\baselineskip}
	So far, we discussed the usual case that a forgetful functor is right adjoint to a free functor. However, forgetful functors can also be left adjoints themselves. \\
\paragraph{Example 15} Consider the following adjunction.
	\[\begin{tikzcd}	
	[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\textbf{Mon} \arrow[r, shift left=5,"F","\bot"'] \arrow[r, shift right=5,"\bot","R"']
	\& \textbf{Grp} \arrow[l, shift left=0, "U" description]
	\end{tikzcd}\] Again, the forgetful functor is inclusion, and it is right adjoint to a free functor $F$. $F$ adds just enough elements to a monoid as needed for it to be a group, namely all the missing inverses. There is a second way to turn a monoid into a group, though: monoid elements which do not have an inverse can be removed. This is realized by $U$'s right adjoint $R$ which maps a monoid to the submonoid of invertible elements. {\begin{flushright}$\square$\end{flushright}}\vspace{\baselineskip}
%TODO mention (co)reflective subcat?
\paragraph{Example 16} Here is another example where the forgetful functor $U$ appears as the left adjoint.
	\[\begin{tikzcd}	
	[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	\textbf{Set} \arrow[r, shift left=2,"U","\bot"']
	\& \textbf{Rel} \arrow[l, shift left=2, "P"]
	\end{tikzcd}\]
	  The right adjoint is the powerset functor $P$ which maps a set $A$ to its powerset $\mathcal{P}(A)$. For sets $A, B$, the functor maps a relation $R \subseteq A \times B$ to the function \[f_R \from \mathcal{P}(A) \to \mathcal{P}(B), \; f_R(X) = \{b \in B \such \exists a \in X. \;(a,b) \in R \}.\]

	By construction, a relation $R \subseteq U(A) \times B = A \times B$ describes the same as $f_R \from A \to \mathcal{P}(B)$. Thus, we have \[\textbf{Rel }(U(A), B) \cong \textbf{ Set}(A, P(B)).\]\\
	Here, we also show how to use the naturality axiom to verify that the isomorphism is natural in $A, B$. \\
	\paragraph{Lemma 17}The isomorphism $\textbf{Rel }(U(A), B) \cong \textbf{ Set}(A, P(B))$ is natural in $A$ and $B$. \begin{proof} For functions $f \from A \to B, h \from A' \to A$,  the constructed relation for $f \of h$ is $R_{f \of h} = \{(x,y) \such  y \in f(h(x))\}.$ 
	The relation $R_f \of P(h)$ is the same one:  \[R_f \of P(h)= R_f \of R_h = \{(x,z) \such \exists y. \; z \in f(y) \land y \in h(x)\} = R_{f \of h}.\]
	For naturality in $B$, take relations $R \subseteq A \times B, R \subseteq B \times B'$, then \[f_{R' \of R} = \{z \in B' \such \exists x \in A. \; (x,z) \in R' \of R \} = \{z \in B' \such \exists x \in A, y \in B. \; (x,y) \in R \land (y,z) \in R' \}\] 
	\[=\{z \in B' \such \exists y \in B. \; (y,z) \in R' \}\of  \{y \in B \such \exists x \in A. \; (x,y) \in R \} =f_{R'} \of f_R = U(R') \of f_R.\,\,\,\,\,\,\,\,\,\,\,\,\,\,\]\end{proof}
\vspace{\baselineskip}
	So as you would expect, composition of functions and relations in \textbf{Set} and \textbf{Rel} behaves in a way that does not violate the naturality condition of the adjunction. In most situations like \begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
		\mathscr{A} \arrow[r, shift left=1,"F"]
		\& \mathscr{B} \arrow[l, shift left=1, "G"] 
	\end{tikzcd} where functors go to a  category and back, adjointness follows automatically because of the way functors preserve the existing structure in the catgories. For this reason, forgetful functors like the ones we have seen usually have left adjoints, regardless of how exactly the free functor is constructed.  \\\\
\subsection*{Use of adjoints to characterize properties of categories}

\paragraph{Example 18} 
	Recall from Example 1.1.4 that if a poset has a Galois connection with the singleton poset $\mathcal{O}$, the set has a maximum or minimum. Now we take some non-empty category $\mathscr C$ and the one-object category \textbf{1} and look at functors $F:  \mathscr C \to$ \textbf{1} and
	$G,H$ : \textbf{1} $\to \mathscr C $. \\\\
	If \textbf{1} has the object $\bullet$ and the map $1{_\bullet}$, $F(A) = \bullet$ and $F(f) = 1{_\bullet}$ for all $f \in \mathscr C (A, B)$ and $A, B \in \mathscr C $.
	 \\ For $F \dashv G$, we need \[\textbf{1}(F(A), \bullet) \; \cong \;  \mathscr C (A, G(\bullet))\]
	 \[\Leftrightarrow \textbf{1}(\bullet, \bullet) \; \cong \;  \mathscr C (A, G(\bullet))  \]
	 \[\Leftrightarrow \textbf{1}_\bullet \; \cong \;  \mathscr C (A, G(\bullet))  \]
	naturally $\forall A \in\mathscr C $. All this means is that maps from $A$ to $G(\bullet)$ are unique, i.e. $G(\bullet)$ is $\mathscr C $'s initial object.	
	\\With duality, $H \dashv F$ if and only if $H(\bullet)$ is terminal in $\mathscr C $.
	\\\\As a result, the existence of an initial (terminal) object in a category $\mathscr C$ is determined by the existence of the right (left) adjoint of the constant functor $F \from \mathscr C \to$ \textbf{1}. {\begin{flushright}$\square$\end{flushright}}
	\vspace{\baselineskip}
	\vspace{\baselineskip}
	 \paragraph{Example 19} 
	For a category $\mathscr C$, the diagonal functor $\Delta \from \mathscr C \to \mathscr C \times \mathscr C$ maps objects $C$ and morphisms $f$ in $\mathscr C$ to pairs $\langle C,C \rangle$ and $\langle f, f \rangle$ in the product category $\mathscr C \times \mathscr C$. For a right adjoint $\Pi \from \mathscr C \times \mathscr C \to \mathscr C$, we need \[(\mathscr C \times \mathscr C)(\langle A, A \rangle, \langle B, C \rangle) \cong \mathscr C (A, \Pi (B,C))\]
\[\Leftrightarrow \mathscr C (A, B) \times \mathscr C (A, C) \cong \mathscr C (A, \Pi (B,C)),\]
by the definition of product categories.\\\\
	For $\mathscr C =$ \textbf{Set}, $\Pi$ is the Cartesian product, and the above equation means that a pair $\langle f \from A \to B, g \from A \to C \rangle$ can be represented by a single function $h: A \to B \times C$. 
	In general, there is an adjunction $\Delta \dashv \Pi$ whenever the category $\mathscr C$ has binary products. {\begin{flushright}$\square$\end{flushright}}
\vspace{\baselineskip}
\vspace{\baselineskip}
\paragraph{Example 20} 
	In 2.4.5, we established that \textbf{Set} and \textbf{Rel} have an adjunction because a function $f \from A \to 2^B$ can be seen as a relation $R \subseteq A \times B$. Note that $R$ can again be interpreted as a function $f \from A \times B \to \{0, 1\}$, so we could have defined this adjunction within the category \textbf{Set}.\\\\In fact, \{0,1\} can be replaced by an arbitrary set $C$ and the resulting functions still describe the same. To see this, consider the following functors:
	 \[(-\times B) \from \textbf{Set} \to \textbf{Set}, A \mapsto A \times B\]
	\[(-)^B \from \textbf{Set} \to \textbf{Set}, C \mapsto C^B\]
	 \\where $B$ is a fixed set and $C^B = B \to C$.\\\\
	 A function $f \from A \times B \to C$ can be mapped to a function $\bar f \from A \to C^B$ that assigns a map $B \to C$ to each element of $A$, instead of assigning a single element of $C$ to each pair in $ A \times B$. It is defined in the natural way by \[(\bar f (a))(b) = f(a, b)\;\; \forall a \in A, b \in B.\] \\Similarly, given a function $g \from A \to C^B$, we set $\bar g \from A \times B \to C$ to \[\bar g (a, b) = (g(a))(b)\;\; \forall a \in A, b \in B.\]\\
	 The above defines a bijection \[\textbf{Set}(A \times B, C) \cong \textbf{Set}(A, C^B),\] resulting in the adjunction
	  	\[\begin{tikzcd}	
	 	[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
	 	\textbf{Set} \arrow[r, shift left=2,"-\times B","\bot"']
	 	\& \textbf{Set} \arrow[l, shift left=2, "\;\;\;(-)^B"]\\
	 \end{tikzcd}\]
	 In other words, a function taking multiple arguments evaluates to the same as a sequence of functions with a single argument each. The process of translating between functions in this way is known as currying and uncurrying.\\\\
	 Note also that we have actually already seen this adjunction before. In Example 3, the sets are logical propositions (or sets containing the propositions that are equivalent). For propositions $\alpha , \beta$, we interpreted the cartesian product $(\alpha , \beta)$ as conjunction $\alpha \land \beta$ and functions $\alpha \rightarrow \beta$ as implication and found that $\land \dashv \; \to$ holds.{\begin{flushright}$\square$\\\end{flushright}}
	 \vspace{\baselineskip}
	 \vspace{\baselineskip}
\section*{Conclusion}
	We have discussed Galois connections and found that similar connections can be defined between mathematical structures other than just partially ordered sets. In fact, adjointness can be found in many other areas of mathematics and it can only be "seen" and formally defined with the use of category theory. \\\\
	Just as Galois connections are an order-isomorphism between sub-posets, adjunctions are an isomorphism of certain subsets of morphisms. So adjointness is a relation between categories that is weaker than isomorphism but still very common. Similarly to an equivalence between categories, an adjunction always consists of two functors, and in situations where there is such a pair of functors between two categories, there is a good chance that the functors are adjoint (Leinster, 2.1).
	\\\\ 
	We have seen that this is the case for some free and forgetful functors. Also, we encountered an example for a category where the product is left adjoint to the exponential, and it was shown the existence of certain adjunctions with a category exhibits some important properties of the category, like whether it has an initial or terminal object or whether it has products.

\end{document}

### 2017-09-20 papers/talk-9/talk-9.tex
copied to github

\documentclass{article}
\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}

\usepackage{todonotes}
\usepackage[utf8]{inputenc}
\usepackage{mathtools}
\usepackage{semantic}

\usetikzlibrary{cd, shapes, tikzmark}

\allowdisplaybreaks

\newtheorem{defn}{Definition}[section]
\newtheorem{exmpl}[defn]{Example}
\newtheorem{lemma}[defn]{Lemma}
\newtheorem{theorem}[defn]{Theorem}

\newcommand{\type}{\vdash}
\newcommand{\term}{1}
\newcommand{\unit}{\mathsf{unit}}
\newcommand{\app}{\mathsf{app}}
\newcommand{\cur}{\mathsf{cur}}
\newcommand{\fst}{\mathsf{fst}\,}
\newcommand{\snd}{\mathsf{snd}\,}
\newcommand{\id}{\mathsf{id}}
\newcommand{\eenv}{\diamond}

\title{Cartesian Closed Categories as a Model for the Simply Typed Lambda Calculus}
\author{Joachim Bard}

\begin{document}

\maketitle

\section{Motivation}
\label{sec:motiv}

Recall the Curry-Howard correspondence which relates logic and type theory.
According to this correspondence proposotions are types and proofs are terms.
A prominent example is the relation between intuitionistic propositional logic and simply typed lambda calculus.
We can extend the correspondence to the so called Curry-Howard-Lawvere/Lambek correspondence by adding category theory as shown in the following table.

\begin{center}
    \begin{tabular}{l|l|l}
        Logic & Type theory & Category theory\\
        \hline
        Propositions & Types & Objects\\
        Proofs & Terms & Morphisms
    \end{tabular}
\end{center}

In the following sections we will draw a connection between the simply typed lambda calculus and cartesian closed categories.

\section{Simply Typed Lambda Calculus}

The simply types lambda calculus has types $A$ and terms $s$ as shown in the following definition.

\begin{defn}{(Simply Typed Lambda Calculus)}
    \label{def:stlc}
    \begin{align*}
        A, B & ::= G \mid \unit \mid A \times B \mid A \to B \\
        s, t & ::= x \mid c^A \mid () \mid (s, t) \mid \fst s \mid \snd s \mid \lambda x: A.\, s \mid s\, t
    \end{align*}
\end{defn}

Types consists of ground types $G$, the unit type $\unit$, product types $A \times B$ and function types $A \to B$.
Possible ground types are \textsf{int} and \textsf{bool}.
Terms are variables $x$, constants $c^A$ annotated by its type $A$, the empty tuple $()$, pairs $(s, t)$, first ($\fst s$) and second ($\snd s$) projections,
function definitions $\lambda x: A.\, s$ and function applications $s\,t$.
Common constants are integers like $3^{\mathsf{int}}$ and booleans ($\mathsf{true}^{\mathsf{bool}}$ and $\mathsf{false}^{\mathsf{bool}}$).
In order to state that a term has some type we use environments.
Environments are lists of variables and their types.
Thus we have the empty environment $\eenv$ and we can extend an environment $\Gamma$ with a variable $x$ of type $A$ (written $\Gamma, x: A$).
Often the empty environment is ommited.
For example the environment $x: \mathsf{int}, y: \mathsf{bool}$ has a variable $x$ of type \textsf{int} and a variable $y$ of type \textsf{bool}.
Variables in environments are always distinct, that means $x: \mathsf{int}, x: \mathsf{int}$ is not an environment.
We use environments to `remember' the type of a variable in a function definition.
The following definition shows how environments $\Gamma$ are used to state that a term $s$ has a type $A$.
We write $\Gamma \type s: A$ for this statement.
If the environment is empty we write just $\type s: A$ instead of $\eenv \type s: A$.
Also, we say that a term $s$ is well-typed, if an  environment $\Gamma$ and a type $A$ exists, such that the judgement $\Gamma \type s: A$ holds.

\begin{defn}{(Typing Rules)}
    \label{def:type}
    \begin{align*}
        &\qquad\qquad \inference{}{\Gamma, x: A \type x: A} \qquad \inference{\Gamma \type x : A}{\Gamma , y: B \type x: A}\\
        \\
        &\inference{}{\Gamma \type c^A: A} \qquad \inference{}{\Gamma \type () : \mathsf{unit}}\qquad
        \inference{\Gamma \type s: A \quad \Gamma \type t:B}{\Gamma \type (s, t) : A \times B}\\
        \\
        & \qquad\qquad \inference{\Gamma \type s : A \times B}{\Gamma \type \mathsf{fst}\,s: A} \qquad
        \inference{\Gamma \type s : A \times B}{\Gamma \type \mathsf{snd}\,s: B}\\
        \\
        & \quad \inference{\Gamma, x:A \type s: B}{\Gamma \type \lambda x: A.\, s: A \to B} \qquad
        \inference{\Gamma \type s: A \to B \quad \Gamma \type t: A}{\Gamma \type s\,t: B}
    \end{align*}
\end{defn}

We observe that the type of a variable is determined by looking into the environment.
The empty tuple $()$ is the only term of type $\unit$.
Furthermore, projections are only applicable to pairs.
As already mentioned above, the defined variable $x$ in a function definition $\lambda x: A.\, s$ is added to the environment
in order to specify the type of the body $s$.
But this directs to a problem if we consider the term $s := \lambda x: A.\,\lambda x: B.\,x$.
Applying the rules starting from the empty environment leads to $x: A, x: B \type x: B$.
The problem is that we have the variable $x$ twice in our `environment' $x: A, x: B$.
This is the reason why variables have to be distinct in environments.
To resolve this problem we consistently rename variables.
We rename the second $x$ to $y$ and obtain $s = \lambda x: A.\,\lambda y: B.\, y$.
Now $s$ has the type $A \to B \to B$, because we can derive $\type \lambda x: A.\,\lambda y: B.\, y: A \to B \to B$.
The consistent renaming of variables does not alter the term significantly.
Therefore, we always allow this renaming in the following sections
and call two terms $\alpha$-equivalent if one can be transformed into the other.

\section{Reminder: Cartesian Closed Category}

Before connecting cartesian closed categories and the simply typed lambda calculus in the next section,
we recall the definition of cartesian closed categories.

\begin{defn}
    \label{def:ccc}
    A cartesian closed category $\cat{C}$ consists of:
    \begin{itemize}
        \item A teminal object $\term \in \cat{C}$: $\forall X \in \cat{C}.\, \exists! \langle\rangle: X \to \term$.
        \item Binary products $X \times Y \in \cat{C}$ with projections $\pi_1 \from X \times Y \to X$ and $\pi_2 \from X \times Y \to Y$
            for all $X, Y \in \cat{C}$:
            For every $Z \in \cat{C}$ and morphisms $f \from Z \to X$ and $g \from Z \to Y$,
            there is exactly one morphism $\langle f, g \rangle \from Z \to X \times Y$, such that the following diagram commutes:
            \[
                \begin{tikzcd}
                    & Z \arrow[bend right]{ddl}{f} \arrow{d}{\langle f, g \rangle} \arrow[bend left]{ddr}{g} & \\
                    & X \times Y \arrow{dl}{\pi_1} \arrow{dr}{\pi_2} & \\
                    X & & Y
                \end{tikzcd}
            \]
            For two morphisms $p \from X \to X'$ and $q \from Y \to Y'$ we define
            \[ p \times q := \langle p \circ \pi_1, q \circ \pi_2 \rangle \from X \times Y \to X' \times Y'.\]
        \item Exponentials $Y^X \in \cat{C}$ together with a morphism $\app \from Y^X \times X \to Y$ for all $X, Y \in \cat{C}$:
            For every $Z \in \cat{C}$ and morphism $f \from Z \times X \to Y$,
            there is a unique morphism $\cur(f) \from Z \to Y^X$, such that following diagram commutes:
            \[
                \begin{tikzcd}
                    Z \times X \arrow{d}[left]{\cur(f) \times \id_X} \arrow{r}{f} & Y \\
                    Y^X \times X \arrow{ur}[right=7pt, below]{\app}
                \end{tikzcd}
            \]
    \end{itemize}
\end{defn}

\section{Semantics in a Cartesian Closed Category}
\label{sec:sem}

Let $\cat{C}$ be a cartesian closed category.
We want to express well-typed terms as morphisms in $\cat{C}$.
To that end, we first map types to objects of $\cat{C}$.
Thus, let $M$ be a function which maps every ground type $G$ to $M(G) \in \cat{C}$.
We can extend $M$ to arbitrary types:

\begin{align*}
    M|[G|] & := M(G) \\
    M|[\unit|] & := \term && (\text{terminal object}) \\
    M|[A \times B |] & := M|[A|] \times M|[B|] && (\text{product}) \\
    M|[A \to B |] & := M|[B|]^{M|[A|]} && (\text{exponential})
\end{align*}

In addition we map environments to objects of $\cat{C}$:

\begin{align*}
    M|[\eenv|] & := \term && (\text{terminal object}) \\
    M|[\Gamma, x: A|] & := M|[\Gamma|] \times M|[A|] && (\text{product})
\end{align*}

Furthermore, if $M$ maps every constant $c^A$ to a morphism $M(c^A) \from \term \to M|[A|]$,
then we define $M|[\Gamma \type s: A|] \from M|[\Gamma|] \to M|[A|]$ by recursion on $\Gamma \type s: A$ as follows.

\begin{defn}
    \label{def:type-C}
    \begin{align*}
        M|[\Gamma, x: A \type x: A|] & := M|[\Gamma|] \times M|[A|] \toby{\pi_2} M|[A|] \\
        M|[\Gamma, y: B \type x: A|] & := M|[\Gamma|] \times M|[B|] \toby{\pi_1} M|[\Gamma|] \xrightarrow{M\llbracket\Gamma \type x: A\rrbracket} M|[A|] \\
        M|[\Gamma \type c^A: A|] & := M|[\Gamma|] \toby{\langle\rangle} \term \xrightarrow{M(c^A)} M|[A|] \\
        M|[\Gamma \type (): \unit|] & := M|[\Gamma|] \toby{\langle\rangle} \term \\
        M|[\Gamma \type (s, t): A \times B|] & := M|[\Gamma|]
        \xrightarrow{\langle M\llbracket\Gamma \type s: A\rrbracket, M\llbracket\Gamma \type s: A\rrbracket\rangle} M|[A|] \times M|[B|] \\
        M|[\Gamma \type \fst s: A|] & := M|[\Gamma|] \xrightarrow{M\llbracket\Gamma \type s: A \times B\rrbracket} M|[A|] \times M|[B|] \toby{\pi_1} M|[A|] \\
        M|[\Gamma \type \snd s: A|] & := M|[\Gamma|] \xrightarrow{M\llbracket\Gamma \type s: A \times B\rrbracket} M|[A|] \times M|[B|] \toby{\pi_2} M|[B|] \\
        M|[\Gamma \type \lambda x: A.\, s: A \to B|] & := \cur(M|[\Gamma|] \times M|[A|] \xrightarrow{M\llbracket\Gamma, x: A \type s : B\rrbracket} M|[B|]) \\
        M|[\Gamma \type s\,t: B|] & := M|[\Gamma|]
        \xrightarrow{\langle M\llbracket\Gamma \type s: A \to B\rrbracket, M\llbracket\Gamma \type t: A\rrbracket\rangle}
        M|[B|]^{M|[A|]} \times M|[A|] \xrightarrow{\app} M|[B|]
    \end{align*}
\end{defn}

%The case for constants 
\todo{explain, esp. $M(c^A)$}

\begin{exmpl}
    \label{exp:sem}
    \begin{align*}
            & M |[ \ \type \lambda f: A \to B.\, \lambda x: A.\, f\,x: (A \to B) \to A \to B |] \\
        = \,& \mathsf{cur}(M |[ f: A \to B \type \lambda x: A. \, f\,x : A \to B |]) \\
        = \,& \mathsf{cur}(\mathsf{cur}( M |[ f: A \to B, x: A \type f \, x : B |])) \\
        = \,& \mathsf{cur}(\mathsf{cur}(\mathsf{app} \circ (M |[\Gamma \type f: A \to B |], M |[\Gamma \type x : A|]))) \\
            &  \text{ where } \Gamma = f: A \to B, x: A \\
        = \,& \mathsf{cur}(\mathsf{cur}(\mathsf{app} \circ (M |[ f: A \to B \type f : A \to B |] \circ \pi_1, \pi_2))) \\
        = \,& \mathsf{cur}(\mathsf{cur}(\mathsf{app} \circ (M |[ f: A \to B \type f : A \to B |] \times \mathsf{id}))) && (\text{definition}~\ref{def:ccc}) \\
        = \,& \mathsf{cur}(M |[ f: A \to B \type f : A \to B |]) && (\text{definition}~\ref{def:ccc})\\
        = \,& M |[ \ \type \lambda f: A \to B.\, f: (A \to B) \to A \to B |]
    \end{align*}
\end{exmpl}

\section[beta-eta-equivalence]{$\beta\eta$-equivalence}

In the previous section we saw how to express well-typed terms as morphisms in a cartesian closed category.
Example~\ref{exp:sem} showed that there are terms which are mapped to the same morphism.
It turns out that this holds for all $\beta$- and $\eta$-equivalent terms.
In order to define this equivalence we need substitutions.
A substitution corresponds to a function application $(\lambda x: A.\, s)\, t$, where the function is called with an argument $t$.
It replaces in the term $s$ all occurences of the variable $x$ with the term $t$.
We write $s[t/x]$ and define it as follows.

\begin{defn}{(Substitution)}
    \begin{align*}
        x[t/x] & := t\\
        y[t/x] & := y \qquad && y \neq x\\
        c^A[t/x] & := c^A\\
        ()[t/x] & := ()\\
        (s_1, s_2)[t/x] & := (s_1[t/x], s_2[t/x])\\
        (\mathsf{fst}\, s)[t/x] & := \mathsf{fst}(s[t/x])\\
        (\mathsf{snd}\, s)[t/x] & := \mathsf{snd}(s[t/x])\\
        (\lambda y:A.\, s)[t/x] & := \lambda y:A.\,(s[t/x]) \qquad && y \neq x \text{ and } y \text{ not in } t \\
        (s_1\, s_2)[t/x] & := (s_1[t/x])\,(s_2[t/x])
    \end{align*}
\end{defn}

The interesting case ($\lambda y: A.\, s)[t/x]$) is only applicable if $y \neq x$ and $y$ does not occur in $t$.
But because of $\alpha$-equivalence, we can always rename $y$ to a \emph{fresh} variable $z$ in order to ensure this condition.
\par
The next lemma illustrates how substitutions are interpreted in the category $\cat{C}$.

\begin{lemma}
    \label{lem:subst}
    If $\Gamma \type t: A$ and $\Gamma, x: A \type s: B$, then the diagram
    \[
        \begin{tikzcd}
            M \llbracket \Gamma \rrbracket \arrow{rr}{\langle \mathsf{id}, M\llbracket\Gamma \type t: A\rrbracket \rangle}
                                           \arrow{drr}[left=1cm, below]{M\llbracket\Gamma \type s[t/x] : B\rrbracket}
            & & M \llbracket\Gamma\rrbracket \times M \llbracket A \rrbracket \arrow{d}{M \llbracket\Gamma, x:A \type s: B\rrbracket} \\
            & & M \llbracket B \rrbracket
        \end{tikzcd}
    \]
    commutes.
\end{lemma}
\begin{proof}
    We omit the proof here.
    It can be found in~\cite{CT-STLC}.
\end{proof}

Using substitutions we can now define $\beta\eta$-equivalence.
If we read the $\beta$-conversions from left to right they correspond to computation steps.
Also the $\eta$-conversions relate terms, which have the same computational behavior.
The equivalence and congruence rules ensure that the $\beta\eta$-equivalence is an congruence relation.

\begin{defn}{($\beta\eta$-equivalence)}
    \begin{description}
        \item[$\beta$-conversions:]
            \begin{align*}
                & \qquad\qquad\qquad \inference{\Gamma, x:A \type s: B \quad \Gamma \type t:A}{\Gamma \type (\lambda x:A.\, s)\, t =_{\beta\eta} s[t/x]: B}\\
                \\
                & \quad \inference{\Gamma \type s: A \quad \Gamma \type t:B}{\Gamma \type \mathsf{fst}(s, t) =_{\beta\eta} s: A}\qquad
                \inference{\Gamma \type s: A \quad \Gamma \type t:B}{\Gamma \type \mathsf{snd}(s, t) =_{\beta\eta} t: B}
            \end{align*}
        \item[$\eta$-conversions:]
            \begin{align*}
                & \qquad\qquad\qquad \inference{\Gamma \type s: A \to B \quad x \text{ not in } s}{\Gamma \type s =_{\beta\eta} \lambda x: A.\,(s\,x): A \to B}\\
                \\
                & \inference{\Gamma \type s: A \times B}{\Gamma \type s =_{\beta\eta} (\mathsf{fst}\,s, \mathsf{snd}\,s): A \times B}\qquad
                \inference{\Gamma \type s: \mathsf{unit}}{\Gamma \type s =_{\beta\eta} () : \mathsf{unit}}
            \end{align*}
        \item[equivalence rules:]
            \begin{align*}
                & \inference{}{\Gamma \type s =_{\beta\eta} s: A} \qquad
                \inference{\Gamma \type s =_{\beta\eta} t: A}{\Gamma \type t =_{\beta\eta} s: A}
                \qquad \inference{\Gamma \type s =_{\beta\eta} t: A \quad \Gamma \type t =_{\beta\eta} u: A}{\Gamma \type s =_{\beta\eta} u: A}
            \end{align*}
        \item[congruence rules:]
            \begin{align*}
                & \inference{\Gamma \type s =_{\beta\eta} s': A \quad \Gamma \type t =_{\beta\eta} t': B}{\Gamma \type (s,t) =_{\beta\eta} (s',t'): A \times B}
                \qquad\inference{\Gamma \type s =_{\beta\eta} t : A \times B}{\Gamma \type \fst s =_{\beta\eta} \fst t: A} \\
                \\
                & \inference{\Gamma \type s =_{\beta\eta} t : A \times B}{\Gamma \type \snd s =_{\beta\eta} \snd t: B} \qquad
                \inference{\Gamma, x: A \type s =_{\beta\eta} t: B}{\Gamma \type \lambda x:A.\, s =_{\beta\eta} \lambda x:A.\, t: A \to B} \\
                \\
                & \qquad\qquad \inference{\Gamma\type s =_{\beta\eta} s': A \to B \quad \Gamma\type t =_{\beta\eta} t': A}{\Gamma\type s\,t =_{\beta\eta} s'\,t': B}
            \end{align*}
    \end{description}
\end{defn}

The following theorem shows that $\beta\eta$-equivalent terms are indeed mapped to the same morphism in a cartesian closed category.

\begin{theorem}{(Soundness)}
    \label{thm:sound}
    Let $\cat{C}$ be a cartesian closed category and
    let $M$ be a function which maps ground types to objects of $\cat{C}$ and constants to morphisms in $\cat{C}$ as required by section~\ref{sec:sem}.
    Every two terms $s$ and $t$ which are $\beta\eta$-equivalent of type $A$ under environment $\Gamma$ lead to equal morphisms in $\cat{C}$, i.\,e.
    if $\Gamma \type s =_{\beta\eta} t: A$, then $M|[\Gamma \type s: A|] = M|[\Gamma \type t: A|]$.
\end{theorem}
\begin{proof}
    The proof goes by induction on $\Gamma \type s =_{\beta\eta} t: A$.
    We will show the case for
    \[\Gamma \type (\lambda x:A.\, s)\, t =_{\beta\eta} s[t/x]: B.\]
    Let $f := M|[\Gamma, x: A \type s: B|]$ and $g := M|[\Gamma \type t: A|]$.
    By definition~\ref{def:type-C} we get \[\cur(f) = M|[\Gamma \type \lambda x: A.\, s: A \to B|].\]
    Using this we obtain
    \begin{align*}
        M|[\Gamma \type (\lambda x: A.\, s)\, t: B|] & = \app \circ \underbrace{\langle \cur(f), g \rangle}_{= (\cur(f) \times \mathsf{id}) \circ \langle \mathsf{id}, g \rangle} && (\text{definition}~\ref{def:type-C}) \\
                                                     & = \underbrace{\app \circ (\cur(f) \times \mathsf{id})}_{= f} \circ \langle \mathsf{id}, g \rangle && (\text{definition}~\ref{def:ccc})\\
                                                     & = f \circ \langle \mathsf{id}, g \rangle && (\text{definition}~\ref{def:ccc})\\
                                                     & = M|[\Gamma \type s[t/x]|] && (\text{lemma}~\ref{lem:subst})
    \end{align*}
\end{proof}

\section{Internal Language of a Cartesian Closed Category}

In the previous section we established a connection between $\beta\eta$-equivalent terms and equal morphisms.
We can exploit this connection to use the simply type lambda calculus to describe cartesian closed categories.
To do so, let $\cat{C}$ be a cartesian closed category.
We want to define a simply type lambda calculus which resembles $\cat{C}$, which we call internal language of $\cat{C}$.
Therefore, we take every object $G$ of $\cat{C}$ as ground type.
Furthermore, for every morphism $f \from \term \to A$, where $\term$ is the terminal object, we have a constant $f^A$ of type $A$.
The mapping from section~\ref{sec:sem} is defined as $M(G) := G$ for every ground type $G$ and $M(f^A) := f$ for every constant $f^A$.
The following example illustrates how the internal language can be used to show properties of $\cat{C}$.

\begin{exmpl}
    We show that $C^{A \times B} \iso (C^B)^A$ for all objects $A, B$ and $C$ from $\cat{C}$.
    Note that $A, B$ and $C$ are types in the internal language of $\cat{C}$.
    First, we  define lambda terms $cas$ and $car$:
    \begin{align*}
        & cas := \lambda f: A \times B \to C.\,\lambda x: A.\,\lambda y: B.\,f(x, y) \\
        & car := \lambda g: A \to B \to C.\,\lambda p: A \times B.\,g\,(\mathsf{fst}\,p)\,(\mathsf{snd}\,p),
    \end{align*}
    such that they satisfy the following typing judgements and $\beta\eta$-equivalences:
    \begin{align*}
        & \vdash cas: (A \times B \to C) \to A \to B \to C \\
        & \vdash car: (A \to B \to C) \to A \times B \to C\\
        & f: A \times B \to C \vdash car\, (cas\, f) =_{\beta\eta} f : A \times B \to C \\
        & g: A \to B \to C \vdash cas\, (car\, g) =_{\beta\eta} g : A \to B \to C.
    \end{align*}
    Furthermore, we define
    \begin{align*}
        h & := M|[f: A \times B \to C \vdash cas\,f: A \to B \to C|] \\
        h' & := M|[g: A \to B \to C \vdash car\,g: A \times B \to C|].
    \end{align*}
    Note that $\langle\rangle \circ h = \pi_1$ and thus
    \begin{align*}
        (\mathsf{cur}(h') \times \id_{\term}) \circ \langle \langle\rangle, \id_{(C^B)^A} \rangle \circ h = \langle \mathsf{cur}(h') \circ \pi_1, h \rangle. \tag{1}
    \end{align*}
    We obtain
    \begin{align*}
        & h' \circ \langle\langle\rangle, \id_{(C^B)^A}\rangle \circ h \\
        = \,& \mathsf{app} \circ (\mathsf{cur}(h') \times \id_\term) \circ \langle\langle\rangle, \id_{(C^B)^A}\rangle \circ h && (\text{definition}~\ref{def:ccc})\\
        = \,& \mathsf{app} \circ \langle \mathsf{cur}(h') \circ \pi_1, h \rangle && (\text{equation}~(1))\\
        = \,& \mathsf{app} \circ \langle M|[\ \vdash car: (A \to B \to C) \to A \times B \to C|] \circ \pi_1, h \rangle && (\text{definition}~\ref{def:type-C})\\
        = \,& \mathsf{app} \circ \langle M|[f: A \times B \to C \vdash car: (A \to B \to C) \to A \times B \to C|] , h \rangle && (\text{definition}~\ref{def:type-C})\\
        = \,& M|[f: A \times B \to C \vdash car\,(cas \,f): A \times B \to C|] && (\text{definition}~\ref{def:type-C}) \\
        = \,& M|[f: A \times B \to C \vdash f: A \times B \to C|] && (\text{theorem}~\ref{thm:sound})\\
        = \,& \pi_2 && (\text{definition}~\ref{def:type-C})
    \end{align*}
    and therefore
    \begin{align*}
        & h' \circ \langle\langle\rangle, \id_{(C^B)^A}\rangle \circ h \circ \langle \langle\rangle, \id_{C^{A \times B}} \rangle \\
        =\, & \pi_2 \circ \langle \langle\rangle, \id_{C^{A \times B}} \rangle \\
        =\, & \id_{C^{A \times B}} && (\text{definition}~\ref{def:ccc}).
    \end{align*}
    Analogously we obtain
    \[ h \circ \langle \langle\rangle, \id_{C^{A \times B}} \rangle \circ h' = \pi_2 \tag{2} \]
    and thus
    \begin{align*}
        & h \circ \langle \langle\rangle, \id_{C^{A \times B}} \rangle \circ h' \circ \langle\langle\rangle, \id_{(C^B)^A}\rangle \\
        =\, & \pi_2 \circ \langle\langle\rangle, \id_{(C^B)^A}\rangle && (\text{equation}~(2))\\
        =\, & \id_{(C^B)^A} && (\text{definition}~\ref{def:ccc}).
    \end{align*}
    This shows $C^{A \times B} \iso (C^B)^A$.
\end{exmpl}

\iffalse
\section{Free Cartesian Closed Categories}
\label{sec:free}

\begin{theorem}
    \label{thm:compl}
    If $M|[\Gamma\type s: A|] = M|[\Gamma\type t: A|]$ for every mapping $M$ of ground types and constants and every cartesian closed category $\cat{C}$,
    then $\Gamma \type s =_{\beta\eta} t: A$.
\end{theorem}
\begin{proof}
    We give a proof sketch.

\end{proof}
\fi

\nocite{*}
\bibliographystyle{plain}
\bibliography{bib}

\end{document}

### 2017-09-20 papers/talk-9/talk-9.tex
added completeness

@@
 
-\iffalse
+%\iffalse
 \section{Free Cartesian Closed Categories}
@@
 
-\begin{theorem}
+The soundness theorem~\ref{thm:sound} has a converse called completeness.
+
+\begin{theorem}{(Completeness)}
     \label{thm:compl}
@@
     We give a proof sketch.
-
+    The idea is to construct a freely generated cartesian closed category $\cat{F}$ and instantiate the assumption with $\cat{F}$.
+    Objects of $\cat{F}$ are types of the simply typed lambda calculus.
+    The set of morphisms from $A$ to  $B$ is the quotient of the set $\{s \mid \ \type s: A \to B\}$, where two terms $s$ and $t$ are equivalent iff
+    $\ \type s =_{\beta\eta} t: A \to B$ holds.
+    Showing that $\cat{F}$ is a cartesian closed category concludes the proof.
 \end{proof}
-\fi
+%\fi
 

### 2017-09-20 papers/talk-9/talk-9.tex
fixes

@@
 Recall the Curry-Howard correspondence which relates logic and type theory.
-According to this correspondence proposotions are types and proofs are terms.
+According to this correspondence propositions are types and proofs are terms.
 A prominent example is the relation between intuitionistic propositional logic and simply typed lambda calculus.
@@
 Thus we have the empty environment $\eenv$ and we can extend an environment $\Gamma$ with a variable $x$ of type $A$ (written $\Gamma, x: A$).
-Often the empty environment is ommited.
+Often the empty environment is omitted.
 For example the environment $x: \mathsf{int}, y: \mathsf{bool}$ has a variable $x$ of type \textsf{int} and a variable $y$ of type \textsf{bool}.
@@
     \begin{itemize}
-        \item A teminal object $\term \in \cat{C}$: $\forall X \in \cat{C}.\, \exists! \langle\rangle: X \to \term$.
+        \item A terminal object $\term \in \cat{C}$: $\forall X \in \cat{C}.\, \exists! \langle\rangle: X \to \term$.
         \item Binary products $X \times Y \in \cat{C}$ with projections $\pi_1 \from X \times Y \to X$ and $\pi_2 \from X \times Y \to Y$
@@
 
-%The case for constants 
-\todo{explain, esp. $M(c^A)$}
+The case for constants illustrates why we choose $M(c^A) \from \term \to M|[A|]$.
+Since the type of a constant does not depend on the environment $\Gamma$, it is natural to take the empty environment.
+We can easily construct morphisms for arbitrary environments, because the empty environment is mapped to the terminal object.
 
@@
 A substitution corresponds to a function application $(\lambda x: A.\, s)\, t$, where the function is called with an argument $t$.
-It replaces in the term $s$ all occurences of the variable $x$ with the term $t$.
+It replaces in the term $s$ all occurrences of the variable $x$ with the term $t$.
 We write $s[t/x]$ and define it as follows.

### 2017-09-20 papers/talk-12/paper-12.tex
Add writeup for talk 12

\documentclass{article}
  \def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
  \usepackage{microtype}

  \graphicspath{{graphics/}}

  \usepackage{mathpartir}
  \theoremstyle{plain}
  \newtheorem{lemma}{Lemma}
  \newtheorem{theorem}{Theorem}
  \theoremstyle{definition}
  \newtheorem{definition}{Definition}

  \usetikzlibrary{fit}
  \usetikzlibrary{knots}

  \tikzstyle{morphism}=[draw, minimum size=0.5cm, semithick]
  \tikzstyle{phantom morphism}=[minimum size=0.5cm]
  \tikzstyle{object}=[semithick]

  \begin{document}
  \title{The Graphical Calculus for Monoidal Categories}
  \author{Felix Rech}
  \maketitle

  In this chapter, we will talk about a certain class of categories that have a binary operation on objects.
  We call them monoidal categories because they combine the concept of a category and that of a monoid.
  Our central goal is to introduce a \emph{graphical calculus} for the construction of morphisms in such categories.
  This calculus hides non-essential details of constructions.
  Hence it is easy to understand what a morphism does and to decide whether two morphisms are equal or not.
  To illustrate this, assume that we have four morphisms $B
  \xrightarrow{f, g} B$, $C \xrightarrow{h} B$ and $A \xrightarrow{k} B \times C$ in the category of sets.
  Can you tell if the following equation is valid?
  \[ \rounddd{f \times (g \of h)} \of k = (1 \times f) \of \rounddd{(g \times h) \of k} \]
  The answer is that in general the equation does not hold.
  But at first glance, this is not obvious, although both constructions are rather simple.
  Now take a look at this graphical representation of the equation:
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
          \node[phantom morphism] (out1) at (0, 4) {};
          \node[morphism] (f) at (0, 2.5) {$f$};
          \node[phantom morphism] (out2) at (1, 4) {};
          \node[morphism] (g) at (1, 2.9) {$g$};
          \node[morphism] (h) at (1, 2.1) {$h$};
          \node[phantom morphism] (k_left) at (0, 1) {};
          \node[phantom morphism] (k_right) at (1, 1) {};
          \node[morphism, fit=(k_left) (k_right), inner sep=0, ] (k) {};
          \node at (k) {$k$};
          \node (in) at (0.5, 0) {};

          \draw[object] (f) to (out1);
          \draw[object] (k_left) to (f);
          \draw[object] (g) to (out2);
          \draw[object] (h) to (g);
          \draw[object] (k_right) to (h);
          \draw[object] (in) to (k);
        \end{tikzpicture}}}
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
          \node[phantom morphism] (out1) at (0, 4) {};
          \node[morphism] (g) at (0, 2) {$g$};
          \node[phantom morphism] (out2) at (1, 4) {};
          \node[morphism] (f) at (1, 3) {$f$};
          \node[morphism] (h) at (1, 2) {$h$};
          \node[phantom morphism] (k_left) at (0, 1) {};
          \node[phantom morphism] (k_right) at (1, 1) {};
          \node[morphism, fit=(k_left) (k_right), inner sep=0, ] (k) {};
          \node at (k) {$k$};
          \node (in) at (0.5, 0) {};

          \draw[object] (g) to (out1);
          \draw[object] (k_left) to (g);
          \draw[object] (f) to (out2);
          \draw[object] (h) to (f);
          \draw[object] (k_right) to (h);
          \draw[object] (in) to (k);
        \end{tikzpicture}}}.
  \]
  Each side of the equation can be read like the drawing of a circuit with outputs at the top and inputs at the bottom.
  Wires and boxes that run in parallel, stand for the Cartesian product of objects and morphisms respectively.
  It is okay if you do not immediately understand how this notation works.
  It will become clear over the next few pages.
  The important message is that circuit-like drawings of morphisms are often easier to understand and compare than algebraic expressions.

  In our example above we used the category of sets with the Cartesian product as the binary operation.
  For the circuit metaphor, this means that inputs and outputs of the circuit consist of signals on all input or output wires simultaneously.
  We could also use the disjoint union as operation, in which case inputs and outputs would consist of a single signal on one of those wires.
  Another option is to use the category of relations with the Cartesian product.
  If we interpret relations as non-deterministic functions then this yields non-deterministic circuits.
  In fact, we could do this with most categories and binary operations that we normally encounter.
  In the end of this chapter, you will see that circuits are just the start and that we can extend or generalize the calculus to represent more concepts in this powerful way.

  This chapter is based on unpublished lecture notes about categorical quantum mechanics by Daniel Marsden and Jamie Vicary from 2016.
  For more details, background and omitted proofs, we refer to those lecture notes or an older version by Chris Heunen and Jamie Vicary~\cite{heunen2015cqm}.

  \section{Graphical Calculus for General Categories}

  We will now start with the introduction of those components of the graphical calculus that can be applied for all categories.
  We draw objects as vertical lines that we also refer to as wires and morphisms as boxes.
  If there is a morphism $A \xrightarrow{f} B$ then we draw it as a box with an output wire at the top and an input wire at the bottom:
  \[ \includegraphics{morphism}. \]
  Often we omit labels on object and morphisms if they are not essential.

  For two morphisms $A \xrightarrow{f} B$ and $B \xrightarrow{g} C$, we use the output wire of the $f$-box as input wire to the $g$-box to denote the composition $g \of f$:
  \[ \includegraphics{vertical_composition}. \]

  So far this is basically the same as the algebraic notation that we are used to, except that we read it top to bottom instead of left to right.
  In the same way as we can do it with the algebraic notation, we usually omit parentheses and do not draw identity morphisms.

  \section{Monoidal Categories}

  We have defined what it means to connect multiple morphisms in a chain, but how do we combine morphisms in parallel?
  As mentioned earlier, we need a binary operation on objects for that which should also operate on morphisms.
  Hence it seems reasonable to require it to be a functor.
  We will call it tensor product because the tensor product of vector spaces is a prominent example.
  For our purposes, we need some minimal structure on such a vector product.
  Inspired by the definition of a monoid, we require associativity and a neutral element that we call unit.
  This leads us to the definition of a strict monoidal category.

  \begin{definition}
    A \demph{strict monoidal category} is a category $C$ with the following:
    \begin{description}
    \item[Tensor product:] A functor $\otimes : C \times C \to C$
    \item[Unit object:] An object $I \in \ob(C)$
    \item[Associativity:] An equality $(A \otimes B) \otimes C = A \otimes (B \otimes C)$ for all objects $A$, $B$ and $C$
    \item[Left unit law:] An equality $I \otimes A = A$ for all objects $A$
    \item[Right unit law:] An equality $A \otimes I = A$ for all objects $A$
    \end{description}
  \end{definition}

  Our original example, that served as motivation for this definition, was the category of sets with the Cartesian product as the tensor product.
  If we check the details, however, we see that it does not actually satisfy the definition because associativity and the unit laws hold only up to isomorphism.
  To admit this, we define monoidal categories as a generalization of strict monoidal categories.

  \begin{definition}
    A \demph{monoidal category} is a category $C$ with the following:
    \begin{description}
    \item[Tensor product:] A functor $\otimes : C \times C \to C$
    \item[Unit object:] An object $I \in \ob(C)$
    \item[Associator:] A natural isomorphism $(A \otimes B) \otimes C \xrightarrow{\smash{\alpha_{A, B, C}}} A \otimes (B \otimes C)$ for all objects $A$, $B$ and $C$
    \item[Left unitor:] A natural isomorphism $I \otimes A \xrightarrow{\smash{\lambda_A}} A$ for all objects $A$
    \item[Right unitor:] A natural isomorphism $A \otimes I \xrightarrow{\rho_A} A$ for all objects $A$
    \end{description}

    Those have to satisfy the \demph{triangle} and \demph{pentagon equalities}:
    \[ \begin{adjustbox}{max totalsize={\textwidth}{\textheight}, center}
      \begin{tikzcd}[column sep=0]
        (A \otimes I) \otimes B
        \arrow[dr, "\rho_A \otimes 1_B"']
        \arrow[rr, "\alpha_{A, I, B}"] &&
        A \otimes (I \otimes B)
        \arrow[dl, "1_A \otimes \lambda_B"] \\
        %
        & A \otimes B
      \end{tikzcd}
      \hspace{1em}
      \begin{tikzpicture}[commutative diagrams/every diagram]
        \node (S) at (0, -2)
        {$(A \otimes B) \otimes (C \otimes D)$};
        \node (E) at (3, 0)
        {$A \otimes (B \otimes (C \otimes D))$};
        \node (NE) at (2.5, 2)
        {$A \otimes ((B \otimes C) \otimes D)$};
        \node (NW) at (-2.5, 2)
        {$(A \otimes (B \otimes C)) \otimes D$};
        \node (W) at (-3, 0)
        {$((A \otimes B) \otimes C) \otimes D$};

        \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C} \otimes 1_D$} (W)
        (W) edge node[anchor=north east] {$\alpha_{A \otimes B, C, D}$} (S)
        (S) edge node[anchor=north west] {$\alpha_{A, B, C \otimes D}$} (E)
        (E) edge node[anchor=west] {$1_A \otimes \alpha_{B, C, D}$} (NE)
        (NW) edge node {$\alpha_{A, B \otimes C, D}$} (NE);
      \end{tikzpicture}
    \end{adjustbox}. \]
  \end{definition}

  It is not necessary that you understand the triangle and pentagon equalities.
  The only reason why we have them is to obtain the following coherence property.

  \begin{theorem}[Coherence for monoidal categories]
    Given the data of a monoidal category, if the pentagon and triangle equations hold, then any well-typed equation built only from $\lambda$, $\rho$, $\alpha$ and their inverses holds.
  \end{theorem}

  An example for the coherence property is the equation $\lambda_I = \rho_I$ which would not follow without the triangle and pentagon equalities.
  Other examples are the triangle and pentagon equations themselves.
  Of course, the statement of the theorem is imprecise because we didn't define what it means for an equation to be well-typed.
  For a counterexample suppose that there is an object $A$ that satisfies the equality $A = (A \otimes A) \otimes A$.
  Then we have morphisms $1_A : A \to A$ and $\alpha_{A, A, A} : A \to A$ and we can write down the equation $1_A = \alpha_{A, A, A}$.
  But this equation is not well-typed in our terms because it uses the extra knowledge about $A$.
  Roughly speaking, an equation is well-typed if it makes sense in every monoidal category and for every choice of objects.

  \section{Graphical Calculus for Monoidal Categories}

  From now on we will talk about monoidal categories and extend our graphical calculus to express the tensor product of morphisms as parallel combination.
  If we have two morphisms $f : A \to B$ and $g : C \to D$, then we write them side-by-side in the graphical representation to denote the tensor product $f \otimes g$:
  \[ \includegraphics{horizontal_composition}. \]
  Usually, we do not draw the unit object or any parentheses for this parallel combination, which is justified by associativity, the unit laws and the following interchange law.

  \begin{theorem}
    In a monoidal category all morphisms $f : A \to B$, $g : B \to C$, $h : D \to E$ and $j : E \to F$ satisfy the equation
    \[ (g \of f) \otimes (j \of h) = (g \otimes j) \of (f \otimes h). \]
    In the graphical calculus this can be written as
    \[ \includegraphics{interchange_law}. \]
  \end{theorem}

  We do not draw $\lambda$, $\rho$ and $\alpha$ either:
  \begin{mathpar}
    \includegraphics{lambda} \and
    \includegraphics{rho} \and
    \includegraphics{alpha}.
  \end{mathpar}
  This can be justified by the their naturality and the coherence law.
  The actual proof is complex but you should get a feeling for it if you consider the following properties.
  \begin{itemize}
  \item Naturality allows us to move other morphisms \textit{through} $\lambda$, $\rho$ and $\alpha$.
    Therefore the exact position where we apply those isomorphisms does not matter.
  \item The coherence theorem tells us that the exact selection and combination of those isomorphisms does not matter as long as we fix the input and the output type.
  \end{itemize}

  We have given some hints, why the graphical calculus makes sense.
  The next theorem makes this precise and summarizes everything you need to know about the graphical calculus.

  \begin{definition}
    Two diagrams are \demph{isotopic} when one can be deformed continuously into the other, keeping the boundaries fixed.
  \end{definition}

  This means essentially that we can move things around in one diagram to obtain the other whereby we are not allowed to move one object through another, to move objects outside of the boundary or to remove inputs or outputs from the boundary.

  \begin{theorem}[Correctness for monoidal categories]
    A well-typed equation between morphisms in a monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{planar} isotopy.
  \end{theorem}

  To make this clear, let us first look at a counterexample:
  \[ \includegraphics{simple_counterexample}. \]
  Those two diagrams do not in general represent the same morphism.
  To check if the diagrams are isotopic, take a look at the following schematic representation that contains the boundaries and represents $f$ as a geometric point.
  \[ \includegraphics{simple_counterexample_geometric} \]
  In the plane, we cannot move $f$ from the left half of the diagram to the right half without going through $A$ or leaving the boundary.
  Thus the diagrams are not isotopic and the morphisms are not in general equal.

  On the other hand, the following diagrams are obviously isotopic and thus represent the same morphisms.

  \[ \includegraphics{simple_example} \]

  \section{Braided Monoidal Categories}

  So far we followed the goal to interpret constructions of circuits as constructions of morphisms in categories and we have nearly achieved that goal, but there is one thing that we can do in circuits but in general not with morphisms in monoidal categories: we cannot \textit{cross wires}:
  There is no general way to construct a morphism $A \otimes B \to B \otimes A$ for arbitrary objects $A$ and $B$.
  We describe monoidal categories that allow us to do this by the following definition.

  \begin{definition}
    A \demph{braided monoidal category} is a monoidal category equipped with a natural isomorphism
    \[ A \otimes B \xrightarrow{\sigma_{A, B}} B \otimes A \]
    satisfying the \demph{hexagon equations}:
    \[ \begin{adjustbox}{max totalsize={\textwidth}{\textheight}, center}
      \begin{tikzpicture}[commutative diagrams/every diagram]
        \node (SW) at (-1.75, -1.5)
        {$(B \otimes A) \otimes C$};
        \node (SE) at (1.75, -1.5)
        {$B \otimes (A \otimes C)$};
        \node (E) at (2.5, 0)
        {$B \otimes (C \otimes A)$};
        \node (NE) at (1.75, 1.5)
        {$(B \otimes C) \otimes A$};
        \node (NW) at (-1.75, 1.5)
        {$A \otimes (B \otimes C)$};
        \node (W) at (-2.5, 0)
        {$(A \otimes B) \otimes C$};

        \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C}^{-1}$} (W)
        (W) edge node[anchor=east] {$\sigma{A, B} \otimes 1_C$} (SW)
        (SW) edge node[anchor=north] {$\alpha_{B, A, C}$} (SE)
        (SE) edge node[anchor=west] {$1_B \otimes \sigma_{A, C}$} (E)
        (E) edge node[anchor=west] {$\alpha_{B, C, A}^{-1}$} (NE)
        (NW) edge node {$\sigma_{A, B \otimes C}$} (NE);
      \end{tikzpicture}
      \hspace{1em}
      \begin{tikzpicture}[commutative diagrams/every diagram]
        \node (SW) at (-1.75, -1.5)
        {$A \otimes (C \otimes B)$};
        \node (SE) at (1.75, -1.5)
        {$(A \otimes C) \otimes B$};
        \node (E) at (2.5, 0)
        {$(C \otimes A) \otimes B$};
        \node (NE) at (1.75, 1.5)
        {$C \otimes (A \otimes B)$};
        \node (NW) at (-1.75, 1.5)
        {$(A \otimes B) \otimes C$};
        \node (W) at (-2.5, 0)
        {$A \otimes (B \otimes C)$};

        \draw[commutative diagrams/.cd, every arrow, every label]
        (NW) edge node[anchor=east] {$\alpha_{A, B, C}$} (W)
        (W) edge node[anchor=east] {$1_A \otimes \sigma_{B, C}$} (SW)
        (SW) edge node[anchor=north] {$\alpha_{A, C, B}^{-1}$} (SE)
        (SE) edge node[anchor=west] {$\sigma_{A, C} \otimes 1_B$} (E)
        (E) edge node[anchor=west] {$\alpha_{C, A, B}$} (NE)
        (NW) edge node {$\sigma_{A \otimes B, C}$} (NE);
      \end{tikzpicture}
    \end{adjustbox}. \]
  \end{definition}

  Similar to the triangle and pentagon equations, the hexagon equations have the only purpose to ensure a coherence property that includes not only $\lambda$, $\rho$ and $\alpha$ but also $\sigma$.
  But this time the coherence statement is more complicated.
  For example the equation $1_{A \otimes B} = \sigma_{B, A} \of \sigma_{A, B}$ does not hold in general, even though it is well-typed.
  Again, the coherence property will be important for the correctness of the graphical calculus, but we will not state it here.

  We extend the graphical calculus with simple notations for $\sigma$ and its inverse that reflect our intuition of crossing wires:
  \begin{mathpar}
    \overset{\includegraphics{sigma}}{\sigma} \and
    \overset{\includegraphics{sigma_inverse}}{\sigma^{-1}}.
  \end{mathpar}

  Naturality of $\sigma$ allows to move morphisms through crossings:
  \begin{mathpar}
    \includegraphics{sigma_naturality} \and
    \includegraphics{sigma_inverse_naturality}.
  \end{mathpar}

  The fact that $\sigma^{-1}$ is inverse to $\sigma$, means that we can move wires over another:
  \begin{mathpar}
    \includegraphics{sigma_inverse_sigma} \and
    \includegraphics{sigma_sigma_inverse}.
  \end{mathpar}

  As you might guess, this together allows us to formulate a new correctness statement that does not consider the graphical representation in the plane but in three-dimensional space.

  \begin{theorem}[Correctness for braided monoidal categories]
    A well-typed equation between morphisms in a braided monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{spatial} isotopy.
  \end{theorem}

  As examples consider the following two pairs of equations between diagrams with their algebraic counterpart.

  \[ \includegraphics{braiding_example_1} \]
  \[ \alpha \of (\sigma \otimes 1) \of \alpha^{-1} \of (1 \otimes f) \of \rho^{-1}
    = (1 \otimes \sigma^{-1}) \of \alpha \of (f \otimes 1) \of \lambda^{-1} \]

  \[ \includegraphics{braiding_example_2} \]
  \[ \alpha \of (\sigma \otimes 1) \of \alpha^{-1} \of (1 \otimes \sigma) \of \alpha \of (\sigma \otimes 1)
    = (1 \otimes \sigma) \of \alpha \of (\sigma \otimes 1) \of \alpha^{-1} \of (1 \otimes \sigma) \of \alpha \]

  With our geometric intuition it is obvious that the compared diagrams are isotopic in three-dimensional space and thus the equations hold.
  If you look at the algebraic form however, it becomes much harder to see.

  \section{Symmetric Monoidal Categories}

  You may have noticed that there is one last thing that distinguishes circuits from our constructions in braided monoidal categories.
  In circuits, it does not matter which wire crosses on top.
  In the language of braided monoidal categories this means that $\sigma_{B, A}^{-1} = \sigma_{A, B}$.

  \begin{definition}
    A braided monoidal category is \demph{symmetric} when
    \[ \sigma_{B, A}^{-1} = \sigma_{A, B} \]
    for all objects $A$ and $B$.
  \end{definition}

  Since we do not need to distinguish between $\sigma$ and $\sigma^{-1}$ anymore, we denote both by the same notation:
  \[ \includegraphics{symmetric_sigma}. \]

  The most important consequence of symmetry is that wires can move through each other:
  \[ \includegraphics{symmetry}. \]
  This means that for the correctness statement we can consider four-dimensional isotopy now.

  \begin{theorem}[Correctness for symmetric monoidal categories]
    A well-typed equation between morphisms in a symmetric monoidal category follows from the axioms if and only if it holds in the graphical language up to \textbf{four-dimensional} isotopy.
  \end{theorem}

  \section{Dual Objects}

  In this section, we go beyond the circuit metaphor and demonstrate the extensibility and power of the graphical calculus with a new notation for so-called dual objects.

  \begin{definition}
    In a monoidal category, an object $L$ is \demph{left-dual} to an object $R$, and $R$ is \demph{right-dual} to $L$, written $L \dashv R$, when there exist
    \begin{itemize}
    \item a \demph{unit} morphism $I \xrightarrow{\eta} R \otimes L$ and
    \item a \demph{counit} morphism $L \otimes R \xrightarrow{\varepsilon} I$
    \end{itemize}
    satisfying the \demph{snake equations}:
    \[ \begin{adjustbox}{max width=\linewidth, center}
      \begin{tikzcd}
        L \arrow[r, "\rho_L^{-1}"] \arrow[d, "1_L"] &
        L \otimes I \arrow[r, "1_L \otimes \eta"] &
        L \otimes (R \otimes L) \arrow[d, "\alpha_{L, R, L}^{-1}"] \\
        %
        L &
        I \otimes L \arrow[l, "\lambda_L"] &
        (L \otimes R) \otimes L \arrow[l, "\varepsilon \otimes 1_L"]
      \end{tikzcd}
      \hspace{1em}
      \begin{tikzcd}
        R \arrow[r, "\lambda_R^{-1}"] \arrow[d, "1_R"] &
        I \otimes R \arrow[r, "\eta \otimes 1_R"] &
        (R \otimes L) \otimes R \arrow[d, "\alpha_{R, L, R}"] \\
        %
        R &
        R \otimes I \arrow[l, "\rho_R"] &
        R \otimes (L \otimes R) \arrow[l, "1_R \otimes \varepsilon"]
      \end{tikzcd}.
    \end{adjustbox} \]
  \end{definition}

  To make it obvious if we talk about a left or right dual, we indicate this by an upward or downward arrow respectively:
  \begin{mathpar}
    \includegraphics{left_dual} \and
    \includegraphics{right_dual}.
  \end{mathpar}

  For the unit and counit we introduce a special notation:
  \begin{mathpar}
    \overset{\includegraphics{unit}}{\lower 3ex \hbox{\centering unit}} \and
    \overset{\includegraphics{counit}}{\lower 3ex \hbox{\centering counit}}.
  \end{mathpar}
  We say that we can bend wires now.
  With this notation the snake equations take a very simple form:
  \[ \includegraphics{snake_equations}. \]
  They let us straighten arbitrarily bent wires.

  We will now present three proofs of basic properties of dual objects to demonstrate how the graphical calculus can make things simple and clear.

  \begin{lemma} 
    The left- and right-dual of an object are unique up to isomorphism.
  \end{lemma}
  \begin{proof}
    We show only that the right-dual is unique.
    Assume that $R$ and $R'$ are both right-dual to $L$.
    For convenience, we use the bent-wire notation for both.
    Then we can define morphisms between $R$ and $R'$ by
    \[ \includegraphics{dual_unique}. \]
    From the snake equations, it is easy to see that those are mutually inverse.
    Thus $R' \simeq R$.
  \end{proof}

  \begin{lemma}
    The unit of a duality determines the counit and vice-versa.
  \end{lemma}
  \begin{proof}
    We show only that the unit determines the counit.
    The other direction is similar.
    Assume that we have a duality $L \dashv R$ with unit $\eta$ and counit $\varepsilon$ and we have another duality $L \dashv R$, also with unit $\eta$ but with counit $\varepsilon'$.
    Then we have the equality
    \[
      \vcenter{ \hbox{ \begin{tikzpicture}
        \node[phantom morphism] (epsilon_left) at (0, 1) {};
        \node[phantom morphism] (epsilon_right) at (1, 1) {};
        \node[morphism, fit=(epsilon_left) (epsilon_right), inner sep=0] (epsilon) {};
        \node at (epsilon) {$\varepsilon$};

        \coordinate (in1) at (0, -0.5);
        \coordinate (in2) at (1, -0.5);
        \draw[object] (in1) to (epsilon_left);
        \draw[object] (in2) to (epsilon_right);
      \end{tikzpicture} } }
      =
      \vcenter{ \hbox{ \begin{tikzpicture}
        \node[phantom morphism] (epsilon_left) at (0, 1) {};
        \node[phantom morphism] (epsilon_right) at (1, 1) {};
        \node[morphism, fit=(epsilon_left) (epsilon_right), inner sep=0] (epsilon) {};
        \node at (epsilon) {$\varepsilon$};

        \node[phantom morphism] (eta_left) at (1, 0) {};
        \node[phantom morphism] (eta_right) at (2, 0) {};
        \node[morphism, fit=(eta_left) (eta_right), inner sep=0] (eta) {};
        \node at (eta) {$\eta$};

        \node[phantom morphism] (epsilon'_left) at (2, 1) {};
        \node[phantom morphism] (epsilon'_right) at (3, 1) {};
        \node[morphism, fit=(epsilon'_left) (epsilon'_right), inner sep=0] (epsilon') {};
        \node at (epsilon') {$\varepsilon'$};

        \coordinate (in1) at (0, -0.5);
        \coordinate (in2) at (3, -0.5);
        \draw[object] (in1) to (epsilon_left);
        \draw[object] (eta_left) to (epsilon_right);
        \draw[object] (eta_right) to (epsilon'_left);
        \draw[object] (in2) to (epsilon'_right);
      \end{tikzpicture} } }
      =
      \vcenter{ \hbox{ \begin{tikzpicture}
        \node[phantom morphism] (epsilon'_left) at (2, 1) {};
        \node[phantom morphism] (epsilon'_right) at (3, 1) {};
        \node[morphism, fit=(epsilon'_left) (epsilon'_right), inner sep=0] (epsilon') {};
        \node at (epsilon') {$\varepsilon'$};

        \coordinate (in1) at (2, -0.5);
        \coordinate (in2) at (3, -0.5);
        \draw[object] (in1) to (epsilon'_left);
        \draw[object] (in2) to (epsilon'_right);
      \end{tikzpicture} } }
    \]
    where we apply one snake equation for $\varepsilon'$ and one snake equation for $\varepsilon$.
  \end{proof}

  \begin{lemma}
    In a braided monoidal category $L \dashv R$ implies $R \dashv L$.
  \end{lemma}
  \begin{proof}
    Assume that $L \dashv R$.
    We define a unit and a counit for $R \dashv L$ by
    \[ \includegraphics{duality_symmetric_unit_and_counit}. \]
    The first snake equation follows by
    \[ \includegraphics{duality_symmetric_snake} \]
    where we apply the correctness theorem for braided monoidal categories in the first step and a snake equation for $L \dashv R$ in the second step.
    The other snake equation follows in the same way.
  \end{proof}

  \section{Outlook}
  There is some obvious similarity between our definition of duality and the definition of an adjunction.
  This is because on the right abstraction level both are the same.
  To explain this, we will briefly introduce an application of the graphical calculus to something other than monoidal categories.
  The most notable difference that you should be aware of before we start, is that we will not only assign meaning to boxes and wires, but to the space between wires too.
  An empty space stands for a category, a wire between spaces stands for a functor and a box between wires stands for a natural transformation.
  In this setting, the definition of adjunction through unit and counit takes exactly the same form as the definition of duality above where the triangle identities correspond to the snake equations.
  The common abstraction between both cases that makes this possible is that of a bicategory.
  Hopefully, this convinces you that the graphical calculus has far more potential than we have seen and that the circuit metaphor was really just the start.

  \bibliographystyle{plain}
  \bibliography{paper-12}

  \end{document}

### 2017-09-21 papers/talk-11/paper-11.tex
Rewrite introduction to dependent types

@@
 \usepackage{semantic}
+\usepackage{mathpartir}
 
@@
 
-Dependent types are types that unlike simple types like functions and products
-may depend on or vary with values.  A classical example of a dependent type is
-a type of lists $List_{A}(N)$ of a given length $N : \NN$ over some
-type $A$. The elements of such type are $Nil_{A} : List(0)$ and
-$Cons_{A}(x, l) : List(N+1)$, where $x : {A}$ and $l : List_{A}(N)$.
-Now let us consider a function that takes $n : \NN$ and returns a list of
-length $n$ containing $0$ in all entries.  This function would have the type
-$\Pi n : \NN.List_\NN(n)$, a type of functions for which the type
-of the return value depends on the argument.  The important point of the
-dependent typing is that it reveals more information about the function.
-Another example is an exception-free head function that returns the first
-element of the list:
+Suppose you want to write a function that returns the head of a given list of
+natural numbers, using the classical definition of the list type:
 \[
-  hd: \Pi x : \mathbb{N}.List_{A}(Suc(x)) \to A
+  List = Nil : Unit \to List~|~Cons : \NN \times List \to List
 \]
-The type of this function prevents it from being applied to an empty list,
-showing clearly that the function is defined only for non empty lists.
+Writing the case for $Cons$ is easy --- just take the first element of the
+pair.  But what about the case of an empty list? This case is tricky because it
+is not clear what should the result be. Obviously, such function (if we do not
+allow definition of partial functions which is the case in many programming
+languages) cannot be well defined on the empty list without taking arbitrary
+decisions. But maybe such problems can be prevent from happening at all?
+
+A solution to this issue is an instance for the application of dependent types.
+Dependent types are such types that include simple types, like products and
+functions, that we have seen in simply-typed lambda calculus (STLC), but also
+add a possibility to define types using values. With dependent types we would
+be able to define a type for lists that includes the size of the list:
+\[
+  List(\NN) = Nil : Unit \to List(0)~|~Cons : \NN \times List(n) \to List(Suc(n))
+\]
+\begin{example}
+  The function to construct a list of length $n$, having only $0$ as elements
+  would then have the type $\Pi n: \NN.List(n)$, indicating that it accepts a
+  natural number and returns a value of a type constructed using that number.
+\end{example}
+\begin{example}
+  Our desired head function would have the type
+  \[
+    Hd: \Pi n: \NN.List(Suc(n)) \to \NN
+  \]
+  indicating that it cannot be applied to the empty list at all.
+\end{example}
+
+The important part is that the dependent typing reveals more information about
+the function.
 
@@
 
-To formalize dependent types we build a system that allows us to make
-judgements of the form $a : A$ (term $a$ has type $A$) and $A~type$ ($A$ is a
-type). We will also want a notion of definitional equality to be built into the
-system, for example we want to consider $0 : \NN$ and $0 + 0 : \NN$ to be
-(definitionally) equal terms and $List_A(0)$, $List_A(0 + 0)$ as
-(definitionally) equal types: if $l : List_A(0)$ then also $l : List_A(0 + 0)$.
-This leads to two more forms of the judgements: $a = b : A$ ($a$ and $b$ are
-definitionally equal terms of type $A$) and $A = B~type$ ($A$ and $B$ are
-definitionally equal types). Finally, we must know types of the free variables
-occurring in a judgement: we cannot assert $x + y : \NN$ unless we know $x :
-\NN$ and $y : \NN$. Since such declarations may depend on each other like in $x
-: \NN$, $y : List_\NN(x)$ it is convenient to make all judgements relative to a
-list of variable declarations including at least the free variables occurring
-inside the judgement. Such lists are called contexts and sometimes also type
-assignments.
+To formalize dependent types we will use a notion of typing context (which is
+almost the same as typing environment as seen for STLC, with the exception that
+it holds dependent types).
 
-Intuitively, a context $x_1 : A_1, \dots, x_n : A_n$ is well-formed if each
-$A_i$ is a type in the context $x_1 : A_1, \dots, x_{n-1} : A_{n-1}$ and all
-the variables $x_i$ are pairwise distinct. Thus we arrive to six kinds of
-judgements:
+\begin{definition}[Context]
+  A \emph{context} is a list of bindings that connect variables to types: $x_1: A_1,
+  \dots, x_n: A_n$. A context is called well-formed if every variable in it is
+  distinct from the other variables and each $A_i$ is a type in the context $x_1
+  : A_1, \dots, x_{n-1} : A_{n-1}$, since our types may now include the variables
+  from the context.
+\end{definition}
+
+Note that we extend the notion of a well-formed context from STLC --- in STLC
+we only needed the variables to be distinct, since they had no influence on the
+types.
+
+Having contexts, we want to work with the following judgements: $a : A$ (term
+$a$ has type $A$) and $A~type$ ($A$ is a type) and consider definitional
+equality to be built-in into our formal system. That is, we want the types
+$List(0)$ and $List(0 + 0)$ to be equal. So we also include the judgements
+about the equality of terms $a = b : A$, types $A = B~type$, and contexts
+$\Gamma = \Delta~ctxt$.
+
+To put it together, we have six types of judgements:
 
@@
 
-We also have the following structural rules:
-\begin{itemize}
-  \item The empty context construction, context extension and context equality
-    rules
-    \[
-      \inference[C-Emp]{}{\vdash \diamond~ctxt}
-      \qquad
-      \inference[C-Ext]{\Gamma \vdash A~type}{\vdash \Gamma, x : A~ctxt}
-    \]
-    \[
-      \inference[C-Ext-Eq]{\vdash \Gamma = \Delta~ctxt \qquad \Gamma \vdash A = B~type}{\vdash \Gamma, x:A = \Delta, y:B~ctxt}
-    \]
-    where $x$ and $y$ are assumed to be fresh variables.
-  \item The variable rule
-    \[
-      \inference[Var]{\vdash \Gamma, x:A,\Delta~ctxt}{\Gamma,x:A,\Delta \vdash x:A}
-    \]
-  \item Rules expressing that definitional equality is an equivalence relation:
-    \[
-      \inference[C-Eq-R]{\vdash \Gamma~ctxt}{\vdash \Gamma = \Gamma~ctxt}
-    \]
-    \[
-      \inference[C-Eq-S]{\vdash \Gamma = \Delta~ctxt}{\vdash \Delta = \Gamma~ctxt}
-    \]
-    \[
-      \inference[C-Eq-T]{\vdash \Gamma = \Delta~ctxt \qquad \vdash \Delta = \Theta~ctxt}{\vdash \Gamma = \Theta~ctxt}
-    \]
-    with analogous rules for terms and types (omitted).
-  \item Rules relating typing and definitional equality:
-    \[
-      \inference[Tm-Conv]{\Gamma \vdash a : A \qquad \vdash \Gamma = \Delta~ctxt \qquad \Gamma \vdash A = B~type}{\Delta \vdash a : B}
-    \]
-    \[
-      \inference[Ty-Conv]{\vdash \Gamma = \Delta~ctxt \qquad \Gamma \vdash A~type}{\Delta \vdash A~type}
-    \]
-  \item And rules for weakening and substitution, where $\mathcal{J}$ ranges over judgements of the form $a : A, A~type, a = b: A, A = B~type$:
-    \[
-      \inference[Weak]{\Gamma, \Delta \vdash \mathcal{J} \qquad \Gamma \vdash C~type}{\Gamma, x : C, \Delta \vdash \mathcal{J}}
-    \]
-    \[
-      \inference[Subst]{\Gamma, x: C, \Delta \vdash \mathcal{J} \qquad \Gamma \vdash c: C}{\Gamma, \Delta[c/x] \vdash \mathcal{J}[c/x]}
-    \]
-    Here $\mathcal{J}[c/x]$ (and $\Delta[c/x]$) denotes the capture-free
-    substitution of $x$ for $c$ in $\mathcal{J}$. This means that bound
-    variables in $\mathcal{J}$ are systematically renamed to prevent any free
-    variables in $c$ from becoming bound in $\mathcal{J}[c/x]$. The contexts,
-    types, and terms will be considered equal if they agree up to names of
-    bound variables and we will assume existence of capture-free substitution
-    function on these equivalence classes.
-\end{itemize}
+We also need the structural rules:
 
-\section*{Type formers}
+\begin{definition}[Structural rules]
+\begin{mathpar}
+  \inferrule*[right=C-Emp]{ }{\vdash \diamond~ctxt}
+  \and
+  \inferrule*[right=C-Ext]{\Gamma \vdash A~type}{\vdash \Gamma, x : A~ctxt}
+  \and
+  \inferrule*[right=C-Ext-Eq]{\vdash \Gamma = \Delta~ctxt \\ \Gamma \vdash A = B~type}{\vdash \Gamma, x:A = \Delta, y:B~ctxt}
+  \and
+  \inferrule*[right=Var]{\vdash \Gamma, x:A,\Delta~ctxt}{\Gamma,x:A,\Delta \vdash x:A}
+  \and
+  \inferrule*[right=C-Eq-R]{\vdash \Gamma~ctxt}{\vdash \Gamma = \Gamma~ctxt}
+  \and
+  \inferrule*[right=C-Eq-S]{\vdash \Gamma = \Delta~ctxt}{\vdash \Delta = \Gamma~ctxt}
+  \and
+  \inferrule*[right=C-Eq-T]{\vdash \Gamma = \Delta~ctxt \\ \vdash \Delta = \Theta~ctxt}{\vdash \Gamma = \Theta~ctxt}
+  \and
+  \inferrule*[right=Tm-Conv]{\Gamma \vdash a : A \\ \vdash \Gamma = \Delta~ctxt \\ \Gamma \vdash A = B~type}{\Delta \vdash a : B}
+  \and
+  \inferrule*[right=Ty-Conv]{\vdash \Gamma = \Delta~ctxt \\ \Gamma \vdash A~type}{\Delta \vdash A~type}
+  \and
+  \inferrule*[right=Weak]{\Gamma, \Delta \vdash \mathcal{J} \\ \Gamma \vdash C~type}{\Gamma, x : C, \Delta \vdash \mathcal{J}}
+  \and
+  \inferrule*[right=Subst]{\Gamma, x: C, \Delta \vdash \mathcal{J} \qquad \Gamma \vdash c: C}{\Gamma, \Delta[c/x] \vdash \mathcal{J}[c/x]}
+\end{mathpar}
 
-Type and term formers are introduced by formation, introduction, elimination,
-and equality rules. We present several type formers to give an example of how
-they may be constructed.
+where $x$ and $y$ in C-Ext and C-Ext-Eq are assumed to be fresh variables and
+$\mathcal{J}$ in Weak and Subst ranges over judgements of the form $a : A,
+A~type, a = b: A, A = B~type$. $[c/x]$ denotes the same substitution as seen in
+STLC, except that it now also applies to the variables that may be contained in
+types.
+\end{definition}
+
+The rules C-Emp and C-Ext give us the empty context ($\diamond$) and allow us
+to extend the context (analogous to the environments from STLC).
+
+C-Ext-Eq gives the meaning to the context equality: if we have two equal
+contexts and one of them contains equal types, then the variables of these
+types are the same in both contexts.
+
+Var says that if a context contains a variable of a certain type, then the
+judgement that this variable has that type in the context is satisfied.
+
+Rules C-Eq define that context equality is an equivalence relation, very
+similar rules for terms and types, stating that their equalities are also
+equivalence relations, are omitted for brevity.
+
+The rule Tm-Conv states that equivalent types and equivalent contexts may be
+used in place of each other in judgements on terms, while the rule Ty-Conv
+states that equivalent contexts can be used in place of each other in
+judgements on types.
+
+The rule Weak says that we can extend our context with a valid type, preserving
+the validity of the typing judgements.
+
+The rule Subst states that the validity of the typing judgements is preserved
+under substitution.
+
+There are two important dependent type formers to consider: dependent product
+and dependent sum. The types formed with them have well-understood analogues in
+set theory and are essential when working with dependent types. These are not
+the only type formers that we can construct and we hope that the general
+principles of their construction will be clear from these two examples.
+
+To present these type formers we will be using formation, introduction,
+elimination, and equality rules.
 
@@
 
-Dependent product (also known as dependent function space or $\Pi$-type)
-corresponds to set-theoretic notion of cartesian product over a family of sets
-$\Pi_{i \in I}B_i$ which has as elements functions mapping an index $i$ to an
-element of a corresponding set $B_i$. In type theory this is expressed as
-follows:
+Dependent product or dependent function type (often abbreviated as $\Pi$-type)
+is a type of a function that returns a result which type may depend on the
+function argument. It is defined by the following rules:
 
-\[
-  \inference[$\Pi$-F]{\Gamma \vdash A~type \qquad \Gamma, x:A \vdash B~type}{\Gamma \vdash \Pi x:A. B~type}
-\]
+\begin{definition}[Dependent product]
+  \begin{mathpar}
+    \inferrule*[right=$\Pi$-F]{\Gamma \vdash A~type \\ \Gamma, x:A \vdash B~type}{\Gamma \vdash \Pi x:A. B~type}
+    \and
+    \inferrule*[right=$\Pi$-Eq]{\Gamma \vdash A_1 = A_2 ~type \\ \Gamma, x:A_1 \vdash B_1 = B_2 ~type}{\Gamma \vdash \Pi x:A_1. B_1 = \Pi x:A_2. B_2~type}
+    \and
+    \inferrule*[right=$\Pi$-I]{\Gamma, x:A \vdash b : B}{\Gamma \vdash \lambda x : A. b : \Pi x:A. B}
+    \and
+    \inferrule*[right=$\Pi$-I-Eq]
+      {\Gamma, x:A_1 \vdash b_1 = b_2 : B_1 \\ \Gamma \vdash A_1 = A_2~type \\ \Gamma, x: A_1 \vdash B_1 = B_2~type}
+      {\Gamma \vdash \lambda x : A_1. b^{B_1} = \lambda x : A_2. b^{B_2} : \Pi x:A_1. B_1}
+    \and
+    \inferrule*[right=$\Pi$-E]{\Gamma \vdash u : \Pi x:A. B \\ \Gamma \vdash a : A}{\Gamma \vdash App_{[x:A]B}(u, a): B[a / x]}
+    \and
+    \inferrule*[right=$\Pi$-C]{\Gamma \vdash \lambda x : A. b : \Pi x:A. B \\ \Gamma \vdash a : A}
+    {\Gamma \vdash App_{[x:A]B}(\lambda x : A. b, a) = b[a / x]: B [a / x]}
+  \end{mathpar}
+\end{definition}
 
-\[
-  \inference[$\Pi$-Eq]{\Gamma \vdash A_1 = A_2 ~type \qquad \Gamma, x:A \vdash B_1 = B_2 ~type}{\Gamma \vdash \Pi x:A_1. B_1 = \Pi x:A_2. B_2~type}
-\]
+The rule $\Pi$-F serves for the formation of the type: if we know that having a
+variable $x: A$ we obtain a type $B$, then we have our $\Pi$-type, making $x$
+bound in that type.
 
-The first rule expresses that a dependent product consists of a type $A$ (that
-possibly depends on other types in $\Gamma$) and a type depending on $A$ (and
-$\Gamma$), namely $B$. The rule $\Pi$-Eq expresses that the definitional
-equality is respected by the $\Pi$-former. Note that the variable $x$ becomes
-bound in $\Pi x:A. B$.
+The rule $\Pi$-Eq expresses that the definitional equality is respected by the
+$\Pi$-former.
 
-To form the elements of the $\Pi$-type we have the introduction rules with
-associated congruence rule:
+To introduce instances of $\Pi$-type we use the rule $\Pi$-I which looks almost
+the same as in the simply-typed case, with the exception that our $\lambda$
+here forms a dependent function type, not the simply-typed one.
 
-\[
-  \inference[$\Pi$-I]{\Gamma, x:A \vdash b : B}{\Gamma \vdash \lambda x : A. b : \Pi x:A. B}
-\]
+The rule $\Pi$-I-Eq serves similar purpose to $\Pi$-Eq and ensures the
+preservation of the definitional equality. The similar equality preservation
+rules henceforth will be omitted for the other rules we introduce, but the
+reader should keep their implicit presence in mind while seeing other term and
+type formers that we introduce.
 
-\[
-  \inference[$\Pi$-I-Eq]{\Gamma, x:A_1 \vdash b_1 = b_2 : B_1 \qquad \Gamma A_1 = A_2~type \qquad \Gamma x: A_1 \vdash B_1 = B_2~type}{\Gamma \vdash \lambda x : A_1.
-b^{B_1} = \lambda x : A_2.
-b^{B_2} : \Pi x:A_1.
-B_1}
-\]
+The rule $\Pi$-E regulates the application of a dependent function to some value
+of the correct type results in the return value of a proper type. Note the rule
+only concerns the resulting type here, not the value and that the bound
+variable $x$ (shown as $[x : A] B$ in subscript) is replaced in the type, not
+the value, which is handled by the next rule.
 
-So to give an element of $\Pi x:A. B$ one must give an element of $B[x]$ in the
-presence of a variable $x$ of type $A$. The congruence rule $\Pi$-I-Eq
-expresses that definitional equality preserves $\Pi$-introduction. In the
-following we will refrain from writing down the congruence rules explicitly,
-having them present implicitly for every type and term formers we introduce.
+The rule $\Pi$-C defines the resulting value of a dependent function
+application. It is almost the same $\beta$-conversion rule as in STLC, with the
+variable substitution also occurring in the type.
 
-Elements of $\Pi$-type are consumed using application:
-\[
-  \inference[$\Pi$-E]{\Gamma \vdash u : \Pi x:A. B \qquad \Gamma \vdash a : A}
-  {\Gamma \vdash App_{[x:A]B}(u, a): B[a / x]}
-\]
-
-Square brackets in the typing annotation $[x : A] B$ here mean that the
-variable $x$ is bound in $B$.
-
-The following rule gives definitional equality to applications --- applying a
-function $\lambda x : A. b^B$ to a term $a : A$ results in $b$ with $x$
-replaced by $a$:
-
-\[
-  \inference[$\Pi$-C]{\Gamma \vdash \lambda x : A. b : \Pi x:A. B \qquad \Gamma \vdash a : A}
-  {\Gamma \vdash App_{[x:A]B}(\lambda x : A. b, a) = b[a / x]: B [a / x]}
-\]
-
-Note that the substitution plays more prominent role here than in simply typed
-lambda calculus: substitution is needed to formulate not only the term
-conversion rules, but also the typing rules.
+The analogue to dependent product in set theory is a cartesian product over a
+family of sets $\Pi_{i \in I}B_i$ where elements are the functions mapping an
+index $i$ to an element of a corresponding set $B_i$.
 
@@
 
-The next type former we consider is the $\Sigma$-type, or dependent sum,
-corresponding to disjoint union in set theory. If we are given a family of sets
-$(B_i)_{i \in I}$ we can form the set $\Sigma_{i \in I} B_i := \{ (i, b)~|~i
-\in I \land b \in B_i\}$ whose elements consist of an index $i$ and an element
-of the corresponding set $B_i$. In type theory the corresponding formation and
-introduction rules look as follows.
-\[
-  \inference[$\Sigma$-F]{\Gamma \vdash A \text{ type}\\ \Gamma, x:A \vdash B \text{ type}}{\Gamma \vdash \Sigma x : A. B \text{ type}}
-\]
-\[
-  \inference[$\Sigma$-I]{\Gamma \vdash a : A\\ \Gamma\vdash b: B[a / x]}{\Gamma \vdash Pair_{[x:A]B} (a,b) : \Sigma x : A. B}
-\]
+The other very important type former is dependent sum type (also known as
+dependent pair or $\Sigma$-type).
 
-Elimination rule:
-\[
-  \inference[$\Sigma$-E]{\Gamma, z : \Sigma x: A. B \vdash C \text{ type}\\ \Gamma, x: A, y: B \vdash c : C[Pair_{[x:A]B}(x, y)/z]\\ \Gamma \vdash u : \Sigma x: A. B}
+\begin{definition}[Dependent sum]
+  \begin{mathpar}
+    \inferrule*[right=$\Sigma$-F]{\Gamma \vdash A \text{ type}\\ \Gamma, x:A \vdash B \text{ type}}{\Gamma \vdash \Sigma x : A. B \text{ type}}
+    \and
+    \inferrule*[right=$\Sigma$-I]{\Gamma \vdash a : A\\ \Gamma\vdash b: B[a / x]}{\Gamma \vdash Pair_{[x:A]B} (a,b) : \Sigma x : A. B}
+    \and
+    \inferrule*[right=$\Sigma$-E]
+      {\Gamma, z : \Sigma x: A. B \vdash C \text{ type} \\ \Gamma, x: A, y: B \vdash c : C[Pair_{[x:A]B}(x, y)/z] \\ \Gamma \vdash u : \Sigma x: A. B}
       {R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, u): C[u/z]}
-\]
+    \and
+    \inferrule*[right=$\Sigma$-C]
+      {\Gamma\vdash R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, Pair_{[x:A]B}(a,b)): C[Pair_{[x:A]B}/z]}
+      {\Gamma\vdash R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, Pair_{[x:A]B}(a,b)) = c[a/x,b/y] : C [Pair_{[x:A]B}/z]}
+  \end{mathpar}
+\end{definition}
 
-And definitional equality:
-\[
-  \inference{\Gamma\vdash R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, Pair_{[x:A]B}(a,b)): C[Pair_{[x:A]B}/z]}
-  {\Gamma\vdash R^\Sigma_{[z : \Sigma x: A. B]C}([x : A, y: B]c, Pair_{[x:A]B}(a,b)) = c[a/x,b/y] : C [Pair_{[x:A]B}/z}
-\]
+The formation rule $\Sigma$-F follows the same pattern as the rule for
+dependent function, but the introduction is a bit different. Note that we take
+such $b$, where the type contains a variable $x : A$ that is substituted by the
+value $a$. Together the values $a$ and $b$ form a pair, where the type of the
+second element depends on the value of the first. We use the notation $[x : A]
+B$ here again to indicate that the variable $x$ is bound in the type $B$.
 
-The projection functions for the dependent sum can be defined using these
-rules:
+The rule $\Sigma$-E is again analogous to the one seen in dependent product: we
+can replace a variable of a dependent pair type with a concrete term if he
+have a term that contains that variable. And again we just say what happens to
+the type, not the value there.
+
+To define the effect on the values we use the rule $\Sigma$-C, which says that
+the elements of the pair will replace the corresponding variables in the body
+of the given term, but not the pair variable itself, such substitution occurs
+only in the type.
+
+We know the projection functions for the simply-typed products in STLC. Let us
+now define the projection functions for the \emph{dependent} products, which we
+call $\mathbf{p}$ for the first projection and $\mathbf{q}$ for the second
+projection:
 \[
@@
 \]
-for the first projection and
 \[
@@
 \]
-for the second projection, assuming $\Gamma \vdash A~type$, $\Gamma, x:A \vdash
-B~type$, and $\Gamma \vdash a : \Sigma x: A. B$.
+where $\Gamma \vdash A~type$, $\Gamma, x:A \vdash B~type$, and $\Gamma \vdash a
+: \Sigma x: A. B$.
 
-The important special cases of $\Pi x: A. B$ and $\Sigma x : A. B$ arise when
-$B$ does not actually depend on $A$. The we write
+The analogue to dependent sum in set theory is a disjoint union. For every
+family of sets $(B_i)_{i \in I}$ we can form the set $\Sigma_{i \in I} B_i :=
+\{ (i, b)~|~i \in I \land b \in B_i\}$ whose elements consist of an index $i$
+and an element of the corresponding set $B_i$.
+
+It is worth notice that we may define the simply-typed products and functions
+in terms of dependent types: we just need the types to not depend on each
+other. So, assuming that $B$ does not depend on $A$, we get:
 \[
@@
 \]
-indicating that in this cases $\Pi$- and $\Sigma$-types correspond to ordinary
-non-dependent function and cartesian product types respectively.
 
@@
 
-We will use categories with families (CwF) to model dependent types. First
-consider what is a category of families. The category of families of sets,
-$\fcat{Fam}$, has as objects $(A, B)$ where $A$ is a set and $B = (B_a~|~a \in
-A)$ is an $A$-indexed family of sets $B_a$, $a \in A$. A morphism $(A, B) \to
-(A', B')$ is given by a pair $(f, g)$ where $f: A \to A'$ and $g$ is an
-$A$-indexed famiy of functions such that $g_a: B_a \to B_{fa}'$.
+There are several ways to model dependent types with category theory, the one
+we show is using Categories with Families.
 
+\begin{definition}[Category of families of sets]
+  The category of families of sets, $\fcat{Fam}$, has as objects $(A, B)$ where
+  $A$ is a set and $B = (B_a~|~a \in A)$ is an $A$-indexed family of sets $B_a$,
+  $a \in A$. A morphism $(A, B) \to (A', B')$ is given by a pair $(f, g)$ where
+  $f: A \to A'$ and $g$ is an $A$-indexed famiy of functions such that $g_a: B_a
+  \to B_{fa}'$.
+\end{definition}
+
+\begin{definition}[Category with Families (CwF)]
 A \emph{category with families} is given by ($\cat{C}$,$F$) where:
@@
 \end{itemize}
+\end{definition}
 
-We say that a category with families supports $\Pi$- and $\Sigma$-types if it's
+We say that a category with families supports $\Pi$- and $\Sigma$-types if it is
 closed under the corresponding rules shown in previous sections.

### 2017-09-21 papers/talk-11/paper-11.tex
Add subsection on context morphisms

@@
 
+\subsection*{Context morphism}
+
+\begin{definition}[Context morphism]
+  Let $\Gamma$ and $\Delta := x_1: A_1, \dots, x_n : A_n$ be valid contexts. If
+  $\sigma := (a_1, \dots, a_n)$ is a sequence of $n$ valid terms we say that
+  $\sigma \from \Gamma \to \Delta$ is a \emph{context morphism} if the
+  following $n$ judgements hold:
+
+  \begin{align*}
+    &\Gamma \vdash a_1 : A_1\\
+    &\Gamma \vdash a_2 : A_2[a_1/x_1]\\
+    &\dots\\
+    &\Gamma \vdash a_n : A_n[a_1/x_1][a_2/x_2]\dots[a_{n-1}/x_{n-1}]
+  \end{align*}
+\end{definition}
+
+\begin{example}
+  We always have the empty context morphism $()$ from any context $\Gamma$ to
+  $\diamond$, it does not contain any terms and is the only context morphism
+  from $\Gamma$ to $\diamond$.
+\end{example}
+\begin{example}
+  We also have the identity context morphism $1$ (from now on we will use $1$
+  as the suitable identity morphism for the surroundings) for any $\Gamma = x_1
+  : A_1, \dots, x_n : A_n$, which is $(x_1, \dots, x_n)$.
+\end{example}
+
+A note on notation: we sometimes write $A\sigma$, where $\Gamma \vdash
+A\sigma~type$, $\Delta \vdash A~type$, and $\sigma \from \Gamma \to \Delta$ for
+some contexts $\Gamma$ and $\Delta$, to denote the application of $\sigma$ on
+the right. That is, $A\sigma$ becomes here a type in the domain context of
+$\sigma$, where $\sigma(A\sigma) = A$. The same notation is sometimes
+analogously used for terms.
+
+\begin{definition}[Context extension]
+  If $\vdash \Gamma~ctxt$ and $\Gamma \vdash A~type$, there is a context
+  $\Gamma, x:A$ (for brevity also written as $\Gamma.A$ from now on), a context
+  morphism $\mathbf{p}: \Gamma.A \to \Gamma$, and a term $\Gamma.A \vdash
+  \mathbf{q} : A \mathbf{p}$. This should satisfy the following property: for
+  each $\vdash \Delta~ctxt$, a context morphism $\sigma \from \Delta \to
+  \Gamma$, and $\Delta \vdash a : A \sigma$, there is a context morphism
+  $(\sigma, a): \Delta \to \Gamma.A$ called \emph{context extension} such that:
+  \[
+    \mathbf{p}(\sigma, a) = \sigma \quad \mathbf{q}(\sigma, a) = a \quad (\sigma, a) \tau =
+    (\sigma \tau, a \tau) \quad (\mathbf{p}, \mathbf{q}) = 1
+  \]
+  (Note that in the last equation $1: \Gamma.A \to \Gamma.A$)
+\end{definition}
+
 \section*{Categories with Families}
@@
 \begin{itemize}
-  \item[$\bullet$] $\cat{C}$ is a category, objects: $\Gamma, \Delta, \dots$ are
-    \emph{contexts}, morphisms:  $\sigma, \tau, \dots$ are
-    \emph{substitutions} or context morphisms
+  \item[$\bullet$] $\cat{C}$ is a category, whose
+
+    objects: $\Gamma, \Delta, \dots$ are \emph{contexts}
+
+    morphisms:  $\sigma, \tau, \dots$ are \emph{context morphisms}
 
@@
     following equations also hold:
-
     \[
@@
     \]
-
-  \item[$\bullet$] There is also the operation of \emph{context extension}: if
-    $\vdash \Gamma~ctxt$ and $\Gamma \vdash A~type$, there is a context
-    $\Gamma, x:A$ (for brevity also written as $\Gamma.A$ from now on), a
-    context morphism $\mathbf{p}: \Gamma.A \to \Gamma$, and a term $\Gamma.A
-    \vdash \mathbf{q} : A \mathbf{p}$. This should satisfy the following
-    property: for each $\vdash \Delta~ctxt$ and a context morphism
-    $\sigma \from \Delta \to \Gamma$ and $\Delta \vdash a : A \sigma$, there is
-    a context morphism $(\sigma, a): \Delta \to \Gamma.A$ such that:
-    \[
-      \mathbf{p}(\sigma, a) = \sigma \quad \mathbf{q}(\sigma, a) = a \quad (\sigma, a) \tau =
-      (\sigma \tau, a \tau) \quad (\mathbf{p}, \mathbf{q}) = 1
-    \]
-    Where in the last equation $1: \Gamma.A \to \Gamma.A$
 \end{itemize}
@@
 We also add the rules for context morphisms:
-
-\[
-  \inference{\Gamma \vdash A~type \qquad \sigma \from \Delta \to \Gamma}{\Delta \vdash A\sigma} \qquad
-  \inference{\Gamma \vdash a : A \qquad \sigma \from \Delta \to \Gamma}{\Delta \vdash a\sigma : A\sigma}
-\]
-\[
-  \inference{\sigma \from \Delta \to \Gamma \qquad \Gamma \vdash A \qquad \Delta \vdash a : A\sigma}{(\sigma, a) \from \Delta \to \Gamma.A}
-\]
 

### 2017-09-21 papers/talk-11/paper-11.tex
Update the writeup11

@@
 \begin{example}
-  We also have the identity context morphism $1$ (from now on we will use $1$
-  as the suitable identity morphism for the surroundings) for any $\Gamma = x_1
-  : A_1, \dots, x_n : A_n$, which is $(x_1, \dots, x_n)$.
+  We also have the identity context morphism $1$ for any $\Gamma = x_1 : A_1,
+  \dots, x_n : A_n$, which is $(x_1, \dots, x_n)$.
 \end{example}
@@
 $\sigma$, where $\sigma(A\sigma) = A$. The same notation is sometimes
-analogously used for terms.
+analogously used for terms: $a\sigma$ to denote a term $\Gamma \vdash a\sigma :
+A\sigma$, where $\Delta \vdash a : A$.
+
+Having this notation, we also want the following equations to hold:
+\[
+  A1=A \quad (A\sigma)\tau = A(\sigma\tau) \quad a1=a \quad (a\sigma)\tau=a(\sigma\tau)
+\]
+For readability, we usually omit writing indices to the identity morphism $1$,
+using it as the most suitable identity morphism for the surroundings. We also
+assume correct types in the equations, i.e. $(A\sigma)\tau = A(\sigma\tau)$
+assumes $\sigma \from \Delta \to \Gamma$, $\tau \from \Theta \to \Delta$, and
+$\Gamma \vdash A~type$.
 
@@
   \]
-  (Note that in the last equation $1: \Gamma.A \to \Gamma.A$)
+  Note that in the last equation $1: \Gamma.A \to \Gamma.A$ and that we leave
+  the indices on $\mathbf{p}$ implicit.
 \end{definition}
@@
     context $\Gamma$ we write $Ty_F(\Gamma)$ for the indexing set of the family
-    $F(\Gamma)$ and its family as $(Ter_F(\Gamma;A) | A \in Ty_F(\Gamma))$;
+    $F(\Gamma)$ and its family as $(Ter_F(\Gamma;A)~|~A \in Ty_F(\Gamma))$;
     $\Gamma \vdash A~type$ means that $A \in Ty_F(\Gamma)$ and $\Gamma \vdash a
-    : A$ means $a \in Ter_F(\Gamma;A)$. For a context morphism $\sigma \from
-    \Delta \to \Gamma$ the morphism $F(\sigma)$ acts on types $\Gamma \vdash
-    A~type$ as $A\sigma$ and on terms $\Gamma \vdash a : A$ as $a\sigma$. The
-    following equations also hold:
-    \[
-      A1=A \quad (A\sigma)\tau = A(\sigma\tau) \quad a1=a \quad (a\sigma)\tau=a(\sigma\tau)
-    \]
+    : A$ means $a \in Ter_F(\Gamma;A)$. Note that for a context morphism
+    $\sigma \from \Delta \to \Gamma$ the morphism $F(\sigma)$ acts on types
+    $\Gamma \vdash A~type$ as $A\sigma$ and on terms $\Gamma \vdash a : A$ as
+    $a\sigma$.
 \end{itemize}
@@
 
-We say that a category with families supports $\Pi$- and $\Sigma$-types if it is
-closed under the corresponding rules shown in previous sections.
-
-We also add the rules for context morphisms:
+A category with families supports $\Pi$- and $\Sigma$-types if it is
+closed under the $\Pi$ and $\Sigma$ rules.
 
@@
 
-We will now show how any presheaf category gives rise to a category with
-families where the contexts are presheaves.  Let $\cat{C}$ be a small category.
-In the following objects of $\cat{C}$ will be denoted as $I,J,K$ and morphisms
-as $f,g,h$.
+Let us now consider the categories of presheaves and see that they form a
+category with families where the contexts are presheaves.  Let $\cat{C}$ be a
+small category. We will refer to the objects of $\cat{C}$  as $I,J,K$ and
+morphisms as $f,g,h$.
 
@@
 
-The empty context then is a terminal presheaf and a context morphism $\sigma
-\from \Delta \to \Gamma$ is a natural transformation making the diagram
-commute.
-
+The empty context $\diamond$ is a terminal presheaf and a context morphism
+$\sigma \from \Delta \to \Gamma$ is a natural transformation that makes the
+diagram commute.
 \[
@@
 
-If we write $\Gamma_f$ and $\Delta_f$ as $f$ acting on the right, the
-equation becomes:
+In the following, we will omit the indices on $\sigma$; Then if we write
+$\Gamma_f$ and $\Delta_f$ as $f$ acting on the right, the equation becomes:
 \[

### 2017-09-21 papers/talk-11/paper-11.tex

@@
 \usepackage{mathpartir}
+
+\bibliographystyle{plainurl}
 
@@
 
-The context $\Gamma$ is then a presheaf on $\cat{C}$, $\Gamma \from
-\cat{C}^\op \to \Set$, that is, we are given a set $\Gamma(I)$ for each
-$I$ in $\ob(\cat{C})$, and functions $\Gamma(I) \to \Gamma(J)$, $\rho \mapsto \rho
-f$ (called \emph{restriction} and written as $f$ acting on the right) for each
-$f: J \to I$ such that
+The context $\Gamma$ is then a presheaf on $\cat{C}$, $\Gamma \from \cat{C}^\op
+\to \Set$, that is, we are given a set $\Gamma(I)$ for each $I$ in
+$\ob(\cat{C})$, and functions $\Gamma(I) \to \Gamma(J)$, $\rho \mapsto \rho f$
+(called \emph{restriction} and written as $f$ acting on the right) for each $f:
+J \to I$ such that
 \[
@@
 In the following, we will omit the indices on $\sigma$; Then if we write
-$\Gamma_f$ and $\Delta_f$ as $f$ acting on the right, the equation becomes:
+$\Gamma_f$ and $\Delta_f$ as $f$ acting on the right (recall the session on
+presheaves), the equation becomes:
 \[
@@
 
-\begin{example}[Ordinary function type]
+\begin{example}
   Let us consider a term of type $A \to A$, and the elements in $Ty(\Gamma)$
   and $Ter(\Gamma, A \to A)$ for some $\vdash \Gamma~ctxt$, $\Gamma \vdash
-  A~type$.
+  A~type$ in the presheaf model.
 
@@
 
-\section*{Conclusion}
+\section*{Credits}
+
+Most of the material on the dependent types bases on \cite{Hofmann}, while
+starting from categories with families we closely follow \cite{Huber}.
+
+\bibliography{main}
 

### 2017-09-21 papers/talk-6/paper-6.tex

@@
 \def\pathToRoot{../../}\input{\pathToRoot headers/header_lecturenotes}
-
-\usepackage{todonotes}
 
@@
   and $P'$ be another product of $X$ and $Y$ together with $p'_1 \from P' \to X, p'_2 \from P' \to Y$.
-  We apply definition \ref{def:prod} multiple times and get the following commuting diagram:
+  We apply Definition \ref{def:prod} multiple times and get the following commuting diagram:
   \[ \xymatrix{
@@
   \label{ex:prod:cart2}
-  For sets we can define another product, where we reverse the order of the elements, i.~e.:
+  For sets we can define another product, where we reverse the order of the elements, i.e.:
   \begin{alignat*}{2}
@@
   \end{alignat*}
-  Now $X \times Y$ and $X \times' Y$ are isomorphic with the involution $i(a, b) := \pair{b}{a}$.
+  Now $X \times Y$ and $X \times' Y$ are obviously isomorphic.
 \end{example}
@@
   Let $\pair{A}{\le}$ be a poset.
-  A \demph{lower bound} for two objects $x, y \in A$ is a element $a \in A$ such that $a \le x$ and $a \le y$.
+  A \demph{lower bound} for two objects $x, y \in A$ is an element $a \in A$ such that $a \le x$ and $a \le y$.
   A \demph{greatest lower bound} of $x, y \in A$ is a lower bound $z \in A$ of $x$ and $y$ such that for all lower bounds $a \in A$ of $x$ and $y$, $a \le z$.
@@
   When we see the poset $\pair{A}{\le}$ as a category $\catA$ and have a greatest lower bound $z$ of $x, y \in A$, then $z$ is the product of $x$ and $y$.
-  To see this, we apply the definition \ref{def:prod}.
+  To see this, we apply the Definition \ref{def:prod}.
   Let $x, y \in \catA$ and $z \in \catA$ be a greatest lower bound of $x$ and $y$.
@@
   Let $K$ be a field and $V, W$ vector spaces over this field.
-  Then we define\footnote{One should verify that the defined structure is indeed a vector field.}
-  the \demph{direct sum} $V \oplus W := \bigl(V \times W, +, \cdot \bigr)$ as a new vector field with the following operations:
+  Then we define\footnote{One should verify that the defined structure is indeed a vector space.}
+  the \demph{direct sum} $V \oplus W := \bigl(V \times W, +, \cdot \bigr)$ as a new vector space with the following operations:
   \begin{alignat*}{4}
@@
   \label{def:arb:prod}
-  Let $\catA$ be a category, $I$ a set (i.~e. of indices) and $\family{X_i}{i \in I}$ be a family of objects of $\catA$.
+  Let $\catA$ be a category, $I$ a set (i.e. of indices) and $\family{X_i}{i \in I}$ be a family of objects of $\catA$.
   A product of $\family{X_i}{i \in I}$ consists of an object $P \in \catA$ and the projection maps $\family{P \toby{p_i} X_i}{i \in I}$, such that:
@@
   An \demph{arbitrary Cartesian tuple} of $\family{X_i}{i \in I}$ consists of a family of elements $\pi = \family{\pi_i \in X_i}{i \in I}$.
-  The \demph{arbitrary Cartesian product} $\prod_{i \in I}{X_i}$ is then set set of all such tuples.
+  The \demph{arbitrary Cartesian product} $\prod_{i \in I}{X_i}$ is then the set set of all such tuples.
   We define the projection functions $p_i(\pi) := \pi_i$ for $i \in I$.
   Let $A$ be a set and $\family{A \toby{f_i} X_i}{i \in I}$ be a family of maps.
-  Then we define $\fbar(a) := \family{f_i(a)}{i \in I}$. $p_i \of \fbar = f_i$ holds by definition for $i \in I$.
+  Then we define $\fbar(a) := \family{f_i(a)}{i \in I}$.
+  Then $p_i \of \fbar = f_i$ holds by definition for $i \in I$.
   When we have another function $\ftilde \from A \to \bigl(\prod_{i \in I}{X_i}\bigr)$ with $p_i \of \ftilde = f_i$ for $i \in I$, then $\fbar = \ftilde$, since for every $i \in I$:
@@
 Now we want to abstract the concept of the disjoint union of sets.
-We will see that the generalized structure is the dual of the product.
+The generalized structure is the dual of the product.
 
@@
   \label{ex:coprod:vspace}
-  In Example \ref{ex:prod:vspace} we have seen that the direct sum of two vector spaces $V$ and $W$ is the product of $V$ and $W$ over a field $K$ in the category $\Vect_K$.
-  We will now see that it is \emph{also} the coproduct.
+  In Example \ref{ex:prod:vspace} we have seen that the direct sum of two vector spaces $V$ and $W$ over a field $K$ is the product of $V$ and $W$ in the category $\Vect_K$.
+  In this example we show that it is \emph{also} the coproduct.
   In general products and coproducts are not the same as we have seen in Examples \ref{ex:prod:cart} and \ref{ex:coprod:dunion}.
@@
   \label{def:arb:coprod}
-  Let $\catA$ be a category, $I$ a set (i.~e. of indices) and $\family{X_i}{i \in I}$ be a family of objects of $\catA$.
+  Let $\catA$ be a category, $I$ a set (i.e. of indices) and $\family{X_i}{i \in I}$ be a family of objects of $\catA$.
   A product of $\family{X_i}{i \in I}$ consists of an object $C \in \catA$ and the coprojection maps $\family{X_i \toby{p_i} C}{i \in I}$, such that:
@@
 In Example \ref{ex:arb:prod:cart} we have generalized the Cartesian product.
-No we will generalize Example \ref{ex:coprod:dunion} and define the disjoint union of a a family of sets.
+No we generalize Example \ref{ex:coprod:dunion} and define the disjoint union of a family of sets.
 
@@
   \label{ex:arb:coprod:dunion}
-  Let $\family{X_i}{i \in I}$ be a familiy of sets. We define the \demph{arbitrary disjoint union}
+  Let $\family{X_i}{i \in I}$ be a family of sets. We define the \demph{arbitrary disjoint union}
   $$\bigdunion_{i \in I}{X_i} := \bigcup_{i \in I} \setMap{\pair{x}{i}}{x \in X_i}.$$
@@
 
-Solutions sets and minus cores (as defined bellow) share the same structure.
-We will generalize that structure and it equalizer.
+Solution sets and minus cores (as defined bellow) share the same structure.
+We generalize that structure and call it \emph{equalizer}.
 
@@
 
-We define coequalizer as the dual of the equalizer. We will see that an example for this structure are quotient sets that are generated by a equivalence relation.
+We define coequalizers as the dual structure of equalizers.
+We show that quotient sets that are generated by a equivalence relation are an example for this structure.
 
@@
   $R^{-1} := \setMap{\pair{b}{a}}{\pair{a}{b} \in R}$.
-  The \demph{symmetric transitive closure} of $R$ is $\sim := \bigcup_{n\in\nat} {\hat R}^n$
-  where ${\hat R}^n$ is recursively defined with
+  The \demph{symmetric reflexive transitive closure} of $R$ is $\sim ~:= \bigcup_{n\in\nat} {\hat R}^n$
+  where ${\hat R}^n$ is recursively defined by
   ${\hat R}^0 := \setMap{\pair{a}{a}}{a \in X}$ and
@@
   be a cofork.
-  By induction we can show that $f$ respects the relation $\sim$, i.~e. $f(y_1) = f(y_2)$ whenever $y_1 \sim y_2$.
+  By induction we can show that $f$ respects the relation $\sim$, i.e. $f(y_1) = f(y_2)$ whenever $y_1 \sim y_2$.
 
-  To define the function $\fbar \from C \to A$ we need a choice function $\choice \cdot \from C \to Y$ that returns for every equivalence class of $\sim$ a member of it, i.~e.
+  To define the function $\fbar \from C \to A$ we need a choice function $\choice \cdot \from C \to Y$ that returns for every equivalence class of $\sim$ a member of it, i.e.
   $\forall c \in C, \forall y \in Y, y \in c \Leftrightarrow y \sim \choice{c}$
@@
 
-% XXX Insert exercise 6.7 as further example?
-
 \section*{Pullbacks and Pushouts}
 
-In this section we will define our last two structures: pullbacks and their duals, pushouts.
+In this section we define our last two structures: pullbacks and their duals, pushouts.
 
@@
   } $
-  be a functions and sets.
+  be a digram in $\Set$.
   We want to define a set $P$ and two functions $f_1 \from X \to A, f_2 \from Y \to A$, such that the diagram
@@
   commutes.
-  We define $P := (X \dunion Y) / \sim$ where $\sim \; \subseteq (X \dunion Y)^2$ is the symmetric reflexive transitive closure (see definition \ref{def:srtc}) over the relation
+  We define $P := (X \dunion Y) / \sim$ where $\sim \; \subseteq (X \dunion Y)^2$ is the symmetric reflexive transitive closure (see Definition \ref{def:srtc}) over the relation
   $R := \setMap{\pair{\pair{s(z)}{0}}{\pair{t(z)}{1}}}{z \in Z} \subseteq (X \dunion Y)^2$.
@@
   \end{alignat*}
-  By induction we can show that $\fbar'$ respects the relatation $\sim$, i.~e. $\forall a,b \in X \dunion Y, a \sim b \Rightarrow \fbar'(a) = \fbar'(b)$.
+  By induction we can show that $\fbar'$ respects the relation $\sim$, i.e. $\forall a,b \in X \dunion Y, a \sim b \Rightarrow \fbar'(a) = \fbar'(b)$.
 

### 2017-09-21 sheets/talk-6/sheet-6.tex

@@
 \begin{answer}
-  Let $(A, \le)$ be a poset concidered as a category and $I$ a set.
+  Let $(A, \le)$ be a poset considered as a category and $I$ a set.
   A \demph{lower bound} of a family $(x_i)_{i\in I}$ is an element $a \in A$, s.\,t. $\forall i \in I: a \le x_i$.
@@
 
-\begin{definition}{Minus-core of linear maps in vector spaces}
+\begin{definition}{Minus-core of linear maps between vector spaces}
   Let $V, W$ be vector spaces over a field $K$. Let $s, t \from V \to W$ be linear functions.
@@
   }$
-  be a diagram in $Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagram.
+  be a diagram in $\Vect_K$ for a field $K$. Show that $core(s-t)$ is the equalizer for this diagram.
 \end{exercise}
@@
   \begin{proof}
-    We show that $S := P$ is the coproduct of $X$ and $Y$ together with the coprojection functions $p_1$ and $p_2$.
+    We show that $P$ is the coproduct of $X$ and $Y$ together with the coprojection functions $p_1$ and $p_2$.
     Let
@@
     } \]
-    would also commute (because $Z$ is initial) but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
+      would also commute (because $Z$ is initial), but $\bar{f}$ is unique, hence $\bar{f} = \tilde{f}$.
   \end{proof}
@@
         } \]
-        commutes. We have to show $j \of \bar{f} = f_2$ and $h \of \bar{f} = f_1$. The first equatation holds by assumption.
+        commutes. We have to show $j \of \bar{f} = f_2$ and $h \of \bar{f} = f_1$. The first equation holds by assumption.
 
@@
         commutes.
-        By showing that this diagram comutes for $\bar{g} = f_1$ and for $\bar{g} = h \of \bar{f}$ we show $h \of \bar{f} = f_2$.
+        By showing that this diagram commutes for $\bar{g} = f_1$ and for $\bar{g} = h \of \bar{f}$ we show $h \of \bar{f} = f_2$.
         \begin{itemize}
@@
         commutes.
-        The diagram comutes for $\bar{g}$:
+        The diagram commutes for $\bar{g}$:
         \begin{align*}

### 2017-09-21 papers/talk-8/CCC.tex
talk 8 added at right place

\def\pathToRoot{../}\input{\pathToRoot headers/uebungHeader}

\begin{document}
\title{Cartesian Closed Categories}
\author{Andreas Meyer}
\maketitle

\section{Introduction}
If we take a look on, for instance, the category of sets \textbf{Sets}, we may observe that the cartesian product $A \times B$, $\forall A, B \in ob(\textbf{Sets})$ generates again objects in \textbf{Sets}. There are more categories equipped with the property, that their products give rise to a new elements in the collection of their objects, in addition, there are often other functors with the same behaviour.
\\
In some cases, where these other functors happen to be left adjoints to the product, these categories' objects posess a natural- number like arithmetic and we talk about cartesian closed categories.
Furthermore, the internal logic of cartesian closed categories is the simply typed lambda calculus, which will not be topic of this write up, but shortly mentioning it provides some more context to argue, why  it is appropriate to have a look onto the abstract concept of cartesian closed categories.
In the following, CCC will serve as an abbrevation.
\\
This writeup will range over a this introduction with a short defintion of CCC, the concept of exponentials, and detail a more compact characterisation as well as the mentioned arithmetical properties.
Examples of CCC's like \textbf{Sets}, \textbf{Posets} and, \textbf{HA}, where \textbf{HA} will be the category of Heyting Algebras, shall provide a better intuition for what CCC looks like, and how they can be identified.
This is complemented by some counterexamples like  $\boldsymbol{Vect_k}$. 
\\

To be cartesian closed, a category $\mathcal{A}$ has to meet two requirements, as coarsly detailed in the subsequent definition.
\begin{definition}{Cartesian Closed Category}
  A Cartesian Closed Category has all finite products and exponentials.

\end{definition}
\\
While it is required, that the reader knows about products, the idea of exponentials is assumed to be new and needs some more explanations.
\\
Regarding the products, just assess at this point, that the notion of all finite products is exhaustively described by terminal objects and binary products, because a terminal object represents the empty product, whereas every other product can be constructed with binary products. Since the whole concept of CCC is based on categories, whose functors are again objects, the existence of all finite products follows inductively, as long as it is clear that binary products are again objects.

\section{Exponentials}

Exponentials can be best described as right adjoints to binary products.
They are equipped with evaluation arrows $\epsilon$ and transposition $\lambda$.
This is an extension of the definition beyond the characterisation as right adjoint and it is necessary to at least mention the existence of both arrows at this point, because some of the examples would not work without introducing them here, for assuming and verifying the existence of exponential arrows for some category $\mathcal{A}$ does not always imply these being elements of $ob(\mathcal{A})$.
So for an exponential of $\mathcal{A}$ to be an object of $\mathcal{A}$, we sometimes need to verify that the categories defining properties hold for the evaluation arrow as well as for the transposition.
So we obtain the following as a "working definition".

\newpage

\begin{definition}{Exponentials}
  An Exponential $A^B$ of a category $\mathcal{A}$ with binary product $A \times B$ is a right adjoint to this product, such that, $\forall B \in \mathcal{A}, \ Hom( - \times B, -) \cong Hom(- , (-)^B)$ holds, which comes $\forall A,B,C \in \mathcal{A}$ with
  \begin{itemize}
    \item an evaluation arrow $\epsilon : C^B \times B \to C$
    \item an transposition $\lambda f: A \to C^B$ for each arrow $f: A \times B \to C$
  \end{itemize}
  such that this diagram commutes:
\[
  \begin{tikzcd}[row sep=huge]
    & A \times B \arrow{dr}{f}  \arrow{rr}{\lambda f \times 1_B} && C^B\times B \arrow{dl}{\epsilon} \\
    &&C 
  \end{tikzcd}
\]
\end{definition}

%\begin{tikzcd}
%      & 0' \arrow{dr}{g} \arrow{rr}{id_{0'}}    &&  0'\\
%      0 \arrow{ur}{f} \arrow{rr}{id_0} && 0 \arrow{ur}{f}
%    \end{tikzcd}

\section{Examples}
Now, we have what is necessary to look at a few examples to justify, that there is a need for the abstract concept of CCC. 
\subsection{\textbf{Sets}}
\begin{itemize}
\item \underline{\textit{Products:}} As already argued in the introduction, the cartesian product yields new objects $A\timesB\in ob(\textbf{Sets})$. Furthermore, it delivers the empty product $\emptyset \times \emptyset \to \emptyset$. Because all larger products can be build form binary products, \textbf{Sets} meets the first requirement to be a CCC. 
\item  \underline{\textit{Exponentials:}} As exponential, we choose the function space $\forall A, B \in ob(\textbf{Sets}), A^B := \{ f: A \to B| f functional\}$.
  \\ As proof obligation arises, that $\textbf{Sets}(A \times B, C ) \cong \textbf{Sets}(A, C^B), \forall A,B,C \in ob(\textbf{Sets})$ holds; that means that there are two arrows $f:A \times B \to C$ and $f':A \to C^B$ with $f (x,y) = f' (x) (y), \forall x \in A, y \in B$.
  \begin{itemize}
\item "$\Rightarrow$" : Assume $f: A \times B \to C$ with $f(x,y) \in A \times B \to C, x \in A, y \in B$.
\\ By fixing $x \in A$, one obtains an arrow $$f_x$$ with $$f_x(y) \in C^B.$$ By unfixing x, $f_x$ is morphed in the curried function $$\mu f:A\toC^B$$ with $$\mu f (x)(y) \in A \to C^B,$$ such that $$\mu f(x)(y)=f(x,y).$$ 
\item "$\Leftarrow$" : Assume $f': A \to C^B$. One simply defines a cartesian function $$g: A \times B \to C$$ as $$g(x,y) := f(x)(y), x \in A, y \in B$$ and is done.
  \end{itemize}
\end{itemize}

\newpage

  \subsection{\textbf{Posets}}
  The example of the category of preordered sets requires to exploit the evaluation arrow $\epsilon$ and the transposition $\lambda$ to verify the existence of exponential objects.
    \begin{itemize}
  \item \underline{\textit{Products:}}
    For all elements P, Q $\in ob(\textbf{Posets})$, the arrows are the monotone functions,
    \\ Hom(P,Q) := $\{f: P \to Q \mid f \quad  monotone \}$ with $\forall p,p', p\leq p' \to f p \leq f p'$.
  We can construct a binary product $P \times Q$, which is again a poset,using pairs $$(p,q) \in ob(P\timesQ)$$ and a partial ordering $$(p,q) \leq (p',q') \quad iff\quad p \leq p' \wedge q \leq q' .$$
  From this follow monotone projections: \\
\[
  \begin{tikzcd}
    & P && P\times Q \arrow{ll}{\pi_2} \arrow{rr}{\pi_1}  &&  Q\\
  \end{tikzcd}
\]
  And a monontone pairing
  $$ <f,g>: X \to P\times Q$$
  if $f:X \to P$ and $g: X\to Q$ are monotone.

\item \underline{\textit{Exponentials}}
  As exponential $Q^P$, we consider a set of monotone functions
  $$Q^P:=\{f:Q\toP \mid f \quad monotone \}$$
  with an pointwise ordering, such that $\forall f,g \in Q^P$,
  $$ f \leq g \leftrightarrow f p \leq g p,\forall p \in P.$$

  The evalution comes as an arrow
  $$\epsilon: Q^P\times P \to Q \quad with \quad \epsilon(f,p):=f(p)$$
  and the transposition is the arrow
  $$\lambda f : X \to Q^P \quad with \quad \lambda f(x) \in Q^P$$
  for some poset X and arrow
  $$f:X \times P \to Q \quad with \quad f(x,p):=\lambda f(x)(p)$$
  If these arrows proof to be monotone, $Q^P$ is an exponential object within \textbf{Posets}
  For this, we take some pairs $(f,p),(f',p')\in Q^P\times P, (f,p)\leq(f',p')$ and plug them into $\epsilon$:
  $$\epsilon(f,p) = f(p)$$
  $$\leq f(p')$$
  $$\leq f'(p') $$
  $$=\epsilon(f',p') $$
  Hence, $\epsilon$ is monotone. To show monotonicity of the transposition, we assume
  $f:X\timesP\to Q$ to be monotone and $x\leq x'$ for some $x,x' \in X$.
  $$\lambda f(x) \leq \lambda f(x') \in Q^P$$
  changes for any $p \in P$ to
  $$\lambda f(x)(p)\leq \lambda f(x')(p)$$
  which is by defintionß
  $$f(x,p)= \lambda f (x)(p) \leq \lambda f(x')(p) = f(x',p),$$
  so the proposition holds by equality $f(x,p)=\lambda f (x)(p)$ and assumption.
  \end{itemize}
  \newpage

 \subsection{\textbf{HA}}
\begin{definition}{Heyting Algebra}
  A Heyting Algebra is a poset with
  \begin{itemize}
  \item finite meets: 1 and $p \land q$,
  \item finite joint: 0 and $p \lor q$,
  \item exponential object $a \Rightarrow b$, such that $a \land b \leq c$ iff $a \leq b \Rightarrow c$.
  \end{itemize}

\end{definition}
Since products and exponentials are given in the definition, the remainung task is, whether
 $$\textbf{HA}(a \land b, c) \cong \textbf{HA}(a, b \Rightarrow c)$$ holds.
\\
Heyting Algbras are partially ordered sets, so as in the example for \textbf{posets}, the monotonicity constraint must hold for exponentials $b^a= a \Rightarrow q$.
In the following, exponentation is written as $a \Rightarrow b = \neg a \lor b$, when necessary.
\\
So the monotonicity constraint imposes $$ a \Rightarrow b \leq b \Rightarrow c \ iff \ a \leq b \ and \ b \leq c.$$
Hence $$a \Rightarrow b  \leq a \Rightarrow b$$
$$ a \Rightarrow b \land \a \leq b$$ by the condition $a \land b \leq c$ iff $a \leq b \Rightarrow c$, which is the evaluation arrow, and it is monotone
$$ a \Rightarrow b \land a = (\neg a \lor b) \land a$$
$$ = (\neg a \land a) \lor (b \land a)$$
$$ = 0 \lor b \land a \leq b$$ what always holds.

\\
The transpose $$a\leq b \Rightarrow c $$ is monotone as well, as shown by these inequations:
$$ a \land b \leq c \ to \neg b \lor (a \land b) \leq \neg b \lor c $$
$$ (\neg b \lor a) \land (\neg b \lor b) \leq b \Rightarrow c$$
$$ \neg b \lor a \leq (\neg b \lor a) \land (\neg b \lor b) $$
$$ a \leq \neg b \lor a$$
Hence
$$ a \leq b \Rightarrow c.
$$
\subsection{Counterexamples}

One striking property of any cartesian closed categories $\mathcal{A}$ with an initial object 0 and some object A is, that the isomorphism $0 \times A \cong 0$ holds,
because $$Hom(0 \times A, -) \cong Hom(0, (-)^A) \cong \{ \ast \}$$
where {$\ast$} denotes the one object, one arrow category, which is isomorphic to 0.
 \\
 So, to dismiss cartesian closedness for some category $\mathcal{B}$ with object A and an initial object C, it suffices to show, that $0 \times A \cong 0$ does not hold.
So in the case of $\boldsymbol{Vect_k}$ with the initial object 0:= ({x},$\lambda \cdot x=x$, x+x=x), the binary product is the direct product of two vectorspaces, but $0 \oplus A \cong A$. Therefore, $\boldsymbol{Vect_k}$ is not cartesian closed.
 \\
The identity above can be exploited to argue, that pointed categories, where the terminal object is identical to the initial object, are only CCC if they are trivial.
Because assuming \mathcal{A} to be a pointed category yields $$Hom(A,B) \cong Hom(1,A^B) \cong Hom(0,A^B) \cong 1$$  $\forall A,B \in ob(\mathcal{A})$.
With this argument, the category of relations \textbf{Rel} or of groups \textbf{Grp} can be ruled out to be a CCC, because both are pointed with the empty set, resp. the trivial group as zero objects.

\section{Equational Characterisation}
Finally, there is a more compact characterisation to identify a CCC $\mathcal{A}$, by offering a few equations which must hold for $\mathcal{A}$.
The first two arise from the first constraint that $\mathcal{A}$ has all finite products, the remaining from the exponentials.
They sometimes offer a shortcut to verifying cartesian closedness.
\subsection{Terminal objects}
$\mathcal{A}$ must have a terminal object 1. The reason is, that the terminal object is the limit of the empty product.
If there is no terminal object, there is no empty product and hence not all finite products.
\subsection{Binary products are objects of $\mathcal{A}$}
$\forall A,B,Z\in \mathcal{A}$, there is given an object $A \times B$ and arrows $p_1:A \times B \to A$, $p_2:A \times B \to B$ and for each pair of arrows <f,g> with $f:Z \to A$ and $g:Z \to B$ such that
$$ p_1<f,g> = f, p_2<f,g>=g$$ and $$<p_1 \circ  h,P_2 \circ h> = h, \forall h : Z \to A \times B $$ holds.
\subsection{Evaluation Arrow and Transpose}
$\forall A, B. Z\in \mathcal{A}$, there is  an object $B^A$ and an arrow $\epsilon: B^A \times A \to B$ and for each arrow $f:Z \times A \to B$, there is an transpose $\lambda f: Z \to B^A$ such that the equation:
$$(i) \quad \epsilon \circ (\lambda f \circ 1_A) = f$$
and $\forall g: Z \to B^A$:
$$(ii) \quad \lambda (\epsilon \circ (g \circ 1_A))=g$$
with the identity functor on A $1_A: A \to A$, hold.
\\
For that these equations indeed imply exponentials, we have to verify the properties of exponentials:
\begin{itemize}
\item  $- \times A$ is left adjoint to $(-)^A$:
\\Define $\lambda^{-1}: Hom(Z, B^A) \to Hom(Z \times A,B)$ with $\lambda^{-1}(f) = \epsilon \circ (f \times 1_A), f: Z \to B^A$.
Now verify the equation $(\lambda \circ \lambda^{-1}) = (\lambda^{-1} \circ \lambda) = 1_f$ by unfolding definitions.
$$(\lambda \circ \lambda^{-1})f=\lambda(\epsilon \circ (f \times 1_A)) =f \quad by \quad ii$$ 
$$(\lambda^{-1} \circ \lambda)f=\epsilon \circ (\lambda f \times 1_A) =f \quad by \quad i$$
\item $Hom(- \times B, -) \cong Hom(-,(-)^B)$ is natural:
  \\
  The naturality axiom consists of two identities:
  \\
  For categories $\mathcal{A}, \mathcal{B}$, for all objects $A,A' \in ob(\mathcal{A}), B,B' \in ob(\mathcal{B})$, functors $F: \mathcal{A} \to \mathcal{B}, G: \mathcal{B} \to \mathcal{A}$,
  mappings $g:F(A) \to B, f: A \to G(B)$  and transpositions $\lambda:A' \to G(B'),\mu: F(A') \to B'$,
  \[
  \begin{tikzcd}[column sep=normal]
    \lambda(F(A) \arrow{r}{g} & B \arrow{r}{q} & B') = A \arrow{r}{\lambda g} & G(B) \arrow{r}{G(q)} & G(B') 
  \end{tikzcd}
  \]

   \[
   \begin{tikzcd}[column sep=normal]
    \mu(A' \arrow{r}{p} & A \arrow{r}{f} & G(B)) = F(A') \arrow{r}{F(p)} & F(A) \arrow{r}{\mu f} & B 
  \end{tikzcd}
   \]

   \\ So we instantiate this with the category $\mathcal{A}$ having exponential objects $B^A, A,B \in ob(\mathcal{A})$ , $\epsilon: B^A \times A \to B$, and an invertible transpose $\lambda$.
   Lets define $$F(Z):=Z \times B, G(Z):= Z^B$$. 
   For g, we take $1_B$, for $\epsilon$ and have
   $$1_B \circ \epsilon: A \times B \to B \to B $$ with the transpose
   $$\lambda (1_B \circ \epsilon) : A \to (B \to B)^B = A \to B^B \to B^B$$ which is the type of
   $$(1_B)^B \circ \lambda\epsilon$$ what instantiates the first axiom as
   \[
   \begin{tikzcd}[column sep=normal]
     \lambda(A \times B \arrow{r}{\epsilon} & B \arrow{r}{1_B} & B) = A \arrow{r}{\lambda \epsilon} & B^B \arrow{r}{(1_B)^B} & B^B 
   \end{tikzcd}
   \]

   \\ For the second axiom, choose $f:A\to B^B$ as and $p:=1_A$.Let $\mu := \lambda^{-1}$
   $$f \circ 1_A : A \to A \to B^B$$
   the transpose has type
   $$\lambda^{-1}: A\timesB \to B $$
   what happens to be the type of
   $$1_A \times A \circ \lambda^{-1} f,$$ with
   $\lambda^{-1}f : A\times B \to B$
   So the diagram
   \[
   \begin{tikzcd}[column sep=normal]
     \lambda^{-1}(A \arrow{r}{1_A} & A \arrow{r}{f} & B^B) = A \times B \arrow{r}{1_A \times B} & A \times B \arrow{r}{\lambda^{-1} f} & B 
   \end{tikzcd}
   \]
   matches the second axiom.

\item $ (-)^B$ is functorial.
  For the sake of the argument, we put our knowledge about exponentials being right adjoints to the side and explore, whether the object $B^A$ is a also a functor from scratch.
  To be functorial, the exponential object must preserve identity arrows and composition.
  So we start by defining a functor $\beta:B \to C$ by transposing the composite $\beta \circ \epsilon$ with
  $$ \beta^A: B^A \to C^A$$
  and
  $$ \beta^A := \lambda (\beta \circ \epsilon),$$
  which is made more explicit with the following diagram:
\[
\begin{tikzcd}[row sep=huge]
    C^A                    & C^A \times A \arrow{r}{\epsilon} & C \\
    B^A \arrow{u}{\beta^A} & B^A \times A \arrow{u}{\beta^A \times 1_A} \arrow{r}{\epsilon} & B \arrow{u}{\beta}
  \end{tikzcd}
  \]

  \begin{itemize}
  \item To show that the exponential object preserves the identity functor, we replace $\beta \quad by \quad 1_B$, what transforms the diagram to:
    \[
\begin{tikzcd}[row sep=huge]
    B^A \times A \arrow{r}{\epsilon} & B \\
    B^A \times A \arrow{u}{\beta^A \times 1_A} \arrow{r}{\epsilon} & B \arrow{u}{\beta}
  \end{tikzcd}
\]
From which follows $(1_B)^A = 1_{B^A} : B^A \to B^A$ because
$$ 1_B \circ \epsilon = \epsilon$$
$$ = \epsilon \circ 1_{B^A \times A}$$
$$ = \epsilon \circ 1_{B^A} \times 1_A.$$
\item To show the preservation of the composite, we add another functor $\gamma$:
  \[
  \begin{tikzcd}[row sep=huge]
    B \arrow{r}{\beta} & C \arrow{r}{\gamma} & D
  \end{tikzcd}
  \]
  Having this diagram:
  \[
  \begin{tikzcd}[row sep=huge]
    D^A \times A \arrow{r}{\epsilon} & D \\
    C^A \times A \arrow{u}{\gamma^A \times 1_A} \arrow{r}{\epsilon} & C \arrow{u}{\gamma} \\
    B^A \times A \arrow{u}{\beta^A \times 1_A} \arrow{r}{\epsilon} & B \arrow{u}{\beta}
  \end{tikzcd}
  \]
  The identity
  $$\gamma^A \circ \beta^A = (\gamma \circ \beta)^A. $$ follows from
  $$\gamma \beta \epsilon = \gamma \epsilon (\beta^A \times 1_A) \ by \ Def. \ of \ \beta^A$$
  $$=\epsilon (\gamma^A \times 1_A)(\beta^A \times 1_A) \ by \ Def. \ of \ \gamma^A$$
  $$=\epsilon ((\gamma^A \circ \beta^A) \times 1_A)$$

  \end{itemize}
\end{itemize}

\subsection{\textbf{Cat}}
There is yet another example, which exemplifies the usage of the third part of the equational characterisation to verify cartesian closedness.
We consider the category of categories with finite products.
Since the first CCC requirement is given by definition, it remains to show, that the category is equipped with exponential objects.

  \\
  So $\forall A,B,C \in \textbf{Cat}$, the exponential object is given as $C^B$, with the evaluation arrow
  $$\epsilon: C^B \times B \to C $$ and the transposition
  $$\lambda f: A \to C^B,\forall f:A \times B \to C$$
  The equation
  $$f = \epsilon \circ (\lambda f \times 1_B)$$ can be verified with the commuting diagram
  \[
  \begin{tikzcd}[row sep=huge]
    A \arrow{d}{\lambda f} & A \times B \arrow{d}{\lambda f \times 1_B} \arrow{dr}{f} \\
    C^B                    & C^B \times B \arrow{r}{\epsilon} & C 
  \end{tikzcd}
  \]
  \\
  For some $g:A\toC^B$, we have the inverse of the transpose
  $$\lambda^{-1}g: A\times B \to C$$ with
  $$\lambda^{-1}g:=\epsilon \circ (g\times 1_B)$$
  and need to show, that
  $$g = \lambda(\epsilon \circ (g\times 1_B).$$
  Since we already know, that $(\lambda \circ \lambda^-1) (g) = g$, there is nothing left to show and \textbf{Cat} is cartesian closed.

\section{Further Properties}

Beyond the already seen isomorphism $0 \times A \cong 0$, many identities knwon from common arithmetics,like for instance that of the natural numbers,
hold for cartesian closed categories.$\forall A, B, C \in $ in some cartesian closed category $\mathcal{A}$ with terminal object "1" and binary products "$\times$": \\
\begin{itemize}
\item $A \times (B \times C) \cong (A \times B) \times C$
\item $A \times B \cong B \times A$
\item $ 1^A \cong 1 $
\item $ A^1 \cong A$
\item $(A \times B)^C \cong A^C \times B^C$
\item $(A^B)^C \cong A^{B \times C} $
\end{itemize}
\\
If $\mathcal{A}$ is bicartesian closed with initial object "0" and co-product "+", the enumeration of axioms above is extended by: 
\begin{itemize}
\item $A+B \cong B+A$
\item $(A+B)+C \cong A+(B+C)$
\item $(A+B)\times C \cong A \times C+B \times C$
\item $A^{B+C} \cong A^B \times A^C $
\item $0+A \cong A$
\item $A^0 \cong 1$
\end{itemize}
\\
We are going to show some of these identities by transformations exploiting right or left adjointness.
Usually it suffices to show, that both sides satisfy the same universal property.
\begin{itemize}
\item $$1^A \cong 1.$$
  holds, because there is only one arrow from every object A $\in ob(\mathcal{A})$ to the terminal object 1 which makes Hom(A,1) containing exactly one arrow. The terminal object itself is merely equipped with the identity $1_1$.
  So both sides satisfy the universal property and are therefore isomorphic.
  \\

\item $$(A^B)^C \cong A^{B \times C}.$$
  Take some X,Y $\in ob(\mathcal{A})$ and set
  $$ X:=Hom(-, (A^B)^C) \quad and \quad Y:= Hom(-,A^{B\timesC})$$ then
  $$ Hom(-, (A^B)^C) \cong Hom(-\times C, A^B) $$
  $$ \cong Hom((- \times C)\times B,A) $$
  With B, associativity and commutativity of the product
  $$ Hom(-,A^{B\times C}) \cong Hom(- \times (B\times C),A) $$
  $$ \cong Hom(-\times (C\times B),A) $$
  $$ \cong Hom((-\times C)\times B,A) $$
  So $X \cong Y$, what shows the claim.
\item $$A^0 \cong 1$$
  The argument is similar to the one above. There is only one arrow from the initial object 0 to any other object A, so Hom(0,A) is isomorphic to the terminal object 1.
\end{itemize}

Other equations like distibutivity appear to be more difficult without any further assumptions and hence will be omitted, for there is an easy way using the Yoneda Lemma which is not the topic of this write up.

\\
Subcategories $\mathcal{B}$ of cartesian closed categories $\mathcal{A}$ are in general not cartesian closed.
A simple counterexample is the category of categories with finite products\textbf{Cat}, which is cartesian closed with binary products and exponentials. 
Yet $\textbf{Rel} \in ob(\textbf{Cat})$ is true, whereas \textbf{Rel} is not a CCC.

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2017-09-21 papers/talk-10/notes-10.tex
Rearranged files into directory papers

--- 2017-07-05 slides/slides-10/notes-10.tex
@@
 \documentclass{article}
+\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
 
-\usepackage[utf8]{inputenc}
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
 \newtheorem{lemma}{Lemma}
@@
 \newtheorem{corollary}{Corollary}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
 
-% \beamertemplatenavigationsymbolsempty
-% \setbeamertemplate{footline}{
-% 	\raisebox{0.5em}{\makebox[\paperwidth-0.5em][r]{
-% 		\color{gray}\scriptsize\insertframenumber}}}
+\usetikzlibrary{cd}
+\usepackage{subfig}
 
-\usepackage{mathtools}
-\usepackage{tikz}
-\usetikzlibrary{cd, shapes, tikzmark}
-\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}
-
+%\setcounter{equation}{0}
+\numberwithin{equation}{section}
+\begin{document}
 \title{Presheaves, Representables and\\the Yoneda Lemma}
@@
 %\date{June 7, 2017}
-
-\begin{document}
+\maketitle
 
@@
   \end{align*}
-\item Left adjoints are unique up to isomorphisms;
+\item Right adjoints are unique up to isomorphisms;
 \item For any small category $\cat A$, the category of presheaves $[\cat A^\op,\Set]$ is cartesian closed.
@@
 
-In order to get a first understanding what this is supposed to mean we consider the example of a partially order set $(P,\leq)$. Let $p,p'\in P$ be arbitrary. Then we clearly have 
+In order to get a first understanding what this is supposed to mean we consider the example of a partially ordered set $(P,\leq)$. Let $p,p'\in P$ be arbitrary. Then we clearly have 
 \begin{align*}
@@
 or in other terms
-\begin{align*}
-  (p\leq p')\iff \downarrow(p)=\{q\in P\mid q\leq p\}\subseteq \{q\in P\mid q\leq p'\}=\downarrow(p'). *
-\end{align*}
+\begin{align}
+  \label{eq:pre}
+  (p\leq p')\iff {\downarrow}(p)=\{q\in P\mid q\leq p\}\subseteq \{q\in P\mid q\leq p'\}={\downarrow}(p').
+\end{align}
 
 How can we lift this to arbitrary categories?
-We can do so by noting that the two sets correspond to the maps into $p$ and $q$ respectively.
-However, since maps are not unique in general categories we have to do a little bit more work.
-Recall the following special functor introduced in \ref[?]:
+We can do so by noting that the two sets correspond to the maps into $p$ and $p'$ respectively.
+However, since maps between two given objects are not unique in general categories, we have to do a little bit more work.
+
+Recall the following special functor introduced in chapter 2:
 
 \begin{definition}[Contravariant Representable Functor]
-  Let $\cat A$ be a locally small category and $A\in \cat A$. The \emph{contravariant representable functor} of $A$
+  Let $\cat A$ be a locally small category and $A\in \cat A$.
+ The \demph{contravariant representable functor} of $A$
   \begin{align*}
@@
   \item $H_A(B)=\cat A(B,A)$ for objects $B\in\cat A$ and
-  \item for maps $B'\xrightarrow{f} B$:
+  \item for maps $f\from B'\to B$:
     \begin{align*}
@@
 
-Clearly, for each $p$ in a preorder regarded as a category $\downarrow(p)$ is essentially the same as $H_p$.
-Furthermore, inclusion are essentially natural transformations.
-Then, * corresponds to the fact that the following functor is full and faithful:
+Clearly, for each $p$ (in a preorder regarded as a category) ${\downarrow}(p)$ is essentially the same as $H_p$.
+Furthermore, the inclusions in equation \eqref{eq:pre} correspond to the existence of a natural transformation:
+\begin{align*}
+  &{\downarrow}(p)=\{q\in P\mid q\leq p\}\subseteq \{q\in P\mid q\leq p'\}={\downarrow}(p')\\
+  \iff &\forall q\in P\ldotp q\leq p\rightarrow q\leq p'\\
+  \iff &\forall q\in P\ldotp\exists \eta_q\from H_p(q)=P(q,p)\to P(q,p')=H_{p'}(q)\ldotp\top\\
+  \iff&\exists \eta\from H_p\to H_q\ldotp\top.
+\end{align*}
+
+Thus, the condition in equation \eqref{eq:pre} corresponds to the fact that the following functor is full and faithful:
 
 \begin{definition}[Yoneda Embedding]
-  Let $\cat A$ be locally small. The \emph{Yoneda embedding} of $\cat A$ is the functor
+  Let $\cat A$ be locally small. The \demph{Yoneda embedding} of $\cat A$ is the functor
   \begin{align*}
@@
   \item $H_-(A)=H_A$ for objects $A\in\cat A$ and
-  \item for maps $A\xrightarrow{f}A'$, $H_f$ is the natural transformation defined for $B\in \cat A$ by
-    \begin{align*}
+  \item for maps $f\from A\to A'$, $H_f$ is the natural transformation defined for $B\in \cat A$ by
+    \begin{align}
+      \label{eq:pre2}
       H_A(B)=\cat A(B,A)&\to H_{A'}(B)=\cat A(B,A')\\
-      g&\mapsto  f\of g **\\
-    \end{align*}
+      g&\mapsto  f\of g \nonumber
+    \end{align}
   \end{itemize}
-  \vspace{-0.8cm}
   \includegraphics{nattrans.pdf}
 \end{definition}
-Basically, for preorders ** constructs a proof from $p\leq p'$ that for all $q\leq p$ also $q\leq p'$ holds.
+Basically, for preorders \eqref{eq:pre2} constructs a proof from $p\leq p'$ that for all $q\leq p$ also $q\leq p'$ holds.
 
-In what follows we prove that this is indeed an embedding, i.e. full and faithful.
+In what follows we prove that this is indeed an embedding, i.e. full and faithful, in the general case.
 
 \section{Yoneda Lemma}
-To do so we state the following technical lemma due to Yoneda:
+To do so, we state the following technical lemma due to Yoneda:
 \begin{lemma}[Yoneda]
@@
   \end{align*}
-  naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$. That is for each $B\xrightarrow{f}A$ in $\cat A$ and for each natural transformation $\theta \from F\to F'$ the following commute
+  naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$. That is for each $f\from B\to A$ in $\cat A$ and for each natural transformation $\theta \from F\to F'$ the following commute
   \begin{align*}
@@
   \end{enumerate}
-  It turns out that the definitions of $(\widehat{-})$ and $(\widetilde{-})$ are straight-forward as there is only one canonical choice that can possibly be made:
+  It turns out that the definitions of $(\widehat{-})$ and $(\widetilde{-})$ are straightforward as there is only one canonical choice that can possibly be made:
   \begin{enumerate}
@@
 In order to define $(\widehat{-})\from[\cat A^\op,\Set](H_A,F)\to F(A)$, we have to come up with an element $\widehat\alpha\in F(A)$.
-However, $\alpha$ gives us a function $\alpha_A\from H_A(A)\to F(A)$ in $\Set$.
+However, $\alpha$ provides us with a function $\alpha_A\from H_A(A)\to F(A)$ in $\Set$.
 Hence, we only have to come up with an element of $H_A(A)=\cat A(A,A)$ to obtain an element of $F(A)$ and we can simply choose $1_A$.
@@
 This is perfect since we already have $a\in F(A)$.
-Thus, $F(f)(a)\in F(A)$ and we can define
+Thus, $F(f)(a)\in F(B)$ and we can define
     \begin{align*}
@@
 
-    We still have to check naturality, i.e. that for any $B'\xrightarrow{g}B$ the following commutes:
+    We still have to check naturality, i.e. that for any $g\from B'\to B$ the following commutes:
     \begin{align*}
-    \xymatrix{
-    H_A(B)=\cat A(B,A) \ar[r]^{H_A(g)=-\of g} \ar[d]^{\widetilde a_B} & H_A(B')=\cat A(B',A) \ar[d]_{\widetilde a_B'} \\
+    \xymatrix@+2pc{
+    H_A(B)=\cat A(B,A) \ar[r]^{H_A(g)=-\of g} \ar[d]^{\widetilde a_B} & H_A(B')=\cat A(B',A) \ar[d]_{\widetilde a_{B'}} \\
     F(B) \ar[r]^{F(g)} & F(B').
@@
   \end{align*}
-  Since this square is in $\Set$, it suffices to check $(\widetilde a_B'\of H_A(g))(f)=(F(g)\of\widetilde a_B)(f)$ for every $f\in H_A(B)=\cat A(B,A)$. Indeed, this holds:
+  Since this square is in $\Set$, it suffices to check $(\widetilde a_{B'}\of H_A(g))(f)=(F(g)\of\widetilde a_B)(f)$ for every $f\in H_A(B)=\cat A(B,A)$. Indeed, this holds:
   \begin{align*}
-    (\widetilde a_B'\of H_A(g))(f)&=(F(f\of g))(a)\\
+    (\widetilde a_{B'}\of H_A(g))(f)&=(F(f\of g))(a)\\
     &=(F(g)\of F(f))(a)&F \text{ contravariant functor}\\
@@
   \end{align*}
-\item[(d)-(f)] The remaining proof obligations are quite straight-forward to show and are thus left as an exercise.
+\item[(d)-(f)] The remaining proof obligations are quite straightforward to show and are thus left as an exercise.
   \end{enumerate}
@@
 
-Now, it is easy to show that $H_-$ is full and faithful by instantiating $F$ with a covariant representable functor.
+Now, it is easy to show that $H_-$ is full and faithful by instantiating $F$ with a contravariant representable functor.
 \begin{theorem}
@@
   \begin{align*}
-    \cat A(A,A')&\to[\cat A^\op,\Set](H_A,H_A')\\
+    \cat A(A,A')&\to[\cat A^\op,\Set](H_A,H_{A'})\\
     f&\mapsto H_f
@@
   \begin{align*}
-    (\widetilde{-})\from F(A)\to[\cat A^\op,\Set](H_A,F)
+    (\widetilde{-})\from H_{A'}(A)\to[\cat A^\op,\Set](H_A,H_{A'})
   \end{align*}
@@
 \begin{lemma}
-  Let $F\from\cat A\to\cat B$ be a full and faithful functor and $A\xrightarrow{f} A'$ in $\cat A$.
+  \label{lem:ff}
+  Let $F\from\cat A\to\cat B$ be a full and faithful functor and $f\from A\to A'$ in $\cat A$.
 
@@
 \end{quote}
-Note that it is really necessary for the isomorphism to be natural in $B\in\cat A$. As the following example illustrates:
-TODO
+Abstractly speaking, the Yoneda embedding represents arbitrary locally small categories as presheaf categories, which have much structure. In particular they are cartesian closed.
+
+Note that it is really necessary for the isomorphism to be natural in $B\in\cat A$. This is illustrated by the following example:
+\begin{example}
+  Consider the following category $\cat A$: 
+\begin{align*}
+  \xymatrix{
+  A \ar@(ul,ur)^{1_A,f} \ar@/^0.3pc/@[][r]^{f_1,f_2} & B \ar@/^0.3pc/@[][l]^{g_1,g_2} \ar@(ul,ur)^{1_B,g}
+                                                       } 
+  \end{align*}
+  i.e. between any pair of object there are two arrows, where the composition of $f_i$ and $g_j$ ($i,j\in\{1,2\}$) is given by
+  \begin{align*}
+    f_i\of g_j&=g
+  \end{align*}
+  and the remaining compositions can be chosen arbitrarily.
+
+  Obviously,
+  \begin{align*}
+    \cat A(C,A)\iso\cat A(C,B)
+  \end{align*}
+  for all $C\in\cat A$ because all homsets contain exactly two elements.
+However $A\not\iso B$ because by construction neither $g_1$ nor $g_2$ (which are the only maps between $B$ and $A$) have a left inverse.
+\end{example}
+
+On the other hand, in special categories it is not even necessary to consider the mappings from \emph{all} other objects.
+$\Set$ is an extreme such case: Due to the fact that $A\iso\Set[1,A]=H_A(1)$, we have $A\iso A'$ if and only if $H_A(1)\iso H_{A'}(1)$, for all sets $A$ and $A'$.
+
 \section{Consequences of the Yoneda Principle}
-Next, we use the insights just gained and we consider (further) consequences of the Yoneda lemma/principle. In particular, we are going to proof the facts mentioned in the introduction.
+Next, we use the insights just gained and we consider (further) consequences of the Yoneda lemma/principle. In particular, we are going to prove the facts mentioned in the introduction.
 
@@
 
-Another rather easy consequence is that left adjoints are unique:
+Another rather easy consequence is that right adjoints are unique:
 
 \begin{proposition}
-  Let $F,F'\from \cat A\to\cat B$ and $G\from\cat B\to\cat A$ such that $F\dashv G$ and $F'\dashv G$. Then $F\iso F'$
+\label{prop:unique}
+  Let $F\from \cat A\to\cat B$ and $G,G'\from\cat B\to\cat A$ such that $F\dashv G$ and $F\dashv G'$. Then $G\iso G'$
 \end{proposition}
-\begin{proof}
-  By adjointness we have 
+With the following lemma the proof becomes very nice, indeed:
+\begin{lemma}
+  \label{lem:comp}
+  Let $\cat B$, $\cat C$ and $\cat D$ be categories and $F\from\cat C\to\cat D$ a functor. There is an induced functor
   \begin{align*}
-    \cat B^\op(B,F(A))\iso\cat B(F(A),B)\iso\cat A(A,G(B))\iso\cat B(F'(A),B)\iso\cat B^\op(B,F'(A))
+    F\of -\from[\cat B,\cat C]\to [\cat B,\cat D]
   \end{align*}
-  naturally in $B\in\cat B$. Hence, by what was stated above we have $F(A)\iso F'(A)$ for all $A\in\cat A$. It can be shown that this isomorphsim is natural in $A$ (why?), hence $F\iso F'$.
+  defined by composition with $F$. 
+Then, $F\of -$ is full and faithful if $F$ itself has these properties.  
+\end{lemma}
+\begin{proof}[Proof of proposition \ref{prop:unique}]
+
+  By adjointness we get
+    \begin{align*}
+      H_{G(B)}(A)=\cat A(A,G(B))\iso\cat B(F(A),B)\iso\cat A(A,G'(B))=H_{G'(B)}(A)
+    \end{align*}
+    naturally in $A\in\cat A$ and $B\in\cat B$. In particular,
+    \begin{align*}
+      (H_-\of G)(B)=H_{G(B)}\iso H_{G'(B)}=(H_-\of G')(B)
+    \end{align*}
+    naturally in $B$, i.e.
+$H_-\of G\iso H_-\of G'$.
+Due to the lemmas \ref{lem:ff} (``isomorphisms in the image of full and faithful functors are the image of isomorphisms'') and \ref{lem:comp}, as well as the fact that the Yoneda embedding $H_-$ is full and faithful, $G\iso G'$ holds.
+  % By adjointness we have 
+  % \begin{align*}
+  %   \cat B^\op(B,F(A))\iso\cat B(F(A),B)\iso\cat A(A,G(B))\iso\cat B(F'(A),B)\iso\cat B^\op(B,F'(A))
+  % \end{align*}
+  % naturally in $B\in\cat B$. Hence, by what was stated above we have $F(A)\iso F'(A)$ for all $A\in\cat A$. It can be shown that this isomorphsim is natural in $A$ (why?), hence $F\iso F'$.
 \end{proof}
+Similarly, it can be shown that left adjoints are unique up to isomorphisms.
 
@@
 Showing that exponentials exist is still non-trivial.
-However the Yoneda lemma provides a straight-forward definition of exponentials and in a special case verifying that this is indeed an exponential is simple.
+However the Yoneda lemma provides a straightforward definition of exponentials and in a special case verifying that this is indeed an exponential is simple.
 In the general case we need a couple of additional lemmas.
@@
 Thus, we are left to show that exponentials exists, i.e. that for functors $G,H\in[\cat A^\op,\Set]$, there is a functor $H^G\in[\cat A^\op,\Set]$ such that
-\begin{align*}
-  [\cat A^\op,\Set](F,H^G)\iso [\cat A^\op,\Set](F\times G,H)***
-\end{align*}
+\begin{align}
+  \label{eq:rem}
+  [\cat A^\op,\Set](F,H^G)\iso [\cat A^\op,\Set](F\times G,H)
+\end{align}
 naturally in $F\in[\cat A^\op,\Set]$.
 
-To define $H^G$ on objects $A\in\cat A$ note that by the Yoneda lemma we have
+To define $H^G$ on objects $A\in\cat A$, note that by the Yoneda lemma we have
 \begin{align*}
@@
 \end{align*}
-Clearly this set exists and we can thus define
+Clearly, this set exists and we can thus define
 \begin{align*}
@@
 on objects.
-For maps $A\xrightarrow{f}A'$ we define $H^G(f)=H^G_f\from [\cat A^\op,\Set](H_{A'}\times G,H)\to[\cat A^\op,\Set](H_A\times G,H)$ by
-$H^G_f(\alpha)\in[\cat A^\op,\Set](H_A\times G,H)$ for $\alpha\in [\cat A^\op,\Set](H_{A'}\times G,H)$ by
+For maps $f\from A\to A'$ we define 
+\begin{align*}
+  H^G(f)=H^G_f\from [\cat A^\op,\Set](H_{A'}\times G,H)\to[\cat A^\op,\Set](H_A\times G,H)
+\end{align*}
+ by
+$H^G_f(\alpha)\in[\cat A^\op,\Set](H_A\times G,H)$ for $\alpha\in [\cat A^\op,\Set](H_{A'}\times G,H)$, which is a natural transformation and which in turn is defined by
 \begin{align*}
@@
 \end{align*}
-for $B\in\cat A$, which is of course natural.
+for $B\in\cat A$. This is of course natural.
 
-Therefore, we are left to verify that (***) holds and that it is natural in $F\in[\cat A^\op,\Set]$.
-In order to understand the proof better let us first consider the special case where $F=H_A$ is a covariant representable functor for some $A\in\cat A$. Then we immediately have
+Therefore, we are left to verify that \eqref{eq:rem} holds and that it is natural in $F\in[\cat A^\op,\Set]$.
+In order to understand the proof better let us first consider the special case where $F=H_A$ is a covariant representable functor for some $A\in\cat A$. Then, we immediately have
 \begin{align*}
   [\cat A^\op,\Set](H_A,H^G)&\iso H^G(A)&\text{Yoneda lemma}\\
-  &=[\cat A^\op,\Set](H_A\times G,H)&\text{definition of }H^G.
+  &=[\cat A^\op,\Set](H_A\times G,H).&\text{definition of }H^G
 \end{align*}
@@
 \begin{proof}
-  First of all, we have to define the shape of the limit, i.e. a small category $\mathbf I$. We take the category of elements of $F$, $F(\cat A)$ with
+  First of all, we have to define the shape of the limit, i.e. a small category $\mathbf I$. We take the \emph{category of elements} of $F$, $\int_{\cat A}F$ with
   \begin{itemize}
@@
   \end{itemize}
-  TODO: figure
 
-  and identities and composites defined in the natural way. Clearly, $F(\cat A)$ is small since $\cat A$ is small.
+\begin{figure}
+  \subfloat[A category $\cat A$ with two objects and one map.]
+  {%
+    \begin{tikzpicture}
+      \draw[draw=white] (-1.2,0) ellipse (0.75cm and 1.3cm);
+      \node[] at (-1.2, 0)     (A) {$A$};
+      \node[] at (1.2, 0)    (B) {$B$};
+      \path[draw,->] (A) -- (B);
+    \end{tikzpicture}
+  }\hfill
+  \subfloat[Its image under a presheaf $F$, where $F(A)=\{a_1,a_2,a_3\}$ and $F(B)=\{b_1,b_2\}$, as well as $b_1\mapsto a_1$ and $b_2\mapsto a_3$.]
+  {
+    \begin{tikzpicture}
+      \draw (-1.2,0) ellipse (0.75cm and 1.3cm);
+      \draw (1.2,0) ellipse (0.75cm and 1.3cm);
+      \node[] at (-1.2, 0)   (fa) {$F(A)$};
+      \node[] at (1.2, 0)   (fb) {$F(B)$};
+      \path[draw,<-] (-0.4,0) -- (0.4,0);
+    \end{tikzpicture}
+  }\hfill
+\subfloat[The corresponding category of elements $\int_{\cat A}F$.]
+  {
+    \begin{tikzpicture}
+      \draw[dashed] (-1.2,0) ellipse (0.75cm and 1.3cm);
+      \draw[dashed] (1.2,0) ellipse (0.75cm and 1.3cm);
+      \node[] at (-1.25, 0.6)   (aA1) {$(a_1,A)$};
+      \node[] at (-1.25, 0)     (aA2) {$(a_2,A)$};
+      \node[] at (-1.25, -0.6)  (aA3) {$(a_3,A)$};
+      \node[] at (1.25, 0.3)    (bB1) {$(b_1,B)$};
+      \node[] at (1.25, -0.3)   (bB2) {$(b_2,B)$};
+      \path[draw,->] (aA1) -- (bB1);
+      %\path[draw,->] (aA2) -- (bB1);
+      \path[draw,->] (aA3) -- (bB2);
+    \end{tikzpicture}
+  }
 
-  Next, we have to define a diagram $F(\cat A)\to[\cat A^\op,\Set]$ such that $F$ is the corresponding limit.
-We simply take $H_-\of\pi$ where $\pi\from F(\cat A)\to\cat A$ is the projection on $\cat A$ defined by
+  \caption{\label{fig:ex}Visualization of a category and its corresponding category of elements (identities are omitted).}
+\end{figure}
+
+  and identities and composites are defined in the natural way.
+Figure \ref{fig:ex} contains a visualization of a category and its corresponding category of elements.
+Clearly, $\int_{\cat A}F$ is small since $\cat A$ is small.
+
+  Next, we have to define a diagram $\int_{\cat A}F\to[\cat A^\op,\Set]$ such that $F$ is the corresponding limit.
+We simply take $H_-\of\pi$ where $\pi\from \int_{\cat A}F\to\cat A$ is the projection on $\cat A$ defined by
   \begin{itemize}
@@
 
-  Now, we have to verify that $F$ is indeed the limit of $H_-\of\pi$ under $F(\cat A)$, i.e. we have to specify a cocone
+  Now, we have to verify that $F$ is indeed the limit of $H_-\of\pi$ under $\int_{\cat A}F$, i.e. we have to specify a cocone
   \begin{align*}
-    (H_{\pi(a,A)}\to F)_{(a,A)\in F(\cat A)}
+    (H_{\pi(a,A)}\to F)_{(a,A)\in \int_{\cat A}F}
   \end{align*}
@@
 
-  To define $H_{\pi(a,A)}=H_A\to F$ for any $(a,A)\in F(\cat A)$ we recall that $[\cat A^\op,\Set](H_A,F)\iso F(A)$, where the isomorphism from the right to the left is given by $(\widetilde -)$ (as defined in the proof of the Yoneda lemma).
+  To define $H_{\pi(a,A)}=H_A\to F$ for any $(a,A)\in \int_{\cat A}F$ we recall that $[\cat A^\op,\Set](H_A,F)\iso F(A)$, where the isomorphism from the right to the left is given by $(\widetilde -)$ (as defined in the proof of the Yoneda lemma).
 Thus, we define the $(a,A)$-component of the cocone to be $\widetilde a$.
-For this to be a cocone we still have to verify that for each $(b,B)\xrightarrow{f}(a,A)$ (i.e.
-$B\xrightarrow{f}A$ such that $F(f)(a)=b$)
+For this to be a cocone we still have to verify that for each $f\from (b,B)\to (a,A)$ (i.e.
+$f\from B\to A$ such that $F(f)(a)=b$)
   \begin{align*}
-    \xymatrix{
+    \xymatrix@+1.5pc{
       H_{\pi(b,B)}=H_B \ar[r]^{\widetilde b} \ar[d]_{H_{\pi(f)}=H_f} & F \\
@@
   \begin{align*}
-    \xymatrix{
+    \xymatrix@+1pc{
     [\cat A^\op,\Set](H_A,F) \ar[r]^{-\of H_f} & [\cat A^\op,\Set](H_B,F) \\
@@
   \end{align*}
-  by definition of arrows in $F(\cat A)$.
+  by definition of arrows in $\int_{\cat A}F$.
 
-  Thus, we have shown that this indeed constitues a cocone. Similarly, it can be shown that this cocone is universal, i.e. that $F$ is the limit of $H_-\of\pi$ under $F(\cat A)$.
+  Thus, we have shown that this indeed constitues a cocone. Similarly, it can be shown that this cocone is universal, i.e. that $F$ is the limit of $H_-\of\pi$ under $\int_{\cat A}F$.
 \end{proof}
 
-Now, are almost in a position where we can prove that presheaf categories are cartesian closed. All we need is the following lemma which states that we can ``pull limits out of products'':
+Now, we are almost in a position where we can prove that presheaf categories are cartesian closed.
+All we need is the following lemma (the proof of which is left as an exercise) which states that we can ``pull limits out of products'':
 \begin{lemma}
@@
 \begin{proof}
-  It easy to prove that $[\cat A^\op,\Set]$ has terminal objects and binary products (by defining them ``pointwise'').
+  It is easy to prove that $[\cat A^\op,\Set]$ has terminal objects and binary products (by defining them ``pointwise'').
 
@@
 
+\section{Conclusion}
+In this chapter we studied the Yoneda lemma and some of its consequences.
+It states that for any presheaf $F$ and object $A$, $F(A)$ is isomorphic to the set of presheaves $H_A\to F$ (and that the isomorphism is natural in $F$ and $A$).
+A consequence is that any locally small category can be embedded in a presheaf category, which has much more structure.
+Informally, two objects are isomorphic if and only if they look the same from all other objects.
+
+As applications, we proved a distributivity-property in cartesian closed categories that have copro\-ducts and we showed that right adjoints are unique up to isomorphisms.
+
+Finally, the proof that presheaf categories are cartesian closed elaborated on the statement above that we gain structure by means of the Yoneda embedding.
+
 \end{document}

### 2017-09-21 papers/talk-8/CCC.tex
Fixed some paths

@@
-\def\pathToRoot{../}\input{\pathToRoot headers/uebungHeader}
+\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}
 
@@
 \end{definition}
-\\
+
 While it is required, that the reader knows about products, the idea of exponentials is assumed to be new and needs some more explanations.
-\\
+
 Regarding the products, just assess at this point, that the notion of all finite products is exhaustively described by terminal objects and binary products, because a terminal object represents the empty product, whereas every other product can be constructed with binary products.
@@
 \begin{itemize}
-\item \underline{\textit{Products:}} As already argued in the introduction, the cartesian product yields new objects $A\timesB\in ob(\textbf{Sets})$.
+\item \underline{\textit{Products:}} As already argued in the introduction, the cartesian product yields new objects $A\times B\in ob(\textbf{Sets})$.
 Furthermore, it delivers the empty product $\emptyset \times \emptyset \to \emptyset$.
@@
 \item "$\Rightarrow$" : Assume $f: A \times B \to C$ with $f(x,y) \in A \times B \to C, x \in A, y \in B$.
-\\ By fixing $x \in A$, one obtains an arrow $$f_x$$ with $$f_x(y) \in C^B.$$ By unfixing x, $f_x$ is morphed in the curried function $$\mu f:A\toC^B$$ with $$\mu f (x)(y) \in A \to C^B,$$ such that $$\mu f(x)(y)=f(x,y).$$
+\\ By fixing $x \in A$, one obtains an arrow $$f_x$$ with $$f_x(y) \in C^B.$$ By unfixing x, $f_x$ is morphed in the curried function $$\mu f:A\to C^B$$ with $$\mu f (x)(y) \in A \to C^B,$$ such that $$\mu f(x)(y)=f(x,y).$$
 \item "$\Leftarrow$" : Assume $f': A \to C^B$. One simply defines a cartesian function $$g: A \times B \to C$$ as $$g(x,y) := f(x)(y), x \in A, y \in B$$ and is done.
@@
     \\ Hom(P,Q) := $\{f: P \to Q \mid f \quad  monotone \}$ with $\forall p,p', p\leq p' \to f p \leq f p'$.
-  We can construct a binary product $P \times Q$, which is again a poset,using pairs $$(p,q) \in ob(P\timesQ)$$ and a partial ordering $$(p,q) \leq (p',q') \quad iff\quad p \leq p' \wedge q \leq q' .$$
+  We can construct a binary product $P \times Q$, which is again a poset,using pairs $$(p,q) \in ob(P\times Q)$$ and a partial ordering $$(p,q) \leq (p',q') \quad iff\quad p \leq p' \wedge q \leq q' .$$
   From this follow monotone projections: \\
@@
   As exponential $Q^P$, we consider a set of monotone functions
-  $$Q^P:=\{f:Q\toP \mid f \quad monotone \}$$
+  $$Q^P:=\{f:Q\to P \mid f \quad monotone \}$$
   with an pointwise ordering, such that $\forall f,g \in Q^P$,
@@
   Hence, $\epsilon$ is monotone. To show monotonicity of the transposition, we assume
-  $f:X\timesP\to Q$ to be monotone and $x\leq x'$ for some $x,x' \in X$.
+  $f:X\times P \to Q$ to be monotone and $x\leq x'$ for some $x,x' \in X$.
   $$\lambda f(x) \leq \lambda f(x') \in Q^P$$
@@
 
-\\
 The transpose $$a\leq b \Rightarrow c $$ is monotone as well, as shown by these inequations:
@@
 The identity above can be exploited to argue, that pointed categories, where the terminal object is identical to the initial object, are only CCC if they are trivial.
-Because assuming \mathcal{A} to be a pointed category yields $$Hom(A,B) \cong Hom(1,A^B) \cong Hom(0,A^B) \cong 1$$  $\forall A,B \in ob(\mathcal{A})$.
+Because assuming $\mathcal{A}$ to be a pointed category yields $$Hom(A,B) \cong Hom(1,A^B) \cong Hom(0,A^B) \cong 1$$  $\forall A,B \in ob(\mathcal{A})$.
 With this argument, the category of relations \textbf{Rel} or of groups \textbf{Grp} can be ruled out to be a CCC, because both are pointed with the empty set, resp. the trivial group as zero objects.
@@
 
-   \\ So we instantiate this with the category $\mathcal{A}$ having exponential objects $B^A, A,B \in ob(\mathcal{A})$ , $\epsilon: B^A \times A \to B$, and an invertible transpose $\lambda$.
+   So we instantiate this with the category $\mathcal{A}$ having exponential objects $B^A, A,B \in ob(\mathcal{A})$ , $\epsilon: B^A \times A \to B$, and an invertible transpose $\lambda$.
    Lets define $$F(Z):=Z \times B, G(Z):= Z^B$$. 

### 2017-09-21 papers/talk-11/paper-11.tex
Improve formatting of the proofs and refer to STLC

@@
 
-As a motivation for dependent products recall how to define exponents in
+\begin{example}
+  The type for the function $A \to A$ (where $A$ is a ground type) is expressed
+  as an exponential $M(A)^{M(A)}$ in STLC with $M$ being the interpretation
+  function and $M(A)$ an object corresponding to the ground type in cartesian
+  closed category.
+\end{example}
+
+To represent the same type in the model with dependent types we will also use
+exponentials, but on the presheaves. Let us consider how to define exponents in
 presheaf categories. Let $\Gamma$ and $\Delta$ be presheaves and suppose we
@@
 \]
-where $\mathbf{y}$ is Yoneda embedding.
+where $\mathbf{y}$ is the Yoneda embedding.
 
-The latter can be taken up as a definition. So an element $w$ of
-$(\Delta^\Gamma)(I)$ is a natural transformation $w \from \mathbf{y}I\times
-\Gamma \to \Delta$, so it is given by functions
+We can take that as a definition. Then an element $w$ of $(\Delta^\Gamma)(I)$
+is a natural transformation $w \from \mathbf{y}I\times \Gamma \to \Delta$ given
+by functions
 \[
@@
 
-We will now show that CwF associated to a presheaf category supports
-$\Pi$-types. Given $\Gamma \vdash A$ and $\Gamma.A \vdash B$ we have to define
-$\Gamma \vdash \Pi x: A. B~type$, that is, we have to define the set $(\Pi x:
-A.B) \rho$ for $I \in \cat{C}$ and $\rho \in \Gamma(I)$. The elements $w$ of
-$(\Pi x:A.B)\rho$ are families $w = (w_f~|~J\in \cat{C}, f \from J \to I)$ of
-(dependent) functions such that
+\begin{lemma}
+  CwF associated to a presheaf category supports $\Pi$-types.
+  \begin{proof}
+    We give a proof sketch as seen in \cite{Huber}. More detailed proof may be
+    found in \cite{Hofmann}.
+
+    Given $\Gamma \vdash A$ and $\Gamma.A \vdash B$ we have to define $\Gamma
+    \vdash \Pi x: A. B~type$, that is, we have to define the set $(\Pi x: A.B)
+    \rho$ for $I \in \cat{C}$ and $\rho \in \Gamma(I)$. The elements $w$ of
+    $(\Pi x:A.B)\rho$ are families $w = (w_f~|~J\in \cat{C}, f \from J \to I)$
+    of (dependent) functions such that
     \[
@@
     \]
-where $g \from K \to J$ and $u \in A(\rho f g)$. This definition satisfies $w 1
-= w$ and $w f g = w(f g)$.
+    where $g \from K \to J$ and $u \in A(\rho f g)$. This definition satisfies
+    $w 1 = w$ and $w f g = w(f g)$.
 
-Now let $\Gamma.A \vdash b : B$; we have to define $\Gamma \vdash \lambda b :
-\Pi x:A.B$, that is , give $(\lambda b)\rho \in (\Pi x:A.B)\rho$. For $f \from
-J \to I$ and $u \in A(\rho f)$ we set
+    Now let $\Gamma.A \vdash b : B$; we have to define $\Gamma \vdash \lambda b
+    : \Pi x:A.B$, that is , give $(\lambda b)\rho \in (\Pi x:A.B)\rho$. For $f
+    \from J \to I$ and $u \in A(\rho f)$ we set
     \[
@@
 
-To define the application let $\Gamma \vdash u : \Pi x:A.B$ and $\Gamma \vdash
-v : A$, and we set $App(u, v)\rho = (u\rho)_1 (v\rho) \in B(\rho, v\rho)$, thus
-$App(u, v)\rho \in B(1, v)\rho$ as $(\rho, v \rho) = (1, v)\rho$.
-$\beta$-equality is readily checked
+    To define the application let $\Gamma \vdash u : \Pi x:A.B$ and $\Gamma
+    \vdash v : A$, and we set $App(u, v)\rho = (u\rho)_1 (v\rho) \in B(\rho,
+    v\rho)$, thus $App(u, v)\rho \in B(1, v)\rho$ as $(\rho, v \rho) = (1,
+    v)\rho$. $\beta$-equality is readily checked
     \[
@@
     and similarly for $\eta$-equality and other equations.
+  \end{proof}
+\end{lemma}
 
@@
 \]
-for $\rho \in \Gamma(I), I \in \ob(\cat{C})$. The restrictions are defined
-componentwise $(a, b)f = (af, bf)$ for $f \from J \to I$. The pairing operation
-$\Gamma \vdash (u, v) : \Sigma x:A.B$ of $\Gamma \vdash u : A$ and $\Gamma.A
-\vdash v : B(1, u)$ is defined by the componentwise pairing: $(u, v)\rho =
-(u\rho, v\rho)$. Likewise for the projections $\mathbf{p}$ and $\mathbf{q}$,
-$(\mathbf{p}w)\rho = a$ and $(\mathbf{q}w)\rho = b$ for $w\rho = (a, b)$. This
-validates the necessary equations.
+for $\rho \in \Gamma(I), I \in \ob(\cat{C})$.
+\begin{lemma}
+  CwF associated to a presheaf category supports $\Sigma$-types.
+  \begin{proof}
+    Again we give a proof sketch as seen in \cite{Huber}. More detailed proof
+    is also found in \cite{Hofmann}.
+
+    The restrictions are defined componentwise:
+    \[
+      (a, b)f = (af, bf)
+    \]
+    for $f \from J \to I$.
+
+    The pairing operation $\Gamma \vdash (u, v) : \Sigma x:A.B$ of $\Gamma
+    \vdash u : A$ and $\Gamma.A \vdash v : B(1, u)$ is defined by the
+    componentwise pairing:
+    \[
+      (u, v)\rho = (u\rho, v\rho)
+    \]
+
+    Likewise for the projections $\mathbf{p}$ and $\mathbf{q}$,
+    $(\mathbf{p}w)\rho = a$ and $(\mathbf{q}w)\rho = b$ for $w\rho = (a, b)$.
+    This validates the necessary equations.
+  \end{proof}
+\end{lemma}
 

### 2017-09-21 papers/talk-8/CCC.tex

@@
 Hence $$a \Rightarrow b  \leq a \Rightarrow b$$
-$$ a \Rightarrow b \land \a \leq b$$ by the condition $a \land b \leq c$ iff $a \leq b \Rightarrow c$, which is the evaluation arrow, and it is monotone
+$$ a \Rightarrow b \land a \leq b$$ by the condition $a \land b \leq c$ iff $a \leq b \Rightarrow c$, which is the evaluation arrow, and it is monotone
 $$ a \Rightarrow b \land a = (\neg a \lor b) \land a$$
@@
 
-   \\ For the second axiom, choose $f:A\to B^B$ as and $p:=1_A$.Let $\mu := \lambda^{-1}$
+   For the second axiom, choose $f:A\to B^B$ as and $p:=1_A$.Let $\mu := \lambda^{-1}$
    $$f \circ 1_A : A \to A \to B^B$$
    the transpose has type
-   $$\lambda^{-1}: A\timesB \to B $$
+   $$\lambda^{-1}: A \times B \to B $$
    what happens to be the type of
@@
 
-  \\
   So $\forall A,B,C \in \textbf{Cat}$, the exponential object is given as $C^B$, with the evaluation arrow
@@
   \\
-  For some $g:A\toC^B$, we have the inverse of the transpose
+  For some $g:A \to C^B$, we have the inverse of the transpose
   $$\lambda^{-1}g: A\times B \to C$$ with
@@
 Beyond the already seen isomorphism $0 \times A \cong 0$, many identities knwon from common arithmetics,like for instance that of the natural numbers,
-hold for cartesian closed categories.$\forall A, B, C \in $ in some cartesian closed category $\mathcal{A}$ with terminal object "1" and binary products "$\times$": \\
+hold for cartesian closed categories.$\forall A, B, C \in $ in some cartesian closed category $\mathcal{A}$ with terminal object "1" and binary products "$\times$": 
 \begin{itemize}
@@
 \end{itemize}
-\\
+
 If $\mathcal{A}$ is bicartesian closed with initial object "0" and co-product "+", the enumeration of axioms above is extended by: 
@@
 \end{itemize}
-\\
+
 We are going to show some of these identities by transformations exploiting right or left adjointness.
@@
   So both sides satisfy the universal property and are therefore isomorphic.
-  \\
 
@@
   Take some X,Y $\in ob(\mathcal{A})$ and set
-  $$ X:=Hom(-, (A^B)^C) \quad and \quad Y:= Hom(-,A^{B\timesC})$$ then
+  $$ X:=Hom(-, (A^B)^C) \ and \ Y:= Hom(-,A^{B \times C})$$
+  then
   $$ Hom(-, (A^B)^C) \cong Hom(-\times C, A^B) $$
@@
   With B, associativity and commutativity of the product
-  $$ Hom(-,A^{B\times C}) \cong Hom(- \times (B\times C),A) $$
+  $$ Hom(-,A^{B\times C}) \cong Hom(- \times (B \times C),A) $$
   $$ \cong Hom(-\times (C\times B),A) $$
@@
 
-\\
 Subcategories $\mathcal{B}$ of cartesian closed categories $\mathcal{A}$ are in general not cartesian closed.

### 2017-09-28 papers/talk-11/paper-11.tex
Rewrite section on the model

@@
 \newtheorem{lemma}{Lemma}
+
+\newcommand{\Fam}{\fcat{Fam}}
 
@@
 \begin{definition}[Category of families of sets]
-  The category of families of sets, $\fcat{Fam}$, has as objects $(A, B)$ where
+  The category of families of sets, $\Fam$, has as objects $(A, B)$ where
   $A$ is a set and $B = (B_a~|~a \in A)$ is an $A$-indexed family of sets $B_a$,
@@
 
-  \item[$\bullet$] $F$ is a functor $F\from \cat{C}^\op \to \fcat{Fam}$; For a
+  \item[$\bullet$] $F$ is a functor $F\from \cat{C}^\op \to \Fam$; For a
     context $\Gamma$ we write $Ty_F(\Gamma)$ for the indexing set of the family
@@
     $a\sigma$.
+
+  \item[$\bullet$] There are operations of context extension as defined above.
 \end{itemize}
@@
 
-Let us now consider the categories of presheaves and see that they form a
-category with families where the contexts are presheaves.  Let $\cat{C}$ be a
-small category. We will refer to the objects of $\cat{C}$  as $I,J,K$ and
-morphisms as $f,g,h$.
+In the following we will give a presheaf model for dependent types. First
+recall the definition of a presheaf:
 
-The context $\Gamma$ is then a presheaf on $\cat{C}$, $\Gamma \from \cat{C}^\op
-\to \Set$, that is, we are given a set $\Gamma(I)$ for each $I$ in
-$\ob(\cat{C})$, and functions $\Gamma(I) \to \Gamma(J)$, $\rho \mapsto \rho f$
-(called \emph{restriction} and written as $f$ acting on the right) for each $f:
-J \to I$ such that
-\[
-  \rho 1 = \rho \quad \text{and} \quad (\rho f) g = \rho (f g)
-\]
-for $g: K \to J$ and $f: J \to I$. We write $\Gamma_f$ for $\rho \mapsto \rho
-f$.
+\begin{definition}[Presheaf]
+  Let $\cat{C}$ be a category. A \emph{presheaf} on $\cat{C}$ is a functor
+  $\cat{C}^\op \to \Set$.
+\end{definition}
 
-The empty context $\diamond$ is a terminal presheaf and a context morphism
-$\sigma \from \Delta \to \Gamma$ is a natural transformation that makes the
-diagram commute.
+Let us see how the category of presheaves over $\cat{C}$ (we denote it as
+$Psh(\cat{C})$) forms a category of families where presheaves are contexts.  If
+we set presheaves to be the contexts, then the context morphisms are morphisms
+in $[\cat{C}^\op, \Set]$, which are natural transformations between contravariant
+functors. These natural transformations satisfy the following naturality
+condition:
 \[
@@
 \]
+where $I, J \in \ob(\cat{C}); \Gamma, \Delta \in Psh(\cat{C}); \sigma
+\from \Delta \to \Gamma$, and $f \from J \to I$. $\Gamma_f$ and $\Delta_f$
+denote the morphisms that we obtain from the morphism $f$ in $\Set$ by the
+functoriality of $\Gamma$ and $\Delta$. For instance, if $\rho \in \Gamma(I)$,
+then we have a morphism $\Gamma_f$ to the corresponding $\rho f \in \Gamma(J)$
+(we write $f$ applied on the right to connect to the initial morphism in
+$\cat{C}$) since the contravariant functor $\Gamma$ preserves morphism
+mappings.
 
-In the following, we will omit the indices on $\sigma$; Then if we write
-$\Gamma_f$ and $\Delta_f$ as $f$ acting on the right (recall the session on
-presheaves), the equation becomes:
+Recall that every context has a morphism from itself to the empty context.
+Hence we set the empty context to be the terminal object in the category of
+presheaves, the terminal context, which is a functor that maps everything to
+the terminal object in $\Set$, the singleton set. Indeed, we need a unique
+natural transformation from every presheaf to the terminal presheaf. This
+natural transformation is a family of unique morphisms to the terminal object
+in $\Set$.
+
+To define $F$ let us consider a category of elements of a presheaf $\Gamma$,
+denoted by $\int\Gamma$. The objects of $\int \Gamma$ are $(I, \rho)$ where
+$I \in \ob(\cat{C})$ and $\rho \in \Gamma(I)$. For $J, I \in \ob(\cat{C})$ the
+morphism between $(J, \rho')$ and $(I, \rho)$ is a morphism $f \from J \to I$
+in $\cat{C}$ denoted as $(f, \rho)$, such that $\rho' = \Gamma_f \rho$. Note
+that this way the composition is inherited from $\cat{C}$.
+
+Now we can define $Ty_F(\Gamma)$ as a set of presheaves over $\int \Gamma$.
+Then $A \in Ty_F(\Gamma)$ is a presheaf and for a morphism in $\cat{C}$, $f
+\from J \to I$, $\rho \in \Gamma(I)$, $x \in A(I, \rho)$ we have $A(f, \rho)(x)
+\in A(J, \Gamma_f \rho)$ since $A$ is a functor. For a context morphism $\sigma
+\from \Delta \to \Gamma$ we define a functor $\int \sigma \from \int \Delta \to
+\int \Gamma$ by
 \[
-  (\sigma \rho) f = \sigma (\rho f)
+ \int \sigma (I, \rho') = (I, \sigma_I \rho'),
 \]
-for $\rho$ in $\Delta(I)$.
+where $\rho' \in \Delta(I)$.
 
-Now we consider how to give a dependent type $\Gamma \vdash A~type$ in context
-$\Gamma$. For each object $I \in \ob(\cat{C})$ and $\rho \in \Gamma(I)$ we require a
-set $A \rho$, and for each $f \from J \to I$, we require a function $A \rho \to
-A (\rho f)$, written as $a \mapsto af$ satisfying $a1 = a$ and $afg = a (f g)$
-if $g \from K \to J$. The dependence on $I$ in $A\rho$ and on $I$ and $\rho$ in
-$af$ is omitted for brevity. Substitution $\Delta \vdash A \sigma$ with $\sigma
-\from \Delta \to \Gamma$ is simply given $(A \sigma) \rho = A (\sigma \rho)$
-for $\rho \in \Gamma(I)$, together with the induced map
+To define terms of a dependent type $\Gamma \vdash A~type$ we require a family
+of elements $a \rho \in A \rho$ for each $I \in \cat{C}$ and $\rho \in
+\Gamma(I)$. These elements should satisfy for any $f \from J \to I$ the
+equation
 \[
-  (A\sigma)\rho = A (\sigma \rho) \to A ((\sigma \rho)f) = A (\sigma (\rho f)) = (A \sigma) (\rho f)
+  A(f, \rho)(a(I, \rho)) = a(J, \Gamma_f \rho).
 \]
-for $f \from J \to I$.
-
-The definition of a dependent type can be rephrased more categorically: $A$ is
-a presheaf on $\int_C \Gamma$, where $\int_C \Gamma$ is a category of elements
-of the presheaf $\Gamma$, defined as follows:
-\begin{itemize}
-  \item objects are pairs $(I, \rho)$, where $I \in \ob(\cat{C})$ and $\rho \in
-    \Gamma(I)$
-  \item a morphism $(J, \rho') \to (I, \rho)$ is a morphism $f \from J \to I$
-    in $\cat{C}$ such that $\rho' = \Gamma_f \rho$
-\end{itemize}
-
-A term $\Gamma \vdash a : A$ of a dependent type $\Gamma \vdash A~type$ is
-given by a family of elements $a \rho \in A \rho$ for each $I \in \ob(\cat{C})$
-and $\rho \in \Gamma(I)$, such that $a \rho f = a (\rho f)$ for each $f \from J
-\to I$. The substitution $\Delta \vdash a \sigma : A \sigma$ with $\sigma \from
-\Delta \to \Gamma$ is given by the family $(a \sigma) \rho = a (\sigma \rho)$.
-
-For $\Gamma \vdash A~type$ the context extension $\vdash \Gamma.A~ctxt$ is
-defined by
-\begin{align*}
-  &(\Gamma.A)(I) = \{(\rho, u)~|~\rho \in \Gamma(I) \land u \in A \rho\} &&
-  \text{for $I \in \cat{C}$}\\
-  &(\rho, u)f = (\rho f, u f) && \text{for $f \from J \to I$}
-\end{align*}
-and the projections are defined by $\mathbf{p} \from \Gamma.A \to \Gamma$,
-$\mathbf{p}(\rho, u) = \rho$, and $\Gamma.A \vdash \mathbf{q} : Ap$ by
-$\mathbf{q}(\rho, u) = u$.  Now assume $\sigma \from \Delta \to \Gamma$ and
-$\Delta \vdash a : A\sigma$; we define $(\sigma, a) \from \Delta \to \Gamma.A$
-by $(\sigma, a)\rho = (\sigma \rho, a \rho)$.
+The context morphism $\sigma \from \Delta \to \Gamma$ then operates on a term
+$\Delta \vdash a\sigma : A\sigma$ by composition
+\[
+  (a\sigma_I)(I, \rho) = a(I, \sigma_I(\rho)).
+\]
+We define the presheaf $\Gamma.A$ corresponding to a context extension as
+\[
+  (\Gamma.A)(I) = \{(\rho, u)~|~\rho \in \Gamma(I) \land u \in A(I, \rho)\}
+\]
+for objects $I \in \cat{C}$ and for morphisms $f \from J \to I$ as
+\[
+  (\Gamma.A)_f(\rho, u) = ((\Gamma.A)_f \rho, A(f, \rho)(u)).
+\]
+The projections are then defined by $\mathbf{p} \from \Gamma.A \to \Gamma$,
+$\mathbf{p}(\rho, u) = \rho$, and $\Gamma.A \vdash \mathbf{q} : A\mathbf{p}$ by
+$\mathbf{q}(\rho, u) = u$. Under context morphism $\sigma \from \Delta \to
+\Gamma$ we define an extension of $\Gamma$ with a term $\Delta \vdash a :
+A\sigma$ $(\sigma, a) \from \Delta \to \Gamma.A$ by $(\sigma, a)\rho = (\sigma
+\rho, a \rho)$.
 
@@
 
-\begin{example}
-  The type for the function $A \to A$ (where $A$ is a ground type) is expressed
-  as an exponential $M(A)^{M(A)}$ in STLC with $M$ being the interpretation
-  function and $M(A)$ an object corresponding to the ground type in cartesian
-  closed category.
-\end{example}
+So far we know how to represent a type $\Gamma \vdash A~type$ and a dependent
+type $\Gamma.A \vdash B$. Now we want to have a representation for a function
+that corresponds to a type $\Gamma \vdash \Pi A. B$.
 

### 2017-09-29 papers/talk-11/paper-11.tex
Update paper11

@@
 Then $A \in Ty_F(\Gamma)$ is a presheaf and for a morphism in $\cat{C}$, $f
-\from J \to I$, $\rho \in \Gamma(I)$, $x \in A(I, \rho)$ we have $A(f, \rho)(x)
+\from J \to I$, $\rho \in \Gamma(I)$, $u \in A(I, \rho)$ we have $A(f, \rho)(u)
 \in A(J, \Gamma_f \rho)$ since $A$ is a functor. For a context morphism $\sigma
@@
 To define terms of a dependent type $\Gamma \vdash A~type$ we require a family
-of elements $a \rho \in A \rho$ for each $I \in \cat{C}$ and $\rho \in
-\Gamma(I)$. These elements should satisfy for any $f \from J \to I$ the
+of elements $a(I, \rho) \in A(I, \rho)$ for each $I \in \ob(\cat{C})$ and $\rho
+\in \Gamma(I)$. These elements should satisfy for any $f \from J \to I$ the
 equation
@@
 \[
-  (a\sigma_I)(I, \rho) = a(I, \sigma_I(\rho)).
+  (a\sigma_I)(I, \rho') = a(I, \sigma_I(\rho')).
 \]
-We define the presheaf $\Gamma.A$ corresponding to a context extension as
+for $\rho' \in \Delta(I)$.
+
+We define the presheaf $\Gamma.A$ corresponding to a context extension for
+objects $I \in \ob(\cat{C})$ as
 \[
@@
 \]
-for objects $I \in \cat{C}$ and for morphisms $f \from J \to I$ as
+and for morphisms $f \from J \to I$ as
 \[
@@
 \Gamma$ we define an extension of $\Gamma$ with a term $\Delta \vdash a :
-A\sigma$ $(\sigma, a) \from \Delta \to \Gamma.A$ by $(\sigma, a)\rho = (\sigma
-\rho, a \rho)$.
+A\sigma$, $(\sigma, a) \from \Delta \to \Gamma.A$ by
+\[
+  (\sigma, a)\rho = (\sigma \rho, a \rho)
+\]
+for the appropriate subscript on $\sigma$.
+
+The second projection $\mathbf{q}$ here is the one that models the variables in
+an extended context. The application of it returns the current term that
+substitutes the variable. Note the similarity to the variable rule in the ccc
+model of STLC, as it is also modeled by a second projection there.
 
@@
 
-So far we know how to represent a type $\Gamma \vdash A~type$ and a dependent
-type $\Gamma.A \vdash B$. Now we want to have a representation for a function
-that corresponds to a type $\Gamma \vdash \Pi A. B$.
-
-To represent the same type in the model with dependent types we will also use
-exponentials, but on the presheaves. Let us consider how to define exponents in
-presheaf categories. Let $\Gamma$ and $\Delta$ be presheaves and suppose we
-already know how the exponent $\Delta^\Gamma$ is constructed. Then we get by
-the Yoneda Lemma and the fact that $-^\Gamma$ should be right adjoint to $-
-\times \Gamma$, that
-\[
-  (\Delta^\Gamma)(I) \cong Hom(\mathbf{y}I, \Delta^\Gamma) \cong Hom(\mathbf{y}I \times \Gamma, \Delta)
-\]
-where $\mathbf{y}$ is the Yoneda embedding.
+So far we have a model for types $\Gamma \vdash A~type$ and $\Gamma.A \vdash
+B~type$. Now we want to find a model for a dependent function type $\Gamma
+\vdash \Pi x:A.B$. This model should connect our representations $A(I, \rho)$
+and $B(\rho, u)$ for these types, where $I \in \ob(\cat{C})$, $\Gamma \in
+\ob(Psh(\cat{C}))$, $\rho \in \Gamma(I)$, $A \in Psh(\int \Gamma)$, $u \in A(I,
+\rho)$.
 

### 2017-09-29 papers/talk-11/paper-11.tex
Proofread context extension model
Rewrite sections on dependent product and sum

@@
   \]
-  Note that in the last equation $1: \Gamma.A \to \Gamma.A$ and that we leave
-  the indices on $\mathbf{p}$ implicit.
+  Note that in the last equation $1: \Gamma.A \to \Gamma.A$.
 \end{definition}
@@
 functoriality of $\Gamma$ and $\Delta$. For instance, if $\rho \in \Gamma(I)$,
-then we have a morphism $\Gamma_f$ to the corresponding $\rho f \in \Gamma(J)$
-(we write $f$ applied on the right to connect to the initial morphism in
-$\cat{C}$) since the contravariant functor $\Gamma$ preserves morphism
-mappings.
+then we have a morphism $\rho \mapsto \rho f$ written as $\Gamma_f$ to the
+corresponding $\rho f \in \Gamma(J)$, (we write $f$ applied on the right to
+connect to the initial morphism in $\cat{C}$) since the contravariant functor
+$\Gamma$ preserves morphism mappings.
 
@@
 of elements $a(I, \rho) \in A(I, \rho)$ for each $I \in \ob(\cat{C})$ and $\rho
-\in \Gamma(I)$. These elements should satisfy for any $f \from J \to I$ the
-equation
+\in \Gamma(I)$. Additionally, since $A$ is a presheaf these elements should
+satisfy for any $f \from J \to I$ the equation
 \[
@@
 \[
-  (\sigma, a)\rho = (\sigma \rho, a \rho)
+  (\sigma, a)(I, \rho') = (\sigma_I \rho', a(I, \rho'))
 \]
-for the appropriate subscript on $\sigma$.
 
@@
 B~type$. Now we want to find a model for a dependent function type $\Gamma
-\vdash \Pi x:A.B$. This model should connect our representations $A(I, \rho)$
-and $B(\rho, u)$ for these types, where $I \in \ob(\cat{C})$, $\Gamma \in
-\ob(Psh(\cat{C}))$, $\rho \in \Gamma(I)$, $A \in Psh(\int \Gamma)$, $u \in A(I,
-\rho)$.
+\vdash \Pi x:A.B$ that should connect a term $u \in A(I, \rho)$ with a term $v
+\in B(I, (\rho, a))$. In other words, the element $\pi$ of $(\Pi x:A.B)(I, \rho)$
+should satisfy
+\[
+  \pi(u) \in B(I, (\rho, u))
+\]
+for some $I \in \ob(\cat{C})$, $\Gamma \in \ob(Psh(\cat{C}))$, $\rho \in
+\Gamma(I)$.
 
-We can take that as a definition. Then an element $w$ of $(\Delta^\Gamma)(I)$
-is a natural transformation $w \from \mathbf{y}I\times \Gamma \to \Delta$ given
-by functions
+However, this condition alone is not enough: since $\Pi x: A.B$ is a presheaf
+we also need to preserve the equalities for the morphisms in $\int \Gamma$
+(which are morphisms in $\cat{C}$). To generalize for the morphism $f \from J
+\to I$ (where $I, J \in \ob(\cat{C}))$) we extend $\pi \in (\Pi x: A.B)(I,
+\rho)$ to a family of functions $\pi = (\pi_f~|~f \from J \to I)$ and require
+the following equation to hold:
 \[
-  w_J \from (\mathbf{y}I)(J) \times \Gamma(J) \to \Delta(J),
+  \pi_f(u) \in B(J, (\Gamma_f \rho, A(f, \rho)(u)))
 \]
-and the naturality condition becomes $(w_J(f, \rho))g = w_K(fg, \rho g)$ for $f
-\from J \to I$, $\rho \in \Gamma(J)$, and $g \from K \to J$.
+where $\rho \in \Gamma(I)$ and $u \in A(J, \Gamma_f \rho)$. Notice that since
+$(f, \rho)$ is a morphism in $\int \Gamma$ we have $A(f, \rho)(u) \in A(I,
+\Gamma_f \rho)$. Another point is that if we take $1 \from I \to I$ instead of
+$f$, then we get our above condition immediately.
 
+We also need to respect compositions:
+\[
+  \pi_{fg}(A(g, \Gamma_f \rho)(u)) = B(g, (\Gamma_f \rho, A(f, \rho)(u)))(\pi_f(u))
+\]
+for some $g \from K \to J$.
+
+And define the necessary equality for operations of the morphisms in $\cat{C}$
+on the right
+\[
+  ((\Pi u: A.B)(h, \Gamma_f \rho)(\pi))_f(u) = \pi_{hf}(u)
+\]
+where $h \from I' \to I$, $\pi \in (\Pi x:A.B)(I, \rho)$, $f \from J \to I'$,
+$u \in A(J, \Gamma_f(\Gamma_h \rho))$, and $(\Pi x: A.B)(h, \Gamma_f \rho)(\pi)
+\in (\Pi x: A.B')(I, \Gamma_h \rho)$.
+
+Let us now define terms $\lambda x: A. b$ of a dependent function type. If we
+have some $b \in B(I, (\rho, x))$, then we define $(\lambda x: a.b)(I, \rho)
+\in (\Pi x: A.B)(I, \rho)$ by:
+\[
+  ((\lambda u: A.b)(I, \rho))_f(u) = b(J, (\Gamma_f \rho, u)) \in B(J, (\Gamma_f \rho, u))
+\]
+where $f \from J \to I$ and $u \in A(J, \Gamma_f \rho)$. Note that this
+definition satisfies the required equalities above.
+
+What remains is to define the application of the lambda terms. First let
+$\Gamma \vdash u : \Pi x:A.B$ and $\Gamma \vdash v : A$. The application
+morphism $App$ is then thus defined for some $I \in \ob(\cat{C})$:
+\[
+  App(u, v)(I, \rho) = (u(I, \rho))_1(v(I, \rho)) \in B(I, (\rho, v(I, \rho)))
+\]
+
+Since the necessary elements of $\Pi$-types are in place, we can state the soundness lemma:
 \begin{lemma}
@@
   \begin{proof}
-    We give a proof sketch as seen in \cite{Huber}. More detailed proof may be
-    found in \cite{Hofmann}.
-
-    Given $\Gamma \vdash A$ and $\Gamma.A \vdash B$ we have to define $\Gamma
-    \vdash \Pi x: A. B~type$, that is, we have to define the set $(\Pi x: A.B)
-    \rho$ for $I \in \cat{C}$ and $\rho \in \Gamma(I)$. The elements $w$ of
-    $(\Pi x:A.B)\rho$ are families $w = (w_f~|~J\in \cat{C}, f \from J \to I)$
-    of (dependent) functions such that
-    \[
-      w_f u \in B(\rho f, u) \qquad \text{for $J \in \cat{C}, f \from J \to I, and u \in A(\rho f)$},
-    \]
-    with the requirement that for $g \from K \to J$
-    \[
-      (w_f u)g = w_{fg}(ug).
-    \]
-
-    For such a family $w \in (\Pi x: A. B)\rho$, the restriction $wf \in (\Pi
-    x:A.B)(\rho f)$ for $f: \from J \to I$ is defined by taking
-    \[
-      (w f)_g u = w_{fg}u \in B(\rho f g, u)
-    \]
-    where $g \from K \to J$ and $u \in A(\rho f g)$. This definition satisfies
-    $w 1 = w$ and $w f g = w(f g)$.
-
-    Now let $\Gamma.A \vdash b : B$; we have to define $\Gamma \vdash \lambda b
-    : \Pi x:A.B$, that is , give $(\lambda b)\rho \in (\Pi x:A.B)\rho$. For $f
-    \from J \to I$ and $u \in A(\rho f)$ we set
-    \[
-      ((\lambda b)\rho)_f u = b (\rho f, u) \in B(\rho f, u)
-    \]
-    This satisfies for $g \from K \to J$
-    \[
-      (((\lambda b)\rho)_f u)g = (b(\rho f, u))g = b(\rho(fg), ug) = ((\lambda b)\rho)_{fg}(ug)
-    \]
-    and defines a term since
-    \[
-      (((\lambda b)\rho)f)_g u = ((\lambda b) \rho)_{fg} u = b(\rho fg, u) = ((\lambda b)(\rho f))_g u.
-    \]
-
-    To define the application let $\Gamma \vdash u : \Pi x:A.B$ and $\Gamma
-    \vdash v : A$, and we set $App(u, v)\rho = (u\rho)_1 (v\rho) \in B(\rho,
-    v\rho)$, thus $App(u, v)\rho \in B(1, v)\rho$ as $(\rho, v \rho) = (1,
-    v)\rho$. $\beta$-equality is readily checked
-    \[
-      App(\lambda b, v)\rho = ((\lambda b)\rho)_1(v\rho)=b(\rho, v\rho)
-    \]
-    and similarly for $\eta$-equality and other equations.
+    Follows after the verification of the $\Pi$-rules on the above equations. A
+    detailed proof is found $\cite{Hofmann}$
   \end{proof}
@@
 \begin{example}
-  Let us consider a term of type $A \to A$, and the elements in $Ty(\Gamma)$
-  and $Ter(\Gamma, A \to A)$ for some $\vdash \Gamma~ctxt$, $\Gamma \vdash
-  A~type$ in the presheaf model.
-
-  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we give the
+  Let us consider a representation of a term of type $A \to A$ for some $\vdash
+  \Gamma~ctxt$, $\Gamma \vdash A~type$ in the presheaf model. For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we give the
   element of $\mathit{Ty}(\Gamma)$ corresponding to the type $A \to A$:
   \[
-    (A \to A)\rho = \{w | w_f u \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
+    (A \to A)\rho = \{\pi~|~\pi_f(u) \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
   \]
-  The required equalities hold trivially for any $g \from K \to J$, $((\lambda
-  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.
-
-  $\Gamma, x : A \vdash x : A$ corresponds to $q$ in the model. Thus $\Gamma
+  Note that the required equalities hold trivially for any $g \from K \to J$, $((\lambda
+  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$. $\Gamma, x : A \vdash x : A$ corresponds to $q$ in the model. Thus $\Gamma
   \vdash \lambda x.x : A -> A$ corresponds to $\lambda q$ being defined as
@@
 \subsection*{Dependent sum}
-The interpretation of $\Gamma \vdash \Sigma x:A.B~type$ for $\Gamma \vdash
-A~type$ and $\Gamma.A \vdash B~type$ is defined by
+Suppose we have some types $\Gamma \vdash A~type$ and $\Gamma.A \vdash B~type$.
+Then we can define the dependent pair type $\Gamma \vdash \Sigma x:A.B~type$
+for some $I \in \ob(\cat{C})$ by
 \[
-  (\Sigma x :A. B)\rho = \{(a, b)~|~a \in A\rho \land b \in B(\rho, a)\}
+  (\Sigma a :A. B)(I, \rho) = \{(a, b)~|~a \in A(I, \rho) \land b \in B(I, (\rho, a))\}
 \]
-for $\rho \in \Gamma(I), I \in \ob(\cat{C})$.
+where $\rho \in \Gamma(I)$.
+
+Note that as is the case with the dependent product we need to take the
+morphisms in $\cat{C}$ into account if we want $\Sigma x:A.B$ to be a presheaf.
+This leads us to the following equation for some $f \from J \to I$:
+\[
+  \Gamma_f(a, b) = (\Gamma_f a, \Gamma_f b)
+\]
+
+Also, for a pair to be a term, both of its elements should terms in the
+following manner:
+\[
+  (u, v)(I, \rho) = (u(I, \rho), v(I, \rho))
+\]
+where $u \in A(I, \rho)$ and $v \in B(I, (\rho, u))$.
+
+Lastly, we need projection morphisms to be defined:
+\[
+  (\mathbf{p}(w))(I, \rho) = a
+\]
+\[
+  (\mathbf{q}(w)(I, \rho) = b
+\]
+where $w(I, \rho) = (a, b)$.
+
 \begin{lemma}
@@
   \begin{proof}
-    Again we give a proof sketch as seen in \cite{Huber}. More detailed proof
-    is also found in \cite{Hofmann}.
-
-    The restrictions are defined componentwise:
-    \[
-      (a, b)f = (af, bf)
-    \]
-    for $f \from J \to I$.
-
-    The pairing operation $\Gamma \vdash (u, v) : \Sigma x:A.B$ of $\Gamma
-    \vdash u : A$ and $\Gamma.A \vdash v : B(1, u)$ is defined by the
-    componentwise pairing:
-    \[
-      (u, v)\rho = (u\rho, v\rho)
-    \]
-
-    Likewise for the projections $\mathbf{p}$ and $\mathbf{q}$,
-    $(\mathbf{p}w)\rho = a$ and $(\mathbf{q}w)\rho = b$ for $w\rho = (a, b)$.
-    This validates the necessary equations.
+    Follows after verification of $\Sigma$-rules. A detailed proof is found in
+    \cite{Hofmann}.
   \end{proof}
 \end{lemma}
-
-\section*{Credits}
-
-Most of the material on the dependent types bases on \cite{Hofmann}, while
-starting from categories with families we closely follow \cite{Huber}.
 

### 2017-09-29 papers/talk-11/paper-11.tex
Update paper11
Fix typos

@@
 The second projection $\mathbf{q}$ here is the one that models the variables in
-an extended context. The application of it returns the current term that
-substitutes the variable. Note the similarity to the variable rule in the ccc
-model of STLC, as it is also modeled by a second projection there.
+an extended context. The application of it returns the current term for the
+variable. Note the similarity to the variable rule in the ccc model of STLC, as
+it is also modeled by a second projection there.
 
@@
 \[
-  (\mathbf{q}(w)(I, \rho) = b
+  (\mathbf{q}(w))(I, \rho) = b
 \]
@@
   \begin{proof}
-    Follows after verification of $\Sigma$-rules. A detailed proof is found in
-    \cite{Hofmann}.
+    Follows after verification of the $\Sigma$-rules. A detailed proof is found
+    in \cite{Hofmann}.
   \end{proof}

### 2017-09-30 papers/talk-11/paper-11.tex
Fix example for A -> A
Correct wording/typos

@@
 is not clear what should the result be. Obviously, such function (if we do not
-allow definition of partial functions which is the case in many programming
+allow definitions of partial functions which is the case in many programming
 languages) cannot be well defined on the empty list without taking arbitrary
-decisions. But maybe such problems can be prevent from happening at all?
+decisions. But maybe such problems can be prevented from happening at all?
 
@@
 \begin{example}
-  The function to construct a list of length $n$, having only $0$ as elements
+  The function to construct a list of a length $n$, having only $0$ as elements
   would then have the type $\Pi n: \NN.List(n)$, indicating that it accepts a
@@
 
-To formalize dependent types we will use a notion of typing context (which is
-almost the same as typing environment as seen for STLC, with the exception that
-it holds dependent types).
+To formalize dependent types we will use a notion of a typing context (which is
+almost the same as the typing environment as seen for STLC, with the exception
+that it holds dependent types).
 
@@
   distinct from the other variables and each $A_i$ is a type in the context $x_1
-  : A_1, \dots, x_{n-1} : A_{n-1}$, since our types may now include the variables
-  from the context.
+  : A_1, \dots, x_{n-1} : A_{n-1}$.
 \end{definition}
@@
 we only needed the variables to be distinct, since they had no influence on the
-types.
+types. We require the additional condition because our types may now include
+the variables from the context.
 
@@
 \end{mathpar}
-
 where $x$ and $y$ in C-Ext and C-Ext-Eq are assumed to be fresh variables and
-$\mathcal{J}$ in Weak and Subst ranges over judgements of the form $a : A,
+$\mathcal{J}$ in Weak and Subst ranges over the judgements of the form $a : A,
 A~type, a = b: A, A = B~type$. $[c/x]$ denotes the same substitution as seen in
 STLC, except that it now also applies to the variables that may be contained in
-types.
+the types.
 \end{definition}
@@
 
-Rules C-Eq define that context equality is an equivalence relation, very
+Rules C-Eq define that context equality is an equivalence relation. Very
 similar rules for terms and types, stating that their equalities are also
-equivalence relations, are omitted for brevity.
+equivalence relations, are omitted for brevity. One should keep in mind however
+that we implicitly include them too.
 
@@
 the same as in the simply-typed case, with the exception that our $\lambda$
-here forms a dependent function type, not the simply-typed one.
+here forms a dependent function type, not a simply-typed one.
 
@@
 reader should keep their implicit presence in mind while seeing other term and
-type formers that we introduce.
+type formers.
 
-The rule $\Pi$-E regulates the application of a dependent function to some value
-of the correct type results in the return value of a proper type. Note the rule
-only concerns the resulting type here, not the value and that the bound
-variable $x$ (shown as $[x : A] B$ in subscript) is replaced in the type, not
-the value, which is handled by the next rule.
+The rule $\Pi$-E regulates that the application of a dependent function to some
+value of a correct type results in the return value of a proper type. Note that
+the rule only concerns the resulting type here --- not the value --- and that
+the bound variable $x$ (shown as $[x : A] B$ in subscript) is replaced in the
+type, not the value, which is handled by the next rule.
 
@@
 The formation rule $\Sigma$-F follows the same pattern as the rule for
-dependent function, but the introduction is a bit different. Note that we take
+dependent functions, but the introduction is a bit different. Note that we take
 such $b$, where the type contains a variable $x : A$ that is substituted by the
@@
 
-We know the projection functions for the simply-typed products in STLC. Let us
-now define the projection functions for the \emph{dependent} products, which we
-call $\mathbf{p}$ for the first projection and $\mathbf{q}$ for the second
+We know the projection functions for the simply-typed pairs in STLC. Let us now
+define the projection functions for the \emph{dependent} pairs, which we call
+$\mathbf{p}$ for the first projection and $\mathbf{q}$ for the second
 projection:
@@
 
-The analogue to dependent sum in set theory is a disjoint union. For every
-family of sets $(B_i)_{i \in I}$ we can form the set $\Sigma_{i \in I} B_i :=
-\{ (i, b)~|~i \in I \land b \in B_i\}$ whose elements consist of an index $i$
-and an element of the corresponding set $B_i$.
+The analogue to dependent sum in set theory is disjoint union. For every family
+of sets $(B_i)_{i \in I}$ we can form the set $\Sigma_{i \in I} B_i := \{ (i,
+b)~|~i \in I \land b \in B_i\}$ whose elements consist of an index $i$ and an
+element of the corresponding set $B_i$.
 
@@
 
-A note on notation: we sometimes write $A\sigma$, where $\Gamma \vdash
-A\sigma~type$, $\Delta \vdash A~type$, and $\sigma \from \Gamma \to \Delta$ for
-some contexts $\Gamma$ and $\Delta$, to denote the application of $\sigma$ on
-the right. That is, $A\sigma$ becomes here a type in the domain context of
-$\sigma$, where $\sigma(A\sigma) = A$. The same notation is sometimes
-analogously used for terms: $a\sigma$ to denote a term $\Gamma \vdash a\sigma :
-A\sigma$, where $\Delta \vdash a : A$.
+We sometimes write $A\sigma$, where $\Gamma \vdash A\sigma~type$, $\Delta
+\vdash A~type$, and $\sigma \from \Gamma \to \Delta$ for some contexts $\Gamma$
+and $\Delta$, to denote the application of $\sigma$ on the right. That is,
+$A\sigma$ becomes here a type in the domain context of $\sigma$, where
+$\sigma(A\sigma) = A$. The same notation is sometimes analogously used for
+terms: $a\sigma$ denotes a term $\Gamma \vdash a\sigma : A\sigma$, where
+$\Delta \vdash a : A$.
 
@@
 using it as the most suitable identity morphism for the surroundings. We also
-assume correct types in the equations, i.e. $(A\sigma)\tau = A(\sigma\tau)$
+assume the correct types in the equations, i.e. $(A\sigma)\tau = A(\sigma\tau)$
 assumes $\sigma \from \Delta \to \Gamma$, $\tau \from \Theta \to \Delta$, and
@@
 There are several ways to model dependent types with category theory, the one
-we show is using Categories with Families.
+we show relies on Categories with Families.
 
@@
 
-A category with families supports $\Pi$- and $\Sigma$-types if it is
-closed under the $\Pi$ and $\Sigma$ rules.
+A category with families supports $\Pi$- and $\Sigma$-types if it is closed
+under the $\Pi$ and $\Sigma$ rules.
 
@@
 we set presheaves to be the contexts, then the context morphisms are morphisms
-in $[\cat{C}^\op, \Set]$, which are natural transformations between contravariant
-functors. These natural transformations satisfy the following naturality
-condition:
+in $[\cat{C}^\op, \Set]$, which are natural transformations between
+contravariant functors. These natural transformations satisfy the following
+naturality condition:
 \[
@@
 \]
-where $I, J \in \ob(\cat{C}); \Gamma, \Delta \in Psh(\cat{C}); \sigma
-\from \Delta \to \Gamma$, and $f \from J \to I$. $\Gamma_f$ and $\Delta_f$
-denote the morphisms that we obtain from the morphism $f$ in $\Set$ by the
-functoriality of $\Gamma$ and $\Delta$. For instance, if $\rho \in \Gamma(I)$,
-then we have a morphism $\rho \mapsto \rho f$ written as $\Gamma_f$ to the
+where $I, J \in \ob(\cat{C}); \Gamma, \Delta \in Psh(\cat{C}); \sigma \from
+\Delta \to \Gamma$, and $f \from J \to I$. $\Gamma_f$ and $\Delta_f$ denote the
+morphisms that we obtain from the morphism $f$ in $\Set$ by the fact that
+$\Gamma$ and $\Delta$ are functors. For instance, if $\rho \in \Gamma(I)$, then
+we have a morphism $\rho \mapsto \rho f$ written as $\Gamma_f$ to the
 corresponding $\rho f \in \Gamma(J)$, (we write $f$ applied on the right to
@@
 Hence we set the empty context to be the terminal object in the category of
-presheaves, the terminal context, which is a functor that maps everything to
+presheaves, the terminal context. That is a functor that maps everything to
 the terminal object in $\Set$, the singleton set. Indeed, we need a unique
@@
 an extended context. The application of it returns the current term for the
-variable. Note the similarity to the variable rule in the ccc model of STLC, as
-it is also modeled by a second projection there.
+variable. Note the similarity to the variable rule in the cartesian closed
+category model of STLC, as it is also modeled by a second projection there.
 
@@
   \[
-    (A \to A)\rho = \{\pi~|~\pi_f(u) \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
+    (A \to A)(I, \rho) = \{\pi~|~\pi_f(u) \in A(I, \Gamma_f \rho), u \in A(I, \Gamma_f \rho)\} = \{\lambda I f u.u\}
   \]
-  Note that the required equalities hold trivially for any $g \from K \to J$, $((\lambda
-  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$. $\Gamma, x : A \vdash x : A$ corresponds to $q$ in the model. Thus $\Gamma
-  \vdash \lambda x.x : A -> A$ corresponds to $\lambda q$ being defined as
-  follows:
+  Note that the required equalities hold trivially for any $g \from K \to J$,
+  $((\lambda I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$. $\Gamma, x : A
+  \vdash x : A$ corresponds to $q$ in the model. Thus $\Gamma \vdash \lambda
+  x.x : A -> A$ corresponds to $\lambda q$ being defined as follows:
   \[
-    ((\lambda q) \rho)_f u = q(\rho f, u) = u
+    ((\lambda q) (I, \rho))_f u = q(\Gamma_f \rho, u) = u
   \]

### 2017-10-11 papers/base-3/paper-3.tex
Update section on functor category

@@
 \section*{Functor Category}
-\begin{defn}[Functor Category]
-  Natural transformations, as other types of maps that we have encountered earlier
-  can be composed. That is, for any two categories $\cat{A}$ and $\cat{B}$,
-  functors $F, G, H$ between them, natural transformations $\alpha \from F \to
-  G$ and $\beta \from G \to H$:
-  \vspace{-2ex}
+
+Consider a composition $\beta \of \alpha$ between some natural transformations
+$\alpha \from F \to G$ and $\beta \from G \to H$.
 \[
@@
 \vspace{-3ex}\\
-  there is a composite natural transformation $\beta \of \alpha$
+We can define such composite natural transformation
 \[
@@
 \]
-  defined by $(\beta\of\alpha)_A = \beta_A \of \alpha_A$ for all $A \in \cat{A}$.
-  There is also an identity natural transformation
+with $(\beta\of\alpha)_A = \beta_A \of \alpha_A$ for all $A \in \cat{A}$.
+(Recall that $\beta_A$ and $\alpha_A$ are moprhisms in $\cat{B}$.)
+
+We can also define an identity natural transformation
 \[
@@
 \]
-  on any functor $F$, defined by $(1_F)_A = 1_{F(A)}$.
+on any functor $F$ with $(1_F)_A = 1_{F(A)}$.
 
-  This gives rives to a category whose objects are the functors from $\cat{A}$ to
-  $\cat{B}$ and whose maps are the natural transformations between them. This is
-  called the \demph{functor category} from $\cat{A}$ to $\cat{B}$, and written as
-  $\ftrcat{\cat{A}}{\cat{B}}$ or $\cat{B}^\cat{A}$.
+Thus we have a definition of a category:
+
+\begin{defn}[Functor Category]
+  The \demph{functor category} from a category $\cat{A}$ to a category
+  $\cat{B}$ (written as $\ftrcat{\cat{A}}{\cat{B}}$ or $\cat{B}^\cat{A}$) is a
+  category whose objects are the functors from $\cat{A}$ to $\cat{B}$. The maps
+  in this category are the natural transformations between functors.
 \end{defn}
 
-\begin{exmp}
-  Let $2$ be the discrete category with two objects. A functor from $2$ to a
-  category $\cat{B}$ is a pair of objects of $\cat{B}$, and a natural
-  transformation is a pair of maps.  The functor category $\ftrcat{2}{\cat{B}}$
-  is therefore isomorphic to the product category $\cat{B} \times \cat{B}$.
-  This fits well with the alternative notation $\cat{B}^2$ for the functor
-  category.
+\begin{exmp}[Functors from a discrete category]
+  Consider the discrete category with two objects, $2$. Any functor from $2$ to
+  some another category $\cat{B}$ will be a pair of objects in $\cat{B}$. A
+  natural transformation between such functors then is a pair of maps between
+  these objects.  Note that the functor category $\ftrcat{2}{\cat{B}}$ is
+  isomorphic to the product category $\cat{B} \times \cat{B}$ and the
+  superscript notation for the functor category $\cat{B}^2$ reflects this
+  nicely.
 \end{exmp}
 
-\begin{exmp}
-  Take ordered sets $A$ and $B$, viewed as categories. Given order-preserving
-  maps $\parpairi{A}{B}{f}{g}$, viewed as functors, there is at most one
-  natural transformation
-  \[
-  \xymatrix{
-  A \rtwocell^f_g &B,
-  }
-  \]
-  and there is one if and only if $f(a) \leq g(a)$ for all $a \in A$.  (The
-  naturality axiom~\eqref{eq:nat} holds automatically, because in an ordered
-  set, all diagrams commute.)  So $\ftrcat{A}{B}$ is an ordered set too; its
-  elements are the order-preserving maps from $A$ to $B$, and $f \leq g$ if and
-  only if $f(a) \leq g(a)$ for all $a \in A$.
+\begin{exmp}[Maps in ordered sets]
+  Consider ordered sets $A$ and $B$, viewed as categories. The maps
+  $\parpairi{A}{B}{f}{g}$ that preserve order can be then considered as
+  functors. Note that there is at most one natural transformation between such
+  functors and it exists if and only if $f(a) \leq g(a)$ for all $a \in A$.  In
+  this case we have the naturality axiom~\eqref{eq:nat} automatically, since
+  all diagrams commute in an ordered set. Moreover, the functor category
+  $\ftrcat{A}{B}$ is an ordered set itself, where elements are the
+  order-preserving maps from $A$ to $B$, and there a morphism $f \leq g$
+  between $f$ and $g$ if and only if $f(a) \leq g(a)$ for all $a \in A$.
 \end{exmp}
 
-\begin{defn}
+\begin{defn}[Natural isomorphism]
   Let $\cat{A}$ and $\cat{B}$ be categories. A \demph{natural isomorphism}
@@
 
-Then, we say that functors $F$ and $G$ are \demph{naturally isomorphic} if
-there is a natural isomorphism from $F$ to $G$. Note that this natural
-isomorphism is just isomorphism in $\ftrcat{\cat{A}}{\cat{B}}$, and we already
-have notation for this: $F \iso G$.
+\begin{defn}[Isomorphic functors]
+  We call functors $F$ and $G$ \demph{naturally isomorphic} if there is a
+  natural isomorphism from $F$ to $G$, which is just an isomorphism in
+  $\ftrcat{\cat{A}}{\cat{B}}$. The notation for isomorphic functors is the same
+  as for isomorphic objects: $F \iso G$.
+\end{defn}
 
-\begin{defn}
+\begin{defn}[Isomorphic naturally in an object]
   Given functors $\parpair{\cat{A}}{\cat{B}}{F}{G}$, we say that
@@
 
-This alternative terminology can be understood as follows.  If $F(A) \iso
-G(A)$ naturally in $A$ then certainly $F(A) \iso G(A)$ for each individual
-$A$, but more is true: we can choose isomorphisms $\alpha_A\from F(A) \to
-G(A)$ in such a way that the naturality axiom~\eqref{eq:nat} is satisfied.
+Note that the definition says that we not only have $F(A) \iso G(A)$ for each
+individual $A$, but also there is an isomorphism $\alpha_A\from F(A) \to G(A)$
+that satisfies the naturality axiom~\eqref{eq:nat}.
 
 \begin{exmp}
-  Let $F, G\from \cat{A} \to \cat{B}$ be functors from a discrete category
-  $\cat{A}$ to a category $\cat{B}$.  Then $F \iso G$ if and only if $F(A) \iso
-  G(A)$ for all $A \in \cat{A}$.
+  Consider two functors $F, G\from \cat{A} \to \cat{B}$ where $\cat{A}$ is a
+  discrete category. We have $F \iso G$ if and only if $F(A) \iso G(A)$ for all
+  $A \in \cat{A}$. If it is the case, then we also have that $F(A) \iso G(A)$
+  naturally in $A$ for all $A$.
 
-  So in \emph{this} case, $F(A) \iso G(A)$ naturally in $A$ if and only if
-  $F(A) \iso G(A)$ for all $A$.  But this is only true because $\cat{A}$ is
-  discrete.  In general, it is emphatically false.  There are many examples of
-  categories and functors $\parpairi{\cat{A}}{\cat{B}}{F}{G}$ such that $F(A)
-  \iso G(A)$ for all $A \in \cat{A}$, but not \emph{naturally} in $A$.
+  Be aware that it is not always so: there are examples of categories and
+  functors $\parpairi{A}{B}{F}{G}$ where $F(A) \iso G(A)$ for all $A \in
+  \cat{A}$, but not naturally in $\cat{A}$. One of the examples arises if one
+  considers two functors that take the objects of the category of finite sets
+  and bijections to the sets of permutations and the sets of total orders
+  correspondingly.  There one can find isomorphisms between the sets of
+  permutations and the sets of total orders, but no natural isomorphism between
+  such functors.
 \end{exmp}

### 2017-10-17 sheets/basis-1/sheet-basis-1.tex
Fixed compilation bugs on sheet 1

--- 2017-05-23 sheets/basis-1/sheet-basis-1.tex
@@
 \begin{document}
-
 % Use Basis x or Talk x, where x is the number of the session
@@
 \end{hint}
+
+%\section{Diagrams}
+%
+%\begin{hint}
+%  This is how you can draw diagrams:
+%  \[
+%    \begin{tikzcd}
+%                                    & A \arrow{dr}{g}    & \\
+%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
+%    \end{tikzcd}
+%  \]
+%\end{hint}
 
@@
 \end{exercise}
+
+\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets 
+(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A_\cdot $ ) 
+An arrow $ m \from A \to B $ function that is monotone, i.e. $\forall a, a' \in A$, 
+\[a \leq^A_\cdot a' \implies m(a) \leq^A_\cdot m(a')\]
+\end{definition}
 
@@
 \end{answer}
-
-\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
-(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A$).
-The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
-\end{definition}
 
@@
 \end{document}
-
 %%% Local Variables:
